### 2022-02-23 docs/Probabilistic Tiling/ProblemAndAlgorithm.tex
Add initial draft of probabilistic tiling document

\section{Introduction and Notation}

The fundamental problem we are faced with is to compute the prediction of a decision tree $T = (V, E, r)$ where $V$ is the set of nodes, $E$ the set of edges and
$r \in V$ is the root. For each node $n \in V$, the following are given.
\begin{enumerate}
    \item $threshold(n) \in \mathbb{R}$ which gives the threshold value for $n$.
    \item $featureIndex(n) \in \mathbb{N}$ which gives the feature index for $n$.
    \item $left(n) \in V$, the left child of $n$ or $\emptyset$ if $n$ is a leaf. If $left(n) \neq \emptyset$, then $(n, left(n)) \in E$.
    \item $right(n) \in V$, the right child of $n$ or $\emptyset$ if $n$ is a leaf. If $right(n) \neq \emptyset$, then $(n, right(n)) \in E$.
\end{enumerate}
We use $L \subset V$ to denote the set of leaves.

To compute the prediction of a tree $T$ on an input row $x$, the procedure in algorithm \ref{TreePredictionAlgo} is followed.
\begin{algorithm}
    \caption{Compute Tree Prdiction}
    \label{TreePredictionAlgo}
    \begin{algorithmic}
        \Procedure{PredictTree}{$T = (V, E, r)$, $x$} 
            \State $n \leftarrow r$
            \While{$n \notin L$}
                \State \textcolor{green}{\textit{//Compute the predicate of the current node.}}
                \State $pred \leftarrow x[featureIndex(n)] < threshold(n)$
                \State \textcolor{green}{\textit{//If predicate is true, move to right child, else left child.}}
                \If{$pred$}
                    \State $n \leftarrow left(n)$
                \Else
                    \State $n \leftarrow right(n)$
                \EndIf
            \EndWhile
            \State \textcolor{green}{\textit{//Return the threshold value of the leaf as the tree prediction.}}
            \State \textbf{return} $threshold(n)$
        \EndProcedure
    \end{algorithmic}
\end{algorithm}

In order to speed up the tree walk, Treebeard aggregates decision tree nodes into tiles and uses vector instructions to evaluate tiles.
Experiments have shown that, in several trees, some leaves are much more likely to be reached than others. Here, we explore how, 
given this fact, the tree can be tiled to \textbf{minimize average inference latency}. Practically, tiling is important from the perspective of both instruction count and locality.

\section{Tree Tiling}
\label{TreeTilingSection}
A tiling $\mathcal{T}$ of the tree $T = (V, E, r)$ with tile size $n_t$ is a partition $\{ T_1, T_2, ... ,T_m \}$ of the set $V$ such that 
\begin{enumerate}
    \item $T_1 \cup T_2 ... \cup T_m = V$
    \item $T_i \cap T_j = \emptyset$ for all $i, j \in [1, m]$ and $i \neq j$
    \item $|T_i| \leq n_t$ for all $i \in [1, m]$
    \item $\forall l \in L$ : $l \in T_i \rightarrow v \notin T_i \;\; \forall v \in V \backslash \{l\}$
    \item Tiles are \textbf{maximal}, i.e. if $|T_i| < n_t$, then there is no $v \in V\backslash \{ T_i \cup L \}$ such that $(u, v) \in E$ for some $u \in T_i$. 
    \item Tiles are \textbf{connected}, i.e. for an $u, v \in T_i$, there is a (undirected) path connecting $u$ and $v$ fully contained in $T_i$.
\end{enumerate}

\section{The Optimization Problem}

We assume that we are given the probabilities of each leaf node of the decision tree (these can easily be computed using the training data). For every leaf $l \in V$, we are given the probability $p_l$ that the leaf $l$ is reached.

We observe that the latency of one tree walk is proportional to the number of tiles that need to be evaluated to reach the leaf.
Therefore, the average inference latency can be modelled as the expected number of tiles that are evaluated to compute one tree prediction. More formally, the problem is to find a tiling $\mathcal{T}$ such that the following objective is minimized.
\[
    \min_{\mathcal{T}}{\sum_{l \in L} p_l.depth_{\mathcal{T}}(l)}
\]
where the minimization is over all valid tilings $\mathcal{T}$ of the tree $T$, $depth_{\mathcal{T}}(l)$ is the depth of the leaf $l$ given tiling ${\mathcal{T}}$.

The constraints on the optimization are the constraints listed in Section \ref{TreeTilingSection}.

\section{Dynamic Programming Formulation}
In order to formulate the above optimization problem as a dynamic program, we define the following.
\begin{enumerate}
    \item For each node $n \in V$, we define the absolute probability $p_v$ as
    \begin{equation}
        p_v = \begin{cases}
        p_l &\text{if $l \in L$}\\
        p_{left(v)} + p_{right(v)} &\text{otherwise}
        \end{cases}
    \end{equation}
    \item For any tree $T$, $\mathcal{C}(T)$ represents the set of all valid tilings of $T$.
    \item For every $v \in V$, we define $S_v$ as the subtree rooted at $v$.
    \item For every $v \in V$, we define $L_v$ as the set of leaves of $T_v$.
    \item For a every tile $T_i$, we define $root(T_i)$ as the node $v \in T_i$ such that $v$ has no incoming edges from any other node $u \in T_i$.
    \item For a tile $T_i$, $out(T_i) \subseteq E$ is the set of edges $(u, v)$ such that $u \in T_i$ and $v \notin T_i$.
\end{enumerate}

For any node $v \in V$, we define
\[
    cost(v, \mathcal{T}) = \sum_{l \in L_v} p(l | v).depth_{\mathcal{T}}(l)
\]
where $\mathcal{T} \in \mathcal{C}(T_v)$.

Then, the objective function, for the tree $T_v$, can be rewritten as 
\[
    opt\_cost(v) = \min_{\mathcal{T} \in \mathcal{C}(T_v)}{cost(v, \mathcal{T})}
\]

The objective function can then be rewritten in the following recursive form.
\[
    opt\_cost(v) = \min_{T_0 \in TileShapes(n_t, v)}{1 + \sum_{(n_1, n_2) \in out(T_0)} p(n_1 | v)p(n_2 | n1)opt\_cost(n_2)}
\]
where $TileShapes(n_t, v)$ is the set of all tile shapes of size $n_t$ with root $v$. A straight forward substitution argument shows why the solution to the subproblems (tiling all sub-trees) needs to to be optimal. The objective is now in a form that can solved using
dynamic programming.

\section{Greedy Algorithm}

Intuitively, it seems like the following greedy algorithm also gives the optimal tiling. The algorithm starts at the root and greedily keeps adding the most probable node to the current tile until the maximum tile size is reached.
\begin{algorithm}
    \caption{Greedy Tree Tiling}
    \label{GreedyTilingAlgo}
    \begin{algorithmic}
        \Procedure{TileTree}{$T = (V, E, r)$, $n_t$} 
            \If {$r \in L$}
                \State \textbf{return}
            \EndIf
            \State $Tile \leftarrow \{ r \}$
            \While{$|Tile| < n_t$}
                \State $e = (u,v) \in Out(Tile)$ such that $p(v | u)$ is maximum and $v \notin L$
                \If{$e = \emptyset$}
                    \State \textbf{break}
                \EndIf
                \State $Tile = Tile \cup \{ v \}$
            \EndWhile 
            \For{$(u,v) \in Out(Tile)$}
                \State $TileTree(T_v, n_t)$
            \EndFor
            \State \textbf{return} $threshold(n)$
        \EndProcedure
    \end{algorithmic}
\end{algorithm}

### 2022-02-23 docs/Probabilistic Tiling/prob-tiling.tex
Add initial draft of probabilistic tiling document

% !TEX TS-program = pdflatex
% !TEX encoding = UTF-8 Unicode

% This is a simple template for a LaTeX document using the "article" class.
% See "book", "report", "letter" for other types of document.

\documentclass[10pt, fleqn]{article} % use larger type; default would be 10pt

\usepackage[utf8]{inputenc} % set input encoding (not needed with XeLaTeX)

%%% Examples of Article customizations
% These packages are optional, depending whether you want the features they provide.
% See the LaTeX Companion or other references for full information.

%%% PAGE DIMENSIONS
\usepackage{geometry} % to change the page dimensions
\geometry{a4paper} % or letterpaper (US) or a5paper or....
% \geometry{margin=2in} % for example, change the margins to 2 inches all round
% \geometry{landscape} % set up the page for landscape
%   read geometry.pdf for detailed page layout information

\usepackage{graphicx} % support the \includegraphics command and options

% \usepackage[parfill]{parskip} % Activate to begin paragraphs with an empty line rather than an indent

%%% PACKAGES
\usepackage{booktabs} % for much better looking tables
\usepackage{array} % for better arrays (eg matrices) in maths
\usepackage{paralist} % very flexible & customisable lists (eg. enumerate/itemize, etc.)
\usepackage{verbatim} % adds environment for commenting out blocks of text & for better verbatim
\usepackage{subfig} % make it possible to include more than one captioned figure/table in a single float
% These packages are all incorporated in the memoir class to one degree or another...
\usepackage{amsmath}
\usepackage{amssymb}
\usepackage{amsthm}

\usepackage{algorithm}
%\usepackage{algorithmic}
\usepackage{algpseudocode}
\usepackage{xcolor}
%%% HEADERS & FOOTERS
\usepackage{fancyhdr} % This should be set AFTER setting up the page geometry
\pagestyle{fancy} % options: empty , plain , fancy
\renewcommand{\headrulewidth}{0pt} % customise the layout...
\lhead{}\chead{}\rhead{}
\lfoot{}\cfoot{\thepage}\rfoot{}

\newtheorem{theorem}{Theorem}[section]
\newtheorem{corollary}{Corollary}[theorem]
\newtheorem{lemma}[theorem]{Lemma}

% \algtext*{EndIf}
%%% SECTION TITLE APPEARANCE
\usepackage{sectsty}
\allsectionsfont{\sffamily\mdseries\upshape} % (See the fntguide.pdf for font help)
% (This matches ConTeXt defaults)

%%% ToC (table of contents) APPEARANCE
\usepackage[nottoc,notlof,notlot]{tocbibind} % Put the bibliography in the ToC
\usepackage[titles,subfigure]{tocloft} % Alter the style of the Table of Contents
\renewcommand{\cftsecfont}{\rmfamily\mdseries\upshape}
\renewcommand{\cftsecpagefont}{\rmfamily\mdseries\upshape} % No bold!

%%% END Article customizations

%%% The "real" document content comes below...

\title{Tiling Decision Trees Using Edge Probabilities}
\author{Ashwin Prasad, PhD CSA \\\texttt{ashwinprasad@iisc.ac.in}}
%\date{} % Activate to display a given date or no date (if empty),
         % otherwise the current date is printed

\begin{document}
\maketitle

\input{ProblemAndAlgorithm}

\end{document}

### 2022-02-25 src/include/DecisionForest.h
Add methods to write and read stat profiles

@@ -125,6 +125,9 @@ public:
     void SetClassId(int32_t classId) { m_classId = classId; }
     int32_t GetClassId() const { return m_classId; }
     int32_t NumFeatures();
+
+    void InitializeInternalNodeHitCounts();
+    int32_t GetSubtreeHitCount(int32_t nodeIndex);
 private:
     std::vector<Node> m_nodes;
     size_t m_numFeatures;
@@ -367,6 +370,20 @@ int32_t DecisionTree<ThresholdType, ReturnType, FeatureIndexType, NodeIndexType>
             featureSet.insert(node.featureIndex);
     }
     return featureSet.size();
+}
+
+template <typename ThresholdType, typename ReturnType, typename FeatureIndexType, typename NodeIndexType>
+void DecisionTree<ThresholdType, ReturnType, FeatureIndexType, NodeIndexType>::InitializeInternalNodeHitCounts() {
+    GetSubtreeHitCount(0);
+}
+
+template <typename ThresholdType, typename ReturnType, typename FeatureIndexType, typename NodeIndexType>
+int32_t DecisionTree<ThresholdType, ReturnType, FeatureIndexType, NodeIndexType>::GetSubtreeHitCount(int32_t nodeIndex) {
+    Node& node = m_nodes.at(nodeIndex);
+    if (node.IsLeaf())
+        return node.hitCount;
+    node.hitCount = GetSubtreeHitCount(node.leftChild) + GetSubtreeHitCount(node.rightChild);
+    return node.hitCount;
 }
 
 template <typename ThresholdType, typename ReturnType, typename FeatureIndexType, typename NodeIndexType>

### 2022-02-25 src/test/StatsTests.cpp

@@ -1,0 +1,140 @@
+#include <vector>
+#include <sstream>
+#include "Dialect.h"
+#include "TestUtilsCommon.h"
+
+#include "mlir/IR/Attributes.h"
+#include "mlir/IR/Builders.h"
+#include "mlir/IR/BuiltinOps.h"
+#include "mlir/IR/BuiltinTypes.h"
+#include "mlir/IR/MLIRContext.h"
+#include "mlir/IR/Verifier.h"
+#include "mlir/Dialect/StandardOps/IR/Ops.h"
+#include "mlir/Dialect/SCF/SCF.h"
+#include "llvm/ADT/STLExtras.h"
+
+#include "xgboostparser.h"
+#include "ExecutionHelpers.h"
+#include "TreeTilingDescriptor.h"
+#include "TreeTilingUtils.h"
+#include "ForestTestUtils.h"
+#include "CompileUtils.h"
+#include "StatsUtils.h"
+
+using namespace mlir;
+
+namespace TreeBeard
+{
+namespace test
+{
+
+decisionforest::DecisionForest<> ConstructForestAndRunInference(const std::string& modelJSONPath, const std::string& csvPath, int32_t numRows) {
+  // std::string csvPath = "/home/ashwin/ML/scikit-learn_bench/xgb_models/airline_xgb_model_save.json.test.csv";
+  mlir::MLIRContext context;
+  TreeBeard::XGBoostJSONParser<> xgBoostParser(context, modelJSONPath, "", 1);
+  xgBoostParser.Parse();
+  auto decisionForest = xgBoostParser.GetForest();
+
+  TreeBeard::test::TestCSVReader csvReader(csvPath);
+  // std::cerr << "Done reading csv..\n";
+  // std::cerr << "Running inference\n";
+  for (size_t i=0  ; i<csvReader.NumberOfRows() ; ++i ) {
+    auto row = csvReader.GetRowOfType<double>(i);
+    row.pop_back();
+    decisionForest->Predict(row);
+  }
+  return *decisionForest;
+}
+
+bool Test_XGBoostModel_StatGenerationAndReading(const std::string& modelJSON, const std::string& inputCSV, const std::string& statsCSV) {
+  TreeBeard::Profile::ComputeForestProbabilityProfile(modelJSON, inputCSV, statsCSV, -1);
+
+  mlir::MLIRContext context;
[…]

### 2022-02-25 src/test/TestMain.cpp

@@ -304,6 +304,15 @@ bool Test_SparseTileSize8_Epsilon_TestInputs_TiledSchedule(TestArgs_t &args);
 bool Test_SparseTileSize8_Higgs_TestInputs_TiledSchedule(TestArgs_t &args);
 bool Test_SparseTileSize8_Year_TestInputs_TiledSchedule(TestArgs_t &args);
 
+// Stats tests
+bool Test_AbaloneStatGenerationAndReading(TestArgs_t &args);
+bool Test_AirlineStatGenerationAndReading(TestArgs_t &args);
+bool Test_AirlineOHEStatGenerationAndReading(TestArgs_t &args);
+bool Test_CovtypeStatGenerationAndReading(TestArgs_t &args);
+bool Test_EpsilonStatGenerationAndReading(TestArgs_t &args);
+bool Test_HiggsStatGenerationAndReading(TestArgs_t &args);
+bool Test_YearStatGenerationAndReading(TestArgs_t &args);
+
 void InitializeVectorWithRandValues(std::vector<double>& vec) {
   for(size_t i=0 ; i<vec.size() ; ++i)
     vec[i] = (double)rand()/RAND_MAX;
@@ -1118,16 +1127,27 @@ TestDescriptor testList[] = {
   TEST_LIST_ENTRY(Test_SparseTileSize8_Epsilon_TestInputs_TiledSchedule),
   TEST_LIST_ENTRY(Test_SparseTileSize8_Higgs_TestInputs_TiledSchedule),
   TEST_LIST_ENTRY(Test_SparseTileSize8_Year_TestInputs_TiledSchedule),
+
+  // Stats tests
+  TEST_LIST_ENTRY(Test_AbaloneStatGenerationAndReading),
+  TEST_LIST_ENTRY(Test_AirlineStatGenerationAndReading),
+  TEST_LIST_ENTRY(Test_AirlineOHEStatGenerationAndReading),
+  TEST_LIST_ENTRY(Test_CovtypeStatGenerationAndReading),
+  TEST_LIST_ENTRY(Test_EpsilonStatGenerationAndReading),
+  TEST_LIST_ENTRY(Test_HiggsStatGenerationAndReading),
+  TEST_LIST_ENTRY(Test_YearStatGenerationAndReading),
 };
 
 #else // RUN_ALL_TESTS
 
 TestDescriptor testList[] = {
[…]

### 2022-02-25 src/utils/StatsUtils.cpp

@@ -3,11 +3,10 @@
 #include "xgboostparser.h"
 #include "TestUtilsCommon.h"
 
-namespace TreeBeard
+namespace
 {
-namespace Profile
-{
-void ComputeForestInferenceStats(const std::string& modelJSONPath, const std::string& csvPath, int32_t numRows) {
+
+void ComputeForestInferenceStatsImpl(const std::string& modelJSONPath, const std::string& csvPath, int32_t numRows, std::ostream& outputStream, bool sortLeaves) {
   // std::string csvPath = "/home/ashwin/ML/scikit-learn_bench/xgb_models/airline_xgb_model_save.json.test.csv";
   mlir::MLIRContext context;
   TreeBeard::XGBoostJSONParser<> xgBoostParser(context, modelJSONPath, "", 1);
@@ -15,16 +14,16 @@ void ComputeForestInferenceStats(const std::string& modelJSONPath, const std::st
   auto decisionForest = xgBoostParser.GetForest();
 
   TreeBeard::test::TestCSVReader csvReader(csvPath);
-  std::cerr << "Done reading csv..\n";
-  std::cerr << "Running inference\n";
+  // std::cerr << "Done reading csv..\n";
+  // std::cerr << "Running inference\n";
   for (size_t i=0  ; i<csvReader.NumberOfRows() ; ++i ) {
     auto row = csvReader.GetRowOfType<double>(i);
     row.pop_back();
     decisionForest->Predict(row);
-    if (i % (csvReader.NumberOfRows()/25) == 0)
-      std::cerr << "-" << std::flush;
+    // if (i % (csvReader.NumberOfRows()/25) == 0)
+    //   std::cerr << "-" << std::flush;
   }
-  std::cout << decisionForest->NumTrees() << ", " << csvReader.NumberOfRows() <<  std::endl;
+  outputStream << decisionForest->NumTrees() << ", " << csvReader.NumberOfRows() <<  std::endl;
   for (size_t i=0 ; i<decisionForest->NumTrees() ; ++i) {
     auto& tree = decisionForest->GetTree(i);
     std::vector<mlir::decisionforest::DecisionTree<>::Node> leaves;
@@ -32,15 +31,27 @@ void ComputeForestInferenceStats(const std::string& modelJSONPath, const std::st
       if (node.IsLeaf())
         leaves.push_back(node);
     }
[…]

### 2022-02-25 src/utils/StatsUtils.h

@@ -11,7 +11,9 @@ namespace Profile
 void ComputeForestInferenceStats(const std::string& modelJSONPath, const std::string& csvPath, int32_t numRows);
 void ComputeForestInferenceStatsOnSampledTestInput(const std::string& model, int32_t numRows);
 void ComputeForestInferenceStatsOnModel(const std::string& model, const std::string& csvPath, int32_t numRows);
+void ComputeForestProbabilityProfile(const std::string& modelJSONPath, const std::string& csvPath, const std::string& statsCSVPath, int32_t numRows);
 
+void ReadProbabilityProfile(mlir::decisionforest::DecisionForest<>& decisionForest, const std::string& statsCSVFile);
 }
 }
 

### 2022-04-08 src/utils/CompileUtils.h
Remove pipelineWidth

@@ -31,7 +31,6 @@ struct CompilerOptions {
 
   mlir::decisionforest::ScheduleManipulator *scheduleManipulator=nullptr;
   std::string statsProfileCSVPath = "";
-  int32_t pipelineWidth = -1;
   int32_t numberOfCores = -1;
 
   CompilerOptions() { }

### 2022-04-08 src/mlir/LowerToMidLevelIR.cpp
Xgboost benchmarks for pipelining + fix bug w/eeling

@@ -546,7 +546,7 @@ struct PredictForestOpLowering: public ConversionPattern {
   Value GeneratePipelinedTreeIndexLeafLoopBody(
     ConversionPatternRewriter &rewriter,
     Location location,
-    const decisionforest::IndexVariable& indexVar,
+    int32_t stepSize,
     std::list<Value> treeIndices,
     PredictOpLoweringState& state,
     Value row,
@@ -562,7 +562,7 @@ struct PredictForestOpLowering: public ConversionPattern {
     std::vector <Value> rows;
     std::vector <Value> trees;
     std::vector <Type> treeResultTypes;
-    for (int32_t i = 0; i < indexVar.GetRange().m_step; i++) {
+    for (int32_t i = 0; i < stepSize; i++) {
       treeIndices.push_back(rewriter.create<arith::ConstantIndexOp>(location, i));
       
       Value treeIndex = SumOfValues(rewriter, location, treeIndices);
@@ -640,7 +640,7 @@ struct PredictForestOpLowering: public ConversionPattern {
       scf::ForOp loop = rewriter.create<scf::ForOp>(location, startConst, stopConst, stepConst, ValueRange{ zeroConst });
       rewriter.setInsertionPointToStart(loop.getBody());
       treeIndices.push_back(loop.getInductionVar());
-      auto accumulatedValue = GeneratePipelinedTreeIndexLeafLoopBody(rewriter, location, indexVar, treeIndices, state, row, rowIndex, loop.getBody()->getArguments()[1]);
+      auto accumulatedValue = GeneratePipelinedTreeIndexLeafLoopBody(rewriter, location, range.m_step, treeIndices, state, row, rowIndex, loop.getBody()->getArguments()[1]);
       rewriter.create<scf::YieldOp>(location, static_cast<Value>(accumulatedValue));
       rewriter.setInsertionPointAfter(loop);
 
@@ -663,7 +663,7 @@ struct PredictForestOpLowering: public ConversionPattern {
         treeIndices.pop_back();
         treeIndices.push_back(peeledLoop.getInductionVar());
         
-        auto accumulatedValue = GeneratePipelinedTreeIndexLeafLoopBody(rewriter, location, indexVar, treeIndices, state, row, rowIndex, peeledLoop.getBody()->getArguments()[1]);
[…]

### 2022-04-08 src/test/XGBoostBenchmarks.cpp

@@ -57,18 +57,25 @@ int64_t Test_CodeGenForJSON_ProbabilityBasedTiling(int64_t batchSize, const std:
                                               const std::string& statsProfileCSV,
                                               int32_t tileSize, int32_t tileShapeBitWidth, 
                                               int32_t childIndexBitWidth, mlir::decisionforest::ScheduleManipulator *scheduleManipulator, 
-                                              bool probTiling, bool parallel) {
+                                              bool probTiling, bool parallel,
+                                              int32_t pipelineSize) {
   // TODO consider changing this so that you use the smallest possible type possible (need to make it a parameter)
   using FeatureIndexType = int16_t;
   using NodeIndexType = int16_t;
 
   mlir::MLIRContext context;
   int32_t floatTypeBitWidth = sizeof(FloatType)*8;
+  bool reorderTrees = probTiling || parallel || pipelineSize > 1;
   TreeBeard::CompilerOptions options(floatTypeBitWidth, sizeof(ReturnType)*8, IsFloatType(ReturnType()), sizeof(FeatureIndexType)*8, sizeof(NodeIndexType)*8,
                                      floatTypeBitWidth, batchSize, tileSize, tileShapeBitWidth, childIndexBitWidth,
                                      (probTiling ? TreeBeard::TilingType::kHybrid : TreeBeard::TilingType::kUniform), 
-                                     false, probTiling || parallel, ((probTiling || parallel) ? nullptr : scheduleManipulator));
+                                     pipelineSize > 1, // make all leaves same depth
+                                     reorderTrees, // reorder trees
+                                     reorderTrees ? nullptr : scheduleManipulator);
+
   options.statsProfileCSVPath = statsProfileCSV;
+  options.SetPipelineSize(pipelineSize);
+
   if (parallel)
     options.numberOfCores = GetNumberOfCores();
   TreeBeard::InitializeMLIRContext(context);
@@ -117,42 +124,51 @@ int64_t Test_CodeGenForJSON_ProbabilityBasedTiling(int64_t batchSize, const std:
 
[…]

### 2022-04-09 src/runtime/runtime.cpp
Add APIs for sparse rep and peeling

@@ -147,6 +147,14 @@ extern "C" int32_t IsSparseRepresentationEnabled() {
   return mlir::decisionforest::UseSparseTreeRepresentation;
 }
 
+extern "C" void SetPeeledCodeGenForProbabilityBasedTiling(int32_t val) {
+  mlir::decisionforest::PeeledCodeGenForProbabiltyBasedTiling = val;
+}
+
+extern "C" int32_t IsPeeledCodeGenForProbabilityBasedTilingEnabled() {
+  return mlir::decisionforest::PeeledCodeGenForProbabiltyBasedTiling;
+}
+
 // ===-------------------------------------------------------------=== //
 // Predefined Schedule Manipulation API
 // ===-------------------------------------------------------------=== //

### 2022-04-09 test/python/run_python_tests.py

@@ -119,15 +119,11 @@ def RunTestOnSingleModelTestInputs_Multibatch(modelName : str, returnType=numpy.
   csvPath = os.path.join(os.path.join(treebeard_repo_dir, "xgb_models"), modelName + "_xgb_model_save.json.test.sampled.csv")
   return RunSingleTest_Multibatch(soPath, globalsJSONPath, csvPath, returnType)
 
-def CompilerOptionsTest() -> Boolean:
-  model_name = "abalone"
-  compilerOptions = treebeard.CompilerOptions(20, 8)
-  jsonPath = os.path.join(os.path.join(treebeard_repo_dir, "xgb_models"), model_name + "_xgb_model_save.json")
-  tempPath = os.path.join(os.path.join(treebeard_repo_dir, "debug"), "temp")
-  globalsJSONPath = os.path.join(tempPath, "abalone_python_test.treebeard-globals.json")
-  llvmIRFile = os.path.join(tempPath, "abalone_python_test.ll")
-  treebeard.GenerateLLVMIRForXGBoostModel(jsonPath, llvmIRFile, globalsJSONPath, compilerOptions)
-  return True
+def SetStatsProfileCSVPath(options, modelName):
+  profilesDir = os.path.join(os.path.join(treebeard_repo_dir, "xgb_models"), "profiles") 
+  csvPath = os.path.join(profilesDir, modelName + ".test.csv")
+  options.SetStatsProfileCSVPath(csvPath)
+  return options
 
 defaultTileSize8Options = treebeard.CompilerOptions(200, 8)
 defaultTileSize8MulticlassOptions = treebeard.CompilerOptions(200, 8)
@@ -159,6 +155,35 @@ assert RunTestOnSingleModelTestInputsJIT("epsilon", invertLoopsTileSize8Options)
 assert RunTestOnSingleModelTestInputsJIT("higgs", invertLoopsTileSize8Options)
 assert RunTestOnSingleModelTestInputsJIT("letters", invertLoopsTileSize8MulticlassOptions, numpy.int8)
 assert RunTestOnSingleModelTestInputsJIT("year_prediction_msd", invertLoopsTileSize8Options)
+
+treebeard.SetEnableSparseRepresentation(1)
+assert RunTestOnSingleModelTestInputsJIT("abalone", invertLoopsTileSize8Options)
[…]

### 2022-04-11 docs/MICRO 2022/abstract.tex
Add paper draft

\begin{abstract}
Abstract goes here!
\end{abstract}

### 2022-04-11 docs/MICRO 2022/main.tex
Add paper draft

%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%
% This is the template for submission to MICRO 2022
% The cls file is modified from 'sig-alternate.cls'
%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%

\documentclass{sig-alternate}
\usepackage{mathptmx} % This is Times font
\usepackage{listings}
\usepackage{fancyhdr}
\usepackage{xcolor}
\usepackage[normalem]{ulem}
\usepackage[hyphens]{url}
\usepackage[sort,nocompress]{cite}
\usepackage[final]{microtype}
\usepackage[keeplastbox]{flushend}
\usepackage{subfig}
% Always include hyperref last
\usepackage[bookmarks=true,breaklinks=true,letterpaper=true,colorlinks,citecolor=blue,linkcolor=blue,urlcolor=blue]{hyperref}

\newcommand{\CommentOut}[1]{}
\newcommand\TODO[1]{\textcolor{red}{TODO #1}}

\definecolor{codegreen}{rgb}{0,0.6,0}
\definecolor{codegray}{rgb}{0.5,0.5,0.5}
\definecolor{codepurple}{rgb}{0.58,0,0.82}
\definecolor{backcolour}{rgb}{0.95,0.95,0.92}

\lstdefinestyle{mystyle}{
    backgroundcolor=\color{backcolour},   
    commentstyle=\color{codegreen},
    keywordstyle=\color{magenta},
    numberstyle=\tiny\color{codegray},
    stringstyle=\color{codepurple},
    basicstyle=\scriptsize,
    breakatwhitespace=false,         
    breaklines=true,                 
    captionpos=b,                    
    keepspaces=true,                 
    numbers=left,                    
    numbersep=5pt,                  
    showspaces=false,                
    showstringspaces=false,
    showtabs=false,                  
    tabsize=2
}

\lstset{style=mystyle}

% Ensure letter paper
\pdfpagewidth=8.5in
\pdfpageheight=11in

%%%%%%%%%%%---SETME-----%%%%%%%%%%%%%
\newcommand{\microsubmissionnumber}{XXX}
%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%

\fancypagestyle{firstpage}{
  \fancyhf{}
  \renewcommand{\headrulewidth}{0pt}
  \fancyhead[C]{\vspace{10pt}\normalsize{MICRO 2022 Submission
      \textbf{\#\microsubmissionnumber} -- Confidential Draft -- Do NOT Distribute!!}\\\vspace{-25pt}} 
  \fancyfoot[C]{\thepage}
}

\pagenumbering{arabic}

%%%%%%%%%%%---SETME-----%%%%%%%%%%%%%
\title{Treebeard : A Compiler for Decision Forest Inference} 
%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%

\begin{document}
\maketitle
\thispagestyle{firstpage}
\pagestyle{plain}

%%%%%% -- PAPER CONTENT STARTS-- %%%%%%%%
\input{abstract}
\input{introduction}
\input{overview}

\CommentOut{
\begin{abstract}

  This document is intended to serve as a sample for submissions to the 55\textsuperscript{th} IEEE/ACM International Symposium on Microarchitecture\textsuperscript{\textregistered} (MICRO 2022). We provide some guidelines that authors should follow when submitting papers to the conference.  This format is derived from the ACM sig-alternate.cls file, and is used with an objective of keeping the submission version similar to the camera-ready version.

\end{abstract}

\section{Introduction}

This document provides instructions for submitting papers to the 55\textsuperscript{th} IEEE/ACM International Symposium on Microarchitecture\textsuperscript{\textregistered} (MICRO 2022).  In an effort to respect the efforts of reviewers and in the interest of fairness to all prospective authors, we request that all submissions to MICRO 2022 follow the formatting and submission rules detailed below. Submissions that violate these instructions may not be reviewed, at the discretion of the program chairs, in order to maintain a review process that is fair to all potential authors.

This document is itself formatted using the MICRO 2022 submission format. The content of this document mirrors that of the submission instructions that appear on the conference website. All questions regarding paper formatting and submission should be directed to the program chairs.

\subsection{Format Highlights}
\begin{itemize}
\item Paper must be submitted in printable PDF format.
\item Text must be in a minimum 10pt font, see Table~\ref{table:formatting}.
\item Papers must be at most 11 pages, not including references.
\item No page limit for references.
\item Each reference must specify {\em all} authors (no {\em et al.}).
\item Author anonymity must be fully preserved, including in any referenced artifacts (e.g., GitHub repository).
\end{itemize}

\subsection{Paper Evaluation Objectives} 
The committee will make every effort to judge each submitted paper on its own merits. There will be no target acceptance rate. We expect to accept a wide range of papers with appropriate expectations for evaluation---while papers that build on significant past work with strong evaluations are valuable, papers that open new areas with less rigorous evaluation are equally welcome and especially encouraged.

\section{Paper Preparation Instructions}

\subsection{Paper Formatting}

Papers must be submitted in printable PDF format and should contain a {\em maximum of 11 pages} of single-spaced two-column text, {\bf not including references}.  You may include any number of pages for references, but see below for more instructions.  If you are using \LaTeX~\cite{lamport94} to typeset your paper, then we suggest that you use the template here: \href{https://www.microarch.org/micro55/submit/micro55-latex-template.zip}{\LaTeX~Template}. This document was prepared with that template. Note that the template and sample paper may render slightly differently on different \LaTeX~engines, due to typesetting changes between versions. If you use a different software package to typeset your paper, then please adhere to the guidelines given in Table~\ref{table:formatting}.

\begin{scriptsize}
\begin{table}[h!]
  \centering
  \begin{tabular}{|l|l|}
    \hline
    \textbf{Field} & \textbf{Value}\\
    \hline
    \hline
    File format & PDF \\
    \hline
    Page limit & 11 pages, {\bf not including}\\
               & {\bf references}\\
    \hline
    Paper size & US Letter 8.5in $\times$ 11in\\
    \hline
    Top margin & 1in\\
    \hline
    Bottom margin & 1in\\
    \hline
    Left margin & 0.75in\\
    \hline
    Right margin & 0.75in\\
    \hline
    Body & 2-column, single-spaced\\
    \hline
    Space between columns & 0.25in\\
    \hline
    Line spacing (leading) & 11pt \\
    \hline
    Body font & 10pt, Times\\
    \hline
    Abstract font & 10pt, Times\\
    \hline
    Section heading font & 12pt, bold\\
    \hline
    Subsection heading font & 10pt, bold\\
    \hline
    Caption font & 9pt (minimum), bold\\
    \hline
    References & 8pt, no page limit, list \\
               & all authors' names\\
    \hline
  \end{tabular}
  \caption{Formatting guidelines for submission.}
  \label{table:formatting}
\end{table}
\end{scriptsize}

{\em Please ensure that you include page numbers with your submission}. This makes it easier for the reviewers to refer to different parts of your paper when they provide comments. Please ensure that your submission has a banner at the top of the title page, similar to this document, which contains the submission number and the notice of confidentiality.  If using the template, just replace XXX with your submission number.

\subsection{Content}

Reviewing will be {\em double blind} (no author list); therefore, please do not include any author names on any submitted documents except in the space provided on the submission form.  You must also ensure that the metadata included in the PDF does not give away the authors. You must fully anonymize any links to artifacts (e.g., GitHub repository) and remove any links to artifacts that cannot be fully anonymized. {\bf Papers that violate the anonymization policy may be rejected without review} at the discretion of the program chairs.

If you are improving upon your prior work, refer to your prior work in the third person and include a full citation for the work in the bibliography.  For example, if you are building on {\em your own} prior work in the papers \cite{nicepaper1,nicepaper2,nicepaper3}, you would say something like: "While the authors of \cite{nicepaper1,nicepaper2,nicepaper3} did X, Y, and Z, this paper additionally does W, and is therefore much better."  Do NOT omit or anonymize references for blind review.  There is one exception to this for your own prior work that appeared in IEEE CAL, arXiv, workshops without archived proceedings, etc.\, as discussed later in this document.

\noindent\textbf{Figures and Tables:} Ensure that the figures and tables are legible.  Please also ensure that you refer to your figures in the main text.  Many reviewers print the papers in gray-scale. Therefore, if you use colors for your figures, ensure that the different colors are highly distinguishable in gray-scale.

\noindent\textbf{References:}  There is no length limit for references. {\em Each reference must explicitly list all authors of the paper.  Papers not meeting this requirement will be rejected.} Authors of NSF proposals should be familiar with this requirement. Knowing all authors of related work will help find the best reviewers. Since there is no length limit for the number of pages used for references, there is no need to save space here.

\section{Paper Submission Instructions}

\subsection{Guidelines for Determining Authorship}
IEEE guidelines dictate that authorship should be based on a {\em substantial intellectual contribution}. It is assumed that all authors have had a significant role in the creation of an article that bears their names. In particular, the authorship credit must be reserved only for individuals who have met each of the following conditions:

\begin{enumerate}
\item Made a significant intellectual contribution to the theoretical development, system or experimental design, prototype development, and/or the analysis and interpretation of data associated with the work contained in the article;

\item Contributed to drafting the article or reviewing and/or revising it for intellectual content; and

\item Approved the final version of the article as accepted for publication, including references.
\end{enumerate}

A detailed description of the IEEE authorship guidelines and responsibilities is available \href{https://www.ieee.org/publications_standards/publications/rights/Section821.html}{here}. Per these guidelines, it is not acceptable to award {\em honorary } authorship or {\em gift} authorship. Please keep these guidelines in mind while determining the author list of your paper.

\subsection{Declaring Authors}
Declare all the authors of the paper upfront. Addition/removal of authors once the paper is accepted will have to be approved by the program chairs, since it potentially undermines the goal of eliminating conflicts for reviewer assignment.

\subsection{Areas and Topics}
Authors should indicate these areas on the submission form as well as specific topics covered by the paper for optimal reviewer match. If you are unsure whether your paper falls within the scope of MICRO, please check with the program chairs -- MICRO is a broad, multidisciplinary conference and encourages new topics.

\subsection{Revision of Previously-Reviewed\\ Manuscript}
If the manuscript has been previously reviewed and rejected and is now being submitted to MICRO, the authors have an option of providing a letter explaining how the paper has been revised for this current submission. We expect this revision information to improve both the submission and the review process. Authors choosing to provide such a letter have control over who has access to it by specifying one of the following options:

\begin{enumerate}
\item Shared with all reviewers of the paper 
\item Shared with reviewers who declare that they reviewed a prior version and who request the revision information
\item Not shared with any PC member but available to the program chairs
\end{enumerate}

We encourage you to keep this letter concise and optionally append additional information, such as a version of the paper that highlights the differences or any other material of your choice.

\subsection{Declaring Conflicts of Interest}
Authors must register all their conflicts for their paper submission. Conflicts are needed to ensure appropriate assignment of reviewers. {\bf If a paper is found to have an undeclared conflict that causes a problem OR if a paper is found to declare false conflicts in order to abuse or ``game'' the review system, the paper may be rejected without review.} We use the following conflict of interest guidelines for determining the conflict period for MICRO 2022.  Please declare a conflict of interest (COI) with the following people for any author of your paper:

\begin{enumerate}
\item Your Ph.D. advisor(s), post-doctoral advisor(s), Ph.D. students,
      and post-doctoral advisees, forever.
\item Family relations by blood or marriage, or their equivalent,
      forever (if they might be potential reviewers).
\item People with whom you have collaborated in the last FOUR years, including:
  \begin{itemize}
  \item co-authors of accepted/rejected/pending papers
  \item co-PIs on accepted/rejected/pending grant proposals
  \end{itemize}
\item Ongoing collaboration that has not yet resulted in a paper or proposal submission. Justification may be queried.
\item When there is a direct funding relationship between an author and the potential reviewer (e.g., the reviewer is a sponsor of an author's research on behalf of his/her company or vice versa) in the last FOUR years.
\item People (including students) who shared your primary institution(s) in the last FOUR years.
\item Other relationships, such as close personal friendship, that you think might tend
to affect your judgment or be seen as doing so by a reasonable person familiar
with the relationship.
\end{enumerate}

We would also like to emphasize that the following scenarios do {\em not} constitute a conflict:
\begin{enumerate}
\item Authors of previously-published, closely related work on that basis alone.
\item ``Service'' collaborations such as co-authoring a report for a professional organization, serving on a program committee, or co-presenting tutorials.
\item Co-authoring a paper that is a compendium of various projects with no true collaboration among the projects.
\item People who work on topics similar to or related to those in your papers.
\item Collaborators on large funded projects where there is no close collaboration and no joint benefit in the paper being accepted.
\end{enumerate}

We hope to draw most reviewers from the program committee, but others
from the community may also write reviews. {\bf Please declare all your conflicts (not just restricted to the PC).} When in doubt, please contact the program chairs.

%Please note that all paper submissions require all authors to electronically sign a statement confirming their best effort to accurately identify potential reviewers with a conflict of interest, and importantly also {\bf assuring that each author will make no explicit attempt to directly or indirectly influence any reviewer opinion or decision about the submitted paper}. Importantly, we do not consider technical discussion of a paper's content or any other sharing of content from the paper to violate the above policy.

\subsection{Concurrent Submissions and Workshops}

By submitting a manuscript to MICRO 2022, the authors guarantee that the manuscript has not been previously published or accepted for publication in a substantially similar form in any conference, journal, or the archived proceedings of a workshop (e.g., in the ACM/IEEE digital library) -- see exceptions below. The authors also guarantee that no paper that contains significant overlap with the contributions of the submitted paper will be under review for any other conference or journal or an archived proceedings of a workshop during the MICRO 2022 review period. Violation of any of these conditions will lead to rejection.

The only exceptions to the above rules are for the authors' own papers in (1) workshops without archived proceedings such as in the ACM/IEEE digital library (or where the authors chose not to have their paper appear in the archived proceedings), or (2) venues such as IEEE CAL or arXiv where there is an explicit policy that such publication does not preclude longer conference submissions.  In all such cases, the submitted manuscript may ignore the above work to preserve author anonymity. This information must, however, be provided on the submission form -- the program chairs will make this information available to reviewers if it becomes necessary to ensure a fair review.  As always, if you are in doubt, it is best to contact the program chairs.

Finally, the ACM/IEEE Plagiarism Policy (\href{http://www.acm.org/publications/policies/plagiarism_policy}{here} and \href{https://www.ieee.org/publications_standards/publications/rights/plagiarism.html}{here}) covers a range of ethical issues concerning the misrepresentation of other works or one's own work.

\section{Ethics}

\begin{enumerate}
\item Authors must abide by the ACM code of ethics and the IEEE code of ethics
\item Authors must not contact reviewers or PC members about any submission, including their own. This includes attempting to sway a reviewer, requesting information about any aspect of the reviewing process, and/or asking about the outcome of a submission. Similarly, authors are not allowed to ask another party to contact the reviewers on their behalf.
\item Authors must not disclose the content of reviews for their paper publicly (e.g., on social media)  before the results are announced. 
\item Authors must report any allegations of submission or reviewing misconduct to the program chairs. The only exception is if the complaint is about the program chairs; in this case, the Steering Committee should be contacted. 
\end{enumerate}

\section*{ACKNOWLEDGMENTS}
This document is derived from previous conferences, in particular MICRO 2013, ASPLOS 2015, MICRO 2015, MICRO 2016, MICRO 2017, MICRO
2018, MICRO 2019, MICRO 2020 and MICRO 2021, as well as SIGARCH/TCCA's Recommended Best Practices for the Conference Reviewing Process.
}

%%%%%%% -- PAPER CONTENT ENDS -- %%%%%%%%

%%%%%%%%% -- BIB STYLE AND FILE -- %%%%%%%%
% \bibliographystyle{IEEEtranS}
% \bibliography{refs}
%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%

\end{document}

### 2022-04-11 docs/MICRO 2022/overview.tex
Add paper draft

\section{Compiler Overview}
Figure \ref{Fig:CompilerStructure} shows the high level structure of Treebeard. The input to the compiler is a JSON file that describes the decision forest model and a set of options. The output of the compiler is a callable function that takes an array of rows and returns an array containing the model's prediction for each row. \TODO{Add a diagram, including IR stages}. The specified set of options includes information such as the batch size, the type of the input features, the type for node thresholds and the type for feature indices \TODO{There are also several optimization related inputs like tile size, type of tiling, pipeline depth etc. Should we mention those?}. Treebeard is written as a dialect in the MLIR framework \cite{MLIR}. At a high level, Treebeard first performs transformations on the trees in the model and subsequently generates and performs optimizations on a more traditional loop based IR. From an implementation perspective, Treebeard first constructs a high level representation of the decision forest inference operation and then progressively optimizes and lowers it to LLVM IR \cite{LLVM}. LLVM is then used to JIT compile the generated IR to executable code. The following sections describe each level of the IR and their lowering in more detail.

\begin{figure}
  \centering
  \includegraphics[scale = 0.3]{figures/CompilerStructure.PNG}
  \caption{Treebeard Compiler Structure}
  \label{Fig:CompilerStructure}
\end{figure}

\TODO{Should we describe the dialect's type system?}

\subsection{High Level IR}
Treebeard parses the input JSON file and generates a function with a single MLIR operation, \texttt{predictForest} that represents inference using the input model on a set of rows. The operation contains within it a tree based representation of the model that can be manipulated by optimizing transformations. Transformations on the model such as tiling, tree reordering and leaf padding are done at this level. The structure of the loop nest to walk the iteration space of trees and inputs is also decided at this level of the IR. \TODO{Should we mention that there is a scheduling language to decide this?}

\begin{lstlisting}[language=C++]
func Predict(float rows[batchSize]) {
  predictions = predictForest(rows) 
  return rows
}
\end{lstlisting}

\subsection{Mid Level IR}
The Mid Level IR makes the loop structures and tree walks more explicit. Firstly, the order in which the iteration space of trees and inputs is walked is explicitly specified in the IR through loop nests. Also, tree specific operations such as \texttt{isLeaf}, \texttt{traverseTile}, \texttt{getLeafValue} are introduced so that the traversal of trees explicitly represented. The following listing shows the IR for inference using a model with four trees on an input batch with two rows. The listed IR walks all trees for one input row before moving to the next row. One important point to note here is that details such as the data structure used for the trees are not explicitly encoded in the IR. This allows us to reuse optimization and lowering passes on this level of the IR regardless of what the final in memory representation of the model is.

\begin{lstlisting}[language=C++]
  // Constant that represents the model being compiled
  forest = ensemble(...)
  for i = 0 to 2 step 1 {
    prediction = 0
    for t = 0 to 4 step 1 {
      tree = getTree(forest, t) 
      node = getRoot(tree)
      while (isLeaf(tree, n)==false)  do {
        node = traverseTreeTile(tree, node, rows[i])
      }
      treePrediction = getLeafValue(tree, node)
      prediction = prediction + treePrediction
    }
    predictions[i] = prediction
  }
\end{lstlisting}

The IR listed above is a simplification of the actual IR. The actual IR is strongly typed and in SSA form.

\subsection{Low Level IR}
The IR is finally lowered to a form where the in memory representation of the model is made explicit. Buffers to hold the model values are inserted into the generated code and all tree operations in the mid-level IR are lowered to explicitly reference these buffers. The semantics of all operations are made explicit. For example, \texttt{traverseTreeTile} is lowered into a series of operations to load thresholds, feature indices and features, compare the features with the thresholds and compute the next node to evaluate. This IR is then lowered directly to LLVM IR and JITted.

### 2022-04-12 docs/MICRO 2022/overview.tex
overview pass

@@
 Figure \ref{Fig:CompilerStructure} shows the high level structure of Treebeard. 
-The input to the compiler is a JSON file that describes the decision forest model and a set of options.
-The output of the compiler is a callable function that takes an array of rows and returns an array containing the model's prediction for each row.
-\TODO{Add a diagram, including IR stages}.
-The specified set of options includes information such as the batch size, the type of the input features, the type for node thresholds and the type for feature indices \TODO{There are also several optimization related inputs like tile size, type of tiling, pipeline depth etc.
+The input to TreeBeard is a JSON serialized decision forest model, it supports popular frameworks like XGBoost and LightGBM and is extensible to other frameworks.
+Given an input model our compiler generates optimized inference code.
+Specifically it generates a callable batch inference function \texttt{predictForest} that given an array of inputs, outputs an array of model predictions.
+
+TreeBeard specializes the code generated for inference by progressively optimizing and lowering a high level representation of the \texttt{predictForest} function down to LLVM IR \cite{LLVM}.
+It is implemented as a dialect in the MLIR framework \cite{MLIR},
+%Treebeard first constructs a high level representation of the decision forest inference operation. 
+ At this level (HIR) it applies transformations that determine how inputs traverse the decision forest.
+It re-order trees in the forest and determines the high level loop structure (process one tree at a time or one input at a time).
+It then lowers to a traditional loop based IR (MIR).
+Several loop level optimizations like unrolling and pipelining are applied at this level.
+At the lowest level of IR TreeBeard applies layout optimizations and generates parallel code in LLVM IR.
+LLVM is then used to JIT compile the generated IR to executable code. The following sections describe each level of the IR and their lowering in more detail.
+
+ %The specified set of options includes information such as the batch size, the type of the input features, the type for node thresholds and the type for feature indices \TODO{There are also several optimization related inputs like tile size, type of tiling, pipeline depth etc.
 Should we mention those?}.
-Treebeard is written as a dialect in the MLIR framework \cite{MLIR}.
-At a high level, Treebeard first performs transformations on the trees in the model and subsequently generates and performs optimizations on a more traditional loop based IR.
-From an implementation perspective, Treebeard first constructs a high level representation of the decision forest inference operation and then progressively optimizes and lowers it to LLVM IR \cite{LLVM}.
-LLVM is then used to JIT compile the generated IR to executable code.
-The following sections describe each level of the IR and their lowering in more detail.
 
@@
   \centering
-  \includegraphics[scale = 0.3]{figures/CompilerStructure.PNG}
+  \includegraphics[width=\linewidth]{figures/compiler.pdf}
   \caption{Treebeard Compiler Structure}
@@
 \TODO{Should we describe the dialect's type system?}
-
+\TODO{Kr : consider focusing on the example instead of verbose description of optimizations. This section can be short, descriptions can come in later sections.}
 \subsection{High Level IR}

### 2022-04-12 docs/MICRO 2022/details.tex
Add tiling section

\section{Optimizations}

\TODO{This section needs a better name!}
\input{tiling}

### 2022-04-12 docs/MICRO 2022/main.tex
Add tiling section

@@
 
-\lstdefinestyle{mystyle}{
+\lstdefinestyle{c++}{language=C++,
+    morekeywords={ensemble, getTree, getRoot, isLeaf, traverseTreeTile, getLeafValue, loadThresholds, loadFeatureIndices, loadTileShape, getChildNode},    
     backgroundcolor=\color{backcolour},   
@@
 
-\lstset{style=mystyle}
+\lstset{style=c++}
 
@@
 \input{overview}
+\input{details}
 

### 2022-04-12 docs/MICRO 2022/overview.tex
Add tiling section

@@
 Firstly, the order in which the iteration space of trees and inputs is walked is explicitly specified in the IR through loop nests.
-Also, tree specific operations such as \texttt{isLeaf}, \texttt{traverseTile}, \texttt{getLeafValue} are introduced so that the traversal of trees explicitly represented.
+Also, operations such as \texttt{isLeaf}, \texttt{traverseTile}, \texttt{getLeafValue} are introduced so that the traversal of trees explicitly represented.
 The following listing shows the IR for inference using a model with four trees on an input batch with two rows.
@@
 
-\begin{lstlisting}[language=C++]
+\begin{lstlisting}[style=c++]
   // Constant that represents the model being compiled

### 2022-04-12 docs/MICRO 2022/tiling.tex
Add tiling section

\subsection{Notation}
\TODO{Notation needs to be introduced in the background section}
We represent a decision tree by $T = (V, E, r)$ where $V$ is the set of nodes, $E$ the set of edges and
$r \in V$ is the root. For each node $n \in V$, the following are given.
\begin{enumerate}
    \item $threshold(n) \in \mathbb{R}$ which gives the threshold value for $n$.
    \item $featureIndex(n) \in \mathbb{N}$ which gives the feature index for $n$.
    \item $left(n) \in V$, the left child of $n$ or $\emptyset$ if $n$ is a leaf. If $left(n) \neq \emptyset$, then $(n, left(n)) \in E$.
    \item $right(n) \in V$, the right child of $n$ or $\emptyset$ if $n$ is a leaf. If $right(n) \neq \emptyset$, then $(n, right(n)) \in E$.
\end{enumerate}
We use $L \subset V$ to denote the set of leaves.

\subsection{Tiling}

Treebeard vectorizes tree walks by grouping nodes of a decision tree into \textbf{\emph{tiles}}. The nodes in a tile are evaluated concurrently using vector instructions. Once the nodes of the current tile are evaluated, a look up table is used to compute which child of the current tile to move to next. The listing below shows at a high level how a tiled tree is walked. 
\begin{lstlisting}[style=c++]
  // A lookup table that determines the child index of
  // the next tile given the tile shape and the outcome
  // of the vector comparison on the current tile
  int16_t LUT[NUM_TILE_SHAPES, pow(2, TileSize)]

  ResultType Prediction_Function(...) {
    // ...
    Node n = getRoot(tree)
    while (isLeaf(tree, n)==false) do {
      thresholds = loadThresholds(tree, n)
      featureIndices = loadFeatureIndices(tree, n)
      // Gather required feature from the current row
      features = rows[i][featureIndices]
      // Vector comparison of features and thresholds
      comparison = features < thresholds

      // Pack bits in comparison vector into an integer
      comparisonIndex = combineBitsIntoInt(comparison)

      // Get child index of tile we need to move to
      tileShape = loadTileShape(tree, n)
      childIndex = LUT[tileShapeID, comparisonIndex]

      // Move to the correct child of the current node
      n = getChildNode(tree, n, childIndex) 
    }
    ThresholdType prediction = getLeafValue(n)
    // ...
  }  
\end{lstlisting}

To evaluate the current tile, the vector of thresholds is first loaded (\texttt{loadThresholds}). This vector contains the thresholds of all nodes in the tile. Then, the features required for comparison are gathered into a vector (lines 11 and 13). The feature vector is compared to the threshold vector and the child tile to move to is determined (lines 15 to 25). More details about tile shapes and the look up table are presented in subsequent sections.

\subsection{Tiles and Tile Shapes}
A \textbf{\emph{tile}} is a collection of connected non-leaf nodes of a decision tree. The path connecting any pair of nodes in the tile must fully be contained within the tile. The tile size $n_t$ is the number of nodes contained in a tile.

Informally, the \textbf{\emph{tile shape}} is the shape of the region that encloses all nodes in a tile in a diagram of the decision tree. More formally, for a tile size $n_t$, each unique legal binary tree containing $n_t$ nodes (nodes being indistinguishable) corresponds to a tile shape.

Figure \ref{Fig:TileSize3Shapes} enumerates all tile shapes with a tile size of 3. There are a total of 5 tile shapes with size 3. The number of tile shapes with a tile size $n_t$, denoted by $NTS(n_t)$ is given by the following equation.

\begin{equation}
  NTS(n) = \sum_{k=0}^{n-1} NTS(k) \times NTS(n-k-1)
\end{equation}

where $NTS(0) = NTS(1) = 1$.

\begin{figure}
  \centering
  \includegraphics[width=\linewidth]{figures/TileShapes_Size3.PNG}
  \caption{All possible tile shapes with a tile size $n_t=3$}
  \label{Fig:TileSize3Shapes}
\end{figure}

\subsection{Valid Tiling of a Tree}
A tiling $\mathcal{T}$ of the tree $T = (V, E, r)$ with tile size $n_t$ is a partition $\{ T_1, T_2, ... ,T_m \}$ of the set $V$ such that 
\begin{enumerate}
    \item $T_1 \cup T_2 ... \cup T_m = V$
    \item $T_i \cap T_j = \emptyset$ for all $i, j \in [1, m]$ and $i \neq j$
    \item $|T_i| \leq n_t$ for all $i \in [1, m]$
    \item $\forall l \in L$ : $l \in T_i \rightarrow v \notin T_i \;\; \forall v \in V \backslash \{l\}$
    \item Tiles are \textbf{maximal}, i.e. if $|T_i| < n_t$, then there is no $v \in V\backslash \{ T_i \cup L \}$ such that $(u, v) \in E$ for some $u \in T_i$. 
    \item Tiles are \textbf{connected}, i.e. for an $u, v \in T_i$, there is a (undirected) path connecting $u$ and $v$ fully contained in $T_i$.
\end{enumerate}

\TODO{AP We need to talk about how tiling is specified in the compiler}

\subsection{Tiled Trees}
A tiling transformation communicates the tiling to the Treebeard infrastructure by assigning a tile ID to each node in the decision tree. Using these tile IDs, Treebeard checks the validity of the tiling and then contructs a tree whose nodes are tiles. We call this tree the \textbf{\emph{tree of tiles}}. \TODO{We need a better name for this}
Figure \ref{Fig:ValidTilingTileSize3} shows a valid tiling with tile size 3 and the tree of tiles constructed by Treebeard. Three nodes are grouped into each of the tiles $t_1$ and $t_2$ as shown. Each tile is collapsed into a single node in the tree of tiles. However, each leaf in the original tree becomes a leaf in the tree of tiles.

\begin{figure}
  \centering
  \includegraphics[width=\linewidth]{figures/TiledTree_Size3.PNG}
  \caption{Example of a valid tree tiling with a tile size $n_t=3$}
  \label{Fig:ValidTilingTileSize3}
\end{figure}

Treebeard maintains the following invariants.
\begin{enumerate}
  \item All tiles in a tree are the same size $n_t$. If the tiling produces any smaller tiles, these are padded by inserting dummy nodes to make them the required size.
  \item Nodes within tiles are always ordered in level order and left to right within a level. The numbering of the nodes in the above diagram shows this node order.
  \item Children of a node are numbered from left to right (regardless of level). For example, $l_1$ is the first child of $t_1$, $l_2$ is the second and so on.
\end{enumerate}

\subsection{Tile Shapes and Decision Tree Inference}
\label{Sec:TileShapesAndDecisionTreeInference}
Treebeard uses vector instructions to accelerate decision tree walks. Vector instructions are used to evaluate the predicates of all the nodes in a tile simultaneously. However, once the predicates of all the nodes in the tile are evaluated, computing the next tile to move to, given the outcome of the comparison depends on the tile shape of the current tile. To illustrate this problem, consider the case of the tiles of size 3 shown in figure \ref{Fig:TileTraversalTileSize3}. 
\begin{figure}
  \centering
  \includegraphics[width=\linewidth]{figures/TileTraversal_Size3.PNG}
  \caption{Example tile traversals with tile size $n_t=3$}
  \label{Fig:TileTraversalTileSize3}
\end{figure}
The diagram shows 3 of the 5 possible tile shapes for a tile size of 3. The nodes drawn in black are members of the tile $t_1$. The nodes in blue are the entry nodes of the children tiles of $t_1$. \TODO{Define entry nodes}

To traverse a tile on an input row, first, the predicate of each node in the tile is computed. Subsequently, we need to determine which of the child tiles to move to next. Note that a true predicte (bit value 1) on a node implies a move to the left child and a false predicate (bit value 0) implies a move to the right child.

In the diagram, the bit strings (written in red) show which child we need to move to given the outcomes of the comparison (the bits represent the comparison outcomes of nodes and are in the order of the nodes in the tile -- marked 0, 1 and 2 in the diagram, i.e., the MSB is the predicate outcome of node 0 and the LSB the predicate outcome of node 2). For example, for the first tile shape, if the predicate of all nodes are true (i.e. the comparison outcome is 111), the next node to evaluate is $a$. However, if the predicate of node 1 is false, then we need to move to $d$ regardless of the outcomes of nodes 2 and 3. It is easy to see from the diagram that, depending on the tile shape, the same predicate outcomes can mean moving to different children. For example, for the outcome "011", the next tile is the 4th child (node $d$) for the first two tile shapes while it is the 3rd child for the other tile shape (node $c$).

\subsection{Lookup Table}
A lookup table (LUT) is used to solve the problem described in section \ref{Sec:TileShapesAndDecisionTreeInference}, i.e. given the outcome of the comparisons on all nodes in a tile, determine the child tile we should evaluate next. The LUT is indexed by the tile shape and the comparison outcome. Formally, the LUT is a map.
\[
LUT : (TileShape, < Boolean \times n_t >) \rightarrow [0, n_t] \subset \mathbb{N}
\]

where $n_t$ is the tile size, $< Boolean \times n_t >$ is a vector of $n_t$ booleans. The value returned by the LUT is the index of the child of the current tile that should be evaluated next. For example, if we are evaluating the first tile $t$ in figure \ref{Fig:TileTraversalTileSize3}, and the result of the comparison is 110, then $LUT(TileShape(t), 110)=1$ since the tile we need to evaluate next is the tile with node $b$, which is the second child of the current tile.

In order to realize this LUT in generated code, Treebeard associates a non-negative integer ID with every unique tile shape of the given tile size. The result of the comparison, a vector of booleans, can be interpreted as a 64-bit integer. Therefore, the LUT can be implemented as a 2 dimensional array.
\begin{lstlisting}{style=c++}
  int16_t LUT[NTS(n_t), pow(2, n_t)]  
\end{lstlisting}

Treebeard computes the values in the LUT statically as the tile size is a compile time constant.

\subsection{In Memory Representation of Tiled Trees}
Treebeard currently has two in memory representations for tiled trees - an array based representation and a sparse representation. Both representations use an array of structs to represent all tiles of the model.

\subsubsection{Array Based Representation}
\label{sec:ArrayBased}
Each tree in the model is represented as an array of tiles using the standard representation of trees as arrays. The root node is at index 0 and for a node at index $n$ in the array, the index of its $i^{th}$ child is given by $(n_t + 1) n + (i + 1)$ (every node in the tree of tiles has $n_t + 1$ children). A tile is represented by an object of the following struct.
\begin{lstlisting}{style=c++}
  struct Tile {
    // A vector of TileSize elements
    <ThresholdType x TileSize> thresholds; 
    <FeatureIndexType x TileSize> featureIndices;
    // Integer that identifies the tile shape
    TileShapeIDType tileShapeID; 
  };  
\end{lstlisting}
\TODO{AP Is this level of detail really needed? Also, the vector type notation needs to be introduced somewhere.}
Even though this representation is simple, the memory required for reasonable sized models is very large. The memory footprint is up to 20X that of the scalar representation. This memory bloat causes a performance degradation because the span of the L1 TLB is not sufficient to efficiently translate addresses for the whole model. \TODO{AP There are also some cache misses. Write this better.} Storing leaves as full tiles (even though leaves just have to represent one value) and the empty space introduced due to the array based representation of trees that are not complete account for most of the increase. The sparse representation described next tries to address these issues.

\subsubsection{Sparse Representation}
\label{sec:SparseRep}
The sparse representation tries to address the large memory footprint of the array based representation by doing the following.

\begin{itemize}
  \item To eliminate the wasted space in the array representation, we add a child pointer to each tile. This points to the first child of the tile. All children of a tile are stored contiguously.
  \item Leaves are stored in a separate array. We found that, across all our benchmarks, a large fraction of leaves are such that all their siblings are also leaves. Such leaves are directly moved into the leaves array. For leaves for which this property does not hold, an additional hop is added by making the leaf tile a comparison tile and all its children are made leaves with the same value as the original leaf.
\end{itemize}

\begin{figure}
  \centering
  \includegraphics[width=\linewidth]{figures/SparseRep_TileSize3.PNG}
  \caption{Sparse representation with tile size $n_t=3$}
  \label{Fig:SparseRep}
\end{figure}

Figure \ref{Fig:SparseRep} shows some of the details of the sparse representation.
The tree on the left of the diagram is the actual decision tree with the nodes grouped into tiles $t_1$ and $t_2$. The tree on the right is the tree of tiles. The arrays depicted below show how the tree is represented in memory. The first array ($\texttt{tree}$) is an array of tiles and has 5 elements. Each element of the array represents a single tile and has the thresholds of the nodes, the feature indices, a tile shape ID and a pointer to the first child (shown explicitly in red).

As a specific example, consider the tile $t_1$. The tile has four children -- $l_1$, $l_2$, $l_3$ and $t_2$ in that order (left to right). These tiles are stored contiguously in the $\texttt{tree}$ array and a pointer to the first of these, $l_1$ is stored in the tile $t_1$ (the index 1 is stored in the tile $t_1$ as shown).

Now consider the tile $t_2$. Since all children of the tile $t_2$ are leaves, they are all moved into the $\texttt{leaves}$ array. To store a pointer into the $\texttt{leaves}$ array, we add $\texttt{len(tree)}$ to the element index in the $\texttt{leaves}$ array. The tile $t_2$'s child is the element at index 12 of the $\texttt{leaves}$ array. Therefore, the index $12 + 5 = 17$ is stored in the tile $t_2$. (Any index $i$ that is greater than the length of the $\texttt{tree}$ array is regarded as an index into the $\texttt{leaves}$ array. The index into the $\texttt{leaves}$ array is $i - \texttt{len(tree)}$.)

The other aspect of the representation is that an extra hop is added for the leaves $l_1$, $l_2$ and $l_3$ in order to simplify code generation. This enforces the invariant that all leaves are stored in the leaves array and  simplifies checking whether we've reached a leaf. Therefore, 4 new leaves are added as children for each of the original leaves $l_1$, $l_2$ and $l_3$. Each of these 12 newly added leaves has the same value as its parent. These are the first 12 elements of the $\texttt{leaves}$ array.

Even though we currently have implementations of the two representations detailed in sections \ref{sec:ArrayBased} and \ref{sec:SparseRep}, support for other representations is not hard to add. All optimizing passes that work on the high level and mid level IR will continue to work as is. Programmers need only provide new lowering passes for a few operations in the low level IR.

### 2022-04-13 docs/MICRO 2022/abstract.tex
Add abstract

@@
 \begin{abstract}
-Abstract goes here!
+Forests of decision trees, called decision tree ensembles, are a very commonly used
+machine learning model. Machine learning techniques like gradient boosting 
+and random forests build forests of decision trees. These models are used in several
+applications ranging from recommendation systems to tabular data analysis. They 
+are also deployed at scale in the enterprise.
+
+Several systems such as XGBoost, LightGBM and Sklearn implement algorithms based on
+decision tree ensembles. However, limited work exists on optimizing the inference 
+performance of decision tree based models. Existing systems perform specific 
+optimizations but these are hard to maintain as hardware evolves. Additionally, these systems
+cannot tailor the inference code to the model being used. Also, no system exists today
+that can perform a systematic search of the optimization space. To address all these
+problems, we build Treebeard, an extensible compiler infrastructure to compile
+decision tree models. The infrastructure is built to allow exploration of optimization 
+and code generation techniques. We demonstrate the power of the infrastructure by 
+implementing several novel optimizations for decision tree ensemble inference.
+
+We evaluate the performance of code generated by Treebeard and find that our
+optimizations improve performance by 2.2X over an unoptimized baseline. 
+We compare the performance of 
+Treebeard with that of XGBoost and Treelite and find that Treebeard improves
+performance over these systems by almost 3X and 5X respectively. \TODO{How do we 
+list speedups over XGBoost and Treelite? They vary across batch sizes. Should we 
+also state serial and parallel speedups separately?}
+
 \end{abstract}

### 2022-04-13 docs/MICRO 2022/uniftiling.tex
Add draft of uniform tiling

\subsection{Uniform Tiling}
% The justification for uniform tiling -- if we assume all paths are equally likely,
% we want to minimize the depth of speculation
The first tree tiling algorithm we implement is uniform tiling. The basic idea of uniform tiling is to construct tiles
with a tile size $n_t$ so that we evaluate 
the least possible number of nodes speculatively. Intuitively, this translates to minimizing the height of the constructed tile. We minimize
the height of the tile by performing a level order traversal starting at the root node of the tile. 
\TODO{AP Should we say this is optimal when probabilities are equal? Problem is we haven't said what optimal is.}
This traversal collects at most $n_t$ nodes
and does not include any leaves in the tile so that the constraints listed in section \ref{sec:ValidTiling} are respected. Algorithm 
\ref{Alg:UnifTilingAlgo} lists the algorithm for uniform tiling. Tiling starts at the root and constructs a tile $Tile$ by performing
a level order traversal. Once the current tile is constructed, the tiling procedure is recursively performed on all nodes that are 
destinations for edges going out of the constructed tile. The set of tiles constructed by algorithm \ref{Alg:UnifTilingAlgo} constitutes 
a uniform tiling of the input tree.
% Algorithm for uniform tiling
\begin{algorithm}
  \caption{Uniform tree tiling}
  \label{Alg:UnifTilingAlgo}
  \begin{algorithmic}
      \Procedure{TileTree}{$T = (V, E, r)$, $n_t$} 
          \If {$r \in L$}
              \State \textbf{return} $\{ r \}$
          \EndIf
          \State \textcolor{codegreen}{\textit{//Level order traversal to collect $n_t$ or fewer nodes. }}
          \State \textcolor{codegreen}{\textit{//Leaves are not included in the constructed tile. }}
          \State $Tile \leftarrow LevelOrderTraveral(r, n_t)$
          \State $Tiles =  \{ Tile \}$
          \For{$(u,v) \in Out(Tile)$}
              \State $Tiles \leftarrow Tiles \cup TileTree(T_v, n_t)$
          \EndFor
          \State \textbf{return} $Tiles$
      \EndProcedure
  \end{algorithmic}
\end{algorithm}

\subsubsection{Further Opimization and Code Generation}
We found that most leaf tiles for a given tree are at the same depth when uniform tiling is used. Furthermore, we see that deeper leaves 
are more likely to be reached.
%\footnote{Intuitively, this is true because training algorithms keep splitting nodes to maximize gain and gain
%will typically be maximized by splitting a large number of inputs.}.
Based on these observations, we pad the tree of tiles generated with uniform tiling so that all leaves are at the same depth.
This transformation is performed on the high level IR after uniform tiling. 
Once the trees have been padded to make all leaves equal depth, the tree walks are fully unrolled to evaluate a fixed 
number of tiles and all leaf checks are elided.

One other complication the code generator needs to handle is the fact that trees potentially have different depths. In order to handle
this, Treebeard sorts the trees by their depth. This ensures that all trees with equal depth are grouped together. Once this is done, 
the loop over the trees is fissed so that each of the resulting loops only walks trees of a single depth. Consider for example a 
forest with 4 trees $T_1$, $T_2$, $T_3$, and $T_4$ in that order. Further, assume that $T_1$ and $T_4$ have depth 2 while $T_2$ and $T_3$
have depth 3. First, Treebeard reorders the trees to be in the order $T_1$, $T_4$, $T_2$, $T_3$. Then, the loop over the trees is fissed
as shown in the following listing.

% loop transformations for uniform tiling (splitting) 
\begin{lstlisting}{style=c++}
  forest = ensemble(...)
  for i = 0 to batchSize step 1 {
    prediction = 0
    for t = 0 to 2 step 1 {
      tree = getTree(forest, t) 
      node = traverseTreeTile(tree, node, rows[i])
      treePrediction = getLeafValue(tree, node)
      prediction = prediction + treePrediction
    }
    for t = 2 to 4 step 1 {
      tree = getTree(forest, t) 
      node = traverseTreeTile(tree, node, rows[i])
      node = traverseTreeTile(tree, node, rows[i])
      treePrediction = getLeafValue(tree, node)
      prediction = prediction + treePrediction
    }
    predictions[i] = prediction
  }  
\end{lstlisting}

\TODO{AP: This listing is unnecesarily long. Can we maybe leave out the loop bodies and say something like "depth 2 walk"?}

### 2022-04-14 docs/MICRO 2022/abstract.tex
Push Kaushik's edits to the abstract

--- 2022-04-11 docs/MICRO 2022/abstract.tex
@@
 \begin{abstract}
-Abstract goes here!
+  Decision tree ensembles, are a commonly used machine learning model. Such models are
+  generated when using machine learning techniques like gradient boosting and random 
+  forests and are used in a wide range of applications from recommendation systems 
+  to tabular data analysis. They are also deployed at scale in the enterprise.
+
+  Several libraries such as XGBoost, LightGBM and Sklearn expose algorithms for both
+  training and inference with decision tree ensembles. These systems incorporate a 
+  limited set of optimizations usually targeted at specific hardware. Further, existing
+  systems do not specialize the inference code to the model being used.
+
+  This paper presents Treebeard, an extensible compiler for decision tree models.
+  TreeBeard progressively lowers inference code to LLVM through multiple intermediate 
+  representations. By applying model specific optimizations at the higher levels, loop
+  optimizations at the middle level and machine specific optimizations lower down,
+  TreeBeard is able to specialize inference code for each model on each supported
+  hardware target. Treebeard performs several novel optimizations such as tree tiling,
+  tree walk unrolling and tree walk interleaving to improve performance of decision
+  tree ensemble inference.
+
+  We demonstrate the utility of Treebeard by evaluating it on a diverse set of
+  tree ensemble models and comparing it against an unpotimized baseline. We report
+  that Treebeard optimizations improve average latency over a batch of inputs by 2.2X.
+  Further, we find that TreeBeard is significantly faster than other franeworks like
+  XGBoost  (3X on average) and Treelite (5X on average) on both single and multi-core
+  settings. 
 \end{abstract}

### 2022-04-14 docs/MICRO 2022/abstract.tex
Change title and add MLIR to abstract

@@
   This paper presents Treebeard, an extensible compiler for decision tree models.
-  TreeBeard progressively lowers inference code to LLVM through multiple intermediate 
+  Treebeard progressively lowers inference code to LLVM IR through multiple intermediate
   representations. By applying model specific optimizations at the higher levels, loop
   optimizations at the middle level and machine specific optimizations lower down,
-  TreeBeard is able to specialize inference code for each model on each supported
+  Treebeard can specialize inference code for each model on each supported
   hardware target. Treebeard performs several novel optimizations such as tree tiling,
@@
 
-  We demonstrate the utility of Treebeard by evaluating it on a diverse set of
-  tree ensemble models and comparing it against an unpotimized baseline. We report
-  that Treebeard optimizations improve average latency over a batch of inputs by 2.2X.
-  Further, we find that TreeBeard is significantly faster than other franeworks like
-  XGBoost  (3X on average) and Treelite (5X on average) on both single and multi-core
+  We implement Treebeard using the MLIR compiler infrastructure and
+  demonstrate the utility of Treebeard by evaluating it on a diverse set of
+  tree ensemble models. We report that Treebeard optimizations improve
+  average latency over a batch of inputs by 2.2X in comparison to an unoptimized baseline.
+  Further, we find that Treebeard is significantly faster than other frameworks like
+  XGBoost (3X on average) and Treelite (5X on average) in both single and multi-core
   settings.

### 2022-04-14 docs/MICRO 2022/abstract.tex
Grammarly suggestions

@@
 \begin{abstract}
-  Decision tree ensembles, are a commonly used machine learning model. Such models are
-  generated when using machine learning techniques like gradient boosting and random 
-  forests and are used in a wide range of applications from recommendation systems 
-  to tabular data analysis. They are also deployed at scale in the enterprise.
-
-  Several libraries such as XGBoost, LightGBM and Sklearn expose algorithms for both
+  Decision tree ensembles are a commonly used machine learning model
+  generated by machine learning techniques like gradient boosting and random
+  forests. These models are used in applications ranging from recommendation systems
+  to tabular data analysis and are also deployed at scale in the enterprise.
+  Several libraries such as XGBoost, LightGBM, and Sklearn expose algorithms for both
   training and inference with decision tree ensembles. These systems incorporate a
@@
   representations. By applying model specific optimizations at the higher levels, loop
-  optimizations at the middle level and machine specific optimizations lower down,
+  optimizations at the middle level, and machine specific optimizations lower down,
   Treebeard can specialize inference code for each model on each supported
-  hardware target. Treebeard performs several novel optimizations such as tree tiling,
-  tree walk unrolling and tree walk interleaving to improve performance of decision
-  tree ensemble inference.
+  hardware target. To improve model inference performance, Treebeard performs several 
+  novel optimizations such as tree tiling, tree walk unrolling, and tree walk interleaving.
 
@@
   tree ensemble models. We report that Treebeard optimizations improve
-  average latency over a batch of inputs by 2.2X in comparison to an unoptimized baseline.
+  average latency over a batch of inputs by 2.2X compared to an unoptimized baseline.
   Further, we find that Treebeard is significantly faster than other frameworks like

### 2022-04-15 docs/MICRO 2022/abstract.tex
Abstract updates. Overview figure update.

@@
   Several libraries such as XGBoost, LightGBM, and Sklearn expose algorithms for both
-  training and inference with decision tree ensembles. These systems incorporate a
-  limited set of optimizations usually targeted at specific hardware. Further, existing
-  systems do not specialize the inference code to the model being used.
+  training and inference with decision tree ensembles. These libraries incorporate a
+  limited set of optimizations usually targeted at specific hardware. Further, they
+  do not specialize the inference code to the model being used.
 
   This paper presents Treebeard, an extensible compiler for decision tree models.
-  Treebeard progressively lowers inference code to LLVM IR through multiple intermediate
-  representations. By applying model specific optimizations at the higher levels, loop
-  optimizations at the middle level, and machine specific optimizations lower down,
+  Treebeard progressively lowers inference code to LLVM IR through 
+  multiple levels of intermediate abstractions.
+  By applying model-specific optimizations at the higher levels, loop
+  optimizations at the middle level, and machine-specific optimizations lower down,
   Treebeard can specialize inference code for each model on each supported
@@
   demonstrate the utility of Treebeard by evaluating it on a diverse set of
-  tree ensemble models. We report that Treebeard optimizations improve
-  average latency over a batch of inputs by 2.2X compared to an unoptimized baseline.
-  Further, we find that Treebeard is significantly faster than other frameworks like
-  XGBoost  (3X on average) and Treelite (5X on average) in both single and multi-core
-  settings.
+  tree ensemble models. Experimental evaluation demonstrates that Treebeard optimizations 
+  improve average latency over a batch of inputs by 2.2X compared to an unoptimized baseline.
+  Further, Treebeard is significantly faster than other frameworks like
+  XGBoost and Treelite in both single-core (2.8X and 5.1X respectively) and multi-core
+  (3.2X and 2.6X respectively) settings.
 \end{abstract}

### 2022-04-15 docs/MICRO 2022/probtiling.tex
Add prob tiling section

\subsection{Probability Based Tiling}
\subsubsection{Motivation}
As discussed in section \ref{sec:UnifTiling}, uniform tiling optimizes tree walks assuming that all paths in the tree from root to leaf are equally likely. 
However, we find that this is not always the case. We performed inference on the complete training set of each of our benchmarks and measured how many
times each leaf was reached. While for some benchmarks all leaves are roughly equally likely, for others, a small number of leaves are extremely likely 
while the rest aren't reached often. We call trees with a small number of extremely likely leaves \textbf{\emph{skewed}}.
\begin{figure}
    \centering
    \includegraphics[width=\linewidth]{figures/airline-ohe.stats.train.txt.png}
    \caption{Statistical profile for airline-ohe}
    \label{Fig:AirlineOHEStats}
\end{figure}
\begin{figure}
    \centering
    \includegraphics[width=\linewidth]{figures/epsilon.stats.train.txt.png}
    \caption{Statistical profile for epsilon}
    \label{Fig:EpsilonStats}
\end{figure}
Figures \ref{Fig:AirlineOHEStats} and \ref{Fig:EpsilonStats} show this. Each line in these graphs corresponds to a fixed fraction of the input (say $f$). 
A point on a line at coordinate $(x, y)$ means that $y$ percent of trees in the model could handle a fraction $f$ of all training inputs with $x$ percent of 
leaves. \TODO{Explain with a specific point in one of the graphs}
Figure \ref{Fig:AirlineOHEStats} shows that very few leaves are reached for a very large fraction of inputs for the benchmark airline-ohe. 
This means that a small fraction of leaves are very likely. On the other hand,
figure \ref{Fig:EpsilonStats} shows that a large fraction of trees need a large fraction of their leaves to handle any fraction of the test input 
for the benchmark epsilon. One other observation we made is that most models have some skewed trees while the rest of the trees have equally likely leaves.
\TODO{AP Define a term for trees with roughly equally likely leaves.} We design the probability based tiling algorithm to take advantage of this property 
of decision tree ensembles. 
\TODO{This section needs to be rewritten. Define ``leaves covering inputs'' to clarify the presentation.}

\subsubsection{Notation}
In order to formulate the probability based tiling algorithm as an optimization problem, we define the following.
\begin{enumerate}
    \item For every leaf $l \in L$, we define $p_l$ as the probability that the leaf $l$ is reached.
    \item For each node $n \in V$, we define the absolute probability $p_v$ as
    \begin{equation}
        p_v = \begin{cases}
        p_l &\text{if $l \in L$}\\
        p_{left(v)} + p_{right(v)} &\text{otherwise}
        \end{cases}
    \end{equation}
    \item For any tree $T$, $\mathcal{C}(T)$ represents the set of all valid tilings of $T$.
    \item For every $v \in V$, we define $S_v$ as the subtree rooted at $v$.
    \item For every $v \in V$, we define $L_v$ as the set of leaves of $T_v$.
    \item For a every tile $T_i$, we define $root(T_i)$ as the node $v \in T_i$ such that $v$ has no incoming edges from any other node $u \in T_i$.
    \item For a tile $T_i$, $out(T_i) \subseteq E$ is the set of edges $(u, v)$ such that $u \in T_i$ and $v \notin T_i$.
\end{enumerate}

\subsubsection{The Optimization Problem}

%We assume that we are given the probabilities of each leaf node of the decision tree (these can easily be computed using the training data). For every leaf $l \in L$, we %are given the probability $p_l$ that the leaf $l$ is reached.

We observe that the latency of one tree walk is proportional to the number of tiles that need to be evaluated to reach the leaf.
Therefore, the average inference latency can be modelled as the expected number of tiles that are evaluated to compute one tree prediction. More formally, the problem is to find a tiling $\mathcal{T}$ such that the following objective is minimized.
\[
    \min_{\mathcal{T} \in \mathcal{C}(T)}{\sum_{l \in L} p_l.depth_{\mathcal{T}}(l)}
\]
where the minimization is over all valid tilings $\mathcal{T}$ of the tree $T$, $depth_{\mathcal{T}}(l)$ is the depth of the leaf $l$ given tiling ${\mathcal{T}}$.

The constraints on the optimization are the constraints listed in Section \ref{sec:ValidTiling}. 
The above optimization problem can be solved optimally using dynamic programming. We leave this out in the interest of space. 
Instead, we use the simple greedy algorithm listed in algorithm \ref{Alg:GreedyTilingAlgo} to construct a valid tiling given the node probabilities.
The algorithm starts at the root and greedily keeps adding the most probable legal node to the current tile until the maximum tile size is reached.
Subsequently, the tiling procedure is recursively performed on all nodes that are destinations for edges going out of the constructed tile.

% \subsubsection{Dynamic Programming Formulation}

% For any node $v \in V$, we define
% \[
%     cost(v, \mathcal{T}) = \sum_{l \in L_v} p(l | v).depth_{\mathcal{T}}(l)
% \]
% where $\mathcal{T} \in \mathcal{C}(T_v)$.

% Then, the objective function, for the tree $T_v$, can be rewritten as 
% \[
%     opt\_cost(v) = \min_{\mathcal{T} \in \mathcal{C}(T_v)}{cost(v, \mathcal{T})}
% \]

% The objective function can then be rewritten in the following recursive form.
% \[
%     opt\_cost(v) = \min_{T_0 \in TileShapes(n_t, v)}{1 + \sum_{(n_1, n_2) \in out(T_0)} p(n_1 | v)p(n_2 | n1)opt\_cost(n_2)}
% \]
% where $TileShapes(n_t, v)$ is the set of all tile shapes of size $n_t$ with root $v$. A straight forward substitution argument shows why the solution to the subproblems (tiling all sub-trees) needs to to be optimal. The objective is now in a form that can solved using
% dynamic programming.

% \subsubsection{Greedy Algorithm}

% Intuitively, it seems like the following greedy algorithm also gives the optimal tiling. The algorithm starts at the root and greedily keeps adding the most probable node to the current tile until the maximum tile size is reached.
\begin{algorithm}
    \caption{Greedy Probability Based Tree Tiling}
    \label{Alg:GreedyTilingAlgo}
    \begin{algorithmic}
        \Procedure{TileTree}{$T = (V, E, r)$, $n_t$} 
            \If {$r \in L$}
                \State \textbf{return} $\{ r \}$
            \EndIf
            \State $Tile \leftarrow \{ r \}$
            \While{$|Tile| < n_t$}
                \State $e = (u,v) \in Out(Tile)$ st $p(v)$ is max and $v \notin L$
                \If{$e = \emptyset$}
                    \State \textbf{break}
                \EndIf
                \State $Tile = Tile \cup \{ v \}$
            \EndWhile
            \State $Tiles =  \{ Tile \}$
            \For{$(u,v) \in Out(Tile)$}
                \State $Tiles \leftarrow Tiles \cup TileTree(T_v, n_t)$
            \EndFor
            \State \textbf{return} $Tiles$
        \EndProcedure
    \end{algorithmic}
\end{algorithm}

% Talk about problems with increasing number of tile shapes and only performing such tiling on skewed trees
When we tried to apply algorithm \ref{Alg:GreedyTilingAlgo} on all trees in our benchmarks, we found
that even minor variations in probability caused the tiling algorithm to generate a large 
number of tile shapes. This in turn caused a loss in performance because the large size of the 
lookup table needed (section \ref{sec:LookupTable}) caused increased L1 cache misses. In order to 
alleviate this, we only perform probability based tiling on trees that are skewed. We say a tree is
skewed if a small fraction of leaves, say $\alpha$, cover a large fraction of training inputs, say $\beta$.
We only perform probability based tiling on skewed trees and fall back to uniform tiling otherwise. 
In our experiments we use $\alpha=0.05$ and $\beta=0.9$.

\subsubsection{Code Generation}
As probability based tiling pulls the most probable leaves of a decision tree nearest the root, it poses 
some implementation challenges. By design, the tiling process makes the tree of tiles 
imbalanced. Therefore, the simple array based representation described in section \ref{sec:ArrayBased}
cannot be used because of the memory footprint blow up (a large part of the tree is empty, but would 
need to be allocated). On the other hand, the sparse representation in section \ref{sec:SparseRep} adds 
an extra hop for leaves that have non-leaf siblings. But this would mean that we add extra hops for 
the most probable leaves after probability based tiling. To address these problems, Treebeard peels 
the tree walk and specializes the leaf checks at higher levels to avoid the extra hop. Currently, 
we determine the maximum depth of leaves needed to cover 90 percent of the inputs and peel the tree 
walk by as many iterations. For example, consider the case where leaves until depth 2 are needed to 
cover 90 percent of the training input. Then, Treebeard generates the following IR.

\begin{lstlisting}{style=c++}
    // ...
    tree = getTree(forest, t)
    node = getRoot(tree)
    node = traverseTreeTile(tree, node, rows[i])
    if (isLeafTile(node)) {
        treePrediction = getLeafValue(tree, node)
    } else {    
        node = traverseTreeTile(tree, node, rows[i])
        if (isLeafTile(node)) {
            treePrediction = getLeafValue(tree, node)
        } else {    
            // Loop based traversal 
        }
    }
    treePrediction = getLeafValue(tree, node)
    // ...
\end{lstlisting}

The if statements check whether a node is a leaf tile and hence avoid the extra hop. The memory 
requirement is also not increased because only a small fraction of leaves are represented as full tiles.
While walk peeling is used to improve the performance of probability based tiling by specializing leaf tests,
the transformation is by itself general and can be used in different contexts. For example, it could be used 
to elide leaf checks until a depth $d$ is reached if we know all leaves are at a depth greater than $d$.

One other issue that the code generator needs to handle is that walks of different trees in the same ensemble may 
need to be peeled to different depths. In order to handle this, trees are reordered so that all trees 
with equal peeling depth are grouped together and the loops in the IR are fissed so that tree walks 
for these groups of trees can be specialized differently. This is very similar to the code generation strategy
used for uniform tiling.

### 2022-04-15 docs/MICRO 2022/tiling.tex
Add draft of uniform tiling
Add prob tiling section

@@
 \subsection{Valid Tiling of a Tree}
+\label{sec:ValidTiling}
 A tiling $\mathcal{T}$ of the tree $T = (V, E, r)$ with tile size $n_t$ is a partition $\{ T_1, T_2, ... ,T_m \}$ of the set $V$ such that 
@@
 \end{enumerate}
-
-\TODO{AP We need to talk about how tiling is specified in the compiler}
 
@@
 \subsection{Lookup Table}
+\label{sec:LookupTable}
 A lookup table (LUT) is used to solve the problem described in section \ref{Sec:TileShapesAndDecisionTreeInference}, i.e.

### 2022-04-15 docs/MICRO 2022/details.tex
Add draft of uniform tiling
Add prob tiling section
Add interleaving, vectorization, parallelization

@@
 \input{tiling}
+\input{uniftiling}
+\input{probtiling}
+\input{unrollandjam}
+
+\subsection{Vectorization}
+Vectorization performed by Treebeard is enabled by the tiling transformations described in section \ref{sec:Tiling}. 
+When the low level IR is translated to LLVM IR, Treebeard generates LLVM instructions that operate on the threholds and feature indices 
+of nodes within a tile in a vector fashion. Therefore, thresholds and feature indices are loaded using vector loads and predicates are 
+evaluated using vector comparisons. These vector LLVM IR instructions are then converted to vector instructions in the target ISA by 
+the LLVM JIT.
+
+\subsection{Parallelization}
+Currently, Treebeard performs a naive parallelization of the inference computation. When parallelism is enabled, the 
+loop over the input rows is parallelized using OpenMP. Treebeard rewrites 
+the mid-level IR by tiling the loop over the input rows with a tile size equal to the number of cores. 
+As a concrete example, consider the case where we intend to perform inference using a model with 4 trees 
+on a batch of 64 rows. Further, assume that we wish to parallelize this computation across 8 cores. 
+Treebeard then generates the following IR.
+\begin{lstlisting}[style=c++]
+  forest = ensemble(...)
+  parallel.for i0 = 0 to 64 step 8 {
+    for i1 = 0 to 8 step 1 {
+      i = i0 + i1
+      prediction = 0
+      for t = 0 to 4 step 1 {
+        tree = getTree(forest, t) 
+        treePrediction = WalkTree(tree, rows[i])
+        prediction = prediction + treePrediction
+      }
+      predictions[i] = prediction
+    }
+  }
+\end{lstlisting}
+Currently, as our focus is on single core performance, we do not 
+do any parallelism related optimizations (such as loop tiling). We leave a more thorough exploration of parallelization 
+to future work.

### 2022-04-15 docs/MICRO 2022/main.tex
Add draft of uniform tiling
Change title and add MLIR to abstract
Add interleaving, vectorization, parallelization

--- 2022-04-11 docs/MICRO 2022/main.tex
@@
 \usepackage[bookmarks=true,breaklinks=true,letterpaper=true,colorlinks,citecolor=blue,linkcolor=blue,urlcolor=blue]{hyperref}
+\usepackage{algorithm}
+%\usepackage{algorithmic}
+\usepackage{algpseudocode}
 
@@
 
-\lstdefinestyle{mystyle}{
+\lstdefinestyle{c++}{language=C++,
+    morekeywords={ensemble, getTree, getRoot, isLeaf, isLeafTile, traverseTreeTile, getLeafValue, loadThresholds, loadFeatureIndices, loadTileShape, getChildNode},    
     backgroundcolor=\color{backcolour},   
@@
 %%%%%%%%%%%---SETME-----%%%%%%%%%%%%%
-\title{Treebeard : A Compiler for Decision Forest Inference} 
+\title{Treebeard : An Optimizing Compiler for Tree Based ML Inference} 
 %%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%

### 2022-04-15 docs/MICRO 2022/overview.tex
overview HIR

@@
 Figure \ref{Fig:CompilerStructure} shows the high level structure of Treebeard. 
-The input to TreeBeard is a JSON serialized decision forest model, it supports popular frameworks like XGBoost and LightGBM and is extensible to other frameworks.
+The input to Treebeard is a JSON serialized decision forest model, it supports popular frameworks like XGBoost and LightGBM and is extensible to other frameworks.
 Given an input model our compiler generates optimized inference code.
@@
 
-TreeBeard specializes the code generated for inference by progressively optimizing and lowering a high level representation of the \texttt{predictForest} function down to LLVM IR \cite{LLVM}.
-It is implemented as a dialect in the MLIR framework \cite{MLIR},
+Treebeard specializes the code generated for inference by progressively optimizing and lowering a high level representation of the \texttt{predictForest} function down to LLVM IR \cite{LLVM}.
 %Treebeard first constructs a high level representation of the decision forest inference operation. 
- At this level (HIR) it applies transformations that determine how inputs traverse the decision forest.
-It re-order trees in the forest and determines the high level loop structure (process one tree at a time or one input at a time).
-It then lowers to a traditional loop based IR (MIR).
-Several loop level optimizations like unrolling and pipelining are applied at this level.
-At the lowest level of IR TreeBeard applies layout optimizations and generates parallel code in LLVM IR.
+At the highest level (HIR) it tiles nodes together to transform a binary tree into an $n$-ary tree and applies transformations that determine how inputs traverse the decision forest.
+It then lowers to a mid-level IR (MIR) where loop optimizations like loop-reordering and unrolling are applied.
+The compiler also figures out which loop nest to parallelize.
+At the lowest level of IR Treebeard applies layout optimizations and generates vectorized code to take advantage of SIMD instructions where applicable.
 LLVM is then used to JIT compile the generated IR to executable code. The following sections describe each level of the IR and their lowering in more detail.
@@
 
-\TODO{Should we describe the dialect's type system?}
+%\TODO{Should we describe the dialect's type system?}
 \TODO{Kr : consider focusing on the example instead of verbose description of optimizations. This section can be short, descriptions can come in later sections.}
 \subsection{High Level IR}
-Treebeard parses the input JSON file and generates a function with a single MLIR operation, \texttt{predictForest} that represents inference using the input model on a set of rows.
-The operation contains within it a tree based representation of the model that can be manipulated by optimizing transformations.
+As a first step Treebeard parses the input and generates a single MLIR operation, \texttt{predictForest} that represents inference using the input model on a set of rows. 
+At this level the operator simply contains a collection of binary trees.
+Two optimizations, tiling and tree ordering are applied at this level.
+The objective of tiling is to group nodes together so that the tree can be walked one tile at a time instead of one node at a time.
+We demonstrate that tiling can specialize the traversal of individual trees to either balance heavily skewed trees or proritize walks leading to higher probability leafs.
+Figure~\ref{ttile} shows examples of the tiling transformation.
+\TODO{kr: draw and explain example}.
+The objective of tree padding and ordering is to group identically structured trees so that they can share the same traversal code.
+The predictForest function is now lowered to MIR with separate loop nests generated for traversing a group of identical trees.
+The generated MIR code is as shown below.
+\TODO{show unoptimized loop nest as a listing}.
+
+%The operation contains within it a tree based representation of the model that can be manipulated by optimizing transformations.
 Transformations on the model such as tiling, tree reordering and leaf padding are done at this level.
@@
 \subsection{Mid Level IR}
-The Mid Level IR makes the loop structures and tree walks more explicit.
-Firstly, the order in which the iteration space of trees and inputs is walked is explicitly specified in the IR through loop nests.
+The Mid Level IR optimizes the loop structures and tree walks.
+Firstly, the order in which the iteration space of trees and inputs is walked is determined and by re-ordering the loop nesting if necessary.
 Also, operations such as \texttt{isLeaf}, \texttt{traverseTile}, \texttt{getLeafValue} are introduced so that the traversal of trees explicitly represented.

### 2022-04-15 docs/MICRO 2022/overview.tex
overview

@@
+\newcommand{\Treebeard}{{\sc {Treebeard}}}
+\newcommand{\op}[1]{{\texttt {#1}}}
+
 \section{Compiler Overview}
-Figure \ref{Fig:CompilerStructure} shows the high level structure of Treebeard. 
-The input to Treebeard is a JSON serialized decision forest model, it supports popular frameworks like XGBoost and LightGBM and is extensible to other frameworks.
+Figure \ref{Fig:CompilerStructure} shows the high level structure of \Treebeard. 
+The input to \Treebeard{} is a JSON serialized decision forest model, it supports popular frameworks like XGBoost and LightGBM and is extensible to other frameworks.
 Given an input model our compiler generates optimized inference code.
-Specifically it generates a callable batch inference function \texttt{predictForest} that given an array of inputs, outputs an array of model predictions.
+Specifically it generates a callable batch inference function \op{predictForest} that given an array of inputs, outputs an array of model predictions.
 
-Treebeard specializes the code generated for inference by progressively optimizing and lowering a high level representation of the \texttt{predictForest} function down to LLVM IR \cite{LLVM}.
+\Treebeard{} specializes the code generated for inference by progressively optimizing and lowering a high level representation of the \op{predictForest} function down to LLVM IR \cite{LLVM}.
 %Treebeard first constructs a high level representation of the decision forest inference operation. 
-At the highest level (HIR) it tiles nodes together to transform a binary tree into an $n$-ary tree and applies transformations that determine how inputs traverse the decision forest.
-It then lowers to a mid-level IR (MIR) where loop optimizations like loop-reordering and unrolling are applied.
-The compiler also figures out which loop nest to parallelize.
-At the lowest level of IR Treebeard applies layout optimizations and generates vectorized code to take advantage of SIMD instructions where applicable.
+At the highest level (HIR) optimizations like tree tiling and ordering are applied to compact and balance the trees and group structurally identical trees in the forest so that they can share the same code.
+It then lowers to a mid-level IR (MIR) where loop optimizations like loop unrolling and parallelization are applied.
+At the lowest level of IR \Treebeard{} applies layout optimizations and generates vectorized code to take advantage of SIMD instructions where applicable.
 LLVM is then used to JIT compile the generated IR to executable code. The following sections describe each level of the IR and their lowering in more detail.
@@
 \subsection{High Level IR}
-As a first step Treebeard parses the input and generates a single MLIR operation, \texttt{predictForest} that represents inference using the input model on a set of rows. 
+As a first step \Treebeard{} parses the input and generates a single MLIR operation, \op{predictForest} that represents inference using the input model on a set of rows. 
 At this level the operator simply contains a collection of binary trees.
@@
 \TODO{kr: draw and explain example}.
+
 The objective of tree padding and ordering is to group identically structured trees so that they can share the same traversal code.
-The predictForest function is now lowered to MIR with separate loop nests generated for traversing a group of identical trees.
+The \op{predictForest} function is now lowered to MIR with separate loop nests generated for traversing a group of identical trees.
 The generated MIR code is as shown below.
@@
 The Mid Level IR optimizes the loop structures and tree walks.
-Firstly, the order in which the iteration space of trees and inputs is walked is determined and by re-ordering the loop nesting if necessary.
-Also, operations such as \texttt{isLeaf}, \texttt{traverseTile}, \texttt{getLeafValue} are introduced so that the traversal of trees explicitly represented.
-The following listing shows the IR for inference using a model with four trees on an input batch with two rows.
-The listed IR walks all trees for one input row before moving to the next row.
-One important point to note here is that details such as the data structure used for the trees are not explicitly encoded in the IR.
-This allows us to reuse optimization and lowering passes on this level of the IR regardless of what the final in memory representation of the model is.
+The order in which the iteration space of trees and inputs is walked is determined and loop nest are re-ordered if necessary.
+Loop peeling makes use of tree specific information about the levels at which leafs start to occur, to split the traversal loop into two parts, one that traverses the tree a constant number of steps (say upto first leaf) and an eplilogue that traverses the rest.
+Unroll and Jam, unrolls the inner loop and interleaves traversal steps so that a single tree is simultaneously traversed by multiple inputs.
+Usually when applied after peeling, it unrolls the first split almost entirely.
+\TODO{add figure with example, and text where appropriate above}.
+Finally, the outermost loop is parallelized by leveraging existing MLIR passes.
+The \op{predictForest} function is now lowered to LIR with placeholder operators \op{isLeaf}, \op{traverseTile}, \op{getLeafValue} as shown below.
+
+\subsection{Low Level IR}
+Optimizations at the LIR specialize the implementation of placeholder operators to the specific hardware. 
+ For example, \op{traverseTile} is lowered into a series of operations to load thresholds, feature indices and features, compare the features with the thresholds and compute the next node to evaluate.
+Low level ISA instructions are picked (instruction selection) to implement these functions efficiently, for example vector instructions are introduced to exploit SIMD capabilities of the underlying hardware.
+This stage also determines the memory layout of the trees, the goal of the layout optimizations is to enable the traversal of the tree and implement checks like \op{isLeaf} at low memory overhead.
+\TODO{example}This IR is then lowered directly to LLVM IR and JITted.\\
+\\
+  \noindent
+Note that all the optimization described so far are orthogonal and can be applied independent of each other.
+However, as we will see in detail over the next few sections, they do compliment each other.
+Padding and tree tiling discussed before help balance the tree, tiling in addition reduces the depth of the tree.
+These help peeling and unrolling eliminate most of the branches and loop checks in the innermost loop, producing mostly simple straightline code.
+This code is usually amenable to low level optimizations like vectorization.
 
@@
 The IR listed above is a simplification of the actual IR. The actual IR is strongly typed and in SSA form.
-
-\subsection{Low Level IR}
-The IR is finally lowered to a form where the in memory representation of the model is made explicit.
-Buffers to hold the model values are inserted into the generated code and all tree operations in the mid-level IR are lowered to explicitly reference these buffers.
-The semantics of all operations are made explicit.
-For example, \texttt{traverseTreeTile} is lowered into a series of operations to load thresholds, feature indices and features, compare the features with the thresholds and compute the next node to evaluate.
-This IR is then lowered directly to LLVM IR and JITted.

### 2022-04-15 docs/MICRO 2022/tiling.tex
Add interleaving, vectorization, parallelization
Add new tiling diagrams

@@
 \subsection{Tiling}
-
+\label{sec:Tiling}
 Treebeard vectorizes tree walks by grouping nodes of a decision tree into \textbf{\emph{tiles}}.
@@
   \includegraphics[width=\linewidth]{figures/TileShapes_Size3.PNG}
-  \caption{All possible tile shapes with a tile size $n_t=3$}
+  \caption{All possible tile shapes with tile size $n_t=3$}
   \label{Fig:TileSize3Shapes}
@@
   \includegraphics[width=\linewidth]{figures/TiledTree_Size3.PNG}
-  \caption{Example of a valid tree tiling with a tile size $n_t=3$}
+  \caption{Example of a valid tree tiling with tile size $n_t=3$}
   \label{Fig:ValidTilingTileSize3}

### 2022-04-16 docs/MICRO 2022/overview.tex
Rewrite overview

--- 2022-04-15 docs/MICRO 2022/overview.tex
@@
 Figure \ref{Fig:CompilerStructure} shows the high level structure of Treebeard. 
-The input to Treebeard is a JSON serialized decision forest model, it supports popular frameworks like XGBoost and LightGBM and is extensible to other frameworks.
-Given an input model our compiler generates optimized inference code.
-Specifically it generates a callable batch inference function \texttt{predictForest} that given an array of inputs, outputs an array of model predictions.
+The input to Treebeard is a serialized decision tree ensemble. Popular frameworks 
+like XGBoost and LightGBM are supported and the system is easily extensible to other frameworks.
+Given an input model our compiler generates optimized inference code. Specifically it generates a 
+callable batch inference function \texttt{predictForest} that given an array of input rows, outputs an
+array of model predictions.
 
-Treebeard specializes the code generated for inference by progressively optimizing and lowering a high level representation of the \texttt{predictForest} function down to LLVM IR \cite{LLVM}.
+Treebeard specializes the code generated for inference by progressively optimizing and lowering a 
+high level representation of the \texttt{predictForest} function down to LLVM IR \cite{LLVM}.
+By \textbf{\emph{lowering}}, we mean the process of transforming an operation at a higher 
+abstraction to a sequence of operations at a lower abstraction. Optimizations in Treebeard 
+are implemented using a combination of annotation and lowering. An Operation at a higher 
+abstraction is annotated with meta-data that indicates what kind of optimization is 
+to be performed while lowering it. The lowering transformation uses this information to generate 
+optimized lower level code. For example, tree tiling and loop order are decided 
+at the highest abstraction. These decisions are communicated to the lowering pass 
+which explicitly encodes them in the lowered IR as shown in figure \ref{Fig:Overview}.
+
 %Treebeard first constructs a high level representation of the decision forest inference operation.
-At the highest level (HIR) it tiles nodes together to transform a binary tree into an $n$-ary tree and applies transformations that determine how inputs traverse the decision forest.
-It then lowers to a mid-level IR (MIR) where loop optimizations like loop-reordering and unrolling are applied.
-The compiler also figures out which loop nest to parallelize.
-At the lowest level of IR Treebeard applies layout optimizations and generates vectorized code to take advantage of SIMD instructions where applicable.
-LLVM is then used to JIT compile the generated IR to executable code. The following sections describe each level of the IR and their lowering in more detail.
+Treebeard's IR has three abstractions as shown in Figure \ref{fig:Overview}.  
+At the highest level (HIR), the input model is represented as a collection of binary trees. This is shown 
+in the second row of figure \ref{fig:Overview}. At this level of abstraction,
+Treebeard tiles nodes together to transform a binary tree into an $n$-ary tree 
+and decides what order trees are to be traversed in. In the example in the figure, 
+trees are tiled with a tile size of 2. Tiling is indicated using colored elipses drawn 
+around nodes that are in the same tile. Tree reordering is another transformation that is performed at this abstraction. 
+One objective of tree reordering is to group identically structured trees so that they can share the same traversal code. 
+With the tiling shown in Figure \ref{fig:Overview}, Tree1 and Tree3 have depth 2, while Tree2 has a depth of 3.
+Therefore, the trees are reordered so that Tree1 and Tree3 are together.
 
- %The specified set of options includes information such as the batch size, the type of the input features, the type for node thresholds and the type for feature indices \TODO{There are also several optimization related inputs like tile size, type of tiling, pipeline depth etc.
-Should we mention those?}.
+The high level IR is then lowered to a mid-level IR (MIR). The aim of MIR is 
+to allow optimizations that are independent of the final memory layout of the model. 
+The lowering to MIR performs loop optimizations like loop-reordering, parallelization and unrolling.
+At this level of the IR (the third row in Figure \ref{fig:Overview}),
+the order in which trees, input row pairs are walked is explicitly represented in the IR using loop nests.
+Figure \ref{fig:Overview} shows two possible ways that loop nests could be generated -- the first
+walks all rows for one tree before moving to the next tree while the second walks all trees for 
+one row before moving to the next row. Another aspect handled by this lowering is the fissing of 
+loops to make sure that trees with the same structure share code. Here, both versions of MIR shown 
+ensure that Tree1 and Tree3 share the same traversal code, while traversal code for Tree2 is different.
+Additionally, tree walk optimizations, such as tree walk unrolling (shown in figure \ref{fig:Overview}) and 
+peeling are performed on the MIR.
+
+Treebeard then further lowers the IR to explicitly represent the memory layout of the model. Buffers to hold model 
+values are inserted into the generated code and all tree operations in the mid-level IR are lowered to explicitly 
+reference these buffers. Additionally, at this level of the IR, Treebeard generates vectorized code to take 
+advantage of SIMD instructions where applicable. Finally, the inference function is translated to LLVM IR and 
+JIT compiled to executable code.
+
+%The following sections describe each level of the IR and their lowering in more detail.
 
@@
 
+% \begin{figure}
+%   \centering
+%   \includegraphics[width=\textwidth]{figures/CompilerOverview.png}
+%   \caption{Treebeard IR lowering and optimization details}
+%   \label{Fig:Overview}
+% \end{figure}
+
 %\TODO{Should we describe the dialect's type system?}
-\TODO{Kr : consider focusing on the example instead of verbose description of optimizations. This section can be short, descriptions can come in later sections.}
-\subsection{High Level IR}
-As a first step Treebeard parses the input and generates a single MLIR operation, \texttt{predictForest} that represents inference using the input model on a set of rows. 
-At this level the operator simply contains a collection of binary trees.
+% \TODO{Kr : consider focusing on the example instead of verbose description of optimizations. This section can be short, descriptions can come in later sections.}
+% \subsection{High Level IR}
+% As a first step Treebeard parses the input and generates a single MLIR operation, \texttt{predictForest} that represents inference using the input model on a set of rows. 
+% At this level the operator simply contains a collection of binary trees.
 Two optimizations, tiling and tree ordering are applied at this level.
@@
 \TODO{kr: draw and explain example}.
-The objective of tree padding and ordering is to group identically structured trees so that they can share the same traversal code.
-The predictForest function is now lowered to MIR with separate loop nests generated for traversing a group of identical trees.
+The objective of tree reordering is to group identically structured trees so that they can share the same traversal code.
+The predictForest function is now lowered to MIR with separate loop nests generated for traversing groups of identical trees.
 The generated MIR code is as shown below.
-\TODO{show unoptimized loop nest as a listing}.
+% \TODO{show unoptimized loop nest as a listing}.
 
@@
 \TODO{Should we mention that there is a scheduling language to decide this?}
+% \begin{lstlisting}[language=C++]
+% func Predict(float rows[batchSize]) {
+%   predictions = predictForest(rows) 
+%   return rows
+% }
+% \end{lstlisting}
 
-\begin{lstlisting}[language=C++]
-func Predict(float rows[batchSize]) {
-  predictions = predictForest(rows) 
-  return rows
-}
-\end{lstlisting}
-
-\subsection{Mid Level IR}
-The Mid Level IR optimizes the loop structures and tree walks.
+% \subsection{Mid Level IR}
+% The Mid Level IR optimizes the loop structures and tree walks.
 Firstly, the order in which the iteration space of trees and inputs is walked is determined and by re-ordering the loop nesting if necessary.
@@
 
-\begin{lstlisting}[style=c++]
-  // Constant that represents the model being compiled
-  forest = ensemble(...)
-  for i = 0 to 2 step 1 {
-    prediction = 0
-    for t = 0 to 4 step 1 {
-      tree = getTree(forest, t) 
-      node = getRoot(tree)
-      while (isLeaf(tree, n)==false)  do {
-        node = traverseTreeTile(tree, node, rows[i])
-      }
-      treePrediction = getLeafValue(tree, node)
-      prediction = prediction + treePrediction
-    }
-    predictions[i] = prediction
-  }
-\end{lstlisting}
+% \begin{lstlisting}[style=c++]
+%   // Constant that represents the model being compiled
+%   forest = ensemble(...)
+%   for i = 0 to 2 step 1 {
+%     prediction = 0
+%     for t = 0 to 4 step 1 {
+%       tree = getTree(forest, t) 
+%       node = getRoot(tree)
+%       while (isLeaf(tree, n)==false)  do {
+%         node = traverseTreeTile(tree, node, rows[i])
+%       }
+%       treePrediction = getLeafValue(tree, node)
+%       prediction = prediction + treePrediction
+%     }
+%     predictions[i] = prediction
+%   }
+% \end{lstlisting}
 
-The IR listed above is a simplification of the actual IR. The actual IR is strongly typed and in SSA form.
+% The IR listed above is a simplification of the actual IR. The actual IR is strongly typed and in SSA form.
 
-\subsection{Low Level IR}
-The IR is finally lowered to a form where the in memory representation of the model is made explicit.
+% \subsection{Low Level IR}
+% The IR is finally lowered to a form where the in memory representation of the model is made explicit.
 Buffers to hold the model values are inserted into the generated code and all tree operations in the mid-level IR are lowered to explicitly reference these buffers.

### 2022-04-16 docs/MICRO 2022/abstract.tex

@@
 \begin{abstract}
-  Decision tree ensembles are a commonly used machine learning model
-  generated by machine learning techniques like gradient boosting and random
-  forests. These models are used in applications ranging from recommendation systems
-  to tabular data analysis and are also deployed at scale in the enterprise.
-  Several libraries such as XGBoost, LightGBM, and Sklearn expose algorithms for both
-  training and inference with decision tree ensembles. These libraries incorporate a
-  limited set of optimizations usually targeted at specific hardware. Further, they
-  do not specialize the inference code to the model being used.
+  Decision tree ensembles are commonly used machine learning models
+  generated by machine learning techniques like gradient boosting and random forests. 
+  These models are used in a wide range of applications and are deployed at scale
+  (run repeatedly on billions of data items). Several libraries such as XGBoost, 
+  LightGBM, and Sklearn expose algorithms for both training and inference with 
+  decision tree ensembles. While these libraries 
+  incorporate a limited set of hardware-specific optimizations,
+  they do not specialize the inference code to the model being used, 
+  leaving significant performance on the table.
 
-  This paper presents Treebeard, an extensible compiler for decision tree models.
-  Treebeard progressively lowers inference code to LLVM IR through 
-  multiple levels of intermediate abstractions.
+  This paper presents a compiler-based approach that automatically generates  
+  efficient code for decision tree inference. 
+  It develops Treebeard, an extensible compiler, which progressively lowers  
+  inference code to LLVM IR through multiple intermediate abstractions.
   By applying model-specific optimizations at the higher levels, loop
@@
   tree ensemble models. Experimental evaluation demonstrates that Treebeard optimizations 
-  improve average latency over a batch of inputs by 2.2X compared to an unoptimized baseline.
-  Further, Treebeard is significantly faster than other frameworks like
+  improve average latency over a batch of inputs by 2.2X over our baseline version. 
+  Further, Treebeard is significantly faster than %% other frameworks like
   XGBoost and Treelite in both single-core (2.8X and 5.1X respectively) and multi-core

### 2022-04-16 docs/MICRO 2022/main.tex

@@
 \input{details}
-
-\CommentOut{
-\begin{abstract}
-
-  This document is intended to serve as a sample for submissions to the 55\textsuperscript{th} IEEE/ACM International Symposium on Microarchitecture\textsuperscript{\textregistered} (MICRO 2022).
-We provide some guidelines that authors should follow when submitting papers to the conference.
-This format is derived from the ACM sig-alternate.cls file, and is used with an objective of keeping the submission version similar to the camera-ready version.
-
-\end{abstract}
-
-\section{Introduction}
-
-This document provides instructions for submitting papers to the 55\textsuperscript{th} IEEE/ACM International Symposium on Microarchitecture\textsuperscript{\textregistered} (MICRO 2022).
-In an effort to respect the efforts of reviewers and in the interest of fairness to all prospective authors, we request that all submissions to MICRO 2022 follow the formatting and submission rules detailed below.
-Submissions that violate these instructions may not be reviewed, at the discretion of the program chairs, in order to maintain a review process that is fair to all potential authors.
-
-This document is itself formatted using the MICRO 2022 submission format.
-The content of this document mirrors that of the submission instructions that appear on the conference website.
-All questions regarding paper formatting and submission should be directed to the program chairs.
-
-\subsection{Format Highlights}
-\begin{itemize}
-\item Paper must be submitted in printable PDF format.
-\item Text must be in a minimum 10pt font, see Table~\ref{table:formatting}.
-\item Papers must be at most 11 pages, not including references.
-\item No page limit for references.
-\item Each reference must specify {\em all} authors (no {\em et al.}).
-\item Author anonymity must be fully preserved, including in any referenced artifacts (e.g., GitHub repository).
-\end{itemize}
-
-\subsection{Paper Evaluation Objectives} 
-The committee will make every effort to judge each submitted paper on its own merits.
-There will be no target acceptance rate.
-We expect to accept a wide range of papers with appropriate expectations for evaluation---while papers that build on significant past work with strong evaluations are valuable, papers that open new areas with less rigorous evaluation are equally welcome and especially encouraged.
-
-\section{Paper Preparation Instructions}
-
-\subsection{Paper Formatting}
-
-Papers must be submitted in printable PDF format and should contain a {\em maximum of 11 pages} of single-spaced two-column text, {\bf not including references}.
-You may include any number of pages for references, but see below for more instructions.
-If you are using \LaTeX~\cite{lamport94} to typeset your paper, then we suggest that you use the template here: \href{https://www.microarch.org/micro55/submit/micro55-latex-template.zip}{\LaTeX~Template}.
-This document was prepared with that template.
-Note that the template and sample paper may render slightly differently on different \LaTeX~engines, due to typesetting changes between versions.
-If you use a different software package to typeset your paper, then please adhere to the guidelines given in Table~\ref{table:formatting}.
-
-\begin{scriptsize}
-\begin{table}[h!]
-  \centering
-  \begin{tabular}{|l|l|}
-    \hline
-    \textbf{Field} & \textbf{Value}\\
-    \hline
-    \hline
-    File format & PDF \\
-    \hline
-    Page limit & 11 pages, {\bf not including}\\
-               & {\bf references}\\
-    \hline
-    Paper size & US Letter 8.5in $\times$ 11in\\
-    \hline
-    Top margin & 1in\\
-    \hline
-    Bottom margin & 1in\\
-    \hline
-    Left margin & 0.75in\\
-    \hline
-    Right margin & 0.75in\\
-    \hline
-    Body & 2-column, single-spaced\\
-    \hline
-    Space between columns & 0.25in\\
-    \hline
-    Line spacing (leading) & 11pt \\
-    \hline
-    Body font & 10pt, Times\\
-    \hline
-    Abstract font & 10pt, Times\\
-    \hline
-    Section heading font & 12pt, bold\\
-    \hline
-    Subsection heading font & 10pt, bold\\
-    \hline
-    Caption font & 9pt (minimum), bold\\
-    \hline
-    References & 8pt, no page limit, list \\
-               & all authors' names\\
-    \hline
-  \end{tabular}
-  \caption{Formatting guidelines for submission.}
-  \label{table:formatting}
-\end{table}
-\end{scriptsize}
-
-{\em Please ensure that you include page numbers with your submission}.
-This makes it easier for the reviewers to refer to different parts of your paper when they provide comments.
-Please ensure that your submission has a banner at the top of the title page, similar to this document, which contains the submission number and the notice of confidentiality.
-If using the template, just replace XXX with your submission number.
-
-\subsection{Content}
-
-Reviewing will be {\em double blind} (no author list); therefore, please do not include any author names on any submitted documents except in the space provided on the submission form.
-You must also ensure that the metadata included in the PDF does not give away the authors.
-You must fully anonymize any links to artifacts (e.g., GitHub repository) and remove any links to artifacts that cannot be fully anonymized.
-{\bf Papers that violate the anonymization policy may be rejected without review} at the discretion of the program chairs.
-
-If you are improving upon your prior work, refer to your prior work in the third person and include a full citation for the work in the bibliography.
-For example, if you are building on {\em your own} prior work in the papers \cite{nicepaper1,nicepaper2,nicepaper3}, you would say something like: "While the authors of \cite{nicepaper1,nicepaper2,nicepaper3} did X, Y, and Z, this paper additionally does W, and is therefore much better."  Do NOT omit or anonymize references for blind review.
-There is one exception to this for your own prior work that appeared in IEEE CAL, arXiv, workshops without archived proceedings, etc.\, as discussed later in this document.
-
-\noindent\textbf{Figures and Tables:} Ensure that the figures and tables are legible.
-Please also ensure that you refer to your figures in the main text.
-Many reviewers print the papers in gray-scale.
-Therefore, if you use colors for your figures, ensure that the different colors are highly distinguishable in gray-scale.
-
-\noindent\textbf{References:}  There is no length limit for references.
-{\em Each reference must explicitly list all authors of the paper.
-Papers not meeting this requirement will be rejected.} Authors of NSF proposals should be familiar with this requirement.
-Knowing all authors of related work will help find the best reviewers.
-Since there is no length limit for the number of pages used for references, there is no need to save space here.
-
-\section{Paper Submission Instructions}
-
-\subsection{Guidelines for Determining Authorship}
-IEEE guidelines dictate that authorship should be based on a {\em substantial intellectual contribution}.
-It is assumed that all authors have had a significant role in the creation of an article that bears their names.
-In particular, the authorship credit must be reserved only for individuals who have met each of the following conditions:
-
-\begin{enumerate}
-\item Made a significant intellectual contribution to the theoretical development, system or experimental design, prototype development, and/or the analysis and interpretation of data associated with the work contained in the article;
-
-\item Contributed to drafting the article or reviewing and/or revising it for intellectual content; and
-
-\item Approved the final version of the article as accepted for publication, including references.
-\end{enumerate}
-
-A detailed description of the IEEE authorship guidelines and responsibilities is available \href{https://www.ieee.org/publications_standards/publications/rights/Section821.html}{here}.
-Per these guidelines, it is not acceptable to award {\em honorary } authorship or {\em gift} authorship.
-Please keep these guidelines in mind while determining the author list of your paper.
-
-\subsection{Declaring Authors}
-Declare all the authors of the paper upfront.
-Addition/removal of authors once the paper is accepted will have to be approved by the program chairs, since it potentially undermines the goal of eliminating conflicts for reviewer assignment.
-
-\subsection{Areas and Topics}
-Authors should indicate these areas on the submission form as well as specific topics covered by the paper for optimal reviewer match.
-If you are unsure whether your paper falls within the scope of MICRO, please check with the program chairs -- MICRO is a broad, multidisciplinary conference and encourages new topics.
-
-\subsection{Revision of Previously-Reviewed\\ Manuscript}
-If the manuscript has been previously reviewed and rejected and is now being submitted to MICRO, the authors have an option of providing a letter explaining how the paper has been revised for this current submission.
-We expect this revision information to improve both the submission and the review process.
-Authors choosing to provide such a letter have control over who has access to it by specifying one of the following options:
-
-\begin{enumerate}
-\item Shared with all reviewers of the paper 
-\item Shared with reviewers who declare that they reviewed a prior version and who request the revision information
-\item Not shared with any PC member but available to the program chairs
-\end{enumerate}
-
-We encourage you to keep this letter concise and optionally append additional information, such as a version of the paper that highlights the differences or any other material of your choice.
-
-\subsection{Declaring Conflicts of Interest}
-Authors must register all their conflicts for their paper submission.
-Conflicts are needed to ensure appropriate assignment of reviewers.
-{\bf If a paper is found to have an undeclared conflict that causes a problem OR if a paper is found to declare false conflicts in order to abuse or ``game'' the review system, the paper may be rejected without review.} We use the following conflict of interest guidelines for determining the conflict period for MICRO 2022.
-Please declare a conflict of interest (COI) with the following people for any author of your paper:
-
-\begin{enumerate}
-\item Your Ph.D. advisor(s), post-doctoral advisor(s), Ph.D. students,
-      and post-doctoral advisees, forever.
-\item Family relations by blood or marriage, or their equivalent,
-      forever (if they might be potential reviewers).
-\item People with whom you have collaborated in the last FOUR years, including:
-  \begin{itemize}
-  \item co-authors of accepted/rejected/pending papers
-  \item co-PIs on accepted/rejected/pending grant proposals
-  \end{itemize}
-\item Ongoing collaboration that has not yet resulted in a paper or proposal submission. Justification may be queried.
-\item When there is a direct funding relationship between an author and the potential reviewer (e.g., the reviewer is a sponsor of an author's research on behalf of his/her company or vice versa) in the last FOUR years.
-\item People (including students) who shared your primary institution(s) in the last FOUR years.
-\item Other relationships, such as close personal friendship, that you think might tend
-to affect your judgment or be seen as doing so by a reasonable person familiar
-with the relationship.
-\end{enumerate}
-
-We would also like to emphasize that the following scenarios do {\em not} constitute a conflict:
-\begin{enumerate}
-\item Authors of previously-published, closely related work on that basis alone.
-\item ``Service'' collaborations such as co-authoring a report for a professional organization, serving on a program committee, or co-presenting tutorials.
-\item Co-authoring a paper that is a compendium of various projects with no true collaboration among the projects.
-\item People who work on topics similar to or related to those in your papers.
-\item Collaborators on large funded projects where there is no close collaboration and no joint benefit in the paper being accepted.
-\end{enumerate}
-
-We hope to draw most reviewers from the program committee, but others
-from the community may also write reviews. {\bf Please declare all your conflicts (not just restricted to the PC).} When in doubt, please contact the program chairs.
-
-%Please note that all paper submissions require all authors to electronically sign a statement confirming their best effort to accurately identify potential reviewers with a conflict of interest, and importantly also {\bf assuring that each author will make no explicit attempt to directly or indirectly influence any reviewer opinion or decision about the submitted paper}.
-Importantly, we do not consider technical discussion of a paper's content or any other sharing of content from the paper to violate the above policy.
-
-\subsection{Concurrent Submissions and Workshops}
-
-By submitting a manuscript to MICRO 2022, the authors guarantee that the manuscript has not been previously published or accepted for publication in a substantially similar form in any conference, journal, or the archived proceedings of a workshop (e.g., in the ACM/IEEE digital library) -- see exceptions below.
-The authors also guarantee that no paper that contains significant overlap with the contributions of the submitted paper will be under review for any other conference or journal or an archived proceedings of a workshop during the MICRO 2022 review period.
-Violation of any of these conditions will lead to rejection.
-
-The only exceptions to the above rules are for the authors' own papers in (1) workshops without archived proceedings such as in the ACM/IEEE digital library (or where the authors chose not to have their paper appear in the archived proceedings), or (2) venues such as IEEE CAL or arXiv where there is an explicit policy that such publication does not preclude longer conference submissions.
-In all such cases, the submitted manuscript may ignore the above work to preserve author anonymity.
-This information must, however, be provided on the submission form -- the program chairs will make this information available to reviewers if it becomes necessary to ensure a fair review.
-As always, if you are in doubt, it is best to contact the program chairs.
-
-Finally, the ACM/IEEE Plagiarism Policy (\href{http://www.acm.org/publications/policies/plagiarism_policy}{here} and \href{https://www.ieee.org/publications_standards/publications/rights/plagiarism.html}{here}) covers a range of ethical issues concerning the misrepresentation of other works or one's own work.
-
-\section{Ethics}
-
-\begin{enumerate}
-\item Authors must abide by the ACM code of ethics and the IEEE code of ethics
-\item Authors must not contact reviewers or PC members about any submission, including their own.
-This includes attempting to sway a reviewer, requesting information about any aspect of the reviewing process, and/or asking about the outcome of a submission.
-Similarly, authors are not allowed to ask another party to contact the reviewers on their behalf.
-\item Authors must not disclose the content of reviews for their paper publicly (e.g., on social media)  before the results are announced. 
-\item Authors must report any allegations of submission or reviewing misconduct to the program chairs.
-The only exception is if the complaint is about the program chairs; in this case, the Steering Committee should be contacted.
-\end{enumerate}
-
-\section*{ACKNOWLEDGMENTS}
-This document is derived from previous conferences, in particular MICRO 2013, ASPLOS 2015, MICRO 2015, MICRO 2016, MICRO 2017, MICRO
-2018, MICRO 2019, MICRO 2020 and MICRO 2021, as well as SIGARCH/TCCA's Recommended Best Practices for the Conference Reviewing Process.
-}
-
 %%%%%%% -- PAPER CONTENT ENDS -- %%%%%%%%
@@
 %%%%%%%%% -- BIB STYLE AND FILE -- %%%%%%%%
-% \bibliographystyle{IEEEtranS}
-% \bibliography{refs}
+\bibliographystyle{IEEEtranS}
+\bibliography{refs}
 %%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%

### 2022-04-16 docs/MICRO 2022/overview.tex
Finalize overview section

@@
+\newcommand{\Treebeard}{{\sc {Treebeard}}}
+\newcommand{\op}[1]{{\texttt {#1}}}
+
 \section{Compiler Overview}
@@
 Given an input model our compiler generates optimized inference code. Specifically it generates a 
-callable batch inference function \texttt{predictForest} that given an array of input rows, outputs an
+callable batch inference function \texttt{predictForest} that, given an array of input rows, computes an
 array of model predictions.
@@
 abstraction to a sequence of operations at a lower abstraction. Optimizations in Treebeard 
-are implemented using a combination of annotation and lowering. An Operation at a higher 
-abstraction is annotated with meta-data that indicates what kind of optimization is 
+are implemented using a combination of annotation and lowering. An operation at a higher 
+abstraction is annotated with attributes that indicate what kind of optimization is 
 to be performed while lowering it. The lowering transformation uses this information to generate 
@@
 %Treebeard first constructs a high level representation of the decision forest inference operation.
-Treebeard's IR has three abstractions as shown in Figure \ref{fig:Overview}.  
+Treebeard's IR has three abstractions as shown in Figure \ref{Fig:Overview}.  
 At the highest level (HIR), the input model is represented as a collection of binary trees. This is shown 
-in the second row of figure \ref{fig:Overview}. At this level of abstraction,
+in the second row of figure \ref{Fig:Overview}. At this level of abstraction,
 Treebeard tiles nodes together to transform a binary tree into an $n$-ary tree 
@@
 One objective of tree reordering is to group identically structured trees so that they can share the same traversal code. 
-With the tiling shown in Figure \ref{fig:Overview}, Tree1 and Tree3 have depth 2, while Tree2 has a depth of 3.
+With the tiling shown in Figure \ref{Fig:Overview}, Tree1 and Tree3 have depth 2, while Tree2 has a depth of 3.
 Therefore, the trees are reordered so that Tree1 and Tree3 are together.
@@
 to allow optimizations that are independent of the final memory layout of the model. 
-The lowering to MIR performs loop optimizations like loop-reordering, parallelization and unrolling.
-At this level of the IR (the third row in Figure \ref{fig:Overview}),
+The lowering to MIR performs loop transformations like loop tiling, permutation, parallelization and unrolling.
+At this level of the IR (the third row in Figure \ref{Fig:Overview}),
 the order in which trees, input row pairs are walked is explicitly represented in the IR using loop nests.
-Figure \ref{fig:Overview} shows two possible ways that loop nests could be generated -- the first
+Figure \ref{Fig:Overview} shows two possible ways that loop nests could be generated -- the first
 walks all rows for one tree before moving to the next tree while the second walks all trees for 
@@
 ensure that Tree1 and Tree3 share the same traversal code, while traversal code for Tree2 is different.
-Additionally, tree walk optimizations, such as tree walk unrolling (shown in figure \ref{fig:Overview}) and 
+Additionally, tree walk optimizations, such as tree walk unrolling (shown in figure \ref{Fig:Overview}) and 
 peeling are performed on the MIR.
@@
 
-% \begin{figure}
-%   \centering
-%   \includegraphics[width=\textwidth]{figures/CompilerOverview.png}
-%   \caption{Treebeard IR lowering and optimization details}
-%   \label{Fig:Overview}
-% \end{figure}
+\begin{figure*}
+  \centering
+  \includegraphics[width=\textwidth]{figures/OverviewExample.pdf}
+  \caption{Treebeard IR lowering and optimization details}
+  \label{Fig:Overview}
+\end{figure*}
 

### 2022-04-16 docs/MICRO 2022/overview.tex
Edit figure captions

@@
   \includegraphics[width=\linewidth]{figures/compiler.pdf}
-  \caption{Treebeard Compiler Structure}
+  \caption{Treebeard compiler structure}
   \label{Fig:CompilerStructure}
@@
   \includegraphics[width=\textwidth]{figures/OverviewExample.pdf}
-  \caption{Treebeard IR lowering and optimization details}
+  \caption{Treebeard IR lowering and optimization details. The three abstraction level in Treebeard's IR are shown. The
+           high level IR is a tree based IR to perform model level optimization, the mid-level IR is for
+           loop optimizations that are independent of memory layout and the low level IR allows us to perform
+           vectorization and other memory layout dependent optimizations.}
   \label{Fig:Overview}

### 2022-04-17 docs/MICRO 2022/uniftiling.tex

@@
 \subsection{Uniform Tiling}
+\label{sec:UnifTiling}
 % The justification for uniform tiling -- if we assume all paths are equally likely,
@@
       tree = getTree(forest, t) 
+      node = getRoot(tree)
       node = traverseTreeTile(tree, node, rows[i])

### 2022-04-17 docs/MICRO 2022/unrollandjam.tex

\subsection{Unroll and Jam}
While tiling and subsequent vectorization gives significant performance gains, profiling 
the generated code showed that true dependencies between instructions were still causing
a significant number of stall cycles. In order to fill these stall cycles, Treebeard does 
an unroll and jam on the inner most loops of the loop nest. This transformation has 
the effect of walking multiple tree and input row pairs in an interleaved fashion. 
The dependency stalls can be hidden by scheduling instructions from independent tree walks 
in the stall cycles.

This optimization is performed across both the mid-level IR and low-level IR. Loops on which 
to perform unroll and jam are identified in the mid-level IR. This information is communicated to the 
lowering passes by replacing these tree walks with specialized tree walk operations. The lowering 
passes that transform the mid-level IR to low-level IR interleave operations across independent
tree walks.

The following listing shows the mid-level IR when the inner loop over the input rows is unrolled 
by a factor of 2 and the two resulting tree walks are jammed together.

\begin{lstlisting}[style=c++]
  forest = ensemble(...)
  for t = 0 to 4 step 1 {
    for i = 0 to 2 step 2 {
      tree = getTree(forest, t)
      prediction1, prediction2 = InterleavedWalk((tree, rows[i]), (tree, rows[i+1]))
    }
  }
\end{lstlisting}

When lowered to low-level IR, the operations to traverse each of the tree, input row pairs 
(the arguments to the \texttt{InterleavedWalk}) are interleaved. One step of the interleaved 
walk is listed below. 
\begin{lstlisting}[style=c++]
  // ... 
  n1 = n2 = getRoot(tree)
  // ...
  threshold1 = loadThresholds(n1)
  threshold2 = loadThresholds(n2)
  featureIndex1 = loadFeatureIndices(n1)
  featureIndex2 = loadFeatureIndices(n2)
  feature1 = rows[i][featureIndex1]
  feature2 = rows[i][featureIndex2]
  pred1 = feature1 < threshold1
  pred2 = feature2 < threshold2
  n1 = getChildNode(n1, pred1)
  n2 = getChildNode(n2, pred2)
  // ...
\end{lstlisting}
These transformations are fairly general and are not aware of the in memory representation of the model. Therefore, they 
are reusable across different in memory representations - the ones that are currently built into Treebeard or ones that 
are added in the future.
\TODO{AP I feel the way this section is currently written makes the optimization seem extremely trivial. Is there a different 
way to present it?}

### 2022-04-17 docs/MICRO 2022/details.tex
Edits to tiling section

@@
 \input{unrollandjam}
-
-\subsection{Vectorization}
-Vectorization performed by Treebeard is enabled by the tiling transformations described in section \ref{sec:Tiling}. 
-When the low level IR is translated to LLVM IR, Treebeard generates LLVM instructions that operate on the threholds and feature indices 
-of nodes within a tile in a vector fashion. Therefore, thresholds and feature indices are loaded using vector loads and predicates are 
-evaluated using vector comparisons. These vector LLVM IR instructions are then converted to vector instructions in the target ISA by 
-the LLVM JIT.
+\input{vectorization}
 

### 2022-04-17 docs/MICRO 2022/probtiling.tex
Add interleaving, vectorization, parallelization
Edits to tiling section

@@
 \subsection{Probability Based Tiling}
+\label{sec:ProbTiling}
 \subsubsection{Motivation}
-As discussed in section \ref{sec:UnifTiling}, uniform tiling optimizes tree walks assuming that all paths in the tree from root to leaf are equally likely. 
-However, we find that this is not always the case. We performed inference on the complete training set of each of our benchmarks and measured how many
+Uniform tiling optimizes tree walks assuming that all root to leaf paths in the tree are equally likely. 
+However, we find that this is not always the case. We performed inference on the complete training set of each of our benchmarks and counted how many
 times each leaf was reached. While for some benchmarks all leaves are roughly equally likely, for others, a small number of leaves are extremely likely 
-while the rest aren't reached often. We call trees with a small number of extremely likely leaves \textbf{\emph{skewed}}.
+while the rest aren't reached often. We call trees with a small number of extremely likely leaves \textbf{\emph{leaf biased}}.
 \begin{figure}
@@
 \end{figure}
-Figures \ref{Fig:AirlineOHEStats} and \ref{Fig:EpsilonStats} show this. Each line in these graphs corresponds to a fixed fraction of the input (say $f$). 
-A point on a line at coordinate $(x, y)$ means that $y$ percent of trees in the model could handle a fraction $f$ of all training inputs with $x$ percent of 
-leaves. \TODO{Explain with a specific point in one of the graphs}
-Figure \ref{Fig:AirlineOHEStats} shows that very few leaves are reached for a very large fraction of inputs for the benchmark airline-ohe. 
-This means that a small fraction of leaves are very likely. On the other hand,
-figure \ref{Fig:EpsilonStats} shows that a large fraction of trees need a large fraction of their leaves to handle any fraction of the test input 
-for the benchmark epsilon. One other observation we made is that most models have some skewed trees while the rest of the trees have equally likely leaves.
-\TODO{AP Define a term for trees with roughly equally likely leaves.} We design the probability based tiling algorithm to take advantage of this property 
+
+We say that an input row $r_i$ is \textbf{\emph{covered}} by a subset of leaves $L' \subseteq L$ of a tree $T$, if the leaf $l$ reached by 
+walking $T$ for row $r_i$ is in $L'$. We show how different models (and even different trees within the same model) behave differently 
+using models for two benchmarks, airline-ohe and epsilon. Consider the graphs shown in   
+figures \ref{Fig:AirlineOHEStats} and \ref{Fig:EpsilonStats}. Each line in these graphs corresponds to a fixed fraction of the input (say $f$). 
+A point on a line at coordinate $(x, y)$ means that a fraction $y$ of trees in the model could cover a fraction $f$ of all training inputs with a fraction $x$ of 
+leaves. For example, the first point on the $f=0.9$ line in figure \ref{Fig:AirlineOHEStats} says that about 52\% of trees ($y$ value) need only 1\% of their
+leaves ($x$ value) to cover 90\% of the training input. 
+In general, Figure \ref{Fig:AirlineOHEStats} shows that very few leaves are needed to cover a very large fraction of inputs for the benchmark airline-ohe. 
+This means that a small fraction of leaves are very likely. On the other hand, for the benchmark epsilon,
+figure \ref{Fig:EpsilonStats} shows that a trees need a much larger fraction of their leaves to cover a significant fraction of the test input.
+This means that most trees in the epsilon model are not leaf biased.
+One other observation we make is that most models have some leaf biased trees while the rest of the trees have equally likely leaves.
+\TODO{AP Maybe define a term for trees with roughly equally likely leaves?} We design the probability based tiling algorithm to take advantage of this property 
 of decision tree ensembles.
-\TODO{This section needs to be rewritten. Define ``leaves covering inputs'' to clarify the presentation.}
-
-\subsubsection{Notation}
-In order to formulate the probability based tiling algorithm as an optimization problem, we define the following.
-\begin{enumerate}
-    \item For every leaf $l \in L$, we define $p_l$ as the probability that the leaf $l$ is reached.
-    \item For each node $n \in V$, we define the absolute probability $p_v$ as
-    \begin{equation}
-        p_v = \begin{cases}
-        p_l &\text{if $l \in L$}\\
-        p_{left(v)} + p_{right(v)} &\text{otherwise}
-        \end{cases}
-    \end{equation}
-    \item For any tree $T$, $\mathcal{C}(T)$ represents the set of all valid tilings of $T$.
-    \item For every $v \in V$, we define $S_v$ as the subtree rooted at $v$.
-    \item For every $v \in V$, we define $L_v$ as the set of leaves of $T_v$.
-    \item For a every tile $T_i$, we define $root(T_i)$ as the node $v \in T_i$ such that $v$ has no incoming edges from any other node $u \in T_i$.
-    \item For a tile $T_i$, $out(T_i) \subseteq E$ is the set of edges $(u, v)$ such that $u \in T_i$ and $v \notin T_i$.
-\end{enumerate}
 
@@
             \For{$(u,v) \in Out(Tile)$}
-                \State $Tiles \leftarrow Tiles \cup TileTree(T_v, n_t)$
+                \State $Tiles \leftarrow Tiles \cup TileTree(S_v, n_t)$
             \EndFor
@@
 lookup table needed (section \ref{sec:LookupTable}) caused increased L1 cache misses. In order to 
-alleviate this, we only perform probability based tiling on trees that are skewed. We say a tree is
-skewed if a small fraction of leaves, say $\alpha$, cover a large fraction of training inputs, say $\beta$.
+alleviate this, we only perform probability based tiling on trees that are leaf biased. We decide that 
+a tree is leaf biased if a small fraction of leaves, say $\alpha$, can cover a large fraction of training inputs, say $\beta$.
 We only perform probability based tiling on skewed trees and fall back to uniform tiling otherwise. 
 In our experiments we use $\alpha=0.05$ and $\beta=0.9$.
-
-\subsubsection{Code Generation}
-As probability based tiling pulls the most probable leaves of a decision tree nearest the root, it poses 
-some implementation challenges. By design, the tiling process makes the tree of tiles 
-imbalanced. Therefore, the simple array based representation described in section \ref{sec:ArrayBased}
-cannot be used because of the memory footprint blow up (a large part of the tree is empty, but would 
-need to be allocated). On the other hand, the sparse representation in section \ref{sec:SparseRep} adds 
-an extra hop for leaves that have non-leaf siblings. But this would mean that we add extra hops for 
-the most probable leaves after probability based tiling. To address these problems, Treebeard peels 
-the tree walk and specializes the leaf checks at higher levels to avoid the extra hop. Currently, 
-we determine the maximum depth of leaves needed to cover 90 percent of the inputs and peel the tree 
-walk by as many iterations. For example, consider the case where leaves until depth 2 are needed to 
-cover 90 percent of the training input. Then, Treebeard generates the following IR.
-
-\begin{lstlisting}{style=c++}
-    // ...
-    tree = getTree(forest, t)
-    node = getRoot(tree)
-    node = traverseTreeTile(tree, node, rows[i])
-    if (isLeafTile(node)) {
-        treePrediction = getLeafValue(tree, node)
-    } else {    
-        node = traverseTreeTile(tree, node, rows[i])
-        if (isLeafTile(node)) {
-            treePrediction = getLeafValue(tree, node)
-        } else {    
-            // Loop based traversal 
-        }
-    }
-    treePrediction = getLeafValue(tree, node)
-    // ...
-\end{lstlisting}
-
-The if statements check whether a node is a leaf tile and hence avoid the extra hop. The memory 
-requirement is also not increased because only a small fraction of leaves are represented as full tiles.
-While walk peeling is used to improve the performance of probability based tiling by specializing leaf tests,
-the transformation is by itself general and can be used in different contexts. For example, it could be used 
-to elide leaf checks until a depth $d$ is reached if we know all leaves are at a depth greater than $d$.
-
-One other issue that the code generator needs to handle is that walks of different trees in the same ensemble may 
-need to be peeled to different depths. In order to handle this, trees are reordered so that all trees 
-with equal peeling depth are grouped together and the loops in the IR are fissed so that tree walks 
-for these groups of trees can be specialized differently. This is very similar to the code generation strategy
-used for uniform tiling.

### 2022-04-17 docs/MICRO 2022/tiling.tex
Edits to tiling section

@@
 We represent a decision tree by $T = (V, E, r)$ where $V$ is the set of nodes, $E$ the set of edges and
-$r \in V$ is the root. For each node $n \in V$, the following are given.
+$r \in V$ is the root. For each node $n \in V$, we define the following.
 \begin{enumerate}
-    \item $threshold(n) \in \mathbb{R}$ which gives the threshold value for $n$.
-    \item $featureIndex(n) \in \mathbb{N}$ which gives the feature index for $n$.
+    \item $threshold(n) \in \mathbb{R}$, the threshold value for $n$.
+    \item $featureIndex(n) \in \mathbb{N}$, the feature index for $n$.
     \item $left(n) \in V$, the left child of $n$ or $\emptyset$ if $n$ is a leaf. If $left(n) \neq \emptyset$, then $(n, left(n)) \in E$.
@@
 \end{enumerate}
-We use $L \subset V$ to denote the set of leaves.
+We use $L \subseteq V$ to denote the set of leaves. % subseteq because tree could have single node
 
@@
 \label{sec:Tiling}
-Treebeard vectorizes tree walks by grouping nodes of a decision tree into \textbf{\emph{tiles}}.
+% Treebeard vectorizes tree walks by grouping nodes of a decision tree into \textbf{\emph{tiles}}.
 The nodes in a tile are evaluated concurrently using vector instructions.
 Once the nodes of the current tile are evaluated, a look up table is used to compute which child of the current tile to move to next.
-The listing below shows at a high level how a tiled tree is walked.
+Treebeard groups nodes of the decision tree into \textbf{\emph{tiles}}. Tiling provides two benefits. 
+\begin{enumerate}
+  \item It allows the compiler to generate vector code to traverse trees. Section \ref{sec:Vectorization} describes how Treebeard does this.
+  \item It enables spatial locality improvements by grouping together nodes that are likely to be accessed together. 
+\end{enumerate}
+Once nodes are grouped into tiles, an $n$-ary tree whose nodes are tiles is constructed. Treebeard then generates 
+optimized code to walk this tree. The listing below shows at a high level how a tiled tree is walked (This is not 
+true IR, but presented for clarity). 
 \begin{lstlisting}[style=c++]
-  // A lookup table that determines the child index of
-  // the next tile given the tile shape and the outcome
-  // of the vector comparison on the current tile
-  int16_t LUT[NUM_TILE_SHAPES, pow(2, TileSize)]
-
   ResultType Prediction_Function(...) {
     // ...
-    Node n = getRoot(tree)
-    while (isLeaf(tree, n)==false) do {
-      thresholds = loadThresholds(tree, n)
-      featureIndices = loadFeatureIndices(tree, n)
-      // Gather required feature from the current row
-      features = rows[i][featureIndices]
-      // Vector comparison of features and thresholds
-      comparison = features < thresholds
+    Tile t = getRootTile(tree)
+    while (!isLeaf(tree, t)) do {
+      // Evaluate conditions of all nodes in the tile
+      predicates = evaluateTilePredicates(t, rows[i])
 
-      // Pack bits in comparison vector into an integer
-      comparisonIndex = combineBitsIntoInt(comparison)
-
-      // Get child index of tile we need to move to
-      tileShape = loadTileShape(tree, n)
-      childIndex = LUT[tileShapeID, comparisonIndex]
-
-      // Move to the correct child of the current node
-      n = getChildNode(tree, n, childIndex) 
+      // Move to the correct child of the current tile
+      t = getChildTile(tree, t, predicates) 
     }
-    ThresholdType prediction = getLeafValue(n)
+    treePrediction = getLeafValue(t)
     // ...
@@
 
-To evaluate the current tile, the vector of thresholds is first loaded (\texttt{loadThresholds}).
-This vector contains the thresholds of all nodes in the tile.
-Then, the features required for comparison are gathered into a vector (lines 11 and 13).
-The feature vector is compared to the threshold vector and the child tile to move to is determined (lines 15 to 25).
-More details about tile shapes and the look up table are presented in subsequent sections.
+To compute the prediction of the tree, the predicates of all nodes in the tile are computed simultaneously (line 6). 
+Then, the computed predicate values are used to determine which child of the current tile to move to (in the 
+$n$-ary tree). This section presents the details of tiling and how Treebeard's general tiling infrastructure 
+can be used to develop tiling algorithms with different objectives (sections \ref{sec:UnifTiling}) and \ref{sec:ProbTiling}).
+The details of how Treebeard lowers predicate evaluation and moving to the correct child to use vector instructions 
+are described in section \ref{sec:Vectorization}.
 
-\subsection{Tiles and Tile Shapes}
-A \textbf{\emph{tile}} is a collection of connected non-leaf nodes of a decision tree.
-The path connecting any pair of nodes in the tile must fully be contained within the tile.
-The tile size $n_t$ is the number of nodes contained in a tile.
+\subsection{Tiles and Tree Tiling}
+\label{sec:ValidTiling}
+A \textbf{\emph{tile}} is a collection of connected non-leaf nodes of a decision tree. The path connecting any pair of nodes
+in the tile must be contained fully within the tile. The \textbf{\emph{tile size}} $n_t$ is the number of nodes contained in a tile.
 
-Informally, the \textbf{\emph{tile shape}} is the shape of the region that encloses all nodes in a tile in a diagram of the decision tree.
-More formally, for a tile size $n_t$, each unique legal binary tree containing $n_t$ nodes (nodes being indistinguishable) corresponds to a tile shape.
-
-Figure \ref{Fig:TileSize3Shapes} enumerates all tile shapes with a tile size of 3.
-There are a total of 5 tile shapes with size 3.
-The number of tile shapes with a tile size $n_t$, denoted by $NTS(n_t)$ is given by the following equation.
-
-\begin{equation}
-  NTS(n) = \sum_{k=0}^{n-1} NTS(k) \times NTS(n-k-1)
-\end{equation}
-
-where $NTS(0) = NTS(1) = 1$.
-
-\begin{figure}
-  \centering
-  \includegraphics[width=\linewidth]{figures/TileShapes_Size3.PNG}
-  \caption{All possible tile shapes with tile size $n_t=3$}
-  \label{Fig:TileSize3Shapes}
-\end{figure}
-
-\subsection{Valid Tiling of a Tree}
-\label{sec:ValidTiling}
-A tiling $\mathcal{T}$ of the tree $T = (V, E, r)$ with tile size $n_t$ is a partition $\{ T_1, T_2, ... ,T_m \}$ of the set $V$ such that 
+A \textbf{\emph{tiling}} $\mathcal{T}$ of the tree $T = (V, E, r)$ with tile size $n_t$ is a partition $\{ T_1, T_2, ... ,T_m \}$ of the set $V$ such that 
 \begin{enumerate}
@@
     \item $T_i \cap T_j = \emptyset$ for all $i, j \in [1, m]$ and $i \neq j$
+    \item Tiles are \textbf{connected}, i.e. for an $u, v \in T_i$, there is a (undirected) path connecting $u$ and $v$ fully contained in $T_i$.
+\end{enumerate}
+
+We say a tiling $\mathcal{T}$ is a \textbf{\emph{valid tiling}} iff
+\begin{enumerate}
   \item $|T_i| \leq n_t$ for all $i \in [1, m]$
@@
   \item Tiles are \textbf{maximal}, i.e. if $|T_i| < n_t$, then there is no $v \in V\backslash \{ T_i \cup L \}$ such that $(u, v) \in E$ for some $u \in T_i$. 
-    \item Tiles are \textbf{connected}, i.e. for an $u, v \in T_i$, there is a (undirected) path connecting $u$ and $v$ fully contained in $T_i$.
 \end{enumerate}
@@
 \end{enumerate}
-
-\subsection{Tile Shapes and Decision Tree Inference}
-\label{Sec:TileShapesAndDecisionTreeInference}
-Treebeard uses vector instructions to accelerate decision tree walks.
-Vector instructions are used to evaluate the predicates of all the nodes in a tile simultaneously.
-However, once the predicates of all the nodes in the tile are evaluated, computing the next tile to move to, given the outcome of the comparison depends on the tile shape of the current tile.
-To illustrate this problem, consider the case of the tiles of size 3 shown in figure \ref{Fig:TileTraversalTileSize3}.
-\begin{figure}
-  \centering
-  \includegraphics[width=\linewidth]{figures/TileTraversal_Size3.PNG}
-  \caption{Example tile traversals with tile size $n_t=3$}
-  \label{Fig:TileTraversalTileSize3}
-\end{figure}
-The diagram shows 3 of the 5 possible tile shapes for a tile size of 3.
-The nodes drawn in black are members of the tile $t_1$.
-The nodes in blue are the entry nodes of the children tiles of $t_1$.
-\TODO{Define entry nodes}
-
-To traverse a tile on an input row, first, the predicate of each node in the tile is computed.
-Subsequently, we need to determine which of the child tiles to move to next.
-Note that a true predicte (bit value 1) on a node implies a move to the left child and a false predicate (bit value 0) implies a move to the right child.
-
-In the diagram, the bit strings (written in red) show which child we need to move to given the outcomes of the comparison (the bits represent the comparison outcomes of nodes and are in the order of the nodes in the tile -- marked 0, 1 and 2 in the diagram, i.e., the MSB is the predicate outcome of node 0 and the LSB the predicate outcome of node 2).
-For example, for the first tile shape, if the predicate of all nodes are true (i.e.
-the comparison outcome is 111), the next node to evaluate is $a$.
-However, if the predicate of node 1 is false, then we need to move to $d$ regardless of the outcomes of nodes 2 and 3.
-It is easy to see from the diagram that, depending on the tile shape, the same predicate outcomes can mean moving to different children.
-For example, for the outcome "011", the next tile is the 4th child (node $d$) for the first two tile shapes while it is the 3rd child for the other tile shape (node $c$).
-
-\subsection{Lookup Table}
-\label{sec:LookupTable}
-A lookup table (LUT) is used to solve the problem described in section \ref{Sec:TileShapesAndDecisionTreeInference}, i.e.
-given the outcome of the comparisons on all nodes in a tile, determine the child tile we should evaluate next.
-The LUT is indexed by the tile shape and the comparison outcome.
-Formally, the LUT is a map.
-\[
-LUT : (TileShape, < Boolean \times n_t >) \rightarrow [0, n_t] \subset \mathbb{N}
-\]
-
-where $n_t$ is the tile size, $< Boolean \times n_t >$ is a vector of $n_t$ booleans.
-The value returned by the LUT is the index of the child of the current tile that should be evaluated next.
-For example, if we are evaluating the first tile $t$ in figure \ref{Fig:TileTraversalTileSize3}, and the result of the comparison is 110, then $LUT(TileShape(t), 110)=1$ since the tile we need to evaluate next is the tile with node $b$, which is the second child of the current tile.
-
-In order to realize this LUT in generated code, Treebeard associates a non-negative integer ID with every unique tile shape of the given tile size.
-The result of the comparison, a vector of booleans, can be interpreted as a 64-bit integer.
-Therefore, the LUT can be implemented as a 2 dimensional array.
-\begin{lstlisting}{style=c++}
-  int16_t LUT[NTS(n_t), pow(2, n_t)]  
-\end{lstlisting}
-
-Treebeard computes the values in the LUT statically as the tile size is a compile time constant.
-
-\subsection{In Memory Representation of Tiled Trees}
-Treebeard currently has two in memory representations for tiled trees - an array based representation and a sparse representation.
-Both representations use an array of structs to represent all tiles of the model.
-
-\subsubsection{Array Based Representation}
-\label{sec:ArrayBased}
-Each tree in the model is represented as an array of tiles using the standard representation of trees as arrays.
-The root node is at index 0 and for a node at index $n$ in the array, the index of its $i^{th}$ child is given by $(n_t + 1) n + (i + 1)$ (every node in the tree of tiles has $n_t + 1$ children).
-A tile is represented by an object of the following struct.
-\begin{lstlisting}{style=c++}
-  struct Tile {
-    // A vector of TileSize elements
-    <ThresholdType x TileSize> thresholds; 
-    <FeatureIndexType x TileSize> featureIndices;
-    // Integer that identifies the tile shape
-    TileShapeIDType tileShapeID; 
-  };  
-\end{lstlisting}
-\TODO{AP Is this level of detail really needed? Also, the vector type notation needs to be introduced somewhere.}
-Even though this representation is simple, the memory required for reasonable sized models is very large.
-The memory footprint is up to 20X that of the scalar representation.
-This memory bloat causes a performance degradation because the span of the L1 TLB is not sufficient to efficiently translate addresses for the whole model.
-\TODO{AP There are also some cache misses.
-Write this better.} Storing leaves as full tiles (even though leaves just have to represent one value) and the empty space introduced due to the array based representation of trees that are not complete account for most of the increase.
-The sparse representation described next tries to address these issues.
-
-\subsubsection{Sparse Representation}
-\label{sec:SparseRep}
-The sparse representation tries to address the large memory footprint of the array based representation by doing the following.
-
-\begin{itemize}
-  \item To eliminate the wasted space in the array representation, we add a child pointer to each tile. This points to the first child of the tile. All children of a tile are stored contiguously.
-  \item Leaves are stored in a separate array.
-We found that, across all our benchmarks, a large fraction of leaves are such that all their siblings are also leaves.
-Such leaves are directly moved into the leaves array.
-For leaves for which this property does not hold, an additional hop is added by making the leaf tile a comparison tile and all its children are made leaves with the same value as the original leaf.
-\end{itemize}
-
-\begin{figure}
-  \centering
-  \includegraphics[width=\linewidth]{figures/SparseRep_TileSize3.PNG}
-  \caption{Sparse representation with tile size $n_t=3$}
-  \label{Fig:SparseRep}
-\end{figure}
-
-Figure \ref{Fig:SparseRep} shows some of the details of the sparse representation.
-The tree on the left of the diagram is the actual decision tree with the nodes grouped into tiles $t_1$ and $t_2$.
-The tree on the right is the tree of tiles.
-The arrays depicted below show how the tree is represented in memory.
-The first array ($\texttt{tree}$) is an array of tiles and has 5 elements.
-Each element of the array represents a single tile and has the thresholds of the nodes, the feature indices, a tile shape ID and a pointer to the first child (shown explicitly in red).
-
-As a specific example, consider the tile $t_1$.
-The tile has four children -- $l_1$, $l_2$, $l_3$ and $t_2$ in that order (left to right).
-These tiles are stored contiguously in the $\texttt{tree}$ array and a pointer to the first of these, $l_1$ is stored in the tile $t_1$ (the index 1 is stored in the tile $t_1$ as shown).
-
-Now consider the tile $t_2$.
-Since all children of the tile $t_2$ are leaves, they are all moved into the $\texttt{leaves}$ array.
-To store a pointer into the $\texttt{leaves}$ array, we add $\texttt{len(tree)}$ to the element index in the $\texttt{leaves}$ array.
-The tile $t_2$'s child is the element at index 12 of the $\texttt{leaves}$ array.
-Therefore, the index $12 + 5 = 17$ is stored in the tile $t_2$.
-(Any index $i$ that is greater than the length of the $\texttt{tree}$ array is regarded as an index into the $\texttt{leaves}$ array.
-The index into the $\texttt{leaves}$ array is $i - \texttt{len(tree)}$.)
-
-The other aspect of the representation is that an extra hop is added for the leaves $l_1$, $l_2$ and $l_3$ in order to simplify code generation.
-This enforces the invariant that all leaves are stored in the leaves array and  simplifies checking whether we've reached a leaf.
-Therefore, 4 new leaves are added as children for each of the original leaves $l_1$, $l_2$ and $l_3$.
-Each of these 12 newly added leaves has the same value as its parent.
-These are the first 12 elements of the $\texttt{leaves}$ array.
-
-Even though we currently have implementations of the two representations detailed in sections \ref{sec:ArrayBased} and \ref{sec:SparseRep}, support for other representations is not hard to add.
-All optimizing passes that work on the high level and mid level IR will continue to work as is.
-Programmers need only provide new lowering passes for a few operations in the low level IR.

### 2022-04-17 docs/MICRO 2022/uniftiling.tex
Edits to tiling section

@@
+\subsubsection{More Notation}
+In order to formulate the probability based tiling algorithm as an optimization problem, we define the following.
+\begin{enumerate}
+    \item For every leaf $l \in L$, we define $p_l$ as the probability that the leaf $l$ is reached.
+    \item For each node $n \in V$, we define the absolute probability $p_v$ as
+    \begin{equation}
+        p_v = \begin{cases}
+        p_l &\text{if $l \in L$}\\
+        p_{left(v)} + p_{right(v)} &\text{otherwise}
+        \end{cases}
+    \end{equation}
+    \item For any tree $T$, $\mathcal{C}(T)$ represents the set of all valid tilings of $T$.
+    \item For every $v \in V$, we define $S_v$ as the subtree rooted at $v$.
+    \item For every $v \in V$, we define $L_v$ as the set of leaves of $S_v$.
+    \item For a every tile $T_i$, we define $root(T_i)$ as the node $v \in T_i$ such that $v$ has no incoming edges from any other node $u \in T_i$.
+    \item For a tile $T_i$, $out(T_i) \subseteq E$ is the set of edges $(u, v)$ such that $u \in T_i$ and $v \notin T_i$.
+\end{enumerate}
+
 \subsection{Uniform Tiling}
@@
 % we want to minimize the depth of speculation
-The first tree tiling algorithm we implement is uniform tiling. The basic idea of uniform tiling is to construct tiles
-with a tile size $n_t$ so that we evaluate 
+The first tree tiling algorithm we implement is \textbf{\emph{uniform tiling}}. The core idea of uniform tiling is to construct tiles,
+with a tile size $n_t$, so that we evaluate 
 the least possible number of nodes speculatively. Intuitively, this translates to minimizing the height of the constructed tile. We minimize
@@
 This traversal collects at most $n_t$ nodes
-and does not include any leaves in the tile so that the constraints listed in section \ref{sec:ValidTiling} are respected. Algorithm 
+and does not include any leaves in the tile so that the constraints listed in section \ref{sec:ValidTiling} are satisfied. Algorithm 
 \ref{Alg:UnifTilingAlgo} lists the algorithm for uniform tiling. Tiling starts at the root and constructs a tile $Tile$ by performing
@@
           \For{$(u,v) \in Out(Tile)$}
-              \State $Tiles \leftarrow Tiles \cup TileTree(T_v, n_t)$
+              \State $Tiles \leftarrow Tiles \cup TileTree(S_v, n_t)$
           \EndFor
@@
 %will typically be maximized by splitting a large number of inputs.}.
-Based on these observations, we pad the tree of tiles generated with uniform tiling so that all leaves are at the same depth.
+Based on these observations, we (optionally) pad the tree of tiles generated with uniform tiling so that all leaves are at the same depth.
 This transformation is performed on the high level IR after uniform tiling. 
 Once the trees have been padded to make all leaves equal depth, the tree walks are fully unrolled to evaluate a fixed 
-number of tiles and all leaf checks are elided.
+number of tiles and all leaf checks are omitted.
 
-One other complication the code generator needs to handle is the fact that trees potentially have different depths. In order to handle
+One other complication the code generator needs to handle is the fact that different trees in the model being 
+compiled potentially have different depths. In order to handle
 this, Treebeard sorts the trees by their depth. This ensures that all trees with equal depth are grouped together. Once this is done, 
@@
 
-\TODO{AP: This listing is unnecesarily long. Can we maybe leave out the loop bodies and say something like "depth 2 walk"?}
+\TODO{AP: This listing is unnecesarily long. Can we maybe leave out the loop bodies and say something like "depth 2 walk"? Should 
+we point to the figure in the overview section instead?}

### 2022-04-17 docs/MICRO 2022/vectorization.tex
Edits to tiling section

\subsection{Vectorization}
\label{sec:Vectorization}
Vectorization performed by Treebeard is enabled by the tiling transformations described in section \ref{sec:Tiling}. 
When the low level IR is translated to LLVM IR, Treebeard generates LLVM instructions that operate on the threholds and feature indices 
of nodes within a tile in a vector fashion. Therefore, thresholds and feature indices are loaded using vector loads and predicates are 
evaluated using vector comparisons. These vector LLVM IR instructions are then converted to vector instructions in the target ISA by 
the LLVM JIT.
\begin{lstlisting}[style=c++]
  // A lookup table that determines the child index of
  // the next tile given the tile shape and the outcome
  // of the vector comparison on the current tile
  int16_t LUT[NUM_TILE_SHAPES, pow(2, TileSize)]

  ResultType Prediction_Function(...) {
    // ...
    Node n = getRoot(tree)
    while (isLeaf(tree, n)==false) do {
      thresholds = loadThresholds(tree, n)
      featureIndices = loadFeatureIndices(tree, n)
      // Gather required feature from the current row
      features = rows[i][featureIndices]
      // Vector comparison of features and thresholds
      comparison = features < thresholds

      // Pack bits in comparison vector into an integer
      comparisonIndex = combineBitsIntoInt(comparison)

      // Get child index of tile we need to move to
      tileShape = loadTileShape(tree, n)
      childIndex = LUT[tileShapeID, comparisonIndex]

      // Move to the correct child of the current node
      n = getChildNode(tree, n, childIndex) 
    }
    ThresholdType prediction = getLeafValue(n)
    // ...
  }  
\end{lstlisting}
To evaluate the current tile, the vector of thresholds is first loaded (\texttt{loadThresholds}). This vector contains the thresholds of all nodes in the tile. Then, the features required for comparison are gathered into a vector (lines 11 and 13). The feature vector is compared to the threshold vector and the child tile to move to is determined (lines 15 to 25). More details about tile shapes and the look up table are presented in subsequent sections.

\subsection{Tile Shapes}
Informally, the \textbf{\emph{tile shape}} is the shape of the region that encloses all nodes in a tile in a diagram of the decision tree. More formally, for a tile size $n_t$, each unique legal binary tree containing $n_t$ nodes (nodes being indistinguishable) corresponds to a tile shape.

Figure \ref{Fig:TileSize3Shapes} enumerates all tile shapes with a tile size of 3. There are a total of 5 tile shapes with size 3. The number of tile shapes with a tile size $n_t$, denoted by $NTS(n_t)$ is given by the following equation.

\begin{equation}
  NTS(n) = \sum_{k=0}^{n-1} NTS(k) \times NTS(n-k-1)
\end{equation}

where $NTS(0) = NTS(1) = 1$.

\begin{figure}
  \centering
  \includegraphics[width=\linewidth]{figures/TileShapes_Size3.PNG}
  \caption{All possible tile shapes with tile size $n_t=3$}
  \label{Fig:TileSize3Shapes}
\end{figure}

\subsection{Tile Shapes and Decision Tree Inference}
\label{Sec:TileShapesAndDecisionTreeInference}
Treebeard uses vector instructions to accelerate decision tree walks. Vector instructions are used to evaluate the predicates of all the nodes in a tile simultaneously. However, once the predicates of all the nodes in the tile are evaluated, computing the next tile to move to, given the outcome of the comparison depends on the tile shape of the current tile. To illustrate this problem, consider the case of the tiles of size 3 shown in figure \ref{Fig:TileTraversalTileSize3}. 
\begin{figure}
  \centering
  \includegraphics[width=\linewidth]{figures/TileTraversal_Size3.PNG}
  \caption{Example tile traversals with tile size $n_t=3$}
  \label{Fig:TileTraversalTileSize3}
\end{figure}
The diagram shows 3 of the 5 possible tile shapes for a tile size of 3. The nodes drawn in black are members of the tile $t_1$. The nodes in blue are the entry nodes of the children tiles of $t_1$. \TODO{Define entry nodes}

To traverse a tile on an input row, first, the predicate of each node in the tile is computed. Subsequently, we need to determine which of the child tiles to move to next. Note that a true predicte (bit value 1) on a node implies a move to the left child and a false predicate (bit value 0) implies a move to the right child.

In the diagram, the bit strings (written in red) show which child we need to move to given the outcomes of the comparison (the bits represent the comparison outcomes of nodes and are in the order of the nodes in the tile -- marked 0, 1 and 2 in the diagram, i.e., the MSB is the predicate outcome of node 0 and the LSB the predicate outcome of node 2). For example, for the first tile shape, if the predicate of all nodes are true (i.e. the comparison outcome is 111), the next node to evaluate is $a$. However, if the predicate of node 1 is false, then we need to move to $d$ regardless of the outcomes of nodes 2 and 3. It is easy to see from the diagram that, depending on the tile shape, the same predicate outcomes can mean moving to different children. For example, for the outcome "011", the next tile is the 4th child (node $d$) for the first two tile shapes while it is the 3rd child for the other tile shape (node $c$).

\subsection{Lookup Table}
\label{sec:LookupTable}
A lookup table (LUT) is used to solve the problem described in section \ref{Sec:TileShapesAndDecisionTreeInference}, i.e. given the outcome of the comparisons on all nodes in a tile, determine the child tile we should evaluate next. The LUT is indexed by the tile shape and the comparison outcome. Formally, the LUT is a map.
\[
LUT : (TileShape, < Boolean \times n_t >) \rightarrow [0, n_t] \subset \mathbb{N}
\]

where $n_t$ is the tile size, $< Boolean \times n_t >$ is a vector of $n_t$ booleans. The value returned by the LUT is the index of the child of the current tile that should be evaluated next. For example, if we are evaluating the first tile $t$ in figure \ref{Fig:TileTraversalTileSize3}, and the result of the comparison is 110, then $LUT(TileShape(t), 110)=1$ since the tile we need to evaluate next is the tile with node $b$, which is the second child of the current tile.

In order to realize this LUT in generated code, Treebeard associates a non-negative integer ID with every unique tile shape of the given tile size. The result of the comparison, a vector of booleans, can be interpreted as a 64-bit integer. Therefore, the LUT can be implemented as a 2 dimensional array.
\begin{lstlisting}{style=c++}
  int16_t LUT[NTS(n_t), pow(2, n_t)]  
\end{lstlisting}

Treebeard computes the values in the LUT statically as the tile size is a compile time constant.

\subsection{In Memory Representation of Tiled Trees}
Treebeard currently has two in memory representations for tiled trees - an array based representation and a sparse representation. Both representations use an array of structs to represent all tiles of the model.

\subsubsection{Array Based Representation}
\label{sec:ArrayBased}
Each tree in the model is represented as an array of tiles using the standard representation of trees as arrays. The root node is at index 0 and for a node at index $n$ in the array, the index of its $i^{th}$ child is given by $(n_t + 1) n + (i + 1)$ (every node in the tree of tiles has $n_t + 1$ children). A tile is represented by an object of the following struct.
\begin{lstlisting}{style=c++}
  struct Tile {
    // A vector of TileSize elements
    <ThresholdType x TileSize> thresholds; 
    <FeatureIndexType x TileSize> featureIndices;
    // Integer that identifies the tile shape
    TileShapeIDType tileShapeID; 
  };  
\end{lstlisting}
\TODO{AP Is this level of detail really needed? Also, the vector type notation needs to be introduced somewhere.}
Even though this representation is simple, the memory required for reasonable sized models is very large. The memory footprint is up to 20X that of the scalar representation. This memory bloat causes a performance degradation because the span of the L1 TLB is not sufficient to efficiently translate addresses for the whole model. \TODO{AP There are also some cache misses. Write this better.} Storing leaves as full tiles (even though leaves just have to represent one value) and the empty space introduced due to the array based representation of trees that are not complete account for most of the increase. The sparse representation described next tries to address these issues.

\subsubsection{Sparse Representation}
\label{sec:SparseRep}
The sparse representation tries to address the large memory footprint of the array based representation by doing the following.

\begin{itemize}
  \item To eliminate the wasted space in the array representation, we add a child pointer to each tile. This points to the first child of the tile. All children of a tile are stored contiguously.
  \item Leaves are stored in a separate array. We found that, across all our benchmarks, a large fraction of leaves are such that all their siblings are also leaves. Such leaves are directly moved into the leaves array. For leaves for which this property does not hold, an additional hop is added by making the leaf tile a comparison tile and all its children are made leaves with the same value as the original leaf.
\end{itemize}

\begin{figure}
  \centering
  \includegraphics[width=\linewidth]{figures/SparseRep_TileSize3.PNG}
  \caption{Sparse representation with tile size $n_t=3$}
  \label{Fig:SparseRep}
\end{figure}

Figure \ref{Fig:SparseRep} shows some of the details of the sparse representation.
The tree on the left of the diagram is the actual decision tree with the nodes grouped into tiles $t_1$ and $t_2$. The tree on the right is the tree of tiles. The arrays depicted below show how the tree is represented in memory. The first array ($\texttt{tree}$) is an array of tiles and has 5 elements. Each element of the array represents a single tile and has the thresholds of the nodes, the feature indices, a tile shape ID and a pointer to the first child (shown explicitly in red).

As a specific example, consider the tile $t_1$. The tile has four children -- $l_1$, $l_2$, $l_3$ and $t_2$ in that order (left to right). These tiles are stored contiguously in the $\texttt{tree}$ array and a pointer to the first of these, $l_1$ is stored in the tile $t_1$ (the index 1 is stored in the tile $t_1$ as shown).

Now consider the tile $t_2$. Since all children of the tile $t_2$ are leaves, they are all moved into the $\texttt{leaves}$ array. To store a pointer into the $\texttt{leaves}$ array, we add $\texttt{len(tree)}$ to the element index in the $\texttt{leaves}$ array. The tile $t_2$'s child is the element at index 12 of the $\texttt{leaves}$ array. Therefore, the index $12 + 5 = 17$ is stored in the tile $t_2$. (Any index $i$ that is greater than the length of the $\texttt{tree}$ array is regarded as an index into the $\texttt{leaves}$ array. The index into the $\texttt{leaves}$ array is $i - \texttt{len(tree)}$.)

The other aspect of the representation is that an extra hop is added for the leaves $l_1$, $l_2$ and $l_3$ in order to simplify code generation. This enforces the invariant that all leaves are stored in the leaves array and  simplifies checking whether we've reached a leaf. Therefore, 4 new leaves are added as children for each of the original leaves $l_1$, $l_2$ and $l_3$. Each of these 12 newly added leaves has the same value as its parent. These are the first 12 elements of the $\texttt{leaves}$ array.

Even though we currently have implementations of the two representations detailed in sections \ref{sec:ArrayBased} and \ref{sec:SparseRep}, support for other representations is not hard to add. All optimizing passes that work on the high level and mid level IR will continue to work as is. Programmers need only provide new lowering passes for a few operations in the low level IR.

\subsubsection{Code Generation for Probability Based Tiling}
As probability based tiling pulls the most probable leaves of a decision tree nearest the root, it poses 
some implementation challenges. By design, the tiling process makes the tree of tiles 
imbalanced. Therefore, the simple array based representation described in section \ref{sec:ArrayBased}
cannot be used because of the memory footprint blow up (a large part of the tree is empty, but would 
need to be allocated). On the other hand, the sparse representation in section \ref{sec:SparseRep} adds 
an extra hop for leaves that have non-leaf siblings. But this would mean that we add extra hops for 
the most probable leaves after probability based tiling. To address these problems, Treebeard peels 
the tree walk and specializes the leaf checks at higher levels to avoid the extra hop. Currently, 
we determine the maximum depth of leaves needed to cover 90 percent of the inputs and peel the tree 
walk by as many iterations. For example, consider the case where leaves until depth 2 are needed to 
cover 90 percent of the training input. Then, Treebeard generates the following IR.

\begin{lstlisting}{style=c++}
    // ...
    tree = getTree(forest, t)
    node = getRoot(tree)
    node = traverseTreeTile(tree, node, rows[i])
    if (isLeafTile(node)) {
        treePrediction = getLeafValue(tree, node)
    } else {    
        node = traverseTreeTile(tree, node, rows[i])
        if (isLeafTile(node)) {
            treePrediction = getLeafValue(tree, node)
        } else {    
            // Loop based traversal 
        }
    }
    treePrediction = getLeafValue(tree, node)
    // ...
\end{lstlisting}

The if statements check whether a node is a leaf tile and hence avoid the extra hop. The memory 
requirement is also not increased because only a small fraction of leaves are represented as full tiles.
While walk peeling is used to improve the performance of probability based tiling by specializing leaf tests,
the transformation is by itself general and can be used in different contexts. For example, it could be used 
to elide leaf checks until a depth $d$ is reached if we know all leaves are at a depth greater than $d$.

One other issue that the code generator needs to handle is that walks of different trees in the same ensemble may 
need to be peeled to different depths. In order to handle this, trees are reordered so that all trees 
with equal peeling depth are grouped together and the loops in the IR are fissed so that tree walks 
for these groups of trees can be specialized differently. This is very similar to the code generation strategy
used for uniform tiling.

### 2022-04-17 docs/MICRO 2022/unrollandjam.tex
Update vectorization and unroll jam

@@
-\subsection{Unroll and Jam}
+\subsection{Tree Walk Interleaving}
 While tiling and subsequent vectorization gives significant performance gains, profiling 
@@
 to perform unroll and jam are identified in the mid-level IR. This information is communicated to the 
-lowering passes by replacing these tree walks with specialized tree walk operations. The lowering 
+lowering passes by annotating these tree walk operations as descibred in section \ref{sec:Overview}. The lowering 
 passes that transform the mid-level IR to low-level IR interleave operations across independent
@@
   forest = ensemble(...)
-  for t = 0 to 4 step 1 {
-    for i = 0 to 2 step 2 {
+  for t = 0 to numTrees step 1 {
+    for i = 0 to batchSize step 2 {
       tree = getTree(forest, t)
@@
 are reusable across different in memory representations - the ones that are currently built into Treebeard or ones that 
-are added in the future.
+maybe added in the future.
 \TODO{AP I feel the way this section is currently written makes the optimization seem extremely trivial. Is there a different 

### 2022-04-17 docs/MICRO 2022/vectorization.tex
Update vectorization and unroll jam

@@
 the LLVM JIT.
+
+The below listing shows some of the details of a vectorized tree walk. 
 \begin{lstlisting}[style=c++]
@@
 
-\subsection{Tile Shapes}
+\subsubsection{Tile Shapes}
 Informally, the \textbf{\emph{tile shape}} is the shape of the region that encloses all nodes in a tile in a diagram of the decision tree.
@@
 
-\subsection{Tile Shapes and Decision Tree Inference}
+\subsubsection{Tile Shapes and Decision Tree Inference}
 \label{Sec:TileShapesAndDecisionTreeInference}
@@
 
-To traverse a tile on an input row, first, the predicate of each node in the tile is computed.
+% To traverse a tile on an input row, first, the predicate of each node in the tile is computed.
 Subsequently, we need to determine which of the child tiles to move to next.
@@
 
-In the diagram, the bit strings (written in red) show which child we need to move to given the outcomes of the comparison (the bits represent the comparison outcomes of nodes and are in the order of the nodes in the tile -- marked 0, 1 and 2 in the diagram, i.e., the MSB is the predicate outcome of node 0 and the LSB the predicate outcome of node 2).
-For example, for the first tile shape, if the predicate of all nodes are true (i.e.
+In the diagram, the bit strings (written in red) show which child we need to move to given the outcomes of the comparison.
+The bits represent the comparison outcomes of nodes and are in the order of the nodes in the tile -- marked 0, 1 and 2 in the diagram, i.e., the MSB is the predicate outcome of node 0 and the LSB the predicate outcome of node 2.
+For example, for the first tile shape, if the predicates of all nodes are true (i.e.
 the comparison outcome is 111), the next node to evaluate is $a$.
-However, if the predicate of node 1 is false, then we need to move to $d$ regardless of the outcomes of nodes 2 and 3.
+% However, if the predicate of node 1 is false, then we need to move to $d$ regardless of the outcomes of nodes 2 and 3.
 It is easy to see from the diagram that, depending on the tile shape, the same predicate outcomes can mean moving to different children.
@@
 
-\subsection{Lookup Table}
+\subsubsection{Lookup Table}
 \label{sec:LookupTable}
 A lookup table (LUT) is used to solve the problem described in section \ref{Sec:TileShapesAndDecisionTreeInference}, i.e.
-given the outcome of the comparisons on all nodes in a tile, determine the child tile we should evaluate next.
+given the outcome of the comparisons of all nodes in a tile, determine the child tile we should evaluate next.
 The LUT is indexed by the tile shape and the comparison outcome.
@@
 Therefore, the LUT can be implemented as a 2 dimensional array.
-\begin{lstlisting}{style=c++}
-  int16_t LUT[NTS(n_t), pow(2, n_t)]  
-\end{lstlisting}
-
+% \begin{lstlisting}{style=c++}
+%   int16_t LUT[NTS(n_t), pow(2, n_t)]  
+% \end{lstlisting}
 Treebeard computes the values in the LUT statically as the tile size is a compile time constant.
+\TODO{AP What comes after subsubsection?}
 
@@
 Each tree in the model is represented as an array of tiles using the standard representation of trees as arrays.
-The root node is at index 0 and for a node at index $n$ in the array, the index of its $i^{th}$ child is given by $(n_t + 1) n + (i + 1)$ (every node in the tree of tiles has $n_t + 1$ children).
+The root node is at index 0 and for a node at index $n$ in the array, the index of its $i^{th}$ child is given by $(n_t + 1) n + (i + 1)$ (nodes in the tree of tiles have $n_t + 1$ children).
 A tile is represented by an object of the following struct.
@@
 \TODO{AP Is this level of detail really needed? Also, the vector type notation needs to be introduced somewhere.}
-Even though this representation is simple, the memory required for reasonable sized models is very large.
-The memory footprint is up to 20X that of the scalar representation.
-This memory bloat causes a performance degradation because the span of the L1 TLB is not sufficient to efficiently translate addresses for the whole model.
-\TODO{AP There are also some cache misses.
-Write this better.} Storing leaves as full tiles (even though leaves just have to represent one value) and the empty space introduced due to the array based representation of trees that are not complete account for most of the increase.
-The sparse representation described next tries to address these issues.
+Even though this representation is simple and efficient for small models, the memory required for bigger models is very large. 
+%The memory footprint is up to 20X that of the scalar representation.
+This memory bloat causes performance problems because the span of the L1 TLB is not sufficient to efficiently translate 
+addresses for the whole model. Storing leaves as full tiles (even though leaves just have to represent one value) and the
+empty space introduced due to the array based representation of trees that are not complete account for most
+of the increase.
+%The sparse representation described next tries to address these issues.
 
@@
 \begin{itemize}
-  \item To eliminate the wasted space in the array representation, we add a child pointer to each tile. This points to the first child of the tile. All children of a tile are stored contiguously.
-  \item Leaves are stored in a separate array.
-We found that, across all our benchmarks, a large fraction of leaves are such that all their siblings are also leaves.
-Such leaves are directly moved into the leaves array.
-For leaves for which this property does not hold, an additional hop is added by making the leaf tile a comparison tile and all its children are made leaves with the same value as the original leaf.
+  \item We add a child pointer to each tile to eliminate the wasted space in the array representation. This points to the first child of the tile. All children of a tile are stored contiguously.
+  \item Leaves are stored as a separate array of scalar values. Across all our benchmarks, after tiling a large fraction of 
+  leaves are such that all their siblings are also leaves. Such leaves are directly moved into the leaves array. For leaves
+  for which this property does not hold, an extra ``hop'' is added by making the original leaf tile a normal tile. All its
+  children are made leaves with the same value as the original leaf.
 \end{itemize}
@@
 Figure \ref{Fig:SparseRep} shows some of the details of the sparse representation.
-The tree on the left of the diagram is the actual decision tree with the nodes grouped into tiles $t_1$ and $t_2$.
-The tree on the right is the tree of tiles.
-The arrays depicted below show how the tree is represented in memory.
-The first array ($\texttt{tree}$) is an array of tiles and has 5 elements.
-Each element of the array represents a single tile and has the thresholds of the nodes, the feature indices, a tile shape ID and a pointer to the first child (shown explicitly in red).
+% The tree on the left of the diagram is the actual decision tree with the nodes grouped into tiles $t_1$ and $t_2$. The tree on the right is the tree of tiles.
+The arrays depicted below show how the tree is represented in memory. The first array ($\texttt{tree}$) is an array of tiles 
+and has 5 elements. Each element of the array represents a single tile and has the thresholds of the nodes, the feature
+indices, a tile shape ID and a pointer to the first child (shown explicitly in red).
 
@@
 
-Now consider the tile $t_2$.
-Since all children of the tile $t_2$ are leaves, they are all moved into the $\texttt{leaves}$ array.
+Now consider the tile $t_2$. Since all children of the tile $t_2$ are leaves, they are all moved into the $\texttt{leaves}$ array.
 To store a pointer into the $\texttt{leaves}$ array, we add $\texttt{len(tree)}$ to the element index in the $\texttt{leaves}$ array.
-The tile $t_2$'s child is the element at index 12 of the $\texttt{leaves}$ array.
-Therefore, the index $12 + 5 = 17$ is stored in the tile $t_2$.
-(Any index $i$ that is greater than the length of the $\texttt{tree}$ array is regarded as an index into the $\texttt{leaves}$ array.
-The index into the $\texttt{leaves}$ array is $i - \texttt{len(tree)}$.)
+The tile $t_2$'s child is the element at index 12 of the $\texttt{leaves}$ array. Therefore, the index $12 + 5 = 17$ is stored in 
+the tile $t_2$. Any index $i$ that is greater than the length of the $\texttt{tree}$ array is regarded as an index into the
+$\texttt{leaves}$ array. The index into the $\texttt{leaves}$ array is $i - \texttt{len(tree)}$.
 
-The other aspect of the representation is that an extra hop is added for the leaves $l_1$, $l_2$ and $l_3$ in order to simplify code generation.
-This enforces the invariant that all leaves are stored in the leaves array and  simplifies checking whether we've reached a leaf.
-Therefore, 4 new leaves are added as children for each of the original leaves $l_1$, $l_2$ and $l_3$.
-Each of these 12 newly added leaves has the same value as its parent.
-These are the first 12 elements of the $\texttt{leaves}$ array.
+The other aspect of the representation is that an extra hop is added for the leaves $l_1$, $l_2$ and $l_3$ in order to simplify
+code generation. This enforces the invariant that all leaves are stored in the leaves array and  simplifies checking whether
+we've reached a leaf. Therefore, 4 new leaves are added as children for each of the original leaves $l_1$, $l_2$ and $l_3$. 
+Each of these 12 newly added leaves has the same value as its parent. These are the first 12 elements of the $\texttt{leaves}$ array.
 
-Even though we currently have implementations of the two representations detailed in sections \ref{sec:ArrayBased} and \ref{sec:SparseRep}, support for other representations is not hard to add.
-All optimizing passes that work on the high level and mid level IR will continue to work as is.
-Programmers need only provide new lowering passes for a few operations in the low level IR.
+Even though we currently have implementations of the two representations detailed in sections \ref{sec:ArrayBased} 
+and \ref{sec:SparseRep}, support for other representations is not hard to add. All optimizing passes that work on 
+the high level and mid level IR will continue to work as is. Programmers need only provide new lowering passes for
+a few operations in the low level IR.
 
@@
 some implementation challenges. By design, the tiling process makes the tree of tiles 
-imbalanced. Therefore, the simple array based representation described in section \ref{sec:ArrayBased}
-cannot be used because of the memory footprint blow up (a large part of the tree is empty, but would 
-need to be allocated). On the other hand, the sparse representation in section \ref{sec:SparseRep} adds 
+imbalanced. The array based representation (section \ref{sec:ArrayBased})
+cannot be used because of the memory footprint increase (a large part of the tree is empty, but would need to be allocated).
+On the other hand, the sparse representation in section \ref{sec:SparseRep} adds 
 an extra hop for leaves that have non-leaf siblings. But this would mean that we add extra hops for 
-the most probable leaves after probability based tiling. To address these problems, Treebeard peels 
+the most probable leaves after probability based tiling which defeats the optimization.
+
+We address these challenges using a code generation strategy. Treebeard peels 
 the tree walk and specializes the leaf checks at higher levels to avoid the extra hop. Currently, 
@@
 
-The if statements check whether a node is a leaf tile and hence avoid the extra hop. The memory 
-requirement is also not increased because only a small fraction of leaves are represented as full tiles.
+The if statements check whether a node is a leaf tile and hence avoid the extra hop. 
+%The memory 
+%requirement is also not increased because only a small fraction of leaves are represented as full tiles.
 While walk peeling is used to improve the performance of probability based tiling by specializing leaf tests,
@@
 One other issue that the code generator needs to handle is that walks of different trees in the same ensemble may 
-need to be peeled to different depths. In order to handle this, trees are reordered so that all trees 
+need to be peeled to different depths. A strategy similar to what is used for uniform tiling is used to handle this.
+Trees are reordered so that all trees 
 with equal peeling depth are grouped together and the loops in the IR are fissed so that tree walks 
-for these groups of trees can be specialized differently. This is very similar to the code generation strategy
-used for uniform tiling.
+for these groups of trees can be specialized differently.
+% This is very similar to the code generation strategy used for uniform tiling.

### 2022-04-17 docs/MICRO 2022/probtiling.tex
unif tililng

@@
 \subsection{Probability Based Tiling}
+\subsubsection{More Notation}
+In order to formulate the probability based tiling algorithm as an optimization problem, we define the following.
+\begin{enumerate}
+    \item For every leaf $l \in L$, we define $p_l$ as the probability that the leaf $l$ is reached.
+    \item For each node $n \in V$, we define the absolute probability $p_v$ as
+    \begin{equation}
+        p_v = \begin{cases}
+        p_l &\text{if $l \in L$}\\
+        p_{left(v)} + p_{right(v)} &\text{otherwise}
+        \end{cases}
+    \end{equation}
+    \item For any tree $T$, $\mathcal{C}(T)$ represents the set of all valid tilings of $T$.
+    \item For every $v \in V$, we define $S_v$ as the subtree rooted at $v$.
+    \item For every $v \in V$, we define $L_v$ as the set of leaves of $S_v$.
+    \item For a every tile $T_i$, we define $root(T_i)$ as the node $v \in T_i$ such that $v$ has no incoming edges from any other node $u \in T_i$.
+    \item For a tile $T_i$, $out(T_i) \subseteq E$ is the set of edges $(u, v)$ such that $u \in T_i$ and $v \notin T_i$.
+\end{enumerate}
+
 \label{sec:ProbTiling}

### 2022-04-17 docs/MICRO 2022/tiling.tex
unif tililng

@@
 \subsection{Notation}
-\TODO{Notation needs to be introduced in the background section}
+\TODO{Notation needs to be introduced in the background section, Tiling should be the first subsection}
 We represent a decision tree by $T = (V, E, r)$ where $V$ is the set of nodes, $E$ the set of edges and
@@
 Once the nodes of the current tile are evaluated, a look up table is used to compute which child of the current tile to move to next.
+
+While a decision tree is naturally represented by a binary tree, 
+it is not the best representation for tree traversal as it (i) requires many memory accesses, 
+(ii) has poor branching structure and (iii) cannot make use of vector instructions. 
+This section proposes a tiling optimization where we group multiple nodes in a decision tree into a single tile, effectively transforming a binary tree into an $n$-ary tiled tree.
+This not only allows the compiler to generate vectorized code (see Section~\ref{sec:Vectorization} ) to traverse trees but also enables spatial locality improvements by grouping nodes that are likely to be accessed together.
+We demonstrate this with two tiling heuristics later in the section.
+
+\CommentOut{
 Treebeard groups nodes of the decision tree into \textbf{\emph{tiles}}. Tiling provides two benefits. 
@@
 true IR, but presented for clarity). 
+}
+
+Once trees are tiled Treebeard generates tree walks with the code structure shown below.
 \begin{lstlisting}[style=c++]
@@
 \end{lstlisting}
+The code is just an abstract representation of a tiled tree walk that enables efficient lowering of specific steps in subsequent stages. 
+\op{evaluateTilePredicates} (speculatively) computes the predicates on all nodes in a tile (line 6).
+Then \op{getChildTile} (line 9), uses the computed predicate values to determine which child of the current tile to move to.
+We defer a description of how these operators are lowered to a later section but focus on tiling algorithms in this section.
 
+\CommentOut{
 To compute the prediction of the tree, the predicates of all nodes in the tile are computed simultaneously (line 6). 
@@
 \label{sec:ValidTiling}
-A \textbf{\emph{tile}} is a collection of connected non-leaf nodes of a decision tree. The path connecting any pair of nodes
-in the tile must be contained fully within the tile. The \textbf{\emph{tile size}} $n_t$ is the number of nodes contained in a tile.
-
-A \textbf{\emph{tiling}} $\mathcal{T}$ of the tree $T = (V, E, r)$ with tile size $n_t$ is a partition $\{ T_1, T_2, ... ,T_m \}$ of the set $V$ such that 
-\begin{enumerate}
-    \item $T_1 \cup T_2 ... \cup T_m = V$
-    \item $T_i \cap T_j = \emptyset$ for all $i, j \in [1, m]$ and $i \neq j$
-    \item Tiles are \textbf{connected}, i.e. for an $u, v \in T_i$, there is a (undirected) path connecting $u$ and $v$ fully contained in $T_i$.
-\end{enumerate}
-
-We say a tiling $\mathcal{T}$ is a \textbf{\emph{valid tiling}} iff
-\begin{enumerate}
-  \item $|T_i| \leq n_t$ for all $i \in [1, m]$
-  \item $\forall l \in L$ : $l \in T_i \rightarrow v \notin T_i \;\; \forall v \in V \backslash \{l\}$
-  \item Tiles are \textbf{maximal}, i.e. if $|T_i| < n_t$, then there is no $v \in V\backslash \{ T_i \cup L \}$ such that $(u, v) \in E$ for some $u \in T_i$. 
-\end{enumerate}
-
+}
+\subsection{Conditions for Tiling}
+\label{sec:ValidTiling}
+While any arbitrary partitioning of the nodes of a tree could be considered for tiling we impose a few intuitive constraints to only allow nodes that are likely to be accessed together to be grouped into a tile.
+\TODO{kr : replace $n_t$ with sz?}
+Specifically given $tree = (V, E, r)$ and a tile size $n_t$ we impose the following constraints on the generated tiles $\{ T_1, T_2, ... ,T_m \}$ .
+\begin{description}
+    \item[Partitioning] $T_1 \cup T_2 ... \cup T_m = V$ and $T_i \cap T_j = \emptyset$ for all $i\neq j$
+    \item[Connectedness] If $u, v \in T_i$, there is a (undirected) path connecting $u$ and $v$ fully contained in $T_i$.
+   \item [Leaf seperation] $\forall l \in L$ : $l \in T_i \rightarrow v \notin T_i \;\; \forall v \in V \backslash \{l\}$
+  \item [Maximal tiling] if there are tiles such that. $|T_i| < n_t$ then there is no $v \in V\backslash \{ T_i \cup L \}$ such that $(u, v) \in E$ for some $u \in T_i$. 
+\end{description}
+The \textbf{partitioning} and \textbf{maximal tiling} constraints together ensure that we group nodes into as few tiles as possible.
+{\textbf{Connectedness}} ensures that each tile is a sub-tree, a natural grouping of nodes that are accessed together.
+{\textbf{Leaf separation}} is critical to ensure that leaf checks can be efficiently performed.
+\TODO{kr: improve the above}
+\CommentOut{
 \subsection{Tiled Trees}
@@
 \end{enumerate}
+   }

### 2022-04-17 docs/MICRO 2022/uniftiling.tex
unif tililng

@@
-\subsubsection{More Notation}
-In order to formulate the probability based tiling algorithm as an optimization problem, we define the following.
-\begin{enumerate}
-    \item For every leaf $l \in L$, we define $p_l$ as the probability that the leaf $l$ is reached.
-    \item For each node $n \in V$, we define the absolute probability $p_v$ as
-    \begin{equation}
-        p_v = \begin{cases}
-        p_l &\text{if $l \in L$}\\
-        p_{left(v)} + p_{right(v)} &\text{otherwise}
-        \end{cases}
-    \end{equation}
-    \item For any tree $T$, $\mathcal{C}(T)$ represents the set of all valid tilings of $T$.
-    \item For every $v \in V$, we define $S_v$ as the subtree rooted at $v$.
-    \item For every $v \in V$, we define $L_v$ as the set of leaves of $S_v$.
-    \item For a every tile $T_i$, we define $root(T_i)$ as the node $v \in T_i$ such that $v$ has no incoming edges from any other node $u \in T_i$.
-    \item For a tile $T_i$, $out(T_i) \subseteq E$ is the set of edges $(u, v)$ such that $u \in T_i$ and $v \notin T_i$.
-\end{enumerate}
-
-\subsection{Uniform Tiling}
+\subsection{Basic Tiling Algorithm}
 \label{sec:UnifTiling}
@@
 % we want to minimize the depth of speculation
-The first tree tiling algorithm we implement is \textbf{\emph{uniform tiling}}. The core idea of uniform tiling is to construct tiles,
-with a tile size $n_t$, so that we evaluate 
-the least possible number of nodes speculatively. Intuitively, this translates to minimizing the height of the constructed tile. We minimize
-the height of the tile by performing a level order traversal starting at the root node of the tile. 
-\TODO{AP Should we say this is optimal when probabilities are equal? Problem is we haven't said what optimal is.}
-This traversal collects at most $n_t$ nodes
-and does not include any leaves in the tile so that the constraints listed in section \ref{sec:ValidTiling} are satisfied. Algorithm 
-\ref{Alg:UnifTilingAlgo} lists the algorithm for uniform tiling. Tiling starts at the root and constructs a tile $Tile$ by performing
-a level order traversal. Once the current tile is constructed, the tiling procedure is recursively performed on all nodes that are 
-destinations for edges going out of the constructed tile. The set of tiles constructed by algorithm \ref{Alg:UnifTilingAlgo} constitutes 
-a uniform tiling of the input tree.
+Algorithm~\ref{Alg:UnifTilingAlgo} shows a simple tiling algorithm that satisfies all the above constraints. 
+Tiling starts at the root and constructs a tile $Tile$ by performing
+a level order traversal.
+The call \op{LevelOrderTraversal(r, $n_t$)} picks the next $n_t$ nodes according to the standard level order tree traversal algorithm (not described here) applied only to internal nodes of the tree.
+Once the current tile is constructed, the tiling procedure is recursively performed on all nodes that are
+destinations for edges going out of the constructed tile.
+It is easy to see that the set of tiles constructed by algorithm \ref{Alg:UnifTilingAlgo} satisfies the constraints above.
+\TODO{Pass a modified tree without leafs to the level order traversal call. Explicitly add leaves as seperate tiles}
 % Algorithm for uniform tiling
@@
   \begin{algorithmic}
-      \Procedure{TileTree}{$T = (V, E, r)$, $n_t$} 
+      \Procedure{TileTree}{$tree = (V, E, r)$, $n_t$} 
           \If {$r \in L$}

### 2022-04-17 docs/MICRO 2022/introduction.tex
Add intro draft

\section{Introduction}
% Decision trees are among the most widely used ML models. Cite Kaggle and looking glass 
Decision tree ensembles are one of the most popular classes of machine learning models \cite{KaggleSurvey,LookingGlass}.
They are generated by machine learning techniques like gradient boosting and random forests. 
The Kaggle state of data science and machine learning survey \cite{KaggleSurvey} shows that 
these are the most widely used classes of models among data scientists. An analysis of 
several machine learning pipelines that are in production use in a real large scale web company showed that 
gradient boosting machines and random forests were the two most widely used ML algorithms in the company \cite{LookingGlass}.
Not only are decision tree based models widely used, they are also used in a diverse set of 
applications\cite{DecisionTreesOverview}. GBM models were used in the CERN large hadron collider
to classify particles based on the data collected from the collider\cite{LHCModel}. Other applications 
include search engines\cite{YahooSearch}, prediction of the financial performance of companies\cite{Finance},
medical diagnostics\cite{Med1, Med2} and recommendation and notification systems\cite{Facebook}.

% Decision tree ensemble overview
Each node of a decision tree has a threshold value $v$ and a feature index (or column index) $i$.
To compute the prediction of a decision tree on an input row $x$\footnote{An input row is a
vector of numbers.}, a path from the root to leaf is computed. At a node, if $x_i < v$, then 
the walk moves to the left child. If not, we move to the right child. When a leaf is reached, 
the value of the leaf is returned as the prediction of the tree. 
Most ML techniques combine several decision trees into a forest (or ensemble) to improve prediction accuracy.
To compute the prediction of the forest, the prediction of each tree is computed and these predictions are 
then combined (usually by adding them).

% Inference on decision trees is hard 
% Pointer chasing -- cache performance, branch prediction, dependency stalls, not easy to vectorize
Given the wide spread use of decision tree ensembles, the ability to perform both high throughput and low latency inference is important.
\TODO{AP Since we don't talk about latency or throughput explicitly, should we be saying this?}
However, the tree walks required for inference are not straight-forward to optimize on a modern CPU. 
Firstly, naively implemented tree walks have poor spatial and temporal locality and therefore poor cache performance \cite{FAST,MilindTreeVectorization}.
Additionally, since the decision of which node to evaluate next can only be made after evaluating the current node,
true dependencies between instructions cause several pipeline stalls. Also, using SIMD instructions to accelerate 
tree walks is extremely challenging \cite{MilindTreeVectorization}.

Several systems like XGBoost\cite{XGBoost}, LightGBM\cite{LightGBM} and Sklearn\cite{Sklearn} implement
decision tree ensemble inference. These are libraries and implement some optimizations. However, these optimizations are hand
coded and need to be manually rewritten or redesigned for every target that is supported. Also, these 
systems cannot apply specialize inference code depending on the model. A compiler based approach is 
needed to address these problems. However, existing compilers only support fixed code generation techniques\cite{Treelite}.
To fill the need for an extensible compiler infrastructure for decision forest ensembles, we design and 
build Treebeard\footnote{In J. R. R. Tolkein's The Lord of the Rings, Treebeard was the oldest of the Ents left in Middle-earth, an 
ancient tree-like being who was a ``shepherd of trees''.}, an optimizing compiler for decision tree ensemble inference.
Our motivations for building Treebeard are as follows. 
\begin{enumerate}
  \item Existing libraries \cite{XGBoost, Treelite, LightGBM, VPred} perform specific optimizations and are hard to maintain/upgrade as hardware evolves. 
  Additionally, they cannot tailor inference code to the model being used. Past work has identified that the optimizations required change with the model
  and architectural parameters like cache size\cite{CacheConscious1, CacheConscious2, VPred}. 
  \item The repeated effort to optimize and maintain libraries on several targets is prohibitive. The proliferation of new architectures
  further exacerbates this problem. Compilers have been successful in alleviating this problem in other domains\cite{Halide, TVM}.
  However, no such compiler infrastructure currently exists for decision tree ensembles.
  \item Currently, no system exists that allows a thorough exploration of the potential optimization space for decision tree ensemble inference. 
  Compilers have been successful in addressing this problem with ML models like DNNs \cite{TVM, Tiramisu, XLA}. However, compiler techniques for
  decision tree ensembles are less well studied.
\end{enumerate}

We make the following contributions.
\begin{enumerate}
  \item We design and build Treebeard, an extensible compiler infrastructure for decision tree model inference. The 
  infrastructure is built to allow exploration of optimization and code generation techniques. \TODO{Is this claim too grand?}
  \item We develop a general infrastructure for the vectorization of decision tree walks based on grouping tree nodes into ``tiles''.
   This includes general support for code generation and the in-memory representation of tiled trees. The infrastructure can be 
   used to tile trees based on different cost functions.
  \item We show that trees can be tiled using different cost functions. We present two novel tiling methods that are implemented
  using the general tiling infrastructure. 
  \item We design and implement various model and loop transformations that significantly improve generated inference code performance 
  to show the power of the proposed framework.
\end{enumerate}

### 2022-04-17 docs/MICRO 2022/probtiling.tex
Optimizations pass

@@
 \subsection{Probability Based Tiling}
+\label{sec:ProbTiling}
+%\subsubsection*{Motivation}
+
+The next algorithm we propose exploits the inherent biases among the leaves of a decision tree.
+In typical machine learning models some leafs (equivalently outcomes or predictions) are more likely to be reached than others.
+In such settings, having balanced tiled trees is not sufficient to minimize expected inference time.
+
+Consider for example two machine learning models \op{airline-ohe} and \op{epsilon} (also used in our evaluation). 
+Consider the graphs shown in   
+figures \ref{Fig:AirlineOHEStats} and \ref{Fig:EpsilonStats} that are generated from training data. Each line in these graphs corresponds to a fixed fraction of the input (say $f$). 
+A point on a line at coordinate $(x, y)$ means that a fraction $y$ of trees in the model could cover a fraction $f$ of all training inputs with a fraction $x$ of 
+leaves. For example, the first point on the $f=0.9$ line in figure \ref{Fig:AirlineOHEStats} says that about 52\% of trees ($y$ value) need only 1\% of their
+leaves ($x$ value) to cover 90\% of the training input. 
+In general, Figure \ref{Fig:AirlineOHEStats} shows that very few leaves are needed to cover a very large fraction of inputs for the benchmark \op{airline-ohe}. 
+This means that a small fraction of leaves are very likely. 
+We call trees with a small number of extremely likely leaves \textbf{\emph{leaf biased}}.
+
+On the other hand, for the benchmark \op{epsilon},
+figure \ref{Fig:EpsilonStats} shows that a trees need a much larger fraction of their leaves to cover a significant fraction of the training input.
+This means that most trees in the \op{epsilon} model are not leaf biased.
+
+\CommentOut{
 \subsubsection{More Notation}
@@
 \end{enumerate}
+}
 
-\label{sec:ProbTiling}
-\subsubsection{Motivation}
-Uniform tiling optimizes tree walks assuming that all root to leaf paths in the tree are equally likely. 
-However, we find that this is not always the case. We performed inference on the complete training set of each of our benchmarks and counted how many
-times each leaf was reached. While for some benchmarks all leaves are roughly equally likely, for others, a small number of leaves are extremely likely 
-while the rest aren't reached often. We call trees with a small number of extremely likely leaves \textbf{\emph{leaf biased}}.
 \begin{figure}
@@
 \end{figure}
-
+\CommentOut{
 We say that an input row $r_i$ is \textbf{\emph{covered}} by a subset of leaves $L' \subseteq L$ of a tree $T$, if the leaf $l$ reached by 
@@
 of decision tree ensembles. 
-
+}
 \subsubsection{The Optimization Problem}
@@
 
-We observe that the latency of one tree walk is proportional to the number of tiles that need to be evaluated to reach the leaf.
-Therefore, the average inference latency can be modelled as the expected number of tiles that are evaluated to compute one tree prediction.
-More formally, the problem is to find a tiling $\mathcal{T}$ such that the following objective is minimized.
+Observe that the latency of one tree walk is proportional to the number of tiles that need to be evaluated to reach the leaf.
+It is easy to see that for a leaf biased tree, basic tilling does not optimize for this objective, it considers all leafs to be equally likely.
+
+The goal of probablistic tiling is to minimize the average inference latency, or equivalently the minimize the expected number of tiles that are evaluated to compute one tree prediction.
+More formally, the problem is to find a \emph{valid} (as defined in Section~\ref{sec:ValidTiling}) tiling $\mathcal{T}$ such that the following objective is minimized.
 \[
@@
 where the minimization is over all valid tilings $\mathcal{T}$ of the tree $T$, $depth_{\mathcal{T}}(l)$ is the depth of the leaf $l$ given tiling ${\mathcal{T}}$.
+$p_l$ is the probability of of reaching leaf $l$ as observed during training.
 
-The constraints on the optimization are the constraints listed in Section \ref{sec:ValidTiling}. 
-The above optimization problem can be solved optimally using dynamic programming. We leave this out in the interest of space. 
-Instead, we use the simple greedy algorithm listed in algorithm \ref{Alg:GreedyTilingAlgo} to construct a valid tiling given the node probabilities.
+The above optimization problem can be solved optimally using dynamic programming. 
+We leave this out in the interest of space. 
+Instead, we use the simple greedy algorithm listed in algorithm \ref{Alg:GreedyTilingAlgo} to construct a valid tiling given the node probabilities\footnote{Probabilites for internal nodes can be computed from probablities for leafs by summing up the probabilities of all leafs that belong to the sub-tree rooted at the internal node.
+Leaf probabilities are collected during training.}.
 The algorithm starts at the root and greedily keeps adding the most probable legal node to the current tile until the maximum tile size is reached.
@@
 % Talk about problems with increasing number of tile shapes and only performing such tiling on skewed trees
+\CommentOut{
 When we tried to apply algorithm \ref{Alg:GreedyTilingAlgo} on all trees in our benchmarks, we found
@@
 lookup table needed (section \ref{sec:LookupTable}) caused increased L1 cache misses. In order to 
-alleviate this, we only perform probability based tiling on trees that are leaf biased. We decide that 
-a tree is leaf biased if a small fraction of leaves, say $\alpha$, can cover a large fraction of training inputs, say $\beta$.
-We only perform probability based tiling on skewed trees and fall back to uniform tiling otherwise. 
-In our experiments we use $\alpha=0.05$ and $\beta=0.9$.
+alleviate this, we only perform probability based tiling on trees that are leaf biased.
+}
+We find probability based tiling is only beneficial for leaf biased trees\footnote{Turns out that probability based tiling produces many more tile shapes (see Section~\ref{sec:tileShapes}) which direclty impacts the cost of \op{getChildTile} making it more expensive than basic tiling}.
+Recall that
+a tree to be leaf biased if a small fraction of leaves, say $\alpha$, can cover a large fraction of training inputs, say $\beta$.
+We only perform probability based tiling on trees with thresholds $\alpha=0.05$ and $\beta=0.9$ and fall back to uniform tiling otherwise.

### 2022-04-17 docs/MICRO 2022/uniftiling.tex
Optimizations pass

@@
 
+One interesting property of this tilling algorithm is that it naturally reduces the imbalance in trees, especially at large tile sizes.
+As the algorithm traverses down to sparser level of the tree, it naturally groups sub-trees containing chains of nodes, thus balancing the trees.
+While its possible to further enhance the algorithm to explicitly balance tiled trees, we find that basic tiling suffices in practice.
+\CommentOut{
 \subsubsection{Further Opimization and Code Generation}
@@
 we point to the figure in the overview section instead?}
+}

### 2022-04-18 docs/MICRO 2022/peelandunroll.tex

\subsection{Loop peeling and walk unrolling}
\Treebeard{} splits the loop that performs a tree walk into two parts. As it is aware of the depth of the first leaf in a tree, it peels and introduces a \op{prologue} loop that walks down the tree a constant number of steps (up to first leaf) and then performs the rest of the tree walk in a separate loop. Further \Treebeard{} unrolls the prologue completely, avoiding all traversal induced branching in it.

Note that for trees where \Treebeard{} has already padded and balanced the tree (Section~\ref{sec:treeorder}), walk unrolling completely avoids all traversal induced spills.
\TODO{kr: Add example?}

### 2022-04-18 docs/MICRO 2022/tiling.tex

@@
 }
-\subsection{Conditions for Tiling}
+\subsection*{Conditions for Efficient Tiling}
 \label{sec:ValidTiling}
-While any arbitrary partitioning of the nodes of a tree could be considered for tiling we impose a few intuitive constraints to only allow nodes that are likely to be accessed together to be grouped into a tile.
+While any arbitrary partitioning of the nodes of a tree could be considered for tiling we impose a few intuitive constraints.
+% to only allow nodes that are likely to be accessed together to be grouped into a tile. 
 \TODO{kr : replace $n_t$ with sz?}
-Specifically given $tree = (V, E, r)$ and a tile size $n_t$ we impose the following constraints on the generated tiles $\{ T_1, T_2, ... ,T_m \}$ .
+Given a $tree = (V, E, r)$ and a tile size $n_t$ we impose the following constraints on the generated tiles $\{ T_1, T_2, ... ,T_m \}$ .
 \begin{description}
@@
 The \textbf{partitioning} and \textbf{maximal tiling} constraints together ensure that we group nodes into as few tiles as possible.
-{\textbf{Connectedness}} ensures that each tile is a sub-tree, a natural grouping of nodes that are accessed together.
-{\textbf{Leaf separation}} is critical to ensure that leaf checks can be efficiently performed.
-\TODO{kr: improve the above}
+{\textbf{Connectedness}} ensures that each tile is a sub-tree, a natural grouping of nodes that are likely to be accessed together.
+The {\textbf{Leaf separation}} constraint ensures that leafs are not tiled along with internal nodes.
+Leafs in a decision tree need special handling, they are used to check for walk termination and to determine the output (prediction).
+This constraint ensures that tiles are homogenous, this in-turn allows us to specialize the in-memory layout of trees and also simplifies code generation.
+We discuss leaf handling and tree layout in Section~\ref{sec:}.
+We refer to any tiling that satisties the above constraints as a \emph{valid} tilling.
 \CommentOut{

### 2022-04-18 docs/MICRO 2022/treeorder.tex

--- 2022-04-17 docs/MICRO 2022/uniftiling.tex
@@
-\subsection{Basic Tiling Algorithm}
-\label{sec:UnifTiling}
-% The justification for uniform tiling -- if we assume all paths are equally likely,
-% we want to minimize the depth of speculation
-Algorithm~\ref{Alg:UnifTilingAlgo} shows a simple tiling algorithm that satisfies all the above constraints. 
-Tiling starts at the root and constructs a tile $Tile$ by performing
-a level order traversal.
-The call \op{LevelOrderTraversal(r, $n_t$)} picks the next $n_t$ nodes according to the standard level order tree traversal algorithm (not described here) applied only to internal nodes of the tree.
-Once the current tile is constructed, the tiling procedure is recursively performed on all nodes that are
-destinations for edges going out of the constructed tile.
-It is easy to see that the set of tiles constructed by algorithm \ref{Alg:UnifTilingAlgo} satisfies the constraints above.
-\TODO{Pass a modified tree without leafs to the level order traversal call. Explicitly add leaves as seperate tiles}
-% Algorithm for uniform tiling
-\begin{algorithm}
-  \caption{Uniform tree tiling}
-  \label{Alg:UnifTilingAlgo}
-  \begin{algorithmic}
-      \Procedure{TileTree}{$tree = (V, E, r)$, $n_t$} 
-          \If {$r \in L$}
-              \State \textbf{return} $\{ r \}$
-          \EndIf
-          \State \textcolor{codegreen}{\textit{//Level order traversal to collect $n_t$ or fewer nodes. }}
-          \State \textcolor{codegreen}{\textit{//Leaves are not included in the constructed tile. }}
-          \State $Tile \leftarrow LevelOrderTraveral(r, n_t)$
-          \State $Tiles =  \{ Tile \}$
-          \For{$(u,v) \in Out(Tile)$}
-              \State $Tiles \leftarrow Tiles \cup TileTree(S_v, n_t)$
-          \EndFor
-          \State \textbf{return} $Tiles$
-      \EndProcedure
-  \end{algorithmic}
-\end{algorithm}
-
-\subsubsection{Further Opimization and Code Generation}
-We found that most leaf tiles for a given tree are at the same depth when uniform tiling is used. Furthermore, we see that deeper leaves 
-are more likely to be reached.
-%\footnote{Intuitively, this is true because training algorithms keep splitting nodes to maximize gain and gain
-%will typically be maximized by splitting a large number of inputs.}.
-Based on these observations, we (optionally) pad the tree of tiles generated with uniform tiling so that all leaves are at the same depth.
-This transformation is performed on the high level IR after uniform tiling. 
-Once the trees have been padded to make all leaves equal depth, the tree walks are fully unrolled to evaluate a fixed 
-number of tiles and all leaf checks are omitted.
-
-One other complication the code generator needs to handle is the fact that different trees in the model being 
-compiled potentially have different depths. In order to handle
-this, Treebeard sorts the trees by their depth. This ensures that all trees with equal depth are grouped together. Once this is done, 
+\subsection{Tree ordering}
+\label{sec:treeorder}
+Specializing the code for each tree in a model comes at a cost.
+First the code generator needs to generate different code for different trees potentially increasing the size of the generated code.
+Second some cross tree optimizations (applied at the lower levels of abstraction) like tree walk interleaving require that the multiple trees share identical code.
+In order to handle
+this, \Treebeard{} pads trees with dummy nodes to make them balanced and then sorts the trees by their depth, so that trees of same depth can share code.
+Padding is only done for almost balanced trees (as generated by basic tiling), this is ensured by only adding up to a fixed fraction of dummy nodes.
+\TODO{Kr : Can we say something better above. If not can we add a threshold say 10\%?. Note : need to mention unrolling of padded sections in MIR.}
+%This ensures that all trees with equal depth are grouped together.
+ Once this is done, 
 the loop over the trees is fissed so that each of the resulting loops only walks trees of a single depth. Consider for example a 
@@
 \end{lstlisting}
-
-\TODO{AP: This listing is unnecesarily long. Can we maybe leave out the loop bodies and say something like "depth 2 walk"? Should 
-we point to the figure in the overview section instead?}

### 2022-04-18 docs/MICRO 2022/overview.tex
Prof. Govind's suggestions, Move macros to main

@@
-\newcommand{\Treebeard}{{\sc {Treebeard}}}
-\newcommand{\op}[1]{{\texttt {#1}}}
-
-\section{Compiler Overview}
-Figure \ref{Fig:CompilerStructure} shows the high level structure of Treebeard. 
-The input to Treebeard is a serialized decision tree ensemble. Popular frameworks 
-like XGBoost and LightGBM are supported and the system is easily extensible to other frameworks.
+\section{Treebeard Overview}
+Figure \ref{Fig:CompilerStructure} shows the high level structure of \Treebeard{}. 
+The input to \Treebeard{} is a serialized decision tree ensemble. 
+%% Popular frameworks like XGBoost and LightGBM are supported and the system is easily extensible to other frameworks.
 Given an input model our compiler generates optimized inference code. Specifically it generates a 
@@
 
-Treebeard specializes the code generated for inference by progressively optimizing and lowering a 
+\Treebeard{} specializes the code generated for inference by progressively optimizing and lowering a 
 high level representation of the \texttt{predictForest} function down to LLVM IR \cite{LLVM}.
-By \textbf{\emph{lowering}}, we mean the process of transforming an operation at a higher 
-abstraction to a sequence of operations at a lower abstraction. Optimizations in Treebeard 
+%% By \textbf{\emph{lowering}}, we mean 
+\textbf{\emph{Lowering}} is the process of transforming an operation at a higher 
+abstraction to a sequence of operations at a lower abstraction. Optimizations in \Treebeard{} 
 are implemented using a combination of annotation and lowering. An operation at a higher 
@@
 to be performed while lowering it. The lowering transformation uses this information to generate 
-optimized lower level code. For example, tree tiling and loop order are decided 
+optimized lower level code. For example, tree tiling and loop nest structure (figure \ref{Fig:Overview} 
+shows two possible loop nest structures -- one tree at a time and one row at a time) are decided 
 at the highest abstraction. These decisions are communicated to the lowering pass 
-which explicitly encodes them in the lowered IR as shown in figure \ref{Fig:Overview}.
+which explicitly encodes them in the lowered IR as shown in Figure \ref{Fig:Overview}.
 
 %Treebeard first constructs a high level representation of the decision forest inference operation.
-Treebeard's IR has three abstractions as shown in Figure \ref{Fig:Overview}.  
+\Treebeard{}'s IR has three abstraction levels as shown in Figure \ref{Fig:Overview}.  
 At the highest level (HIR), the input model is represented as a collection of binary trees. This is shown 
-in the second row of figure \ref{Fig:Overview}. At this level of abstraction,
-Treebeard tiles nodes together to transform a binary tree into an $n$-ary tree 
+in the second row of Figure \ref{Fig:Overview}. At this level of abstraction,
+\Treebeard{} tiles nodes together to transform a binary tree into an $n$-ary tree 
 and decides what order trees are to be traversed in. In the example in the figure, 
@@
 around nodes that are in the same tile. Tree reordering is another transformation that is performed at this abstraction. 
-One objective of tree reordering is to group identically structured trees so that they can share the same traversal code. 
-With the tiling shown in Figure \ref{Fig:Overview}, Tree1 and Tree3 have depth 2, while Tree2 has a depth of 3.
+One objective of tree reordering is to group identically structured trees so that they can share the same traversal code
+to improve locality in the instruction cache.  Alternative objectives for reordering can easily be supported.
+With the tiling shown in Figure \ref{Fig:Overview}, Tree1 and Tree3 have depth 1, while Tree2 has a depth of 2.
 Therefore, the trees are reordered so that Tree1 and Tree3 are together.
@@
 
-Treebeard then further lowers the IR to explicitly represent the memory layout of the model. Buffers to hold model 
+\Treebeard{} then further lowers the IR to explicitly represent the memory layout of the model. Buffers to hold model 
 values are inserted into the generated code and all tree operations in the mid-level IR are lowered to explicitly 
-reference these buffers. Additionally, at this level of the IR, Treebeard generates vectorized code to take 
+reference these buffers. Additionally, at this level of the IR, \Treebeard{} generates vectorized code to take 
 advantage of SIMD instructions where applicable. Finally, the inference function is translated to LLVM IR and 
@@
   \includegraphics[width=\textwidth]{figures/OverviewExample.pdf}
-  \caption{Treebeard IR lowering and optimization details. The three abstraction level in Treebeard's IR are shown. The
+  \caption{\Treebeard{} IR lowering and optimization details. The three abstraction levels in Treebeard's IR are shown. The
            high level IR is a tree based IR to perform model level optimization, the mid-level IR is for

### 2022-04-18 docs/MICRO 2022/treeorder.tex
pass on HIR optimizationswq

@@
 \label{sec:treeorder}	
+%Another objective at the HIR is to reduce the number of trees that need to be specialized by the lower levels.   
+\TODO{kr : Not sure if i got this para right}
 Specializing the code for each tree in a model comes at a cost.
@@
 Second some cross tree optimizations (applied at the lower levels of abstraction) like tree walk interleaving require that the multiple trees share identical code.
+
 In order to handle
@@
 Padding is only done for almost balanced trees (as generated by basic tiling), this is ensured by only adding up to a fixed fraction of dummy nodes.
-\TODO{Kr : Can we say something better above. If not can we add a threshold say 10\%?. Note : need to mention unrolling of padded sections in MIR.}
+\TODO{Kr : Can we add a threshold say 10\%?}
 %This ensures that all trees with equal depth are grouped together.

### 2022-04-18 docs/MICRO 2022/probtiling.tex

@@
 figure \ref{Fig:EpsilonStats} shows that a trees need a much larger fraction of their leaves to cover a significant fraction of the training input.
-This means that most trees in the \op{epsilon} model are not leaf biased.
+This means that most trees in \op{epsilon} are not leaf biased.
 
@@
 }
-We find probability based tiling is only beneficial for leaf biased trees\footnote{Turns out that probability based tiling produces many more tile shapes (see Section~\ref{sec:tileShapes}) which direclty impacts the cost of \op{getChildTile} making it more expensive than basic tiling}.
+We find probability based tiling is only beneficial for leaf biased trees\footnote{Turns out that for trees that are not leaf biased, probability based tiling produces many more tile shapes (see Section~\ref{sec:tileShapes}) which direclty impacts the cost of \op{getChildTile} making it more expensive than basic tiling.}.
 Recall that

### 2022-04-18 docs/MICRO 2022/introduction.tex
Minor intro edits

@@
 gradient boosting machines and random forests were the two most widely used ML algorithms in the company \cite{LookingGlass}.
-Not only are decision tree based models widely used, they are also used in a diverse set of 
+Not only are decision tree based models widely used, they are also used in a diverse range of 
 applications\cite{DecisionTreesOverview}. GBM models were used in the CERN large hadron collider
@@
 % Decision tree ensemble overview
+To compute the prediction of a decision tree on an input row $x$\footnote{An input row is a
+vector of numbers.}, a path from the root to leaf is computed.
 Each node of a decision tree has a threshold value $v$ and a feature index (or column index) $i$.
-To compute the prediction of a decision tree on an input row $x$\footnote{An input row is a
-vector of numbers.}, a path from the root to leaf is computed. At a node, if $x_i < v$, then 
+At a node, if $x_i < v$, then 
 the walk moves to the left child. If not, we move to the right child. When a leaf is reached, 
 the value of the leaf is returned as the prediction of the tree. 
-Most ML techniques combine several decision trees into a forest (or ensemble) to improve prediction accuracy.
+Most decision tree based techniques combine several trees into a forest (or ensemble) to improve prediction accuracy.
 To compute the prediction of the forest, the prediction of each tree is computed and these predictions are 
@@
 % Pointer chasing -- cache performance, branch prediction, dependency stalls, not easy to vectorize
-Given the wide spread use of decision tree ensembles, the ability to perform both high throughput and low latency inference is important.
+Given the wide spread use of decision tree ensembles, the ability to perform high throughput and low latency inference is important.
 \TODO{AP Since we don't talk about latency or throughput explicitly, should we be saying this?}
-However, the tree walks required for inference are not straight-forward to optimize on a modern CPU. 
+However, optimizing the tree walks required for decision tree inference on modern CPUs is not straight-forward. 
 Firstly, naively implemented tree walks have poor spatial and temporal locality and therefore poor cache performance \cite{FAST,MilindTreeVectorization}.
@@
 systems cannot apply specialize inference code depending on the model. A compiler based approach is 
-needed to address these problems. However, existing compilers only support fixed code generation techniques\cite{Treelite}.
+needed to address these problems. However, existing compilers only support fixed code generation techniques\cite{Treelite, Hummingbird}.
 To fill the need for an extensible compiler infrastructure for decision forest ensembles, we design and 
-build Treebeard\footnote{In J. R. R. Tolkein's The Lord of the Rings, Treebeard was the oldest of the Ents left in Middle-earth, an 
+build \Treebeard{}\footnote{In J. R. R. Tolkein's The Lord of the Rings, Treebeard was the oldest of the Ents left in Middle-earth, an 
 ancient tree-like being who was a ``shepherd of trees''.}, an optimizing compiler for decision tree ensemble inference.
-Our motivations for building Treebeard are as follows. 
+Our motivations for building \Treebeard{} are as follows. 
 \begin{enumerate}
-  \item Existing libraries \cite{XGBoost, Treelite, LightGBM, VPred} perform specific optimizations and are hard to maintain/upgrade as hardware evolves. 
-  Additionally, they cannot tailor inference code to the model being used. Past work has identified that the optimizations required change with the model
+  \item Existing libraries \cite{XGBoost, Treelite, LightGBM, VPred} perform specific optimizations and are hard to maintain as hardware evolves. 
+  Additionally, they cannot tailor inference code to the specific model being used. Past work has identified that the optimizations required change with the model
   and architectural parameters like cache size\cite{CacheConscious1, CacheConscious2, VPred}. 
@@
   However, no such compiler infrastructure currently exists for decision tree ensembles.
-  \item Currently, no system exists that allows a thorough exploration of the potential optimization space for decision tree ensemble inference. 
+  \item Currently, no system exists that allows a thorough exploration of the optimization space for decision tree ensemble inference. 
   Compilers have been successful in addressing this problem with ML models like DNNs \cite{TVM, Tiramisu, XLA}. However, compiler techniques for
@@
 \begin{enumerate}
-  \item We design and build Treebeard, an extensible compiler infrastructure for decision tree model inference. The 
+  \item We design and build \Treebeard{}, an extensible compiler infrastructure for decision tree model inference. The 
   infrastructure is built to allow exploration of optimization and code generation techniques. \TODO{Is this claim too grand?}

### 2022-04-18 docs/MICRO 2022/MIR.tex
MIR pass

\subsection{Tree Walk Interleaving}
A key bottleneck we found when we profiled code generated from tiled walks (even after vectorization) was that true dependencies between instructions were still causing
a significant number of processor stalls. Performing a walk with a single input-tree pair was not providing enough independent instructions to keep the processor busy. IN order to address this \Treebeard{} applies an unroll and jam transformation on the inner most loops of the loop nest. This has 
the effect of walking multiple tree and input row pairs in an interleaved fashion. 
This mitigates the dependency stalls by enabling scheduling of instructions from independent tree walks.

This optimization is performed in two steps. 
First a pass on the mid-level IR transforms the loop structure. 
The following listing shows the mid-level IR when the inner loop over the input rows is unrolled 
by a factor of 2 and the two resulting tree walks are jammed together.

\begin{lstlisting}[style=c++]
  forest = ensemble(...)
  for t = 0 to numTrees step 1 {
    for i = 0 to batchSize step 2 {
      tree = getTree(forest, t)
      prediction1, prediction2 = InterleavedWalk((tree, rows[i]), (tree, rows[i+1]))
    }
  }
\end{lstlisting}
Next when lowering, the operations to traverse each of the tree, input row pairs 
(the arguments to the \texttt{InterleavedWalk}) are interleaved. One step of the interleaved 
walk is listed below. 
\begin{lstlisting}[style=c++]
  // ... 
  n1 = n2 = getRoot(tree)
  // ...
  threshold1 = loadThresholds(n1)
  threshold2 = loadThresholds(n2)
  featureIndex1 = loadFeatureIndices(n1)
  featureIndex2 = loadFeatureIndices(n2)
  feature1 = rows[i][featureIndex1]
  feature2 = rows[i][featureIndex2]
  pred1 = feature1 < threshold1
  pred2 = feature2 < threshold2
  n1 = getChildNode(n1, pred1)
  n2 = getChildNode(n2, pred2)
  // ...
\end{lstlisting}

\CommentOut{
These transformations are fairly general and are not aware of the in memory representation of the model. Therefore, they 
are reusable across different in memory representations - the ones that are currently built into Treebeard or ones that 
maybe added in the future.
\TODO{AP I feel the way this section is currently written makes the optimization seem extremely trivial. Is there a different 
way to present it?}
}
\subsection{Loop peeling and walk unrolling}
\Treebeard{} splits the loop that performs a tree walk into two parts. As can be made aware (through a simple pass on the IR) of the depth of the first leaf in a tree, it peels and introduces a \op{prologue} loop that walks down the tree a constant number of steps (up to first leaf) and then performs the rest of the tree walk in a separate loop. Next \Treebeard{} unrolls the prologue completely, avoiding all traversal induced branching in it.

Note that for trees where \Treebeard{} has already padded and balanced the tree (Section~\ref{sec:treeorder}), walk unrolling completely avoids all traversal induced spills.
\TODO{kr: Add example?}

\subsection{Parallelization}
Currently, \Treebeard{} performs a naive parallelization of the inference computation. When parallelism is enabled, the 
loop over the input rows is parallelized using a parallel for construct in MLIR. Treebeard rewrites 
the mid-level IR by tiling the loop over the input rows with a tile size equal to the number of cores. 
As a concrete example, consider the case where we intend to perform inference using a model with 4 trees 
on a batch of 64 rows. Further, assume that we wish to parallelize this computation across 8 cores. 
Treebeard then generates the following IR.
\begin{lstlisting}[style=c++]
  forest = ensemble(...)
  parallel.for i0 = 0 to 64 step 8 {
    for i1 = 0 to 8 step 1 {
      i = i0 + i1
      prediction = 0
      for t = 0 to 4 step 1 {
        tree = getTree(forest, t) 
        treePrediction = WalkTree(tree, rows[i])
        prediction = prediction + treePrediction
      }
      predictions[i] = prediction
    }
  }
\end{lstlisting}
Currently, \Treebeard{} does not perform other standard parallelization optimizations  (such as loop tiling) as they are generic and independent of the problem domain. We leave a more thorough exploration of parallelizing decision trees
to future work.

### 2022-04-18 docs/MICRO 2022/introduction.tex

@@
 several machine learning pipelines that are in production use in a real large scale web company showed that 
-gradient boosting machines and random forests were the two most widely used ML algorithms in the company \cite{LookingGlass}.
+gradient boosting machines and random forests were the two most widely used ML algorithms at the company \cite{LookingGlass}.
 Not only are decision tree based models widely used, they are also used in a diverse range of 
@@
   \item We design and build \Treebeard{}, an extensible compiler infrastructure for decision tree model inference. The 
-  infrastructure is built to allow exploration of optimization and code generation techniques. \TODO{Is this claim too grand?}
+  infrastructure is built to allow exploration of optimization and code generation techniques. % \TODO{Is this claim too grand?}
   \item We develop a general infrastructure for the vectorization of decision tree walks based on grouping tree nodes into ``tiles''.
@@
 \end{enumerate}
+
+We implement \Treebeard{} in the MLIR compiler infrastructure\cite{MLIR}. MLIR provides shared infrastructure
+to develop intermediate representations for domain specific compilers. Firstly, to enable compiler writers to 
+quickly design and build intermediate representations, MLIR has a large set of 
+\textbf{\emph{dialects}} that provide operations from various domains. Each dialect 
+contains a logically related set of operations and types and the transformations to translate these operations to LLVM IR. 
+For example, the \texttt{scf} dialect provides control flow constructs like loops and if-else statements. 
+The \texttt{memref} dialect provides functionality to manage buffers. We build \Treebeard{} using a custom
+MLIR dialect and a combination of dialects provided in MLIR. Secondly, MLIR provides an extremely powerful 
+operator rewriting infrastructure. We make extensive use of this to implement transformations.
+
+Our evaluation of \Treebeard{} on a diverse set of decision tree based models shows that 
+the optimizations we implement improve performance significantly over a naive baseline 
+on both Intel (by 2.2X) and AMD (by 1.8X) processors. We also compare the performance of \Treebeard{}
+with XGBoost and Treelite. Code generated by \Treebeard{} is significantly faster than 
+these state of the art systems. We find that single-core code generated by \Treebeard{} is 
+2.8X and 5.1X faster than XGBoost and Treelite respectively, while geomean speedups for parallel code are
+3.2X and 2.6X respectively.

### 2022-04-18 docs/MICRO 2022/related.tex

\section{Related Work}

\emph{Decision Tree Ensemble Compilers:}
Treelite\cite{Treelite} only supports generation of if-else style code and is not extensible.
Hummingbird\cite{Hummingbird} tries to compile decision tree models to tensor primitives so 
they can be integrated into tensor based frameworks like TensorFlow \cite{TensorFlow}.

\emph{Libraries:} Currently, the most popular systems for decision tree based models 
are libraries. XGBoost\cite{XGBoost} and LightGBM\cite{LightGBM} are the most  
popular gradient boosting libraries while scikit-learn\cite{Sklearn} is 
extremely popular for random forest models. These libraries implement both 
training and inference.

\emph{Other Systems and Techniques}: 
Ren et al \cite{PortableVM} build an intermediate language and VM to 
enable SIMD execution of decision tree inference. The SIMD execution itself is implemented 
by hand in the VM and the VM needs to be reimplemented for every supported target architecture.
Additionally, even though they perform layout optimizations, their system doesn't perform 
any model specific optimizations.

Tahoe, Vpred, CacheConscious1, QuickScorer, QuickScorer1, MilindTreeVectorization
Bliss?

Decision tree training parallelization references

### 2022-04-18 docs/MICRO 2022/tiling.tex

@@
    }
+\input{uniftiling}
+\input{probtiling}
+
+\subsection{A note on implementation}
+\TODO{Kr : Not sure if this is needed}
+The tiling algorithms generate \op{TileId} attributes per tree and this information is used when lowering to the mid level abstraction in the form of loops.
+The \op{TileId} attribute maintains a mapping from a Node to the TileId asigned to it.
+A sample of the lowered MIR code is shown in Figure~\ref{Fig:Overview}.

### 2022-04-18 docs/MICRO 2022/uniftiling.tex

@@
 % we want to minimize the depth of speculation
-Algorithm~\ref{Alg:UnifTilingAlgo} shows a simple tiling algorithm that satisfies all the above constraints. 
+Algorithm~\ref{Alg:UnifTilingAlgo} shows a simple tiling algorithm that satisfies the above constraints and produces a valid tiling. 
 Tiling starts at the root and constructs a tile $Tile$ by performing
@@
 destinations for edges going out of the constructed tile.
-It is easy to see that the set of tiles constructed by algorithm \ref{Alg:UnifTilingAlgo} satisfies the constraints above.
+It is easy to see that the tiled tree constructed by algorithm \ref{Alg:UnifTilingAlgo} is valid.
 \TODO{Pass a modified tree without leafs to the level order traversal call. Explicitly add leaves as seperate tiles}
@@
 One interesting property of this tilling algorithm is that it naturally reduces the imbalance in trees, especially at large tile sizes.
-As the algorithm traverses down to sparser level of the tree, it naturally groups sub-trees containing chains of nodes, thus balancing the trees.
+As the algorithm traverses down to sparser levels of the tree, it naturally groups sub-trees containing chains of nodes, thus balancing the trees.
 While its possible to further enhance the algorithm to explicitly balance tiled trees, we find that basic tiling suffices in practice.

### 2022-04-18 docs/MICRO 2022/overview.tex
Add related work section

@@
 to be performed while lowering it. The lowering transformation uses this information to generate 
-optimized lower level code. For example, tree tiling and loop nest structure (figure \ref{Fig:Overview} 
-shows two possible loop nest structures -- one tree at a time and one row at a time) are decided 
+optimized lower level code. For example, tree tiling and loop ordering (figure \ref{Fig:Overview} 
+shows two possible loop orders -- one tree at a time and one row at a time) are decided 
 at the highest abstraction. These decisions are communicated to the lowering pass 

### 2022-04-18 docs/MICRO 2022/related.tex
Add related work section

@@
 \emph{Decision Tree Ensemble Compilers:}
-Treelite\cite{Treelite} only supports generation of if-else style code and is not extensible.
-Hummingbird\cite{Hummingbird} tries to compile decision tree models to tensor primitives so 
+Treelite\cite{Treelite} is a model compiler for decision tree ensembles and is the system most
+closely related to our work. Even though Treelite compiles ensembles, it only supports 
+generation of if-else style code and is not easily extensible. It cannot perform any of the optimizations 
+\Treebeard{} is designed to perform. 
+Hummingbird\cite{Hummingbird} is a compiler for traditional ML models that can also compile decision trees. 
+It compiles decision tree models to tensor primitives so 
 they can be integrated into tensor based frameworks like TensorFlow \cite{TensorFlow}.
+Hummingbird does not perform any optimizations tailored to decision trees. It 
+also does not perform model specific optimizations.
 
@@
 extremely popular for random forest models. These libraries implement both 
-training and inference.
+training and inference. However, as mentioned in section \ref{sec:Intro}, porting 
+optimizations in these libraries to newer architectures and maintaining them is 
+extremely difficult. Additionally, the inference routines in these libraries have 
+to be general and cannot be specialized to a specific model.
+
+Tahoe\cite{Tahoe} is a library and a performance model that performs high 
+performance tree inference on GPUs. Even though it performs some model 
+specific optimizations, the techniques are GPU specific and cannot easily 
+be ported to architectures like CPUs. Asadi et. al.\cite{VPred} optimizes tree walks
+by hiding dependency stalls by interleaving tree walks. However, their system
+does not do perform any vectorization, tree reordering and specialized code
+generation. They also do not build an easily extensible end-to-end system.
+QuickScorer\cite{QuickScorer, QuickScorer1} is an algorithm that uses 
+bit manipulation to compute tree predictions. Even though QuickScorer is 
+extremely fast for smaller models, it does not scale well to larger 
+models\cite{ProbBasedLayout}. The goal of QuickScorer is orthogonal 
+to the goals of \Treebeard{} and the QuickScorer algorithm can easily  
+be integrated into \Treebeard{} as another traversal strategy for the 
+system to explore. Tang et. al.\cite{CacheConscious1} and Jin et. al.\cite{CacheConscious2}
+build models to predict cache performance of decision tree ensembles on CPUs.
+This work is again orthogonal to the work described in this paper and 
+can easily be incorporated into \Treebeard{}. Some systems have been proposed 
+to parallelize decision tree training on CPUs and GPUs\cite{Jansson2014gpuRFAG, Nasridinov2013DecisionTC}.
 
@@
 Additionally, even though they perform layout optimizations, their system doesn't perform 
-any model specific optimizations.
+any model specific optimizations. Jo et. al.\cite{MilindTreeVectorization} describe techniques 
+to vectorize tree based applications. They do not study optimizations specific to decision trees.
 
-Tahoe, Vpred, CacheConscious1, QuickScorer, QuickScorer1, MilindTreeVectorization
-Bliss?
-
-Decision tree training parallelization references
+\emph{Code Generation Systems from Other Domains:}
+Several compilers and code generation techniques exist for other domains.
+TVM\cite{TVM}, Tiramisu\cite{Tiramisu} and Tensor 
+Comprehensions\cite{TensorComprehensions} are domain specific compilers 
+for deep learning models. Halide\cite{Halide} is a DSL and compiler 
+primarily designed for image processing applications. Several systems
+that generate optimized processing routines have also been designed 
+for BLAS and signal processing. BLIS\cite{BLIS} and ATLAS\cite{atlas_sc98}
+are systems to instantiate high performance BLAS routines on multiple 
+target architectures. CUTLASS\cite{CUTLASS} provides building blocks 
+in the form of C++ templates to quickly instantiate high performance 
+BLAS functions on different GPU architectures. SPIRAL\cite{SPIRAL}
+is a domain specific compilers for signal processing applications
+that instantiates high performance routines for linear transformations 
+like FFTs and FIR filters. FFTW\cite{FFTW} is a fast fourier transform 
+compiler that generates high performance FFT routines by customizing 
+the routine based on FFT size and the target machine.

### 2022-04-18 docs/MICRO 2022/related.tex
Add FAST to related work

@@
 to vectorize tree based applications. They do not study optimizations specific to decision trees.
+Both these works vectorize tree walks by performing different tree walks on each vector lane.
+The main issue with this approach is the divergence of tree walks. Another issue is that 
+memory accesses are gathers rather than vector loads. \Treebeard{}'s approach to 
+vectorization solves both these issues. Also, these approaches are not precluded 
+by \Treebeard{}'s tiling based vectorization. Multiple tiled tree walks can be combined
+into a single vectorized walk. We leave an exploration of this to future work. 
+FAST\cite{FAST} is a system that accelerates tree structured index search 
+on CPUs and GPUs. FAST defines a layout for the index tree that enables vectorization 
+of the tree walk. FAST uses a tree tiling approach to vectorize tree walks. However, 
+FAST only uses a single triangular tile shape. \Treebeard{}'s basic tiling algorithm 
+is a generalization of the tiling used in FAST. If given a perfectly balanced tree,
+the basic tiling algorithm would return exactly the tiling used by FAST. 
+Also, the tree walks in FAST are hand coded using intrinsics on the CPU and CUDA 
+on the GPU and therefore need repeated effort to implement on each target.
 

### 2022-04-18 docs/MICRO 2022/related.tex
related pass

@@
 \section{Related Work}
-
+While decision trees are heavily used there is little prior work on building optimizing compilers for tree based inference. Below we discuss some related work and contrast it to \Treebeard{}.
 \emph{Decision Tree Ensemble Compilers:}
-Treelite\cite{Treelite} is a model compiler for decision tree ensembles and is the system most
-closely related to our work. Even though Treelite compiles ensembles, it only supports 
-generation of if-else style code and is not easily extensible. It cannot perform any of the optimizations 
+Treelite\cite{Treelite} is a model compiler for decision tree ensembles and is the system
+ most closely related to our work. Even though Treelite compiles ensembles, it only supports 
+generation of simple \op{if-else} style code, it does not perform the rich 
+optimizations \Treebeard{} perform and is not easily extensible.
+% It cannot perform any of the optimizations 
 \Treebeard{} is designed to perform. 
-Hummingbird\cite{Hummingbird} is a compiler for traditional ML models that can also compile decision trees. 
-It compiles decision tree models to tensor primitives so 
+Hummingbird\cite{Hummingbird} compiles traditional ML models to make use 
+of tensor primitives so 
 they can be integrated into tensor based frameworks like TensorFlow \cite{TensorFlow}.
@@
 specific optimizations, the techniques are GPU specific and cannot easily 
-be ported to architectures like CPUs. Asadi et. al.\cite{VPred} optimizes tree walks
-by hiding dependency stalls by interleaving tree walks. However, their system
-does not do perform any vectorization, tree reordering and specialized code
-generation. They also do not build an easily extensible end-to-end system.
+be ported to CPUs. Asadi et. al.\cite{VPred} optimizes tree walks
+by hiding dependency stalls by interleaving tree walks.
+\Treebeard{} is an extensible optimizing compiler that carefully implements this and many other optimizations at different levels of abstractions.
 QuickScorer\cite{QuickScorer, QuickScorer1} is an algorithm that uses 
@@
 build models to predict cache performance of decision tree ensembles on CPUs.
-This work is again orthogonal to the work described in this paper and 
-can easily be incorporated into \Treebeard{}. Some systems have been proposed 
+This work is again orthogonal to the work described in this paper. Some systems have been proposed 
 to parallelize decision tree training on CPUs and GPUs\cite{Jansson2014gpuRFAG, Nasridinov2013DecisionTC}.
@@
 by hand in the VM and the VM needs to be reimplemented for every supported target architecture.
-Additionally, even though they perform layout optimizations, their system doesn't perform 
+Additionally, even though they perform layout optimizations, their system does not perform 
 any model specific optimizations. Jo et. al.\cite{MilindTreeVectorization} describe techniques 

### 2022-04-19 docs/MICRO 2022/probtiling.tex
Change symbol for tree

@@
 \[
-    \min_{\mathcal{T} \in \mathcal{C}(T)}{\sum_{l \in L} p_l.depth_{\mathcal{T}}(l)}
+    \min_{\mathcal{T} \in \mathcal{C}(T)}{\sum_{l \in L_{\Tree}} p_l.depth_{\mathcal{T}}(l)}
 \]
-where the minimization is over all valid tilings $\mathcal{T}$ of the tree $T$, $depth_{\mathcal{T}}(l)$ is the depth of the leaf $l$ given tiling ${\mathcal{T}}$.
+where the minimization is over all valid tilings $\mathcal{T}$ of the tree $\Tree$, $depth_{\mathcal{T}}(l)$ is the depth of the leaf $l$ given tiling ${\mathcal{T}}$.
 $p_l$ is the probability of of reaching leaf $l$ as observed during training.
@@
     \begin{algorithmic}
-        \Procedure{TileTree}{$T = (V, E, r)$, $n_t$} 
-            \If {$r \in L$}
+        \Procedure{TileTree}{$\Tree = (V, E, r)$, $n_t$} 
+            \If {$r \in L_{\Tree}$}
                 \State \textbf{return} $\{ r \}$

### 2022-04-19 docs/MICRO 2022/tiling.tex
Change symbol for tree

@@
 \TODO{Notation needs to be introduced in the background section, Tiling should be the first subsection}
-We represent a decision tree by $T = (V, E, r)$ where $V$ is the set of nodes, $E$ the set of edges and
+We represent a decision tree by $\Tree = (V, E, r)$ where $V$ is the set of nodes, $E$ the set of edges and
 $r \in V$ is the root. For each node $n \in V$, we define the following.
@@
 \end{enumerate}
-We use $L \subseteq V$ to denote the set of leaves. % subseteq because tree could have single node
+We use $L_{\Tree} \subseteq V$ to denote the set of leaves. % subseteq because tree could have single node
 
@@
     while (!isLeaf(tree, t)) do {
-      // Evaluate conditions of all nodes in the tile
+      // Evaluate predicates of all nodes in the tile
       predicates = evaluateTilePredicates(t, rows[i])
@@
 \TODO{kr : replace $n_t$ with sz?}
-Given a $tree = (V, E, r)$ and a tile size $n_t$ we impose the following constraints on the generated tiles $\{ T_1, T_2, ... ,T_m \}$ .
+Given a tree $\Tree = (V, E, r)$ and a tile size $n_t$ we impose the following constraints on the generated tiles $\{ T_1, T_2, ... ,T_m \}$ .
 \begin{description}
@@
     \item[Connectedness] If $u, v \in T_i$, there is a (undirected) path connecting $u$ and $v$ fully contained in $T_i$.
-   \item [Leaf seperation] $\forall l \in L$ : $l \in T_i \rightarrow v \notin T_i \;\; \forall v \in V \backslash \{l\}$
-  \item [Maximal tiling] if there are tiles such that. $|T_i| < n_t$ then there is no $v \in V\backslash \{ T_i \cup L \}$ such that $(u, v) \in E$ for some $u \in T_i$. 
+   \item [Leaf seperation] $\forall l \in L_{\Tree}$ : $l \in T_i \rightarrow v \notin T_i \;\; \forall v \in V \backslash \{l\}$
+  \item [Maximal tiling] if there are tiles such that. $|T_i| < n_t$ then there is no $v \in V\backslash \{ T_i \cup L_{\Tree} \}$ such that $(u, v) \in E$ for some $u \in T_i$. 
 \end{description}
@@
 We refer to any tiling that satisties the above constraints as a \emph{valid} tilling.
+
+%%% COMMENT %%%% 
 \CommentOut{
@@
 }
+%%% END COMMENT %%%%
+
 \input{uniftiling}

### 2022-04-19 docs/MICRO 2022/uniftiling.tex
Change symbol for tree
Add level order, uniform -> basic

@@
 \label{sec:UnifTiling}
-% The justification for uniform tiling -- if we assume all paths are equally likely,
+% The justification for basic tiling -- if we assume all paths are equally likely,
 % we want to minimize the depth of speculation
@@
 Tiling starts at the root and constructs a tile $Tile$ by performing
-a level order traversal.
-The call \op{LevelOrderTraversal(r, $n_t$)} picks the next $n_t$ nodes according to the standard level order tree traversal algorithm (not described here) applied only to internal nodes of the tree.
-Once the current tile is constructed, the tiling procedure is recursively performed on all nodes that are
-destinations for edges going out of the constructed tile.
+a level order traversal. The call \op{LevelOrderTraversal(r, $n_t$)} picks the next $n_t$ nodes according to the standard 
+level order tree traversal algorithm (not described here) applied only to internal nodes of the tree. Once the current tile is constructed, 
+the tiling procedure is recursively performed on all nodes that are destinations for edges going out of the constructed tile.
 It is easy to see that the tiled tree constructed by algorithm \ref{Alg:UnifTilingAlgo} is valid.
 \TODO{Pass a modified tree without leafs to the level order traversal call. Explicitly add leaves as seperate tiles}
-% Algorithm for uniform tiling
+% Algorithm for basic tiling
 \begin{algorithm}
-  \caption{Uniform tree tiling}
+  \caption{Basic tree tiling}
   \label{Alg:UnifTilingAlgo}
   \begin{algorithmic}
-      \Procedure{TileTree}{$tree = (V, E, r)$, $n_t$} 
-          \If {$r \in L$}
+      \Procedure{LevelOrderTraversal}{$\Tree = (V, E, r)$, $n_t$}
+        \State $queue = \{ r \}$
+        \State $Tile = \emptyset$
+        \While{$\neg queue.empty() \wedge |Tile|<n_t$}
+          \State $n = queue.pop()$
+          \If {$n \in L_{\Tree}$}
+              \State \textbf{continue}
+          \EndIf
+          \State $Tile = Tile \cup \{ n \}$
+          \State $queue.push(left(n), right(n))$
+        \EndWhile
+      \EndProcedure
+      \State {}
+      \Procedure{TileTree}{$\Tree = (V, E, r)$, $n_t$} 
+          \If {$r \in L_{\Tree}$}
               \State \textbf{return} $\{ r \}$
           \EndIf
-          \State \textcolor{codegreen}{\textit{//Level order traversal to collect $n_t$ or fewer nodes. }}
-          \State \textcolor{codegreen}{\textit{//Leaves are not included in the constructed tile. }}
-          \State $Tile \leftarrow LevelOrderTraveral(r, n_t)$
+          % \State \textcolor{codegreen}{\textit{//Level order traversal to collect $n_t$ or fewer nodes. }}
+          % \State \textcolor{codegreen}{\textit{//Leaves are not included in the constructed tile. }}
+          \State $Tile \leftarrow LevelOrderTraveral(\Tree, n_t)$
           \State $Tiles =  \{ Tile \}$
@@
 While its possible to further enhance the algorithm to explicitly balance tiled trees, we find that basic tiling suffices in practice.
+
+%%%% COMMENT %%%%% 
 \CommentOut{
 \subsubsection{Further Opimization and Code Generation}
-We found that most leaf tiles for a given tree are at the same depth when uniform tiling is used. Furthermore, we see that deeper leaves 
+We found that most leaf tiles for a given tree are at the same depth when basic tiling is used. Furthermore, we see that deeper leaves 
 are more likely to be reached.
@@
 %will typically be maximized by splitting a large number of inputs.}.
-Based on these observations, we (optionally) pad the tree of tiles generated with uniform tiling so that all leaves are at the same depth.
-This transformation is performed on the high level IR after uniform tiling. 
+Based on these observations, we (optionally) pad the tree of tiles generated with basic tiling so that all leaves are at the same depth.
+This transformation is performed on the high level IR after basic tiling. 
 Once the trees have been padded to make all leaves equal depth, the tree walks are fully unrolled to evaluate a fixed 
@@
 
-% loop transformations for uniform tiling (splitting) 
+% loop transformations for basic tiling (splitting) 
 \begin{lstlisting}{style=c++}
@@
 }
+%%%% COMMENT END %%%%%

### 2022-04-19 docs/MICRO 2022/vectorization.tex
Optimizations pass
Add level order, uniform -> basic

@@
 the most probable leaves after probability based tiling which defeats the optimization.
-
+\TODO{kr: connect to Peeling in MIR}
 We address these challenges using a code generation strategy. Treebeard peels 
@@
 One other issue that the code generator needs to handle is that walks of different trees in the same ensemble may 
-need to be peeled to different depths. A strategy similar to what is used for uniform tiling is used to handle this.
+need to be peeled to different depths. A strategy similar to what is used for basic tiling is used to handle this.
 Trees are reordered so that all trees 
@@
 for these groups of trees can be specialized differently.
-% This is very similar to the code generation strategy used for uniform tiling.
+% This is very similar to the code generation strategy used for basic tiling.

### 2022-04-19 docs/MICRO 2022/MIR.tex
Address todos in HIR and MIR

@@
 Performing a walk with a single input-tree pair was not providing enough independent instructions to keep the processor busy.
-IN order to address this \Treebeard{} applies an unroll and jam transformation on the inner most loops of the loop nest.
+In order to address this \Treebeard{} applies an unroll and jam transformation on the inner most loops of the loop nest.
 This has
@@
 }
-\subsection{Loop peeling and walk unrolling}
+\subsection{Loop Peeling and Walk Unrolling}
 \Treebeard{} splits the loop that performs a tree walk into two parts.
-As can be made aware (through a simple pass on the IR) of the depth of the first leaf in a tree, it peels and introduces a \op{prologue} loop that walks down the tree a constant number of steps (up to first leaf) and then performs the rest of the tree walk in a separate loop.
-Next \Treebeard{} unrolls the prologue completely, avoiding all traversal induced branching in it.
+As can be made aware (through a simple pass on the IR) of the depth of the first leaf in a tree, it peels and introduces a \op{prologue} loop that walks down the tree a constant number of steps (for example, up to the depth of the first leaf) and then performs the rest of the tree walk in a separate loop.
+% Next \Treebeard{} unrolls the prologue completely, avoiding all traversal induced branching in it.
 
-Note that for trees where \Treebeard{} has already padded and balanced the tree (Section~\ref{sec:treeorder}), walk unrolling completely avoids all traversal induced spills.
-\TODO{kr: Add example?}
+Several rewrites of the peeled loop are possible. \Treebeard{} completely unrolls the prologue if the peeled loop 
+walks the tree upto the depth of the first leaf. In cases where \Treebeard{} has already padded and balanced the 
+tree (Section~\ref{sec:treeorder}), walk unrolling completely avoids all traversal induced spills. In the case of 
+probability based tiling, the prologue loop is specialized to optimize leaf checks so that these checks 
+are faster (and less general) for the most likely nodes.
+
+% \TODO{kr: Add example?}
 
 \subsection{Parallelization}
-Currently, \Treebeard{} performs a naive parallelization of the inference computation. When parallelism is enabled, the 
-loop over the input rows is parallelized using a parallel for construct in MLIR. Treebeard rewrites 
+Currently, \Treebeard{} performs a na\"ive parallelization of the inference computation. When parallelism is enabled, the 
+loop over the input rows is parallelized using a \op{parallel.for} construct in MLIR. Treebeard rewrites 
 the mid-level IR by tiling the loop over the input rows with a tile size equal to the number of cores. 
@@
 on a batch of 64 rows. Further, assume that we wish to parallelize this computation across 8 cores. 
-Treebeard then generates the following IR.
+\Treebeard{} then generates the following IR.
 \begin{lstlisting}[style=c++]
@@
         tree = getTree(forest, t) 
-        treePrediction = WalkTree(tree, rows[i])
+        treePrediction = WalkDecisionTree(tree, rows[i])
         prediction = prediction + treePrediction
@@
 \end{lstlisting}
-Currently, \Treebeard{} does not perform other standard parallelization optimizations  (such as loop tiling) as they are generic and independent of the problem domain.
-We leave a more thorough exploration of parallelizing decision trees
+Currently, \Treebeard{} does not perform other standard parallelization optimizations  (such as loop tiling) as they
+are generic and independent of the problem domain. We leave a more thorough exploration of parallelizing decision trees
 to future work.

### 2022-04-19 docs/MICRO 2022/tiling.tex
Address todos in HIR and MIR

@@
 
-\subsection{A note on implementation}
+\subsection{A Note on Implementation}
 \TODO{Kr : Not sure if this is needed}
-The tiling algorithms generate \op{TileId} attributes per tree and this information is used when lowering to the mid level abstraction in the form of loops.
-The \op{TileId} attribute maintains a mapping from a Node to the TileId asigned to it.
-A sample of the lowered MIR code is shown in Figure~\ref{Fig:Overview}.
+The tiling algorithms generate a \op{TileId} attribute per tree. The \op{TileId} attribute contains a mapping from a Node to the TileId asigned to it.
+This information is used when lowering to the mid level abstraction in the form of loops. A sample of the lowered MIR code is shown in Figure~\ref{Fig:Overview}.

### 2022-04-19 docs/MICRO 2022/uniftiling.tex
Address todos in HIR and MIR

@@
 Tiling starts at the root and constructs a tile $Tile$ by performing
-a level order traversal. The call \op{LevelOrderTraversal(r, $n_t$)} picks the next $n_t$ nodes according to the standard 
-level order tree traversal algorithm (not described here) applied only to internal nodes of the tree. Once the current tile is constructed, 
-the tiling procedure is recursively performed on all nodes that are destinations for edges going out of the constructed tile.
-It is easy to see that the tiled tree constructed by algorithm \ref{Alg:UnifTilingAlgo} is valid.
-\TODO{Pass a modified tree without leafs to the level order traversal call. Explicitly add leaves as seperate tiles}
+a level order traversal. The call \op{LevelOrderTraversal($\Tree$, $n_t$)} picks the next $n_t$ non-leaf nodes according to the standard 
+level order tree traversal algorithm. Once the current tile is constructed, 
+the tiling procedure is recursively performed on subtrees rooted at each node that is a destination of an edge going out of the constructed 
+tile. These edges are denoted by $Out(Tile)$ and the subtree rooted at a node $v$ is denoted by $S_v$.
+It is easy to see that the tiling constructed by algorithm \ref{Alg:UnifTilingAlgo} is valid.
+% \TODO{Pass a modified tree without leafs to the level order traversal call. Explicitly add leaves as seperate tiles}
 % Algorithm for basic tiling

### 2022-04-19 docs/MICRO 2022/main.tex
Outline for related work
Prof. Govind's suggestions, Move macros to main
Add related work section
Change symbol for tree
Tiling changes

@@
 \usepackage{subfig}
-% Always include hyperref last
-\usepackage[bookmarks=true,breaklinks=true,letterpaper=true,colorlinks,citecolor=blue,linkcolor=blue,urlcolor=blue]{hyperref}
 \usepackage{algorithm}
@@
 \usepackage{algpseudocode}
+% Always include hyperref last
+\usepackage[bookmarks=true,breaklinks=true,letterpaper=true,colorlinks,citecolor=blue,linkcolor=blue,urlcolor=blue]{hyperref}
 
@@
 \newcommand\TODO[1]{\textcolor{red}{TODO #1}}
+\newcommand\COMMENT[1]{\textcolor{green}{COMMENT #1}}
+\newcommand{\Treebeard}{{\sc {Treebeard}}}
+\newcommand{\op}[1]{{\texttt {#1}}}
+\DeclareMathOperator{\Tree}{\ensuremath{\tau}}
 
@@
 \lstdefinestyle{c++}{language=C++,
-    morekeywords={ensemble, getTree, getRoot, isLeaf, isLeafTile, traverseTreeTile, getLeafValue, loadThresholds, loadFeatureIndices, loadTileShape, getChildNode},    
+    morekeywords={ensemble, getTree, getRoot, isLeaf, isLeafTile, traverseTreeTile, getLeafValue, loadThresholds,
+                  loadFeatureIndices, loadTileShape, getChildNode, parallel, InterleavedWalk, evaluateTilePredicates,
+                  getChildTile, WalkDecisionTree},    
     backgroundcolor=\color{backcolour},   
     commentstyle=\color{codegreen},
-    keywordstyle=\color{magenta},
+    keywordstyle=\color{blue},
     numberstyle=\tiny\color{codegray},
@@
 %%%%%%%%%%%---SETME-----%%%%%%%%%%%%%
-\newcommand{\microsubmissionnumber}{XXX}
+\newcommand{\microsubmissionnumber}{654}
 %%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%
@@
 \input{details}
+\input{related}
 %%%%%%% -- PAPER CONTENT ENDS -- %%%%%%%%

### 2022-04-19 docs/MICRO 2022/tiling.tex
Tiling changes

@@
 \subsection{Notation}
-\TODO{Notation needs to be introduced in the background section, Tiling should be the first subsection}
+%TODO{Notation needs to be introduced in the background section, Tiling should be the first subsection}
 We represent a decision tree by $\Tree = (V, E, r)$ where $V$ is the set of nodes, $E$ the set of edges and
@@
 }
-\subsection*{Conditions for Efficient Tiling}
+\subsubsection*{Conditions for Efficient Tiling}
 \label{sec:ValidTiling}
@@
 % to only allow nodes that are likely to be accessed together to be grouped into a tile. 
-\TODO{kr : replace $n_t$ with sz?}
+% \TODO{kr : replace $n_t$ with sz?}
 Given a tree $\Tree = (V, E, r)$ and a tile size $n_t$ we impose the following constraints on the generated tiles $\{ T_1, T_2, ... ,T_m \}$ .
@@
 This constraint ensures that tiles are homogenous, this in-turn allows us to specialize the in-memory layout of trees and also simplifies code generation.
-We discuss leaf handling and tree layout in Section~\ref{sec:}.
-We refer to any tiling that satisties the above constraints as a \emph{valid} tilling.
+We discuss leaf handling and tree layout in section ~\ref{Sec:MemoryRep}.
+We refer to any tiling that satisfies the above constraints as a \emph{valid} tilling.
 

### 2022-04-19 docs/MICRO 2022/MIR.tex

@@
       tree = getTree(forest, t)
-      prediction1, prediction2 = InterleavedWalk((tree, rows[i]), (tree, rows[i+1]))
+      pred1, pred2 = InterleavedWalk((tree, rows[i]), (tree, rows[i+1]))
     }
@@
 walks the tree upto the depth of the first leaf. In cases where \Treebeard{} has already padded and balanced the 
-tree (Section~\ref{sec:treeorder}), walk unrolling completely avoids all traversal induced spills. In the case of 
-probability based tiling, the prologue loop is specialized to optimize leaf checks so that these checks 
-are faster (and less general) for the most likely nodes.
+tree (Section~\ref{sec:treeorder}), walk unrolling completely avoids all traversal induced branching. In the case of 
+probability based tiling, the prologue loop enables specialization of leaf checks so that these checks 
+are faster (and less general) for the most probable leaves.
 

### 2022-04-19 docs/MICRO 2022/related.tex

@@
 \section{Related Work}
-While decision trees are heavily used there is little prior work on building optimizing compilers for tree based inference. Below we discuss some related work and contrast it to \Treebeard{}.
+While decision trees are heavily used there is little prior work on building optimizing compilers for tree 
+based inference. Below we discuss some related work and contrast it to \Treebeard{}.
+
 \emph{Decision Tree Ensemble Compilers:}
@@
 most closely related to our work. Even though Treelite compiles ensembles, it only supports 
-generation of simple \op{if-else} style code, it does not perform the rich 
-optimizations \Treebeard{} perform and is not easily extensible.
-% It cannot perform any of the optimizations 
-\Treebeard{} is designed to perform. 
+generation of simple \op{if-else} style code. It does not perform the rich 
+optimizations \Treebeard{} performs and is not easily extensible.
+% It cannot perform any of the optimizations \Treebeard{} is designed to perform. 
 Hummingbird\cite{Hummingbird} compiles traditional ML models to make use 
@@
 specific optimizations, the techniques are GPU specific and cannot easily 
-be ported to CPUs. Asadi et. al.\cite{VPred} optimizes tree walks
-by hiding dependency stalls by interleaving tree walks.
-\Treebeard{} is an extensible optimizing compiler that carefully implements this and many other optimizations at different levels of abstractions.
+be ported to CPUs. Asadi et. al.\cite{VPred} optimize tree walks
+by hiding dependency stalls by interleaving tree walks. In contrast, \Treebeard{} is an extensible 
+optimizing compiler that carefully implements this and many other optimizations at different levels of abstractions.
 QuickScorer\cite{QuickScorer, QuickScorer1} is an algorithm that uses 

### 2022-04-19 docs/MICRO 2022/treeorder.tex
Add level order, uniform -> basic

@@
 %Another objective at the HIR is to reduce the number of trees that need to be specialized by the lower levels.   
-\TODO{kr : Not sure if i got this para right}
-Specializing the code for each tree in a model comes at a cost.
-First the code generator needs to generate different code for different trees potentially increasing the size of the generated code.
-Second some cross tree optimizations (applied at the lower levels of abstraction) like tree walk interleaving require that the multiple trees share identical code.
+% \TODO{kr : Not sure if i got this para right}
+Specializing the code for each tree in a model comes at a cost. First, the size of the generated code increases if the
+code generator needs to generate different code for different trees. This causes several front end stalls like instruction 
+cache misses and delays in instruction decoding. Second, some cross tree optimizations 
+like tree walk interleaving (applied at the lower levels of abstraction) are more effective when multiple trees share identical code.
 
-In order to handle
-this, \Treebeard{} pads trees with dummy nodes to make them balanced and then sorts the trees by their depth, so that trees of same depth can share code.
-Padding is only done for almost balanced trees (as generated by basic tiling), this is ensured by only adding up to a fixed fraction of dummy nodes.
-\TODO{Kr : Can we add a threshold say 10\%?}
+In order to handle this, \Treebeard{} groups trees by their structure so that trees with the same structure can 
+share code. For example, the compiler pads trees with dummy tiles to make them balanced and then sorts the trees by 
+their depth, so that trees of same depth can share the same unrolled tree walk. Padding is only done for almost balanced trees (as generated 
+by basic tiling). This is shown in the example in figure \ref{Fig:Overview}. Other metrics, like feature set commonality, 
+could also be used to reorder trees\cite{Tahoe}. However, we leave the exploration of such optimizations to future work.
+
+% , this is ensured by only adding up to a fixed fraction of dummy nodes. \TODO{Kr : Can we add a threshold say 10\%?}
 %This ensures that all trees with equal depth are grouped together.
- Once this is done, 
-the loop over the trees is fissed so that each of the resulting loops only walks trees of a single depth. Consider for example a 
+\CommentOut{ 
+Once this is done, the loop over the trees is fissed so that each of the resulting loops only walks trees of a single depth. Consider for example a 
 forest with 4 trees $T_1$, $T_2$, $T_3$, and $T_4$ in that order. Further, assume that $T_1$ and $T_4$ have depth 2 while $T_2$ and $T_3$
@@
 \end{lstlisting}
+}

### 2022-04-19 docs/MICRO 2022/vectorization.tex
Address todos in HIR and MIR

@@
 Informally, the \textbf{\emph{tile shape}} is the shape of the region that encloses all nodes in a tile in a diagram of the decision tree.
-More formally, for a tile size $n_t$, each unique legal binary tree containing $n_t$ nodes (nodes being indistinguishable) corresponds to a tile shape.
+More formally, for a tile size $n_t$, each unique legal binary tree containing $n_t$ nodes (nodes being indistinguishable) corresponds to
+a tile shape. Figure \ref{Fig:TileSize3Shapes} enumerates all tile shapes with a tile size of 3.
 
-Figure \ref{Fig:TileSize3Shapes} enumerates all tile shapes with a tile size of 3.
-There are a total of 5 tile shapes with size 3.
-The number of tile shapes with a tile size $n_t$, denoted by $NTS(n_t)$ is given by the following equation.
+% There are a total of 5 tile shapes with size 3. The number of tile shapes with a tile size $n_t$, denoted by $NTS(n_t)$ is given by the following equation.
 
-\begin{equation}
-  NTS(n) = \sum_{k=0}^{n-1} NTS(k) \times NTS(n-k-1)
-\end{equation}
+% \begin{equation}
+%   NTS(n) = \sum_{k=0}^{n-1} NTS(k) \times NTS(n-k-1)
+% \end{equation}
 
-where $NTS(0) = NTS(1) = 1$.
+% where $NTS(0) = NTS(1) = 1$.
 
@@
 Treebeard computes the values in the LUT statically as the tile size is a compile time constant.
-\TODO{AP What comes after subsubsection?}
 
 \subsection{In Memory Representation of Tiled Trees}
+\label{Sec:MemoryRep}
 Treebeard currently has two in memory representations for tiled trees - an array based representation and a sparse representation.
@@
 
+\CommentOut{
 \subsubsection{Code Generation for Probability Based Tiling}
@@
 % This is very similar to the code generation strategy used for basic tiling.
+}

### 2022-04-25 docs/MICRO 2022/results/result_parser.py
Benchmark (#36)
* Add results for multiple batch sizes run

* Changes to benchmark code

* Add xgboost and treelite compare results

* Add AMD results

* Add parallel performance results

* code to generate graphs + modified result numbers

* Fix bug in plotting code

* Add grid to line graphs

* Add grid lines

* Add results for varying number of cores

* Run parallel tests

* Add AMD parallel results

* Update graph generator

* Add scaling results for AMD

* Fix images, alter font sizes and axes

* Just use Probability based tiling

* Update result_parser.py

* Update script to use Intel pr

@@ -1,0 +1,414 @@
+from stat import FILE_ATTRIBUTE_NORMAL
+import pandas
+from scipy.stats.mstats import gmean
+import matplotlib.pyplot as plt
+import numpy as np
+from typing import List
+import math
+import os
+
+current_directory = os.path.dirname(os.path.realpath(__file__))
+
+class SingleBatchSizeResult:
+  def __init__(self, batch_size, min_scalar_times, unif_tiling_times) -> None:
+    self.batch_size = batch_size
+    self.scalar_times = min_scalar_times
+    self.unif_tiling_times = unif_tiling_times
+    self.unif_tiling_speedups = [i / j for i, j in zip(min_scalar_times, unif_tiling_times)]
+    self.unif_tiling_mean_speedup = gmean(self.unif_tiling_speedups)
+    self.prob_tiling_times = []
+    self.prob_tiling_speedups = []
+    self.prob_tiling_mean_speedup = -1
+    self.pipelining_times = []
+    self.pipelining_speedups_over_unif = []
+    self.pipelining_mean_speedup = -1
+    self.overall_speedups = []
+    self.overall_mean_speedup = -1
+  
+  def set_prob_tiling_times(self, times):
+    self.prob_tiling_times = times
+    self.prob_tiling_speedups = [i / j for i, j in zip(self.unif_tiling_times, times)]
+    self.prob_tiling_speedups[4] = 1
+    self.prob_tiling_mean_speedup = gmean(self.prob_tiling_speedups)
+  
+  def set_pipelining_times(self, times):
+    self.pipelining_times = times
+    self.pipelining_speedups_over_unif = [i / j for i, j in zip(self.unif_tiling_times, times)]
+    self.pipelining_mean_speedup = gmean(self.pipelining_speedups_over_unif)
+    self.best_times = [min(i, j, k) for i, j, k in zip(self.unif_tiling_times, self.prob_tiling_times, self.pipelining_times)]
+    self.overall_speedups = [i / j for i, j in zip(self.scalar_times, self.best_times)]
+    self.overall_mean_speedup = gmean(self.overall_speedups)
+
+class OptimizationSpeedups:
+  def __init__(self, uniform_tiling_speedups, prob_tiling_speedups, pipelining_speedups) -> None:
+    self.uniform_tiling_speedups = uniform_tiling_speedups
[…]

### 2022-04-25 src/python/treebeard.py

@@ -83,8 +83,8 @@ class TreebeardAPI:
       self.runtime_lib.Set_tilingType.argtypes = [ctypes.c_int64, ctypes.c_int32]
       self.runtime_lib.Set_tilingType.restype = None
 
-      self.runtime_lib.Set_pipelineWidth.argtypes = [ctypes.c_int64, ctypes.c_int32]
-      self.runtime_lib.Set_pipelineWidth.restype = None
+      self.runtime_lib.Set_pipelineSize.argtypes = [ctypes.c_int64, ctypes.c_int32]
+      self.runtime_lib.Set_pipelineSize.restype = None
 
       self.runtime_lib.Set_numberOfCores.argtypes = [ctypes.c_int64, ctypes.c_int32]
       self.runtime_lib.Set_numberOfCores.restype = None
@@ -179,7 +179,7 @@ class CompilerOptions:
     treebeardAPI.runtime_lib.Set_tilingType(self.optionsPtr, val)
   
   def SetPipelineWidth(self, val : int) :
-    treebeardAPI.runtime_lib.Set_pipelineWidth(self.optionsPtr, val)
+    treebeardAPI.runtime_lib.Set_pipelineSize(self.optionsPtr, val)
 
   def SetNumberOfCores(self, val : int) :
     treebeardAPI.runtime_lib.Set_numberOfCores(self.optionsPtr, val)

### 2022-04-25 src/test/TestUtilsCommon.cpp

@@ -37,6 +37,7 @@ TestCSVReader::TestCSVReader(const std::string& filename, int32_t numRows) {
   while (!fin.eof() && (numRows == -1 || numRowsRead<numRows)) {
     auto row = getNextLineAndSplitIntoTokens(fin);
     m_data.push_back(row);
+    ++numRowsRead;
   }
 }
 

### 2022-04-25 src/test/XGBoostBenchmarks.cpp

@@ -43,21 +43,21 @@ std::string GetTypeName(float f) {
 // #define PROFILE_MODE
 
 #ifndef PROFILE_MODE
-constexpr int32_t NUM_RUNS = 100;
+constexpr int32_t NUM_RUNS = 500;
 #else
 constexpr int32_t NUM_RUNS = 1000;
 #endif
 
-int32_t GetNumberOfCores() { 
-  return 8;
-}
+// int32_t GetNumberOfCores() { 
+//   return 16;
+// }
 
 template<typename FloatType, typename ReturnType=FloatType>
-int64_t Test_CodeGenForJSON_ProbabilityBasedTiling(int64_t batchSize, const std::string& modelJsonPath, 
+double Test_CodeGenForJSON_ProbabilityBasedTiling(int64_t batchSize, const std::string& modelJsonPath, 
                                               const std::string& statsProfileCSV,
                                               int32_t tileSize, int32_t tileShapeBitWidth, 
                                               int32_t childIndexBitWidth, mlir::decisionforest::ScheduleManipulator *scheduleManipulator, 
-                                              bool probTiling, bool parallel,
+                                              bool probTiling, int32_t numberOfCores,
                                               int32_t pipelineSize) {
   // TODO consider changing this so that you use the smallest possible type possible (need to make it a parameter)
   using FeatureIndexType = int16_t;
@@ -65,7 +65,7 @@ int64_t Test_CodeGenForJSON_ProbabilityBasedTiling(int64_t batchSize, const std:
 
   mlir::MLIRContext context;
   int32_t floatTypeBitWidth = sizeof(FloatType)*8;
-  bool reorderTrees = probTiling || parallel || pipelineSize > 1;
+  bool reorderTrees = probTiling || (numberOfCores!=-1) || pipelineSize > 1;
   TreeBeard::CompilerOptions options(floatTypeBitWidth, sizeof(ReturnType)*8, IsFloatType(ReturnType()), sizeof(FeatureIndexType)*8, sizeof(NodeIndexType)*8,
                                      floatTypeBitWidth, batchSize, tileSize, tileShapeBitWidth, childIndexBitWidth,
                                      (probTiling ? TreeBeard::TilingType::kHybrid : TreeBeard::TilingType::kUniform), 
[…]

### 2022-04-25 test/python/run_python_benchmarks.py

@@ -2,13 +2,16 @@ import os
 import sys
 
 # This is a hack to make XGBoost run single threaded. 
-os.environ['OMP_NUM_THREADS'] = "1"
+# os.environ['OMP_NUM_THREADS'] = "1"
+os.environ['TREELITE_BIND_THREADS']="0"
 
 import numpy
 import pandas
 import time
 import math
 from scipy.stats.mstats import gmean
+import treelite
+import treelite_runtime
 
 filepath = os.path.abspath(__file__)
 treebeard_repo_dir = os.path.dirname(os.path.dirname(os.path.dirname(filepath)))
@@ -22,39 +25,46 @@ import treebeard
 import xgboost as xgb
 
 num_repeats = 100
-use_inplace_predict = False
-run_on_same_array_repeatedly = False
+use_inplace_predict = True
+run_on_same_array_repeatedly = True
+run_parallel = False
 
-def RunSingleTest_Multibatch(soPath, globalsJSONPath, csvPath, returnType) -> float:
-  data_df = pandas.read_csv(csvPath, header=None)
-  data = numpy.array(data_df, order='C') # numpy.genfromtxt(csvPath, ',')
+# def RunSingleTest_Multibatch(soPath, globalsJSONPath, csvPath, returnType) -> float:
+#   data_df = pandas.read_csv(csvPath, header=None)
+#   data = numpy.array(data_df, order='C') # numpy.genfromtxt(csvPath, ',')
 
-  inferenceRunner = treebeard.TreebeardInferenceRunner(soPath, globalsJSONPath)
+#   inferenceRunner = treebeard.TreebeardInferenceRunner(soPath, globalsJSONPath)
+#   batch_size = inferenceRunner.batchSize
+#   num_batches = int(math.floor(data.shape[0]/inferenceRunner.batchSize))
+#   numRows = int(math.floor(data.shape[0]/inferenceRunner.batchSize) * inferenceRunner.batchSize)
+#   inputs = numpy.array(data[0:numRows, :-1], numpy.float32, order='C')
+#   repeated_inputs = numpy.tile(inputs, (num_repeats, 1))
 
-  numRows = int(math.floor(data.shape[0]/inferenceRunner.batchSize) * inferenceRunner.batchSize)
-  inputs = numpy.array(data[0:numRows, :-1], numpy.float32, order='C')
-  repeated_inputs = numpy.tile(inputs, (num_repeats, 1))
+#   start = time.time()
+#   if run_on_same_array_repeatedly:
+#     for i in range(num_repeats):
[…]

### 2022-04-25 test/python/run_python_tests.py

@@ -125,6 +125,30 @@ def SetStatsProfileCSVPath(options, modelName):
   options.SetStatsProfileCSVPath(csvPath)
   return options
 
+def RunPipeliningTests():
+  invertLoopsTileSize8Options = treebeard.CompilerOptions(200, 8)
+  # invertLoopsTileSize8Options.SetOneTreeAtATimeSchedule()
+  invertLoopsTileSize8Options.SetPipelineWidth(8)
+  invertLoopsTileSize8Options.SetReorderTreesByDepth(True)
+  invertLoopsTileSize8Options.SetMakeAllLeavesSameDepth(1)
+  
+  invertLoopsTileSize8MulticlassOptions = treebeard.CompilerOptions(200, 8)
+  invertLoopsTileSize8MulticlassOptions.SetReturnTypeWidth(8)
+  invertLoopsTileSize8MulticlassOptions.SetReturnTypeIsFloatType(False)
+  # invertLoopsTileSize8MulticlassOptions.SetOneTreeAtATimeSchedule();
+  invertLoopsTileSize8MulticlassOptions.SetPipelineWidth(8)
+  invertLoopsTileSize8MulticlassOptions.SetReorderTreesByDepth(True)
+  invertLoopsTileSize8MulticlassOptions.SetMakeAllLeavesSameDepth(1)
+
+  assert RunTestOnSingleModelTestInputsJIT("abalone", invertLoopsTileSize8Options)
+  assert RunTestOnSingleModelTestInputsJIT("airline", invertLoopsTileSize8Options)
+  assert RunTestOnSingleModelTestInputsJIT("airline-ohe", invertLoopsTileSize8Options)
+  assert RunTestOnSingleModelTestInputsJIT("covtype", invertLoopsTileSize8MulticlassOptions, numpy.int8)
+  assert RunTestOnSingleModelTestInputsJIT("epsilon", invertLoopsTileSize8Options)
+  assert RunTestOnSingleModelTestInputsJIT("higgs", invertLoopsTileSize8Options)
+  assert RunTestOnSingleModelTestInputsJIT("letters", invertLoopsTileSize8MulticlassOptions, numpy.int8)
+  assert RunTestOnSingleModelTestInputsJIT("year_prediction_msd", invertLoopsTileSize8Options)
+
 defaultTileSize8Options = treebeard.CompilerOptions(200, 8)
 defaultTileSize8MulticlassOptions = treebeard.CompilerOptions(200, 8)
[…]

### 2022-04-25 docs/MICRO 2022/MIR.tex
Micro paper (#37)
* Add results for multiple batch sizes run

* Changes to benchmark code

* Add xgboost and treelite compare results

* Add AMD results

* Add parallel performance results

* code to generate graphs + modified result numbers

* Fix bug in plotting code

* Add grid to line graphs

* Add grid lines

* Add results for varying number of cores

* Run parallel tests

* Add AMD parallel results

* Update graph generator

* Add scaling results for AMD

* Fix images, alter font sizes and axes

* Just use Probability based tiling

* Update images - Change axis font size, fix wrong graph

@@
+\section{Optimizations on mid-level IR}
+\label{Sec:MIR}
+In this section, we present various tree walk optimizations performed by
+\Treebeard{} on the mid-level IR.
+
 \subsection{Tree Walk Interleaving}
-A key bottleneck we found when we profiled code generated from tiled walks (even after vectorization) was that true dependencies between instructions were still causing
-a significant number of processor stalls.
-Performing a walk with a single input-tree pair was not providing enough independent instructions to keep the processor busy.
-In order to address this \Treebeard{} applies an unroll and jam transformation on the inner most loops of the loop nest.
-This has
-the effect of walking multiple tree and input row pairs in an interleaved fashion. 
-This mitigates the dependency stalls by enabling scheduling of instructions from independent tree walks.
+A key bottleneck we found when we profiled code generated from tiled walks (even 
+after vectorization) was that true dependencies between instructions were still 
+causing a significant number of processor stalls. Performing a walk with a 
+single input-tree pair did not provide enough independent instructions to keep 
+the processor busy. In order to address this, \Treebeard{} applies an 
+unroll-and-jam transformation on the innermost loops of the loop nest. This has 
+the effect of walking multiple tree and input row pairs in an interleaved 
+fashion.  This mitigates the dependency stalls by enabling scheduling of 
+instructions from independent tree walks.
 
-This optimization is performed in two steps. 
-First a pass on the mid-level IR transforms the loop structure. 
-The following listing shows the mid-level IR when the inner loop over the input rows is unrolled 
-by a factor of 2 and the two resulting tree walks are jammed together.
+This optimization is performed in two steps. First, a pass on the mid-level IR 
+transforms the loop structure.  It unrolls the innermost loops of the loop nest 
+a specified number of times and jams together tree walks from the different 
+iterations.  The following listing shows the mid-level IR when the inner loop 
+over the input rows is unrolled by a factor of two and the two resulting tree 
+walks are jammed together.
 
 \begin{lstlisting}[style=c++]
-  forest = ensemble(...)
   for t = 0 to numTrees step 1 {
@@
       tree = getTree(forest, t)
-      pred1, pred2 = InterleavedWalk((tree, rows[i]), (tree, rows[i+1]))
+      pred1, pred2 = InterleavedWalk((tree, rows[i]),
+                                     (tree, rows[i+1]))
     }
@@
   // ... 
-  n1 = n2 = getRoot(tree)
+  tile1 = tile2 = getRoot(tree)
   // ...
-  threshold1 = loadThresholds(n1)
-  threshold2 = loadThresholds(n2)
-  featureIndex1 = loadFeatureIndices(n1)
-  featureIndex2 = loadFeatureIndices(n2)
+  threshold1 = loadThresholds(tree, tile1)
+  threshold2 = loadThresholds(tree, tile2)
+  featureIndex1 = loadFeatureIndices(tree, tile1)
+  featureIndex2 = loadFeatureIndices(tree, tile2)
   feature1 = rows[i][featureIndex1]
@@
   pred2 = feature2 < threshold2
-  n1 = getChildNode(n1, pred1)
-  n2 = getChildNode(n2, pred2)
+  tile1 = getChildTile(tile1, pred1)
+  tile2 = getChildTile(tile2, pred2)
   // ...
@@
 \CommentOut{
-These transformations are fairly general and are not aware of the in memory representation of the model. Therefore, they 
-are reusable across different in memory representations - the ones that are currently built into Treebeard or ones that 
+These transformations are fairly general and are not aware of the in-memory representation of the model. Therefore, they 
+are reusable across different in-memory representations - the ones that are currently built into Treebeard or ones that 
 maybe added in the future.
@@
 }
-\subsection{Loop Peeling and Walk Unrolling}
-\Treebeard{} splits the loop that performs a tree walk into two parts.
-As can be made aware (through a simple pass on the IR) of the depth of the first leaf in a tree, it peels and introduces a \op{prologue} loop that walks down the tree a constant number of steps (for example, up to the depth of the first leaf) and then performs the rest of the tree walk in a separate loop.
+\subsection{Tree Walk Peeling and Tree Walk Unrolling}
+\Treebeard{} splits the loop that performs a tree walk into two parts. It peels 
+and introduces a \op{prologue} loop that walks down the tree a constant number 
+of steps (for example, up to the depth of the first leaf) and then performs the 
+rest of the tree walk in a separate loop.
+
+%% As can be made aware (through a simple pass on the IR) of the depth of the first leaf in a tree, 
+%First it identifies the leaf with the minimum depth through a simple pass.  Then
+
 % Next \Treebeard{} unrolls the prologue completely, avoiding all traversal induced branching in it.
 
-Several rewrites of the peeled loop are possible. \Treebeard{} completely unrolls the prologue if the peeled loop 
-walks the tree upto the depth of the first leaf. In cases where \Treebeard{} has already padded and balanced the 
-tree (Section~\ref{sec:treeorder}), walk unrolling completely avoids all traversal induced branching. In the case of 
-probability based tiling, the prologue loop enables specialization of leaf checks so that these checks 
-are faster (and less general) for the most probable leaves.
+Several rewrites of the peeled loop are possible. \Treebeard{} completely 
+unrolls the prologue if the peeled loop walks the tree upto the depth of the 
+first leaf. In cases where \Treebeard{} has already padded and balanced the tree 
+(Section~\ref{sec:treeorder}), unrolling the prologue loop completely avoids all 
+traversal induced branching.  In the case of probability-based tiling, the 
+prologue loop enables specialization of leaf checks so that these checks are 
+faster (and less general) for the most probable leaves.
 
@@
 Currently, \Treebeard{} performs a na\"ive parallelization of the inference computation. When parallelism is enabled, the 
-loop over the input rows is parallelized using a \op{parallel.for} construct in MLIR. Treebeard rewrites 
+loop over the input rows is parallelized using MLIR's OpenMP support. \Treebeard{} rewrites 
 the mid-level IR by tiling the loop over the input rows with a tile size equal to the number of cores. 
-As a concrete example, consider the case where we intend to perform inference using a model with 4 trees 
-on a batch of 64 rows. Further, assume that we wish to parallelize this computation across 8 cores. 
-\Treebeard{} then generates the following IR.
+As a concrete example, consider the case where we intend to perform inference 
+using a model with four trees on a batch of 64 rows. Further, assume that we 
+wish to parallelize this computation across 8 cores.  \Treebeard{} then 
+generates the following IR:
 \begin{lstlisting}[style=c++]
-  forest = ensemble(...)
   parallel.for i0 = 0 to 64 step 8 {
@@
 \end{lstlisting}
-Currently, \Treebeard{} does not perform other standard parallelization optimizations  (such as loop tiling) as they
-are generic and independent of the problem domain. We leave a more thorough exploration of parallelizing decision trees
-to future work.
+Currently, \Treebeard{} does not perform other standard parallelization 
+optimizations as they are generic and independent of the 
+problem domain. We leave a more thorough exploration of parallelizing decision 
+trees to future work.

### 2022-04-25 docs/MICRO 2022/abstract.tex
Micro paper (#37)
* Add results for multiple batch sizes run

* Changes to benchmark code

* Add xgboost and treelite compare results

* Add AMD results

* Add parallel performance results

* code to generate graphs + modified result numbers

* Fix bug in plotting code

* Add grid to line graphs

* Add grid lines

* Add results for varying number of cores

* Run parallel tests

* Add AMD parallel results

* Update graph generator

* Add scaling results for AMD

* Fix images, alter font sizes and axes

* Just use Probability based tiling

* Update images - Change axis font size, fix wrong graph

--- 2022-04-14 docs/MICRO 2022/abstract.tex
@@
 \begin{abstract}
-  Decision tree ensembles are a commonly used machine learning model
-  generated by machine learning techniques like gradient boosting and random
-  forests. These models are used in applications ranging from recommendation systems
-  to tabular data analysis and are also deployed at scale in the enterprise.
-  Several libraries such as XGBoost, LightGBM, and Sklearn expose algorithms for both
-  training and inference with decision tree ensembles. These systems incorporate a
-  limited set of optimizations usually targeted at specific hardware. Further, existing
-  systems do not specialize the inference code to the model being used.
+  Decision tree ensembles are among the most commonly used machine learning
+  models. These models are used in a wide range of applications and are deployed
+  at scale. Inference with decision tree ensembles is usually performed with
+  libraries such as XGBoost, LightGBM, and Sklearn. These libraries incorporate
+  a fixed set of optimizations for the hardware targets they support. However,
+  maintaining these optimizations is prohibitively expensive with the evolution
+  of hardware.  Further, they do not specialize the inference
+  code to the model being used, leaving significant performance on the table.
 
-  This paper presents Treebeard, an extensible compiler for decision tree models.
-  Treebeard progressively lowers inference code to LLVM IR through multiple intermediate
-  representations. By applying model specific optimizations at the higher levels, loop
-  optimizations at the middle level, and machine specific optimizations lower down,
-  Treebeard can specialize inference code for each model on each supported
-  hardware target. To improve model inference performance, Treebeard performs several 
-  novel optimizations such as tree tiling, tree walk unrolling, and tree walk interleaving.
+  This paper presents \Treebeard{}, an optimizing compiler that progressively
+  lowers  inference code to LLVM IR through multiple intermediate abstractions.
+  By applying model-specific optimizations at the higher levels, tree walk
+  optimizations at the middle level, and machine-specific optimizations lower
+  down, \Treebeard{} can specialize inference code for each model on each
+  supported hardware target. \Treebeard{} combines several novel optimizations
+  at various abstraction levels to mitigate architectural bottlenecks and enable
+  SIMD vectorization of tree walks.
 
-  We implement Treebeard using the MLIR compiler infrastructure and
-  demonstrate the utility of Treebeard by evaluating it on a diverse set of
-  tree ensemble models. We report that Treebeard optimizations improve
-  average latency over a batch of inputs by 2.2X compared to an unoptimized baseline.
-  Further, we find that Treebeard is significantly faster than other frameworks like
-  XGBoost  (3X on average) and Treelite (5X on average) in both single and multi-core
-  settings.
+  We implement \Treebeard{} using the MLIR compiler infrastructure and
+  demonstrate its utility by evaluating it on a diverse set of benchmarks.
+  \Treebeard{} is significantly faster than state-of-the-art systems, XGBoost
+  and Treelite: by 2.6$\times$ and 4.7$\times$ respectively in a single-core
+  execution setting, and by 2.3$\times$ and 2.7$\times$ respectively in
+  multi-core settings.
+
 \end{abstract}
+
+  %generated by techniques like gradient boosting and random forests.
+
+  % (run repeatedly on billions of data items).
+ % such as tree tiling, tree walk unrolling, and tree walk interleaving.
+ % Experimental evaluation demonstrates that \Treebeard{} optimizations
+  % improve average latency over a batch of inputs by 2.2$\times$ over our baseline version.
+ %% other frameworks like

### 2022-04-25 docs/MICRO 2022/conclusions.tex
Micro paper (#37)
* Add results for multiple batch sizes run

* Changes to benchmark code

* Add xgboost and treelite compare results

* Add AMD results

* Add parallel performance results

* code to generate graphs + modified result numbers

* Fix bug in plotting code

* Add grid to line graphs

* Add grid lines

* Add results for varying number of cores

* Run parallel tests

* Add AMD parallel results

* Update graph generator

* Add scaling results for AMD

* Fix images, alter font sizes and axes

* Just use Probability based tiling

* Update images - Change axis font size, fix wrong graph

\section{Conclusions}
This paper presents \Treebeard{}, a compiler that automatically generates  
efficient code for decision tree inference. \Treebeard{} gradually lowers  
inference code to LLVM IR through multiple intermediate abstractions.
It composes several novel optimizations to specialize inference code 
for each model on each supported hardware target. 
Experimental evaluation demonstrates that \Treebeard{} is 
significantly faster than state-of-the-art frameworks XGBoost and Treelite.

%% when executing 
% on both single core (by 2.6$\times$ and 4.7$\times$ respectively) and %% on 
% multiple cores (2.3$\times$ and 2.7$\times$ respectively). 
% The optimizations implemented in \Treebeard{} optimizations improved average latency over a batch of inputs by 2.2$\times$ over its
% own baseline version. %% lays a strong foundation to
%% automatically delivering high performance with high productivity through
%% compiler infrastructure for decision tree inference.
\Treebeard{} is a first step in automatically generating high performance
inference code for decision tree ensembles. It significantly simplifies retargeting 
high performance inference routines to new hardware targets. We believe that this methodology
can be extended to benefit a richer ensemble of ML models.

### 2022-04-25 docs/MICRO 2022/introduction.tex
Micro paper (#37)
* Add results for multiple batch sizes run

* Changes to benchmark code

* Add xgboost and treelite compare results

* Add AMD results

* Add parallel performance results

* code to generate graphs + modified result numbers

* Fix bug in plotting code

* Add grid to line graphs

* Add grid lines

* Add results for varying number of cores

* Run parallel tests

* Add AMD parallel results

* Update graph generator

* Add scaling results for AMD

* Fix images, alter font sizes and axes

* Just use Probability based tiling

* Update images - Change axis font size, fix wrong graph

--- 2022-04-17 docs/MICRO 2022/introduction.tex
@@
 \section{Introduction}
-% Decision trees are among the most widely used ML models. Cite Kaggle and looking glass 
-Decision tree ensembles are one of the most popular classes of machine learning models \cite{KaggleSurvey,LookingGlass}.
-They are generated by machine learning techniques like gradient boosting and random forests. 
-The Kaggle state of data science and machine learning survey \cite{KaggleSurvey} shows that 
-these are the most widely used classes of models among data scientists. An analysis of 
-several machine learning pipelines that are in production use in a real large scale web company showed that 
-gradient boosting machines and random forests were the two most widely used ML algorithms in the company \cite{LookingGlass}.
-Not only are decision tree based models widely used, they are also used in a diverse set of 
-applications\cite{DecisionTreesOverview}. GBM models were used in the CERN large hadron collider
-to classify particles based on the data collected from the collider\cite{LHCModel}. Other applications 
-include search engines\cite{YahooSearch}, prediction of the financial performance of companies\cite{Finance},
-medical diagnostics\cite{Med1, Med2} and recommendation and notification systems\cite{Facebook}.
+\label{sec:intro}
+Decision tree ensembles are among the most popular classes of machine learning (ML) models~\cite{KaggleSurvey,LookingGlass}. The Kaggle state of ML survey 2021 \cite{KaggleSurvey} shows that 
+more than 70\% of data scientists use decision tree ensembles while less than 40\% use neural networks. 
+Such ensembles are generated by ML techniques like gradient boosting and random forests, reported to be the top
+two algorithms used in production pipelines at a large scale web company\cite{LookingGlass}. Such models also 
+have many diverse use cases\cite{DecisionTreesOverview}. Not only are they used in computer science applications 
+like search~\cite{YahooSearch}, recommendation and notification systems\cite{Facebook}, but also in 
+financial~\cite{Finance} and medical~\cite{Med1, Med2} applications. Gradient boosted models (GBM) are
+even used in the CERN large hadron collider to classify particles\cite{LHCModel}.
 
-% Decision tree ensemble overview
-Each node of a decision tree has a threshold value $v$ and a feature index (or column index) $i$.
+\CommentOut{
+Decision tree ensembles simply aggregate predictions made by a set of decision trees.
 To compute the prediction of a decision tree on an input row $x$\footnote{An input row is a
-vector of numbers.}, a path from the root to leaf is computed. At a node, if $x_i < v$, then 
-the walk moves to the left child. If not, we move to the right child. When a leaf is reached, 
+vector of numbers.}, a path from the root to leaf is traversed.
+Each node of a decision tree has a threshold value $v$ and a feature index $i$.
+At a node, if $x_i < v$, then
+the walk moves to the left child. If not, it moves to the right child. When a leaf is reached,
 the value of the leaf is returned as the prediction of the tree.
-Most ML techniques combine several decision trees into a forest (or ensemble) to improve prediction accuracy.
-To compute the prediction of the forest, the prediction of each tree is computed and these predictions are 
-then combined (usually by adding them).
 
-% Inference on decision trees is hard 
-% Pointer chasing -- cache performance, branch prediction, dependency stalls, not easy to vectorize
-Given the wide spread use of decision tree ensembles, the ability to perform both high throughput and low latency inference is important.
-\TODO{AP Since we don't talk about latency or throughput explicitly, should we be saying this?}
-However, the tree walks required for inference are not straight-forward to optimize on a modern CPU. 
-Firstly, naively implemented tree walks have poor spatial and temporal locality and therefore poor cache performance \cite{FAST,MilindTreeVectorization}.
-Additionally, since the decision of which node to evaluate next can only be made after evaluating the current node,
-true dependencies between instructions cause several pipeline stalls. Also, using SIMD instructions to accelerate 
-tree walks is extremely challenging \cite{MilindTreeVectorization}.
+%Most decision tree-based techniques combine several trees into a forest (or ensemble) to improve prediction accuracy.
+%The prediction of each tree is computed and these predictions are 
+%then combined (usually by adding them). 
+}
 
-Several systems like XGBoost\cite{XGBoost}, LightGBM\cite{LightGBM} and Sklearn\cite{Sklearn} implement
-decision tree ensemble inference. These are libraries and implement some optimizations. However, these optimizations are hand
-coded and need to be manually rewritten or redesigned for every target that is supported. Also, these 
-systems cannot apply specialize inference code depending on the model. A compiler based approach is 
-needed to address these problems. However, existing compilers only support fixed code generation techniques\cite{Treelite}.
-To fill the need for an extensible compiler infrastructure for decision forest ensembles, we design and 
-build Treebeard\footnote{In J. R. R. Tolkein's The Lord of the Rings, Treebeard was the oldest of the Ents left in Middle-earth, an 
-ancient tree-like being who was a ``shepherd of trees''.}, an optimizing compiler for decision tree ensemble inference.
-Our motivations for building Treebeard are as follows. 
+A recent survey~\cite{SageMaker} that breaks down the cost of ownership of a
+data science solution among different stages indicates that inference costs
+have now become the dominant (in the range 45\%-65\%) component~\cite{Hummingbird}. This motivates
+a careful study of the efficiency of inference with decision tree ensembles.
+Such inference involves a simultaneous traversal of multiple decision trees.
+Given an input row, a path from the root to leaf is traversed for each tree and
+the values at the leafs are aggregated together to produce a prediction.  At
+each internal node an input feature $x_i$ is compared with a threshold value $v$ for that node
+to determine whether the walk moves left or right\footnote{  
+The condition $x_i<v$ is the \emph{node predicate} or the \emph{node condition}.
+At each node, if the node predicate is true, then
+the walk moves to the left child. If not, it moves to the right child. When a leaf is reached,
+the value $v$ of the leaf is returned as the prediction of the tree.}.
+
+Optimizing tree walks on a CPU is challenging because they have irregular access
+patterns and are sensitive to various architectural parameters.  A profile of a
+basic traversal shows that a na\"ively implemented tree walk has poor spatial
+and temporal locality, and therefore poor cache performance~\cite{FAST,MilindTreeVectorization}.  Frequent branching and true dependencies
+between instructions cause several pipeline stalls. Furthermore, accelerating tree
+walks with low level optimizations like vectorization using SIMD instructions is
+extremely challenging~\cite{MilindTreeVectorization}.
+
+Today, the most popular systems for performing inference for tree-based models
+are libraries like XGBoost\cite{XGBoost}, LightGBM\cite{LightGBM} and
+Sklearn\cite{Sklearn}.  As one would expect, these libraries implement a fixed
+set of optimizations for the hardware configurations they support. As hardware
+evolves, specializing the library to newer generations gets prohibitively
+expensive and usually has a high lead time. Further, these libraries do not
+specialize inference to the specifics of the model being used for inference.
+
+Recent research has seen the advent of optimizing compilers for deep neural
+networks~\cite{Halide, TVM, TensorComprehensions, Tiramisu, XLA}. Such compilers
+have been shown to perform much better than domain-specific libraries in several
+cases.  Surprisingly, despite their widespread usage, very few compilers exist
+for decision tree ensembles~\cite{Treelite, Hummingbird}.  One 
+system, Hummingbird~\cite{Hummingbird} builds on the success of DNN compilers by
+transforming decision tree traversals to tensor operations. However, it sometimes
+introduces more expensive operations like matrix multiplication
+just to be able to leverage compilers
+like TVM~\cite{TVM} for efficient code generation. As their results indicate,
+when it comes to tree-based models this strategy does not always succeed: some
+benchmarks' performance degrades when compared to a state-of-the-art library. 
+Treelite\cite{Treelite}, the only other compiler for tree inference, uses 
+a simple compilation strategy. It aggressively expands all trees in the 
+model into \op{if-else} statements. As we show later in the paper, this strategy
+is quite limited in applying various optimizations.
+
+This paper presents the design and implementation of \Treebeard{}\footnote{In 
+J. R. R. Tolkein's The Lord of the Rings, Treebeard was the oldest of the Ents left in Middle-earth, an 
+ancient tree-like being who was a ``shepherd of trees''.}, an extensible optimizing compiler infrastructure for decision tree ensemble
+inference. \Treebeard{} enables three different classes of optimizations: (i) pertaining to the nature of the algorithm, i.e., simultaneous traversal of several trees.
+(ii) pertaining to the properties of the tree being traversed, like imbalances in structure and probabilities of reaching leaves. 
+(iii) targeting the characteristics of the CPU, like support for vector instructions. 
+We carefully construct the compiler so that different optimizations are applied at different intermediate
+representations of the computation. By doing so, \Treebeard{} facilitates the selection and composition of
+several optimizations at each level of abstraction, as well as automatic and efficient code generation.
+
+We implement \Treebeard{} using the MLIR compiler infrastructure\cite{MLIR} and evaluate it 
+on a diverse set of decision tree-based models. We report that the code generated by
+\Treebeard{} is significantly faster than state-of-the-art systems: it is on average
+$2.6\times$ and $4.7\times$ faster than XGBoost and Treelite respectively on a single core. 
+Even with a very simple parallelization strategy, \Treebeard{} achieves speedups of $2.3\times$ and $2.7\times$
+compared to XGBoost and Treelite on a 16-core system.
+% \TODO{kr: is the below line needed}.
+% Our optimizations improve performance significantly over a simple \Treebeard{} generated baseline (with comparable single-core performance as XGBoost but better than treelite) 
+% on both Intel (by $2.2\times$) and AMD (by $1.8\times$) processors.
+
+In summary, we make the following contributions.
 \begin{enumerate}
-  \item Existing libraries \cite{XGBoost, Treelite, LightGBM, VPred} perform specific optimizations and are hard to maintain/upgrade as hardware evolves. 
-  Additionally, they cannot tailor inference code to the model being used. Past work has identified that the optimizations required change with the model
-  and architectural parameters like cache size\cite{CacheConscious1, CacheConscious2, VPred}. 
-  \item The repeated effort to optimize and maintain libraries on several targets is prohibitive. The proliferation of new architectures
-  further exacerbates this problem. Compilers have been successful in alleviating this problem in other domains\cite{Halide, TVM}.
-  However, no such compiler infrastructure currently exists for decision tree ensembles.
-  \item Currently, no system exists that allows a thorough exploration of the potential optimization space for decision tree ensemble inference. 
-  Compilers have been successful in addressing this problem with ML models like DNNs \cite{TVM, Tiramisu, XLA}. However, compiler techniques for
-  decision tree ensembles are less well studied.
+\item We design and implement \Treebeard{}, an extensible compiler infrastructure for decision tree model inference. 
+\Treebeard{} is built to allow exploration of several optimizations and code generation techniques. % \TODO{Is this claim too grand?}
+\item We propose novel tiling transformations that reduce the cost of tree walks by grouping tree nodes into ``tiles''.
+We propose tiling algorithms that can utilize specific properties of the model being compiled.
+\item We design and implement various model and loop transformations that significantly improve generated inference
+code performance. \Treebeard{} also enables parallelization of the inference computation.
+\item We develop a general infrastructure for the vectorization of decision tree walks.
+This includes general support for code generation and memory-layout optimizations for tiled trees.
 \end{enumerate}
-
-We make the following contributions.
-\begin{enumerate}
-  \item We design and build Treebeard, an extensible compiler infrastructure for decision tree model inference. The 
-  infrastructure is built to allow exploration of optimization and code generation techniques. \TODO{Is this claim too grand?}
-  \item We develop a general infrastructure for the vectorization of decision tree walks based on grouping tree nodes into ``tiles''.
-   This includes general support for code generation and the in-memory representation of tiled trees. The infrastructure can be 
-   used to tile trees based on different cost functions.
-  \item We show that trees can be tiled using different cost functions. We present two novel tiling methods that are implemented
-  using the general tiling infrastructure. 
-  \item We design and implement various model and loop transformations that significantly improve generated inference code performance 
-  to show the power of the proposed framework.
-\end{enumerate}

### 2022-04-25 docs/MICRO 2022/looporder.tex
Micro paper (#37)
* Add results for multiple batch sizes run

* Changes to benchmark code

* Add xgboost and treelite compare results

* Add AMD results

* Add parallel performance results

* code to generate graphs + modified result numbers

* Fix bug in plotting code

* Add grid to line graphs

* Add grid lines

* Add results for varying number of cores

* Run parallel tests

* Add AMD parallel results

* Update graph generator

* Add scaling results for AMD

* Fix images, alter font sizes and axes

* Just use Probability based tiling

* Update images - Change axis font size, fix wrong graph

\subsection{Loop Rewriting}
\label{Sec:LoopReordering}
\Treebeard{} supports a wide range of loop rewrites in the process of 
lowering from the high level IR to the mid level IR. \Treebeard{} 
supports fissing and permuting of the loop nest 
over the tree, input row pairs. Figure~\ref{Fig:Overview} shows the specific
case of permuting loops. Two versions of MIR generated for two different 
loop orders are shown -- ``one tree at a time'' that walks one tree for 
all input rows before moving to the next tree and ``one row at a time'' that
walks all trees for an input row before moving to the next row. The 
structure of the loop nest to be generated in MIR is decided at the HIR level and 
communicated to the lowering pass through attributes (as mention in 
section~\ref{Sec:Overview}).

### 2022-04-25 docs/MICRO 2022/main.tex
Micro paper (#37)
* Add results for multiple batch sizes run

* Changes to benchmark code

* Add xgboost and treelite compare results

* Add AMD results

* Add parallel performance results

* code to generate graphs + modified result numbers

* Fix bug in plotting code

* Add grid to line graphs

* Add grid lines

* Add results for varying number of cores

* Run parallel tests

* Add AMD parallel results

* Update graph generator

* Add scaling results for AMD

* Fix images, alter font sizes and axes

* Just use Probability based tiling

* Update images - Change axis font size, fix wrong graph

@@
 \usepackage{xcolor}
+\usepackage{tabularx,booktabs}
+\usepackage{multirow}
 \usepackage[normalem]{ulem}
@@
 \usepackage[keeplastbox]{flushend}
-\usepackage{subfig}
+%\usepackage{subfig}
 \usepackage{algorithm}
@@
 \usepackage{algpseudocode}
+\usepackage{caption}
+\usepackage{subcaption}
+\usepackage{authblk}
 % Always include hyperref last
@@
     morekeywords={ensemble, getTree, getRoot, isLeaf, isLeafTile, traverseTreeTile, getLeafValue, loadThresholds,
-                  loadFeatureIndices, loadTileShape, getChildNode, parallel, InterleavedWalk, evaluateTilePredicates,
-                  getChildTile, WalkDecisionTree},    
+                  loadFeatureIndices, loadTileShape, getChildTile, parallel, InterleavedWalk, evaluateTilePredicates,
+                  getChildTile, WalkDecisionTree, moveToChildTile},
     backgroundcolor=\color{backcolour},
@@
 %%%%%%%%%%%---SETME-----%%%%%%%%%%%%%
-\title{Treebeard : An Optimizing Compiler for Tree Based ML Inference} 
+\title{Treebeard : An Optimizing Compiler for Decision Tree Based ML Inference}
 %%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%
+
+\author[1]{Ashwin Prasad}
+\author[2]{Sampath Rajendra}
+\author[2]{Kaushik Rajan}
+\author[1]{R Govindarajan}
+\author[1]{Uday Bondhugula}
+\affil[1]{Department of Computer Science And Automation, Indian Institute of Science}
+\affil[2]{Microsoft Research}
+
+%\numberofauthors{5} %  in this sample file, there are a *total*
+% of EIGHT authors. SIX appear on the 'first-page' (for formatting
+% reasons) and the remaining two appear in the \additionalauthors section.
+%
+
+% \author{Ashwin Prasad}
+% \affiliation{%
+%  \institution{Indian Institute of Science}
+%  \city{Bangalore}
+%  \state{KA}
+%  \country{India}}
+% \email{ashwinprasad@iisc.ac.in}
+
+% \author{
+% % You can go ahead and credit any number of authors here,
+% % e.g. one 'row of three' or two rows (consisting of one row of three
+% % and a second row of one, two or three).
+% %
+% % The command \alignauthor (no curly braces needed) should
+% % precede each author name, affiliation/snail-mail address and
+% % e-mail address. Additionally, tag each line of
+% % affiliation/address with \affaddr, and tag the
+% % e-mail address with \email.
+% %
+% \alignauthor
+% Ashwin Prasad\\
+%       \affaddr{Indian Institute of Science, Bangalore}
+%       \email{ashwinprasad@iisc.ac.in}
+% \alignauthor
+% Sampath Rajendra\\
+%       \affaddr{Microsoft Research}
+%        \email{srajendra@microsoft.com}
+% \alignauthor
+% Kaushik Rajan\\
+%       \affaddr{Microsoft Research}
+%        \email{krajan@microsoft.com}
+% % \and
+% \alignauthor
+% R Govindarajan\\
+%       \affaddr{Indian Institute of Science, Bangalore}
+%       \email{govind@iisc.ac.in}
+% \alignauthor
+% Uday Bondhugula\\
+%       \affaddr{Indian Institute of Science, Bangalore}
+%       \email{udayb@iisc.ac.in}
+% }
+
+\renewcommand\Affilfont{\itshape\small}
 
@@
 \input{overview}
-\input{details}
+\input{opts}
+\input{results}
 \input{related}
+\input{conclusions}
+
 %%%%%%% -- PAPER CONTENT ENDS -- %%%%%%%%

### 2022-04-25 docs/MICRO 2022/overview.tex
Micro paper (#37)
* Add results for multiple batch sizes run

* Changes to benchmark code

* Add xgboost and treelite compare results

* Add AMD results

* Add parallel performance results

* code to generate graphs + modified result numbers

* Fix bug in plotting code

* Add grid to line graphs

* Add grid lines

* Add results for varying number of cores

* Run parallel tests

* Add AMD parallel results

* Update graph generator

* Add scaling results for AMD

* Fix images, alter font sizes and axes

* Just use Probability based tiling

* Update images - Change axis font size, fix wrong graph

@@
 \section{Treebeard Overview}
+\label{Sec:Overview}
 Figure \ref{Fig:CompilerStructure} shows the high level structure of \Treebeard{}. 
@@
 %% Popular frameworks like XGBoost and LightGBM are supported and the system is easily extensible to other frameworks.
-Given an input model our compiler generates optimized inference code. Specifically it generates a 
-callable batch inference function \texttt{predictForest} that, given an array of input rows, computes an
+Given an input model our compiler generates optimized inference code. Specifically, it generates a 
+batch inference function, \texttt{predictForest} that, given an array of input rows, computes an
 array of model predictions.
@@
 are implemented using a combination of annotation and lowering. An operation at a higher 
-abstraction is annotated with attributes that indicate what kind of optimization is 
-to be performed while lowering it. The lowering transformation uses this information to generate 
-optimized lower level code. For example, tree tiling and loop ordering (figure \ref{Fig:Overview} 
+abstraction is annotated with attributes that guide how code should be lowered. 
+Actual transformation happens while lowering.
+%The lowering transformation uses this information to generate optimized lower level code.
+ For example, tree tiling and loop ordering (Figure \ref{Fig:Overview} 
 shows two possible loop orders -- one tree at a time and one row at a time) are decided 
 at the highest abstraction. These decisions are communicated to the lowering pass 
-which explicitly encodes them in the lowered IR as shown in Figure \ref{Fig:Overview}.
+which explicitly encodes them in the lowered IR as shown in the Figure.
 
 %Treebeard first constructs a high level representation of the decision forest inference operation.
-\Treebeard{}'s IR has three abstraction levels as shown in Figure \ref{Fig:Overview}.  
+\Treebeard{}'s IR has three abstraction levels as shown in Figure \ref{Fig:CompilerStructure}.  
 At the highest level (HIR), the input model is represented as a collection of binary trees. This is shown 
 in the second row of Figure \ref{Fig:Overview}. At this level of abstraction,
-\Treebeard{} tiles nodes together to transform a binary tree into an $n$-ary tree 
-and decides what order trees are to be traversed in. In the example in the figure, 
-trees are tiled with a tile size of 2. Tiling is indicated using colored elipses drawn 
-around nodes that are in the same tile. Tree reordering is another transformation that is performed at this abstraction. 
+\Treebeard{} tiles nodes together to transform a binary tree into an $n$-ary tree. 
+In the example in the Figure, 
+trees are tiled with a tile size of 2, as indicated using colored ellipses drawn 
+around nodes that are in the same tile.
+
+Tree reordering is another transformation that is performed at this abstraction. 
 One objective of tree reordering is to group identically structured trees so that they can share the same traversal code
@@
 
-The high level IR is then lowered to a mid-level IR (MIR). The aim of MIR is 
-to allow optimizations that are independent of the final memory layout of the model. 
-The lowering to MIR performs loop transformations like loop tiling, permutation, parallelization and unrolling.
+The high level IR is then lowered to a mid-level IR (MIR).
+The lowering to MIR also performs loop transformations like loop tiling, permutation, parallelization and unrolling.
 At this level of the IR (the third row in Figure \ref{Fig:Overview}),
 the order in which trees, input row pairs are walked is explicitly represented in the IR using loop nests.
-Figure \ref{Fig:Overview} shows two possible ways that loop nests could be generated -- the first
+Figure \ref{Fig:Overview} shows two possible ways that loop nests could be generated.
+\CommentOut{
+ -- the first
 walks all rows for one tree before moving to the next tree while the second walks all trees for 
-one row before moving to the next row. Another aspect handled by this lowering is the fissing of 
+one row before moving to the next row.
+}
+ Another aspect handled by this lowering is the fissing of 
 loops to make sure that trees with the same structure share code. Here, both versions of MIR shown 
 ensure that Tree1 and Tree3 share the same traversal code, while traversal code for Tree2 is different.
-Additionally, tree walk optimizations, such as tree walk unrolling (shown in figure \ref{Fig:Overview}) and 
+
+ The aim of MIR is 
+to allow optimizations that are independent of the final memory layout of the model. 
+Tree walk optimizations, such as tree walk unrolling (shown in Figure \ref{Fig:Overview}) and 
 peeling are performed on the MIR.
@@
 values are inserted into the generated code and all tree operations in the mid-level IR are lowered to explicitly 
-reference these buffers. Additionally, at this level of the IR, \Treebeard{} generates vectorized code to take 
-advantage of SIMD instructions where applicable. Finally, the inference function is translated to LLVM IR and 
+reference these buffers. Additionally, at this level of the IR, \Treebeard{} generates vectorized code. 
+Figure~\ref{Fig:Overview} shows operations for which SIMD instructions are eventually generated. 
+Finally, the inference function is translated to LLVM IR and 
 JIT compiled to executable code.
+
+We implement \Treebeard{} using the MLIR compiler infrastructure\cite{MLIR}. MLIR provides modular and reusable infrastructure
+to develop intermediate representations for domain-specific compilers. Firstly, to enable compiler writers to 
+quickly design and build intermediate representations, MLIR has a large set of 
+\textbf{\emph{dialects}} that provide operations from various domains. Each dialect 
+contains a logically related set of operations and types; dialect transformations and conversions allow lowering and conversion of these operations and types to other lower-level dialects and subsequently translation to LLVM IR.
+For example, the \texttt{scf} dialect provides control flow constructs like loops and if-else operations. 
+The \texttt{memref} dialect provides functionality to manage buffers. We build \Treebeard{} using a custom
+MLIR dialect and a combination of dialects provided in MLIR. Secondly, MLIR provides an extremely powerful 
+operator rewriting infrastructure. We make extensive use of this to implement transformations.
 
@@
 
-\begin{figure}
+\begin{figure}[htb]
   \centering
-  \includegraphics[width=\linewidth]{figures/compiler.pdf}
-  \caption{Treebeard compiler structure}
+  \includegraphics[width=\linewidth]{figures/compiler}
+  \caption{\Treebeard{} compiler structure.}
   \label{Fig:CompilerStructure}
@@
 
-\begin{figure*}
+\begin{figure*}[htb]
   \centering
-  \includegraphics[width=\textwidth]{figures/OverviewExample.pdf}
-  \caption{\Treebeard{} IR lowering and optimization details. The three abstraction levels in Treebeard's IR are shown. The
-           high level IR is a tree based IR to perform model level optimization, the mid-level IR is for
+  \includegraphics[width=\linewidth]{figures/OverviewExample_New.pdf}
+  \vskip 10pt
+  \caption{\Treebeard{} IR lowering and optimization details: the three abstraction levels in \Treebeard{}'s IR are shown. The
+           high level IR is a tree-based IR to perform model level optimization, the mid-level IR is for
            loop optimizations that are independent of memory layout and the low level IR allows us to perform
@@
 One important point to note here is that details such as the data structure used for the trees are not explicitly encoded in the IR.
-This allows us to reuse optimization and lowering passes on this level of the IR regardless of what the final in memory representation of the model is.
+This allows us to reuse optimization and lowering passes on this level of the IR regardless of what the final in-memory representation of the model is.
 
@@
 % \subsection{Low Level IR}
-% The IR is finally lowered to a form where the in memory representation of the model is made explicit.
+% The IR is finally lowered to a form where the in-memory representation of the model is made explicit.
 Buffers to hold the model values are inserted into the generated code and all tree operations in the mid-level IR are lowered to explicitly reference these buffers.

### 2022-04-25 docs/MICRO 2022/probtiling.tex
Micro paper (#37)
* Add results for multiple batch sizes run

* Changes to benchmark code

* Add xgboost and treelite compare results

* Add AMD results

* Add parallel performance results

* code to generate graphs + modified result numbers

* Fix bug in plotting code

* Add grid to line graphs

* Add grid lines

* Add results for varying number of cores

* Run parallel tests

* Add AMD parallel results

* Update graph generator

* Add scaling results for AMD

* Fix images, alter font sizes and axes

* Just use Probability based tiling

* Update images - Change axis font size, fix wrong graph

@@
-\subsection{Probability Based Tiling}
+\subsection{Probability-Based Tiling}
 \label{sec:ProbTiling}
@@
 
+\CommentOut{
+%% We have already given the motivation for probabilistic tiling!
 The next algorithm we propose exploits the inherent biases among the leaves of a decision tree.
-In typical machine learning models some leafs (equivalently outcomes or predictions) are more likely to be reached than others.
+In typical machine learning models some leaves (equivalently outcomes or predictions) are more likely to be reached than others.
 In such settings, having balanced tiled trees is not sufficient to minimize expected inference time.
@@
 This means that most trees in \op{epsilon} are not leaf biased.
+}
 
@@
 \subsubsection{More Notation}
-In order to formulate the probability based tiling algorithm as an optimization problem, we define the following.
+In order to formulate the probability-based tiling algorithm as an optimization problem, we define the following.
 \begin{enumerate}
@@
 
+\CommentOut{
 \begin{figure}
+  \subfloat[airline-ohe]{
     \centering
     \includegraphics[width=\linewidth]{figures/airline-ohe.stats.train.txt.png}
-    \caption{Statistical profile for airline-ohe}
     \label{Fig:AirlineOHEStats}
-\end{figure}
-\begin{figure}
+  }
+  \subfloat[epsilon]{
     \centering
     \includegraphics[width=\linewidth]{figures/epsilon.stats.train.txt.png}
-    \caption{Statistical profile for epsilon}
     \label{Fig:EpsilonStats}
+  }
+  \caption{Statistical profiles.}
 \end{figure}
+}
+
 \CommentOut{
@@
 One other observation we make is that most models have some leaf biased trees while the rest of the trees have equally likely leaves.
-\TODO{AP Maybe define a term for trees with roughly equally likely leaves?} We design the probability based tiling algorithm to take advantage of this property 
+\TODO{AP Maybe define a term for trees with roughly equally likely leaves?} We design the probability-based tiling algorithm to take advantage of this property 
 of decision tree ensembles. 
 }
+
+\CommentOut{
 \subsubsection{The Optimization Problem}
@@
 Observe that the latency of one tree walk is proportional to the number of tiles that need to be evaluated to reach the leaf.
-It is easy to see that for a leaf biased tree, basic tilling does not optimize for this objective, it considers all leafs to be equally likely.
+It is easy to see that for a leaf biased tree, basic tilling does not optimize for this objective, it considers all leaves to be equally likely.
 
@@
 We leave this out in the interest of space. 
-Instead, we use the simple greedy algorithm listed in algorithm \ref{Alg:GreedyTilingAlgo} to construct a valid tiling given the node probabilities\footnote{Probabilites for internal nodes can be computed from probablities for leafs by summing up the probabilities of all leafs that belong to the sub-tree rooted at the internal node.
+Instead, we use the simple greedy algorithm listed in algorithm \ref{Alg:GreedyTilingAlgo} to construct a valid tiling given the node probabilities\footnote{Probabilites for internal nodes can be computed from probablities for leaves by summing up the probabilities of all leaves that belong to the sub-tree rooted at the internal node.
 Leaf probabilities are collected during training.}.
@@
 Subsequently, the tiling procedure is recursively performed on all nodes that are destinations for edges going out of the constructed tile.
-
+}
 % \subsubsection{Dynamic Programming Formulation}
@@
 \begin{algorithm}
-    \caption{Greedy Probability Based Tree Tiling}
+    \caption{Greedy probability-based tree tiling.}
     \label{Alg:GreedyTilingAlgo}
-    \begin{algorithmic}
+    \small{
+    \begin{algorithmic}[1]
         \Procedure{TileTree}{$\Tree = (V, E, r)$, $n_t$} 
@@
     \end{algorithmic}
+    }
 \end{algorithm}
+
+The goal of probability-based tiling is to minimize the average inference latency, or equivalently to minimize the expected
+number of tiles that are evaluated to compute one tree prediction
+We present a simple greedy heuristic algorithm (Algorithm~\ref{Alg:GreedyTilingAlgo}) to construct a valid tiling given
+the node probabilities\footnote{Probabilities for internal nodes can be computed from probabilities for leaves by summing
+up the probabilities of all leaves that belong to the sub-tree rooted at the internal node. Leaf probabilities are collected
+during training.}. The algorithm starts at the root and greedily keeps adding the most probable legal node to the current
+tile until the maximum tile size is reached. Subsequently, the tiling procedure is recursively performed on all nodes 
+that are destinations for edges going out of the constructed tile. 
+In the algorithm, these edges are denoted by $Out(Tile)$ and the subtree rooted at a node $v$ is denoted by $S_v$.
+It is easy to see that the tiling constructed by Algorithm \ref{Alg:GreedyTilingAlgo} is valid.
 
@@
 lookup table needed (section \ref{sec:LookupTable}) caused increased L1 cache misses. In order to 
-alleviate this, we only perform probability based tiling on trees that are leaf biased.
+alleviate this, we only perform probability-based tiling on trees that are leaf biased.
 }
-We find probability based tiling is only beneficial for leaf biased trees\footnote{Turns out that for trees that are not leaf biased, probability based tiling produces many more tile shapes (see Section~\ref{sec:tileShapes}) which direclty impacts the cost of \op{getChildTile} making it more expensive than basic tiling.}.
-Recall that
-a tree to be leaf biased if a small fraction of leaves, say $\alpha$, can cover a large fraction of training inputs, say $\beta$.
-We only perform probability based tiling on trees with thresholds $\alpha=0.05$ and $\beta=0.9$ and fall back to uniform tiling otherwise.
+
+As  probability-based tiling is only beneficial for leaf-biased trees,
+%\footnote{Turns out that for trees that are not leaf biased, probability-based tiling produces many more tile shapes (see Section~\ref{sec:tileShapes}) which direclty impacts the cost of \op{getChildTile} making it more expensive than basic tiling.}.
+%% Recall that a tree to be leaf biased if a small fraction of leaves, say $\alpha$, can cover a large fraction of training inputs, say $\beta$.
+we  perform probability-based tiling on trees only when a small fraction ($\alpha$) of leaves cover a large part ($\beta$) of the training inputs.
+% In our experiments we set $\alpha=0.075$ and $\beta=0.9$ and fall back to basic tiling otherwise.

### 2022-04-25 docs/MICRO 2022/related.tex
Micro paper (#37)
* Add results for multiple batch sizes run

* Changes to benchmark code

* Add xgboost and treelite compare results

* Add AMD results

* Add parallel performance results

* code to generate graphs + modified result numbers

* Fix bug in plotting code

* Add grid to line graphs

* Add grid lines

* Add results for varying number of cores

* Run parallel tests

* Add AMD parallel results

* Update graph generator

* Add scaling results for AMD

* Fix images, alter font sizes and axes

* Just use Probability based tiling

* Update images - Change axis font size, fix wrong graph

@@
 \section{Related Work}
-While decision trees are heavily used there is little prior work on building optimizing compilers for tree 
-based inference. Below we discuss some related work and contrast it to \Treebeard{}.
+While decision trees are heavily used there is little prior work on building 
+optimizing compilers for tree-based inference. We discuss below related work and 
+contrast it to \Treebeard{}.
 
@@
 of tensor primitives so 
-they can be integrated into tensor based frameworks like TensorFlow \cite{TensorFlow}.
+they can be integrated into tensor-based frameworks like TensorFlow \cite{TensorFlow}.
 Hummingbird does not perform any optimizations tailored to decision trees. It 
@@
 
-\emph{Libraries:} Currently, the most popular systems for decision tree based models 
+\emph{Libraries:} Currently, the most popular systems for decision tree-based models 
 are libraries. XGBoost\cite{XGBoost} and LightGBM\cite{LightGBM} are the most  
@@
 extremely popular for random forest models. These libraries implement both 
-training and inference. However, as mentioned in section \ref{sec:Intro}, porting 
+training and inference. However, as mentioned in Section \ref{sec:intro}, porting 
 optimizations in these libraries to newer architectures and maintaining them is 
@@
 any model specific optimizations. Jo et. al.\cite{MilindTreeVectorization} describe techniques 
-to vectorize tree based applications. They do not study optimizations specific to decision trees.
+to vectorize tree-based applications. They do not study optimizations specific to decision trees.
 Both these works vectorize tree walks by performing different tree walks on each vector lane.
 The main issue with this approach is the divergence of tree walks. Another issue is that 
-memory accesses are gathers rather than vector loads. \Treebeard{}'s approach to 
+memory accesses are \op{gather}'s rather than vector loads. \Treebeard{}'s approach to 
 vectorization solves both these issues. Also, these approaches are not precluded 
@@
 the routine based on FFT size and the target machine. 
+However, no such frameworks for decision tree ensembles exist currently.
+\Treebeard{} is a first step in this direction and unlocks several future 
+optimization opportunities.

### 2022-04-25 docs/MICRO 2022/results.tex
Micro paper (#37)
* Add results for multiple batch sizes run

* Changes to benchmark code

* Add xgboost and treelite compare results

* Add AMD results

* Add parallel performance results

* code to generate graphs + modified result numbers

* Fix bug in plotting code

* Add grid to line graphs

* Add grid lines

* Add results for varying number of cores

* Run parallel tests

* Add AMD parallel results

* Update graph generator

* Add scaling results for AMD

* Fix images, alter font sizes and axes

* Just use Probability based tiling

* Update images - Change axis font size, fix wrong graph

\section{Experimental Evaluation}
\label{sec:evaluation}

We evaluate \Treebeard{} on two machines. The first one has an Intel Core
i9-11900K (Rocket Lake) processor with 8 physical cores (16 logical cores with
hyperthreading), 128~GB of RAM and Ubuntu~20.04.3 LTS.  The second system has an
AMD Ryzen~7 4700G process with 8 physical cores (16 logical cores with
hyperthreading), 64~GB of RAM, and runs CentOS Linux release 7.9.2009. For
comparisons with XGBoost and Treelite, we used Python version 3.10, XGBoost
version 1.5.0 and Treelite version 2.3.0.

% Linux kernel version 5.11.0.

The benchmark models we used are trained using datasets listed in Table~\ref{Tab:Benchmarks}. These datasets were also used by the Intel Machine Learning
Benchmark suite~\cite{MLBenchmarks}. We used
the hyper-parameters from this suite for all the datasets. All models were trained
with XGBoost~\cite{XGBoost}.  For all experiments, we run inference on 2 million test
inputs (divided into the required number of batches). To compute speedups, we
compare the mean time taken per input row between configurations.

\begin{table}
  \centering
  \small{
  \begin{tabularx}{\linewidth}{c r r r r}
   \toprule
   \textbf{Dataset} & \textbf{\#Features} & \textbf{\#Trees} & \textbf{Max Depth} & \textbf{\#Leaf-biased}\\
   \midrule
   \texttt{abalone} & 8 & 1000 & 7 & 438\\
   \texttt{airline} & 13 & 100 & 9 & 8 \\
   \texttt{airline-ohe} & 692 & 1000 & 9 & 976 \\
   \texttt{covtype} & 54 & 800 & 9 & 283\\
   \texttt{epsilon} & 2000 & 100 & 9 & 0\\
   \texttt{letter} & 16 & 2600 & 7 & 0 \\
   \texttt{higgs} & 28 & 100 & 9 & 8\\
   \texttt{year} & 90 & 100 & 9 & 0\\
   \bottomrule
  \end{tabularx}
  \vskip 5pt
  \caption{\label{Tab:Benchmarks}List of benchmark datasets and their 
  parameters. 
  The column \op{Leaf-biased} reports the number of leaf-biased trees per benchmark with $\langle\alpha =0.075, \beta =0.9 \rangle$ . }
  }
\end{table}

%\TODO{AP How do we say that we tried several optimizations and picked the best?}
In order to find the best combination of optimizations, several configurations were explored
for each benchmark at different batch sizes. The grid of optimizations explored is listed in Table~\ref{Tab:Optimizations}. 
%% RG
We report performance results for inference performed using (i) \Treebeard{} baseline code without performing any optimizations
presented in Sections~\ref{Sec:HIR}, \ref{Sec:MIR} and \ref{Sec:LIR}, (ii) \Treebeard{} code with the combination of optimizations that performs best,
(iii) XGBoost, and (iv) Treelite. We refer to the unoptimized \Treebeard{} code as the scalar baseline.
% \begin{enumerate}
%   \item \emph{Loop order} : One tree at a time vs. one row at a time
%   \item \emph{Tile size} : 1, 2, 4, 8
%   \item \emph{Tiling type} : Basic tiling vs probability-based tiling
%   \item \emph{Tree padding and unrolling} : Yes or no
%   \item \emph{Tree walk interleaving} : 2, 4, 8 (Number of walks interleaved)
% \end{enumerate}
\begin{table}[htbp]
  \small{
  \begin{tabularx}{\linewidth}{l l}
   \toprule
   \textbf{Optimization} & \textbf{Configurations}\\
   \midrule
   \multirow{2}{*}{\texttt{Loop order}} &  One tree at a time\\
                       &  One row at a time\\
   \midrule
   \texttt{Tile size} & 1, 2, 4, 8 \\
   \midrule
   \multirow{2}{*}{\texttt{Tiling type}} & Basic tiling\\
   & Probability-based tiling \\
   \midrule
   \texttt{Tree padding and unrolling} & Yes, No \\
   \midrule
   \texttt{Tree walk interleaving} & 2, 4, 8 \\
   \midrule
   \texttt{$\langle \alpha, \beta\rangle$ for leaf-bias} &  $\langle0.05, 0.9 \rangle$, $\langle0.075, 0.9 \rangle$ , $\langle 0.1, 0.9 \rangle$ \\
     \bottomrule
  \end{tabularx}
  \vskip 5pt
  \caption{\label{Tab:Optimizations}Space of optimizations explored.}
  }
\end{table}
\CommentOut{
\TODO{kr: this is out of place. Future work?}
Even though \Treebeard{} has the ability to perform non-trivial rewrites of the loop nest over
tree, input row pairs, we do not explore anything beyond the simple loop interchange listed above
in the current work. We leave a more thorough exploration of loop optimizations to future work.
}
% \subsection{Outline}
% \begin{enumerate}
%   \item Summary plot -- speedups over unopt baseline for both Intel and AMD. One bar graph for specific batch size over all benchmarks + geo mean. (AMD + Intel)
%   \item Effectiveness of optimizations (AMD + Intel)
%   \begin{enumerate}
%     \item Speedup with basic + prob tiling
%     \item Speedup with unroll etc etc
%   \end{enumerate}
%   \item Parallel scaling
%   \begin{enumerate}
%     \item Speedup over batch sizes with all cores -- one line plot and one bar graph for a specific batch size (Intel + AMD)
%     \item Mean speedup scaling with number of cores (Intel + AMD)
%   \end{enumerate}
%   \item Speedup over xgboost and treelite (intel only) -- parallel and serial. For each, one line graph and one bar graph.
% \end{enumerate}

% \begin{figure}
%   \centering
%   \includegraphics[width=\linewidth]{figures/result/scalar_overall_speedup_benchmark_intel_amd.png}
%   \caption{Overall single-core speedup of optimized \Treebeard{} code over a na\"ive scalar baseline at batch size of 1024.}
%   \label{Fig:OverallScalar}
% \end{figure}

% \begin{figure}
%   \centering
%   \includegraphics[width=\linewidth]{figures/result/parallel_overall_speedup_benchmark_intel_amd.png}
%   \caption{Overall multi-core speedup of optimized \Treebeard{} code over a na\"ive scalar baseline at batch size of 1024 and with 16 cores.}
%   \label{Fig:OverallParallel}
% \end{figure}

\begin{figure*}[tbp]
  \centering
  \subfloat[Single-core]{
    \centering
    \includegraphics[width=0.48\linewidth]{figures/result/scalar_overall_speedup_benchmark_intel_amd.png}
    \label{Fig:OverallScalar}
  }
  \subfloat[16 cores]{
    \centering
    \includegraphics[width=0.48\linewidth]{figures/result/parallel_overall_speedup_benchmark_intel_amd.png}
    \label{Fig:OverallParallel}
  }
  \caption{Speedup of \Treebeard{} optimized code over the scalar baseline 
           at batch size of 1024. Number over each bar is the mean time to 
           perform inference on one input row in $\mu$s. Numbers on the geomean bars 
           are the geomean speedups over all benchmarks.}
  \label{fig:Summary}
\end{figure*}

\subsection{Summary of Improvements on Different Hardware}
Figure~\ref{Fig:OverallScalar} compares the single core performance of \Treebeard{} optimized code with the scalar baseline, on different hardware. The plot has two bars corresponding to the Intel and AMD machines,
each showing the speedup of the optimized version over the corresponding baseline.
As can be seen, \Treebeard{} does consistently better on all benchmarks, attaining up to $3.5\times$ speedup on both machines.
Between Intel and AMD we find that on average the speedups are a little better on Intel. This is because the Intel machine has a much
more efficient implementation of the \op{gather} instruction. Our vectorized implementation uses this to load features.
Because of this and other differences in low level architecture we find that the best set of optimizations on the two machines are completely different. On the
Intel machine we aggressively tile the tree with a tile size of 8 for all benchmarks. While on the AMD machine optimal
performance is sometimes achieved at much lower tile sizes (\op{airline} prefers a tile size of 1,
\op{airline-ohe}, \op{covtype} and \op{epsilon} prefer a tile size of 4). Similarly, we interleave walks more aggressively
on Intel than on AMD, interleaving by a factor of eight works best on the Intel
system on all benchmarks, while on the AMD one, two benchmarks perform optimally
with an interleaving of four.

Figure~\ref{Fig:OverallParallel} compares multi-core performance with 16 cores on different hardware. All speedups are
reported with respected to an unoptimized single core run. As can be seen even with simple parallelization 
we are able to achieve good scalability.
In fact, some optimizations combine well with parallelizations to achieve super-linear speedup (see \op{epsilon} on AMD).
This particular benchmark has the largest number of features and we suspect that on a single core it is mainly bottlenecked
on the memory sub-system when accessing them.

%% RG
%% Figure~\ref{fig:OverallScalarOverBatchSizes} reports the geomean speedup (over all benchmarks) of \Treebeard{} over baseline version on single core processor for different batch sizes. It can be seen that the  proposed optimizations improve the performance consistently across different batch sizes. 
In summary by using a different set of optimizations and parameters for different benchmark-hardware
combinations \Treebeard{} is able to alleviate several bottlenecks and achieve a significant speedup over the baseline.
\Treebeard{} optimizations give a geomean speedup of $2.45\times$ over all benchmarks on Intel RocketLake and 
$2.06\times$ on AMD Ryzen 7. 
% \TODO{kr: i think we should report the geomean from here on serial and parallell, inaddition to XGBoost and Treelite.}
% \TODO{AP : We don't define what ``scalar baseline'' means anywhere. Should we define it?}

\subsection{Comparison with XGBoost and Treelite}

% \begin{figure}
%   \centering
%   \includegraphics[width=\linewidth]{figures/result/serial_speedup_per_benchmark_xgoost_treelite.png}
%   \caption{Speedup of \Treebeard{} optimized code compared to XGBoost and Treelite on single core at batch size of 1024.}
%   \label{Fig:XGBoostSerialCompare}
% \end{figure}

\begin{figure*}
  \centering
  \subfloat[Single-core]{
    \centering
    \includegraphics[width=0.48\textwidth]{figures/result/serial_speedup_per_benchmark_xgoost_treelite.png}
    \label{Fig:XGBoostSerialCompare}
  }
  \subfloat[16 cores]{
    \centering
    \includegraphics[width=0.48\textwidth]{figures/result/parallel_speedup_per_benchmark_xgoost_treelite.png}
    \label{Fig:XGBoostParallelCompare}
  }
  \caption{Comparison of \Treebeard{} with XGBoost and Treelite with batch size 1024. Bars show speedup of 
           \Treebeard{} optimized code relative to XGBoost and Treelite. Numbers on XGBoost speedup bars are the 
           mean inference time per row for XGBoost in $\mu$s. Numbers on the Treelite bars are mean inference times 
           for Treelite. Numbers on the geomean bars are the geomean speedups over all benchmarks.}
  \label{fig:XGBoostCompare}
\end{figure*}

\begin{figure}[tbp]
  \centering
  \includegraphics[width=0.8\linewidth]{figures/result/scalar_Intel_mean_speedups_treebeard_vs_xgboost_treelite.png}
  \caption{Geomean speedup (over all benchmarks) of \Treebeard{} over XGBoost
  and Treelite on single-core over several batch sizes.}
\label{Fig:XGBoostBatchSizeCompare} \end{figure}

Figures~\ref{Fig:XGBoostSerialCompare} and~\ref{Fig:XGBoostParallelCompare} report a comparison of \Treebeard{} with
two state-of-the-art frameworks, XGBoost~\cite{XGBoost} which is widely used by ML practitioners and Treelite~\cite{Treelite} a
basic compiler for decision tree inference. As the plots show, \Treebeard{} is significantly better than both these systems.
It is atleast $2\times$ faster on most benchmarks over either systems in both single-core and multi-core settings.
Figure~\ref{Fig:XGBoostBatchSizeCompare} shows that these performance improvements are consistent across a wide range of input batch sizes. 
\Treebeard{} performs several novel optimizations that neither of these systems perform. In fact, apart from parallelization,
all the other optimizations (tiling, tree-ordering, walk interleaving, walk peeling, vectorization and layout optimizations)
in \Treebeard{} are new. These results demonstrate the utility of building a generic compiler to implement different optimizations.

\subsection{Sensitivity Analysis}
This section reports the impact of individual optimizations and reports the sensitivity of results to both batch size and degree of parallelism. All speedup's are with respect to the scalar baseline running on a single core.
\subsubsection{Impact of Optimizations}

\begin{figure*}
  \centering
  \subfloat[Tiling]{
    \centering
    \includegraphics[width=0.48\linewidth]{figures/result/scalar_Intel_max_speedup_unif_prob_tiling.png}
    \label{Fig:IntelTiling1024}
    }
  \subfloat[Walk Unrolling, Walk Interleaving]{
    \centering
    \includegraphics[width=0.48\linewidth]{figures/result/scalar_Intel_max_speedup_unif_tiling_pipelining.png}
    \label{Fig:IntelInterleaving1024}
    }
  \caption{Impact of individual \Treebeard{} optimizations at batch size 1024 (Intel RocketLake).
           Number on the bar is the average inference time per row in $\mu$s. Numbers on the geomean bars are the geomean
           speedups over all benchmarks.}
  \label{fig:OptSpeedup}
\end{figure*}

Figure~\ref{Fig:IntelTiling1024} compares different tiling algorithms on Intel (the trends on AMD are identical,
though as explained before the best tile sizes differ), the plot has two bars per benchmarks one shows the speedup
over baseline when all trees use the basic tiling algorithm, while the second tiles leaf-biased trees
($\langle\alpha =0.075, \beta =0.9 \rangle$ as defined in Section~\ref{sec:ProbTiling})  with probability-based tiling
(other trees continue to be tiled with basic tilling). The number of leaf-biased trees in each model 
is given in the last column of Table~\ref{Tab:Benchmarks}. For these plots we disable all mid-level IR optimizations,
but apply the low level optimizations as they go hand-in-hand with tiling.

As can be seen, even basic tiling is highly efficient, speeding up benchmarks by $1.3 - 2.5\times$.
Probability-based tilling does even better. Recall that probability-based tiling makes use of additional information about leaf probabilities (how
often a particular leaf is encountered while performing inference on the training data). As seen from the graph,
specializing the tiling to this property of the model yields additional gains. The speedup increases from
$2\times$ to $3.1\times$ for \op{airline-ohe}, the benchmark with the highest fraction of leaf-biased trees.
Several other benchmarks see an increase in speedup of $0.2- 0.4\times$. Three benchmarks~\op{epsilon, year}
and \op{letter} see little or no impact because these benchmarks have no leaf-biased trees.

Figure~\ref{Fig:IntelInterleaving1024} shows the additional benefit over and above tiling 
that \op{tree walk interleaving} and \op{tree walk peeling \& unrolling} achieve.
As can be seen these optimizations bring in significant additional improvements (on average the speedup improves from $1.5\times$ to $2.4\times$).
As discussed earlier these optimizations target different bottlenecks; while tiling improves locality and enables vectorization, these
optimizations eliminate different sources of pipeline stalls.
This clearly highlights the need for an extensible compiler framework where independent optimizations can be added at different levels of abstraction.

\subsubsection{Impact of Batch Size}
Figure~\ref{Fig:OverallScalarOverBatchSizes} shows the geomean speedup over all benchmarks for various batch size when all \Treebeard{} optimizations
are performed. The graph shows that the speedups due to \Treebeard{}'s 
optimizations work equally well across all batch sizes on both the Intel and the 
AMD machines. We infer that it would be useful to specialize tiling over the 
batch size to the target machine from the slightly different trends for the 
Intel and AMD machines. However, we leave this for future work.  
We also find that performance of \Treebeard{}'s generated 
code scales well with an increasing number of cores (plot not shown). Even though \Treebeard{} 
currently only implements a na\"ive parallelization strategy, performance scales 
reasonably well with increasing number of cores.

Overall, our results show that \Treebeard{} provides significant performance gains compared to existing systems like XGBoost and Treelite.
The optimizations designed and implemented in \Treebeard{} provide significant speedups. Also, these speedups are robust to changes
in hardware and parameters like batch size.

% \begin{figure}[htbp]
%   \centering
%   \includegraphics[width=0.8\linewidth]{figures/result/scalar_Intel_max_speedup_unif_tiling_pipelining.png}
%   \caption{Speedups (single core) due to walk unrolling and walk interleaving at batch size 1024 (Intel RocketLake).
%            Number on the bar is the average inference time per row in $\mu$s. Numbers on the geomean bar are 
%            the geomean speedups over all benchmarks. }
%   \label{Fig:IntelInterleaving1024}
% \end{figure}

\begin{figure}[htbp]
  \centering
  \includegraphics[width=0.8\linewidth]{figures/result/scalar_overall_mean_speedups_intel_amd.png}
  \caption{Geomean speedup (over all benchmarks) of optimized \Treebeard{} code over the scalar baseline on single-core
           over several batch sizes.}
  \label{Fig:OverallScalarOverBatchSizes}
\end{figure}

% \begin{figure}
%   \centering
%   \includegraphics[width=\linewidth]{figures/result/parallel_core_scaling_intel_amd.png}
%   \caption{\Treebeard{} scaling with varying number of cores. Speedup is computed over the na\"ive \Treebeard{} scalar baseline.}
%   \label{Fig:CoreScaling}
% \end{figure}

### 2022-04-25 docs/MICRO 2022/tiling.tex
Micro paper (#37)
* Add results for multiple batch sizes run

* Changes to benchmark code

* Add xgboost and treelite compare results

* Add AMD results

* Add parallel performance results

* code to generate graphs + modified result numbers

* Fix bug in plotting code

* Add grid to line graphs

* Add grid lines

* Add results for varying number of cores

* Run parallel tests

* Add AMD parallel results

* Update graph generator

* Add scaling results for AMD

* Fix images, alter font sizes and axes

* Just use Probability based tiling

* Update images - Change axis font size, fix wrong graph

@@
 
-\subsection{Tiling}
+\subsection{Tree Tiling}
 \label{sec:Tiling}
@@
 
-While a decision tree is naturally represented by a binary tree, 
-it is not the best representation for tree traversal as it (i) requires many memory accesses, 
-(ii) has poor branching structure and (iii) cannot make use of vector instructions. 
-This section proposes a tiling optimization where we group multiple nodes in a decision tree into a single tile, effectively transforming a binary tree into an $n$-ary tiled tree.
-This not only allows the compiler to generate vectorized code (see Section~\ref{sec:Vectorization} ) to traverse trees but also enables spatial locality improvements by grouping nodes that are likely to be accessed together.
-We demonstrate this with two tiling heuristics later in the section.
+While a decision tree is naturally represented as a binary tree, 
+this is not an efficient representation for tree traversal as it (i) requires many memory accesses, 
+(ii) has poor branching structure which may cause high branch misprediction rates and (iii) cannot use vector instructions. 
+This section proposes a tiling optimization where multiple nodes of a decision tree are grouped into a single tile,
+effectively transforming a binary tree into an $n$-ary tiled tree. 
+Note that an ideal decision tree traversal visits only a subset of nodes in a tile, with strict dependency between nodes at different levels.
+Unfortunately, this dependency prevents parallelization/vectorization of the tree walk. However, by speculatively evaluating conditions 
+of all nodes in a tile, we can eliminate this dependency and enable
+the compiler to generate vectorized code (see Section~\ref{sec:Vectorization}) to traverse trees.
+Secondly, this enables spatial locality improvements by grouping together nodes that are likely to be accessed together (Section~\ref{Sec:MemoryRep}).   
+We present two different tiling heuristics later in this section.
 
@@
 
-Once trees are tiled Treebeard generates tree walks with the code structure shown below.
+Once trees are tiled \Treebeard{} generates tree walks with the code structure shown below.
 \begin{lstlisting}[style=c++]
-  ResultType Prediction_Function(...) {
-    // ...
-    Tile t = getRootTile(tree)
-    while (!isLeaf(tree, t)) do {
+  WalkDecisionTree(...) {
+    tile = getRootTile(tree)
+    while (!isLeaf(tree, tile)) do {
       // Evaluate predicates of all nodes in the tile
-      predicates = evaluateTilePredicates(t, rows[i])
+      predicates = evaluateTilePredicates(tile, rows[i])
 
       // Move to the correct child of the current tile
-      t = getChildTile(tree, t, predicates) 
+      tile = moveToChildTile(tree, tile, predicates) 
     }
-    treePrediction = getLeafValue(t)
-    // ...
+    treePrediction = getLeafValue(tile)
   }  
 \end{lstlisting}
-The code is just an abstract representation of a tiled tree walk that enables efficient lowering of specific steps in subsequent stages. 
-\op{evaluateTilePredicates} (speculatively) computes the predicates on all nodes in a tile (line 6).
-Then \op{getChildTile} (line 9), uses the computed predicate values to determine which child of the current tile to move to.
-We defer a description of how these operators are lowered to a later section but focus on tiling algorithms in this section.
+The code is an abstract representation of a tiled tree walk that enables efficient lowering of specific steps in subsequent stages. 
+\op{evaluateTilePredicates} (speculatively) computes the predicates of all nodes in a tile (line 6)\footnote{ 
+\op{evaluateTilePredicates} expands to the operations from line 10 to 15 in the listing in Section~\ref{sec:Vectorization}.}.
+Then \op{moveToChildTile} (line 9), 
+uses the computed predicate values to determine which child of the current tile to move to\footnote{\op{moveToChildTile} expands 
+to the operations from line 18 to 25 in the listing in Section~\ref{sec:Vectorization}.}. We defer a description of how these 
+operators are lowered to Section~\ref{sec:Vectorization} and instead focus on tiling algorithms in this section.
 
@@
 
-\subsection{Tiles and Tree Tiling}
+\subsubsection{Tiles and Tree Tiling}
 \label{sec:ValidTiling}
 }
-\subsubsection*{Conditions for Efficient Tiling}
+\subsubsection{Conditions for Valid Tiling}
 \label{sec:ValidTiling}
-While any arbitrary partitioning of the nodes of a tree could be considered for tiling we impose a few intuitive constraints.
+While any arbitrary partitioning of the nodes of a tree could be considered for tiling, we impose a 
+few simple constraints to simplify the design of the compiler.
 % to only allow nodes that are likely to be accessed together to be grouped into a tile. 
 % \TODO{kr : replace $n_t$ with sz?}
-Given a tree $\Tree = (V, E, r)$ and a tile size $n_t$ we impose the following constraints on the generated tiles $\{ T_1, T_2, ... ,T_m \}$ .
+Given a tree $\Tree = (V, E, r)$ and a tile size $n_t$ we impose the following constraints on the generated tiles $\{ T_1, T_2, ... ,T_m \}$.
 \begin{description}
-    \item[Partitioning] $T_1 \cup T_2 ... \cup T_m = V$ and $T_i \cap T_j = \emptyset$ for all $i\neq j$
-    \item[Connectedness] If $u, v \in T_i$, there is a (undirected) path connecting $u$ and $v$ fully contained in $T_i$.
-   \item [Leaf seperation] $\forall l \in L_{\Tree}$ : $l \in T_i \rightarrow v \notin T_i \;\; \forall v \in V \backslash \{l\}$
-  \item [Maximal tiling] if there are tiles such that. $|T_i| < n_t$ then there is no $v \in V\backslash \{ T_i \cup L_{\Tree} \}$ such that $(u, v) \in E$ for some $u \in T_i$. 
-\end{description}
+    \item[Partitioning:] $T_1 \cup T_2 ... \cup T_m = V$ and $T_i \cap T_j = 
+    \emptyset$ for all $i\neq j$.
+    \item[Connectedness:] If $u, v \in T_i$, there is a (undirected) path connecting $u$ and $v$ fully contained in $T_i$.
+   \item [Leaf separation:] $\forall l \in L_{\Tree}$ : $l \in T_i \rightarrow v \notin T_i \;\; \forall v \in V \backslash \{l\}$.
+  \item [Maximal tiling:] if there are tiles such that $|T_i| < n_t$, then there 
+is no $v \in V\backslash \{ T_i \cup L_{\Tree} \}$ such that $(u, v) \in E$ for 
+some $u \in T_i$.  \end{description}
 The \textbf{partitioning} and \textbf{maximal tiling} constraints together ensure that we group nodes into as few tiles as possible.
 {\textbf{Connectedness}} ensures that each tile is a sub-tree, a natural grouping of nodes that are likely to be accessed together.
-The {\textbf{Leaf separation}} constraint ensures that leafs are not tiled along with internal nodes.
-Leafs in a decision tree need special handling, they are used to check for walk termination and to determine the output (prediction).
-This constraint ensures that tiles are homogenous, this in-turn allows us to specialize the in-memory layout of trees and also simplifies code generation.
-We discuss leaf handling and tree layout in section ~\ref{Sec:MemoryRep}.
-We refer to any tiling that satisfies the above constraints as a \emph{valid} tilling.
+The {\textbf{leaf separation}} constraint ensures that leaves are not grouped with internal nodes, as leaves in a decision tree need special handling and are used to check for walk termination and to determine the output (prediction).
+Thus, the {\textbf{leaf separation}} constraint ensures that tiles are homogeneous.
+This in-turn allows \Treebeard{}  to specialize the in-memory layout of trees and also simplifies code generation.
+We discuss leaf handling and tree layout in Section ~\ref{Sec:MemoryRep}.
+We refer to any tiling that satisfies the above constraints as a \emph{valid} tiling.
 
@@
 
+\subsubsection{Rationale for Different Tiling Methods}
+\label{sec:rationale}
+
+Before we present different tiling methods, we first discuss an interesting property of decision trees and 
+explore whether it can be exploited to optimize tree traversal. We find that the probabilities of reaching different leaves
+in a tree can vary significantly, making some root to leaf paths more frequently traversed than others. As the set of 
+decision trees that are walked for a given ML model is fixed, we ask whether this statistical property can be 
+exploited to generate efficient traversal code.% We address this question by developing different decision tree tiling methods.
+
+To understand how often different leaves are reached, let us look at the percentage of leaves required to 
+cover a significant part (say 80\% or 90\%) of the training data\footnote{We use the training data to compute tree statistics 
+since the test data must have a distribution similar to the training data.}.
+% in the decision tree ensemble. 
+This is captured in Figures~\ref{Fig:AirlineOHEStats} and~\ref{Fig:EpsilonStats}.
+%% that are generated from training data. 
+Each line in these graphs corresponds to a fixed fraction  (say $f$) of the training data. 
+A point on this line at coordinate $(x, y)$ means that a fraction $y$ of trees in the model could cover a
+fraction $f$ of all training inputs with a fraction $x$ of leaves. For example, the first point on the
+$f=0.9$ line in Figure \ref{Fig:AirlineOHEStats}(at the bottom left of the plot) indicates that about 52\% of trees ($y$ value) need only 1\% of their
+leaves ($x$ value) to cover 90\% of the training input. 
+In general, Figure~\ref{Fig:AirlineOHEStats} shows that very few leaves are needed to cover a very large fraction 
+of inputs for the benchmark \op{airline-ohe}. This means that a small fraction of leaves are very likely. 
+We call trees with a small number of extremely likely leaves \textbf{\emph{leaf-biased}}.
+On the other hand, for the benchmark \op{epsilon},
+Figure \ref{Fig:EpsilonStats} shows that trees need a much larger fraction of their leaves to cover a significant fraction of the training input.
+This means that most trees in \op{epsilon} are not leaf-biased. For leaf-biased trees, it would be beneficial to tile nodes such that the
+depth of the most probable leaves are minimized even at the expense of 
+increasing the depth of the less probable leaves. We call the tiling algorithm 
+designed to do this \emph{probability-based tiling} (Section \ref{sec:ProbTiling}).
+
+% \TODO{AP One hole here is why doesn't this method of tiling work for unbiased trees.}
+
+\begin{figure*}[tbp]
+  \centering
+  \subfloat[\texttt{airline-ohe}]{
+    \centering
+    \includegraphics[width=0.35\linewidth]{figures/airline-ohe.png}
+    \label{Fig:AirlineOHEStats}
+  }
+  %\newline
+  \hspace{2cm}
+  \subfloat[\texttt{epsilon}]{
+    \centering
+    \includegraphics[width=0.35\linewidth]{figures/epsilon.png}
+    \label{Fig:EpsilonStats}
+  }
+  \caption{Statistical profiles for \texttt{airline-ohe} and \texttt{epsilon}.}
+\end{figure*}
+
+% From the above discussion, we can infer that when the decision tree ensemble is not leaf-biased, it would be beneficial to tile the trees in 
+% such a way that the depth of the tree (in terms of the number of tiles) is minimized across all the leaves (assuming there is no major skew 
+% in the probabilities of different leaves).
+
+However, when there is no clear leaf-bias, a reasonable objective is to minimize the number of nodes executed 
+speculatively. An intuitive heuristic to achieve this objective is to minimize the depth of each tile 
+that is constructed. \emph{Basic tiling} (Section \ref{sec:UnifTiling}) is a heuristic designed to do this. 
+It is possible to define other variants of tiling(e.g., minimize the maximum leaf depth), but we leave this for future work.
+We believe that \Treebeard{}'s tiling infrastructure can support and optimize all such variants that produce valid tilings.
+
+\CommentOut{
+\subsubsection{The Optimization Problem}
+
+%We assume that we are given the probabilities of each leaf node of the decision tree (these can easily be computed using the training data).
+For every leaf $l \in L$, we %are given the probability $p_l$ that the leaf $l$ is reached.
+
+Observe that the latency of one tree walk is proportional to the number of tiles that need to be evaluated to reach the leaf.
+It is easy to see that for a leaf biased tree, basic tilling does not optimize for this objective, it considers all leaves to be equally likely.
+
+The goal of probablistic tiling is to minimize the average inference latency, or equivalently the minimize the expected number of tiles that are evaluated to compute one tree prediction.
+More formally, the problem is to find a \emph{valid} (as defined in Section~\ref{sec:ValidTiling}) tiling $\mathcal{T}$ such that the following objective is minimized.
+\[
+    \min_{\mathcal{T} \in \mathcal{C}(T)}{\sum_{l \in L_{\Tree}} p_l.depth_{\mathcal{T}}(l)}
+\]
+where the minimization is over all valid tilings $\mathcal{T}$ of the tree $\Tree$, $depth_{\mathcal{T}}(l)$ is the depth of the leaf $l$ given tiling ${\mathcal{T}}$.
+$p_l$ is the probability of of reaching leaf $l$ as observed during training.
+
+The above optimization problem can be solved optimally using dynamic programming. 
+We leave this out in the interest of space. 
+Instead, we use the simple greedy algorithm listed in algorithm \ref{Alg:GreedyTilingAlgo} to construct a valid tiling given the node probabilities\footnote{Probabilites for internal nodes can be computed from probablities for leaves by summing up the probabilities of all leaves that belong to the sub-tree rooted at the internal node.
+Leaf probabilities are collected during training.}.
+The algorithm starts at the root and greedily keeps adding the most probable legal node to the current tile until the maximum tile size is reached.
+Subsequently, the tiling procedure is recursively performed on all nodes that are destinations for edges going out of the constructed tile.
+}
+
+\input{probtiling}
 \input{uniftiling}
-\input{probtiling}
 
 \subsection{A Note on Implementation}
-\TODO{Kr : Not sure if this is needed}
-The tiling algorithms generate a \op{TileId} attribute per tree. The \op{TileId} attribute contains a mapping from a Node to the TileId asigned to it.
-This information is used when lowering to the mid level abstraction in the form of loops. A sample of the lowered MIR code is shown in Figure~\ref{Fig:Overview}.
+%\TODO{Kr : Not sure if this is needed}
+The tiling algorithms generate a \op{TileId} attribute per tree. The \op{TileId} attribute 
+contains a mapping from a Node to the TileId assigned to it.
+This information is used when lowering to the mid level abstraction in the form of loops. 
+%A sample of the lowered MIR code is shown in Figure~\ref{Fig:Overview}.

### 2022-04-25 docs/MICRO 2022/treeorder.tex
Micro paper (#37)
* Add results for multiple batch sizes run

* Changes to benchmark code

* Add xgboost and treelite compare results

* Add AMD results

* Add parallel performance results

* code to generate graphs + modified result numbers

* Fix bug in plotting code

* Add grid to line graphs

* Add grid lines

* Add results for varying number of cores

* Run parallel tests

* Add AMD parallel results

* Update graph generator

* Add scaling results for AMD

* Fix images, alter font sizes and axes

* Just use Probability based tiling

* Update images - Change axis font size, fix wrong graph

@@
-\subsection{Tree ordering}
+\subsection{Tree Reordering}
 \label{sec:treeorder}	
@@
 % \TODO{kr : Not sure if i got this para right}
-Specializing the code for each tree in a model comes at a cost. First, the size of the generated code increases if the
+Specializing the code, as we do later in MIR for example by unrolling tree walks, for each tree in a model comes at a cost.
+First, the size of the generated code increases if the
 code generator needs to generate different code for different trees. This causes several front end stalls like instruction 
@@
 
-In order to handle this, \Treebeard{} groups trees by their structure so that trees with the same structure can 
-share code. For example, the compiler pads trees with dummy tiles to make them balanced and then sorts the trees by 
-their depth, so that trees of same depth can share the same unrolled tree walk. Padding is only done for almost balanced trees (as generated 
-by basic tiling). This is shown in the example in figure \ref{Fig:Overview}. Other metrics, like feature set commonality, 
+In order to handle this, \Treebeard{} groups trees by tree structure so that they can 
+share traversal code. For example, the compiler pads trees with dummy tiles to make them balanced and then sorts the trees by 
+their depth, so that isomorphic trees can share the same unrolled tree walk. Padding is only done for almost balanced trees (as generated 
+by basic tiling). Additionally, the loop rewriting infrastructure described in Section~\ref{Sec:LoopReordering} is used 
+to modify the loop nest so that the same code walks all trees in a group. 
+This is shown in the example in the third row of Figure \ref{Fig:Overview}. Other metrics, like feature set commonality, 
 could also be used to reorder trees\cite{Tahoe}. However, we leave the exploration of such optimizations to future work.

### 2022-04-25 docs/MICRO 2022/uniftiling.tex
Micro paper (#37)
* Add results for multiple batch sizes run

* Changes to benchmark code

* Add xgboost and treelite compare results

* Add AMD results

* Add parallel performance results

* code to generate graphs + modified result numbers

* Fix bug in plotting code

* Add grid to line graphs

* Add grid lines

* Add results for varying number of cores

* Run parallel tests

* Add AMD parallel results

* Update graph generator

* Add scaling results for AMD

* Fix images, alter font sizes and axes

* Just use Probability based tiling

* Update images - Change axis font size, fix wrong graph

@@
-\subsection{Basic Tiling Algorithm}
+\subsection{Basic Tiling}
 \label{sec:UnifTiling}
@@
 % we want to minimize the depth of speculation
-Algorithm~\ref{Alg:UnifTilingAlgo} shows a simple tiling algorithm that satisfies the above constraints and produces a valid tiling. 
+Algorithm~\ref{Alg:UnifTilingAlgo} shows the basic tiling algorithm that produces a valid tiling. It 
+attempts to minimize the depths of all constructed tiles. 
 Tiling starts at the root and constructs a tile $Tile$ by performing
@@
 the tiling procedure is recursively performed on subtrees rooted at each node that is a destination of an edge going out of the constructed 
-tile. These edges are denoted by $Out(Tile)$ and the subtree rooted at a node $v$ is denoted by $S_v$.
-It is easy to see that the tiling constructed by algorithm \ref{Alg:UnifTilingAlgo} is valid.
-% \TODO{Pass a modified tree without leafs to the level order traversal call. Explicitly add leaves as seperate tiles}
+tile. It is easy to see that the tiling constructed by Algorithm \ref{Alg:UnifTilingAlgo} is valid.
+% \TODO{Pass a modified tree without leaves to the level order traversal call. Explicitly add leaves as seperate tiles}
 % Algorithm for basic tiling
@@
   \label{Alg:UnifTilingAlgo}
-  \begin{algorithmic}
+  \small{
+  \begin{algorithmic}[1]
       \Procedure{LevelOrderTraversal}{$\Tree = (V, E, r)$, $n_t$}
@@
         \While{$\neg queue.empty() \wedge |Tile|<n_t$}
-          \State $n = queue.pop()$
+          \State $n = queue.dequeue()$
           \If {$n \in L_{\Tree}$}
@@
           \State $Tile = Tile \cup \{ n \}$
-          \State $queue.push(left(n), right(n))$
+          \State $queue.enqueue([left(n), right(n)])$
         \EndWhile
@@
   \end{algorithmic}
+  }
 \end{algorithm}
 
-One interesting property of this tilling algorithm is that it naturally reduces the imbalance in trees, especially at large tile sizes.
-As the algorithm traverses down to sparser levels of the tree, it naturally groups sub-trees containing chains of nodes, thus balancing the trees.
-While its possible to further enhance the algorithm to explicitly balance tiled trees, we find that basic tiling suffices in practice.
+One interesting property of this tiling algorithm is that it naturally reduces 
+the imbalance in trees, especially at large tile sizes. As the algorithm 
+traverses down to sparser levels of the tree, it naturally groups sub-trees 
+containing chains of nodes, thus balancing the trees. While it is possible to 
+further enhance the algorithm to explicitly balance tiled trees, we find that 
+basic tiling suffices in practice.
 

### 2022-04-25 docs/MICRO 2022/vectorization.tex
Micro paper (#37)
* Add results for multiple batch sizes run

* Changes to benchmark code

* Add xgboost and treelite compare results

* Add AMD results

* Add parallel performance results

* code to generate graphs + modified result numbers

* Fix bug in plotting code

* Add grid to line graphs

* Add grid lines

* Add results for varying number of cores

* Run parallel tests

* Add AMD parallel results

* Update graph generator

* Add scaling results for AMD

* Fix images, alter font sizes and axes

* Just use Probability based tiling

* Update images - Change axis font size, fix wrong graph

@@
+The optimizations performed by \Treebeard{} at the low-level IR are presented in this section.
+
 \subsection{Vectorization}
 \label{sec:Vectorization}
-Vectorization performed by Treebeard is enabled by the tiling transformations described in section \ref{sec:Tiling}. 
-When the low level IR is translated to LLVM IR, Treebeard generates LLVM instructions that operate on the threholds and feature indices 
-of nodes within a tile in a vector fashion. Therefore, thresholds and feature indices are loaded using vector loads and predicates are 
-evaluated using vector comparisons. These vector LLVM IR instructions are then converted to vector instructions in the target ISA by 
+\Treebeard{} specializes the tree walk over a tiled tree to make use of vector instructions. 
+%Vectorization performed by Treebeard is enabled by the tiling transformations described in section \ref{sec:Tiling}. 
+This specialization happens as we lower low level IR to LLVM IR.  These are then converted to vector instructions in the target ISA by 
 the LLVM JIT.
 
-The below listing shows some of the details of a vectorized tree walk. 
+The below listing shows the low-level IR for a vectorized tree walk in detail. 
 \begin{lstlisting}[style=c++]
@@
 
-  ResultType Prediction_Function(...) {
-    // ...
-    Node n = getRoot(tree)
-    while (isLeaf(tree, n)==false) do {
-      thresholds = loadThresholds(tree, n)
-      featureIndices = loadFeatureIndices(tree, n)
+  WalkDecisionTree(...) {
+    tile = getRoot(tree)
+    while (isLeaf(tree, tile)==false) do {
+      thresholds = loadThresholds(tree, tile)
+      featureIndices = loadFeatureIndices(tree, tile)
       // Gather required feature from the current row
@@
       // Get child index of tile we need to move to
-      tileShape = loadTileShape(tree, n)
+      tileShape = loadTileShape(tree, tile)
       childIndex = LUT[tileShapeID, comparisonIndex]
@@
       // Move to the correct child of the current node
-      n = getChildNode(tree, n, childIndex) 
+      tile = getChildTile(tree, tile, childIndex) 
     }
-    ThresholdType prediction = getLeafValue(n)
-    // ...
+    prediction = getLeafValue(tile)
   }  
 \end{lstlisting}
+
+When lowering to LLVM, operators \op{loadThresholds} and \op{loadFeatureIndices} (lines 11 and 13) are lowered to use vectorized load instructions.
+Predicates are then evaluated using vector comparison instructions (lines 15 to 18).
+The next step is to determine which child to visit (lines 21 to 25), as we explain below this depends not only on the result of the comparison but also on the shape of the sub-tree that the tile represents.
+
+\CommentOut{
 To evaluate the current tile, the vector of thresholds is first loaded (\texttt{loadThresholds}).
-This vector contains the thresholds of all nodes in the tile.
-Then, the features required for comparison are gathered into a vector (lines 11 and 13).
-The feature vector is compared to the threshold vector and the child tile to move to is determined (lines 15 to 25).
+Then, the features required for comparison are gathered into a vector. The feature 
+vector is compared to the threshold vector and the child tile to move to is determined (lines 15 to 25).
 More details about tile shapes and the look up table are presented in subsequent sections.
+}
 
-\subsubsection{Tile Shapes}
-Informally, the \textbf{\emph{tile shape}} is the shape of the region that encloses all nodes in a tile in a diagram of the decision tree.
-More formally, for a tile size $n_t$, each unique legal binary tree containing $n_t$ nodes (nodes being indistinguishable) corresponds to
-a tile shape. Figure \ref{Fig:TileSize3Shapes} enumerates all tile shapes with a tile size of 3.
+\begin{figure}
+  \centering
+  \includegraphics[width=\linewidth]{figures/TileShapes_Size3.PNG}
+  \caption{All possible tile shapes with tile size $n_t=3$.}
+  \label{Fig:TileSize3Shapes}
+\end{figure}
 
+\subsubsection{Tile Shapes and Tree Traversal}
+\label{Sec:TileShapesAndDecisionTreeInference}
+%We refer to the shape of the region that encloses all nodes in a tile in a diagram of the decision tree as the \textbf{\emph{tile shape}}.
+For a given tile size $n_t$, each unique legal binary tree containing $n_t$ nodes (nodes being indistinguishable) corresponds to
+a \textbf{\emph{tile shape}}. Figure \ref{Fig:TileSize3Shapes} enumerates all tile shapes with a tile size of 3. 
+\CommentOut{
 % There are a total of 5 tile shapes with size 3. The number of tile shapes with a tile size $n_t$, denoted by $NTS(n_t)$ is given by the following equation.
@@
 
-\begin{figure}
-  \centering
-  \includegraphics[width=\linewidth]{figures/TileShapes_Size3.PNG}
-  \caption{All possible tile shapes with tile size $n_t=3$}
-  \label{Fig:TileSize3Shapes}
-\end{figure}
-
 \subsubsection{Tile Shapes and Decision Tree Inference}
 \label{Sec:TileShapesAndDecisionTreeInference}
-Treebeard uses vector instructions to accelerate decision tree walks.
-Vector instructions are used to evaluate the predicates of all the nodes in a tile simultaneously.
-However, once the predicates of all the nodes in the tile are evaluated, computing the next tile to move to, given the outcome of the comparison depends on the tile shape of the current tile.
-To illustrate this problem, consider the case of the tiles of size 3 shown in figure \ref{Fig:TileTraversalTileSize3}.
+\Treebeard{} uses vector instructions to accelerate decision tree walks. Vector instructions are used to evaluate the predicates
+of all the nodes in a tile simultaneously. However, once the predicates of all the nodes in the tile are evaluated, computing
+the next tile to move to, given the outcome of the comparison depends on the tile shape of the current tile. To illustrate 
+this problem, consider the case of the tiles of size 3 shown in Figure \ref{Fig:TileTraversalTileSize3}.
+}
+
+%% To understand how tile shapes influence the outcome of tile evaluation, consider the case of the tiles of size 3 shown in figure \ref{Fig:TileTraversalTileSize3}
 \begin{figure}
@@
   \includegraphics[width=\linewidth]{figures/TileTraversal_Size3.PNG}
-  \caption{Example tile traversals with tile size $n_t=3$}
+  \caption{Example tile traversals with tile size $n_t=3$.}
   \label{Fig:TileTraversalTileSize3}
 \end{figure}
-The diagram shows 3 of the 5 possible tile shapes for a tile size of 3.
-The nodes drawn in black are members of the tile $t_1$.
-The nodes in blue are the entry nodes of the children tiles of $t_1$.
-\TODO{Define entry nodes}
-
+%% The diagram 
+Given the  comparison vector evaluated for a tile, the tile that needs to be traversed next depends
+on the shape of the tile. To understand this, consider Figure~\ref{Fig:TileTraversalTileSize3} that 
+shows 3 of the 5 possible tile shapes for a tile size of 3. The nodes drawn in black are members of the tile.
+The nodes in blue are the root nodes of the children tiles.
 % To traverse a tile on an input row, first, the predicate of each node in the tile is computed.
@@
 Note that a true predicte (bit value 1) on a node implies a move to the left child and a false predicate (bit value 0) implies a move to the right child.
-
-In the diagram, the bit strings (written in red) show which child we need to move to given the outcomes of the comparison.
-The bits represent the comparison outcomes of nodes and are in the order of the nodes in the tile -- marked 0, 1 and 2 in the diagram, i.e., the MSB is the predicate outcome of node 0 and the LSB the predicate outcome of node 2.
-For example, for the first tile shape, if the predicates of all nodes are true (i.e.
-the comparison outcome is 111), the next node to evaluate is $a$.
+The bit strings (written in red) show which child 
+%% we need to move to
+needs to be traversed next,  given the outcomes of the comparison.
+The bits represent the comparison outcomes of nodes -- the MSB is the predicate outcome of node 0 and the LSB the predicate
+outcome of node 2. For example, for the first tile shape, if the comparison outcome is 111, the next node to evaluate is $a$. 
 % However, if the predicate of node 1 is false, then we need to move to $d$ regardless of the outcomes of nodes 2 and 3.
-It is easy to see from the diagram that, depending on the tile shape, the same predicate outcomes can mean moving to different children.
-For example, for the outcome "011", the next tile is the 4th child (node $d$) for the first two tile shapes while it is the 3rd child for the other tile shape (node $c$).
+It is easy to see that, depending on the tile shape, the same comparison outcomes can mean moving to different children.
+For example, for the outcome 011, the next tile is the 4th child (node $d$) for the first two tile shapes while it is 
+the 3rd child for the other tile shape (node $c$)\footnote{Children of a tile are ordered left to right regardless of depth.}.
 
@@
 \label{sec:LookupTable}
-A lookup table (LUT) is used to solve the problem described in section \ref{Sec:TileShapesAndDecisionTreeInference}, i.e.
-given the outcome of the comparisons of all nodes in a tile, determine the child tile we should evaluate next.
-The LUT is indexed by the tile shape and the comparison outcome.
-Formally, the LUT is a map.
+As illustrated above a combination of tile shape and the comparison outcome can be used to determine the child to visit. 
+We introduce an additional map called a lookup table (LUT) to encode this information:
 \[
-LUT : (TileShape, < Boolean \times n_t >) \rightarrow [0, n_t] \subset \mathbb{N}
+LUT : (TileShape, \{0, 1\}^{n_t}) \rightarrow [0, n_t] \subset \mathbb{N}.
 \]
+The LUT is indexed by 
+the tile shape and the comparison outcome. 
+where $n_t$ is the tile size, $\{0, 1\}^{n_t}$ is a vector of $n_t$ booleans. The value returned by the LUT is the index of
+the child of the current tile that should be evaluated next. For example, if we are evaluating the first tile $T_1$ in 
+Figure \ref{Fig:TileTraversalTileSize3}, and the result of the comparison is 110, then $LUT(TileShape(T_1), 110)=2$ since 
+the tile %% we need to evaluate 
+to be traversed next is the tile with node $b$, which is the second child of the current tile.
 
-where $n_t$ is the tile size, $< Boolean \times n_t >$ is a vector of $n_t$ booleans.
-The value returned by the LUT is the index of the child of the current tile that should be evaluated next.
-For example, if we are evaluating the first tile $t$ in figure \ref{Fig:TileTraversalTileSize3}, and the result of the comparison is 110, then $LUT(TileShape(t), 110)=1$ since the tile we need to evaluate next is the tile with node $b$, which is the second child of the current tile.
-
-In order to realize this LUT in generated code, Treebeard associates a non-negative integer ID with every unique tile shape of the given tile size.
-The result of the comparison, a vector of booleans, can be interpreted as a 64-bit integer.
-Therefore, the LUT can be implemented as a 2 dimensional array.
+In order to realize this LUT in generated code, \Treebeard{} associates a non-negative integer ID with every unique tile shape of the
+given tile size. The result of the comparison, a vector of booleans, can be interpreted as a 64-bit integer. Therefore, the LUT
+can be implemented as a 2 dimensional array. \Treebeard{} computes the values in the LUT statically as the tile size is a compile time constant.
 % \begin{lstlisting}{style=c++}
@@
 % \end{lstlisting}
-Treebeard computes the values in the LUT statically as the tile size is a compile time constant.
 
-\subsection{In Memory Representation of Tiled Trees}
+\subsection{In-Memory Representation of Tiled Trees}
 \label{Sec:MemoryRep}
-Treebeard currently has two in memory representations for tiled trees - an array based representation and a sparse representation.
-Both representations use an array of structs to represent all tiles of the model.
+\Treebeard{} currently has two in-memory representations for tiled trees - an array based representation and a sparse representation.
+Both representations use an array of \texttt{structs} to represent the model.
 
-\subsubsection{Array Based Representation}
+\subsubsection{Array-Based Representation}
 \label{sec:ArrayBased}
-Each tree in the model is represented as an array of tiles using the standard representation of trees as arrays.
-The root node is at index 0 and for a node at index $n$ in the array, the index of its $i^{th}$ child is given by $(n_t + 1) n + (i + 1)$ (nodes in the tree of tiles have $n_t + 1$ children).
-A tile is represented by an object of the following struct.
-\begin{lstlisting}{style=c++}
-  struct Tile {
-    // A vector of TileSize elements
-    <ThresholdType x TileSize> thresholds; 
-    <FeatureIndexType x TileSize> featureIndices;
-    // Integer that identifies the tile shape
-    TileShapeIDType tileShapeID; 
-  };  
-\end{lstlisting}
-\TODO{AP Is this level of detail really needed? Also, the vector type notation needs to be introduced somewhere.}
+Each tree in the model is represented as an array of tiles using the standard representation of trees as arrays. The root node is 
+at index 0 and for a node at index $n$, the index of its $i^{th}$ child\footnote{Nodes
+in the tree of tiles have $n_t + 1$ children.} is given by $(n_t + 1) n + (i + 1)$.
+% A tile is represented by an object of the following struct.
+% \begin{lstlisting}{style=c++}
+%   struct Tile {
+%     // A vector of TileSize elements
+%     <ThresholdType x TileSize> thresholds; 
+%     <FeatureIndexType x TileSize> featureIndices;
+%     // Integer that identifies the tile shape
+%     TileShapeIDType tileShapeID; 
+%   };  
+% \end{lstlisting}
+% \TODO{AP Is this level of detail really needed? Also, the vector type notation needs to be introduced somewhere.}
 Even though this representation is simple and efficient for small models, the memory required for bigger models is very large. 
 %The memory footprint is up to 20X that of the scalar representation.
-This memory bloat causes performance problems because the span of the L1 TLB is not sufficient to efficiently translate 
-addresses for the whole model. Storing leaves as full tiles (even though leaves just have to represent one value) and the
-empty space introduced due to the array based representation of trees that are not complete account for most
-of the increase.
+This memory bloat causes performance problems due to L1 cache and L1 TLB misses.
+%because the span of the L1 TLB is not sufficient to efficiently translate addresses for the whole model.
+Storing leaves as full tiles (even though they represent a single value) and the
+empty space introduced due to the array based representation of trees that 
+are not complete account for most of the increase.
 %The sparse representation described next tries to address these issues.
@@
 \label{sec:SparseRep}
-The sparse representation tries to address the large memory footprint of the array based representation by doing the following.
+We reduce the large memory footprint of the array based representation by doing the following.
 
 \begin{itemize}
-  \item We add a child pointer to each tile to eliminate the wasted space in the array representation. This points to the first child of the tile. All children of a tile are stored contiguously.
+  \item We add a child pointer to each tile to eliminate the wasted space in the array representation. This 
+  points to the first child of the tile. All children of a tile are stored contiguously.
   \item Leaves are stored as a separate array of scalar values. Across all our benchmarks, after tiling a large fraction of 
@@
   for which this property does not hold, an extra ``hop'' is added by making the original leaf tile a normal tile. All its
-  children are made leaves with the same value as the original leaf.
+  children are leaves with the same value as the original leaf.
 \end{itemize}
@@
   \includegraphics[width=\linewidth]{figures/SparseRep_TileSize3.PNG}
-  \caption{Sparse representation with tile size $n_t=3$}
+  \caption{Sparse representation with tile size $n_t=3$. Leaves $l_4$, $l_5$, $l_6$ and $l_7$ are moved into the \op{leaves} array.
+  Extra hops are added for $l_1$, $l_2$ and $l_3$ as $t_2$ is a non-leaf tile. The new leaves added as children of $l_1$, $l_2$ and $l_3$
+  are moved to the \op{leaves} array.}
   \label{Fig:SparseRep}
@@
 
-Figure \ref{Fig:SparseRep} shows some of the details of the sparse representation.
+Figure \ref{Fig:SparseRep} shows the details of the sparse representation.
 % The tree on the left of the diagram is the actual decision tree with the nodes grouped into tiles $t_1$ and $t_2$. The tree on the right is the tree of tiles.
@@
 and has 5 elements. Each element of the array represents a single tile and has the thresholds of the nodes, the feature
-indices, a tile shape ID and a pointer to the first child (shown explicitly in red).
+indices, a tile shape ID and a child pointer (shown as red arrows). The second array, \op{leaves}, contains
+all leaf values.
 
+%%%% COMMENT START %%%%%
+
+\CommentOut{
 As a specific example, consider the tile $t_1$.
@@
 
-\CommentOut{
-\subsubsection{Code Generation for Probability Based Tiling}
-As probability based tiling pulls the most probable leaves of a decision tree nearest the root, it poses 
+\subsubsection{Code Generation for Probability-Based Tiling}
+As probability-based tiling pulls the most probable leaves of a decision tree nearest the root, it poses 
 some implementation challenges. By design, the tiling process makes the tree of tiles 
@@
 an extra hop for leaves that have non-leaf siblings. But this would mean that we add extra hops for 
-the most probable leaves after probability based tiling which defeats the optimization.
+the most probable leaves after probability-based tiling which defeats the optimization.
 \TODO{kr: connect to Peeling in MIR}
@@
 %requirement is also not increased because only a small fraction of leaves are represented as full tiles.
-While walk peeling is used to improve the performance of probability based tiling by specializing leaf tests,
+While walk peeling is used to improve the performance of probability-based tiling by specializing leaf tests,
 the transformation is by itself general and can be used in different contexts. For example, it could be used 
@@
 }
+
+%%%% COMMENT END %%%%%

### 2022-04-25 docs/MICRO 2022/results/result_parser.py

@@ -1,4 +1,3 @@
-from stat import FILE_ATTRIBUTE_NORMAL
 import pandas
 from scipy.stats.mstats import gmean
 import matplotlib.pyplot as plt
@@ -7,7 +6,10 @@ from typing import List
 import math
 import os
 
-current_directory = os.path.dirname(os.path.realpath(__file__))
+CURRENT_DIRECTORY = os.path.dirname(os.path.realpath(__file__))
+GRAPH_BATCH_SIZE = 1024
+BATCH_SIZES = [64, 128, 256, 512, 1024, 2000]
+CORE_COUNTS = [2, 4, 8, 16]
 
 class SingleBatchSizeResult:
   def __init__(self, batch_size, min_scalar_times, unif_tiling_times) -> None:
@@ -46,7 +48,11 @@ class OptimizationSpeedups:
     self.pipelining_speedups = [pipelining_speedups[i] * speedup for i, speedup in enumerate(uniform_tiling_speedups)]
 
 class FrameworkSpeedups:
-  def __init__(self, xgboost_speedups, treelite_speedups):
+  def __init__(self, xgboost_times, treelite_times, xgboost_speedups, treelite_speedups):
+    # the times are in seconds. Convert to microseconds
+    self.xgboost_times = [time * 10 ** 6 for time in xgboost_times]
+    self.treelite_times = [time * 10 ** 6 for time in treelite_times]
+
     self.xgboost_speedups = xgboost_speedups
     self.treelite_speedups = treelite_speedups
 
@@ -57,53 +63,71 @@ class FrameworkSpeedups:
       return "XGBoost Speedup: " + self.mean_speedup_over_xgboost + ", Treelite speedup: " + self.mean_speedup_over_treelite
 # print (data_df)
 
-
-def PlotSpeedups(batch_sizes, title, plot_xlabel, plot_ylabel, plot_name, **speedups_data):
-  maxSpeedup = 0
-  minSpeedup = float('INF')
-  for speed_up, name, marker in zip(speedups_data['speedups'], speedups_data['names'], speedups_data['markers']):
-    maxSpeedup = max(maxSpeedup, max(speed_up))
-    minSpeedup = min(minSpeedup, min(speed_up))
-    plt.plot(batch_sizes, np.array(speed_up), marker=marker, label=name)
-
-  plt.xlabel(plot_xlabel, fontsize=12)
[…]

### 2022-04-28 src/include/TiledTree.h
Upgrade MLIR. OMP not tested

@@ -344,11 +344,6 @@ TiledTree* DecisionTree<ThresholdType, ReturnType, FeatureIndexType, NodeIndexTy
   return m_tiledTree.get();
 }
 
-template <typename ThresholdType, typename ReturnType, typename FeatureIndexType, typename NodeIndexType>
-DecisionTree<ThresholdType, ReturnType, FeatureIndexType, NodeIndexType>::~DecisionTree() {
-  // delete m_tiledTree;
-}
-
 } // decisionforest
 } // mlir
 #endif // _TILEDTREE_H_

### 2022-04-28 src/json/modeljsonparser.h

@@ -15,7 +15,8 @@
 #include "mlir/IR/BuiltinTypes.h"
 #include "mlir/IR/MLIRContext.h"
 #include "mlir/IR/Verifier.h"
-#include "mlir/Dialect/StandardOps/IR/Ops.h"
+#include "mlir/Dialect/Arithmetic/IR/Arithmetic.h"
+#include "mlir/Dialect/Func/IR/FuncOps.h"
 #include "llvm/ADT/STLExtras.h"
 
 using json = nlohmann::json;
@@ -135,12 +136,12 @@ protected:
         auto resultType = GetFunctionResultType();
         return m_builder.getFunctionType({argType, resultType}, resultType);
     }
-    mlir::FuncOp GetFunctionPrototype() {
+    mlir::func::FuncOp GetFunctionPrototype() {
         auto location = m_builder.getUnknownLoc();
         auto functionType = GetFunctionType();
         // TODO the function name needs to be an input or derived from the input
         // return mlir::FuncOp::create(location, std::string("Prediction_Function"), functionType);
-        return m_builder.create<mlir::FuncOp>(location, std::string("Prediction_Function"), functionType, m_builder.getStringAttr("public"));
+        return m_builder.create<mlir::func::FuncOp>(location, std::string("Prediction_Function"), functionType, m_builder.getStringAttr("public"));
     }
     virtual mlir::decisionforest::TreeEnsembleType GetEnsembleType() {
         // All trees have the default tiling to start with.
@@ -204,7 +205,7 @@ public:
         if (m_statsProfileCSV != "")
             TreeBeard::Profile::ReadProbabilityProfile(*m_forest, m_statsProfileCSV);
         
-        mlir::FuncOp function(GetFunctionPrototype());
+        mlir::func::FuncOp function(GetFunctionPrototype());
         if (!function)
             return nullptr;
 
@@ -227,7 +228,7 @@ public:
             static_cast<mlir::Value>(entryBlock.getArguments()[0]),
             entryBlock.getArguments()[1], scheduleAttribute);
 
-        m_builder.create<mlir::ReturnOp>(m_builder.getUnknownLoc(), static_cast<mlir::Value>(predictOp));
+        m_builder.create<mlir::func::ReturnOp>(m_builder.getUnknownLoc(), static_cast<mlir::Value>(predictOp));
         if (failed(mlir::verify(m_module))) {
             m_module.emitError("Module verification error");
             return nullptr;

### 2022-04-28 src/mlir/CodeGenStateMachine.cpp

@@ -29,20 +29,20 @@ namespace decisionforest
         }
 
         auto leftShift = rewriter.create<arith::ShLIOp>(location, i32VectorType, comparisonExtended, shiftVector);
-        auto kind = rewriter.getStringAttr("add");
-        auto sum = rewriter.create<vector::ReductionOp>(location, rewriter.getI32Type(), kind, static_cast<Value>(leftShift), ValueRange{ });
+        auto kind = vector::CombiningKind::ADD;
+        auto sum = rewriter.create<vector::ReductionOp>(location, kind, static_cast<Value>(leftShift));
         auto index = rewriter.create<arith::IndexCastOp>(location, rewriter.getIndexType(), static_cast<Value>(sum));
         return index;
     }
 
     Value ReduceComparisonResultVectorToInt_Bitcast(Value comparisonResult, int32_t tileSize, ConversionPatternRewriter &rewriter, Location location) {
-    auto bitcastVectorType = VectorType::get(1, rewriter.getIntegerType(tileSize));
-    auto bitcastOp = rewriter.create<vector::BitCastOp>(location, bitcastVectorType, comparisonResult);
-    auto zeroConst = rewriter.create<arith::ConstantIntOp>(location, int64_t(0), rewriter.getI32Type());
-    auto integerResult = rewriter.create<vector::ExtractElementOp>(location, static_cast<Value>(bitcastOp), static_cast<Value>(zeroConst));
-    auto zeroExtend = rewriter.create<arith::ExtUIOp>(location, integerResult, rewriter.getI64Type()); 
-    auto index = rewriter.create<arith::IndexCastOp>(location, rewriter.getIndexType(), static_cast<Value>(zeroExtend));
-    return index;
+      auto bitcastVectorType = VectorType::get(1, rewriter.getIntegerType(tileSize));
+      auto bitcastOp = rewriter.create<vector::BitCastOp>(location, bitcastVectorType, comparisonResult);
+      auto zeroConst = rewriter.create<arith::ConstantIntOp>(location, int64_t(0), rewriter.getI32Type());
+      auto integerResult = rewriter.create<vector::ExtractElementOp>(location, static_cast<Value>(bitcastOp), static_cast<Value>(zeroConst));
+      auto zeroExtend = rewriter.create<arith::ExtUIOp>(location, rewriter.getI64Type(), integerResult); 
[…]

### 2022-04-28 src/mlir/ConvertNodeTypeToIndexType.cpp

@@ -5,7 +5,8 @@
 #include "mlir/Dialect/Affine/IR/AffineOps.h"
 #include "mlir/Dialect/MemRef/IR/MemRef.h"
 #include "mlir/Dialect/Tensor/IR/Tensor.h"
-#include "mlir/Dialect/StandardOps/IR/Ops.h"
+#include "mlir/Dialect/Arithmetic/IR/Arithmetic.h"
+#include "mlir/Dialect/Func/IR/FuncOps.h"
 #include "mlir/Dialect/SCF/SCF.h"
 #include "mlir/Dialect/Math/IR/Math.h"
 
@@ -80,10 +81,10 @@ struct IndexToNodeOpLowering : public ConversionPattern {
   }
 };
 
-struct ConvertNodeTypeToIndexTypePass : public PassWrapper<ConvertNodeTypeToIndexTypePass, FunctionPass> {
+struct ConvertNodeTypeToIndexTypePass : public PassWrapper<ConvertNodeTypeToIndexTypePass, OperationPass<mlir::func::FuncOp>> {
   
   void getDependentDialects(DialectRegistry &registry) const override {
-    registry.insert<AffineDialect, memref::MemRefDialect, tensor::TensorDialect, StandardOpsDialect, scf::SCFDialect>();
+    registry.insert<AffineDialect, memref::MemRefDialect, tensor::TensorDialect, scf::SCFDialect>();
   }
   
   void VisitOp(Operation* op, OpBuilder& builder) {
@@ -98,7 +99,7 @@ struct ConvertNodeTypeToIndexTypePass : public PassWrapper<ConvertNodeTypeToInde
             }
 
             for (auto index : indexArgInsertionIndices){
-                auto newArgument = block.insertArgument(index, builder.getIndexType());
+                auto newArgument = block.insertArgument(index, builder.getIndexType(), op->getLoc());
                 auto oldArgument = block.getArgument(index+1);
                 assert(oldArgument.getType().isa<decisionforest::NodeType>());
                 oldArgument.replaceAllUsesWith(newArgument);
@@ -132,15 +133,16 @@ struct ConvertNodeTypeToIndexTypePass : public PassWrapper<ConvertNodeTypeToInde
 
   void AddIndexArgumentsForAllBlocks() {
       OpBuilder builder(&getContext());
-      auto func = this->getFunction();
+      auto func = this->getOperation();
       VisitOp(func, builder);
   }
 
-  void runOnFunction() final {
+  void runOnOperation() final {
[…]

### 2022-04-28 src/mlir/ExecutionHelpers.cpp

@@ -342,7 +342,8 @@ llvm::Expected<std::unique_ptr<mlir::ExecutionEngine>> InferenceRunner::CreateEx
 #endif
   // Create an MLIR execution engine. The execution engine eagerly JIT-compiles
   // the module.
-  auto maybeEngine = mlir::ExecutionEngine::create(module, /*llvmModuleBuilder=*/nullptr, optPipeline, llvm::None, executionEngineLibs);
+  mlir::ExecutionEngineOptions options{nullptr, {}, llvm::None, executionEngineLibs};
+  auto maybeEngine = mlir::ExecutionEngine::create(module, options);
   assert(maybeEngine && "failed to construct an execution engine");
   return maybeEngine;
 }

### 2022-04-28 src/mlir/LowerDebugHelpers.cpp

@@ -2,9 +2,8 @@
 
 #include "mlir/Conversion/LLVMCommon/ConversionTarget.h"
 #include "mlir/Conversion/LLVMCommon/TypeConverter.h"
-#include "mlir/Conversion/StandardToLLVM/ConvertStandardToLLVM.h"
-#include "mlir/Conversion/StandardToLLVM/ConvertStandardToLLVMPass.h"
-#include "mlir/Dialect/StandardOps/IR/Ops.h"
+#include "mlir/Dialect/Arithmetic/IR/Arithmetic.h"
+#include "mlir/Dialect/Func/IR/FuncOps.h"
 #include "mlir/Dialect/MemRef/IR/MemRef.h"
 #include "mlir/Dialect/LLVMIR/LLVMDialect.h"
 #include "mlir/Pass/Pass.h"
@@ -61,7 +60,7 @@ struct PrintTreePredictionOpLowering: public ConversionPattern {
       auto sextOp = rewriter.create<LLVM::SExtOp>(op->getLoc(), rewriter.getI64Type(), treeIndex);
       treeIndex = static_cast<Value>(sextOp);
     }
-    rewriter.create<CallOp>(op->getLoc(), printFunctionRef, rewriter.getI64Type(), ArrayRef<Value>({ prediction, treeIndex }));
+    rewriter.create<func::CallOp>(op->getLoc(), printFunctionRef, rewriter.getI64Type(), ArrayRef<Value>({ prediction, treeIndex }));
     rewriter.eraseOp(op);
     return mlir::success();
   }
@@ -94,7 +93,7 @@ struct PrintTreeNodeOpLowering: public ConversionPattern {
       auto sextOp = rewriter.create<LLVM::SExtOp>(op->getLoc(), rewriter.getI64Type(), nodeIndex);
       nodeIndex = static_cast<Value>(sextOp);
     }
-    rewriter.create<CallOp>(op->getLoc(), printFunctionRef, rewriter.getI64Type(), ArrayRef<Value>({ nodeIndex }));
+    rewriter.create<func::CallOp>(op->getLoc(), printFunctionRef, rewriter.getI64Type(), ArrayRef<Value>({ nodeIndex }));
     rewriter.eraseOp(op);
     return mlir::success();
   }
@@ -138,7 +137,7 @@ struct PrintComparisonOpLowering: public ConversionPattern {
[…]

### 2022-04-28 src/mlir/LowerEnsembleToMemrefs.cpp

@@ -6,7 +6,8 @@
 #include "mlir/Dialect/Affine/IR/AffineOps.h"
 #include "mlir/Dialect/MemRef/IR/MemRef.h"
 #include "mlir/Dialect/Arithmetic/IR/Arithmetic.h"
-#include "mlir/Dialect/StandardOps/IR/Ops.h"
+#include "mlir/Dialect/Arithmetic/IR/Arithmetic.h"
+#include "mlir/Dialect/Func/IR/FuncOps.h"
 #include "mlir/Dialect/SCF/SCF.h"
 #include "mlir/Dialect/Math/IR/Math.h"
 
@@ -275,13 +276,13 @@ struct EnsembleConstantOpLowering: public ConversionPattern {
     SaveAndRestoreInsertionPoint saveAndRestoreEntryPoint(rewriter);
     auto getMemrefFuncType = rewriter.getFunctionType(TypeRange({}), memrefType);
     std::string funcName = "Get_" + globalName;
-    NamedAttribute visibilityAttribute{module.sym_visibilityAttrName(), rewriter.getStringAttr("public")};
-    auto getGlobalMemrefFunc = FuncOp::create(location, funcName, getMemrefFuncType, ArrayRef<NamedAttribute>(visibilityAttribute));
+    NamedAttribute visibilityAttribute{module.getSymVisibilityAttrName(), rewriter.getStringAttr("public")};
+    auto getGlobalMemrefFunc = mlir::func::FuncOp::create(location, funcName, getMemrefFuncType, ArrayRef<NamedAttribute>(visibilityAttribute));
     auto &entryBlock = *getGlobalMemrefFunc.addEntryBlock();
     rewriter.setInsertionPointToStart(&entryBlock);
 
     auto getGlobalOffsets = rewriter.create<memref::GetGlobalOp>(location, memrefType, globalName);
-    rewriter.create<mlir::ReturnOp>(location, static_cast<Value>(getGlobalOffsets));
+    rewriter.create<mlir::func::ReturnOp>(location, static_cast<Value>(getGlobalOffsets));
 
     module.push_back(getGlobalMemrefFunc);
   }
@@ -297,8 +298,8 @@ struct EnsembleConstantOpLowering: public ConversionPattern {
     auto tileShapeIDArgType = MemRefType::get(memrefType.getShape(), modelMemrefElementType.getTileShapeType());
[…]

### 2022-04-28 src/mlir/LowerToLLVM.cpp

@@ -5,10 +5,11 @@
 #include "mlir/Conversion/LLVMCommon/ConversionTarget.h"
 #include "mlir/Conversion/LLVMCommon/TypeConverter.h"
 #include "mlir/Conversion/MemRefToLLVM/MemRefToLLVM.h"
-#include "mlir/Conversion/SCFToStandard/SCFToStandard.h"
+#include "mlir/Conversion/ControlFlowToLLVM/ControlFlowToLLVM.h"
+#include "mlir/Conversion/FuncToLLVM/ConvertFuncToLLVM.h"
+#include "mlir/Conversion/FuncToLLVM/ConvertFuncToLLVMPass.h"
+#include "mlir/Conversion/SCFToControlFlow/SCFToControlFlow.h"
 #include "mlir/Conversion/SCFToOpenMP/SCFToOpenMP.h"
-#include "mlir/Conversion/StandardToLLVM/ConvertStandardToLLVM.h"
-#include "mlir/Conversion/StandardToLLVM/ConvertStandardToLLVMPass.h"
 #include "mlir/Conversion/VectorToLLVM/ConvertVectorToLLVM.h"
 #include "mlir/Conversion/ArithmeticToLLVM/ArithmeticToLLVM.h"
 #include "mlir/Conversion/OpenMPToLLVM/ConvertOpenMPToLLVM.h"
@@ -20,7 +21,8 @@
 #include "mlir/Dialect/Arithmetic/IR/Arithmetic.h"
 #include "mlir/Dialect/OpenMP/OpenMPDialect.h"
 
-#include "mlir/Dialect/StandardOps/IR/Ops.h"
+#include "mlir/Dialect/Arithmetic/IR/Arithmetic.h"
+#include "mlir/Dialect/Func/IR/FuncOps.h"
 #include "mlir/Pass/Pass.h"
 #include "mlir/Pass/PassManager.h"
 #include "mlir/Transforms/DialectConversion.h"
@@ -322,7 +324,7 @@ struct GetModelMemrefSizeOpLowering : public ConversionPattern {
 struct DecisionForestToLLVMLoweringPass : public PassWrapper<DecisionForestToLLVMLoweringPass, OperationPass<ModuleOp>> {
   void getDependentDialects(DialectRegistry &registry) const override {
     registry.insert<LLVM::LLVMDialect, scf::SCFDialect, AffineDialect, memref::MemRefDialect, 
[…]

### 2022-04-28 src/mlir/LowerToMidLevelIR.cpp

@@ -4,7 +4,7 @@
 #include "mlir/Dialect/Affine/IR/AffineOps.h"
 #include "mlir/Dialect/MemRef/IR/MemRef.h"
 #include "mlir/Dialect/Arithmetic/IR/Arithmetic.h"
-#include "mlir/Dialect/StandardOps/IR/Ops.h"
+#include "mlir/Dialect/Func/IR/FuncOps.h"
 #include "mlir/Dialect/SCF/SCF.h"
 #include "mlir/Transforms/DialectConversion.h"
 #include "mlir/Dialect/Math/IR/Math.h"
@@ -44,11 +44,11 @@ void InsertPrintMemrefOp(
     auto index = rewriter.create<arith::ConstantIndexOp>(location, i);
     auto ithValue = rewriter.create<memref::LoadOp>(location, elementType, memref, static_cast<Value>(index));
     if (kind == 0)  { // Integer
-      auto i64Value = rewriter.create<arith::IndexCastOp>(location, ithValue, rewriter.getI64Type());
+      auto i64Value = rewriter.create<arith::ExtSIOp>(location, rewriter.getI64Type(), static_cast<Value>(ithValue));
       vectorValues.push_back(i64Value);
     }
     else {
-      auto doubleValue = rewriter.create<arith::ExtFOp>(location, ithValue, rewriter.getF64Type());
+      auto doubleValue = rewriter.create<arith::ExtFOp>(location, rewriter.getF64Type(), static_cast<Value>(ithValue));
       vectorValues.push_back(doubleValue);
     }
   }
@@ -375,9 +375,9 @@ struct PredictForestOpLowering: public ConversionPattern {
         auto predTransform = forestAttribute.GetDecisionForest().GetPredictionTransformation();
         Value transformedValue;
         if (predTransform == decisionforest::PredictionTransformation::kIdentity)
-          transformedValue = static_cast<Value>(treeLoop.results()[0]);
+          transformedValue = static_cast<Value>(treeLoop.getResults()[0]);
         else if (predTransform == decisionforest::PredictionTransformation::kSigmoid) {
-          transformedValue = GenSigmoid(rewriter, static_cast<Value>(treeLoop.results()[0]), location);
+          transformedValue = GenSigmoid(rewriter, static_cast<Value>(treeLoop.getResults()[0]), location);
         }
         else if (predTransform == decisionforest::PredictionTransformation::kSoftMax) {
[…]

### 2022-04-28 src/mlir/MemrefTypes.h

@@ -11,7 +11,7 @@
 #include "mlir/IR/DialectImplementation.h"
 #include "mlir/Dialect/MemRef/IR/MemRef.h"
 #include "mlir/Interfaces/DataLayoutInterfaces.h"
-#include "mlir/Dialect/Vector/VectorOps.h"
+#include "mlir/Dialect/Vector/IR/VectorOps.h"
 
 namespace mlir
 {

### 2022-04-28 src/test/XGBoostBenchmarks.cpp

@@ -14,7 +14,8 @@
 #include "mlir/IR/BuiltinTypes.h"
 #include "mlir/IR/MLIRContext.h"
 #include "mlir/IR/Verifier.h"
-#include "mlir/Dialect/StandardOps/IR/Ops.h"
+#include "mlir/Dialect/Arithmetic/IR/Arithmetic.h"
+#include "mlir/Dialect/Func/IR/FuncOps.h"
 #include "mlir/Dialect/SCF/SCF.h"
 #include "llvm/ADT/STLExtras.h"
 

### 2022-04-28 src/test/XGBoostTests.cpp

@@ -9,7 +9,8 @@
 #include "mlir/IR/BuiltinTypes.h"
 #include "mlir/IR/MLIRContext.h"
 #include "mlir/IR/Verifier.h"
-#include "mlir/Dialect/StandardOps/IR/Ops.h"
+#include "mlir/Dialect/Arithmetic/IR/Arithmetic.h"
+#include "mlir/Dialect/Func/IR/FuncOps.h"
 #include "mlir/Dialect/SCF/SCF.h"
 #include "llvm/ADT/STLExtras.h"
 

### 2024-03-09 docs/SOSP 2024/overview.tex
Add latex version of paper

\section{Compiler Overview}
\label{Sec:Overview}
\Treebeard{} takes a serialized decision tree ensemble as input (For example
XGBoost JSON, ONNX etc.) and generates an optimized inference function. 
\Treebeard{} automatically generates an optimized inference function from 
the serialized model and can either target CPUs or GPUs. 
Figure \ref{Fig:CompilerStructure} shows the structure of the \Treebeard{} compiler. 
The inference computation is lowered through three intermediate representations
-- high-level IR (HIR), mid-level IR (MIR) and low-level IR (LIR). The LIR is
finally lowered to LLVM and then JIT'ed to the specified target processor.

\begin{figure}[htb]
  \centering
  \includegraphics[width=\linewidth]{figures/compiler.png}
  \caption{\Treebeard{} compiler structure.}
  \label{Fig:CompilerStructure}
\end{figure}

\begin{figure*}[htb]
  \centering
  \includegraphics[width=\linewidth]{figures/OverviewExample_New.png}
  \vskip 10pt
  \caption{\Treebeard{} IR lowering and optimization details: the three abstraction levels in \Treebeard{}'s IR are shown. The
           high level IR is a tree-based IR to perform model level optimization, the mid-level IR is for
           loop optimizations that are independent of memory layout and the low level IR allows us to perform
           vectorization and other memory layout dependent optimizations.}
  \label{Fig:LoweringExample}
\end{figure*}

In HIR, the model is represented as a collection of binary trees. This abstraction
allows the implementation of optimizations that require the manipulation of the model
or its constituent trees. In Figure \ref{Fig:LoweringExample}, \circled{A}
shows this representation for a model with three trees. 
In HIR, \Treebeard{} tiles tree nodes together to convert
the binary tree to an n-ary tree as shown in \circled{B}. Trees are also 
reordered to enable better code generation and padded to allow more 
efficient traversal as shown in \circled{C}. While these are the optimizations currently 
implemented in \Treebeard{}, these are by no means the only ones that 
are enabled by HIR. It is not difficult to imagine optimizations 
such as tree pruning for specified accuracy levels for example.
These optimizations and rewrites that are performed on the domain-specific 
HIR would have been much on a traditional loop based IR or even in 
other IRs within \Treebeard{}.

After these model-level optimizations are performed on the HIR, the 
code is lowered to the mid-level IR (MIR) as dictated by a user specified schedule
(\circled{C} to \bluecircled{D}) in Figure \ref{Fig:LoweringExample}). The schedule specifies how
the iteration space that goes over the trees and input rows is to be traversed. 
It specifies how the iteration space is to be tiled, which loops are to be
parallelized, which loops are to be mapped to GPU grid and block dimensions etc.
(Details in Section \ref{Sec:SchedulingLang}). MIR is a loop-based IR that 
explicitly encodes details of the iteration space has to be traversed. However, 
it still abstracts details about the in-memory representation of the model. 
Optimizations such as tree-walk unrolling and interleaving are performed 
on the MIR (\bluecircled{E}). Subsequently, reduction operations are split and rewritten 
to correctly and explicitly implement reduction in the presence of 
parallel loops (\bluecircled{F}). More details of this process that we call 
\emph{legalization} are in Section \ref{Sec:Reduction}. Another important 
point to note is that MIR is independent of the target processor and therefore
all optimizations on MIR can be reused across CPU and GPU compilation.

The MIR is then further lowered to a low-level IR (LIR). This is the
level at which the compilation pipeline diverges for CPUs and GPUs. In 
the GPU compilation pipeline, the required memory transfers and kernel
invocations are inserted into the LIR (\circled{J}). Additionally, buffers 
to hold model values are inserted and tree operations are lowered to
explicitly refer to these buffers. This lowering is controlled by 
a plugin mechanism where different in-memory representations can 
be added to the compiler by implementing an interface. These plugins
provide information required for the lowering of MIR to LIR as well as 
the lowering to LLVM IR as shown in Figure \ref{Fig:LoweringExample}.
Again, a significant amount of code is shared between the
CPU and GPU pipelines for representations that are common between them
(Array and Sparse representations). Vectorization of tree traversals 
is also explicitly represented in LIR.

To reiterate, the following are the salient points of \Treebeard{}'s design.
\begin{enumerate}
  \item The compiler uses three intermediate representations (IRs) to
  represent the inference computation at different levels of abstraction. This
  allows different optimizations to be performed and also allows us to share
  infrastructure between compilation pipelines for different target processors.

  \item The specification of how the inference computation is to be lowered to 
  loops is not encoded directly in the compiler. Instead, this 
  is specified as an input to the compiler using a scheduling language that is 
  specialized for decision tree inference computations (Section \ref{Sec:SchedulingLang}).
  This separation allows us to build optimizations and schedule exploration 
  mechanisms independent of the core compiler (Section \ref{AutoTuneHeuristic}). 
  The scheduling language also exposes details like parallelization, optimization
  of reductions (vectorization, use atomics, shared memory on GPUs etc.).

  \item \Treebeard{} has been designed to keep the optimization passes and code 
  generator independent of the in-memory representation finally used for the model. To achieve 
  this, \Treebeard{} specifies an interface to implement that provides the necessary 
  capabilities to the code generator as a plugin. This interface abstracts several details
  on how model values are stored. In particular, it abstracts the actual layout of 
  the memory buffers, how to load model values like thresholds and feature indices, how 
  to move from a node to its children and determining whether a node is a leaf.
  This design allows us to write each memory representation as a standalone plugin 
  and reuse the rest of the compiler infrastructure.
\end{enumerate}

### 2024-03-09 docs/SOSP 2024/schedule.tex
Add latex version of paper

\section{Scheduling Language}
The goal of the scheduling language is to express the following
* The order in which the iteration space over the batch of inputs (batch) and trees in the forest (tree) is to be traversed. Note that there is a reduction over the tree dimension.
* Intra or inter tree optimizations that are to be performed on a tree or set of trees (tree walk unrolling, pipelining, SIMDize etc).

The reasons to use a scheduling language rather than a hard-coded lowering are as follows
* Making the scheduling specification external to the compiler allows us to 
more easily build auto-schedulers and auto-tuners.
* It is very hard to come up with a template loop nest that works for all models (for example, tree sizes may vary across the model making it necessary to iterate different number of trees at different times). 
* A scheduling language will make writing newer locality optimizations faster since no changes to the compiler infrastructure will be needed.
* Adding support for additional hardware targets (GPUs, FPGAs), will be much easier with a scheduling language.

There are some simplifying assumptions and limitations in the current design
* Tree traversals are considered atomic. There is no way to express partial tree traversals or schedule individual node/level computations. 
* Accumulation of tree predictions is done immediately (as opposed to, for example, collecting all predictions and performing a reduction later).

\subsection{Language Definition}
We broadly have three classes of directives in the language. The first is a set of loop nest modifiers that are used to specify the structure of the loop nest to walk the iteration space. The second is a set of clauses that specify intra and inter tree optimizations. Finally, we have a class of attributes that control how reductions are performed.

\subsubsection{Loop Modifiers}
There are two special index variables -- \op{batch} and \op{tree}. The clauses modify these index variables or 
index variables derived from these (through the application of clauses).
\begin{itemize}
  \item \textbf{tile}: Tile the passed index variable using a fixed tile size.
  \item \textbf{split}: Split the range of the passed index variable into two parts. The range of the first part is specified
  by an argument.
  \item \textbf{unroll}: Unroll an index completely
  \item \textbf{reorder}: Reorder the specified indices. The specified indices must be successive indices in the current loop nest.
  \item \textbf{specialize}: Generate separate code for each iteration of the specified index variable. This is useful 
  while parallelizing across trees and these trees have different depths.
  \item \textbf{gpuDimension}: Maps the specified index variable to represent a dimension in either the GPU kernel grid or thread block.
\end{itemize}

The default loop order (as currently generated by the compiler) is (batch, tree), i.e, for each row in the input batch, go over all trees.

The following are examples of how the loop modifiers can be used.
\begin{itemize}
  \item The loop order used by XGBoost\cite{XGBoost} is (tree, batch) -- walk one tree for all inputs in the batch before 
  moving to the next tree. The corresponding schedule would be
\begin{lstlisting}[style=c++]
  reorder(tree, batch)
\end{lstlisting}

  \item The below schedule computes 2 trees at a time over the whole batch.
\begin{lstlisting}[style=c++]
  tile(tree, t0, t1, 2)
  reorder(t0, batch, t1)
\end{lstlisting}

  \item If we additionally only want to compute over 4 input rows (rather than the whole batch) for
  every 2 tree, and then move onto the next 2 trees for the same set of inputs, then the schedule is as follows. 
\begin{lstlisting}[style=c++]
  tile(batch, b0, b1, 4) 
  tile(tree, t0, t1, 2)
  reorder(b0, t0, b1, t1) 
\end{lstlisting}

\end{itemize}

\subsubsection{Optimizations}
The following clauses provide ways to optimize the inference routine being generated.
\begin{itemize}
  \item \textbf{cache}: Cache the working set of one iteration of the specified loop corresponding
  to this index. This can be specified on either batch or tree loops. Specifying it 
  on a batch loop leads to all rows accessed in a single iteration of the loop 
  being cached. Similarly, specifying it on a tree loop leads to all trees accessed in 
  one iteration of that loop being cached.
  \item \textbf{parallel}: Execute the loop corresponding to this index in parallel.
  \item \textbf{interleave}: Interleave the execution of the tree walks within the current index (must be applied on an inner most index).
  \item \textbf{unrollWalk}: Unroll tree walks at the current index. 
  \item \textbf{peelWalk}: Peel the first n steps of the specified tree walk and don't check for leaves for that number of steps.
\end{itemize}

\subsubsection{Reduction Optimization}
\begin{itemize}
  \item \textbf{atomicReduce}: Use atomic memory operations to accumulate values across 
  parallel iterations of the specified loop. 
  \item \textbf{sharedReduce}: Only applies to GPU compilation. Specifies that intermediate
  results are to be stored in shared memory.
  \item \textbf{vectorReduce}: Use vector instructions with the specified vector width 
  to reduce intermediate values across parallel iterations of the specified loop.
\end{itemize}

\TODO{Add examples for RAPIDS, Tahoe (Maybe show some strategies can be encoded?)}

### 2024-03-11 docs/SOSP 2024/schedule.tex
Add xgb and Tahoe schedules

@@
 The goal of the scheduling language is to express the following
-* The order in which the iteration space over the batch of inputs (batch) and trees in the forest (tree) is to be traversed. Note that there is a reduction over the tree dimension.
-* Intra or inter tree optimizations that are to be performed on a tree or set of trees (tree walk unrolling, pipelining, SIMDize etc).
+\begin{itemize}
+  \item The order in which the iteration space over the batch of inputs (batch) and trees in the forest (tree) is to be traversed. Note that there is a reduction over the tree dimension.
+  \item Intra or inter tree optimizations that are to be performed on a tree or set of trees (tree walk unrolling, pipelining, SIMDize etc).
+\end{itemize}
 
 The reasons to use a scheduling language rather than a hard-coded lowering are as follows
-* Making the scheduling specification external to the compiler allows us to 
+\begin{itemize}
+  \item Making the scheduling specification external to the compiler allows us to 
   more easily build auto-schedulers and auto-tuners.
-* It is very hard to come up with a template loop nest that works for all models (for example, tree sizes may vary across the model making it necessary to iterate different number of trees at different times).
-* A scheduling language will make writing newer locality optimizations faster since no changes to the compiler infrastructure will be needed.
-* Adding support for additional hardware targets (GPUs, FPGAs), will be much easier with a scheduling language.
+  \item It is very hard to come up with a template loop nest that works for all models 
+  (for example, tree sizes may vary across the model making it necessary to iterate 
+  different number of trees at different times).
+  \item A scheduling language will make writing newer locality optimizations faster since 
+  no changes to the compiler infrastructure will be needed.
+  \item Adding support for additional hardware targets (GPUs, FPGAs), will be much easier
+  with a scheduling language.
+\end{itemize}
 
 There are some simplifying assumptions and limitations in the current design
-* Tree traversals are considered atomic. There is no way to express partial tree traversals or schedule individual node/level computations. 
-* Accumulation of tree predictions is done immediately (as opposed to, for example, collecting all predictions and performing a reduction later).
+\begin{itemize}
+  \item Tree traversals are considered atomic. There is no way to express partial
+   tree traversals or schedule individual node/level computations. 
+  \item Accumulation of tree predictions is done immediately (as opposed to, 
+  for example, collecting all predictions and performing a reduction later).
+\end{itemize}
 
 \subsection{Language Definition}
-We broadly have three classes of directives in the language.
-The first is a set of loop nest modifiers that are used to specify the structure of the loop nest to walk the iteration space.
-The second is a set of clauses that specify intra and inter tree optimizations.
-Finally, we have a class of attributes that control how reductions are performed.
+We broadly have three classes of directives in the language. The first is a set 
+of loop nest modifiers that are used to specify the structure of the loop nest to
+walk the iteration space. The second is a set of clauses that specify intra and 
+inter tree optimizations. Finally, we have a class of attributes that control how 
+reductions are performed.
 
@@
 \TODO{Add examples for RAPIDS, Tahoe (Maybe show some strategies can be encoded?)}
+
+\subsection{The XBoost Schedule}
+XGBoost\cite{XGBoost} is a very popular gradient boosting library. 
+It implements inference on the CPU by going over a fixed number of rows (64 in the previous version)
+for every tree and then moving to the next tree. When all trees have been walked for this 
+set of rows, the next set of rows is taken up. Also, different sets of rows are 
+processed in parallel.
+
+The schedule used by XGBoost can be represented in \Treebeard{}'s scheduling 
+language as follows.
+\begin{lstlisting}[style=c++]
+  tile(batch, b0, b1, CHUNK_SIZE)
+  reorder(b0, tree, b1)
+  parallel(b0)
+\end{lstlisting}
+
+\subsection{Tahoe Schedules}
+Tahoe\cite{Tahoe} has four strategies that it picks from for a given model. 
+Each of these strategies can be encoded using \Treebeard{}'s scheduling language
+as we show below. 
+\begin{itemize}
+  \item \textbf{Direct Method}: In this strategy, a single GPU thread walks all trees
+  for a given input row. The schedule for this strategy is as follows.
+\begin{lstlisting}[style=c++]
+  tile(batch, b0, b1, ROWS_PER_TB)
+  reorder(b0, b1, tree)
+  gpuDimension(b0, grid.x)
+  gpuDimension(b1, block.x)
+\end{lstlisting}
+  Here, \op{ROWS\_PER\_TB} is the number of rows that are processed by a single thread block.
+  \item \textbf{Shared Data}: In this strategy, a thread block walks all the trees 
+  for a given row in parallel. If threads walk multiple trees, each thread accumulates
+  partial results. Finally, a thread block wide reduction is performed to compute 
+  the prediction. The schedule for this strategy is as follows.
+\begin{lstlisting}[style=c++]
+  reorder(batch, tree)
+  gpuDimension(batch, grid.x)
+  gpuDimension(tree, block.x)
+  cache(batch)
+\end{lstlisting}
+  \item \textbf{Shared Forest}: In this strategy, the whole model is loaded into 
+  shared memory and subsequently, a single thread walks all trees for a particular
+  row. The schedule for this strategy is as follows.
+\begin{lstlisting}[style=c++]
+  tile(batch, b0, b1, ROWS_PER_TB)
+  tile(tree, t0, t1, N_TREES)
+  reorder(b0, b1, t0, t1)
+  cache(t0)
+  gpuDimension(b0, grid.x)
+  gpuDimension(b1, block.x)
+\end{lstlisting}
+  Here, we create a placeholder single iteration loop \op{t0} so that we can 
+  specify that all trees are to be cached.
+  \item \textbf{Shared Partial Forest}: In case the model is too large to fit into
+  shared memory, the model is split into chunks and each chunk is loaded into shared
+  memory. Again, as in the previous strategy, one thread walks all trees assigned to
+  a thread block for a row. The schedule for this strategy is as follows.
+\begin{lstlisting}[style=c++]
+  tile(batch, b0, b1, ROWS_PER_TB)
+
+  tile(tree, t0, t0Inner, TREES_PER_TB)
+  tile(t0Inner, t1, t2, TREES_PER_TB)
+  cache(t1)
+  reorder(b0, t0, b1, t1, t2)
+
+  gpuDimension(b0, grid.x)
+  gpuDimension(t0, grid.y)
+  gpuDimension(b1, block.x)
+\end{lstlisting}
+\end{itemize}

### 2024-03-13 src/main.cpp
Add code to generate random benchmarks

@@ -13,7 +13,8 @@
 namespace TreeBeard {
 namespace test {
 void TestTileStringGen();
-}
+void generateRandomXGBoostModels(const std::string &dirName);
+} // namespace test
 } // namespace TreeBeard
 
 bool EqualsString(char *arg, const std::string &str) {
@@ -369,6 +370,22 @@ bool ComputeProbabilityProfileIfNeeded(int argc, char *argv[]) {
   return true;
 }
 
+bool GenerateRandomXGBoostBenchmarksIfNeeded(int argc, char *argv[]) {
+  bool generateBenchmarks = false;
+  for (int32_t i = 0; i < argc; ++i)
+    if (std::string(argv[i]).find(std::string("--generateBenchmarks")) !=
+        std::string::npos) {
+      generateBenchmarks = true;
+      break;
+    }
+  if (!generateBenchmarks)
+    return false;
+  std::string dirName = "/home/ashwin/mlir-build/llvm-project/mlir/examples/"
+                        "treebeard/xgb_models/test/GPUBenchmarks";
+  TreeBeard::test::generateRandomXGBoostModels(dirName);
+  return true;
+}
+
 int main(int argc, char *argv[]) {
   SetInsertDebugHelpers(argc, argv);
   SetInsertPrintVectors(argc, argv);
@@ -386,6 +403,8 @@ int main(int argc, char *argv[]) {
   else if (ComputeInferenceStatsIfNeeded(argc, argv))
     return 0;
   else if (ComputeProbabilityProfileIfNeeded(argc, argv))
+    return 0;
+  else if (GenerateRandomXGBoostBenchmarksIfNeeded(argc, argv))
     return 0;
   else {
     std::cout

### 2024-03-13 docs/SOSP 2024/motivation.tex

\section{Motivation}

Our main aim while designing \Treebeard{} was to unify the diverse set of implementation
strategies that have been used in existing systems for decision tree inference. Some 
differences in these systems are as follows:
\begin{itemize}
  \item Decision tree inference is run on several platforms including CPUs and GPUs. The 
  implementations used on each of these platforms are different and the techniques used
  to optimize them are also different.
  \item A diverse set techniques have been proposed for optimization of decision tree 
  inference on CPUs and GPUs \cite{VPred, Tahoe, Treelite, XGBoost, Hummingbird, QuickScorer, FIL}. 
  No system exists that unifies the disparate optimizations implemented in these systems.
  \item A very extensive design space of optimizations exists for decision tree inference
  outside the few that have been proposed in the literature. However, currently no 
  system exists that is capable of exploring this space and identifying the best set 
  of parameters to use for a given model and platform.
  \item Different systems use different in-memory representations for the model. For example,
  XGBoost uses a sparse representation, RAPIDs FIL uses what is called the reorg representation 
  and Tahoe uses a variation of the reorg representation. Currently, systems implement 
  inference kernels that are tied to a single representation of the model. Again, this means
  that no current system can explore different combinations of in-memory representations 
  and optimizations.
\end{itemize}
At high-level, to make \Treebeard{} capable of unifying these differences, we design 1)  
expressive intermediate representations that can represent and compose several proposed 
optimizations 2) a scheduling language that specifies the structure of the
generated code and 3) a plugin mechanism with which different in-memory representations
can be composed with different optimizations. Finally, we develop a heuristic to
explore the extensive optimization space that \Treebeard{}'s design enables.

### 2024-03-13 docs/SOSP 2024/reduction.tex

\section{Reductions : Representation, Optimization and Lowering}

Currently, existing reduction support in MLIR is insufficient 
to code generate and optimize the reductions \Treebeard{} needs
to perform while performing inference (sum up individual 
tree predictions to compute the prediction of the model). 
MLIR only supports reductions of value types and cannot 
directly represent and optimize inplace reductions of 
several elements of a memory buffer.

We design a mechanism to specify accumulating values into
an element of a multi-dimensional array inplace. 
The accumulation is performed inside an arbitrary 
loop nest where several surrounding loops maybe parallel and 
the ultimate target machine maybe a CPU or a GPU. Because this problem 
is of general interest, we design this as an MLIR dialect.

The main abstraction we introduce is the \op{reduce} op. It models 
atomically accumulating values into an element of a 
multi-dimensional array (represented by a MLIR \op{memref}).
The following example sums up the elements of the 1D memref \op{arr}
into the first element of the memref \op{result}. It does this
using in two concurrent iterations of a surrounding 
parallel loop.

\begin{lstlisting}[style=c++]
  builtin.func @ReduceVector(%arr: memref<num_elemsxf64>, %result: memref<1xf64>) -> void {
    par.for i0 in range(0 : num_elems/2 : num_elems) {
      for i1 in range(0 : num_elems/2) 
        reduce(%result, 0, arr[i0 + i1]) <"+", 0.0>
    }
  }
\end{lstlisting}

The semantics of the \op{reduce} op guarantee that all elements are 
correctly added and that there is no race between the parallel 
iterations of the loop.

The \op{reduce} op is defined for all associative and commutative
reduction operations with a well-defined initial value. The 
reduction operator and the initial value are attributes applied
on the \op{reduce} op.

The main differences between our \op{reduce} and the existing 
reductions in MLIR are the following:
1. Existing reductions only support scalar, by value reductions. 
It does not support accumulating inplace into memref elements.
2. Existing reduction support in MLIR does not provide a unified
way to handle reductions in GPUs and CPUs. To the best of our 
knowledge, there is currently no way to model reductions on GPUs
in MLIR.
2. Existing reductions have strict rules about where they can be 
written. For example, \op{scf.reduce} needs to be an immediate 
child \op{scf.parallel}. However, our \op{reduce} op can be
generated anywhere in the loop nest.

Having modeled the reductions with an abstract op, the 
aim now is to lower this to a correct and optimized 
implementation on both CPU and GPU. In order to do this,
we make the following observations. 
1. The \op{reduce} op has a loop carried dependency on itself
and loop carried dependences on other \op{reduce} ops that
accumulate into the same target array. 
A simple lowering to a sequence of load-add-store instructions is 
incorrect if any of these dependences are carried by a parallel 
loop. We call any such surrounding parallel loop a 
\textbf{conflicting loop} (\TODO{Change the name!}) for the reduction.
2. There is a race between the parallel iterations of such a loop 
when naively accumulating values into target memref elements. 
To avoid this race, the result memref can be \textbf{privatized} wrt each 
surrounding conflicting loop. Subsequently, each privatized dimension 
can be reduced at the end of the conflicting loop it was inserted for. 
\TODO{We cannot do better than this in terms of memory usage} \TODO{Need a proof}.

\begin{definition}
A parallel loop surrounding one or more \op{reduce} ops is 
a \textbf{conflicting loop} for a target multi-dimensional array if this 
loop has a non-zero dependence distance for the dependence between
any of the contained \op{reduce} ops.
\end{definition}

In the context of \Treebeard{}, this set of loops is exactly the set 
of surrounding parallel loops that are iterating over trees. The 
results can be privatized for each conflicting loop iteratively 
and reductions along each privatized dimension can be inserted  
immediately following the loop the dimension was inserted due to.

We illustrate this process through the example above. The \op{i0} loop is 
a conflicting loop for the reduction into the \op{result} array.
We would therefore privatize the \op{result} memref for each iteration of 
the \op{i0} loop. 
\begin{lstlisting}[style=c++]
  builtin.func @ReduceVector(%arr: memref<num_elemsxf64>, %result: memref<1xf64>) -> void {
    results_1 = memref<2x1xf64>
    par.for i0 = range(0 : num_elems/2 : num_elems) {
      for i1 = range(0 : num_elems/2) 
        reduce(%result_1[i0/(num_elems/2), 0], arr[i0 + i1]) <"+", 0.0>
    }
    results = reduce_dimension(results_1, 0) <"+", 0.0>
  }
\end{lstlisting}
The op \op{reduce\_dimension} reduces values across the specified
dimension of an n-dimensional memref. In the above example, 
the \op{reduce\_dimension} op is reducing across all elements 
of the first dimension (index 0). Therefore, in this case, it 
produces a result memref with a single element (the first dimension
with size 2 is collapsed).

\begin{definition}
  \op{\textbf{reduce\_dimension(targetMemref, memref, dim, [indices], [rangeStart], [rangeEnd])}}:
   Computes the reduction over the dimension specified by \op{dimension} and stores the 
   result in \op{targetMemref}. \op{[indices]} must be a vector of \op{dim} elements
   (or empty if the dimension being reduced is the first dimension). \op{[rangeStart]} 
   and \op{[rangeEnd]} represent the range of indices following the reduction dimension and 
   must have the same number of elements. If both are \op{null} (not passed), 
   all elements of these dimensions are reduced. The computation performed by the op is as follows.

  $targetMemref[\vec{\boldsymbol{indices}}, \vec{\boldsymbol{k}}] = \sum_{i=0}^{shape[dim]} memref[\vec{\boldsymbol{indices}}, i, \vec{\boldsymbol{k}}]\quad   \forall \vec{\boldsymbol{k}} \in \left[[rangeStart_0, rangeEnd_0), ... , [rangeStart_n, rangeEnd_n)\right]$
\end{definition}

Consider the following code with nested parallel loops. 
(A situation where trees are split across both
threads and thread blocks could result in such generated 
code in \Treebeard{}.)
\begin{lstlisting}[style=c++]
  builtin.func @ReduceVector(%arr: memref<num_elemsxf64>, %result: memref<1xf64>) -> void {
    par.for i0 = range(0 : num_elems/2 : num_elems) {
      par.for i1 = range(0 : num_elems/4 : num_elems/2) {
        for i2 = range(0 : num_elems/4) 
          reduce(%result, 0, arr[i0 + i1 + i2]) <"+", 0.0>
      }
    }
  }
\end{lstlisting}

Here, the \op{i0} and \op{i1} loops are conflicting loops wrt 
the \op{result} memref. We \textbf{legalize} the reduction 
by privatizing the \op{result} array wrt the \op{i0} and \op{i1} loops.
However, there are now two privatized dimensions and therefore, 
two dimensions need to be reduced to compute the final result.
This multi-stage reduction is what enables us to model 
hierarchical reductions.

The following code shows how the reduction above is legalized. 
We introduce a new op, \op{reduce\_dimension\_inplace} which 
reduces a dimension of the input memref and stores results 
in the same array. This helps saves memory by removing the 
need to create multiple intermediate arrays to store results.
Only the final dimension reduction uses the \op{reduce\_dimension} op.

\begin{lstlisting}[style=c++]
  builtin.func @ReduceVector(%arr: memref<num_elemsxf64>, %result: memref<1xf64>) -> void {
    results_1 = memref<2x2x1xf64>
    par.for i0 = range(0 : num_elems/2 : num_elems) {
      par.for i1 = range(0 : num_elems/4 : num_elems/2) {
        for i2 = range(0 : num_elems/4) 
          index0 = i0/(num_elems/2)
          index1 = i1/(num_elems/4)
          reduce(%result_1[index0, index1, 0], arr[i0 + i1 + i2]) <"+", 0.0>
      }
      // result_1[i0/(num_elems/2), 0] = sum(result_1[i0/(num_elems/2), :])
      reduce_dimension_inplace(%result_1, 1, i0/(num_elems/2)) 
    }
    // result = sum(result[:, 0])
    %result = reduce_dimension(%result_1, 0)
  }
\end{lstlisting}

The behavior of the \op{reduce\_dimension\_inplace} op is similar to the 
\op{reduce\_dimension} op except that it updates the input array inplace
rather than writing results to a target array. The definition of the 
op is as follows.

\begin{definition}
 \textbf{\op{reduce\_dimension\_inplace(memref, dim, [indices], [rangeStart], [rangeEnd])}}:
  Computes the reduction over the dimension specified by \op{dimension} and stores the 
  result at index 0 of that dimension. \op{[indices]} must be a vector of \op{dim} elements
   (or empty if the dimension being reduced is the first dimension). \op{[rangeStart]} 
   and \op{[rangeEnd]} must have the same number of elements. If both are \op{null} (not passed), 
   all elements of the corresponding dimension are reduced.

  The computation performed by the op is defined by the following equation.

  $memref[\vec{\boldsymbol{indices}}, 0, \vec{\boldsymbol{k}}] = \sum_{i=0}^{shape[dim]} memref[\vec{\boldsymbol{indices}}, i, \vec{\boldsymbol{k}}]\quad   \forall \vec{\boldsymbol{k}} \in \left[[rangeStart_0, rangeEnd_0), ... , [rangeStart_n, rangeEnd_n)\right]$  
\end{definition}

\subsection{Lowering Reduction Operations}
We implement lowering of the operations defined above to both the CPU and GPU.
Since the compilation pipeline diverges after the reductions are legalized, 
we can implement lowering and optimization of our reduction dialect to CPUs and
GPUs simply using different MLIR rewrite patterns. We now briefly describe 
how these operations are lowered to the CPU and GPU.

\subsubsection{Lowering to CPU}
The lowering of the reduction operations to CPU is fairly straightforward. We lower the 
two operations listed above, \op{reduce\_dimension\_inplace} and 
\op{reduce\_dimension} to a simple loop nest that goes over the specified
subset of the input array, performs the reduction and writes 
the result into the appropriate location of the target array. 
If the schedule specifies that the reduction is to be vectorized,
then as many elements as specified by the vector width are read 
from the input array as a vector, accumulated as a vector, and 
finally written back to the target array. In general, this works 
well because reductions are typically being performed on dimensions
other than the inner-most dimension and therefore, this strategy
loads successive elements from memory maximizing memory bandwidth 
utilization.

\TODO{explain atomic reduction}

\subsubsection{Lowering to GPU}
The lowering on GPU is slightly more involved than the lowering on CPUs.
However, we can lower the same abstractions to efficient implementations
and therefore simplify high-level code generation. The lowering for 
the inplace and non-inplace operations are essentially the same, except 
for the target array and we do not distinguish between them except 
for finally storing the result.

The lowering of the \op{reduce\_dimension\_*} ops is distinct from existing 
work on implementing reductions efficiently on GPUs \cite{NVIDIAReductions}
because our abstractions potentially represent several independent reductions
(independent for different output elements).
Therefore, we can either exploit parallelism across the independent reductions or 
the inherent parallelism in the reduction by performing a divide and conquer 
reduction.

The reduction pass for GPU can follow one of two paths. If the lowering pass
determines that there are enough independent reductions to keep all threads
in a thread block busy, then it simply generates code that performs one (or 
multiple) reductions completely in a thread. If however there are not 
enough independent reductions, then the lowering pass generates a tree 
style reduction where multiple threads cooperate to perform a single reduction
using inter-thread shuffles.

Another feature specific to GPU reductions is the use of shared memory. 
If the schedule specifies that the reduction needs to be performed 
using shared memory, the privatized buffer is allocated in shared memory. 
Also, the compiler ensures that only as much shared memory is allocated 
as needed to hold values processed by a single thread-block and 
index offsets are appropriately rewritten to handle the differences between 
the indexing of the target memref and the shared memory array.
Our abstractions allow our lowering passes to be written completely 
independent of whether we use shared memory and therefore allow 
us to enable or disable shared memory use independently from the other 
parts of the compiler.

\subsection{Use in \Treebeard{}}
We now present an example specific to \Treebeard{}. The schedule with
which code is generated is as below. \op{N\_t} is the number of trees 
and \op{batch\_size} is the batch size. The schedule tiles both the 
batch loop and the tree loop and parallelizes the outer batch 
and tree loops.

\begin{lstlisting}[style=c++]
IndexVar i0, i1, t0, t1;
auto& batch = schedule.GetBatchIndex();
auto& tree = schedule.GetTreeIndex();
schedule.Tile(batch, i0, i1, batch_size/2);
schedule.Tile(tree, t0, t1, N_t/2);
schedule.Reorder({i0, t0, t1, i1});
schedule.Parallel(t0);
schedule.Parallel(i0);
\end{lstlisting}

The loop-nest generated by \Treebeard{} for the above schedule is as follows. 
\begin{lstlisting}[style=c++]
  builtin.func @Prediction_Function(%arg0: memref<batch_sizexnum_featuresxf64>) -> memref<batch_sizexf64> {
    %result = memref.alloc <batch_sizexf64>
    %0 = #decisionforest<ReductionType = 0, #Trees = N_t, resultType = memref<batch_sizexf64>> 
    par.for i0 = range(0 : batch_size/2 : batch_size) {
      par.for t0 = range(0 : N_t/2 : N_t) {
        for t1 = range(0 : N_t/2) {
          for i1 = range(0 : batch_size/2) {
            %2 = GetTree(%0, t0 + t1) 
            %3 = WalkDecisionTree(%2, %arg0[i0+i1])
            reduce(%result, i0+i1, %3)
          }
        }
      }
    }
  }
\end{lstlisting}

\Treebeard{} determines that the \op{t0} loop is a conflicting loop for the \op{result}
array and therefore legalizes the reduction by inserting a privatized array 
\op{result\_1}. The privatized dimension of this array is reduced at the end 
of the \op{t0} loop.

\begin{lstlisting}[style=c++]
  builtin.func @Prediction_Function(%arg0: memref<batch_sizexnum_featuresxf64>) -> memref<batch_sizexf64> {
    %result = memref.alloc <batch_sizexf64>
    %result_1 = memref.alloc <2xbatch_sizexf64>
    %0 = #decisionforest<ReductionType = 0, #Trees = N_t, resultType = memref<batch_sizexf64>> 
    par.for i0 = range(0 : batch_size/2 : batch_size) {
      par.for t0 = range(0 : N_t/2 : N_t) {
        for t1 = range(0 : N_t/2) {
          for i1 = range(0 : batch_size/2) {
            %2 = GetTree(%0, t0 + t1) 
            %3 = WalkDecisionTree(%2, %arg0[i0+i1])
            reduce(%result, i0+i1, %3)
          }
        }
      }
      %result[i0 : i0+step] = reduce_dimension(%result_1, 0, i0 : i0+step)
    }
  }
\end{lstlisting}

While legalizing the reduction, the compiler determines that the 
\op{reduce\_dimension} operation must only process a subset of the final 
result that is computed within the current parallel iteration of the 
\op{i0} loop. Once this process is complete, the \op{reduce} ops in 
the result IR can be lowered to a simple ``read-accumulate-write''
sequence of instructions

Finally, we note that in our experiments, we found that our 
current implementation of lowering the reduction operations 
was sufficient and reduction is not the bottleneck in our 
generated code. However, we believe this approach to enabling 
higher level code generators to easily generate reductions 
through simple abstractions and then having the compiler 
automatically lower them to efficient implementation is an
important area for future work with applicability in several 
domains.

\TODO{Should we mention how we handle multi-class models?}

### 2024-03-13 src/python/treebeard.py
Add kernel timing for auto-tuning heuristic

@@ -11,6 +11,7 @@ treebeardAPI = TreebeardAPI()
 class GPUAutoScheduleOptions:
   def __init__(self) -> None:
       self.optionsPtr = treebeardAPI.runtime_lib.CreateGPUAutoScheduleOptions()
+      self.unrollTreeWalks = False
   
   def __del__(self):
       treebeardAPI.runtime_lib.DeleteGPUAutoScheduleOptions(self.optionsPtr)
@@ -37,6 +38,7 @@ class GPUAutoScheduleOptions:
     treebeardAPI.runtime_lib.Set_cacheTrees(self.optionsPtr, ctypes.c_int32(1 if val else 0))
 
   def UnrollTreeWalks(self, val : bool):
+    self.unrollTreeWalks = val
     treebeardAPI.runtime_lib.Set_unrollTreeWalks(self.optionsPtr, ctypes.c_int32(1 if val else 0))
 
   def TreeWalkInterleaveFactor(self, val : int):

### 2024-03-13 test/python/RAPIDs/rapids_test.py

@@ -6,6 +6,7 @@ import numpy
 import pandas
 import xgboost as xgb
 import treebeard
+import matplotlib.pyplot as plt
 from cuml import ForestInference
 from functools import partial
 
@@ -26,17 +27,17 @@ def get_model_config(model_name : str, batch_size: int):
     256: (treebeard.GPUAutoScheduleOptions.Construct(8, 1, 50, True, False, True, 2, True), "gpu_array"),
     512: (treebeard.GPUAutoScheduleOptions.Construct(8, 1, 50, True, False, True, 2, True), "gpu_reorg"),
     1024: (treebeard.GPUAutoScheduleOptions.Construct(8, 1, 20, True, False, True, 2, True), "gpu_reorg"),
-    2048: (treebeard.GPUAutoScheduleOptions.Construct(8, 1, 20, True, False, True, 2), "gpu_reorg"),
+    2048: (treebeard.GPUAutoScheduleOptions.Construct(8, 1, 20, True, False, True, 4), "gpu_reorg"),
     4096: (treebeard.GPUAutoScheduleOptions.Construct(32, 1, 20, True, False, True, 2), "gpu_array"),
     8192: (treebeard.GPUAutoScheduleOptions.Construct(32, 1, 10, True, False, True, 4), "gpu_array"),
-    16384: (treebeard.GPUAutoScheduleOptions.Construct(64, 1, 2, False, False, True, 2), "gpu_array"),
+    16384: (treebeard.GPUAutoScheduleOptions.Construct(64, 1, 10, True, False, True, 4, True), "gpu_array"),
   }
   airline_best_configs = {
     256: (treebeard.GPUAutoScheduleOptions.Construct(8, 1, 50, False, False, True, -1, True), "gpu_array"),
     512: (treebeard.GPUAutoScheduleOptions.Construct(8, 1, 50, False, False, False, -1, True), "gpu_array"),
     1024: (treebeard.GPUAutoScheduleOptions.Construct(32, 1, 20, True, False, True, -1, True), "gpu_reorg"),
-    2048: (treebeard.GPUAutoScheduleOptions.Construct(32, 2, 10, True, False, True, -1), "gpu_array"),
-    4096: (treebeard.GPUAutoScheduleOptions.Construct(32, 2, 10, True, False, True, -1), "gpu_array"),
+    2048: (treebeard.GPUAutoScheduleOptions.Construct(32, 1, 20, True, False, True, 4, True), "gpu_array"),
[…]

### 2024-03-13 test/python/RAPIDs/rapids_test.py
1. Add RAPIDs compare results 2. Change best configs in python benchmark script

@@ -34,15 +34,15 @@ def get_model_config(model_name : str, batch_size: int):
   }
   airline_best_configs = {
     256: (treebeard.GPUAutoScheduleOptions.Construct(8, 1, 50, False, False, True, -1, True), "gpu_array"),
-    512: (treebeard.GPUAutoScheduleOptions.Construct(8, 1, 50, False, False, False, -1, True), "gpu_array"),
-    1024: (treebeard.GPUAutoScheduleOptions.Construct(32, 1, 20, True, False, True, -1, True), "gpu_reorg"),
+    512: (treebeard.GPUAutoScheduleOptions.Construct(32, 1, 20, True, False, True, 4, True), "gpu_array"),
+    1024: (treebeard.GPUAutoScheduleOptions.Construct(32, 1, 20, True, False, True, 4, True), "gpu_array"),
     2048: (treebeard.GPUAutoScheduleOptions.Construct(32, 1, 20, True, False, True, 4, True), "gpu_array"),
-    4096: (treebeard.GPUAutoScheduleOptions.Construct(32, 2, 10, True, False, True, 2, True), "gpu_sparse"),
-    8192: (treebeard.GPUAutoScheduleOptions.Construct(64, 1, 10, True, False, True, 2), "gpu_array"),
+    4096: (treebeard.GPUAutoScheduleOptions.Construct(32, 1, 10, True, False, True, 2, True), "gpu_sparse"),
+    8192: (treebeard.GPUAutoScheduleOptions.Construct(32, 1, 10, True, False, True, 4), "gpu_array"),
     16384: (treebeard.GPUAutoScheduleOptions.Construct(64, 1, 2, True, False, True, 2), "gpu_array"),
   }
   airline_ohe_best_configs = {
-    256: (treebeard.GPUAutoScheduleOptions.Construct(8, 1, 50, True, False, True, 4), "gpu_array"),
+    256: (treebeard.GPUAutoScheduleOptions.Construct(8, 1, 50, True, False, True, 4, True), "gpu_array"),
     512: (treebeard.GPUAutoScheduleOptions.Construct(8, 1, 50, True, False, True, 2), "gpu_array"),
     1024: (treebeard.GPUAutoScheduleOptions.Construct(8, 1, 50, True, False, True, 2), "gpu_reorg"),
     2048: (treebeard.GPUAutoScheduleOptions.Construct(8, 1, 50, True, False, True, 2), "gpu_array"),
[…]

### 2024-03-29 docs/SOSP 2024/caching.tex
Add caching section

\section{Representations}
\TODO{Talk about array, sparse and reorg representations here.}

\section{Caching Implementation}
\label{sec:caching}

% \begin{itemize}
%   \item The compiler exposes caching of both trees and input rows in a unified manner.
%   \item This is independent of the final target on which the inference is to be run.
%   \item The caching is done at the granularity of a tree or a row.
%   \item Caching is encoded in the mid-level IR using the \op{cacheTrees} and \op{cacheRows} operations. 
%   These operations are generated when the HIR lowered to MIR and \op{Cache()} is specified on an 
%   index variable in the schedule. While the HIR is being lowered and a cached index variable is 
%   encountered, the compiler generates a \op{cacheTrees} or \op{cacheRows} operation depending on 
%   whether the index variable is a tree or a batch index variable.
%   \TODO{Should we talk about how the lowering from HIR to MIR is actually implemented as a tree walk?}
%   \item The lowering of these ops is done by the target-specific code generator.
%   \item For the \op{cacheRows} operation, \Treebeard{} uses pre-implemented lowerings for both
%   CPU and GPU. This is possible because the input is currently assumed to be a dense 
%   array format. 
%   \item For the \op{cacheTrees} operation, the lowering is representation-specific.
%   Each representation provides a lowering to the target-specific code generator
%   to lower the \op{cacheTrees} op when that representation is used.
%   \item The cache operations are lowered to reads into shared memory while compiling to GPU 
%   and to prefetches while compiling to CPU.
% \end{itemize}

\Treebeard{} provides mechanisms to cache both trees and input rows 
on both the CPU and GPU. As described in Section \ref{sec:schedule}, 
the user can specify that the working set of an iteration of an index
variable needs to be cached using the \op{Cache()} directive. This 
provides a unified way to specify caching of both trees and input rows.

\Treebeard{} implements caching at the granularity of a tree or a row.
Also, the semantics of caching depends on the target processor. For the CPU,
caching is implemented as prefetching, while for the GPU, caching is implemented
using shared memory.

\subsection{IR Representation of Caching}
Caching is encoded in the mid-level IR using the \op{cacheTrees} and \op{cacheRows} operations. 
These operations are generated when the HIR lowered to MIR and \op{Cache()} is specified on an 
index variable in the schedule. While the HIR is being lowered and a cached index variable is 
encountered, the compiler generates a \op{cacheTrees} or \op{cacheRows} operation depending on 
whether the index variable is a tree or a batch index variable. Additionally, as a part of the 
lowering process, \Treebeard{} determines the working set of the loop with caching enabled 
and generates a caching operation with the appropriate limits.

Each of the caching operations take parameters that specify the set of trees or rows that need 
to be cached. The caching operations are defined as follows. 
\begin{itemize}
  \item \textbf{\op{cacheTrees(forest, start, end)}:} This operation caches the trees in ensemble
  \op{forest} from \op{start} to \op{end}. The trees are cached in the order in which they are
  specified in the ensemble. 
  \item \textbf{\op{cacheRows(data, start, end)}:} This operation caches the rows in the input 
  array \op{data} from \op{start} to \op{end}. The rows are cached in the same order as in the
  input array.
\end{itemize}

\subsection{Lowering of Caching Operations}
When the MIR is lowered to LIR, the cache ops are lowered to target-specific code. Each of 
the two caching operations is lowered differently for the CPU and the GPU.

For the \op{cacheRows} operation, \Treebeard{} uses pre-implemented lowerings for both
CPU and GPU. This is possible because the input is currently assumed to be a dense
array format. Therefore, regardless of any other configuration choices (like what representation 
is used for the model itself), the lowering of the \op{cacheRows} operation is the same. 
For the CPU, the \op{cacheRows} operation is lowered to prefetches. For the GPU, a 
series of coalesced loads read the rows into shared memory.

For the \op{cacheTrees} operation, the lowering is representation-specific. Each representation
provides a lowering to the target-specific code generator to lower the \op{cacheTrees} op when
that representation is used. However, the \Treebeard{} infrastructure does provide helpers 
to generate caching code to cache contiguous regions of memory. These helpers are reused 
as required across different representations.

\section{Exploring the Schedule Space}
\TODO{Motivate the need for schedule space exploration (maybe some initial numbers on how different the perf is with different schedules?
The histograms maybe?). Then motivate the need for a search space exploration tool (present some numbers on how long schedule exploration takes).}

### 2024-03-29 docs/SOSP 2024/hir-mir-opts.tex
Add caching section

\section{HIR and MIR Optimizations}

\begin{itemize}
  \item The compiler performs a number of optimizations on the HIR and MIR to improve the performance of the generated code.
  \item Optimizations in HIR mainly transform the model itself. Tiling, tree reordering and padding are some of the 
  optimizations that are performed on the HIR. \TODO{Should we use the MICRO paper overview for this part?}
  \item Explain lowering to MIR using the schedule here.
  \item Optimizations on the MIR include tree walk unrolling, tree walk interleaving, loop tiling, parallelization etc.
\end{itemize}

### 2024-03-31 docs/SOSP 2024/caching.tex
Add sections for schedule space exploration and introduction

@@
 \section{Exploring the Schedule Space}
-\TODO{Motivate the need for schedule space exploration (maybe some initial numbers on how different the perf is with different schedules?
-The histograms maybe?). Then motivate the need for a search space exploration tool (present some numbers on how long schedule exploration takes).}
+\label{sec:exploring}
+
+\begin{itemize}
+  \item The scheduling language described in Section \ref{sec:schedule} allows for numerous possible schedules for a given model. 
+  Finding a schedule with good performance is a non-trivial task.
+  \item We identify a reasonable template schedule for GPUs which encompasses all strategies published in prior work.
+  \item Describe the template schedule.
+  \item Even within the variants of this template schedule, there is a significant amount of variation in performance. Show the histogram. 
+  \item Extensively exploring all possible parameter values for the template schedule is infeasible due to the large number of parameters.
+  Cite some numbers on how long it takes to explore the schedule space.
+  \item List the observations we make and the heuristic we design based on these.
+\end{itemize}
+
+The set of schedules that can be constructed using the scheduling language described in 
+Section \ref{sec:schedule} is vast. Searching this schedule space to find a schedule that
+provides good performance is a non-trivial task. To simplify this process, we identify a
+reasonable template schedule for GPUs that encompasses all strategies published in prior work.
+The template schedule exposes many parameters that can be tuned to find a 
+high-performance schedule.
+\begin{itemize}
+  \item \textbf{Number of rows per thread block (Integer):} The number of rows that are processed by each thread block.
+  \item \textbf{Number of rows per thread (Integer):} The number of rows processed by each thread.
+  \item \textbf{Number of tree threads (Integer):} The number of threads across which the trees are distributed.
+  \item \textbf{Cache rows (Boolean):} Whether the input rows are cached in shared memory.
+  \item \textbf{Cache trees (Boolean):} Whether the trees are cached in shared memory.
+  \item \textbf{Unroll walks (Boolean):} Whether the tree walks are unrolled.
+  \item \textbf{Tree walk interleave factor:} The number of tree walks that are interleaved.
+  \item \textbf{Shared memory reduction:} Whether the reduction across tree threads is done in shared memory.
+\end{itemize}
+
+While the template schedule simplifies optimization of generated inference code, it is 
+important to note that the \Treebeard{} compiler itself does not place any 
+restrictions on the schedule. The user is free to specify any schedule they wish.
+The auto-scheduler that implements the template schedule is implemented as a 
+module outside the core \Treebeard{} compiler. Users are also free to implement 
+other auto-schedulers that generate schedules different from the template schedule.

### 2024-03-31 docs/SOSP 2024/intro.tex
Add sections for schedule space exploration and introduction

\section{Introduction}
\label{sec:intro}

\subsection{Contributions}
\begin{itemize}
  \item We present the design for a multi-target decision tree compiler infrastructure and implement several optimizations within this framework.
   We are also the first to implement an optimizing compiler for decision tree inference on GPUs.
  \item We identify that an extensive optimization space exists for the problem of decision tree inference. We design a scheduling language that 
  allows us to effectively represent this solution space abstractly. This scheduling language is expressive enough to represent a wide range of  
  implementation strategies proposed by prior work.
  \item To the best of our knowledge, we perform the first extensive characterization of the optimization space for decision tree inference 
  on GPUs. Using some of the characteristics we identify, we design and implement a heuristic that is able to quickly find high-performance 
  schedules for the model being compiled. 
  \item We design and implement a general framework for expressing and optimizing reductions within MLIR. To the best of our knowledge, this is the first
  such framework.
  \item We evaluate our implementation by comparing it against RAPIDs and Tahoe, the state-of-the-art decision tree inference frameworks for GPU and 
  report significant speedups. We also show that our compiler can effectively target different GPUs, including both NVIDIA and AMD GPUs.
\end{itemize}

### 2024-04-01 results/python/analyze_random_model_benchmark_results.py
Add random model results on thedeep

@@ -1,0 +1,78 @@
+import os
+
+batch_size = 4096
+# file_path = "/home/ashwin/mlir-build/llvm-project/mlir/examples/treebeard/results/thedeep/gpu/random_models/20240320/random_cmp_output_mp.txt"
+# file_path = "/home/ashwin/mlir-build/llvm-project/mlir/examples/treebeard/results/thedeep/gpu/random_models/20240320/random_cmp_output_mp_b512.txt"
+file_path = "/home/ashwin/mlir-build/llvm-project/mlir/examples/treebeard/results/thedeep/gpu/random_models/20240320/random_cmp_output_mp_b4k_ft500.txt"
+
+# read the file line by line and filter out lines that don't start with "/"
+with open(file_path, "r") as file:
+    lines = file.readlines()
+    lines = [line for line in lines if line.startswith("/")]
+
+# split each line by the space char
+lines = [line.split(" ") for line in lines]
+
+# construct a pandas dataframe from the list of lists
+import pandas as pd
+cols = ["model_path", "batch_size", "fil_time", "tb_time", "total_speedup", "fil_kernel_time", "tb_kernel_time", "tb_kernel_speedup"]
+df = pd.DataFrame(lines, columns=cols)
+
+# cast the tb_kernel_speedup and total_speedup column to float
+df["tb_kernel_speedup"] = df["tb_kernel_speedup"].astype(float)
+df["total_speedup"] = df["total_speedup"].astype(float)
+
+# compute it from the model_path as follows
+# - split the model_path by the "/" char and get the file name
+# - split the file name by the "_" char and get the third element
+def get_num_trees(model_path):
+    return int(model_path.split("/")[-1].split("_")[2])
+
+# add a column to the dataframe called "num_trees". 
+# This column contains the number of trees in the model
+df["num_trees"] = df["model_path"].apply(get_num_trees)
+
+def get_num_features(model_path):
+    return int(model_path.split("/")[-1].split("_")[3])
+
+df["num_features"] = df["model_path"].apply(get_num_features)
+
+def get_tree_depth(model_path):
+    return int(model_path.split("/")[-1].split("_")[4])
+
+df["tree_depth"] = df["model_path"].apply(get_tree_depth)
+
+print(df.shape)
+# print the min, max and geomean of the tb_kernel_speedup column
[…]

### 2024-04-02 docs/SOSP 2024/results.tex
Add results section

\section{Experimental Evaluation}

\begin{figure}[htb]
  \centering
  \includegraphics[width=\linewidth]{figures/TBvsRAPIDs_4060_Speedup.png}
  \caption{\Treebeard{} vs RAPIDs Speedup on NVIDIA RTX 4060.}
  \label{Fig:TBvsRAPIDs_4060_Speedup}
\end{figure}

\begin{figure}[htb]
  \centering
  \includegraphics[width=\linewidth]{figures/TBvsTahoe_4060_KernelTimeSpeedup.png}
  \caption{\Treebeard{} vs Tahoe Kernel Time Speedup.}
  \label{Fig:TBvsTahoe_4060_KernelTimeSpeedup}
\end{figure}

\begin{figure}[htb]
  \centering
  \includegraphics[width=\linewidth]{figures/TBvsRAPIDs_AllBatchSizes_T400.png}
  \caption{\Treebeard{} vs RAPIDs Speedup on T400.}
  \label{Fig:TBvsRAPIDs_AllBatchSizes_T400}
\end{figure}

\begin{figure}[htb]
  \centering
  \includegraphics[width=\linewidth]{figures/AutotuningSpeedupvs4060Sched_T400.png}
  \caption{Autotuning heuristics speedup vs best 4060 schedule on T400.}
  \label{Fig:AutotuningSpeedupvs4060Sched_T400}
\end{figure}

\begin{figure}[htb]
  \centering
  \includegraphics[width=\linewidth]{figures/HeuristicVsFullExplore_Speedup.png}
  \caption{Autotuning heuristic compile time speedup vs full schedule exploration.}
  \label{Fig:HeuristicVsFullExplore_Speedup}
\end{figure}

\begin{figure}[htb]
  \centering
  \includegraphics[width=\linewidth]{figures/AMD_MI210_ATHeuristicVs4060Sched_speedup.png}
  \caption{Autotuning heuristics speedup vs best 4060 schedule on MI210.}
  \label{Fig:AMD_MI210_ATHeuristicVs4060Sched_speedup}
\end{figure}

\begin{figure*}[ht]
  \centering
  \begin{subfigure}[b]{.3\textwidth}
    \subcaptionbox*{}{\includegraphics[width=\textwidth]{figures/RandomModels/kernel_speedup_b512_depth6.png}}
    \caption{Kernel Speedup for batch size 512, depth 6.}
  \end{subfigure}
  \hfill  % Add horizontal space between subfigures
  \begin{subfigure}[b]{.3\textwidth}
    \subcaptionbox*{}{\includegraphics[width=\textwidth]{figures/RandomModels/kernel_speedup_b512_depth7.png}}
    \caption{Kernel Speedup for batch size 512, depth 7.}
  \end{subfigure}
  \hfill
  \begin{subfigure}[b]{.3\textwidth}
    \subcaptionbox*{}{\includegraphics[width=\textwidth]{figures/RandomModels/kernel_speedup_b512_depth8.png}}
    \caption{Kernel Speedup for batch size 512, depth 8.}
  \end{subfigure}
  \\  % Line break for the second row
  \begin{subfigure}[b]{.3\textwidth}
    \subcaptionbox*{}{\includegraphics[width=\textwidth]{figures/RandomModels/kernel_speedup_b4096_depth6.png}}
    \caption{Kernel Speedup for batch size 4096, depth 6.}
  \end{subfigure}
  \hfill
  \begin{subfigure}[b]{.3\textwidth}
    \subcaptionbox*{}{\includegraphics[width=\textwidth]{figures/RandomModels/kernel_speedup_b4096_depth7.png}}
    \caption{Kernel Speedup for batch size 4096, depth 7.}
  \end{subfigure}
  \hfill
  \begin{subfigure}[b]{.3\textwidth}
    \subcaptionbox*{}{\includegraphics[width=\textwidth]{figures/RandomModels/kernel_speedup_b4096_depth8.png}}
    \caption{Kernel Speedup for batch size 4096, depth 8.}
  \end{subfigure}
\end{figure*}

### 2024-04-02 src/test/GPUBenchmarkUtils.h
1. Tests to compare random models with RAPIDs 2. Changes needed to autotune random models 3. Flip loops of RAPIDs kernel benchmark

@@ -44,7 +44,13 @@ struct GPUTimes {
 
 class TestDataReader {
   const bool useSameRowForAllInputs = false;
+  struct InputData {
+    std::vector<double> data;
+    int32_t rowSize;
+    int32_t numRows;
+  };
   std::map<std::string, std::vector<double>> m_benchmarkToDataMap;
+  std::map<std::string, InputData> m_modelPathToDataMap;
 
   std::vector<double> readDataForBenchmark(const std::string &benchmarkName) {
 
@@ -58,15 +64,44 @@ class TestDataReader {
     for (size_t i = 0; i < csvReader.NumberOfRows(); ++i) {
       if (useSameRowForAllInputs) {
         auto row = csvReader.GetRowOfType<double>(0);
+        row.pop_back();
         data.insert(data.end(), row.begin(), row.end());
       } else {
         auto row = csvReader.GetRowOfType<double>(i);
+        row.pop_back();
         data.insert(data.end(), row.begin(), row.end());
       }
     }
 
     m_benchmarkToDataMap[benchmarkName] = data;
     m_benchmarkToDataMap[csvFilename] = data;
+
+    return data;
+  }
+
+  std::vector<double>
+  readDataForNonBenchmarkModel(const std::string &csvFilename) {
+    TestCSVReader csvReader(csvFilename, -1 /*num lines*/, true /*hasNumRows*/,
+                            ' ');
+
+    std::vector<double> data;
+    int32_t rowSize = -1;
+    for (size_t i = 0; i < csvReader.NumberOfRows(); ++i) {
+      if (useSameRowForAllInputs) {
+        auto row = csvReader.GetRowOfType<double>(0);
+        row.pop_back();
+        data.insert(data.end(), row.begin(), row.end());
+        rowSize = row.size();
+      } else {
+        auto row = csvReader.GetRowOfType<double>(i);
+        row.pop_back();
+        data.insert(data.end(), row.begin(), row.end());
+        rowSize = row.size();
+      }
+    }
+
+    m_modelPathToDataMap[csvFilename] = InputData{
+        data, rowSize, static_cast<int32_t>(csvReader.NumberOfRows())};
 
     return data;
   }
@@ -90,19 +125,45 @@ public:
 
   int32_t getRowSize(const std::string &benchmarkName) {
     auto iter = m_benchmarkToDataMap.find(benchmarkName);
-    assert(iter != m_benchmarkToDataMap.end());
-    return iter->second.size() / 2000;
[…]

### 2024-04-02 src/test/TestUtilsCommon.cpp

@@ -12,7 +12,8 @@
 namespace TreeBeard {
 namespace test {
 
-std::vector<double> getNextLineAndSplitIntoTokens(std::istream &str) {
+std::vector<double> getNextLineAndSplitIntoTokens(std::istream &str,
+                                                  char delimiter) {
   std::vector<double> result;
   std::string line;
   std::getline(str, line);
@@ -20,7 +21,7 @@ std::vector<double> getNextLineAndSplitIntoTokens(std::istream &str) {
   std::stringstream lineStream(line);
   std::string cell;
 
-  while (std::getline(lineStream, cell, ',')) {
+  while (std::getline(lineStream, cell, delimiter)) {
     result.push_back(std::stof(cell));
   }
   // This checks for a trailing comma with no data after it.
@@ -31,12 +32,16 @@ std::vector<double> getNextLineAndSplitIntoTokens(std::istream &str) {
   return result;
 }
 
-TestCSVReader::TestCSVReader(const std::string &filename, int32_t numRows) {
+TestCSVReader::TestCSVReader(const std::string &filename, int32_t numRows,
+                             bool hasNumRows, char delimiter) {
   std::ifstream fin(filename);
   assert(fin);
+  if (hasNumRows) {
+    fin >> numRows;
+  }
   int numRowsRead = 0;
   while (!fin.eof() && (numRows == -1 || numRowsRead < numRows)) {
-    auto row = getNextLineAndSplitIntoTokens(fin);
+    auto row = getNextLineAndSplitIntoTokens(fin, delimiter);
     m_data.push_back(row);
     ++numRowsRead;
   }

### 2024-04-02 src/test/TestUtilsCommon.h

@@ -116,7 +116,8 @@ class TestCSVReader {
   std::string m_filename;
 
 public:
-  TestCSVReader(const std::string &filename, int32_t numLines = -1);
+  TestCSVReader(const std::string &filename, int32_t numLines = -1,
+                bool hasNumLines = false, char delimiter = ',');
   std::vector<double> &GetRow(size_t index) { return m_data[index]; }
 
   template <typename T> std::vector<T> GetRowOfType(size_t index) {

### 2024-04-02 test/python/RAPIDs/rapids_test.py

@@ -196,8 +196,8 @@ def RunSingleTest_RAPIDs_KernelTime(modelJSONPath, csvPath, output_class, num_re
   gpu_array = cp.array(inputs, dtype=numpy.float32)
 
   start = time.time()
-  for i in range(num_repeats):
-    for j in range(0, num_batches):
+  for j in range(0, num_batches):
+    for i in range(num_repeats):
       start_index = j*batchSize
       stop_index = start_index + batchSize
       batch = gpu_array[start_index:stop_index, :]
@@ -465,3 +465,4 @@ if __name__ == "__main__":
     print("Geomean speedup (Total):", batchSize, numpy.prod(total_time_speedups)**(1/len(total_time_speedups)))
     print("Geomean speedup (Kernel):", batchSize, numpy.prod(kernel_time_speedups)**(1/len(kernel_time_speedups)))
     print("Geomean speedup (AutoTune Total):", batchSize, numpy.prod(autotune_total_time_speedups)**(1/len(autotune_total_time_speedups)))
+    print("Geomean speedup (AutoTune Kernel):", batchSize, numpy.prod(autotune_kernel_time_speedups)**(1/len(autotune_kernel_time_speedups)))

### 2024-04-03 docs/SOSP 2024/intro.tex
intro sketch

@@
 \label{sec:intro}
+We are in the midst of a hardware revolution, a new golden age for computer architecture~\cite{TuringLecture}.
+The last decade has seen a shift in architectural paradigms, with the rise of GPUs and accelerators.
+This shift has been driven by the necessity to innovate in the post Moore's law and Dennard's scaling era.
+This transformation has also played a significant role in the success of modern deep learning models, as they enable scaling model training and inference to a massive number of threads.
+Such scalability would be essential for all performance critical applications, including other machine learning models that need to scale with increasing data sizes and model complexities.
+%need to accelerate machine learning workloads.
+The success of modern AI models is due to their ability to scale both training and inference to a massive number of threads.
+This has been made possible by the rise of GPUs and accelerators.
 
+%Despite these recent advancements, one dominant family of models, namely decision forest models have not seen the same level of acceleration.
+Decision forest models remain the mainstay for machine learning over tabular data~\cite{tabular}. 
+Their robustness, interpretability, and ability to handle missing data make them a popular choice for a wide range of applications. 
+\TODO{a few more lines, classification, regression, etc.}
+%We observe that while scalable performance on neural network models has received significant attention, decision tree inference has not seen the same level of acceleration.
+A recent survey~\cite{kaggle}\dots. 
+The survey also finds that, the cost of inference is the most critical factor in the overall cost of deploying a machine learning model.
+This is because, in production settings, each model is trained once and often used for inference a few thousand times. 
+%
+
+This paper is motivated by the need to accelerate decision tree inference to achieve portable performance on a variety of client hardware.
+In particular, we focus on portable performance across a range of commodity CPUs and low-to-mid-range GPUs, a class of hardware that has seen widespread adoption across client/edge devices used for inference.
+
+Decision forest models are composed of a large collection of decision trees (100-1000), and inference involves traversing down each tree in the forest and aggregating the predictions.
+Inference is typically done in a batched setting, where multiple inputs are processed simultaneously.
+
+%Despite the simplicity of the model, decision tree inference is computationally expensive, and is surprisingly hard to scale to a large number of threads
+Despite the simplicity of the model and the availability of multiple sources of coarse grain parallelism (parallelism across inputs in a batch and parallelism across trees), existing systems do not consistently scale well across different models.
+This is because existing systems exploit a limited set of optimizations and often specialize the implementation to a specific hardware platform, which limits their portability.
+
+%\TODO{cover Rapids, Tahoe, XGBoost and Treebeard}.  
+Evaluation on a diverse set of models highlights that the best implementation often requires a careful combination of many optimization strategies.
+For example, managing the working set of nodes across a set of trees is critical to get scalable performance.
+This requires a combination of techniques like data layout optimizations, loop transformations, and memory access optimizations.
+Existing systems each use specific data layouts for the underlying tree (XGBoost uses a sparse representation, RAPIDs FIL uses what is called the reorg representation
+and Tahoe uses a variation of the reorg representation) and propose simple loop transformations (XGBoost uses \dots()) that together determine the accesses patterns and the working set.
+They do not perform consistently well across different models and hardware platforms.
+Managing the working set is just one of several optimizations that are critical for high-performance decision tree inference.
+ \TODO{Improve the next line}
+ Others include tree ordering,  incremental reduction of predicted values, interleaving across multiple trees, and so on.
+
+This paper presents \Treebeard{}, a novel schedule guided compilation infrastructure for decision tree inference on multiple target hardware.
+\Treebeard{} is able to generate high-performance code for decision tree inference by exploring a large optimization space.
+The code generation is guided by a \emph{schedule} object that allows us to effectively represent this solution space abstractly.
+The schedule object is written in \Treebeard{} custom scheduling language that is expressive enough to represent a wide range of implementation strategies.
+We demonstrate that the language is sufficient to express the various optimizations proposed by prior work, and further generalizes them and incorporates several new optimizations.
+We also design and implement a heuristic that is able to quickly find high-performance schedules for the model being compiled.
+\TODO{Performance evaluation summary}
+\TODO{Oganization}
 \subsection{Contributions}

### 2024-04-03 docs/SOSP 2024/caching.tex
Intermediate submission

@@
 \begin{itemize}
-  \item The scheduling language described in Section \ref{sec:schedule} allows for numerous possible schedules for a given model. 
+  \item The scheduling language described in Section \ref{sec:schedule} allows
+  for numerous possible schedules for a given model. 
   Finding a schedule with good performance is a non-trivial task.
-  \item We identify a reasonable template schedule for GPUs which encompasses all strategies published in prior work.
+  \item We identify a reasonable template schedule for GPUs which encompasses
+  all strategies published in prior work.
   \item Describe the template schedule.
-  \item Even within the variants of this template schedule, there is a significant amount of variation in performance. Show the histogram. 
-  \item Extensively exploring all possible parameter values for the template schedule is infeasible due to the large number of parameters.
+  \item Even within the variants of this template schedule, there is a significant 
+  amount of variation in performance. Show the histogram. 
+  \item Extensively exploring all possible parameter values for the template schedule 
+  is infeasible due to the large number of parameters.
   Cite some numbers on how long it takes to explore the schedule space.
@@
 provides good performance is a non-trivial task. To simplify this process, we identify a
-reasonable template schedule for GPUs that encompasses all strategies published in prior work.
-The template schedule exposes many parameters that can be tuned to find a 
+template schedule for GPUs that encompasses several strategies published in prior work.
+Our template schedule assigns a fixed number of rows to each thread block and to each thread.
+It distributes the trees across a fixed number of threads and can cache trees and 
+input rows if required. Unrolling and interleaving of tree walks is also supported.
+
+The template schedule exposes a number of parameters that can be tuned to find a 
 high-performance schedule.
@@
 other auto-schedulers that generate schedules different from the template schedule.
+
+\begin{figure}[htb]
+  \centering
+  \includegraphics[width=\linewidth]{figures/normalized_kernel_histogram_lt5.png}
+  \caption{Distribution of normalized execution times for all benchmark models
+  with different parameter values for the template schedule. \TODO{We need to 
+  describe exactly what the considered set of parameter values are.}}
+  \label{Fig:ExecTimeDistribution}
+\end{figure}
+
+Figure \ref{Fig:ExecTimeDistribution} shows the distribution of normalized execution times
+for all benchmark models with different parameter values for the template schedule.
+The normalized execution time is the inference time of the model with a given set of
+parameter values divided by the best inference time of the model.
+The histogram shows that even within the variants of the template schedule, there is
+a significant amount of variation in performance. Very few schedules perform close to the best
+while a vast majority of schedules perform poorly.
+
+Exploring the schedule space extensively even for a reasonable set of parameter values
+is very expensive. We explored the set of parameter values listed in Table \ref{table:ParamVals}
+for our benchmarks and found that it took anywhere between thirty minutes up to a few hours
+to explore the entire space. Performing this extensive search for every model being compiled 
+is infeasible in practise. We therefore need a better mechanism to guide the search for a good schedule.
+
+We design a heuristic to narrow down the set of schedules to explore based on the following 
+observations on high-performance schedules.
+\begin{itemize}
+  \item For small batch sizes, the best schedules tend to have a small number of rows
+  per thread block and partition the trees across a larger number of threads. This is 
+  intuitive since the amount of data parallelism across the rows is limited for small batch sizes.
+  \item Always cache rows in shared memory and never cache trees. We find that caching rows 
+  when possible (i.e., when the number of features is small enough to fit in shared memory)
+  almost always improves performance. Caching trees on the other hand almost always degrades 
+  performance. This is because the one time cost of loading trees into shared memory is 
+  not sufficiently amortized when the whole of the tree is not accessed during inference. 
+  \item Models with a large number of features tend to benefit from partitioning the 
+  trees across more threads even at larger batch sizes. This is because processing fewer rows at a time 
+  allows us to keep them in shared memory. We empirically find that the threshold for when 
+  we should start partitioning the trees across more threads is when the number of features
+  is greater than 100. 
+  \item We find that when a model prefers schedules with shared reduction, the same schedules 
+  without shared reduction are among the best performing schedules without shared reduction.
+  We therefore are able to separate the evaluation of shared reduction by collecting the 
+  best schedules without shared reduction and only evaluating shared reduction on them. 
+  Evaluating the top 3 schedules for shared reduction is sufficient in practice.  
+\end{itemize}
+
+\Treebeard{} uses these observations to narrow down the set of schedules to explore. 
+The pseudo-code for the current heurisitic is shown in Algorithm \ref{alg:heuristic}.
+We find that this heuristic is able to find schedules that are close to the best schedules
+but improves the search time by two orders of magnitude as we show in Section \ref{sec:HeuristicResults}.
+
+% \begin{algorithm}
+%   \caption{Heuristic to guide the search for a good schedule}
+%   \label{alg:heuristic}
+%   \begin{algorithmic}
+%     \State $bestSchedules \gets \emptyset$
+%     \For{$rowsPerBlock \in \{32, 64, 128\}$}
+%       \For{$rowsPerThread \in \{1, 2, 4\}$}
+%         \For{$treeThreads \in \{32, 64, 128\}$}
+%           \If{$rowsPerBlock \times rowsPerThread \times treeThreads \leq 1024$}
+%             \State $schedule \gets \{rowsPerBlock, rowsPerThread, treeThreads, \text{True}, \text{False}, \text{False}, 1, \text{False}\}$
+%             \State $bestSchedules \gets bestSchedules \cup \{schedule\}$
+%           \EndIf
+%         \EndFor
+%       \EndFor
+%     \EndFor
+%     \State $bestSchedules \gets \text{SortByPerformance}(bestSchedules)$
+%     \State $bestSchedules \gets \text{EvaluateSharedReduction}(bestSchedules)$
+%     \State \Return $bestSchedules$
+%   \end{algorithmic}

### 2024-04-03 docs/SOSP 2024/caching.tex
More autotuning stuff

@@
 \Treebeard{} uses these observations to narrow down the set of schedules to explore. 
-The pseudo-code for the current heurisitic is shown in Algorithm \ref{alg:heuristic}.
+The pseudo-code for the current heuristic is shown in Algorithm \ref{alg:heuristic}.
+The algorithm first computes a subset of thread block configurations in the function 
+\op{TBConfigs}. A set of schedules based on these thread block configurations 
+is then computed (\op{schedules}). The model is compiled with each of these schedules 
+and then the resulting inference code is profiled. The three best performing schedules 
+are collected and shared reduction is enabled on them and the resulting schedules evaluated.
+The best schedule among all the evaluated schedules is selected as the schedule to use.
 We find that this heuristic is able to find schedules that are close to the best schedules
@@
 
-% \begin{algorithm}
-%   \caption{Heuristic to guide the search for a good schedule}
-%   \label{alg:heuristic}
-%   \begin{algorithmic}
-%     \State $bestSchedules \gets \emptyset$
-%     \For{$rowsPerBlock \in \{32, 64, 128\}$}
-%       \For{$rowsPerThread \in \{1, 2, 4\}$}
-%         \For{$treeThreads \in \{32, 64, 128\}$}
-%           \If{$rowsPerBlock \times rowsPerThread \times treeThreads \leq 1024$}
-%             \State $schedule \gets \{rowsPerBlock, rowsPerThread, treeThreads, \text{True}, \text{False}, \text{False}, 1, \text{False}\}$
-%             \State $bestSchedules \gets bestSchedules \cup \{schedule\}$
-%           \EndIf
-%         \EndFor
-%       \EndFor
-%     \EndFor
-%     \State $bestSchedules \gets \text{SortByPerformance}(bestSchedules)$
-%     \State $bestSchedules \gets \text{EvaluateSharedReduction}(bestSchedules)$
-%     \State \Return $bestSchedules$
-%   \end{algorithmic}
+\begin{algorithm}
+  \caption{Heuristic to guide the search for a good schedule}
+  \label{alg:heuristic}
+  \begin{algorithmic}[1]
+  \Procedure{TBConfigs}{$N_{batch}$, $N_f$}
+    \State $T_{batch} \gets 2048, T_f \gets 128$
+    \If {$N_{batch} \leq T_{batch}$}
+      \State $rowsPerBlock \gets \{8, 32\}$
+      \State $treeThreads \gets \{20, 50\}$
+    \Else
+      \If {$N_{f} \leq T_{f}$}
+        \State $rowsPerBlock \gets \{32, 64\}$
+        \State $treeThreads \gets \{2, 10\}$
+      \Else
+        \State $rowsPerBlock \gets \{8, 32\}$
+        \State $treeThreads \gets \{20, 50\}$
+      \EndIf
+    \EndIf
+    \State \Return $rowsPerBlock, treeThreads$
+  \EndProcedure
+  \\
+  \State $bestSchedules \gets PriorityQueue(3)$
+  \State $rowsPerTB, treeThreads \gets TBConfigs(N_{batch}, N_f)$
+  \State $cacheRows \gets \text{True}$
+  \State $cacheTrees \gets \text{False}$
+  \State $interleaveFactors \gets \{1, 2, 4\}$
+  \State $reps \gets \{array, sparse, reorg\}$
+  \State $schedules \gets (rowsPerTB, treeThreads,$ \\
+                          \hspace{1.5cm}$cacheRows, cacheTrees,$\\ 
+                          \hspace{1.5cm}$interleaveFactors)$
+  \For{$(sched, rep) \in schedules \times reps$}
+    \State $time \gets EvaluateSchedule(sched, rep)$
+    \State $bestSchedules.insert(time, sched, rep)$
+  \EndFor
+  \State $shMemSchedules \gets \emptyset$
+  \For{$sched, rep \in bestSchedules$}
+    \State $EnableSharedReduction(sched)$
+    \State $time \gets EvaluateSchedule(sched, rep)$
+    \State $shMemSchedules.insert(time, sched, rep)$
+  \EndFor
+  \State \Return $min(shMemSchedules \cup bestSchedules)$
+  \end{algorithmic} 
+\end{algorithm}

### 2024-04-04 docs/SOSP 2024/caching.tex
Add section on representations

@@
 \section{Representations}
-\TODO{Talk about array, sparse and reorg representations here.}
+\label{sec:representations}
+
+The design of the \Treebeard{} compiler allows the implementation of different strategies for 
+the in-memory representation of the model. The compiler currently has implementations for 
+the three representations shown in Figure \ref{Fig:Representations}. The array and sparse 
+representations are the ones described in the \Treebeard{} paper\cite{Treebeard}. The reorg 
+representation is the representation used by the RAPIDs library\cite{RAPIDs}. 
+The \textbf{array representation} is the simplest representation where the trees are stored in an array
+in level order. The \textbf{sparse representation} stores the trees in a sparse format where only the
+non-zero nodes are stored and nodes contain pointers to their children. The \textbf{reorg representation}
+interleaves the array representation of each tree in the model: all root nodes are stored first, then 
+the left children of all the roots and so on.
+\begin{figure}[htb]
+  \centering
+  \includegraphics[width=\linewidth]{figures/representations.png}
+  \caption{The three representations supported by \Treebeard{}.}
+  \label{Fig:Representations}
+\end{figure}
+
+One of the major changes we make to the original design of \Treebeard{} \cite{Treebeard} is to
+separate the implementation of representations from the rest of the compiler. This allows
+us to implement representations as plugins to the compiler. We do this by defining an interface 
+that each representation needs to implement. The code generator is implemented using methods 
+on the representation interface thus hiding details of the actual representation from the core 
+compiler. The interface abstracts the following details. 
+\begin{itemize}
+  \item \textbf{Buffer allocation:} The representation object exposes methods that generate 
+  buffer allocations in the IR. 
+  \item \textbf{Moving to child nodes:} Given the value of the predicate at a node (or tile),
+  generate code to move to the appropriate child node of the current node.
+  \item \textbf{Details of leaves:} The interface abstracts determining whether the current
+  node is a leaf (which is needed to terminate walks) and how to get the value of the leaf
+  (which is needed to get a tree's prediction).
+  \item \textbf{Caching trees:} Since reading the trees into shared memory on GPU (or prefetching)
+  them on CPU require knowledge of the buffer layout, the task of generating caching code for trees
+  is delegated to the representation object. 
+  \item \textbf{Loading thresholds and feature indices:} The representation object provides MLIR 
+  rewrite patterns to lower operations that load thresholds and feature indices to LLVM IR. This
+  is necessary because the details of how to load the thresholds and feature indices from the model 
+  buffers is representation-specific.
+\end{itemize}
+
+In summary, the representation interface abstracts the details of how the model is stored in memory
+and allows the compiler to generate code without having to explicitly know the details of the
+representation. This design allows us to implement new representations without changing the core
+compiler infrastructure. Implementing the representations as plugins also allows us to reuse
+the implementations across different lowering pipelines. For example, the code for the array 
+and sparse representations is almost fully shared between the CPU and GPU lowering pipelines.
 
@@
 
-\begin{itemize}
-  \item The scheduling language described in Section \ref{sec:schedule} allows
-  for numerous possible schedules for a given model. 
-  Finding a schedule with good performance is a non-trivial task.
-  \item We identify a reasonable template schedule for GPUs which encompasses
-  all strategies published in prior work.
-  \item Describe the template schedule.
-  \item Even within the variants of this template schedule, there is a significant 
-  amount of variation in performance. Show the histogram. 
-  \item Extensively exploring all possible parameter values for the template schedule 
-  is infeasible due to the large number of parameters.
-  Cite some numbers on how long it takes to explore the schedule space.
-  \item List the observations we make and the heuristic we design based on these.
-\end{itemize}
+% \begin{itemize}
+%   \item The scheduling language described in Section \ref{sec:schedule} allows
+%   for numerous possible schedules for a given model. 
+%   Finding a schedule with good performance is a non-trivial task.
+%   \item We identify a reasonable template schedule for GPUs which encompasses
+%   all strategies published in prior work.
+%   \item Describe the template schedule.
+%   \item Even within the variants of this template schedule, there is a significant 
+%   amount of variation in performance. Show the histogram. 
+%   \item Extensively exploring all possible parameter values for the template schedule 
+%   is infeasible due to the large number of parameters.
+%   Cite some numbers on how long it takes to explore the schedule space.
+%   \item List the observations we make and the heuristic we design based on these.
+% \end{itemize}
 

### 2024-04-05 docs/SOSP 2024/caching.tex
Add HIR MIR optimization section

@@
   generate code to move to the appropriate child node of the current node.
-  \item \textbf{Details of leaves:} The interface abstracts determining whether the current
+  \item \textbf{Leaf representation:} The interface abstracts determining whether the current
   node is a leaf (which is needed to terminate walks) and how to get the value of the leaf

### 2024-04-05 docs/SOSP 2024/hir-mir-opts.tex
Add HIR MIR optimization section

@@
 
-\begin{itemize}
-  \item The compiler performs a number of optimizations on the HIR and MIR to improve the performance of the generated code.
-  \item Optimizations in HIR mainly transform the model itself. Tiling, tree reordering and padding are some of the 
-  optimizations that are performed on the HIR. \TODO{Should we use the MICRO paper overview for this part?}
-  \item Explain lowering to MIR using the schedule here.
-  \item Optimizations on the MIR include tree walk unrolling, tree walk interleaving, loop tiling, parallelization etc.
-\end{itemize}
+% \begin{itemize}
+%   \item The compiler performs a number of optimizations on the HIR and MIR to improve the performance of the generated code.
+%   \item Optimizations in HIR mainly transform the model itself. Tiling, tree reordering and padding are some of the 
+%   optimizations that are performed on the HIR. \TODO{Should we use the MICRO paper overview for this part?}
+%   \item Explain lowering to MIR using the schedule here.
+%   \item Optimizations on the MIR include tree walk unrolling, tree walk interleaving, loop tiling, parallelization etc.
+% \end{itemize}
+
+The \Treebeard{} infrastructure was originally designed to target CPUs \cite{Treebeard}. 
+However, it implements several optimizations on the HIR and MIR that 
+can be leveraged across target processors and we find that these 
+are beneficial for GPUs as well. This reuse of optimizations is possible 
+because the intermediate representations on which these optimizations are performed 
+are abstract and are designed to be target-independent. We briefly review theese 
+optimizations below.
+
+\subsection{Optimizations on High-Level IR}
+We augment the existing \Treebeard{} infrastructure with loop rewrites on 
+the HIR that are implemented through the scheduling language (Section \ref{sec:schedule}).
+We use these to implement the automatic scheduling described in Section \ref{sec:exploring}. 
+Additionally, the \Treebeard{} infrastructure implements HIR transformations to reorder and pad 
+trees. It also implements tree tiling transformations on the HIR \cite{Treebeard}. We reuse the 
+reordering and padding transformations on the HIR for GPUs. However, we found that 
+tiling trees was not beneficial for GPUs. This is because the tiling transformations
+introduces redundant computation inorder to vectorize computation on CPUs where 
+all lanes need to follow the same control flow. However, on SIMT GPUs, we find that 
+the benefits of tiling (coalescing memory accesses) do not outweigh the cost of
+redundant computation. We leave an investigation of this for future work.
+
+\subsection{Optimizations on Mid-Level IR}
+The original \Treebeard{} infrastructure implements optimizations like 
+tree-walk unrolling, tree-walk interleaving, and parallelization
+on the MIR. These optimizations are beneficial for GPUs as well.
+and the design of \Treebeard{} allows us to reuse the tree-walk 
+unrolling and tree-walk interleaving optimizations on the MIR for GPUs.
+\TODO{Should we talk about how interleaving is implemented as a 
+statemachine and therefore it can be used across representations 
+and tile traversal techniques?}
+
+\subsection{A Note on Low-level IR}
+Some changes to the original \Treebeard{} design were required to get
+LIR to correctly lower to GPU code. The most important of these was 
+the change to how the compiler implements support for in-memory 
+representations of models (Section \ref{sec:representations}).
+With these design changes, we were able to reuse much of the CPU 
+implementation while customizing some parts for GPUs (for example,
+buffers need to be allocated differently for CPU and GPU, caching 
+is implemented differently etc.).

### 2024-04-05 docs/SOSP 2024/ir-description.tex
Add IR description

\section{\Treebeard{} IR Description}

% \begin{table*}[htb]
%   \centering
%   \resizebox{\linewidth}{!}{
%   \begin{tabularx}{\linewidth}{c l l l l}
%    \toprule
%    \textbf{Operation} & \textbf{Inputs} & \textbf{Outputs} & \textbf{Attributes} & \textbf{Description} \\
%    \midrule
%    \multirow{4}{*}{\texttt{predict\_ensemble}} & \textbf{data} & \textbf{result} & \textbf{ensemble} & Performs inference on the \textbf{data} using the model specified by\\
%                                                &               &                 & \textbf{predicate} & the \textbf{ensemble} attribute. The \textbf{schedule} attribute contains the \\
%                                                &               &                 & \textbf{schedule} & schedule described in Section \ref{sec:schedule}. \textbf{predicate} specifies the \\
%                                                 &               &                &                   & operator to use to evaluate nodes (Eg: $<$, $\leq$). \\
%    \midrule
%    \multirow{3}{*}{\texttt{walk\_decision\_tree}} & \textbf{trees[]} & \textbf{results[] }& \textbf{predicate}   & Represents an interleaved walk on all the element-wise pairs  \\ 
%                                                   & \textbf{rows[]}  &                    & \textbf{unrollDepth} & of \textbf{trees} and \textbf{rows}. \textbf{unrollDepth} specifies the number of \\
%                                                   &                  &                    &                      & hops to unroll. An array of tree walk results is returned.\\
%    \bottomrule
%   \end{tabularx}
%   }
%   \vskip 5pt
%   \caption{\label{Tab:Benchmarks}List of benchmark datasets and their parameters. 
%   The column \op{Leaf-biased} reports the number of leaf-biased trees per benchmark with $\langle\alpha =0.075, \beta =0.9 \rangle$. }
% \end{table*}

\begin{table*}[htb]
  \centering
  \resizebox{\linewidth}{!}{
  \begin{tabularx}{\linewidth}{c | l | l | l | l}
   \toprule
   \textbf{Operation} & \textbf{Inputs} & \textbf{Outputs} & \textbf{Attributes} & \textbf{Description} \\
   \midrule
   \multirow{4}{*}{\texttt{predictEnsemble}} & \textbf{rows[]} & \textbf{result} & \textbf{ensemble} & \multirow{4}{*}{\parbox{0.47\linewidth}{Performs inference on the data in \textbf{rows[]} 
                                                                                                     using the model specified by
                                                                                                     the \textbf{ensemble} attribute. The \textbf{schedule} attribute contains the 
                                                                                                     schedule described in Section \ref{sec:schedule}. \textbf{predicate} specifies the
                                                                                                     operator to use to evaluate nodes (Eg: $<$, $\leq$).}} \\
                                               &               &                 & \textbf{predicate} & \\
                                               &               &                 & \textbf{schedule} &  \\
                                               &               &                 &                   &  \\
   \midrule
   \midrule

   \multirow{3}{*}{\texttt{walkDecisionTree}} & \textbf{trees[]} & \textbf{results[] }& \textbf{predicate}   & \multirow{3}{*}{\parbox{0.47\linewidth}{Represents an interleaved walk on all the element-wise pairs 
                                                                                                                    of \textbf{trees} and \textbf{rows}. \textbf{unrollDepth} specifies the number of
                                                                                                                    hops to unroll. An array of tree walk results is returned.}} \\
                                                  & \textbf{rows[]}  &                    & \textbf{unrollDepth} &  \\
                                                  &                  &                    &                      & \\
   \midrule
   \multirow{2}{*}{\texttt{ensemble}} &  & \textbf{ensemble} & \textbf{model} & \multirow{2}{*}{\parbox{0.47\linewidth}{Represents the forest of trees that constitute the model. The   
                                                                                                                       \textbf{model} attribute contains the actual trees model.}} \\
                                      &  &                 &   & \\
   \midrule
   \multirow{2}{*}{\texttt{getTree}} & \textbf{ensemble} & \textbf{tree} &  & \multirow{2}{*}{\parbox{0.47\linewidth}{Get the tree at the specified index (\textbf{treeIndex}) from the \textbf{ensemble}.}} \\

                                      & \textbf{treeIndex} &                 &   & \\

   \midrule
   \multirow{2}{*}{\texttt{getTreeClassId}} & \textbf{ensemble} & \textbf{classId} &  & \multirow{2}{*}{\parbox{0.47\linewidth}{Get the class ID for the tree at index \textbf{treeIndex} 
                                                                                                                               in the \textbf{ensemble}. This is used for multi-class models.}} \\    
                                            & \textbf{treeIndex} &                 &   & \\

   \midrule
   \multirow{1}{*}{\texttt{getRoot}} & \textbf{tree} & \textbf{rootNode} &  & \multirow{1}{*}{\parbox{0.47\linewidth}{Get the root node of the specified tree.}} \\

   \midrule
   \multirow{2}{*}{\texttt{isLeaf}} & \textbf{tree} & \textbf{bool} &  & \multirow{2}{*}{\parbox{0.47\linewidth}{Returns a boolean value indicating whether \textbf{node} is a leaf of \textbf{tree}.}} \\   
                                    & \textbf{node} &                 &   & \\

   \midrule
   \multirow{2}{*}{\texttt{getLeafValue}} & \textbf{tree} & \textbf{value} &  & \multirow{2}{*}{\parbox{0.47\linewidth}{Returns the value of the leaf \textbf{node} in \textbf{tree}.}} \\   
                                    & \textbf{node} &                 &   & \\

   \midrule
   \multirow{3}{*}{\texttt{traverseTreeTile}} & \textbf{trees[]} & \textbf{nodes[] }& \textbf{predicate}   & \multirow{3}{*}{\parbox{0.47\linewidth}{Represents an interleaved traversal of the   
                                                                                                                    nodes in \textbf{nodes} based on the data in \textbf{rows}. \textbf{predicate} specifies 
                                                                                                                    the operator to use to evaluate nodes.}} \\
                                                  & \textbf{nodes[]}  &             &               &  \\
                                                  & \textbf{rows[]}   &             &               & \\

   \midrule
   \multirow{3}{*}{\texttt{cacheTrees}} & \textbf{ensemble} & \textbf{ensemble}&     & \multirow{3}{*}{\parbox{0.47\linewidth}{Cache the trees in the \textbf{ensemble} between the specified \textbf{start} and \textbf{end} indices.   
                                                                                                                    The returned \textbf{ensemble} has the specified trees cached.}} \\
                                                  & \textbf{start}  &             &               &  \\
                                                  & \textbf{end}   &             &               & \\

   \midrule
   \multirow{3}{*}{\texttt{cacheRows}} & \textbf{rows[]} & \textbf{cachedRows[]}&     & \multirow{3}{*}{\parbox{0.47\linewidth}{Cache the rows in \textbf{rows[]} between the specified \textbf{start} and \textbf{end} indices.   
                                                                                                                    Returns an array of cached rows \textbf{cachedRows[]}.}} \\
                                                  & \textbf{start}  &             &               &  \\
                                                  & \textbf{end}   &             &               & \\

   \midrule
   \midrule
   \multirow{3}{*}{\texttt{loadThreshold}} & \textbf{buffer} & \textbf{threshold}&     & \multirow{3}{*}{\parbox{0.47\linewidth}{Load the threshold value for the node specified by \textbf{nodeIndex} 
                                                                                                                                      in the tree specified by \textbf{treeIndex} from \textbf{buffer}.
                                                                                                                                      Returns the loaded threshold.}} \\   
                                                  & \textbf{treeIndex}  &             &               &  \\
                                                  & \textbf{nodeIndex}   &             &               & \\

   \midrule
   \multirow{3}{*}{\texttt{loadFeatureIndex}} & \textbf{buffer} & \textbf{threshold}&     & \multirow{3}{*}{\parbox{0.47\linewidth}{Load the feature index for the node specified by \textbf{nodeIndex} 
                                                                                                                                      in the tree specified by \textbf{treeIndex} from \textbf{buffer}.
                                                                                                                                      Returns the loaded feature index.}} \\   
                                                  & \textbf{treeIndex}  &             &               &  \\
                                                  & \textbf{nodeIndex}   &             &               & \\

  \bottomrule
  \end{tabularx}
  }
  \vskip 5pt
  \caption{\label{Tab:IRSpecification} List of all the operations in the \Treebeard{} MLIR dialect. These operations are used 
  in conjunction with operations from other MLIR dialects like \op{scf}, \op{arith}, \op{gpu} etc. to represent and optimize 
  decision tree inference.}
\end{table*}

### 2024-04-06 docs/SOSP 2024/motivation.tex
Rewrite motivation

@@
 
-Our main aim while designing \Treebeard{} was to unify the diverse set of implementation
-strategies that have been used in existing systems for decision tree inference. Some 
-differences in these systems are as follows:
+% Our main aim while designing \Treebeard{} was to unify the diverse set of implementation
+% strategies that have been used in existing systems for decision tree inference. Some 
+% differences in these systems are as follows:
+% \begin{itemize}
+%   \item Decision tree inference is run on several platforms including CPUs and GPUs. The 
+%   implementations used on each of these platforms are different and the techniques used
+%   to optimize them are also different.
+%   \item A diverse set techniques have been proposed for optimization of decision tree 
+%   inference on CPUs and GPUs \cite{VPred, Tahoe, Treelite, XGBoost, Hummingbird, QuickScorer, FIL}. 
+%   No system exists that unifies the disparate optimizations implemented in these systems.
+%   \item A very extensive design space of optimizations exists for decision tree inference
+%   outside the few that have been proposed in the literature. However, currently no 
+%   system exists that is capable of exploring this space and identifying the best set 
+%   of parameters to use for a given model and platform.
+%   \item Different systems use different in-memory representations for the model. For example,
+%   XGBoost uses a sparse representation, RAPIDs FIL uses what is called the reorg representation 
+%   and Tahoe uses a variation of the reorg representation. Currently, systems implement 
+%   inference kernels that are tied to a single representation of the model. Again, this means
+%   that no current system can explore different combinations of in-memory representations 
+%   and optimizations.
+% \end{itemize}
+% At high-level, to make \Treebeard{} capable of unifying these differences, we design 1)  
+% expressive intermediate representations that can represent and compose several proposed 
+% optimizations 2) a scheduling language that specifies the structure of the
+% generated code and 3) a plugin mechanism with which different in-memory representations
+% can be composed with different optimizations. Finally, we develop a heuristic to
+% explore the extensive optimization space that \Treebeard{}'s design enables.
+
+While a diverse set techniques have been proposed for optimization of decision tree 
+inference on CPUs and GPUs \cite{VPred, Tahoe, Treelite, XGBoost, Hummingbird, QuickScorer, FIL},
+a very extensive design space of optimizations exists 
+outside what has been proposed in the literature. Furthermore, decision tree inference 
+is run on several platforms including CPUs and GPUs. The implementations used on each of 
+these platforms are different and the techniques used to optimize them are different.
+To make matters even more complicated, several in-memory representations
+have been proposed for decision tree models. For example, XGBoost\cite{XGBoost} uses a sparse representation,
+RAPIDs FIL\cite{FIL} uses what is called the reorg representation and Tahoe uses a variation of the reorg
+representation. 
+\TODO{Can we add some numbers here to show that different models/batch sizes need different optimizations?}
+
+To solve the problems of exploring the design space of optimizations for decision tree
+inference and enabling portable performance, we build several techniques in \Treebeard{}, 
+an open source compiler infrastructure for decision tree inference. To make \Treebeard{}
+capable of unifying these different techniques and targets, we do the following. 
 \begin{itemize}
-  \item Decision tree inference is run on several platforms including CPUs and GPUs. The 
-  implementations used on each of these platforms are different and the techniques used
-  to optimize them are also different.
-  \item A diverse set techniques have been proposed for optimization of decision tree 
-  inference on CPUs and GPUs \cite{VPred, Tahoe, Treelite, XGBoost, Hummingbird, QuickScorer, FIL}. 
-  No system exists that unifies the disparate optimizations implemented in these systems.
-  \item A very extensive design space of optimizations exists for decision tree inference
-  outside the few that have been proposed in the literature. However, currently no 
-  system exists that is capable of exploring this space and identifying the best set 
-  of parameters to use for a given model and platform.
-  \item Different systems use different in-memory representations for the model. For example,
-  XGBoost uses a sparse representation, RAPIDs FIL uses what is called the reorg representation 
-  and Tahoe uses a variation of the reorg representation. Currently, systems implement 
-  inference kernels that are tied to a single representation of the model. Again, this means
-  that no current system can explore different combinations of in-memory representations 
-  and optimizations.
+  \item We design a scheduling language that encapsulates various optimization techniques
+  and controls the structure of the generated code.
+  \item We design an MLIR dialect to represent and optimize reductions and use this 
+  dialect within \Treebeard{} to enable the generation of different variants of 
+  inference routines.
+  \item We extend \Treebeard{}'s intermediate representations to include operations like caching.
+  We were able to easily reuse and extend \Treebeard{}'s IR as it was built as an MLIR dialect.
+  \item We design a plugin mechanism with which different in-memory representations
+  can be composed with different optimizations.  
 \end{itemize}
-At high-level, to make \Treebeard{} capable of unifying these differences, we design 1)  
-expressive intermediate representations that can represent and compose several proposed 
-optimizations 2) a scheduling language that specifies the structure of the
-generated code and 3) a plugin mechanism with which different in-memory representations
-can be composed with different optimizations. Finally, we develop a heuristic to
-explore the extensive optimization space that \Treebeard{}'s design enables.

### 2024-04-08 docs/SOSP 2024/overview.tex
Add GPU compilation overview figure
Add HIR MIR optimization section
Change overview section

@@
 
-\begin{figure*}[htb]
+\begin{figure}[htb]
   \centering
-  \includegraphics[width=\linewidth]{figures/OverviewExample_New.png}
-  \vskip 10pt
-  \caption{\Treebeard{} IR lowering and optimization details: the three abstraction levels in \Treebeard{}'s IR are shown. The
-           high level IR is a tree-based IR to perform model level optimization, the mid-level IR is for
-           loop optimizations that are independent of memory layout and the low level IR allows us to perform
-           vectorization and other memory layout dependent optimizations.}
-  \label{Fig:LoweringExample}
-\end{figure*}
+  \includegraphics[width=\linewidth]{figures/HIR.PNG}
+  \caption{The representation of a model in high-level IR and the model 
+  after trees are padded and reordered. }
+  \label{Fig:HIRExample}
+\end{figure}
 
+Table \ref{Tab:IRSpecification} lists the operations in the three IRs. 
 In HIR, the model is represented as a collection of binary trees. This abstraction
 allows the implementation of optimizations that require the manipulation of the model
-or its constituent trees. In Figure \ref{Fig:LoweringExample}, \circled{A}
-shows this representation for a model with three trees. 
-In HIR, \Treebeard{} tiles tree nodes together to convert
-the binary tree to an n-ary tree as shown in \circled{B}. Trees are also 
-reordered to enable better code generation and padded to allow more 
-efficient traversal as shown in \circled{C}. While these are the optimizations currently 
-implemented in \Treebeard{}, these are by no means the only ones that 
-are enabled by HIR. It is not difficult to imagine optimizations 
-such as tree pruning for specified accuracy levels for example.
-These optimizations and rewrites that are performed on the domain-specific 
-HIR would have been much on a traditional loop based IR or even in 
-other IRs within \Treebeard{}.
+or its constituent trees. Figure \ref{Fig:HIRExample} shows this representation for a model 
+with four trees and how the model can be transformed. We use this model as a running example 
+for the rest of the section.
+After these model-level optimizations are performed on the HIR, the 
+code is lowered to the mid-level IR (MIR) as dictated by a user 
+specified \emph{\textbf{schedule}}. The schedule determines how
+the iteration space that goes over the trees and input rows is to be traversed
+by specifying how loops are to be tiled, which loops are to be
+parallelized, which loops are to be mapped to GPU grid and block dimensions etc.
+(Details in Section \ref{sec:schedule}). While MIR is a loop-based IR that 
+explicitly encodes details of the iteration space has to be traversed, 
+it still abstracts details about the in-memory representation of the model.
 
-After these model-level optimizations are performed on the HIR, the 
-code is lowered to the mid-level IR (MIR) as dictated by a user specified schedule
-(\circled{C} to \bluecircled{D}) in Figure \ref{Fig:LoweringExample}). The schedule specifies how
-the iteration space that goes over the trees and input rows is to be traversed. 
-It specifies how the iteration space is to be tiled, which loops are to be
-parallelized, which loops are to be mapped to GPU grid and block dimensions etc.
-(Details in Section \ref{Sec:SchedulingLang}). MIR is a loop-based IR that 
-explicitly encodes details of the iteration space has to be traversed. However, 
-it still abstracts details about the in-memory representation of the model. 
-Optimizations such as tree-walk unrolling and interleaving are performed 
-on the MIR (\bluecircled{E}). Subsequently, reduction operations are split and rewritten 
-to correctly and explicitly implement reduction in the presence of 
-parallel loops (\bluecircled{F}). More details of this process that we call 
-\emph{legalization} are in Section \ref{Sec:Reduction}. Another important 
-point to note is that MIR is independent of the target processor and therefore
-all optimizations on MIR can be reused across CPU and GPU compilation.
+In the lowered MIR, the compiler uses the \op{reduce} op from 
+the \op{reduction} dialect we design (details in Section \ref{sec:reduction})
+to represent reduction operations. The lowering of the \op{reduce} operation 
+involves introducing temporary buffers and splitting the reduction 
+to correctly implement reduction in the presence of 
+parallel loops. This process, that we call \emph{legalization}, is 
+described in Section \ref{sec:reduction}.
 
 The MIR is then further lowered to a low-level IR (LIR). This is the
-level at which the compilation pipeline diverges for CPUs and GPUs. In 
+level at which the compilation pipeline diverges for different targets. In 
 the GPU compilation pipeline, the required memory transfers and kernel
-invocations are inserted into the LIR (\circled{J}). Additionally, buffers 
+invocations are inserted into the LIR. Additionally, buffers 
 to hold model values are inserted and tree operations are lowered to
@@
 a plugin mechanism where different in-memory representations can 
-be added to the compiler by implementing an interface. These plugins
-provide information required for the lowering of MIR to LIR as well as 
-the lowering to LLVM IR as shown in Figure \ref{Fig:LoweringExample}.
-Again, a significant amount of code is shared between the
-CPU and GPU pipelines for representations that are common between them
-(Array and Sparse representations). Vectorization of tree traversals 
+be added to the compiler by implementing an interface 
+(Section \ref{sec:representations}). Vectorization of tree traversals
 is also explicitly represented in LIR.
+
+In the remainder of this section, we show using our running example
+from Figure \ref{Fig:HIRExample} how different schedules can be used to 
+lower the model to MIR and LIR. We also show how the \op{reduce} operation
+is lowered and legalized.
+
+First, we consider the the schedule that processes one tree at a time 
+for all input rows and unrolls the tree walks for all trees in the model. 
+We assume that the trees have been reordered so that all trees with 
+the same depth are together as shown in Figure \ref{Fig:HIRExample}.
+The schedule splits the loop over the trees into two loops -- one that
+iterates over the first two trees and the second that iterates over the
+last two trees. The schedule the unrolls the tree walks for each tree.
+\begin{lstlisting}[style=c++]
+  reorder(tree, batch)  
+  split(tree, t0, t1, 2)
+  unrollWalk(t0, 1)
+  unrollWalk(t1, 2)
+\end{lstlisting}
+
+\Treebeard{} represents the set of loops as a tree of loops, where the children 
+of a loop node are the nodes for the immediately contained loops. Each schedule 
+primitive modifies this tree in some way. Additionally,
+the lineage of each of the loops is maintained. This allows 
+the compiler to automatically infer the ranges for all loops. To lower the
+HIR to MIR, the compiler traverses the schedule tree and generates the
+corresponding loops. The MIR for the example model, generated using the above   
+schedule is as follows. 
+\begin{lstlisting}[style=c++]
+  model = ensemble(...)
+  for t0 = 0 to 2 step 1 {
+    T = getTree(ensemble, t0)
+    for batch = 0 to BATCH_SIZE step 1 {
+      treePred = walkDecisionTree(T, 
+                    input[batch]) <unrollDepth = 1>
+      reduce(result[batch], treePred)
+    }
+  }
+  for t1 = 2 to 4 step 1 {
+    T = getTree(ensemble, t0)
+    for batch = 0 to BATCH_SIZE step 1 {
+      treePred = walkDecisionTree(T,
+                    input[batch]) <unrollDepth = 2>
+      reduce(result[batch], treePred) <'+', 0.0>
+    }
+  }
+\end{lstlisting}
+Subsequent lowering rewrites the \op{walkDecisionTree} operations to a 
+series of \op{traverseTile} operations, in this case one for the 
+first instance and two for the second (as specified by the unroll depth
+attribute). \Treebeard{} also determines that the \op{reduce} operation
+can be implemented by a simple addition operation as there are no 
+surrounding parallel loops (The legalization of reductions is described 
+in detail in Section \ref{sec:reduction}).
+
+We now show how we would compile our example model to a GPU. 
+Additionally, we divide the trees into two sets and process each set
+in parallel. We also unroll the tree walks for each tree.
+The schedule to accomplish this is as follows.
+\begin{lstlisting}[style=c++]
+  tile(tree, t0, t1, 2)
+  reorder(batch, t1, t0)
+  split(t0, t0_1, t0_2, 2)
+  unrollWalk(t0_1, 1)
+  unrollWalk(t0_2, 2)
+  gpuDimension(batch, grid.x)
+  gpuDimension(t1, block.x)
+\end{lstlisting}
+This schedule processes on input row per thread block 
+and splits the trees into two sets, each of which contains 
+two trees. The MIR generated is as follows. 
+\begin{lstlisting}[style=c++]
+  model = ensemble(...)
+  par.for batch = 0 to BATCH_SIZE step 1 <grid.x> {
+    par.for t1 = 0 to 2 step 1 <block.x> {
+      for t0_1 = 0 to 2 step 2 {
+        T = getTree(ensemble, t0_1 + t1)
+        treePred = walkDecisionTree(T, 
+                        input[batch]) <unrollDepth = 1>
+        reduce(result[batch], treePred)
+      }
+      for t0_2 = 2 to 4 step 2 {
+        T = getTree(ensemble, t0_2 + t1)
+        treePred = walkDecisionTree(T,
+                        input[batch]) <unrollDepth = 2>
+        reduce(result[batch], treePred) <'+', 0.0>
+      }
+    }
+  }
+\end{lstlisting}
+
+In the case of this example, the lowering passes for the \op{reduce}
+operation determines that several paralel iterations of the \op{t1} 
+loop accumulate into the same element of the \op{result} array.
+Therefore, the reduction is rewritten so that each parallel 
+iteration accumulates into a different array element by introducing 
+a temporary buffer as follows.
+\begin{lstlisting}[style=c++]
+  float tempResults[2][BATCH_SIZE]
+  model = ensemble(...)
+  par.for batch = 0 to BATCH_SIZE step 1 <grid.x> {
+    par.for t1 = 0 to 2 step 1 <block.x> {
+      for t0_1 = 0 to 2 step 2 {
+        T = getTree(ensemble, t0_1 + t1)
+        treePred = walkDecisionTree(T, 
+                      input[batch]) <unrollDepth = 1>
+        reduce(tempResults[t1][batch], treePred)
+      }
+      for t0_2 = 2 to 4 step 2 {
+        T = getTree(ensemble, t0_2 + t1)
+        treePred = walkDecisionTree(T,
+                      input[batch]) <unrollDepth = 2>
+        reduce(tempResults[t1][batch], treePred) <'+', 0.0>
+      }
+    }
+    result[batch] = reduce_dimension(result[0], 1)
+  }
+\end{lstlisting}
+Here, partial results are accumulated into \op{tempResults} and then
+reduced across the \op{t1} dimension (represented by the \op{reduce\_dimension}
+where the second argument represents the dimension number to reduce across)
+to get the final result.
+
+Finally, since this computation is being targeted to the GPU, the 
+parallel loops are rewritten into kernel calls and the  required 
+GPU memory allocations and memory transfers are introduced.
+
+As is evident from these examples, the structure of the loop 
+nest in the inference routine can get quite complex even 
+for simple optimizations. Writing these routines by hand 
+is error prone and time-consuming. Building a principled code 
+generator for these routines is the best way to explore the
+vast design space. 
+% Show simple one tree lowering for CPU with walk unrolling
+%    Schedule, MIR, 
+% Show parallelization across trees and GPU target
 
@@
 \end{enumerate}
+
+%**************ROUGH*********************
+
+% Another important 
+% point to note is that MIR is independent of the target processor and therefore
+% all optimizations on MIR can be reused across CPU and GPU compilation.
+
+% \begin{figure*}[htb]
+%   \centering
+%   \includegraphics[width=\linewidth]{figures/OverviewExample_New.png}
+%   \vskip 10pt
+%   \caption{\Treebeard{} IR lowering and optimization details: the three abstraction levels in \Treebeard{}'s IR are shown. The
+%            high level IR is a tree-based IR to perform model level optimization, the mid-level IR is for
+%            loop optimizations that are independent of memory layout and the low level IR allows us to perform
+%            vectorization and other memory layout dependent optimizations.}
+%   \label{Fig:LoweringExample}
+% \end{figure*}
+
+% \begin{figure*}[htb]
+%   \centering
+%   \includegraphics[width=\linewidth]{figures/GPUCompilationOverview-cropped.pdf}
+%   \vskip 10pt
+%   \caption{\Treebeard{} IR lowering and optimization details: the three abstraction levels in \Treebeard{}'s IR are shown. The
+%            high level IR is a tree-based IR to perform model level optimization, the mid-level IR is for
+%            loop optimizations that are independent of memory layout and the low level IR allows us to perform
+%            vectorization and other memory layout dependent optimizations.}
+%   \label{Fig:GPULoweringExample}
+% \end{figure*}
+
+% In HIR, the model is represented as a collection of binary trees. This abstraction
+% allows the implementation of optimizations that require the manipulation of the model
+% or its constituent trees. In Figure \ref{Fig:LoweringExample}, \circled{A}
+% shows this representation for a model with three trees. 
+% In HIR, \Treebeard{} tiles tree nodes together to convert
+% the binary tree to an n-ary tree as shown in \circled{B}. Trees are also 
+% reordered to enable better code generation and padded to allow more 
+% efficient traversal as shown in \circled{C}. While these are the optimizations currently 
+% implemented in \Treebeard{}, these are by no means the only ones that 
+% are enabled by HIR. It is not difficult to imagine optimizations 
+% such as tree pruning for specified accuracy levels for example.
+% These optimizations and rewrites that are performed on the domain-specific 
+% HIR would have been much on a traditional loop based IR or even in 
+% other IRs within \Treebeard{}.
+
+% After these model-level optimizations are performed on the HIR, the 
+% code is lowered to the mid-level IR (MIR) as dictated by a user specified schedule
+% (\circled{C} to \bluecircled{D}) in Figure \ref{Fig:LoweringExample}). The schedule specifies how
+% the iteration space that goes over the trees and input rows is to be traversed. 
+% It specifies how the iteration space is to be tiled, which loops are to be
+% parallelized, which loops are to be mapped to GPU grid and block dimensions etc.
+% (Details in Section \ref{Sec:SchedulingLang}). MIR is a loop-based IR that 
+% explicitly encodes details of the iteration space has to be traversed. However, 
+% it still abstracts details about the in-memory representation of the model. 
+% Optimizations such as tree-walk unrolling and interleaving are performed 
+% on the MIR (\bluecircled{E}). Subsequently, reduction operations are split and rewritten 
+% to correctly and explicitly implement reduction in the presence of 
+% parallel loops (\bluecircled{F}). More details of this process that we call 
+% \emph{legalization} are in Section \ref{Sec:Reduction}. Another important 
+% point to note is that MIR is independent of the target processor and therefore
+% all optimizations on MIR can be reused across CPU and GPU compilation.
+
+% The MIR is then further lowered to a low-level IR (LIR). This is the
+% level at which the compilation pipeline diverges for CPUs and GPUs. In 
+% the GPU compilation pipeline, the required memory transfers and kernel
+% invocations are inserted into the LIR (\circled{J}). Additionally, buffers 
+% to hold model values are inserted and tree operations are lowered to
+% explicitly refer to these buffers. This lowering is controlled by 
+% a plugin mechanism where different in-memory representations can 
+% be added to the compiler by implementing an interface. These plugins
+% provide information required for the lowering of MIR to LIR as well as 
+% the lowering to LLVM IR as shown in Figure \ref{Fig:LoweringExample}.
+% Again, a significant amount of code is shared between the
+% CPU and GPU pipelines for representations that are common between them
+% (Array and Sparse representations). Vectorization of tree traversals 
+% is also explicitly represented in LIR.

### 2024-04-08 docs/SOSP 2024/reduction.tex
Change overview section

@@
 \section{Reductions : Representation, Optimization and Lowering}
-
+\label{sec:reduction}
 Currently, existing reduction support in MLIR is insufficient 

### 2024-04-08 docs/SOSP 2024/abstract.tex
Add abstract and XGB comparison

\begin{abstract}
  Decision tree based models are generated by some of the most commonly 
  used machine learning algorithms like gradient boosting and random forests.
  These algorithms are used very widely in practice and the resulting models
  are deployed to production at scale. The models are served on a variety of 
  machines including CPUs and GPUs and inference performance on all these 
  machines is an important consideration. On CPUs, models are served using
  either libraries such as XGBoost, Sklearn, and LightGBM or specialized 
  compilers like \TreebeardOLD{}, Treelite, and Hummingbird. On GPUs, XGBoost 
  and RAPIDs FIL are popular libraries. However, these libraries and compilers
  do not provide portable performance across the targets of interest. They 
  also leave significant optimization opportunities unexplored.

  In this paper, we present the design of \Treebeard{}, a compiler infrastructure 
  for decision tree based models that significantly expands the design space 
  that can be explored and provides portable performance across CPU and GPU 
  targets. We rearchitect the open-source \Treebeard{} infrastructure 
  and significantly extend it to enable high-performance code generation 
  across target processors. Specifically, we design a scheduling language 
  that encapsulates the large optimization space for decision tree inference
  and develop techniques to explore this space. Further,
  we design techniques to represent and optimize parallel reductions, extend
  \TreebeardOLD{}'s intermediate representations to include operations like caching
  and rearchitect \TreebeardOLD{} to allow in-memory representations to be built 
  as plugins. \TODO{Need to say we investigate how to structure a compiler for 
  decision tree inference that can target multiple hardware platforms}

  We implement \Treebeard{} using the MLIR compiler infrastructure and
  compare the performance of the generated code against state-of-the-art
  systems on CPUs and GPUs. On GPUs we compare performance against XGBoost,
  RAPIDs FIL, and Tahoe. We find that \Treebeard{} is significantly faster
  than other systems that target GPUs across a wide range of batch sizes 
  and models on different NVIDIA GPUs. \Treebeard{} is an order of magnitude faster than XGBoost and
  about 2-3$\times$ faster than RAPIDs FIL and Tahoe on average. \TODO{Make 
  this more precise. But how?}. We also show how \Treebeard{} can generate  
  high-performance code on other GPUs by running the benchmarks on an AMD GPU.
  On CPUs, we compare against \TreebeardOLD{} and show how the  
  optimization opportunities exposed by \Treebeard{}'s scheduling 
  language allow it to generate higher performance code by specializing  
  code for different batch sizes.
\end{abstract}

### 2024-04-08 docs/SOSP 2024/intro.tex
Add abstract and XGB comparison

@@
 \label{sec:intro}
-We are in the midst of a hardware revolution, a new golden age for computer architecture~\cite{TuringLecture}.
-The last decade has seen a shift in architectural paradigms, with the rise of GPUs and accelerators.
-This shift has been driven by the necessity to innovate in the post Moore's law and Dennard's scaling era.
-This transformation has also played a significant role in the success of modern deep learning models, as they enable scaling model training and inference to a massive number of threads.
-Such scalability would be essential for all performance critical applications, including other machine learning models that need to scale with increasing data sizes and model complexities.
+We are in the midst of a hardware revolution, a new golden age for computer architecture~\cite{GoldenAge}. The last 
+decade has seen a shift in architectural paradigms, with the rise of GPUs and accelerators. This shift has been driven
+by the necessity to innovate in the post Moore's law and Dennard's scaling era. This transformation has also played 
+a significant role in the success of modern deep learning models, as they enable scaling model training and inference 
+to a massive number of threads. Such scalability would be essential for all performance critical applications, including
+other machine learning models that need to scale with increasing data sizes and model complexities. 
 %need to accelerate machine learning workloads.
@@
 %Despite these recent advancements, one dominant family of models, namely decision forest models have not seen the same level of acceleration.
-Decision forest models remain the mainstay for machine learning over tabular data~\cite{tabular}. 
+Decision forest models remain the mainstay for machine learning over tabular data~\cite{DLNotAllYouNeed,TreebasedOutperformDL}. 
 Their robustness, interpretability, and ability to handle missing data make them a popular choice for a wide range of applications. 

### 2024-04-09 docs/SOSP 2024/abstract.tex
abstract/title pass

@@
+\begin{comment}
 \begin{abstract}
@@
 \end{abstract}
+\end{comment}
+\begin{abstract}
+
+  With the rapid proliferation of machine learning, the demand for on-device model inference has surged.
+Concurrently, the hardware ecosystem is evolving rapidly, leading to an increasing heterogeneity in client machines.
+  %While CPUs have been the mainstay for machine learning inference, the availability of GPUs holds promise to scale to bigger and more powerful models. 
+  This paper is motivated by the problems encountered when targeting inference of decision tree based models, the most popular models on tabular data, to run at peak performance on diverse client hardware.
+We evaluated existing solutions and found that they do not provide portable performance across different hardware targets.
+  %Decision tree based models are widely used in practice due to their robustness, interpretability, and ability to handle missing data.
+
+  To address this we present the design of \Treebeard{}, a schedule guided compiler infrastructure 
+  for decision tree based models that searches over a large design space 
+  to find high performance schedules on a diverse set of CPU and GPU 
+  targets. 
+  %We re-architect the open-source \TreebeardOLD{} infrastructure and significantly extend it to enable high-performance code generation across target processors. 
+  \Treebeard{} has two core components. A scheduling language 
+  that encapsulates the large optimization space for decision tree inference, 
+  and techniques to efficiently explore this space.
+  \TODO{include some keywords from the language and space complexity}.
+  %We also design a set of heuristics that can find near-optimal solutions in quick time.
+  Second, an optimizing multi-level compiler that can generate code based on the schedule. 
+  For the latter, we re-architect the open-source \TreebeardOLD{} CPU compiler to support schedule guided compilation and
+add support for GPU code generation.
+We further enhance the compiler with fundamental new optimization interfaces for caching, parallel reduction, and a plug-in mechanism to explore different in-memory representations of trees.
+
+  %The compiler builds on an existing compiler for CPUs, \TreebeardOLD{}, also add support for caching, parallel reduction and a plug-in mechanism to explore different in-memory representations of trees.
+
+  We evaluate \Treebeard{} on \TODO{kr:Can we quote a number} a large number of diverse models and demonstrate that the best schedule varies drastically with model, batch size, and target hardware. 
+  Despite this, the scheduling heuristic is able to quickly find near optimal schedules while searching over a small fraction of the total search space.
+  In terms of performance, \Treebeard{} generated code  
+   is an order of magnitude faster than XGBoost and
+  about 2-3$\times$ faster on average than RAPIDs FIL and Tahoe. \TODO{Make 
+  this more precise. But how?}. While these systems only generate code for Nvidia, \Treebeard{} achieves competent performance on AMD GPUs as well. 
+  On CPUs, we compare against \TreebeardOLD{} and show how the  
+  optimization opportunities exposed by \Treebeard{}'s scheduling 
+  language allow it to generate higher performance code by specializing  
+  code for different batch sizes.
+  \TODO{(numbers for CPU performance?)}
+\end{abstract}

### 2024-04-09 docs/SOSP 2024/intro.tex
abstract/title pass

@@
 This paper is motivated by the need to accelerate decision tree inference to achieve portable performance on a variety of client hardware.
-In particular, we focus on portable performance across a range of commodity CPUs and low-to-mid-range GPUs, a class of hardware that has seen widespread adoption across client/edge devices used for inference.
+In particular, we focus on a range of commodity CPUs and low-to-mid-range GPUs, a class of hardware that has seen widespread adoption across client/edge devices used for inference.
 
@@
 Despite the simplicity of the model and the availability of multiple sources of coarse grain parallelism (parallelism across inputs in a batch and parallelism across trees), existing systems do not consistently scale well across different models.
-This is because existing systems exploit a limited set of optimizations and often specialize the implementation to a specific hardware platform, which limits their portability.
+This is because existing systems exploit a limited set of optimizations and often specialize the implementation to a specific hardware platform, and this limits their portability.
 
@@
 Evaluation on a diverse set of models highlights that the best implementation often requires a careful combination of many optimization strategies.
-For example, managing the working set of nodes across a set of trees is critical to get scalable performance.
+For example, as we simultaneously traverse multiple trees, managing the working set of nodes is critical to get scalable performance.
 This requires a combination of techniques like data layout optimizations, loop transformations, and memory access optimizations.
-Existing systems each use specific data layouts for the underlying tree (XGBoost uses a sparse representation, RAPIDs FIL uses what is called the reorg representation
-and Tahoe uses a variation of the reorg representation) and propose simple loop transformations (XGBoost uses \dots()) that together determine the accesses patterns and the working set.
+Existing systems each use specific data layouts for the underlying tree (XGBoost~\cite{XGBoost} uses a sparse representation, RAPIDs~\cite{Rapids} FIL~\cite{FTL} uses what is called the reorg representation
+and Tahoe~\cite{Tahoe} uses a variation of the reorg representation) and propose simple loop transformations (XGBoost uses \dots() \TODO{complete}) that together determine the accesses patterns and the working set.
 They do not perform consistently well across different models and hardware platforms.
 Managing the working set is just one of several optimizations that are critical for high-performance decision tree inference.
- \TODO{Improve the next line}
+ \TODO{Improve/drop the next line}
  Others include tree ordering,  incremental reduction of predicted values, interleaving across multiple trees, and so on.

### 2024-04-09 docs/SOSP 2024/related.tex
Add related work section

\section{Related Work}
\label{Sec:Related}
While several optimization strategies for decision tree based models have been 
studied in the literature, to the best of our knowledge, no systems that are 
capable of exploring the extensive optimization space exist. We describe related work 
and compare these systems to \Treebeard{} in this section.

\emph{Decision Tree Inference Systems for GPUs:} 
Tahoe\cite{Tahoe} is a system that implements high-performance library routines and a 
performance model for tree inference on GPUs. Tahoe is a library-based system that picks 
between four predefined strategies to implement decision tree inference on GPUs.
In comparison, \Treebeard{} explores a much larger set of implementation options 
because it is a compiler. \Treebeard{} can also explore different in-memory 
representations for models. Also, \Treebeard{} generates code that is specific to 
a particular model, specializing both the parallelism (by deciding the thread block 
structure on a per model basis) and the kernel code itself by 
performing optimizations like tree walk unrolling and interleaving. In contrast, Tahoe 
uses a library-based approach, and cannot generate code tailored to a model 
like \Treebeard{} does.

RAPIDS FIL\cite{FIL} is a library that implements decision tree inference on GPUs
and is the most widely used production system for decision tree inference. While 
FIL does implement some heuristics to pick a good configuration for every model, 
these techniques are limited and the library essentially uses a single strategy 
and in-memory representation for all models. XGBoost \cite{XGBoost} also implements GPU 
support\cite{XGBGPU}. Again, XGBoost uses a single strategy and in-memory representation
for all models. In contrast, \Treebeard{} is a compiler that can explore a much larger
optimization space and can generate code that is tailored to a specific model.

% Finally, all of Tahoe's optimizations are designed to address GPU specific problems. 
% For example, it uses an elaborate and opaque heuristic based on locality sensitive hashing 
% to coalesce accesses to the GPU global memory. We believe that the tree tiling
% infrastructure described in this paper will allow us to deterministically 
% coalesce accesses to GPU global memory.
%As described previously, Hummingbird\cite{Hummingbird} uses tensor operations to perform decision tree inference. 
% While compiling decision tree inference to GPUs presents a distinct set of challenges,
% we believe, we can achieve this while reusing much of \Treebeard{}'s current infrastructure.
% Specifically, we expect that most of the HIR and MIR
% optimizations described in this paper will carry over directly while LIR 
% optimizations and in-memory representations will need to be retargeted.
% This can be the subject of a separate future work.

\emph{Decision Tree Ensemble Compilers:}
Several compilers for decision tree ensembles have been proposed in the 
literature \cite{Treelite, Treebeard, Hummingbird}. \TreebeardOLD{} and Treelite
exclusively target CPUs and all their optimizations are designed purely for 
performance on CPUs. 
Treelite\cite{Treelite} is a model compiler that only supports 
generation of simple \op{if-else} style code. Its code generator is hard-coded 
to expand trees in an ensemble into a series of \op{if-else} statements (one for 
each node in a tree). Due to this, extending Treelite and reusing it to 
target GPUs is not possible.

\TreebeardOLD{} is the work most closely related to \Treebeard{}. While we 
build on top of \TreebeardOLD{}, \Treebeard{} is a significant enhancement 
over \TreebeardOLD{}. Most importantly, we re-architect \TreebeardOLD{} in 
order to generate code for multiple target processors. Specifically, we 
introduce the scheduling language and schedule exploration 
while also enhancing the intermediate representations and support for parallelizing 
across trees through the implementation of a novel MLIR reduction dialect.
% It cannot perform any of the optimizations \Treebeard{} is designed to perform.

Hummingbird\cite{Hummingbird} is a compiler that compiles traditional ML models
to tensor operations thereby enabling them to be run on tensor-based frameworks like
TensorFlow. Hummingbird can target both CPUs
and GPUs. Its stated aim is to allow traditional ML based applications to easily 
leverage progress in tensor compilers. However, as was shown earlier~\cite{Treebeard},
tensor operations are not the most efficient way to implement decision tree inference
and the performance of code generated by Hummingbird is significantly lower than 
the code generated by other frameworks.

% Hummingbird\cite{Hummingbird} compiles traditional ML models to make use of tensor primitives so that
% they can be integrated into tensor-based frameworks like TensorFlow \cite{TensorFlow}.
% While Hummingbird does compile decision tree based models to both CPUs and GPUs, the primary aim of 
% the system is to leverage progress in tensor compilers by representing ML models as tensor operations. 
% Consequently, Hummingbird does not perform any optimizations tailored to decision trees. It 
% also does not perform model specific optimizations that are enabled by \Treebeard{}'s abstractions.

\emph{Libraries:} The most popular systems for decision tree-based models 
are libraries. On CPUs, XGBoost\cite{XGBoost}, LightGBM\cite{LightGBM} and
scikit-learn\cite{Sklearn} are extremely popular. These libraries implement both 
training and inference. On GPUs, RAPIDs FIL\cite{FIL} is the most commonly used 
library. However, as mentioned in Section \ref{sec:intro}, none of these systems
provide portable performance across different target machines.
\TODO{Write about the PACT paper and whether our scheduling language can represent 
all the schedules they propose.}
Other systems that hide dependency stalls by interleaving tree walks\cite{VPred},
implement optimized algorithms for tree inference\cite{QuickScorer, QuickScorer1}
and predict cache performance of decision tree ensembles on CPUs\cite{CacheConscious1, CacheConscious2}
have been proposed in prior work. However, these systems are limited to CPUs.
Some systems have been proposed 
to parallelize decision tree training on CPUs and GPUs\cite{Jansson2014gpuRFAG, Nasridinov2013DecisionTC}.

% Asadi et. al.\cite{VPred} optimize tree walks by hiding dependency stalls 
% by interleaving tree walks. In contrast, \Treebeard{} is an extensible 
% optimizing compiler that carefully implements this and many other
% optimizations at different levels of abstraction.
% QuickScorer\cite{QuickScorer, QuickScorer1} is an algorithm that uses 
% bit manipulation to compute tree predictions. Even though QuickScorer is 
% extremely fast for smaller models, it does not scale well to larger 
% models\cite{ProbBasedLayout}. The goal of QuickScorer is orthogonal 
% to the goals of \Treebeard{} and the QuickScorer algorithm can easily  
% be integrated into \Treebeard{} as another traversal strategy for the 
% system to explore. Tang et. al.\cite{CacheConscious1} and Jin et. al.\cite{CacheConscious2}
% build models to predict cache performance of decision tree ensembles on CPUs.
% This work is again orthogonal to the work described in this paper. Some systems have been proposed 
% to parallelize decision tree training on CPUs and GPUs\cite{Jansson2014gpuRFAG, Nasridinov2013DecisionTC}.

%Two recently published systems explore 
%using GPUs for decision tree inference\cite{Tahoe, Hummingbird}.

\emph{Other Systems and Techniques:} Ren et. al.~\cite{PortableVM} design an
intermediate language and a virtual machine to enable vector execution of decision tree
inference. In contrast with \Treebeard{}, the virtual machine that performs the SIMD 
execution is itself implemented by hand on different target processors. This is clearly 
more expensive than \Treebeard{}'s approach since the VM
needs to be reimplemented for every supported target architecture.
% Additionally, even though they perform layout optimizations, their system does
% not perform any model specific optimizations.
Jo et. al.\cite{MilindTreeVectorization} describe code transformations and runtime 
techniques that help vectorize tree-based applications. However, they do not 
study optimizations specific to decision trees.
% Both these works vectorize tree walks by performing different tree walks on each
% vector lane. The main issue with this approach is the divergence of tree walks.
% Another issue is that memory accesses are \op{gather}'s rather than vector
% loads. \Treebeard{}'s approach to vectorization solves both these issues. Also,
% these approaches are not precluded by \Treebeard{}'s tiling based vectorization.
% Multiple tiled tree walks can be combined into a single vectorized walk. We
% leave an exploration of this to future work.  FAST\cite{FAST} is a system that
% accelerates tree structured index search on CPUs and GPUs. FAST defines a layout
% for the index tree that enables vectorization of the tree walk. FAST uses a tree
% tiling approach to vectorize tree walks. However, FAST only uses a single
% triangular tile shape. \Treebeard{}'s basic tiling algorithm is a generalization
% of the tiling used in FAST. If given a perfectly balanced tree, the basic tiling
% algorithm would return exactly the tiling used by FAST.  Also, the tree walks in
% FAST are hand coded using intrinsics on the CPU and CUDA on the GPU and
% therefore need repeated effort to implement on each target. 
Inspector-executor systems \cite{TaoOfParallelism,HybridCPUGPU} have been 
developed to parallelize tree walks. However, these techniques are not a good fit 
for decision tree inference. In decision trees, the individual node predicates are very 
simple and the overhead of an inspector-executor system would be prohibitive.

\emph{Code Generation Systems from Other Domains:}
Several optimizing compilers and code generation techniques have been developed 
for other domains. TVM\cite{TVM}, Tiramisu\cite{Tiramisu}, and Tensor 
Comprehensions\cite{TensorComprehensions} are optimizing compilers 
for DNNs that can target a variety of processors. Similarly, 
Halide\cite{Halide} is a DSL and compiler primarily designed for 
image processing applications. The concept of separating the computation 
from the schedule was pioneered by Halide and has since been adopted 
by several other systems~\cite{TVM,Tiramisu,GraphIt}. However, to 
the best of our knowledge, \Treebeard{} is the first system to design
a scheduling language for decision tree inference optimization
and to build a system capable of state-of-the-art performance 
across different processors.
% {\sc {Cortex}} transforms 
% recursive computations in DNNs to loop based computations and optimizes them.
% However, the model properties {\sc {Cortex}} assumes to optimize
% generated code do not hold in the context of decision tree inference. For example, 
% {\sc {Cortex}} assumes that control flow depends purely on data structure 
% connectivity. This is not the case with decision trees. Control flow is determined 
% by the input value and not by the structure of the tree. This makes the 
% problems addressed by \Treebeard{} very different from the ones handled
% by {\sc {Cortex}}.

Libraries that compose or generate optimized implementations  
for BLAS\cite{BLIS, atlas_sc98, CUTLASS} and signal processing\cite{FFTW, SPIRAL}.
% BLIS\cite{BLIS} and ATLAS\cite{atlas_sc98}
% are systems to instantiate high performance BLAS routines on multiple 
% target architectures. CUTLASS\cite{CUTLASS} provides building blocks 
% in the form of C++ templates to quickly instantiate high performance 
% BLAS functions on different GPU architectures. SPIRAL\cite{SPIRAL}
% is a domain specific compiler for signal processing applications
% that instantiates high performance routines for linear transformations 
% like FFTs and FIR filters. FFTW\cite{FFTW} is a fast fourier transform 
% compiler that generates high performance FFT routines by customizing 
% the routine based on FFT size and the target machine. 
% However, no such frameworks for decision tree ensembles exist currently.
% \Treebeard{} is a first step in this direction and unlocks several future 
% optimization opportunities.

\emph{Reductions:} CUB\cite{CUB} and Thrust\cite{Thrust} are libraries 
that implement high-performance parallel reductions on GPUs. While they 
provide highly-tuned implementations to perform large reductions, 
it is not possible to fuse these functions with other computations  
as required in \Treebeard{}. Reddy et. al.~\cite{ChandanReduction}
describe language constructs in {\sc {Pencil}}~\cite{Pencil}
to express reductions and to represent and optimize them using the polyhedral 
framework. It is not clear how these techniques can be fused with 
other computations in arbitrary loop nests as required in \Treebeard{}.
Additionally, their system does not express the hierarchical nature of 
reductions and also only targets GPUs. Suriana et. al.~\cite{HalideReductions}
extend Halide to add support for factoring reductions in the Halide 
scheduling language and to synthesize reduction operators. De Gonzalo 
et. al.~\cite{TangramReduction} describe a system based on Tangram 
that composes several partial reduction implementations into different 
reduction implementations for GPUs and then searches through these
alternate implementations to find the best ones. In summary, none of 
these systems provide abstractions and a general framework to generate 
and optimize reductions across different target processors as \Treebeard{} does.

### 2024-04-10 docs/SOSP 2024/schedule.tex
Add HIR MIR optimization section
Changes to scheduling language section

@@
 \section{Scheduling Language}
-The goal of the scheduling language is to express the following
-\begin{itemize}
-  \item The order in which the iteration space over the batch of inputs (batch) and trees in the forest (tree) is to be traversed. Note that there is a reduction over the tree dimension.
-  \item Intra or inter tree optimizations that are to be performed on a tree or set of trees (tree walk unrolling, pipelining, SIMDize etc).
-\end{itemize}
+\label{sec:schedule}
+As we articulated in Section \ref{sec:intro}, there are several 
+different configurations and optimizations strategies for decision 
+tree inference. The best ones are significantly different across
+models, batch sizes and hardware platforms. Therefore, designing
+any one hard-coded lowering strategy is not feasible as this would 
+make portable performance impossible. To address this problem, we 
+design a scheduling language for \Treebeard{}. The scheduling 
+language provides an abstract way to specify loop structure and 
+other optimizations as an input to the compiler. Making the 
+scheduling specification external to the compiler allows to build 
+auto-schedulers and auto-tuners (Section \ref{sec:exploring}).
+Using a scheduling language will also significantly simplify 
+adding support for new hardware as this will likely require
+different locality optimizations and loop transformations.
 
-The reasons to use a scheduling language rather than a hard-coded lowering are as follows
-\begin{itemize}
-  \item Making the scheduling specification external to the compiler allows us to 
-  more easily build auto-schedulers and auto-tuners.
-  \item It is very hard to come up with a template loop nest that works for all models 
-  (for example, tree sizes may vary across the model making it necessary to iterate 
-  different number of trees at different times).
-  \item A scheduling language will make writing newer locality optimizations faster since 
-  no changes to the compiler infrastructure will be needed.
-  \item Adding support for additional hardware targets (GPUs, FPGAs), will be much easier
-  with a scheduling language.
-\end{itemize}
-
-There are some simplifying assumptions and limitations in the current design
-\begin{itemize}
-  \item Tree traversals are considered atomic. There is no way to express partial
-   tree traversals or schedule individual node/level computations. 
-  \item Accumulation of tree predictions is done immediately (as opposed to, 
-  for example, collecting all predictions and performing a reduction later).
-\end{itemize}
+The goal of \Treebeard{}'s scheduling language is to declaratively
+express loop structures and the application of other optimizations 
+(tree walk unrolling, tree walk interleaving etc.).
 
 \subsection{Language Definition}
-We broadly have three classes of directives in the language. The first is a set 
-of loop nest modifiers that are used to specify the structure of the loop nest to
-walk the iteration space. The second is a set of clauses that specify intra and 
-inter tree optimizations. Finally, we have a class of attributes that control how 
-reductions are performed.
+The core construct of \Treebeard{}'s scheduling language is an 
+\textbf{\emph{index variable}} which abstractly represents a loop. 
+The language then provides directives to manipulate these index 
+variables. There are two special index variables -- \op{batch} and
+\op{tree} that are used to represent the batch and tree loops and all 
+other index variables are derived from these. A schedule derives 
+new index variables from these root index variables by applying
+directives.
 
-\subsubsection{Loop Modifiers}
-There are two special index variables -- \op{batch} and \op{tree}. The clauses modify these index variables or 
-index variables derived from these (through the application of clauses).
-\begin{itemize}
-  \item \textbf{tile}: Tile the passed index variable using a fixed tile size.
-  \item \textbf{split}: Split the range of the passed index variable into two parts. The range of the first part is specified
-  by an argument.
-  \item \textbf{unroll}: Unroll an index completely
-  \item \textbf{reorder}: Reorder the specified indices. The specified indices must be successive indices in the current loop nest.
-  \item \textbf{specialize}: Generate separate code for each iteration of the specified index variable. This is useful 
-  while parallelizing across trees and these trees have different depths.
-  \item \textbf{gpuDimension}: Maps the specified index variable to represent a dimension in either the GPU kernel grid or thread block.
-\end{itemize}
+\Treebeard{}'s scheduling language has three classes of directives. The first is a set 
+of loop modifiers that are used to specify the structure of the loop nest to
+walk the iteration space (Table \ref{Tab:LoopModifiers}). The second is a set of 
+directives that enable optimizations on a loop (Table \ref{Tab:Optimizations}). 
+Finally, we have a class of attributes that enable reduction specific optimizations
+(Table \ref{Tab:ReductionOpts}).
+
+% \subsubsection{Loop Modifiers}
+% The clauses modify these index variables or 
+% index variables derived from these (through the application of clauses).
+% \begin{itemize}
+%   \item \textbf{tile}: Tile the passed index variable using a fixed tile size.
+%   \item \textbf{split}: Split the range of the passed index variable into two parts. The range of the first part is specified
+%   by an argument.
+%   % \item \textbf{unroll}: Unroll an index completely
+%   \item \textbf{reorder}: Reorder the specified indices. The specified indices must be successive indices in the current loop nest.
+%   \item \textbf{specialize}: Generate separate code for each iteration of the specified index variable. This is useful 
+%   while parallelizing across trees and these trees have different depths.
+%   \item \textbf{gpuDimension}: Maps the specified index variable to represent a dimension in either the GPU kernel grid or thread block.
+% \end{itemize}
+
+\begin{table}[htb]
+  \centering
+  \resizebox{\linewidth}{!}{
+  \begin{tabularx}{\linewidth}{c | l | l}
+   \toprule
+   \textbf{Directive} & \textbf{Inputs} & \textbf{Description} \\
+   \midrule
+   \multirow{4}{*}{\texttt{tile}} & \textbf{indexVar}  & \multirow{4}{*}{\parbox{0.55\linewidth}{Tile the loop corresponding to \textbf{indexVar}
+   with the specified tile size. Resulting loops will be represented by \textbf{outer} and \textbf{inner}.}} \\
+                                               &  \textbf{outer} & \\
+                                               &  \textbf{inner} &  \\
+                                               &  \textbf{tileSize} &  \\
+   \midrule
+   \multirow{6}{*}{\texttt{split}} & \textbf{indexVar}  & \multirow{6}{*}{\parbox{0.55\linewidth}{Fiss the loop represented by \textbf{indexVar}
+   at iteration \textbf{splitIter}. Resulting loops will be represented by \textbf{first} and \textbf{second}. Returns a maps from nested 
+   index variables to new ones created by splitting.}} \\
+                                               &  \textbf{first} & \\
+                                               &  \textbf{second} &  \\
+                                               &  \textbf{splitIter} &  \\
+                                               & & \\
+                                               & & \\
+
+   \midrule
+   \multirow{4}{*}{\texttt{reorder}} & \textbf{indices[]}  & \multirow{4}{*}{\parbox{0.55\linewidth}{Permute loops corresponding to the specified index variables.
+   The loops must be perfectly nested in the current loop structure.}} \\
+                                               &   & \\
+                                               &   &  \\
+                                               &   &  \\
+   \midrule
+   \multirow{4}{*}{\texttt{specialize}} & \textbf{indexVar}  & \multirow{4}{*}{\parbox{0.55\linewidth}{Generate specialized code for each iteration of the loop. 
+   Useful when different iterations of a loop need to execute different code.}} \\
+                                               &   & \\
+                                               &   &  \\
+                                               &   &  \\
+   \midrule
+   \multirow{3}{*}{\texttt{gpuDimension}} & \textbf{indexVar}  & \multirow{3}{*}{\parbox{0.55\linewidth}{Map the passed index variable to a dimension of either the grid
+   or thread block.}} \\
+                                               & \textbf{gpuDim}  & \\
+                                               &   &  \\
+
+  \bottomrule
+  \end{tabularx}
+  }
+  \vskip 5pt
+  \caption{\label{Tab:LoopModifiers} List of all the loop modifiers in \Treebeard{}'s scheduling language. We use \emph{index variable}
+  and \emph{loop} interchangeably in descriptions for clarity of exposition.}
+\end{table}
+
+% The following are examples of how the loop modifiers can be used.
+% \begin{itemize}
+%   \item The loop order used by XGBoost\cite{XGBoost} is (tree, batch) -- walk one tree for all inputs in the batch before 
+%   moving to the next tree. The corresponding schedule would be
+% \begin{lstlisting}[style=c++]
+%   reorder(tree, batch)
+% \end{lstlisting}
+
+%   \item The below schedule computes 2 trees at a time over the whole batch.
+% \begin{lstlisting}[style=c++]
+%   tile(tree, t0, t1, 2)
+%   reorder(t0, batch, t1)
+% \end{lstlisting}
+
+%   \item If we additionally only want to compute over 4 input rows (rather than the whole batch) for
+%   every 2 tree, and then move onto the next 2 trees for the same set of inputs, then the schedule is as follows. 
+% \begin{lstlisting}[style=c++]
+%   tile(batch, b0, b1, 4) 
+%   tile(tree, t0, t1, 2)
+%   reorder(b0, t0, b1, t1) 
+% \end{lstlisting}
+
+% \end{itemize}
+
+% \subsubsection{Optimizations}
+% The following clauses provide ways to optimize the inference routine being generated.
+% \begin{itemize}
+%   \item \textbf{cache}: Cache the working set of one iteration of the specified loop corresponding
+%   to this index. This can be specified on either batch or tree loops. Specifying it 
+%   on a batch loop leads to all rows accessed in a single iteration of the loop 
+%   being cached. Similarly, specifying it on a tree loop leads to all trees accessed in 
+%   one iteration of that loop being cached.
+%   \item \textbf{parallel}: Execute the loop corresponding to this index in parallel.
+%   \item \textbf{interleave}: Interleave the execution of the tree walks within the current index (must be applied on an inner most index).
+%   \item \textbf{unrollWalk}: Unroll tree walks at the current index. 
+%   \item \textbf{peelWalk}: Peel the first n steps of the specified tree walk and don't check for leaves for that number of steps.
+% \end{itemize}
+
+\begin{table}[htb]
+  \centering
+  \resizebox{\linewidth}{!}{
+  \begin{tabularx}{\linewidth}{c | l | l}
+   \toprule
+   \textbf{Directive} & \textbf{Inputs} & \textbf{Description} \\
+   \midrule
+   \multirow{4}{*}{\texttt{cache}} & \textbf{indexVar}  & \multirow{4}{*}{\parbox{0.55\linewidth}{Cache the working set of one iteration of the specified loop. 
+   Cache rows for a batch loop and trees for a tree loop.}} \\
+                                               &  &  \\
+                                               &  &  \\
+                                               &  &  \\
+   \midrule
+   \multirow{2}{*}{\texttt{parallel}} & \textbf{indexVar}  & \multirow{2}{*}{\parbox{0.55\linewidth}{Execute the iterations of the specified 
+   loop in parallel.}} \\
+                                               &  & \\
+
+   \midrule                                               
+   \multirow{3}{*}{\texttt{interleave}} & \textbf{indexVar}  & \multirow{3}{*}{\parbox{0.55\linewidth}{Interleave tree walks within the specified 
+   loop (must be innermost loop).}} \\
+                                               &  & \\
+                                               &  & \\
+
+   \midrule                                               
+   \multirow{3}{*}{\texttt{unrollWalk}} & \textbf{indexVar}  & \multirow{3}{*}{\parbox{0.55\linewidth}{Unroll tree walks at the specified loop 
+   for \textbf{unrollDepth} hops. Loop must be an innermost loop.}} \\
+                                               & \textbf{unrollDepth} & \\
+                                               &  & \\
+
+  \bottomrule
+  \end{tabularx}
+  }
+  \vskip 5pt
+  \caption{\label{Tab:Optimizations} List of optimization directives in \Treebeard{}'s scheduling language. 
+  We use \emph{index variable} and \emph{loop} interchangeably in descriptions for clarity of exposition.}
+\end{table}
+
+% \subsubsection{Reduction Optimization}
+% \begin{itemize}
+%   \item \textbf{atomicReduce}: Use atomic memory operations to accumulate values across 
+%   parallel iterations of the specified loop. 
+%   \item \textbf{sharedReduce}: Only applies to GPU compilation. Specifies that intermediate
+%   results are to be stored in shared memory.
+%   \item \textbf{vectorReduce}: Use vector instructions with the specified vector width 
+%   to reduce intermediate values across parallel iterations of the specified loop.
+% \end{itemize}
+
+\begin{table}[htb]
+  \centering
+  \resizebox{\linewidth}{!}{
+  \begin{tabularx}{\linewidth}{c | l | l}
+   \toprule
+   \textbf{Directive} & \textbf{Inputs} & \textbf{Description} \\
+   \midrule
+   \multirow{3}{*}{\texttt{atomicReduce}} & \textbf{indexVar}  & \multirow{3}{*}{\parbox{0.55\linewidth}{Use atomic memory operations to accumulate values across 
+   parallel iterations of the specified loop.}} \\
+                                               &  &  \\
+                                               &  &  \\
+   \midrule
+   \multirow{3}{*}{\texttt{sharedReduce}} & \textbf{indexVar}  & \multirow{3}{*}{\parbox{0.55\linewidth}{Specifies that intermediate
+   results are to be stored in shared memory (GPU only).}} \\
+                                               &  & \\
+                                               &  & \\
+
+   \midrule                                               
+   \multirow{4}{*}{\texttt{vectorReduce}} & \textbf{indexVar}  & \multirow{4}{*}{\parbox{0.55\linewidth}{Use vector instructions with the specified vector width 
+   to reduce intermediate values across parallel iterations of the specified loop.}} \\
+                                               & \textbf{width} & \\
+                                               &  & \\
+                                               &  & \\
+
+  \bottomrule
+  \end{tabularx}
+  }
+  \vskip 5pt
+  \caption{\label{Tab:ReductionOpts} List of reduction optimization directives in \Treebeard{}'s scheduling language. 
+  We use \emph{index variable} and \emph{loop} interchangeably in descriptions for clarity of exposition.}
+\end{table}
+
+%\TODO{Add examples for RAPIDS, Tahoe (Maybe show some strategies can be encoded?)}
 
 The default loop order (as currently generated by the compiler) is (batch, tree), i.e, for each row in the input batch, go over all trees.
-
-The following are examples of how the loop modifiers can be used.
-\begin{itemize}
-  \item The loop order used by XGBoost\cite{XGBoost} is (tree, batch) -- walk one tree for all inputs in the batch before 
-  moving to the next tree. The corresponding schedule would be
-\begin{lstlisting}[style=c++]
-  reorder(tree, batch)
-\end{lstlisting}
-
-  \item The below schedule computes 2 trees at a time over the whole batch.
-\begin{lstlisting}[style=c++]
-  tile(tree, t0, t1, 2)
-  reorder(t0, batch, t1)
-\end{lstlisting}
-
-  \item If we additionally only want to compute over 4 input rows (rather than the whole batch) for
-  every 2 tree, and then move onto the next 2 trees for the same set of inputs, then the schedule is as follows. 
-\begin{lstlisting}[style=c++]
-  tile(batch, b0, b1, 4) 
-  tile(tree, t0, t1, 2)
-  reorder(b0, t0, b1, t1) 
-\end{lstlisting}
-
-\end{itemize}
-
-\subsubsection{Optimizations}
-The following clauses provide ways to optimize the inference routine being generated.
-\begin{itemize}
-  \item \textbf{cache}: Cache the working set of one iteration of the specified loop corresponding
-  to this index. This can be specified on either batch or tree loops. Specifying it 
-  on a batch loop leads to all rows accessed in a single iteration of the loop 
-  being cached. Similarly, specifying it on a tree loop leads to all trees accessed in 
-  one iteration of that loop being cached.
-  \item \textbf{parallel}: Execute the loop corresponding to this index in parallel.
-  \item \textbf{interleave}: Interleave the execution of the tree walks within the current index (must be applied on an inner most index).
-  \item \textbf{unrollWalk}: Unroll tree walks at the current index. 
-  \item \textbf{peelWalk}: Peel the first n steps of the specified tree walk and don't check for leaves for that number of steps.
-\end{itemize}
-
-\subsubsection{Reduction Optimization}
-\begin{itemize}
-  \item \textbf{atomicReduce}: Use atomic memory operations to accumulate values across 
-  parallel iterations of the specified loop. 
-  \item \textbf{sharedReduce}: Only applies to GPU compilation. Specifies that intermediate
-  results are to be stored in shared memory.
-  \item \textbf{vectorReduce}: Use vector instructions with the specified vector width 
-  to reduce intermediate values across parallel iterations of the specified loop.
-\end{itemize}
-
-\TODO{Add examples for RAPIDS, Tahoe (Maybe show some strategies can be encoded?)}
 
@@
 \end{itemize}
+
+There are some simplifying assumptions and limitations in the current design
+\begin{itemize}
+  \item Tree traversals are considered atomic. There is no way to express partial
+   tree traversals or schedule individual node/level computations. 
+  \item Accumulation of tree predictions is done immediately (as opposed to, 
+  for example, collecting all predictions and performing a reduction later).
+\end{itemize}

### 2024-04-10 docs/SOSP 2024/abstract.tex
Edit abstract

@@
-\begin{comment}
-\begin{abstract}
-  Decision tree based models are generated by some of the most commonly 
-  used machine learning algorithms like gradient boosting and random forests.
-  These algorithms are used very widely in practice and the resulting models
-  are deployed to production at scale. The models are served on a variety of 
-  machines including CPUs and GPUs and inference performance on all these 
-  machines is an important consideration. On CPUs, models are served using
-  either libraries such as XGBoost, Sklearn, and LightGBM or specialized 
-  compilers like \TreebeardOLD{}, Treelite, and Hummingbird. On GPUs, XGBoost 
-  and RAPIDs FIL are popular libraries. However, these libraries and compilers
-  do not provide portable performance across the targets of interest. They 
-  also leave significant optimization opportunities unexplored.
-
-  In this paper, we present the design of \Treebeard{}, a compiler infrastructure 
-  for decision tree based models that significantly expands the design space 
-  that can be explored and provides portable performance across CPU and GPU 
-  targets. We rearchitect the open-source \Treebeard{} infrastructure 
-  and significantly extend it to enable high-performance code generation 
-  across target processors. Specifically, we design a scheduling language 
-  that encapsulates the large optimization space for decision tree inference
-  and develop techniques to explore this space. Further,
-  we design techniques to represent and optimize parallel reductions, extend
-  \TreebeardOLD{}'s intermediate representations to include operations like caching
-  and rearchitect \TreebeardOLD{} to allow in-memory representations to be built 
-  as plugins. \TODO{Need to say we investigate how to structure a compiler for 
-  decision tree inference that can target multiple hardware platforms}
-
-  We implement \Treebeard{} using the MLIR compiler infrastructure and
-  compare the performance of the generated code against state-of-the-art
-  systems on CPUs and GPUs. On GPUs we compare performance against XGBoost,
-  RAPIDs FIL, and Tahoe. We find that \Treebeard{} is significantly faster
-  than other systems that target GPUs across a wide range of batch sizes 
-  and models on different NVIDIA GPUs. \Treebeard{} is an order of magnitude faster than XGBoost and
-  about 2-3$\times$ faster than RAPIDs FIL and Tahoe on average. \TODO{Make 
-  this more precise. But how?}. We also show how \Treebeard{} can generate  
-  high-performance code on other GPUs by running the benchmarks on an AMD GPU.
-  On CPUs, we compare against \TreebeardOLD{} and show how the  
-  optimization opportunities exposed by \Treebeard{}'s scheduling 
-  language allow it to generate higher performance code by specializing  
-  code for different batch sizes.
-\end{abstract}
-\end{comment}
 \begin{abstract}
 
-  With the rapid proliferation of machine learning, the demand for on-device model inference has surged.
-Concurrently, the hardware ecosystem is evolving rapidly, leading to an increasing heterogeneity in client machines.
+  The rapid proliferation of machine learning coupled with accelerating evolution of the hardware ecosystem has led to a surge in the demand for model inference on a variety of hardware.
   %While CPUs have been the mainstay for machine learning inference, the availability of GPUs holds promise to scale to bigger and more powerful models. 
-  This paper is motivated by the problems encountered when targeting inference of decision tree based models, the most popular models on tabular data, to run at peak performance on diverse client hardware.
-We evaluated existing solutions and found that they do not provide portable performance across different hardware targets.
+  This paper is motivated by the problems encountered when targeting inference of decision tree based models, the most popular models on tabular data, to run at peak performance on 
+  diverse CPU and GPU targets. We evaluated existing solutions and found that they do not provide portable performance across different hardware targets.
   %Decision tree based models are widely used in practice due to their robustness, interpretability, and ability to handle missing data.
 
-  To address this we present the design of \Treebeard{}, a schedule guided compiler infrastructure 
+  To address this we present the design of \Treebeard{}, a schedule guided compiler  
   for decision tree based models that searches over a large design space 
-  to find high performance schedules on a diverse set of CPU and GPU 
-  targets. 
+  to automatically generate high-performance inference routines.
   %We re-architect the open-source \TreebeardOLD{} infrastructure and significantly extend it to enable high-performance code generation across target processors. 
@@
   and techniques to efficiently explore this space.
-  \TODO{include some keywords from the language and space complexity}.
+  \TODO{Change large optimization space}.
   %We also design a set of heuristics that can find near-optimal solutions in quick time.
@@
   For the latter, we re-architect the open-source \TreebeardOLD{} CPU compiler to support schedule guided compilation and
-add support for GPU code generation.
-We further enhance the compiler with fundamental new optimization interfaces for caching, parallel reduction, and a plug-in mechanism to explore different in-memory representations of trees.
+  add support for GPU code generation. GPU code generation required fundamental new optimization interfaces for 
+  caching, parallel reduction, and support for multiple in-memory representations of trees.
 
@@
 
-  We evaluate \Treebeard{} on \TODO{kr:Can we quote a number} a large number of diverse models and demonstrate that the best schedule varies drastically with model, batch size, and target hardware. 
-  Despite this, the scheduling heuristic is able to quickly find near optimal schedules while searching over a small fraction of the total search space.
-  In terms of performance, \Treebeard{} generated code  
-   is an order of magnitude faster than XGBoost and
-  about 2-3$\times$ faster on average than RAPIDs FIL and Tahoe. \TODO{Make 
-  this more precise. But how?}. While these systems only generate code for Nvidia, \Treebeard{} achieves competent performance on AMD GPUs as well. 
-  On CPUs, we compare against \TreebeardOLD{} and show how the  
-  optimization opportunities exposed by \Treebeard{}'s scheduling 
-  language allow it to generate higher performance code by specializing  
-  code for different batch sizes.
+  We evaluate \Treebeard{} on over seven hundred diverse models and demonstrate that the best schedule varies drastically with model, batch size, and target hardware. 
+  Our scheduling heuristic is able to quickly find near optimal schedules while searching over a small number (\~50) of schedules.
+  In terms of performance, \Treebeard{} generated code is an order of magnitude faster than XGBoost and
+  about 2-3$\times$ faster on average than RAPIDs FIL and Tahoe. While these systems only target NVIDIA GPUs, \Treebeard{} achieves competent performance on AMD GPUs as well. 
+  On CPUs, \Treebeard{} achieves better scaling compared to \TreebeardOLD{}.
+  For models where \TreebeardOLD{} was only able to achieve diminishing returns with an 
+  increasing number of threads, \Treebeard{} is able to scale linearly with the number of threads.
   \TODO{(numbers for CPU performance?)}

### 2024-04-10 docs/SOSP 2024/intro.tex
Edit intro

@@
 The survey also finds that, the cost of inference is the most critical factor in the overall cost of deploying a machine learning model.
-This is because, in production settings, each model is trained once and often used for inference a few thousand times. 
-%
+This is because, in production settings, each model is trained once and often used for inference millions of times. 
+However, inference is run on a variety of hardware platforms, ranging from low to high-end CPUs and GPUs. 
+This paper is motivated by the need to accelerate decision tree inference to achieve portable performance on a variety of hardware.
+% In particular, we focus on a range of commodity CPUs and GPUs, a class of hardware that has seen widespread 
+% adoption across client/edge devices used for inference.
 
-This paper is motivated by the need to accelerate decision tree inference to achieve portable performance on a variety of client hardware.
-In particular, we focus on a range of commodity CPUs and low-to-mid-range GPUs, a class of hardware that has seen widespread adoption across client/edge devices used for inference.
-
-Decision forest models are composed of a large collection of decision trees (100-1000), and inference involves traversing down each tree in the forest and aggregating the predictions.
-Inference is typically done in a batched setting, where multiple inputs are processed simultaneously.
-
-%Despite the simplicity of the model, decision tree inference is computationally expensive, and is surprisingly hard to scale to a large number of threads
-Despite the simplicity of the model and the availability of multiple sources of coarse grain parallelism (parallelism across inputs in a batch and parallelism across trees), existing systems do not consistently scale well across different models.
-This is because existing systems exploit a limited set of optimizations and often specialize the implementation to a specific hardware platform, and this limits their portability.
+Decision forest models are composed of a large collection of decision trees (100-1000), and inference involves 
+traversing down each tree in the forest and aggregating the predictions. Inference is typically done in a batched
+setting, where multiple inputs are processed simultaneously.
+Despite the simplicity of the model and the availability of multiple sources of coarse grain parallelism (parallelism 
+across inputs in a batch and parallelism across trees), existing systems do not consistently scale well across different
+models even on the limited set of targets they support. 
+This is because existing systems exploit a limited set of optimizations and often specialize the implementation 
+to a specific hardware platform, and this limits their portability.
 
 %\TODO{cover Rapids, Tahoe, XGBoost and Treebeard}.  
-Evaluation on a diverse set of models highlights that the best implementation often requires a careful combination of many optimization strategies.
-For example, as we simultaneously traverse multiple trees, managing the working set of nodes is critical to get scalable performance.
-This requires a combination of techniques like data layout optimizations, loop transformations, and memory access optimizations.
-Existing systems each use specific data layouts for the underlying tree (XGBoost~\cite{XGBoost} uses a sparse representation, RAPIDs~\cite{Rapids} FIL~\cite{FTL} uses what is called the reorg representation
-and Tahoe~\cite{Tahoe} uses a variation of the reorg representation) and propose simple loop transformations (XGBoost uses \dots() \TODO{complete}) that together determine the accesses patterns and the working set.
-They do not perform consistently well across different models and hardware platforms.
-Managing the working set is just one of several optimizations that are critical for high-performance decision tree inference.
- \TODO{Improve/drop the next line}
- Others include tree ordering,  incremental reduction of predicted values, interleaving across multiple trees, and so on.
+Evaluation on a diverse set of models highlights that the best implementation often requires a careful combination 
+of many optimization strategies. For example, simultaneously traversing multiple trees with many threads, while managing the working set is 
+critical to get scalable performance. This requires a combination of techniques like data layout optimizations, 
+loop transformations, and memory access optimizations. Existing systems each use specific data layouts for the underlying
+tree and use fixed loop structures. 
+Parallelism, loop structure and model representation together determine the accesses patterns and the working set. A fixed configuration 
+as used by existing systems does not perform consistently well across different models and hardware platforms.
+XGBoost~\cite{XGBoost} uses a sparse representation for the model and a loop structure that processes one tree for a block of rows before moving to 
+the next tree. RAPIDS FIL~\cite{FIL} uses a reorg representation and partitions trees across a fixed number of threads. Tahoe~\cite{Tahoe} uses a
+variation of the reorg representation and has four predefined inference strategies from which it picks one based on an analytical model.
+These systems are all library based and are hard to extend. 
+\TreebeardOLD{}, the state-of-the-art model compiler for CPUs, supports two fixed loop structures and does not scale well with increasing number of threads.
 

### 2024-04-11 docs/SOSP 2024/intro.tex
intro joint pass

@@
 models even on the limited set of targets they support. 
-This is because existing systems exploit a limited set of optimizations and often specialize the implementation 
-to a specific hardware platform, and this limits their portability.
+%This is because existing systems exploit a limited set of optimizations and often specialize the implementation to a specific hardware platform, and this limits their portability.
 
 %\TODO{cover Rapids, Tahoe, XGBoost and Treebeard}.  
-Evaluation on a diverse set of models highlights that the best implementation often requires a careful combination 
-of many optimization strategies. For example, simultaneously traversing multiple trees with many threads, while managing the working set is 
-critical to get scalable performance. This requires a combination of techniques like data layout optimizations, 
-loop transformations, and memory access optimizations. Existing systems each use specific data layouts for the underlying
-tree and use fixed loop structures. 
-Parallelism, loop structure and model representation together determine the accesses patterns and the working set. A fixed configuration 
-as used by existing systems does not perform consistently well across different models and hardware platforms.
+Evaluation on a diverse set of models highlights that the best implementation often requires a careful composition 
+of many optimization strategies like data layout optimizations, 
+loop transformations, parallelization, and memory access optimizations. 
+%For example, simultaneously traversing multiple trees with many threads, while managing the working set is 
+%critical to get scalable performance. 
+%This requires a combination of techniques like data layout optimizations, loop transformations, and memory access optimizations. 
+%Existing systems each use specific data layouts for the underlying tree and use fixed loop structures. 
+%Parallelism, loop structure and model representation together determine the accesses patterns and the working set. 
+Existing systems today are mostly library based, and only support a predefined combination of optimizations. 
 XGBoost~\cite{XGBoost} uses a sparse representation for the model and a loop structure that processes one tree for a block of rows before moving to 
@@
 variation of the reorg representation and has four predefined inference strategies from which it picks one based on an analytical model.
-These systems are all library based and are hard to extend. 
+%These systems are all library based and are hard to extend. 
 \TreebeardOLD{}, the state-of-the-art model compiler for CPUs, supports two fixed loop structures and does not scale well with increasing number of threads.
+Additionally, it lacks GPU specific optimizations that are critical to scale performance to massive number of threads.
 
@@
 \Treebeard{} is able to generate high-performance code for decision tree inference by exploring a large optimization space.
-The code generation is guided by a \emph{schedule} object that allows us to effectively represent this solution space abstractly.
-The schedule object is written in \Treebeard{} custom scheduling language that is expressive enough to represent a wide range of implementation strategies.
+The code generation is guided by a \emph{schedule} written in a custom scheduling language that can represent a wide range of implementation strategies.
 We demonstrate that the language is sufficient to express the various optimizations proposed by prior work, and further generalizes them and incorporates several new optimizations.

### 2024-04-11 docs/SOSP 2024/hir-mir-opts.tex
Edit schedule and hir-mir section

@@
 
-The \Treebeard{} infrastructure was originally designed to target CPUs \cite{Treebeard}. 
+The \TreebeardOLD{} infrastructure was originally designed to target CPUs \cite{Treebeard}. 
 However, it implements several optimizations on the HIR and MIR that 
-can be leveraged across target processors and we find that these 
+can be leveraged across target processors and we find that some these 
 are beneficial for GPUs as well. This reuse of optimizations is possible 
 because the intermediate representations on which these optimizations are performed 
-are abstract and are designed to be target-independent. We briefly review theese 
+are abstract and are designed to be target-independent. We briefly review these 
 optimizations below.
@@
 \subsection{Optimizations on High-Level IR}
-We augment the existing \Treebeard{} infrastructure with loop rewrites on 
+We augment the existing \TreebeardOLD{} infrastructure with loop rewrites on 
 the HIR that are implemented through the scheduling language (Section \ref{sec:schedule}).
 We use these to implement the automatic scheduling described in Section \ref{sec:exploring}. 
-Additionally, the \Treebeard{} infrastructure implements HIR transformations to reorder and pad 
+Additionally, the \TreebeardOLD{} infrastructure implements HIR transformations to reorder and pad 
 trees. It also implements tree tiling transformations on the HIR \cite{Treebeard}. We reuse the 
 reordering and padding transformations on the HIR for GPUs. However, we found that 
-tiling trees was not beneficial for GPUs. This is because the tiling transformations
+tiling trees was not beneficial for GPUs. This is because the tiling transformation
 introduces redundant computation inorder to vectorize computation on CPUs where 
@@
 \subsection{Optimizations on Mid-Level IR}
-The original \Treebeard{} infrastructure implements optimizations like 
+The original \TreebeardOLD{} infrastructure implements optimizations like 
 tree-walk unrolling, tree-walk interleaving, and parallelization
 on the MIR. These optimizations are beneficial for GPUs as well.
-and the design of \Treebeard{} allows us to reuse the tree-walk 
+and the design of \TreebeardOLD{} allows us to reuse the tree-walk 
 unrolling and tree-walk interleaving optimizations on the MIR for GPUs.
+
+While building \Treebeard{}, we found that one of the performance bottlenecks on the 
+GPU was that warps spent significant time being stalled. Since GPUs 
+implement scoreboarding~\cite{HennesseyPatterson}, we were able to alleviate this bottleneck by
+interleaving tree walks. This significantly improved performance of generated 
+code. We found it surprising that the use of ILP could benefit 
+performance on the GPU.  
 \TODO{Should we talk about how interleaving is implemented as a 
@@
 \subsection{A Note on Low-level IR}
-Some changes to the original \Treebeard{} design were required to get
+Significant changes to the original \TreebeardOLD{} design were required to get
 LIR to correctly lower to GPU code. The most important of these was 

### 2024-04-11 docs/SOSP 2024/schedule.tex
Edit schedule and hir-mir section

@@
 
-The default loop order (as currently generated by the compiler) is (batch, tree), i.e, for each row in the input batch, go over all trees.
+We now show how the implementation strategies of several existing systems can 
+be encoded in \Treebeard{}'s scheduling language. First, we note that the 
+default loop order is [\op{batch}, \op{tree}], i.e, for each row in the input
+batch, go over all trees.
 
-\subsection{The XBoost Schedule}
-XGBoost\cite{XGBoost} is a very popular gradient boosting library. 
-It implements inference on the CPU by going over a fixed number of rows (64 in the previous version)
-for every tree and then moving to the next tree. When all trees have been walked for this 
-set of rows, the next set of rows is taken up. Also, different sets of rows are 
-processed in parallel.
-
-The schedule used by XGBoost can be represented in \Treebeard{}'s scheduling 
-language as follows.
+% \subsection{The XBoost Schedule}
+XGBoost\cite{XGBoost} implements inference on the CPU by going 
+over a fixed number of rows (64 in the previous version)
+for every tree and then moving to the next tree. When all trees 
+have been walked for this set of rows, the next set of rows is 
+taken up. Also, different sets of rows are processed in parallel.
+This schedule can be implemented in \Treebeard{}'s scheduling language 
+as follows.
 \begin{lstlisting}[style=c++]
@@
 
-\subsection{Tahoe Schedules}
-Tahoe\cite{Tahoe} has four strategies that it picks from for a given model. 
-Each of these strategies can be encoded using \Treebeard{}'s scheduling language
-as we show below. 
+Tahoe\cite{Tahoe} has four strategies for inference on the GPU that it picks from for a given model. 
+We show how two of these strategies can be encoded using \Treebeard{}'s scheduling language.
+The rest can be encoded similarly. 
 \begin{itemize}
@@
 \end{lstlisting}
-  \item \textbf{Shared Forest}: In this strategy, the whole model is loaded into 
-  shared memory and subsequently, a single thread walks all trees for a particular
-  row. The schedule for this strategy is as follows.
-\begin{lstlisting}[style=c++]
-  tile(batch, b0, b1, ROWS_PER_TB)
-  tile(tree, t0, t1, N_TREES)
-  reorder(b0, b1, t0, t1)
-  cache(t0)
-  gpuDimension(b0, grid.x)
-  gpuDimension(b1, block.x)
-\end{lstlisting}
-  Here, we create a placeholder single iteration loop \op{t0} so that we can 
-  specify that all trees are to be cached.
-  \item \textbf{Shared Partial Forest}: In case the model is too large to fit into
-  shared memory, the model is split into chunks and each chunk is loaded into shared
-  memory. Again, as in the previous strategy, one thread walks all trees assigned to
-  a thread block for a row. The schedule for this strategy is as follows.
-\begin{lstlisting}[style=c++]
-  tile(batch, b0, b1, ROWS_PER_TB)
+%   \item \textbf{Shared Forest}: In this strategy, the whole model is loaded into 
+%   shared memory and subsequently, a single thread walks all trees for a particular
+%   row. The schedule for this strategy is as follows.
+% \begin{lstlisting}[style=c++]
+%   tile(batch, b0, b1, ROWS_PER_TB)
+%   tile(tree, t0, t1, N_TREES)
+%   reorder(b0, b1, t0, t1)
+%   cache(t0)
+%   gpuDimension(b0, grid.x)
+%   gpuDimension(b1, block.x)
+% \end{lstlisting}
+%   Here, we create a placeholder single iteration loop \op{t0} so that we can 
+%   specify that all trees are to be cached.
+%   \item \textbf{Shared Partial Forest}: In case the model is too large to fit into
+%   shared memory, the model is split into chunks and each chunk is loaded into shared
+%   memory. Again, as in the previous strategy, one thread walks all trees assigned to
+%   a thread block for a row. The schedule for this strategy is as follows.
+% \begin{lstlisting}[style=c++]
+%   tile(batch, b0, b1, ROWS_PER_TB)
 
-  tile(tree, t0, t0Inner, TREES_PER_TB)
-  tile(t0Inner, t1, t2, TREES_PER_TB)
-  cache(t1)
-  reorder(b0, t0, b1, t1, t2)
+%   tile(tree, t0, t0Inner, TREES_PER_TB)
+%   tile(t0Inner, t1, t2, TREES_PER_TB)
+%   cache(t1)
+%   reorder(b0, t0, b1, t1, t2)
 
-  gpuDimension(b0, grid.x)
-  gpuDimension(t0, grid.y)
-  gpuDimension(b1, block.x)
-\end{lstlisting}
+%   gpuDimension(b0, grid.x)
+%   gpuDimension(t0, grid.y)
+%   gpuDimension(b1, block.x)
+% \end{lstlisting}
 \end{itemize}
@@
 There are some simplifying assumptions and limitations in the current design
-\begin{itemize}
-  \item Tree traversals are considered atomic. There is no way to express partial
-   tree traversals or schedule individual node/level computations. 
-  \item Accumulation of tree predictions is done immediately (as opposed to, 
+of the scheduling language. Mainly, tree traversals are considered atomic and
+accumulation of tree predictions is done immediately (as opposed to, 
 for example, collecting all predictions and performing a reduction later).
-\end{itemize}
+However, we find that these are not significant limitations in practice as
+the current design is able to express most strategies of interest.

### 2024-04-11 src/reduction/LowerReduceOps.cpp
Fix atomic reduce bug

@@ -173,6 +173,87 @@ void addLoopIfNeededAndUpdateIndices(ConversionPatternRewriter &rewriter,
   }
 }
 
+/*
+Generates the following snippet (roughly)
+
+scf.if %notZeroIndex {
+  %byteTrue = arith.constant 1 : i8
+  scf.while (%arg = %zeroIdxDone) {
+    %c2 = memref.atomic.rmw andi %byteTrue, %arg[0]
+    %c7 = arith.cmpi "ne", %c2, %byteTrue : i1
+    scf.condition(%c7) %arg : memref<1xi8>
+  } do {
+    scf.yield %arg : memref<1xi8>
+  }
+} else {
+  memref.atomic.rmw assign %byteTrue, %zeroIdxDone[0]
+}
+
+%and = memref.atomic.rmw andi %byteTrue, %zeroIdxDone[0]
+arith.cmpi "eq", %and, %byteTrue : i1
+*/
+
+Value genWaitForZeroIdxReduction(Location location,
+                                 ConversionPatternRewriter &rewriter,
+                                 Value zeroIdxDone, Value notZeroIndex) {
+
+  auto zeroIndexConst = rewriter.create<arith::ConstantIndexOp>(location, 0);
+
+  auto i8True = rewriter.create<arith::ConstantIntOp>(
+      location, 1, rewriter.getIntegerType(8));
+  {
+
+    auto ifZeroIdxDone =
+        rewriter.create<scf::IfOp>(location, TypeRange({}), notZeroIndex, true);
+
+    auto thenBuilder = ifZeroIdxDone.getThenBodyBuilder();
+    auto elseBuilder = ifZeroIdxDone.getElseBodyBuilder();
+    // busy wait until threadZero is finished.
+    {
+      auto whileOp = thenBuilder.create<scf::WhileOp>(
+          location, zeroIdxDone.getType(), zeroIdxDone);
+
+      Block *before = thenBuilder.createBlock(&whileOp.getBefore(), {},
+                                              zeroIdxDone.getType(), location);
+      Block *after = thenBuilder.createBlock(&whileOp.getAfter(), {},
+                                             zeroIdxDone.getType(), location);
+      // before, check if threadZero is finished.
+      {
+        thenBuilder.setInsertionPointToStart(&whileOp.getBefore().front());
+
+        auto zeroIdxDoneVal = thenBuilder.create<memref::AtomicRMWOp>(
[…]

### 2024-04-11 src/test/TestMain.cpp

@@ -2096,273 +2096,273 @@ TestDescriptor testList[] = {
     TEST_LIST_ENTRY(Test_PeeledHybridProbabilisticTiling_TileSize8_Abalone),
 
 #ifdef TREEBEARD_GPU_SUPPORT
-    // GPU model buffer initialization tests (scalar)
-    TEST_LIST_ENTRY(Test_GPUModelInit_LeftHeavy_Scalar_DoubleInt),
-    TEST_LIST_ENTRY(Test_GPUModelInit_RightHeavy_Scalar_DoubleInt),
-    TEST_LIST_ENTRY(Test_GPUModelInit_Balanced_Scalar_DoubleInt),
-    TEST_LIST_ENTRY(Test_GPUModelInit_LeftAndRightHeavy_Scalar_DoubleInt),
-    TEST_LIST_ENTRY(Test_GPUModelInit_LeftHeavy_Scalar_FloatInt),
-    TEST_LIST_ENTRY(Test_GPUModelInit_RightHeavy_Scalar_FloatInt),
-    TEST_LIST_ENTRY(Test_GPUModelInit_Balanced_Scalar_FloatInt),
-    TEST_LIST_ENTRY(Test_GPUModelInit_LeftAndRightHeavy_Scalar_FloatInt),
-    TEST_LIST_ENTRY(Test_GPUModelInit_LeftHeavy_Scalar_FloatInt16),
-    TEST_LIST_ENTRY(Test_GPUModelInit_RightHeavy_Scalar_FloatInt16),
-    TEST_LIST_ENTRY(Test_GPUModelInit_Balanced_Scalar_FloatInt16),
-    TEST_LIST_ENTRY(Test_GPUModelInit_LeftAndRightHeavy_Scalar_FloatInt16),
+// // GPU model buffer initialization tests (scalar)
+// TEST_LIST_ENTRY(Test_GPUModelInit_LeftHeavy_Scalar_DoubleInt),
+// TEST_LIST_ENTRY(Test_GPUModelInit_RightHeavy_Scalar_DoubleInt),
+// TEST_LIST_ENTRY(Test_GPUModelInit_Balanced_Scalar_DoubleInt),
+// TEST_LIST_ENTRY(Test_GPUModelInit_LeftAndRightHeavy_Scalar_DoubleInt),
+// TEST_LIST_ENTRY(Test_GPUModelInit_LeftHeavy_Scalar_FloatInt),
[…]

### 2024-04-11 results/python/analyze_gpu_auto_schedule_results.py
Add schedule comparison across batch sizes

@@ -78,6 +78,34 @@ def write_best_config_maps_to_file(df_best_total_time, benchmark_names, batch_si
     f.write("}\n")
     f.close()
 
+def compare_across_batch_sizes(df_best_kernel_time, df_sorted_by_kernel_time, batch_sizes, benchmarks):
+    # for every benchmark and batch size, get the best schedule from df_best_kernel_time 
+    for batch_size in batch_sizes:
+        for benchmark in benchmarks:
+            best_schedule = df_best_kernel_time[(df_best_kernel_time['model'] == benchmark) & (df_best_kernel_time['batch_size'] == batch_size)]
+            # print(best_schedule)
+            assert (best_schedule.shape[0] == 1)
+            best_schedule_time = best_schedule.iloc[0]['kernel']
+            best_schedule_rep = best_schedule.iloc[0]['representation']
+            for cmp_batch_size in batch_sizes:
+                entry_at_batch_size = df_sorted_by_kernel_time[(df_sorted_by_kernel_time['model'] == benchmark) & 
+                                                               (df_sorted_by_kernel_time['batch_size'] == cmp_batch_size) &
+                                                               (df_sorted_by_kernel_time['representation'] == best_schedule_rep) &
+                                                               (df_sorted_by_kernel_time['rowsPerTB'] == best_schedule.iloc[0]['rowsPerTB']) &
+                                                               (df_sorted_by_kernel_time['rowsPerT'] == best_schedule.iloc[0]['rowsPerT']) &
+                                                               (df_sorted_by_kernel_time['numTreeThreads'] == best_schedule.iloc[0]['numTreeThreads']) &
+                                                               (df_sorted_by_kernel_time['treeInterleaveDepth'] == best_schedule.iloc[0]['treeInterleaveDepth']) &
+                                                               (df_sorted_by_kernel_time['cache_rows'] == best_schedule.iloc[0]['cache_rows']) &
+                                                               (df_sorted_by_kernel_time['cache_trees'] == best_schedule.iloc[0]['cache_trees']) &
[…]

### 2024-04-11 results/python/analyze_gpu_auto_schedule_results.py
print degradation across batch sizes and benchmarks

@@ -1,6 +1,7 @@
 from matplotlib import pyplot as plt
 import pandas as pd
 import numpy as np
+import os
 
 def time_diff_computation(x):
     five_diff = (x.iloc[9] - x.iloc[0])/x.iloc[0]
@@ -80,6 +81,7 @@ def write_best_config_maps_to_file(df_best_total_time, benchmark_names, batch_si
 
 def compare_across_batch_sizes(df_best_kernel_time, df_sorted_by_kernel_time, batch_sizes, benchmarks):
     # for every benchmark and batch size, get the best schedule from df_best_kernel_time 
+    print("benchmark,cur_batch_size,cmp_batch_size,degradation")
     for batch_size in batch_sizes:
         for benchmark in benchmarks:
             best_schedule = df_best_kernel_time[(df_best_kernel_time['model'] == benchmark) & (df_best_kernel_time['batch_size'] == batch_size)]
@@ -101,11 +103,44 @@ def compare_across_batch_sizes(df_best_kernel_time, df_sorted_by_kernel_time, ba
                                                                (df_sorted_by_kernel_time['sharedReduce'] == best_schedule.iloc[0]['sharedReduce'])]
                 # print(entry_at_batch_size)
                 if (entry_at_batch_size.shape[0] != 1):
+                    # print(f'Schedule {best_schedule} not present for {benchmark} at batch size {cmp_batch_size}')
+                    print(f"{benchmark},{batch_size},{cmp_batch_size},NaN")
                     continue
                 cmp_time = entry_at_batch_size.iloc[0]['kernel']
                 best_time_at_cmp_batch_size = df_best_kernel_time[(df_best_kernel_time['model'] == benchmark) & (df_best_kernel_time['batch_size'] == cmp_batch_size)].iloc[0]['kernel']
                 degradation = (cmp_time - best_time_at_cmp_batch_size)/best_time_at_cmp_batch_size
-                print(f"Degradation for {benchmark} at batch size {batch_size} compared to batch size {cmp_batch_size}: {degradation}")
+                print(f"{benchmark},{batch_size},{cmp_batch_size},{degradation}")
+                #print(f"Degradation for {benchmark} at batch size {batch_size} compared to batch size {cmp_batch_size}: {degradation}")
+
[…]

### 2024-04-11 src/reduction/LowerReduceOps.cpp
Fix build error and uncomment gpu tests

@@ -1433,8 +1433,7 @@ struct LowerReductionOps
     : public PassWrapper<LowerReductionOps, OperationPass<mlir::func::FuncOp>> {
 
   void getDependentDialects(DialectRegistry &registry) const override {
-    registry.insert<memref::MemRefDialect, arith::ArithDialect,
-                    omp::OpenMPDialect>();
+    registry.insert<memref::MemRefDialect, arith::ArithDialect>();
   }
 
   void runOnOperation() final {
@@ -1444,7 +1443,7 @@ struct LowerReductionOps
                            decisionforest::DecisionForestDialect,
                            math::MathDialect, arith::ArithDialect,
                            func::FuncDialect, gpu::GPUDialect,
-                           vector::VectorDialect, omp::OpenMPDialect>();
+                           vector::VectorDialect>();
 
     target.addIllegalOp<decisionforest::ReduceOp,
                         decisionforest::ReduceDimensionInplaceOp,

### 2024-04-11 src/test/TestMain.cpp

@@ -2096,273 +2096,273 @@ TestDescriptor testList[] = {
     TEST_LIST_ENTRY(Test_PeeledHybridProbabilisticTiling_TileSize8_Abalone),
 
 #ifdef TREEBEARD_GPU_SUPPORT
-// // GPU model buffer initialization tests (scalar)
-// TEST_LIST_ENTRY(Test_GPUModelInit_LeftHeavy_Scalar_DoubleInt),
-// TEST_LIST_ENTRY(Test_GPUModelInit_RightHeavy_Scalar_DoubleInt),
-// TEST_LIST_ENTRY(Test_GPUModelInit_Balanced_Scalar_DoubleInt),
-// TEST_LIST_ENTRY(Test_GPUModelInit_LeftAndRightHeavy_Scalar_DoubleInt),
-// TEST_LIST_ENTRY(Test_GPUModelInit_LeftHeavy_Scalar_FloatInt),
-// TEST_LIST_ENTRY(Test_GPUModelInit_RightHeavy_Scalar_FloatInt),
-// TEST_LIST_ENTRY(Test_GPUModelInit_Balanced_Scalar_FloatInt),
-// TEST_LIST_ENTRY(Test_GPUModelInit_LeftAndRightHeavy_Scalar_FloatInt),
-// TEST_LIST_ENTRY(Test_GPUModelInit_LeftHeavy_Scalar_FloatInt16),
-// TEST_LIST_ENTRY(Test_GPUModelInit_RightHeavy_Scalar_FloatInt16),
-// TEST_LIST_ENTRY(Test_GPUModelInit_Balanced_Scalar_FloatInt16),
-// TEST_LIST_ENTRY(Test_GPUModelInit_LeftAndRightHeavy_Scalar_FloatInt16),
+    // GPU model buffer initialization tests (scalar)
+    TEST_LIST_ENTRY(Test_GPUModelInit_LeftHeavy_Scalar_DoubleInt),
+    TEST_LIST_ENTRY(Test_GPUModelInit_RightHeavy_Scalar_DoubleInt),
+    TEST_LIST_ENTRY(Test_GPUModelInit_Balanced_Scalar_DoubleInt),
+    TEST_LIST_ENTRY(Test_GPUModelInit_LeftAndRightHeavy_Scalar_DoubleInt),
+    TEST_LIST_ENTRY(Test_GPUModelInit_LeftHeavy_Scalar_FloatInt),
[…]

### 2024-04-11 docs/SOSP 2024/reduction.tex
Edit reduction section

--- 2024-03-13 docs/SOSP 2024/reduction.tex
@@
 \section{Reductions : Representation, Optimization and Lowering}
-
-Currently, existing reduction support in MLIR is insufficient 
-to code generate and optimize the reductions \Treebeard{} needs
-to perform while performing inference (sum up individual 
-tree predictions to compute the prediction of the model). 
-MLIR only supports reductions of value types and cannot 
-directly represent and optimize inplace reductions of 
-several elements of a memory buffer.
-
-We design a mechanism to specify accumulating values into
-an element of a multi-dimensional array inplace. 
-The accumulation is performed inside an arbitrary 
-loop nest where several surrounding loops maybe parallel and 
-the ultimate target machine maybe a CPU or a GPU. Because this problem 
-is of general interest, we design this as an MLIR dialect.
+\label{sec:reduction}
+\Treebeard{} needs to sum up individual tree predictions to compute 
+the prediction of the model while performing inference. However,
+generating fused reductions within arbitrary loop nests specified 
+using \Treebeard{}'s scheduling language is non-trivial. We found 
+that existing reduction support in MLIR is insufficient to code
+generate and optimize these reductions. MLIR only supports reductions
+of value types and does not provide ways to lower reductions to GPUs. 
+To address this gap, we design an MLIR dialect that allows us to
+specify accumulating values into an element of a multi-dimensional array
+and can be lowered to CPU or GPU.
 
@@
 atomically accumulating values into an element of a 
-multi-dimensional array (represented by a MLIR \op{memref}).
-The following example sums up the elements of the 1D memref \op{arr}
-into the first element of the memref \op{result}. It does this
-using in two concurrent iterations of a surrounding 
-parallel loop.
+multi-dimensional array (represented by an MLIR \op{memref}).
+The following example shows how the \op{reduce} op can be used to 
+sum up the elements of an array in parallel.
 
 \begin{lstlisting}[style=c++]
-  builtin.func @ReduceVector(%arr: memref<num_elemsxf64>, %result: memref<1xf64>) -> void {
-    par.for i0 in range(0 : num_elems/2 : num_elems) {
-      for i1 in range(0 : num_elems/2) 
-        reduce(%result, 0, arr[i0 + i1]) <"+", 0.0>
-    }
+  float arr[10], result[1]
+  par.for i0 = 0  to 10 step 5 {
+    for i1 = 0 to 5 
+      reduce(result[0], arr[i0 + i1]) <"+", 0.0>
   }
@@
 
-The semantics of the \op{reduce} op guarantee that all elements are 
-correctly added and that there is no race between the parallel 
-iterations of the loop.
-
+The semantics of the \op{reduce} op is exactly the semantics of 
+an atomic accumulation, i.e. it guarantees that all accumulations 
+are correctly performed even in the presence of parallel loops. 
 The \op{reduce} op is defined for all associative and commutative
@@
 
-The main differences between our \op{reduce} and the existing 
-reductions in MLIR are the following:
-1. Existing reductions only support scalar, by value reductions. 
-It does not support accumulating inplace into memref elements.
-2. Existing reduction support in MLIR does not provide a unified
-way to handle reductions in GPUs and CPUs. To the best of our 
-knowledge, there is currently no way to model reductions on GPUs
-in MLIR.
-2. Existing reductions have strict rules about where they can be 
-written. For example, \op{scf.reduce} needs to be an immediate 
-child \op{scf.parallel}. However, our \op{reduce} op can be
-generated anywhere in the loop nest.
-
-Having modeled the reductions with an abstract op, the 
+Having modeled the reductions with an abstract operation, the 
 aim now is to lower this to a correct and optimized 
-implementation on both CPU and GPU. In order to do this,
-we make the following observations. 
-1. The \op{reduce} op has a loop carried dependency on itself
-and loop carried dependences on other \op{reduce} ops that
-accumulate into the same target array. 
-A simple lowering to a sequence of load-add-store instructions is 
-incorrect if any of these dependences are carried by a parallel 
-loop. We call any such surrounding parallel loop a 
-\textbf{conflicting loop} (\TODO{Change the name!}) for the reduction.
-2. There is a race between the parallel iterations of such a loop 
-when naively accumulating values into target memref elements. 
-To avoid this race, the result memref can be \textbf{privatized} wrt each 
-surrounding conflicting loop. Subsequently, each privatized dimension 
-can be reduced at the end of the conflicting loop it was inserted for. 
+implementation on both CPU and GPU. In order to do this, we 
+first determine if any parallel loop iterations can accumulate 
+into the same array element. We call such loops 
+\emph{\textbf{reduction loops}}. If such loops exist, we 
+\textbf{\emph{privatize}} the array for each iteration of the
+loop. We call this process \textbf{\emph{legalization}}.
+Subsequently, each privatized dimension 
+can be reduced at the end of the reduction loop it was inserted for. 
 \TODO{We cannot do better than this in terms of memory usage} \TODO{Need a proof}.
 
-\begin{definition}
-A parallel loop surrounding one or more \op{reduce} ops is 
-a \textbf{conflicting loop} for a target multi-dimensional array if this 
-loop has a non-zero dependence distance for the dependence between
-any of the contained \op{reduce} ops.
-\end{definition}
+In our example above, parallel iterations of the \op{i0} loop accumulate 
+into the same element of the \op{result} array.
+We would therefore privatize the \op{result} array for each iteration of 
+the \op{i0} loop as follows. 
+\begin{lstlisting}[style=c++]
+  float arr[10], result[1]
+  float resultPriv[2][1]
+  par.for i0 = 0  to 10 step 5 {
+    for i1 = 0 to 5 
+      reduce(resultPriv[i0/5][0], arr[i0 + i1]) <"+", 0.0>
+  }
+  result = reduce_dimension(resultPriv, 0)
+\end{lstlisting}
 
-In the context of \Treebeard{}, this set of loops is exactly the set 
-of surrounding parallel loops that are iterating over trees. The 
-results can be privatized for each conflicting loop iteratively 
-and reductions along each privatized dimension can be inserted  
-immediately following the loop the dimension was inserted due to.
-
-We illustrate this process through the example above. The \op{i0} loop is 
-a conflicting loop for the reduction into the \op{result} array.
-We would therefore privatize the \op{result} memref for each iteration of 
-the \op{i0} loop. 
-\begin{lstlisting}[style=c++]
-  builtin.func @ReduceVector(%arr: memref<num_elemsxf64>, %result: memref<1xf64>) -> void {
-    results_1 = memref<2x1xf64>
-    par.for i0 = range(0 : num_elems/2 : num_elems) {
-      for i1 = range(0 : num_elems/2) 
-        reduce(%result_1[i0/(num_elems/2), 0], arr[i0 + i1]) <"+", 0.0>
-    }
-    results = reduce_dimension(results_1, 0) <"+", 0.0>
-  }
-\end{lstlisting}
 The op \op{reduce\_dimension} reduces values across the specified
-dimension of an n-dimensional memref. In the above example, 
+dimension of an n-dimensional array. In the above example, 
 the \op{reduce\_dimension} op is reducing across all elements 
-of the first dimension (index 0). Therefore, in this case, it 
+of the first dimension (dimension 0). Therefore, in this case, it 
 produces a result memref with a single element (the first dimension
@@
 
-\begin{definition}
-  \op{\textbf{reduce\_dimension(targetMemref, memref, dim, [indices], [rangeStart], [rangeEnd])}}:
-   Computes the reduction over the dimension specified by \op{dimension} and stores the 
-   result in \op{targetMemref}. \op{[indices]} must be a vector of \op{dim} elements
-   (or empty if the dimension being reduced is the first dimension). \op{[rangeStart]} 
-   and \op{[rangeEnd]} represent the range of indices following the reduction dimension and 
-   must have the same number of elements. If both are \op{null} (not passed), 
-   all elements of these dimensions are reduced. The computation performed by the op is as follows.
+% \begin{definition}
+%   A parallel loop surrounding one or more \op{reduce} ops is 
+%   a \textbf{conflicting loop} for a target multi-dimensional array if this 
+%   loop has a non-zero dependence distance for the dependence between
+%   any of the contained \op{reduce} ops.
+% \end{definition}
 
-  $targetMemref[\vec{\boldsymbol{indices}}, \vec{\boldsymbol{k}}] = \sum_{i=0}^{shape[dim]} memref[\vec{\boldsymbol{indices}}, i, \vec{\boldsymbol{k}}]\quad   \forall \vec{\boldsymbol{k}} \in \left[[rangeStart_0, rangeEnd_0), ...
+% \begin{definition}
+%   \op{\textbf{reduce\_dimension(targetMemref, memref, dim, [indices], [rangeStart], [rangeEnd])}}:
+%    Computes the reduction over the dimension specified by \op{dimension} and stores the 
+%    result in \op{targetMemref}. \op{[indices]} must be a vector of \op{dim} elements
+%    (or empty if the dimension being reduced is the first dimension). \op{[rangeStart]} 
+%    and \op{[rangeEnd]} represent the range of indices following the reduction dimension and 
+%    must have the same number of elements. If both are \op{null} (not passed), 
+%    all elements of these dimensions are reduced. The computation performed by the op is as follows.
+
+%   $targetMemref[\vec{\boldsymbol{indices}}, \vec{\boldsymbol{k}}] = \sum_{i=0}^{shape[dim]} memref[\vec{\boldsymbol{indices}}, i, \vec{\boldsymbol{k}}]\quad   \forall \vec{\boldsymbol{k}} \in \left[[rangeStart_0, rangeEnd_0), ...
 , [rangeStart_n, rangeEnd_n)\right]$
-\end{definition}
+% \end{definition}
 
-Consider the following code with nested parallel loops. 
-(A situation where trees are split across both
-threads and thread blocks could result in such generated 
-code in \Treebeard{}.)
-\begin{lstlisting}[style=c++]
-  builtin.func @ReduceVector(%arr: memref<num_elemsxf64>, %result: memref<1xf64>) -> void {
-    par.for i0 = range(0 : num_elems/2 : num_elems) {
-      par.for i1 = range(0 : num_elems/4 : num_elems/2) {
-        for i2 = range(0 : num_elems/4) 
-          reduce(%result, 0, arr[i0 + i1 + i2]) <"+", 0.0>
-      }
-    }
-  }
-\end{lstlisting}
+% Consider the following code with nested parallel loops. 
+% (A situation where trees are split across both
+% threads and thread blocks could result in such generated 
+% code in \Treebeard{}.)
+% \begin{lstlisting}[style=c++]
+%   builtin.func @ReduceVector(%arr: memref<num_elemsxf64>, %result: memref<1xf64>) -> void {
+%     par.for i0 = range(0 : num_elems/2 : num_elems) {
+%       par.for i1 = range(0 : num_elems/4 : num_elems/2) {
+%         for i2 = range(0 : num_elems/4) 
+%           reduce(%result, 0, arr[i0 + i1 + i2]) <"+", 0.0>
+%       }
+%     }
+%   }
+% \end{lstlisting}
 
-Here, the \op{i0} and \op{i1} loops are conflicting loops wrt 
-the \op{result} memref. We \textbf{legalize} the reduction 
-by privatizing the \op{result} array wrt the \op{i0} and \op{i1} loops.
-However, there are now two privatized dimensions and therefore, 
-two dimensions need to be reduced to compute the final result.
-This multi-stage reduction is what enables us to model 
-hierarchical reductions.
+% Here, the \op{i0} and \op{i1} loops are conflicting loops wrt 
+% the \op{result} memref. We \textbf{legalize} the reduction 
+% by privatizing the \op{result} array wrt the \op{i0} and \op{i1} loops.
+% However, there are now two privatized dimensions and therefore, 
+% two dimensions need to be reduced to compute the final result.
+% This multi-stage reduction is what enables us to model 
+% hierarchical reductions.
 
-The following code shows how the reduction above is legalized. 
-We introduce a new op, \op{reduce\_dimension\_inplace} which 
-reduces a dimension of the input memref and stores results 
-in the same array. This helps saves memory by removing the 
-need to create multiple intermediate arrays to store results.
-Only the final dimension reduction uses the \op{reduce\_dimension} op.
+% The following code shows how the reduction above is legalized. 
+% We introduce a new op, \op{reduce\_dimension\_inplace} which 
+% reduces a dimension of the input memref and stores results 
+% in the same array. This helps saves memory by removing the 
+% need to create multiple intermediate arrays to store results.
+% Only the final dimension reduction uses the \op{reduce\_dimension} op.
 
-\begin{lstlisting}[style=c++]
-  builtin.func @ReduceVector(%arr: memref<num_elemsxf64>, %result: memref<1xf64>) -> void {
-    results_1 = memref<2x2x1xf64>
-    par.for i0 = range(0 : num_elems/2 : num_elems) {
-      par.for i1 = range(0 : num_elems/4 : num_elems/2) {
-        for i2 = range(0 : num_elems/4) 
-          index0 = i0/(num_elems/2)
-          index1 = i1/(num_elems/4)
-          reduce(%result_1[index0, index1, 0], arr[i0 + i1 + i2]) <"+", 0.0>
-      }
-      // result_1[i0/(num_elems/2), 0] = sum(result_1[i0/(num_elems/2), :])
-      reduce_dimension_inplace(%result_1, 1, i0/(num_elems/2)) 
-    }
-    // result = sum(result[:, 0])
-    %result = reduce_dimension(%result_1, 0)
-  }
-\end{lstlisting}
+% \begin{lstlisting}[style=c++]
+%   builtin.func @ReduceVector(%arr: memref<num_elemsxf64>, %result: memref<1xf64>) -> void {
+%     results_1 = memref<2x2x1xf64>
+%     par.for i0 = range(0 : num_elems/2 : num_elems) {
+%       par.for i1 = range(0 : num_elems/4 : num_elems/2) {
+%         for i2 = range(0 : num_elems/4) 
+%           index0 = i0/(num_elems/2)
+%           index1 = i1/(num_elems/4)
+%           reduce(%result_1[index0, index1, 0], arr[i0 + i1 + i2]) <"+", 0.0>
+%       }
+%       // result_1[i0/(num_elems/2), 0] = sum(result_1[i0/(num_elems/2), :])
+%       reduce_dimension_inplace(%result_1, 1, i0/(num_elems/2)) 
+%     }
+%     // result = sum(result[:, 0])
+%     %result = reduce_dimension(%result_1, 0)
+%   }
+% \end{lstlisting}
 
@@
 how these operations are lowered to the CPU and GPU.
+
+In the context of \Treebeard{}, the set of reduction loops is exactly the set 
+of surrounding parallel loops that are iterating over trees. The 
+results can be privatized for each loop iteratively 
+and reductions along each privatized dimension can be inserted  
+immediately following the loop the dimension was inserted due to.
 

### 2024-04-11 src/python/treebeard.py
Intermediate submission for tree parallelization on CPU

@@ -132,7 +132,10 @@ class CompilerOptions:
 
   def SetBasicGPUSchedule(self, tileSize : int) :
     treebeardAPI.runtime_lib.SetBasicGPUSchedule(self.optionsPtr, tileSize)
-       
+
+  def SetNumberOfParallelTreeBatches(self, val : int) :
+    treebeardAPI.runtime_lib.Set_numParallelTreeBatches(self.optionsPtr, val)
+
 #### ---------------------------------------------------------------- ####
 #### Index Variable
 #### ---------------------------------------------------------------- ####

### 2024-04-11 src/python/treebeard_runtime_api.py

@@ -92,6 +92,9 @@ class TreebeardAPI:
       self.runtime_lib.Set_numberOfCores.argtypes = [ctypes.c_int64, ctypes.c_int32]
       self.runtime_lib.Set_numberOfCores.restype = None
 
+      self.runtime_lib.Set_numParallelTreeBatches.argtypes = [ctypes.c_int64, ctypes.c_int32]
+      self.runtime_lib.Set_numParallelTreeBatches.restype = None
+
       self.runtime_lib.Set_statsProfileCSVPath.argtypes = [ctypes.c_int64, ctypes.c_char_p]
       self.runtime_lib.Set_statsProfileCSVPath.restype = None
 

### 2024-04-11 src/runtime/runtime.cpp

@@ -156,6 +156,7 @@ COMPILER_OPTION_SETTER(statsProfileCSVPath, const char *)
 COMPILER_OPTION_SETTER(pipelineSize, int32_t)
 COMPILER_OPTION_SETTER(numberOfCores, int32_t)
 // COMPILER_OPTION_SETTER(compileToGPU, bool)
+COMPILER_OPTION_SETTER(numParallelTreeBatches, int32_t)
 
 extern "C" void Set_tilingType(intptr_t options, int32_t val) {
   TreeBeard::CompilerOptions *optionsPtr =

### 2024-04-11 test/python/treebeard_parallel_benchmarks.py

@@ -154,6 +154,33 @@ def ConstructCompilerOptionsList(tile_size, pipeline_width):
       optionsList.append(invertLoopsTileSize8Options)
   return optionsList
 
+def ConstructTreeParallelCompilerOptionsList(tile_size, pipeline_width):
+  invertLoopsTileSize8Options = treebeard.CompilerOptions(batchSize, tile_size)
+  invertLoopsTileSize8Options.SetPipelineWidth(pipeline_width)
+  invertLoopsTileSize8Options.SetReorderTreesByDepth(True)
+  invertLoopsTileSize8Options.SetMakeAllLeavesSameDepth(1)
+  if run_parallel:
+    invertLoopsTileSize8Options.SetNumberOfCores(int(num_cores/4))
+    invertLoopsTileSize8Options.SetNumberOfParallelTreeBatches(4)
+
+  invertLoopsTileSize8MulticlassOptions = treebeard.CompilerOptions(batchSize, tile_size)
+  invertLoopsTileSize8MulticlassOptions.SetReturnTypeWidth(8)
+  invertLoopsTileSize8MulticlassOptions.SetReturnTypeIsFloatType(False)
+  invertLoopsTileSize8MulticlassOptions.SetPipelineWidth(pipeline_width)
+  invertLoopsTileSize8MulticlassOptions.SetReorderTreesByDepth(True)
+  invertLoopsTileSize8MulticlassOptions.SetMakeAllLeavesSameDepth(1)
+  if run_parallel:
+    invertLoopsTileSize8MulticlassOptions.SetNumberOfCores(int(num_cores/4))
+    invertLoopsTileSize8MulticlassOptions.SetNumberOfParallelTreeBatches(4)
+
+  optionsList = []
+  for model in modelNames:
+    if model=="covtype" or model=="letters":
+      optionsList.append(invertLoopsTileSize8MulticlassOptions)
+    else:
+      optionsList.append(invertLoopsTileSize8Options)
+  return optionsList
+
 def print_title_row():
   print("Title", "Batch Size", sep=", ", end="")
   for model in modelNames:
@@ -173,28 +200,29 @@ def run_benchmarks_for_batchsize():
   xgboost_speedups = []
   treelite_speedups = []
   options = ConstructCompilerOptionsList(8, 8)
+  optionsTreePar = ConstructTreeParallelCompilerOptionsList(8, 8)
   i = 0
   for model in modelNames:
[…]

### 2024-04-13 docs/SOSP 2024/abstract.tex
Update abstract

@@
 
-  The rapid proliferation of machine learning coupled with accelerating evolution of the hardware ecosystem has led to a surge in the demand for model inference on a variety of hardware.
-  %While CPUs have been the mainstay for machine learning inference, the availability of GPUs holds promise to scale to bigger and more powerful models. 
-  This paper is motivated by the problems encountered when targeting inference of decision tree based models, the most popular models on tabular data, to run at peak performance on 
-  diverse CPU and GPU targets. We evaluated existing solutions and found that they do not provide portable performance across different hardware targets.
-  %Decision tree based models are widely used in practice due to their robustness, interpretability, and ability to handle missing data.
+  This paper is motivated by the growing demand for the increased performance
+  of machine learning applications on different hardware platforms
+  including CPUs and GPUs. We focus on accelerating the inference of decision tree based models, 
+  which are the most popular models for tabular data. Existing solutions do not achieve 
+  the highest possible performance because they do not explore different optimization
+  configurations. And since these systems are hand-written, they are not portable either.
 
-  To address this we present the design of \Treebeard{}, a schedule guided compiler  
-  for decision tree based models that searches over a large design space 
-  to automatically generate high-performance inference routines.
-  %We re-architect the open-source \TreebeardOLD{} infrastructure and significantly extend it to enable high-performance code generation across target processors. 
-  \Treebeard{} has two core components. A scheduling language 
+  We address these problems by designing \Treebeard{}, a \emph{schedule-guided},
+  \emph{retargetable} compiler infrastructure for decision tree based models.  
+  \Treebeard{} has two core components. The first is a scheduling language 
   that encapsulates the large optimization space for decision tree inference, 
-  and techniques to efficiently explore this space.
-  \TODO{Change large optimization space}.
-  %We also design a set of heuristics that can find near-optimal solutions in quick time.
-  Second, an optimizing multi-level compiler that can generate code based on the schedule. 
-  For the latter, we re-architect the open-source \TreebeardOLD{} CPU compiler to support schedule guided compilation and
-  add support for GPU code generation. GPU code generation required fundamental new optimization interfaces for 
-  caching, parallel reduction, and support for multiple in-memory representations of trees.
+  and techniques to efficiently explore this space. \TODO{Change large optimization space}.
+  The second is an optimizing retargetable multi-level compiler that 
+  can generate code for any specified schedule. \Treebeard{}'s retargetability 
+  is based not only on being able to generate code for different target architectures (CPU vs. GPU),
+  but also on its ability to use different data layouts, caching strategies,
+  parallel reduction schemes etc. To accomplish this level of configurability, we re-architect 
+  and significantly extend the open-source \TreebeardOLD{} CPU compiler
+  to support (i) schedule-guided compilation, (ii) retargetable GPU code
+  generation, and (iii) GPU-specific optimizations. 
+  %The scheduler and retargetable code generator components work hand-in-glove, synergizing and mutually benefiting from each other.
 
-  %The compiler builds on an existing compiler for CPUs, \TreebeardOLD{}, also add support for caching, parallel reduction and a plug-in mechanism to explore different in-memory representations of trees.
-
-  We evaluate \Treebeard{} on over seven hundred diverse models and demonstrate that the best schedule varies drastically with model, batch size, and target hardware. 
-  Our scheduling heuristic is able to quickly find near optimal schedules while searching over a small number (\~50) of schedules.
+  We demonstrate that \Treebeard{} can generate high-performance inference 
+  code, for several hundred decision tree models across different batch sizes
+  and target architectures. Our scheduling heuristic is able to quickly find near-optimal schedules 
+  \TODO{[how do we argue that the schedule is near-optimal? Do we have some 
+  exptl. evidence that we can show in the results section?]}
+  schedule while searching over a small number ($\sim$50) of schedules.
   In terms of performance, \Treebeard{} generated code is an order of magnitude faster than XGBoost and
-  about 2-3$\times$ faster on average than RAPIDs FIL and Tahoe. While these systems only target NVIDIA GPUs, \Treebeard{} achieves competent performance on AMD GPUs as well. 
+  about 2-3$\times$ faster on average than RAPIDs FIL and Tahoe. While these systems only 
+  target NVIDIA GPUs, \Treebeard{} achieves competent performance on AMD GPUs as well. 
   On CPUs, \Treebeard{} achieves better scaling compared to \TreebeardOLD{}.

### 2024-04-13 docs/SOSP 2024/reduction.tex
Edit reduction section

@@
 
-The behavior of the \op{reduce\_dimension\_inplace} op is similar to the 
+To reduce the amount of memory used by arrays introduced for reduction,
+we introduce the \op{reduce\_dimension\_inplace} operation. It is similar to the 
 \op{reduce\_dimension} op except that it updates the input array inplace
-rather than writing results to a target array. The definition of the 
-op is as follows.
+rather than writing results to a target array. It writes results to the 
+zeroth index of the dimension being reduced.
 
-\begin{definition}
- \textbf{\op{reduce\_dimension\_inplace(memref, dim, [indices], [rangeStart], [rangeEnd])}}:
-  Computes the reduction over the dimension specified by \op{dimension} and stores the 
-  result at index 0 of that dimension. \op{[indices]} must be a vector of \op{dim} elements
-   (or empty if the dimension being reduced is the first dimension). \op{[rangeStart]} 
-   and \op{[rangeEnd]} must have the same number of elements. If both are \op{null} (not passed), 
-   all elements of the corresponding dimension are reduced.
+% \begin{definition}
+%  \textbf{\op{reduce\_dimension\_inplace(memref, dim, [indices], [rangeStart], [rangeEnd])}}:
+%   Computes the reduction over the dimension specified by \op{dimension} and stores the 
+%   result at index 0 of that dimension. \op{[indices]} must be a vector of \op{dim} elements
+%    (or empty if the dimension being reduced is the first dimension). \op{[rangeStart]} 
+%    and \op{[rangeEnd]} must have the same number of elements. If both are \op{null} (not passed), 
+%    all elements of the corresponding dimension are reduced.
 
-  The computation performed by the op is defined by the following equation.
+%   The computation performed by the op is defined by the following equation.
 
-  $memref[\vec{\boldsymbol{indices}}, 0, \vec{\boldsymbol{k}}] = \sum_{i=0}^{shape[dim]} memref[\vec{\boldsymbol{indices}}, i, \vec{\boldsymbol{k}}]\quad   \forall \vec{\boldsymbol{k}} \in \left[[rangeStart_0, rangeEnd_0), ...
+%   $memref[\vec{\boldsymbol{indices}}, 0, \vec{\boldsymbol{k}}] = \sum_{i=0}^{shape[dim]} memref[\vec{\boldsymbol{indices}}, i, \vec{\boldsymbol{k}}]\quad   \forall \vec{\boldsymbol{k}} \in \left[[rangeStart_0, rangeEnd_0), ...
 , [rangeStart_n, rangeEnd_n)\right]$
-\end{definition}
+% \end{definition}
 
@@
 We implement lowering of the operations defined above to both the CPU and GPU.
-Since the compilation pipeline diverges after the reductions are legalized, 
-we can implement lowering and optimization of our reduction dialect to CPUs and
-GPUs simply using different MLIR rewrite patterns. We now briefly describe 
+Since the lowering pipeline from MIR to LIR are different for CPU and GPU 
+compilation, we implement lowering and optimization of our reduction dialect to CPUs and
+GPUs simply using different MLIR rewrite patterns. In this section, we briefly describe 
 how these operations are lowered to the CPU and GPU.
 
-In the context of \Treebeard{}, the set of reduction loops is exactly the set 
-of surrounding parallel loops that are iterating over trees. The 
-results can be privatized for each loop iteratively 
-and reductions along each privatized dimension can be inserted  
-immediately following the loop the dimension was inserted due to.
-
 \subsubsection{Lowering to CPU}
-The lowering of the reduction operations to CPU is fairly straightforward. We lower the 
-two operations listed above, \op{reduce\_dimension\_inplace} and 
+The lowering of the reduction operations to CPU is fairly straightforward. We lower 
+\op{reduce\_dimension\_inplace} and 
 \op{reduce\_dimension} to a simple loop nest that goes over the specified
@@
 from the input array as a vector, accumulated as a vector, and 
-finally written back to the target array. In general, this works 
-well because reductions are typically being performed on dimensions
-other than the inner-most dimension and therefore, this strategy
-loads successive elements from memory maximizing memory bandwidth 
-utilization.
+finally written back to the target array.
 
-\TODO{explain atomic reduction}
+% In general, this works 
+% well because reductions are typically being performed on dimensions
+% other than the inner-most dimension and therefore, this strategy
+% loads successive elements from memory maximizing memory bandwidth 
+% utilization.
+
+% \TODO{explain atomic reduction}
 
 \subsubsection{Lowering to GPU}
-The lowering on GPU is slightly more involved than the lowering on CPUs.
-However, we can lower the same abstractions to efficient implementations
-and therefore simplify high-level code generation. The lowering for 
+The same abstractions can be lowered to efficient GPU implementations
+and therefore, simplify higher-level code generation. The lowering for 
 the inplace and non-inplace operations are essentially the same, except 
@@
 
-The lowering of the \op{reduce\_dimension\_*} ops is distinct from existing 
-work on implementing reductions efficiently on GPUs \cite{NVIDIAReductions}
-because our abstractions potentially represent several independent reductions
-(independent for different output elements).
-Therefore, we can either exploit parallelism across the independent reductions or 
+The lowering of the \op{reduce\_dimension\_*} ops can either exploit
+parallelism across the independent reductions or 
 the inherent parallelism in the reduction by performing a divide and conquer 
-reduction.
-
-The reduction pass for GPU can follow one of two paths. If the lowering pass
-determines that there are enough independent reductions to keep all threads
-in a thread block busy, then it simply generates code that performs one (or 
-multiple) reductions completely in a thread. If however there are not 
+reduction. If there are enough independent reductions to keep all threads
+in a thread block busy, then the lowering pass can generate code that performs one (or 
+multiple) reductions in each thread. If, however, there are not 
 enough independent reductions, then the lowering pass generates a tree 
@@
 using shared memory, the privatized buffer is allocated in shared memory. 
-Also, the compiler ensures that only as much shared memory is allocated 
-as needed to hold values processed by a single thread-block and 
-index offsets are appropriately rewritten to handle the differences between 
-the indexing of the target memref and the shared memory array.
+The compiler only allocates as much shared memory 
+as needed to hold values processed by a single thread-block.
+%  and 
+% index offsets are appropriately rewritten to handle the differences between 
+% the indexing of the target memref and the shared memory array.
 Our abstractions allow our lowering passes to be written completely 
@@
 \subsection{Use in \Treebeard{}}
-We now present an example specific to \Treebeard{}. The schedule with
-which code is generated is as below. \op{N\_t} is the number of trees 
+We now show how \Treebeard{} uses the reduction dialect to generate 
+code for decision tree inference using an example. 
+In our example, \op{N\_t} is the number of trees 
 and \op{batch\_size} is the batch size. The schedule tiles both the 
-batch loop and the tree loop and parallelizes the outer batch 
-and tree loops.
+batch and tree loops and parallelizes the outer batch 
+and tree loops. The schedule with which code is generated is as follows.
+% In the context of \Treebeard{} we note that, the set of reduction loops is exactly the set 
+% of surrounding parallel loops that are iterating over trees. The 
+% results can be privatized for each loop iteratively 
+% and reductions along each privatized dimension can be inserted  
+% immediately following the loop the dimension was inserted due to.
 
 \begin{lstlisting}[style=c++]
-IndexVar i0, i1, t0, t1;
-auto& batch = schedule.GetBatchIndex();
-auto& tree = schedule.GetTreeIndex();
-schedule.Tile(batch, i0, i1, batch_size/2);
-schedule.Tile(tree, t0, t1, N_t/2);
-schedule.Reorder({i0, t0, t1, i1});
-schedule.Parallel(t0);
-schedule.Parallel(i0);
+tile(batch, i0, i1, batch_size/2);
+tile(tree, t0, t1, N_t/2);
+reorder({i0, t0, t1, i1});
+parallel(t0);
+parallel(i0);
 \end{lstlisting}
 
-The loop-nest generated by \Treebeard{} for the above schedule is as follows. 
+The MIR generated by \Treebeard{} for the above schedule is as follows. 
 \begin{lstlisting}[style=c++]
-  builtin.func @Prediction_Function(%arg0: memref<batch_sizexnum_featuresxf64>) -> memref<batch_sizexf64> {
-    %result = memref.alloc <batch_sizexf64>
-    %0 = #decisionforest<ReductionType = 0, #Trees = N_t, resultType = memref<batch_sizexf64>> 
-    par.for i0 = range(0 : batch_size/2 : batch_size) {
-      par.for t0 = range(0 : N_t/2 : N_t) {
-        for t1 = range(0 : N_t/2) {
-          for i1 = range(0 : batch_size/2) {
-            %2 = GetTree(%0, t0 + t1) 
-            %3 = WalkDecisionTree(%2, %arg0[i0+i1])
-            reduce(%result, i0+i1, %3)
-          }
+  float result[batch_size]
+  model = ensemble(...) 
+  par.for i0 = 0 to batch_size step batch_size/2 {
+    par.for t0 = 0 to N_t step N_t/2 {
+      for t1 = 0 to N_t/2 {
+        for i1 = 0 to batch_size/2 {
+          t = getTree(model, t0 + t1) 
+          p = walkDecisionTree(t, rows[i0+i1])
+          reduce(result[i0+i1], p)
         }
@@
 
-\Treebeard{} determines that the \op{t0} loop is a conflicting loop for the \op{result}
-array and therefore legalizes the reduction by inserting a privatized array 
-\op{result\_1}. The privatized dimension of this array is reduced at the end 
-of the \op{t0} loop.
+\Treebeard{} determines that the \op{t0} loop is a reduction 
+loop w.r.t the \op{result} array and therefore legalizes 
+the reduction by inserting a privatized array 
+\op{partResults}. The privatized dimension of this array 
+is reduced at the end of the \op{t0} loop.
 
 \begin{lstlisting}[style=c++]
-  builtin.func @Prediction_Function(%arg0: memref<batch_sizexnum_featuresxf64>) -> memref<batch_sizexf64> {
-    %result = memref.alloc <batch_sizexf64>
-    %result_1 = memref.alloc <2xbatch_sizexf64>
-    %0 = #decisionforest<ReductionType = 0, #Trees = N_t, resultType = memref<batch_sizexf64>> 
-    par.for i0 = range(0 : batch_size/2 : batch_size) {
-      par.for t0 = range(0 : N_t/2 : N_t) {
-        for t1 = range(0 : N_t/2) {
-          for i1 = range(0 : batch_size/2) {
-            %2 = GetTree(%0, t0 + t1) 
-            %3 = WalkDecisionTree(%2, %arg0[i0+i1])
-            reduce(%result, i0+i1, %3)
+  float result[batch_size], partResults[2][batch_size]
+  model = ensemble(...) 
+  par.for i0 = 0 to batch_size step batch_size/2 {
+    par.for t0 = 0 to N_t step N_t/2 {
+      for t1 = 0 to N_t/2 {
+        for i1 = 0 to batch_size/2 {
+          t = getTree(model, t0 + t1) 
+          p = walkDecisionTree(t, rows[i0+i1])
+          reduce(result[i0+i1], p)
         }
@@
     }
-      %result[i0 : i0+step] = reduce_dimension(%result_1, 0, i0 : i0+step)
-    }
+    results[i0:i0+batch_size/2] = reduce_dimension(partResults[:, i0:i0+batch_size/2], 0)
   }
@@
 While legalizing the reduction, the compiler determines that the 
-\op{reduce\_dimension} operation must only process a subset of the final 
-result that is computed within the current parallel iteration of the 
-\op{i0} loop. Once this process is complete, the \op{reduce} ops in 
-the result IR can be lowered to a simple ``read-accumulate-write''
-sequence of instructions
+\op{reduce\_dimension} operation can only compute a subset of the final 
+result (the subset that is computed within the current parallel iteration of the 
+\op{i0} loop). 
+% Once this process is complete, the \op{reduce} ops in 
+% the result IR can be lowered to a simple ``read-accumulate-write''
+% sequence of instructions
 
@@
 
-\TODO{Should we mention how we handle multi-class models?}
+%\TODO{Should we mention how we handle multi-class models?}

### 2024-04-14 docs/SOSP 2024/caching.tex
Fix compile errors. Add latest pdf
Edit caching section

--- 2024-04-04 docs/SOSP 2024/caching.tex
@@
-\section{Representations}
+\section{Model Representations}
 \label{sec:representations}
@@
 the three representations shown in Figure \ref{Fig:Representations}. The array and sparse 
-representations are the ones described in the \Treebeard{} paper\cite{Treebeard}. The reorg 
-representation is the representation used by the RAPIDs library\cite{RAPIDs}. 
+representations are the ones described in the \TreebeardOLD{} paper\cite{Treebeard}. The reorg 
+representation is the representation used by the RAPIDs library\cite{FIL}. 
 The \textbf{array representation} is the simplest representation where the trees are stored in an array
-in level order. The \textbf{sparse representation} stores the trees in a sparse format where only the
-non-zero nodes are stored and nodes contain pointers to their children. The \textbf{reorg representation}
+in level order. The \textbf{sparse representation} stores the trees in a sparse format where 
+memory is allocated only for nodes present in the tree and nodes contain pointers to their children. The \textbf{reorg representation}
 interleaves the array representation of each tree in the model: all root nodes are stored first, then 
-the left children of all the roots and so on.
+the left children of all the roots and so on. This representation was 
+designed to improve memory coalescing when tree nodes are being loaded. 
 \begin{figure}[htb]
@@
 
-One of the major changes we make to the original design of \Treebeard{} \cite{Treebeard} is to
+One of the major changes we make to the original design of \TreebeardOLD{} \cite{Treebeard} is to
 separate the implementation of representations from the rest of the compiler. This allows
-us to implement representations as plugins to the compiler. We do this by defining an interface 
-that each representation needs to implement. The code generator is implemented using methods 
-on the representation interface thus hiding details of the actual representation from the core 
-compiler. The interface abstracts the following details. 
-\begin{itemize}
-  \item \textbf{Buffer allocation:} The representation object exposes methods that generate 
-  buffer allocations in the IR. 
-  \item \textbf{Moving to child nodes:} Given the value of the predicate at a node (or tile),
-  generate code to move to the appropriate child node of the current node.
-  \item \textbf{Details of leaves:} The interface abstracts determining whether the current
-  node is a leaf (which is needed to terminate walks) and how to get the value of the leaf
-  (which is needed to get a tree's prediction).
-  \item \textbf{Caching trees:} Since reading the trees into shared memory on GPU (or prefetching)
-  them on CPU require knowledge of the buffer layout, the task of generating caching code for trees
-  is delegated to the representation object. 
-  \item \textbf{Loading thresholds and feature indices:} The representation object provides MLIR 
-  rewrite patterns to lower operations that load thresholds and feature indices to LLVM IR. This
-  is necessary because the details of how to load the thresholds and feature indices from the model 
-  buffers is representation-specific.
-\end{itemize}
+us to implement representations as plugins to the compiler. We define an interface 
+that representations implement. The code generator is implemented using
+this interface thus hiding details of the actual representation from the core 
+compiler. Curcially, the interface abstracts how and what buffers 
+are allocated, how to move from a node to its child, how trees 
+are cached, reading the value of leaves and now threshold and 
+feature indices are read from the allocated buffers.
+% The interface abstracts the following details. 
+% \begin{itemize}
+%   \item \textbf{Buffer allocation:} The representation object exposes methods that generate 
+%   buffer allocations in the IR. 
+%   \item \textbf{Moving to child nodes:} Given the value of the predicate at a node (or tile),
+%   generate code to move to the appropriate child node of the current node.
+%   \item \textbf{Leaf representation:} The interface abstracts determining whether the current
+%   node is a leaf (which is needed to terminate walks) and how to get the value of the leaf
+%   (which is needed to get a tree's prediction).
+%   \item \textbf{Caching trees:} Since reading the trees into shared memory on GPU (or prefetching)
+%   them on CPU require knowledge of the buffer layout, the task of generating caching code for trees
+%   is delegated to the representation object. 
+%   \item \textbf{Loading thresholds and feature indices:} The representation object provides MLIR 
+%   rewrite patterns to lower operations that load thresholds and feature indices to LLVM IR. This
+%   is necessary because the details of how to load the thresholds and feature indices from the model 
+%   buffers is representation-specific.
+% \end{itemize}
 
@@
 compiler infrastructure. Implementing the representations as plugins also allows us to reuse
-the implementations across different lowering pipelines. For example, the code for the array 
-and sparse representations is almost fully shared between the CPU and GPU lowering pipelines.
+the implementations across different lowering pipelines.
 
-\section{Caching Implementation}
+% For example, the code for the array 
+% and sparse representations is almost fully shared between the CPU and GPU lowering pipelines.
+
+\section{Caching}
 \label{sec:caching}
@@
 on both the CPU and GPU. As described in Section \ref{sec:schedule}, 
-the user can specify that the working set of an iteration of an index
-variable needs to be cached using the \op{Cache()} directive. This 
+the user can specify that the working set of an iteration of a loop
+needs to be cached using the \op{cache} directive. This 
 provides a unified way to specify caching of both trees and input rows.
+\Treebeard{} implements caching at the granularity of a tree or a row.
 
-\Treebeard{} implements caching at the granularity of a tree or a row.
-Also, the semantics of caching depends on the target processor. For the CPU,
-caching is implemented as prefetching, while for the GPU, caching is implemented
-using shared memory.
-
-\subsection{IR Representation of Caching}
-Caching is encoded in the mid-level IR using the \op{cacheTrees} and \op{cacheRows} operations. 
-These operations are generated when the HIR lowered to MIR and \op{Cache()} is specified on an 
+% \subsection{IR Representation of Caching}
+Caching is encoded in the mid-level IR using the \op{cacheTrees} and \op{cacheRows} operations (Table \ref{Tab:IRSpecification}). 
+These operations are generated when the HIR lowered to MIR and \op{cache} is specified on an 
 index variable in the schedule. While the HIR is being lowered and a cached index variable is 
 encountered, the compiler generates a \op{cacheTrees} or \op{cacheRows} operation depending on 
-whether the index variable is a tree or a batch index variable. Additionally, as a part of the 
-lowering process, \Treebeard{} determines the working set of the loop with caching enabled 
+whether the index variable is a tree or a batch index variable.
+\Treebeard{} also determines the working set of the loop  
 and generates a caching operation with the appropriate limits.
 
-Each of the caching operations take parameters that specify the set of trees or rows that need 
-to be cached. The caching operations are defined as follows. 
-\begin{itemize}
-  \item \textbf{\op{cacheTrees(forest, start, end)}:} This operation caches the trees in ensemble
-  \op{forest} from \op{start} to \op{end}. The trees are cached in the order in which they are
-  specified in the ensemble. 
-  \item \textbf{\op{cacheRows(data, start, end)}:} This operation caches the rows in the input 
-  array \op{data} from \op{start} to \op{end}. The rows are cached in the same order as in the
-  input array.
-\end{itemize}
+% Each of the caching operations take parameters that specify the set of trees or rows that need 
+% to be cached. The caching operations are defined as follows. 
+% \begin{itemize}
+%   \item \textbf{\op{cacheTrees(forest, start, end)}:} This operation caches the trees in ensemble
+%   \op{forest} from \op{start} to \op{end}. The trees are cached in the order in which they are
+%   specified in the ensemble. 
+%   \item \textbf{\op{cacheRows(data, start, end)}:} This operation caches the rows in the input 
+%   array \op{data} from \op{start} to \op{end}. The rows are cached in the same order as in the
+%   input array.
+% \end{itemize}
 
-\subsection{Lowering of Caching Operations}
+%\subsection{Lowering of Caching Operations}
 When the MIR is lowered to LIR, the cache ops are lowered to target-specific code. Each of 
-the two caching operations is lowered differently for the CPU and the GPU.
+the two caching operations is lowered differently for the CPU and the GPU. On CPU, the cahce operations are lowered to preteches while
+on the GPU they are lowered to reads into shared memory.
 
-For the \op{cacheRows} operation, \Treebeard{} uses pre-implemented lowerings for both
-CPU and GPU. This is possible because the input is currently assumed to be a dense
-array format. Therefore, regardless of any other configuration choices (like what representation 
-is used for the model itself), the lowering of the \op{cacheRows} operation is the same. 
-For the CPU, the \op{cacheRows} operation is lowered to prefetches. For the GPU, a 
-series of coalesced loads read the rows into shared memory.
+Lowering the \op{cacheRows} operation is straightforward
+% \Treebeard{} uses pre-implemented lowerings for both
+% CPU and GPU. This is possible 
+because the input is currently assumed to be a dense
+array format. The lowering for the \op{cacheRows} operation is implemented directly in the \Treebeard{} compiler.
 
 For the \op{cacheTrees} operation, the lowering is representation-specific. Each representation
-provides a lowering to the target-specific code generator to lower the \op{cacheTrees} op when
-that representation is used. However, the \Treebeard{} infrastructure does provide helpers 
-to generate caching code to cache contiguous regions of memory. These helpers are reused 
-as required across different representations.
+provides a lowering to the target-specific code 
+generator to lower the \op{cacheTrees} op when
+that representation is used.
+%  However, the \Treebeard{} infrastructure does provide helpers 
+% to generate caching code to cache contiguous regions of memory. These helpers are reused 
+% as required across different representations.
 

### 2024-04-14 docs/SOSP 2024/intro.tex
Prof. Govind's intro edits

@@
 by the necessity to innovate in the post Moore's law and Dennard's scaling era. This transformation has also played 
-a significant role in the success of modern deep learning models, as they enable scaling model training and inference 
-to a massive number of threads. Such scalability would be essential for all performance critical applications, including
+a significant role in the success of modern deep learning models, as they enable scaling 
+%% RG - minor rewording
+%% model training and inference to a massive number of threads. 
+training and inference to models with billions of parameters across a massive number of threads. 
+Such scalability would be essential for all performance critical applications, including
 other machine learning models that need to scale with increasing data sizes and model complexities. 
-%need to accelerate machine learning workloads.
-The success of modern AI models is due to their ability to scale both training and inference to a massive number of threads.
-This has been made possible by the rise of GPUs and accelerators.
+%need to accelerate machine learning workloads. The success of modern AI models is due to their 
+%ability to scale both training and inference to a massive number of threads. This has been made
+%possible by the rise of GPUs and accelerators.
 
@@
 This is because, in production settings, each model is trained once and often used for inference millions of times. 
-However, inference is run on a variety of hardware platforms, ranging from low to high-end CPUs and GPUs. 
-This paper is motivated by the need to accelerate decision tree inference to achieve portable performance on a variety of hardware.
+%% RG: Minor reword
+%% However, 
+Further, inference is run on a variety of hardware platforms, ranging from low to high-end CPUs and GPUs. 
+This paper is motivated by the need to accelerate decision tree inference to achieve portable performance on 
+%% RG: reword -- specifically say CPUs and GPUs.
+%% a variety of hardware.
+commodity platforms with CPUs and GPUs. 
 % In particular, we focus on a range of commodity CPUs and GPUs, a class of hardware that has seen widespread 
@@
 setting, where multiple inputs are processed simultaneously.
-Despite the simplicity of the model and the availability of multiple sources of coarse grain parallelism (parallelism 
+Despite the simplicity of the model and the availability of multiple sources of coarse-grain parallelism (parallelism 
 across inputs in a batch and parallelism across trees), existing systems do not consistently scale well across different
@@
 %Parallelism, loop structure and model representation together determine the accesses patterns and the working set. 
-Existing systems today are mostly library based, and only support a predefined combination of optimizations. 
+Existing systems today are mostly library based, and only support a predefined combination of optimizations
+and typically only target a single platform. 
 XGBoost~\cite{XGBoost} uses a sparse representation for the model and a loop structure that processes one tree for a block of rows before moving to 
-the next tree. RAPIDS FIL~\cite{FIL} uses a reorg representation and partitions trees across a fixed number of threads. Tahoe~\cite{Tahoe} uses a
+the next tree. RAPIDS FIL~\cite{FIL} uses a reorg representation and partitions trees across a fixed number of threads. 
+%% RG: 
+% \TODO{Should we say RAPIDS is only for nVidia GPUs?} 
+%%
+Tahoe~\cite{Tahoe} uses a
 variation of the reorg representation and has four predefined inference strategies from which it picks one based on an analytical model.
+All these systems are CUDA based and only work on NVIDIA based GPUs.
 %These systems are all library based and are hard to extend. 
-\TreebeardOLD{}, the state-of-the-art model compiler for CPUs, supports two fixed loop structures and does not scale well with increasing number of threads.
-Additionally, it lacks GPU specific optimizations that are critical to scale performance to massive number of threads.
+\TreebeardOLD{}, the state-of-the-art decision tree model compiler for CPUs 
+%% RG
+built using the MLIR infrastructure~\cite{MLIR}, supports two fixed loop structures 
+and does not scale well with increasing number of threads. Additionally, it lacks 
+GPU specific optimizations that are critical to scale performance to massive number of threads.
+\TODO{Shouldn't we reverse this? First say no GPU support and then the scaling issue.}
 
-This paper presents \Treebeard{}, a novel schedule guided compilation infrastructure for decision tree inference on multiple target hardware.
-\Treebeard{} is able to generate high-performance code for decision tree inference by exploring a large optimization space.
-The code generation is guided by a \emph{schedule} written in a custom scheduling language that can represent a wide range of implementation strategies.
-We demonstrate that the language is sufficient to express the various optimizations proposed by prior work, and further generalizes them and incorporates several new optimizations.
-We also design and implement a heuristic that is able to quickly find high-performance schedules for the model being compiled.
+This paper presents \Treebeard{}, a novel schedule guided compilation infrastructure
+for decision tree inference on multiple target hardware. \Treebeard{} is able to 
+generate high-performance code for decision tree inference by exploring a large optimization space. 
+%% RG - rewording
+%% The code generation is guided by a \emph{schedule} written in a custom scheduling language that can represent a wide range of implementation strategies. 
+This is achieved by a compilation framework consisting of a 
+custom scheduling language that can represent a wide range of 
+implementation strategies and techniques to efficiently explore
+the optimization space. We demonstrate that the language is 
+sufficient to express the various optimizations proposed by prior work
+and that our schedule exploration heuristic can quickly find a near 
+optimal schedule for the model being compiled. 
+%% Further we generalizes them and incorporates several new optimizations. 
+\TODO{RG: it would be good to expand on this and also talk about the retargetable component. 
+We could also say that the schedule framework and the retargetable compiler work in an 
+interwined manner, each benefitting from the other.}  
+The second component of \Treebeard{} is a \emph{retargetable} multi-level compiler that can generate efficient code 
+for any specified schedule for both CPUs and GPUs. For this purpose, we re-architect \TreebeardOLD{} to
+support schedule-guided code generation, and incorporate several new optimizations. These two components 
+of the proposed \Treebeard{} compiler infrastructure are interwined, each benefitting from the other.
+%Last, to address the challenge of selecting a near-optimal schedule from the vast optimization space, 
+%%
+%we design and implement a heuristic that is able to quickly find high-performance schedules for the model being compiled.
 \TODO{Performance evaluation summary}
-\TODO{Oganization}
+%% RG: This may not be needed -- can be removed to save space!
+%% \TODO{Oganization}
 \subsection{Contributions}
 \begin{itemize}
-  \item We present the design for a multi-target decision tree compiler infrastructure and implement several optimizations within this framework.
-   We are also the first to implement an optimizing compiler for decision tree inference on GPUs.
-  \item We identify that an extensive optimization space exists for the problem of decision tree inference. We design a scheduling language that 
-  allows us to effectively represent this solution space abstractly. This scheduling language is expressive enough to represent a wide range of  
-  implementation strategies proposed by prior work.
-  \item To the best of our knowledge, we perform the first extensive characterization of the optimization space for decision tree inference 
-  on GPUs. Using some of the characteristics we identify, we design and implement a heuristic that is able to quickly find high-performance 
+  \item We present the design of a multi-target compiler infrastructure for decision tree 
+   inferencing  and implement several optimizations within this framework.
+   We are also the first to implement an optimizing compiler for decision tree 
+   inference on GPUs. \TODO{AP: Given that there is Hummingbird, can we really say this?}
+  \item We identify that an extensive optimization space exists for the problem of 
+  decision tree inference. We design a scheduling language that 
+  allows us to effectively represent this solution space abstractly. This 
+  scheduling language is expressive enough to represent a wide range of  
+  implementation strategies,  %%.  proposed by prior work.
+%% RG: added CPUs and GPUs. 
+  such as different data layouts, caching strategies, parallel reduction schemes, 
+  that work across CPUs and GPUs. 
+  \item To the best of our knowledge, we perform the first extensive characterization
+  of the optimization space for decision tree inference 
+  on GPUs. Using some of the characteristics we identify, we design and implement 
+  a heuristic that is able to quickly find high-performance 
   schedules for the model being compiled. 
-  \item We design and implement a general framework for expressing and optimizing reductions within MLIR. To the best of our knowledge, this is the first
+  \item We design and implement a general framework for expressing and optimizing
+  reductions within MLIR. To the best of our knowledge, this is the first
   such framework.
-  \item We evaluate our implementation by comparing it against RAPIDs and Tahoe, the state-of-the-art decision tree inference frameworks for GPU and 
-  report significant speedups. We also show that our compiler can effectively target different GPUs, including both NVIDIA and AMD GPUs.
+  \item We evaluate our implementation by comparing it against RAPIDs and Tahoe, 
+  the state-of-the-art decision tree inference frameworks for GPU and 
+  report significant speedups. We also show that our compiler can effectively 
+  target different GPUs, including both NVIDIA and AMD GPUs.
 \end{itemize}

### 2024-04-14 results/python/analyze_schedule_sensitivity_data.py
Add sensitivity results

@@ -1,0 +1,129 @@
+# plot a 3-d graph of num_trees, num_features and tb_kernel_speedup
+import pandas as pd
+import numpy as np
+import xarray as xr
+import matplotlib.pyplot as plt
+from mpl_toolkits.mplot3d import Axes3D
+
+# get the path of the current script
+import os
+script_path = os.path.dirname(os.path.realpath(__file__))
+# append ../thedeep/gpu/sensitivity to the path
+sensitivity_path = os.path.join(script_path, "..", "thedeep", "gpu", "sensitivity")
+
+def get_degradation_vals(df_benchmark, X, Y, x_col="cur_batch_size", y_col="cmp_batch_size"):
+    # get the degradation values for the given benchmark
+    degradation_vals = np.zeros((len(X), len(Y)))
+    for i in range(len(X)):
+        for j in range(len(Y)):
+            x = X[i]
+            y = Y[j]
+            df_depth = df_benchmark[(df_benchmark[x_col] == x) & (df_benchmark[y_col] == y)]
+            # assert df_depth.shape[0] == 1, f"Expected 1 row for {x}, {y}, got {df_depth.shape[0]}"
+            if df_depth.shape[0] == 0:
+                degradation_vals[i, j] = 5
+                continue
+            degradation_vals[i, j] = df_depth["degradation"].values[0]
+            # if x < y:
+            #     degradation_vals[i, j] = 10 * degradation_vals[i, j]
+    return degradation_vals
+
+def plot_batch_sensitivity_for_single_benchmark(df_depth, benchmark_name):
+    # draw an intensity plot of the degradation values
+    # x axis is cur_batch_size, y axis is cmp_batch_size, intensity is degradation
+    # x and y axis in log scale
+
+    # get a numpy array for the x, y and z axis
+    x = df_depth["cur_batch_size"].unique()
+    y = df_depth["cmp_batch_size"].unique()
+    x.sort()
+    y.sort()
+    Z = get_degradation_vals(df_depth, x, y) # df_depth['degradation'].values.reshape(len(y), len(x))
+    # print(Z)
+
+    da = xr.DataArray(
+        Z,
+        coords={"Batch Size": x, "Comparison Batch Size": y},
+        dims=("Batch Size", "Comparison Batch Size"),
+    )
+
+    da.plot(xscale="log", yscale="log")
+    # plt.show()
[…]

### 2024-04-15 docs/SOSP 2024/caching.tex
Move schedule exploration to different file

@@
 % as required across different representations.
-
-\section{Exploring the Schedule Space}
-\label{sec:exploring}
-
-% \begin{itemize}
-%   \item The scheduling language described in Section \ref{sec:schedule} allows
-%   for numerous possible schedules for a given model. 
-%   Finding a schedule with good performance is a non-trivial task.
-%   \item We identify a reasonable template schedule for GPUs which encompasses
-%   all strategies published in prior work.
-%   \item Describe the template schedule.
-%   \item Even within the variants of this template schedule, there is a significant 
-%   amount of variation in performance. Show the histogram. 
-%   \item Extensively exploring all possible parameter values for the template schedule 
-%   is infeasible due to the large number of parameters.
-%   Cite some numbers on how long it takes to explore the schedule space.
-%   \item List the observations we make and the heuristic we design based on these.
-% \end{itemize}
-
-The set of schedules that can be constructed using the scheduling language described in 
-Section \ref{sec:schedule} is vast. Searching this schedule space to find a schedule that
-provides good performance is a non-trivial task. To simplify this process, we identify a
-template schedule for GPUs that encompasses several strategies published in prior work.
-Our template schedule assigns a fixed number of rows to each thread block and to each thread.
-It distributes the trees across a fixed number of threads and can cache trees and 
-input rows if required. Unrolling and interleaving of tree walks is also supported.
-
-The template schedule exposes a number of parameters that can be tuned to find a 
-high-performance schedule.
-\begin{itemize}
-  \item \textbf{Number of rows per thread block (Integer):} The number of rows that are processed by each thread block.
-  \item \textbf{Number of rows per thread (Integer):} The number of rows processed by each thread.
-  \item \textbf{Number of tree threads (Integer):} The number of threads across which the trees are distributed.
-  \item \textbf{Cache rows (Boolean):} Whether the input rows are cached in shared memory.
-  \item \textbf{Cache trees (Boolean):} Whether the trees are cached in shared memory.
-  \item \textbf{Unroll walks (Boolean):} Whether the tree walks are unrolled.
-  \item \textbf{Tree walk interleave factor:} The number of tree walks that are interleaved.
-  \item \textbf{Shared memory reduction:} Whether the reduction across tree threads is done in shared memory.
-\end{itemize}
-
-While the template schedule simplifies optimization of generated inference code, it is 
-important to note that the \Treebeard{} compiler itself does not place any 
-restrictions on the schedule. The user is free to specify any schedule they wish.
-The auto-scheduler that implements the template schedule is implemented as a 
-module outside the core \Treebeard{} compiler. Users are also free to implement 
-other auto-schedulers that generate schedules different from the template schedule.
-
-\begin{figure}[htb]
-  \centering
-  \includegraphics[width=\linewidth]{figures/normalized_kernel_histogram_lt5.png}
-  \caption{Distribution of normalized execution times for all benchmark models
-  with different parameter values for the template schedule. \TODO{We need to 
-  describe exactly what the considered set of parameter values are.}}
-  \label{Fig:ExecTimeDistribution}
-\end{figure}
-
-Figure \ref{Fig:ExecTimeDistribution} shows the distribution of normalized execution times
-for all benchmark models with different parameter values for the template schedule.
-The normalized execution time is the inference time of the model with a given set of
-parameter values divided by the best inference time of the model.
-The histogram shows that even within the variants of the template schedule, there is
-a significant amount of variation in performance. Very few schedules perform close to the best
-while a vast majority of schedules perform poorly.
-
-Exploring the schedule space extensively even for a reasonable set of parameter values
-is very expensive. We explored the set of parameter values listed in Table \ref{table:ParamVals}
-for our benchmarks and found that it took anywhere between thirty minutes up to a few hours
-to explore the entire space. Performing this extensive search for every model being compiled 
-is infeasible in practise. We therefore need a better mechanism to guide the search for a good schedule.
-
-We design a heuristic to narrow down the set of schedules to explore based on the following 
-observations on high-performance schedules.
-\begin{itemize}
-  \item For small batch sizes, the best schedules tend to have a small number of rows
-  per thread block and partition the trees across a larger number of threads. This is 
-  intuitive since the amount of data parallelism across the rows is limited for small batch sizes.
-  \item Always cache rows in shared memory and never cache trees. We find that caching rows 
-  when possible (i.e., when the number of features is small enough to fit in shared memory)
-  almost always improves performance. Caching trees on the other hand almost always degrades 
-  performance. This is because the one time cost of loading trees into shared memory is 
-  not sufficiently amortized when the whole of the tree is not accessed during inference. 
-  \item Models with a large number of features tend to benefit from partitioning the 
-  trees across more threads even at larger batch sizes. This is because processing fewer rows at a time 
-  allows us to keep them in shared memory. We empirically find that the threshold for when 
-  we should start partitioning the trees across more threads is when the number of features
-  is greater than 100. 
-  \item We find that when a model prefers schedules with shared reduction, the same schedules 
-  without shared reduction are among the best performing schedules without shared reduction.
-  We therefore are able to separate the evaluation of shared reduction by collecting the 
-  best schedules without shared reduction and only evaluating shared reduction on them. 
-  Evaluating the top 3 schedules for shared reduction is sufficient in practice.  
-\end{itemize}
-
-\Treebeard{} uses these observations to narrow down the set of schedules to explore. 
-The pseudo-code for the current heuristic is shown in Algorithm \ref{alg:heuristic}.
-The algorithm first computes a subset of thread block configurations in the function 
-\op{TBConfigs}. A set of schedules based on these thread block configurations 
-is then computed (\op{schedules}). The model is compiled with each of these schedules 
-and then the resulting inference code is profiled. The three best performing schedules 
-are collected and shared reduction is enabled on them and the resulting schedules evaluated.
-The best schedule among all the evaluated schedules is selected as the schedule to use.
-We find that this heuristic is able to find schedules that are close to the best schedules
-but improves the search time by two orders of magnitude as we show in Section \ref{sec:HeuristicResults}.
-
-\begin{algorithm}
-  \caption{Heuristic to guide the search for a good schedule}
-  \label{alg:heuristic}
-  \begin{algorithmic}[1]
-  \Procedure{TBConfigs}{$N_{batch}$, $N_f$}
-    \State $T_{batch} \gets 2048, T_f \gets 128$
-    \If {$N_{batch} \leq T_{batch}$}
-      \State $rowsPerBlock \gets \{8, 32\}$
-      \State $treeThreads \gets \{20, 50\}$
-    \Else
-      \If {$N_{f} \leq T_{f}$}
-        \State $rowsPerBlock \gets \{32, 64\}$
-        \State $treeThreads \gets \{2, 10\}$
-      \Else
-        \State $rowsPerBlock \gets \{8, 32\}$
-        \State $treeThreads \gets \{20, 50\}$
-      \EndIf
-    \EndIf
-    \State \Return $rowsPerBlock, treeThreads$
-  \EndProcedure
-  \\
-  \State $bestSchedules \gets PriorityQueue(3)$
-  \State $rowsPerTB, treeThreads \gets TBConfigs(N_{batch}, N_f)$
-  \State $cacheRows \gets \text{True}$
-  \State $cacheTrees \gets \text{False}$
-  \State $interleaveFactors \gets \{1, 2, 4\}$
-  \State $reps \gets \{array, sparse, reorg\}$
-  \State $schedules \gets (rowsPerTB, treeThreads,$ \\
-                          \hspace{1.5cm}$cacheRows, cacheTrees,$\\ 
-                          \hspace{1.5cm}$interleaveFactors)$
-  \For{$(sched, rep) \in schedules \times reps$}
-    \State $time \gets EvaluateSchedule(sched, rep)$
-    \State $bestSchedules.insert(time, sched, rep)$
-  \EndFor
-  \State $shMemSchedules \gets \emptyset$
-  \For{$sched, rep \in bestSchedules$}
-    \State $EnableSharedReduction(sched)$
-    \State $time \gets EvaluateSchedule(sched, rep)$
-    \State $shMemSchedules.insert(time, sched, rep)$
-  \EndFor
-  \State \Return $min(shMemSchedules \cup bestSchedules)$
-  \end{algorithmic} 
-\end{algorithm}

### 2024-04-15 docs/SOSP 2024/exploring.tex
Move schedule exploration to different file

\section{Exploring the Schedule Space}
\label{sec:exploring}

% \begin{itemize}
%   \item The scheduling language described in Section \ref{sec:schedule} allows
%   for numerous possible schedules for a given model. 
%   Finding a schedule with good performance is a non-trivial task.
%   \item We identify a reasonable template schedule for GPUs which encompasses
%   all strategies published in prior work.
%   \item Describe the template schedule.
%   \item Even within the variants of this template schedule, there is a significant 
%   amount of variation in performance. Show the histogram. 
%   \item Extensively exploring all possible parameter values for the template schedule 
%   is infeasible due to the large number of parameters.
%   Cite some numbers on how long it takes to explore the schedule space.
%   \item List the observations we make and the heuristic we design based on these.
% \end{itemize}

The set of schedules that can be constructed using the scheduling language described in 
Section \ref{sec:schedule} is vast. Searching this schedule space to find a schedule that
provides good performance is a non-trivial task. To simplify this process, we identify a
template schedule for GPUs that encompasses several strategies published in prior work.
Our template schedule assigns a fixed number of rows to each thread block and to each thread.
It distributes the trees across a fixed number of threads and can cache trees and 
input rows if required. Unrolling and interleaving of tree walks is also supported.

The template schedule exposes a number of parameters that can be tuned to find a 
high-performance schedule.
\begin{itemize}
  \item \textbf{Number of rows per thread block (Integer):} The number of rows that are processed by each thread block.
  \item \textbf{Number of rows per thread (Integer):} The number of rows processed by each thread.
  \item \textbf{Number of tree threads (Integer):} The number of threads across which the trees are distributed.
  \item \textbf{Cache rows (Boolean):} Whether the input rows are cached in shared memory.
  \item \textbf{Cache trees (Boolean):} Whether the trees are cached in shared memory.
  \item \textbf{Unroll walks (Boolean):} Whether the tree walks are unrolled.
  \item \textbf{Tree walk interleave factor:} The number of tree walks that are interleaved.
  \item \textbf{Shared memory reduction:} Whether the reduction across tree threads is done in shared memory.
\end{itemize}

While the template schedule simplifies optimization of generated inference code, it is 
important to note that the \Treebeard{} compiler itself does not place any 
restrictions on the schedule. The user is free to specify any schedule they wish.
The auto-scheduler that implements the template schedule is implemented as a 
module outside the core \Treebeard{} compiler. Users are also free to implement 
other auto-schedulers that generate schedules different from the template schedule.

\begin{figure}[htb]
  \centering
  \includegraphics[width=\linewidth]{figures/normalized_kernel_histogram_lt5.png}
  \caption{Distribution of normalized execution times for all benchmark models
  with different parameter values for the template schedule. \TODO{We need to 
  describe exactly what the considered set of parameter values are.}}
  \label{Fig:ExecTimeDistribution}
\end{figure}

Figure \ref{Fig:ExecTimeDistribution} shows the distribution of normalized execution times
for all benchmark models with different parameter values for the template schedule.
The normalized execution time is the inference time of the model with a given set of
parameter values divided by the best inference time of the model.
The histogram shows that even within the variants of the template schedule, there is
a significant amount of variation in performance. Very few schedules perform close to the best
while a vast majority of schedules perform poorly.

Exploring the schedule space extensively even for a reasonable set of parameter values
is very expensive. We explored the set of parameter values listed in Table \ref{table:ParamVals}
for our benchmarks and found that it took anywhere between thirty minutes up to a few hours
to explore the entire space. Performing this extensive search for every model being compiled 
is infeasible in practise. We therefore need a better mechanism to guide the search for a good schedule.

We design a heuristic to narrow down the set of schedules to explore based on the following 
observations on high-performance schedules.
\begin{itemize}
  \item For small batch sizes, the best schedules tend to have a small number of rows
  per thread block and partition the trees across a larger number of threads. This is 
  intuitive since the amount of data parallelism across the rows is limited for small batch sizes.
  \item Always cache rows in shared memory and never cache trees. We find that caching rows 
  when possible (i.e., when the number of features is small enough to fit in shared memory)
  almost always improves performance. Caching trees on the other hand almost always degrades 
  performance. This is because the one time cost of loading trees into shared memory is 
  not sufficiently amortized when the whole of the tree is not accessed during inference. 
  \item Models with a large number of features tend to benefit from partitioning the 
  trees across more threads even at larger batch sizes. This is because processing fewer rows at a time 
  allows us to keep them in shared memory. We empirically find that the threshold for when 
  we should start partitioning the trees across more threads is when the number of features
  is greater than 100. 
  \item We find that when a model prefers schedules with shared reduction, the same schedules 
  without shared reduction are among the best performing schedules without shared reduction.
  We therefore are able to separate the evaluation of shared reduction by collecting the 
  best schedules without shared reduction and only evaluating shared reduction on them. 
  Evaluating the top 3 schedules for shared reduction is sufficient in practice.  
\end{itemize}

\Treebeard{} uses these observations to narrow down the set of schedules to explore. 
The pseudo-code for the current heuristic is shown in Algorithm \ref{alg:heuristic}.
The algorithm first computes a subset of thread block configurations in the function 
\op{TBConfigs}. A set of schedules based on these thread block configurations 
is then computed (\op{schedules}). The model is compiled with each of these schedules 
and then the resulting inference code is profiled. The three best performing schedules 
are collected and shared reduction is enabled on them and the resulting schedules evaluated.
The best schedule among all the evaluated schedules is selected as the schedule to use.
We find that this heuristic is able to find schedules that are close to the best schedules
but improves the search time by two orders of magnitude as we show in Section \ref{sec:HeuristicResults}.

\begin{algorithm}
  \caption{Heuristic to guide the search for a good schedule}
  \label{alg:heuristic}
  \begin{algorithmic}[1]
  \Procedure{TBConfigs}{$N_{batch}$, $N_f$}
    \State $T_{batch} \gets 2048, T_f \gets 128$
    \If {$N_{batch} \leq T_{batch}$}
      \State $rowsPerBlock \gets \{8, 32\}$
      \State $treeThreads \gets \{20, 50\}$
    \Else
      \If {$N_{f} \leq T_{f}$}
        \State $rowsPerBlock \gets \{32, 64\}$
        \State $treeThreads \gets \{2, 10\}$
      \Else
        \State $rowsPerBlock \gets \{8, 32\}$
        \State $treeThreads \gets \{20, 50\}$
      \EndIf
    \EndIf
    \State \Return $rowsPerBlock, treeThreads$
  \EndProcedure
  \\
  \State $bestSchedules \gets PriorityQueue(3)$
  \State $rowsPerTB, treeThreads \gets TBConfigs(N_{batch}, N_f)$
  \State $cacheRows \gets \text{True}$
  \State $cacheTrees \gets \text{False}$
  \State $interleaveFactors \gets \{1, 2, 4\}$
  \State $reps \gets \{array, sparse, reorg\}$
  \State $schedules \gets (rowsPerTB, treeThreads,$ \\
                          \hspace{1.5cm}$cacheRows, cacheTrees,$\\ 
                          \hspace{1.5cm}$interleaveFactors)$
  \For{$(sched, rep) \in schedules \times reps$}
    \State $time \gets EvaluateSchedule(sched, rep)$
    \State $bestSchedules.insert(time, sched, rep)$
  \EndFor
  \State $shMemSchedules \gets \emptyset$
  \For{$sched, rep \in bestSchedules$}
    \State $EnableSharedReduction(sched)$
    \State $time \gets EvaluateSchedule(sched, rep)$
    \State $shMemSchedules.insert(time, sched, rep)$
  \EndFor
  \State \Return $min(shMemSchedules \cup bestSchedules)$
  \end{algorithmic} 
\end{algorithm}

### 2024-04-15 docs/SOSP 2024/exploring.tex
Edit schedule epxloration section

@@
 The set of schedules that can be constructed using the scheduling language described in 
-Section \ref{sec:schedule} is vast. Searching this schedule space to find a schedule that
-provides good performance is a non-trivial task. To simplify this process, we identify a
+Section \ref{sec:schedule} is unbounded. Searching this schedule space to find a 
+high-performance schedule is a non-trivial task. To simplify this process, we design a
 template schedule for GPUs that encompasses several strategies published in prior work.
-Our template schedule assigns a fixed number of rows to each thread block and to each thread.
-It distributes the trees across a fixed number of threads and can cache trees and 
-input rows if required. Unrolling and interleaving of tree walks is also supported.
+Our template schedule assigns a configurable number of rows to each thread block and to each thread.
+It distributes the trees across a specified number of threads and can cache trees and 
+input rows if required while unrolling and interleaving of tree walks.
+Table \ref{tab:schedparams} lists the parameter values we tried. 
+% \begin{itemize}
+%   \item \textbf{Number of rows per thread block (Integer):} The number of rows that are processed by each thread block.
+%   \item \textbf{Number of rows per thread (Integer):} The number of rows processed by each thread.
+%   \item \textbf{Number of tree threads (Integer):} The number of threads across which the trees are distributed.
+%   \item \textbf{Cache rows (Boolean):} Whether the input rows are cached in shared memory.
+%   \item \textbf{Cache trees (Boolean):} Whether the trees are cached in shared memory.
+%   \item \textbf{Unroll walks (Boolean):} Whether the tree walks are unrolled.
+%   \item \textbf{Tree walk interleave factor:} The number of tree walks that are interleaved.
+%   \item \textbf{Shared memory reduction:} Whether the reduction across tree threads is done in shared memory.
+% \end{itemize}
 
-The template schedule exposes a number of parameters that can be tuned to find a 
-high-performance schedule.
-\begin{itemize}
-  \item \textbf{Number of rows per thread block (Integer):} The number of rows that are processed by each thread block.
-  \item \textbf{Number of rows per thread (Integer):} The number of rows processed by each thread.
-  \item \textbf{Number of tree threads (Integer):} The number of threads across which the trees are distributed.
-  \item \textbf{Cache rows (Boolean):} Whether the input rows are cached in shared memory.
-  \item \textbf{Cache trees (Boolean):} Whether the trees are cached in shared memory.
-  \item \textbf{Unroll walks (Boolean):} Whether the tree walks are unrolled.
-  \item \textbf{Tree walk interleave factor:} The number of tree walks that are interleaved.
-  \item \textbf{Shared memory reduction:} Whether the reduction across tree threads is done in shared memory.
-\end{itemize}
+\begin{table}[htb]
+  \centering
+  \resizebox{\linewidth}{!}{
+  \begin{tabularx}{\linewidth}{c | c }
+  \toprule
+  \textbf{Parameter} & \textbf{Values} \\
+  \midrule
+  \textbf{Rows per thread block} & $\{8, 32, 64\}$ \\
+  \textbf{Rows per thread} & $\{1, 2, 4\}$ \\
+  \textbf{Number of tree threads} & $\{2, 10, 20, 50\}$ \\
+  \textbf{Cache rows} & $\{True, False\}$ \\
+  \textbf{Cache trees} & $\{True, False\}$ \\
+  \textbf{Unroll walks} & $\{True, False\}$ \\
+  \textbf{Tree walk interleave factor} & $\{1, 2, 4\}$ \\
+  \textbf{Shared memory reduction} & $\{True, False\}$ \\
+  \bottomrule
+  \end{tabularx}
+  }
+  \vskip 5pt
+  \caption{\label{tab:schedparams} List of parameter values we explored for the template GPU schedule.}
+\end{table}
 
-While the template schedule simplifies optimization of generated inference code, it is 
-important to note that the \Treebeard{} compiler itself does not place any 
+%While the template schedule simplifies optimization of generated inference code,
+It is important to note that the \Treebeard{} compiler itself does not place any 
 restrictions on the schedule. The user is free to specify any schedule they wish.
-The auto-scheduler that implements the template schedule is implemented as a 
-module outside the core \Treebeard{} compiler. Users are also free to implement 
-other auto-schedulers that generate schedules different from the template schedule.
+The compiler pass that implements the template schedule is also implemented as a 
+module outside the core \Treebeard{} compiler. 
+% Users are also free to implement 
+% other auto-schedulers that generate schedules different from the template schedule.
 
@@
   \caption{Distribution of normalized execution times for all benchmark models
-  with different parameter values for the template schedule. \TODO{We need to 
-  describe exactly what the considered set of parameter values are.}}
+  with the template schedule using parameter values as shown in Table \ref{tab:schedparams}}. 
   \label{Fig:ExecTimeDistribution}
@@
 
+While the template schedule simplifies code generation, finding a good set of 
+parameter values is still hard.
 Figure \ref{Fig:ExecTimeDistribution} shows the distribution of normalized execution times
-for all benchmark models with different parameter values for the template schedule.
-The normalized execution time is the inference time of the model with a given set of
-parameter values divided by the best inference time of the model.
-The histogram shows that even within the variants of the template schedule, there is
-a significant amount of variation in performance. Very few schedules perform close to the best
-while a vast majority of schedules perform poorly.
+for all benchmark models with different parameter values for the template schedule (inference
+times normalized w.r.t fastest time for that model).
+There is a significant amount of variation in performance even within the variants of the
+template schedule,. Very few schedules perform close to the best while a vast majority of
+schedules perform poorly.
 
 Exploring the schedule space extensively even for a reasonable set of parameter values
-is very expensive. We explored the set of parameter values listed in Table \ref{table:ParamVals}
+is very expensive. We explored the set of parameter values listed in Table \ref{tab:schedparams}
 for our benchmarks and found that it took anywhere between thirty minutes up to a few hours
-to explore the entire space. Performing this extensive search for every model being compiled 
-is infeasible in practise. We therefore need a better mechanism to guide the search for a good schedule.
-
-We design a heuristic to narrow down the set of schedules to explore based on the following 
-observations on high-performance schedules.
+to explore the entire space for each model. 
+%Performing this extensive search for every model being compiled is infeasible in practise. 
+We therefore
+%need a better mechanism to guide the search for a good schedule.
+design a heuristic to narrow down the set of schedules to explore based on the following 
+observations.
 \begin{itemize}
@@
   \item Always cache rows in shared memory and never cache trees. We find that caching rows 
-  when possible (i.e., when the number of features is small enough to fit in shared memory)
+  %when possible (i.e., when the number of features is small enough to fit in shared memory)
   almost always improves performance. Caching trees on the other hand almost always degrades 
@@
   trees across more threads even at larger batch sizes. This is because processing fewer rows at a time 
-  allows us to keep them in shared memory. We empirically find that the threshold for when 
-  we should start partitioning the trees across more threads is when the number of features
-  is greater than 100. 
+  allows us to keep them in shared memory.
+  % We empirically find that the threshold for when 
+  % we should start partitioning the trees across more threads is when the number of features
+  % is greater than 100. 
   \item We find that when a model prefers schedules with shared reduction, the same schedules 
   without shared reduction are among the best performing schedules without shared reduction.
-  We therefore are able to separate the evaluation of shared reduction by collecting the 
-  best schedules without shared reduction and only evaluating shared reduction on them. 
+  % We therefore are able to separate the evaluation of shared reduction by collecting the 
+  % best schedules without shared reduction and only evaluating shared reduction on them. 
   Evaluating the top 3 schedules for shared reduction is sufficient in practice.  
@@
 \Treebeard{} uses these observations to narrow down the set of schedules to explore. 
-The pseudo-code for the current heuristic is shown in Algorithm \ref{alg:heuristic}.
+The pseudo-code for the heuristic is shown in Algorithm \ref{alg:heuristic}.
 The algorithm first computes a subset of thread block configurations in the function 
@@
 We find that this heuristic is able to find schedules that are close to the best schedules
-but improves the search time by two orders of magnitude as we show in Section \ref{sec:HeuristicResults}.
+but improves the search time by two orders of magnitude as we show in Section \ref{sec:results}.
 
 \begin{algorithm}
-  \caption{Heuristic to guide the search for a good schedule}
+  \caption{Heuristic to find a good schedule}
   \label{alg:heuristic}
@@
   \Procedure{TBConfigs}{$N_{batch}$, $N_f$}
-    \State $T_{batch} \gets 2048, T_f \gets 128$
-    \If {$N_{batch} \leq T_{batch}$}
+    \State $T_{batch} \gets 2048,\; T_f \gets 128$
+    \If {$N_{batch} \leq T_{batch}$ \textbf{or} $N_{f} > T_{f}$}
       \State $rowsPerBlock \gets \{8, 32\}$
@@
     \Else
-      \If {$N_{f} \leq T_{f}$}
       \State $rowsPerBlock \gets \{32, 64\}$
       \State $treeThreads \gets \{2, 10\}$
-      \Else
-        \State $rowsPerBlock \gets \{8, 32\}$
-        \State $treeThreads \gets \{20, 50\}$
-      \EndIf
     \EndIf
@@
   \\
-  \State $bestSchedules \gets PriorityQueue(3)$
-  \State $rowsPerTB, treeThreads \gets TBConfigs(N_{batch}, N_f)$
-  \State $cacheRows \gets \text{True}$
-  \State $cacheTrees \gets \text{False}$
-  \State $interleaveFactors \gets \{1, 2, 4\}$
-  \State $reps \gets \{array, sparse, reorg\}$
-  \State $schedules \gets (rowsPerTB, treeThreads,$ \\
-                          \hspace{1.5cm}$cacheRows, cacheTrees,$\\ 
-                          \hspace{1.5cm}$interleaveFactors)$
-  \For{$(sched, rep) \in schedules \times reps$}
+  \State $bestSchedules \gets shMemSchedules \gets \emptyset$
+  \State $rowsPerTB, treeThds \gets TBConfigs(N_{batch}, N_f)$
+  \State $cacheRows \gets \text{True},\; cacheTrees \gets \text{False}$
+  \State $interleave \gets \{1, 2, 4\}$
+  % \State $reps \gets \{array, sparse, reorg\}$
+  \State $schedules \gets (rowsPerTB, treeThds, cacheRows,$ \\
+                          \hspace{2cm}$cacheTrees, interleave)$
+                          %\hspace{2cm}$$
+  \For{$(sched, rep) \in schedules \times \{array, sparse, reorg\}$}
     \State $time \gets EvaluateSchedule(sched, rep)$
@@
   \EndFor
-  \State $shMemSchedules \gets \emptyset$
-  \For{$sched, rep \in bestSchedules$}
+  \State 
+  \For{$sched, rep \in Top3(bestSchedules)$}
     \State $EnableSharedReduction(sched)$

### 2024-04-15 docs/SOSP 2024/related.tex
Edit related work

@@
 studied in the literature, to the best of our knowledge, no systems that are 
-capable of exploring the extensive optimization space exist. We describe related work 
+capable of exploring the full optimization space exist. We describe related work 
 and compare these systems to \Treebeard{} in this section.
 
-\emph{Decision Tree Inference Systems for GPUs:} 
+\emph{Decision Tree Inference Systems:} 
 Tahoe\cite{Tahoe} is a system that implements high-performance library routines and a 
@@
 structure on a per model basis) and the kernel code itself by 
-performing optimizations like tree walk unrolling and interleaving. In contrast, Tahoe 
-uses a library-based approach, and cannot generate code tailored to a model 
-like \Treebeard{} does.
+performing optimizations like tree walk unrolling and interleaving.
+% In contrast, Tahoe 
+% uses a library-based approach, and cannot generate code tailored to a model 
+% like \Treebeard{} does.
 
@@
 and in-memory representation for all models. XGBoost \cite{XGBoost} also implements GPU 
-support\cite{XGBGPU}. Again, XGBoost uses a single strategy and in-memory representation
-for all models. In contrast, \Treebeard{} is a compiler that can explore a much larger
-optimization space and can generate code that is tailored to a specific model.
+support\cite{XGBGPU} but uses a single strategy and in-memory representation. 
+% In contrast, \Treebeard{} is a compiler that can explore a much larger
+% optimization space and can generate code that is tailored to a specific model.
 
@@
 % This can be the subject of a separate future work.
+On CPUs, XGBoost\cite{XGBoost}, LightGBM\cite{LightGBM} and
+scikit-learn\cite{Sklearn} are extremely popular. However, 
+as mentioned in Section \ref{sec:intro}, none of these systems
+provide portable performance across different target machines.
+\TODO{Write about the PACT paper and whether our scheduling language can represent 
+all the schedules they propose.}
+Other systems that hide dependency stalls by interleaving tree walks\cite{VPred},
+implement optimized algorithms for tree inference\cite{QuickScorer, QuickScorer1}
+and improve cache performance of decision tree ensembles on CPUs\cite{CacheConscious1, CacheConscious2}
+have been proposed in prior work. However, these systems are limited to CPUs.
+Some systems have been proposed to parallelize decision tree training 
+on CPUs and GPUs\cite{Jansson2014gpuRFAG, Nasridinov2013DecisionTC}.
 
@@
 performance on CPUs. 
-Treelite\cite{Treelite} is a model compiler that only supports 
-generation of simple \op{if-else} style code. Its code generator is hard-coded 
-to expand trees in an ensemble into a series of \op{if-else} statements (one for 
-each node in a tree). Due to this, extending Treelite and reusing it to 
-target GPUs is not possible.
+Treelite\cite{Treelite} is a model compiler that only  
+generates \op{if-else} code for each tree in the model. 
+% Its code generator is hard-coded 
+% to expand trees in an ensemble into a series of \op{if-else} statements (one for 
+% each node in a tree). Due to this, extending Treelite and reusing it to 
+% target GPUs is not possible.
 
@@
 build on top of \TreebeardOLD{}, \Treebeard{} is a significant enhancement 
-over \TreebeardOLD{}. Most importantly, we re-architect \TreebeardOLD{} in 
-order to generate code for multiple target processors. Specifically, we 
-introduce the scheduling language and schedule exploration 
-while also enhancing the intermediate representations and support for parallelizing 
+over \TreebeardOLD{}. 
+% Most importantly, we re-architect \TreebeardOLD{} in 
+% order to generate code for multiple target processors.
+Specifically, we introduce the scheduling language and schedule exploration 
+while also enhancing the IRs and support for parallelizing 
 across trees through the implementation of a novel MLIR reduction dialect.
@@
 Hummingbird\cite{Hummingbird} is a compiler that compiles traditional ML models
-to tensor operations thereby enabling them to be run on tensor-based frameworks like
-TensorFlow. Hummingbird can target both CPUs
-and GPUs. Its stated aim is to allow traditional ML based applications to easily 
-leverage progress in tensor compilers. However, as was shown earlier~\cite{Treebeard},
+to tensor operations, thereby enabling them to be run on tensor-based frameworks like
+TensorFlow\cite{TensorFlow}. Hummingbird can target both CPUs
+and GPUs, but, as was shown earlier~\cite{Treebeard},
 tensor operations are not the most efficient way to implement decision tree inference
-and the performance of code generated by Hummingbird is significantly lower than 
-the code generated by other frameworks.
+and the performance of Hummingbird is significantly lower than 
+that of other frameworks.
 
@@
 % also does not perform model specific optimizations that are enabled by \Treebeard{}'s abstractions.
-
-\emph{Libraries:} The most popular systems for decision tree-based models 
-are libraries. On CPUs, XGBoost\cite{XGBoost}, LightGBM\cite{LightGBM} and
-scikit-learn\cite{Sklearn} are extremely popular. These libraries implement both 
-training and inference. On GPUs, RAPIDs FIL\cite{FIL} is the most commonly used 
-library. However, as mentioned in Section \ref{sec:intro}, none of these systems
-provide portable performance across different target machines.
-\TODO{Write about the PACT paper and whether our scheduling language can represent 
-all the schedules they propose.}
-Other systems that hide dependency stalls by interleaving tree walks\cite{VPred},
-implement optimized algorithms for tree inference\cite{QuickScorer, QuickScorer1}
-and predict cache performance of decision tree ensembles on CPUs\cite{CacheConscious1, CacheConscious2}
-have been proposed in prior work. However, these systems are limited to CPUs.
-Some systems have been proposed 
-to parallelize decision tree training on CPUs and GPUs\cite{Jansson2014gpuRFAG, Nasridinov2013DecisionTC}.
 
@@
 intermediate language and a virtual machine to enable vector execution of decision tree
-inference. In contrast with \Treebeard{}, the virtual machine that performs the SIMD 
-execution is itself implemented by hand on different target processors. This is clearly 
-more expensive than \Treebeard{}'s approach since the VM
-needs to be reimplemented for every supported target architecture.
+inference. However, this virtual machine is itself implemented by 
+hand on different target processors. This is clearly 
+more expensive than \Treebeard{}'s approach.
 % Additionally, even though they perform layout optimizations, their system does
@@
 Inspector-executor systems \cite{TaoOfParallelism,HybridCPUGPU} have been 
-developed to parallelize tree walks. However, these techniques are not a good fit 
-for decision tree inference. In decision trees, the individual node predicates are very 
+developed to parallelize tree walks but are not a good fit 
+for decision tree inference as the individual node predicates are 
 simple and the overhead of an inspector-executor system would be prohibitive.
@@
 % by {\sc {Cortex}}.
-
 Libraries that compose or generate optimized implementations  
-for BLAS\cite{BLIS, atlas_sc98, CUTLASS} and signal processing\cite{FFTW, SPIRAL}.
+for BLAS\cite{BLIS, atlas_sc98, CUTLASS} and signal processing\cite{FFTW, SPIRAL}
+have also been developed.
 % BLIS\cite{BLIS} and ATLAS\cite{atlas_sc98}

### 2024-04-15 results/python/analyze_schedule_sensitivity_data.py
fix model sensitivity plots + add new plots

@@ -21,11 +21,17 @@ def get_degradation_vals(df_benchmark, X, Y, x_col="cur_batch_size", y_col="cmp_
             df_depth = df_benchmark[(df_benchmark[x_col] == x) & (df_benchmark[y_col] == y)]
             # assert df_depth.shape[0] == 1, f"Expected 1 row for {x}, {y}, got {df_depth.shape[0]}"
             if df_depth.shape[0] == 0:
-                degradation_vals[i, j] = 5
+                degradation_vals[i, j] = -1
                 continue
             degradation_vals[i, j] = df_depth["degradation"].values[0]
             # if x < y:
             #     degradation_vals[i, j] = 10 * degradation_vals[i, j]
+
+    max_val = np.max(degradation_vals)
+    for deg_row in degradation_vals:
+        for i in range(len(deg_row)):
+            if deg_row[i] == -1:
+                deg_row[i] = max_val
     return degradation_vals
 
 def plot_batch_sensitivity_for_single_benchmark(df_depth, benchmark_name):
@@ -62,20 +68,34 @@ def plot_model_sensitivity_for_single_benchmark(df_depth, model_name):
     # get a numpy array for the x, y and z axis
     x = df_depth["benchmark"].unique()
     y = df_depth["cmp_benchmark"].unique()
+
     x.sort()
     y.sort()
     Z = get_degradation_vals(df_depth, x, y, "benchmark", "cmp_benchmark") # df_depth['degradation'].values.reshape(len(y), len(x))
     # print(Z)
 
+    x_positions = [(i+1) * 10 for i in range(len(x))]
+    y_positions = [(i+1) * 10 for i in range(len(y))]
+    
     da = xr.DataArray(
         Z,
-        coords={"Model": x, "Comparison Model": y},
+        coords={"Model": x_positions, "Comparison Model": y_positions},
         dims=("Model", "Comparison Model"),
     )
+    da.plot(aspect=1.5, size=10)
+    plt.axis('off')
 
-    # da.plot(xscale="log", yscale="log")
+    for name, x_pos in zip(x, x_positions):
+        name = name if name != "year_prediction_msd" else "year" 
+        plt.text(x_pos, -1.5, name, rotation=90, fontsize=10, wrap=True)
+    
+    for name, y_pos in zip(y, y_positions):
+        name = name if name != "year_prediction_msd" else "year"
[…]

### 2024-04-15 docs/SOSP 2024/intro.tex
Update "widely used" part of intro

@@
 Decision forest models remain the mainstay for machine learning over tabular data~\cite{DLNotAllYouNeed,TreebasedOutperformDL}. 
-Their robustness, interpretability, and ability to handle missing data make them a popular choice for a wide range of applications. 
-\TODO{a few more lines, classification, regression, etc.}
-%We observe that while scalable performance on neural network models has received significant attention, decision tree inference has not seen the same level of acceleration.
-A recent survey~\cite{kaggle}\dots. 
-The survey also finds that, the cost of inference is the most critical factor in the overall cost of deploying a machine learning model.
+Their robustness, interpretability, and ability to handle missing data make them a popular choice for a wide 
+range of applications\cite{DecisionTreesOverview, Med1, Med2, Facebook, LHCModel, Finance}. 
+% \TODO{a few more lines, classification, regression, etc.}
+% We observe that while scalable performance on neural network models has received significant attention, decision tree inference has not seen the same level of acceleration.
+A recent survey~\cite{KaggleSurvey} found that about three quarters of data scientists use decision tree
+based models. An analysis of ML workloads at a large scale web company found that these models are most widely used~\cite{LookingGlass}. 
+Recent work has noted that the cost of inference is the most critical factor in the overall cost of 
+deploying a machine learning model~\cite{Hummingbird, SageMaker}.
 This is because, in production settings, each model is trained once and often used for inference millions of times. 

### 2024-04-15 docs/SOSP 2024/motivation.tex
Motivation section draft

@@
 \section{Motivation}
+
+In this section, we first motivate the need for a scheduling language 
+by showing how a model can be compiled in different ways using the example
+in Figure \ref{Fig:HIRExample} and subsequently, we show how drastically 
+performance can vary across these variants for real benchmarks.
+
+First, we consider the schedule that processes one tree at a time 
+for all input rows and unrolls all tree walks. 
+The schedule splits the loop over the trees into two loops -- one that
+iterates over the first two trees (Trees 1 and 2 with depth 1) and 
+the second that iterates over the last two trees (Trees 3 and 4 with
+depth 2). The schedule then unrolls the tree walks for each tree.
+\begin{lstlisting}[style=c++]
+  reorder(tree, batch)  
+  split(tree, t0, t1, 2)
+  unrollWalk(t0, 1)
+  unrollWalk(t1, 2)
+\end{lstlisting}
+The concrete implementation of this schedule (in one of \Treebeard{}'s IRs) 
+is as follows.
+\begin{lstlisting}[style=c++]
+  model = ensemble(...)
+  for t0 = 0 to 2 step 1 {
+    T = getTree(ensemble, t0)
+    for batch = 0 to BATCH_SIZE step 1 {
+      treePred = walkDecisionTree(T, 
+                    input[batch]) <unrollDepth = 1>
+      reduce(result[batch], treePred)
+    }
+  }
+  for t1 = 2 to 4 step 1 {
+    T = getTree(ensemble, t0)
+    for batch = 0 to BATCH_SIZE step 1 {
+      treePred = walkDecisionTree(T,
+                    input[batch]) <unrollDepth = 2>
+      reduce(result[batch], treePred) <'+', 0.0>
+    }
+  }
+\end{lstlisting}
+This schedule is ideally suited for a single-core CPU. It maximizes 
+the reuse of trees in the L1 cache and also minimizes the amount of
+branching by unrolling tree walks. However, it doesn't exploit  
+any parallelism and is therefore ill-suited for parallel processors.
+
+One form of parallelism that can be exploited is to process rows in 
+parallel. However, with massively parallel processors like GPUs,
+this may not yield sufficient parallel work. Another option is to also 
+parallelize across trees. A possible schedule to accomplish this is as follows.
+\begin{lstlisting}[style=c++]
+  tile(tree, t0, t1, 2)
+  reorder(batch, t1, t0)
+  split(t0, t0_1, t0_2, 2)
+  unrollWalk(t0_1, 1)
+  unrollWalk(t0_2, 2)
+  gpuDimension(batch, grid.x)
+  gpuDimension(t1, block.x)
+\end{lstlisting}
+This schedule generates an inference function that runs on the GPU. 
+The inference routine processes one input row per thread block (since the \op{batch}
+loop is mapped directly to \op{grid.x}).
+It also splits the trees into two sets by tiling the \op{tree} loop.
+Each of the two sets is processed in parallel. We unroll the tree walks 
+for each tree. The IR generated is as follows. 
+\begin{lstlisting}[style=c++]
+  model = ensemble(...)
+  par.for batch = 0 to BATCH_SIZE step 1 <grid.x> {
+    par.for t1 = 0 to 2 step 1 <block.x> {
+      for t0_1 = 0 to 2 step 2 {
+        T = getTree(ensemble, t0_1 + t1)
+        treePred = walkDecisionTree(T, 
+                        input[batch]) <unrollDepth = 1>
+        reduce(result[batch], treePred)
+      }
+      for t0_2 = 2 to 4 step 2 {
+        T = getTree(ensemble, t0_2 + t1)
+        treePred = walkDecisionTree(T,
+                        input[batch]) <unrollDepth = 2>
+        reduce(result[batch], treePred) <'+', 0.0>
+      }
+    }
+  }
+\end{lstlisting}
+
+In the case of this schedule, the \op{reduce}
+operation needs special consideration. In order to correctly generate 
+code for this schedule, the compiler needs to determine that parallel 
+iterations of the \op{t1} 
+loop accumulate into the same element of the \op{result} array.
+One possible solution is to rewrite the reduction so that each parallel 
+iteration accumulates into a different array element by introducing 
+a temporary buffer (\op{tempResults}) as follows.
+\begin{lstlisting}[style=c++]
+  float tempResults[2][BATCH_SIZE]
+  model = ensemble(...)
+  par.for batch = 0 to BATCH_SIZE step 1 <grid.x> {
+    par.for t1 = 0 to 2 step 1 <block.x> {
+      for t0_1 = 0 to 2 step 2 {
+        T = getTree(ensemble, t0_1 + t1)
+        treePred = walkDecisionTree(T, 
+                      input[batch]) <unrollDepth = 1>
+        reduce(tempResults[t1][batch], treePred)
+      }
+      for t0_2 = 2 to 4 step 2 {
+        T = getTree(ensemble, t0_2 + t1)
+        treePred = walkDecisionTree(T,
+                      input[batch]) <unrollDepth = 2>
+        reduce(tempResults[t1][batch], treePred) <'+', 0.0>
+      }
+    }
+    result[batch] = reduce_dimension(tempResults[:][batch], 0)
+  }
+\end{lstlisting}
+Here, partial results are accumulated into \op{tempResults} and then
+reduced across the \op{t1} dimension (represented by the \op{reduce\_dimension}
+operation) to get the final result.
+
+As is evident from these examples, it is possible to optimize 
+the inference routine in different ways. Also, the structure of the loop 
+nest in the inference routine can get quite complex even 
+for simple schedules. Writing these routines by hand 
+is error-prone and time-consuming.  We believe that designing a 
+scheduling language to encapsulate these strategies and a principled code 
+generator to automatically generate code based on the schedule is the 
+best approach.
+
+We find that these different strategies can have significantly different
+performance. Figure \ref{fig:sensitivity} shows the\dots.
+
+Several problems need to be solved in order to build the kind of 
+schedule guided retargetable compiler required. We had to enable the
+compiler to represent and optimize reductions, deal uniformly with different
+in-memory representations of the model, design optimizations to effectively 
+use the memory hierarchy of the target processor (shared memory on GPUs and 
+cache on the CPU) and finally be able to generate target specific code. 
+The rest of the paper describes these challenges in detail and how we
+solved them in \Treebeard{}.
 
@@
 
-While a diverse set techniques have been proposed for optimization of decision tree 
-inference on CPUs and GPUs \cite{VPred, Tahoe, Treelite, XGBoost, Hummingbird, QuickScorer, FIL},
-a very extensive design space of optimizations exists 
-outside what has been proposed in the literature. Furthermore, decision tree inference 
-is run on several platforms including CPUs and GPUs. The implementations used on each of 
-these platforms are different and the techniques used to optimize them are different.
-To make matters even more complicated, several in-memory representations
-have been proposed for decision tree models. For example, XGBoost\cite{XGBoost} uses a sparse representation,
-RAPIDs FIL\cite{FIL} uses what is called the reorg representation and Tahoe uses a variation of the reorg
-representation. 
-\TODO{Can we add some numbers here to show that different models/batch sizes need different optimizations?}
+% While a diverse set techniques have been proposed for optimization of decision tree 
+% inference on CPUs and GPUs \cite{VPred, Tahoe, Treelite, XGBoost, Hummingbird, QuickScorer, FIL},
+% a very extensive design space of optimizations exists 
+% outside what has been proposed in the literature. Furthermore, decision tree inference 
+% is run on several platforms including CPUs and GPUs. The implementations used on each of 
+% these platforms are different and the techniques used to optimize them are different.
+% To make matters even more complicated, several in-memory representations
+% have been proposed for decision tree models. For example, XGBoost\cite{XGBoost} uses a sparse representation,
+% RAPIDs FIL\cite{FIL} uses what is called the reorg representation and Tahoe uses a variation of the reorg
+% representation. 
+% \TODO{Can we add some numbers here to show that different models/batch sizes need different optimizations?}
 
-To solve the problems of exploring the design space of optimizations for decision tree
-inference and enabling portable performance, we build several techniques in \Treebeard{}, 
-an open source compiler infrastructure for decision tree inference. To make \Treebeard{}
-capable of unifying these different techniques and targets, we do the following. 
-\begin{itemize}
-  \item We design a scheduling language that encapsulates various optimization techniques
-  and controls the structure of the generated code.
-  \item We design an MLIR dialect to represent and optimize reductions and use this 
-  dialect within \Treebeard{} to enable the generation of different variants of 
-  inference routines.
-  \item We extend \Treebeard{}'s intermediate representations to include operations like caching.
-  We were able to easily reuse and extend \Treebeard{}'s IR as it was built as an MLIR dialect.
-  \item We design a plugin mechanism with which different in-memory representations
-  can be composed with different optimizations.  
-\end{itemize}
+% To solve the problems of exploring the design space of optimizations for decision tree
+% inference and enabling portable performance, we build several techniques in \Treebeard{}, 
+% an open source compiler infrastructure for decision tree inference. To make \Treebeard{}
+% capable of unifying these different techniques and targets, we do the following. 
+% \begin{itemize}
+%   \item We design a scheduling language that encapsulates various optimization techniques
+%   and controls the structure of the generated code.
+%   \item We design an MLIR dialect to represent and optimize reductions and use this 
+%   dialect within \Treebeard{} to enable the generation of different variants of 
+%   inference routines.
+%   \item We extend \Treebeard{}'s intermediate representations to include operations like caching.
+%   We were able to easily reuse and extend \Treebeard{}'s IR as it was built as an MLIR dialect.
+%   \item We design a plugin mechanism with which different in-memory representations
+%   can be composed with different optimizations.  
+% \end{itemize}

### 2024-04-16 docs/SOSP 2024/motivation.tex
Edit motivation section

@@
 In this section, we first motivate the need for a scheduling language 
-by showing how a model can be compiled in different ways using the example
-in Figure \ref{Fig:HIRExample} and subsequently, we show how drastically 
+by showing how a model can be compiled in different ways
+and subsequently, we show how drastically 
 performance can vary across these variants for real benchmarks.
 
-First, we consider the schedule that processes one tree at a time 
+As an example, consider a model with four trees, two complete
+trees of depth 1 and two of depth 2 (Figure \ref{Fig:HIRExample}).
+% We use this model as our running example throughout this section.
+We first describe a simple strategy that processes one tree at a time 
 for all input rows and unrolls all tree walks. 
-The schedule splits the loop over the trees into two loops -- one that
+The loop over the trees is split into two loops -- one that
 iterates over the first two trees (Trees 1 and 2 with depth 1) and 
 the second that iterates over the last two trees (Trees 3 and 4 with
-depth 2). The schedule then unrolls the tree walks for each tree.
+depth 2). The \Treebeard{} schedule then unrolls the tree walks for each tree.
 \begin{lstlisting}[style=c++]
   reorder(tree, batch)
-  split(tree, t0, t1, 2)
-  unrollWalk(t0, 1)
-  unrollWalk(t1, 2)
+  // Fiss tree loop so trees with equal depth 
+  // are processed together  
+  split(tree, t_depth1, t_depth2, 2)
+  // Unroll the tree walks
+  unrollWalk(t_depth1, 1)
+  unrollWalk(t_depth2, 2)
 \end{lstlisting}
@@
   model = ensemble(...)
-  for t0 = 0 to 2 step 1 {
-    T = getTree(ensemble, t0)
+  for t_depth1 = 0 to 2 step 1 {
+    T = getTree(ensemble, t_depth1)
     for batch = 0 to BATCH_SIZE step 1 {
@@
   }
-  for t1 = 2 to 4 step 1 {
-    T = getTree(ensemble, t0)
+  for t_depth2 = 2 to 4 step 1 {
+    T = getTree(ensemble, t_depth2)
     for batch = 0 to BATCH_SIZE step 1 {
@@
 this may not yield sufficient parallel work. Another option is to also 
-parallelize across trees. A possible schedule to accomplish this is as follows.
+parallelize across trees. A possible strategy to accomplish this is 
+encoded in the following schedule.
 \begin{lstlisting}[style=c++]
+  // Split the trees into two sets
   tile(tree, t0, t1, 2)
   reorder(batch, t1, t0)
-  split(t0, t0_1, t0_2, 2)
-  unrollWalk(t0_1, 1)
-  unrollWalk(t0_2, 2)
+  // Fiss loop so that trees with equal 
+  // depth are processed together
+  split(t0, t0_depth1, t0_depth2, 2)
+  unrollWalk(t0_depth1, 1)
+  unrollWalk(t0_depth2, 2)
+  // Configure the GPU kernel dimensions
   gpuDimension(batch, grid.x)
@@
     par.for t1 = 0 to 2 step 1 <block.x> {
-      for t0_1 = 0 to 2 step 2 {
-        T = getTree(ensemble, t0_1 + t1)
+      for t0_depth1 = 0 to 2 step 2 {
+        T = getTree(ensemble, t0_depth1 + t1)
         treePred = walkDecisionTree(T, 
@@
       }
-      for t0_2 = 2 to 4 step 2 {
-        T = getTree(ensemble, t0_2 + t1)
+      for t0_depth2 = 2 to 4 step 2 {
+        T = getTree(ensemble, t0_depth2 + t1)
         treePred = walkDecisionTree(T,
@@
 iteration accumulates into a different array element by introducing 
-a temporary buffer (\op{tempResults}) as follows.
+a temporary buffer (\op{temp}) as follows.
 \begin{lstlisting}[style=c++]
-  float tempResults[2][BATCH_SIZE]
+  float temp[2][BATCH_SIZE]
   model = ensemble(...)
@@
     par.for t1 = 0 to 2 step 1 <block.x> {
-      for t0_1 = 0 to 2 step 2 {
-        T = getTree(ensemble, t0_1 + t1)
+      for t0_depth1 = 0 to 2 step 2 {
+        T = getTree(ensemble, t0_depth1 + t1)
         treePred = walkDecisionTree(T, 
                       input[batch]) <unrollDepth = 1>
-        reduce(tempResults[t1][batch], treePred)
+        reduce(temp[t1][batch], treePred)
       }
-      for t0_2 = 2 to 4 step 2 {
-        T = getTree(ensemble, t0_2 + t1)
+      for t0_depth2 = 2 to 4 step 2 {
+        T = getTree(ensemble, t0_depth2 + t1)
         treePred = walkDecisionTree(T,
                       input[batch]) <unrollDepth = 2>
-        reduce(tempResults[t1][batch], treePred) <'+', 0.0>
+        reduce(temp[t1][batch], treePred) <'+', 0.0>
       }
     }
-    result[batch] = reduce_dimension(tempResults[:][batch], 0)
+    result[batch] = reduce_dimension(temp[:][batch], 0)
   }
 \end{lstlisting}
-Here, partial results are accumulated into \op{tempResults} and then
-reduced across the \op{t1} dimension (represented by the \op{reduce\_dimension}
-operation) to get the final result.
+Here, partial results are accumulated into \op{temp} and then
+reduced across the \op{t1} dimension to get the final result.
 
@@
 
-We find that these different strategies can have significantly different
-performance. Figure \ref{fig:sensitivity} shows the\dots.
+\begin{figure}[]
+  \centering
+  \begin{subfigure}[b]{.3\textwidth}
+    \subcaptionbox*{}{\includegraphics[width=\textwidth]{figures/batch_sensitivity_covtype.png}}
+    \caption{\label{fig:sensitivitya} Batch sensitivity for \op{covtype}. Each point shows the slowdown when best schedule for the 
+    x-axis batch size is used for the y-axis batch size.}
+  \end{subfigure}
+  % \hfill  % Add horizontal space between subfigures
+  \begin{subfigure}[b]{.3\textwidth}
+    \subcaptionbox*{}{\includegraphics[width=\textwidth]{figures/model_sensitivity_4096.png}}
+    \caption{\label{fig:sensitivityb} Model sensitivity for batch size 4096. Each point shows the slowdown when best schedule for the
+    x-axis model is used for the y-axis model.}
+  \end{subfigure}
+\end{figure}
 
-Several problems need to be solved in order to build the kind of 
-schedule guided retargetable compiler required. We had to enable the
+To further complicate matters, we find that these different strategies can have significantly different
+performance. Figures \ref{fig:sensitivitya} and \ref{fig:sensitivityb} show the
+variation in performance when the same schedules are used across batch sizes 
+and models respectively. These diagrams show that the best schedules to use 
+vary across batch sizes and models. The largest slowdown is 5$\times$ when 
+schedules are used across different batch sizes and 6$\times$ when schedules
+are used across different models. Therefore, no one strategy can be used 
+across all models and batch sizes. 
+A system that is capable of specializing generated code for both batch 
+size and model is required for the best performance.
+
+Building such a configurable compiler and supporting code generation for CPUs and GPUs 
+required us to solve several fundamental problems. We had to enable the
 compiler to represent and optimize reductions, deal uniformly with different

### 2024-04-16 docs/SOSP 2024/schedule.tex
Edit scheduling lang section

@@
 directives.
+
+\Treebeard{} represents loops (index variables) as nodes in a tree where the children
+of a node represent immediately contained loops. Each schedule 
+primitive modifies this tree in some way. The compiler tracks 
+the lineage of each of the loops. This allows the compiler to 
+automatically infer the ranges for all loops.
 

### 2024-04-16 results/python/analyze_gpu_speedups_results.py
Add speedup graphs, script, results as csv

@@ -1,0 +1,173 @@
+import matplotlib.pyplot as plt
+import numpy as np
+import pandas as pd
+
+from scipy.stats import gmean
+
+# Read the CSV file into a dataframe
+results_4060_df = pd.read_csv('../4060_thedeep_results.csv')
+results_t400_df = pd.read_csv('../T400_holmes_results.csv')
+
+PLOT_FILE_EXTENSION = 'png'
+
+def plot_bar_graph_of_abs_times_for_batch_size(df, batch_size):
+    filtered_df = df[df['Batch size'] == batch_size]
+    bar_names = filtered_df['Benchmark'].to_list()
+    tb_kernel_time = filtered_df['TBKernel(AT)'].to_list()
+    rapids_kernel_time = filtered_df['RAPIDS(kernel)'].to_list()
+    tb_total_time = filtered_df['TB (AT)'].to_list()
+    rapids_total_time = filtered_df['RAPIDS (Total)'].to_list()
+
+    # Calculate the difference between kernel time and total time
+    tb_diff = [tb_total - tb_kernel for tb_total, tb_kernel in zip(tb_total_time, tb_kernel_time)]
+    rapids_diff = [rapids_total - rapids_kernel for rapids_total, rapids_kernel in zip(rapids_total_time, rapids_kernel_time)]
+
+    # Normalize w.r.t rapids_total_time
+    norm_tb_kernel_time = [time/ rapids_total for time, rapids_total in zip(tb_kernel_time, rapids_total_time)]
+    norm_rapids_kernel_time = [time/ rapids_total for time, rapids_total in zip(rapids_kernel_time, rapids_total_time)]
+    norm_tb_diff = [time/ rapids_total for time, rapids_total in zip(tb_diff, rapids_total_time)]
+    norm_rapids_diff = [time/ rapids_total for time, rapids_total in zip(rapids_diff, rapids_total_time)]
+    norm_tb_total_time = [time/ rapids_total for time, rapids_total in zip(tb_total_time, rapids_total_time)]
+    norm_rapids_total_time = [time/ rapids_total for time, rapids_total in zip(rapids_total_time, rapids_total_time)]
+
+    # Prepare data for plotting
+    width = 0.35
+    pos = np.arange(len(bar_names))
+    _, ax = plt.subplots()
[…]

### 2024-04-16 docs/SOSP 2024/hir-mir-opts.tex
Edit schedule and IR description sections. Remove Overview

@@
-\section{HIR and MIR Optimizations}
+\section{\Treebeard{} Multi-Level Compiler}
 
@@
 
-The \TreebeardOLD{} infrastructure was originally designed to target CPUs \cite{Treebeard}. 
-However, it implements several optimizations on the HIR and MIR that 
-can be leveraged across target processors and we find that some these 
-are beneficial for GPUs as well. This reuse of optimizations is possible 
-because the intermediate representations on which these optimizations are performed 
-are abstract and are designed to be target-independent. We briefly review these 
-optimizations below.
+\Treebeard{} takes a serialized decision tree ensemble as input (E.g.:
+XGBoost JSON, ONNX etc.) and automatically generates an optimized inference function
+that can either target CPUs or GPUs. 
+Figure \ref{Fig:CompilerStructure} shows the structure of the \Treebeard{} compiler. 
+The inference computation is lowered through three intermediate representations
+-- high-level IR (HIR), mid-level IR (MIR) and low-level IR (LIR). The LIR is
+finally lowered to LLVM and then JIT'ed to the specified target processor. 
+\Treebeard{} is built using the open-source \TreebeardOLD{} infrastructure~\cite{Treebeard}.
+Since the \TreebeardOLD{} infrastructure was originally designed to target CPUs,
+significant extensions  (shown using dark boxes in Figure \ref{Fig:CompilerStructure})
+were required to support schedule guided compilation across CPUs and GPUs. 
+% However, it implements several optimizations on the HIR and MIR that 
+% can be leveraged across target processors and we find that some these 
+% are beneficial for GPUs as well. 
+% This reuse of optimizations is possible 
+% because the intermediate representations on which these optimizations are performed 
+% are abstract and are designed to be target-independent. We briefly review these 
+% optimizations below.
 
-\subsection{Optimizations on High-Level IR}
-We augment the existing \TreebeardOLD{} infrastructure with loop rewrites on 
+\begin{figure}[htb]
+  \centering
+  \includegraphics[width=\linewidth]{figures/compiler.png}
+  \caption{\Treebeard{} compiler structure.}
+  \label{Fig:CompilerStructure}
+\end{figure}
+
+Table \ref{Tab:IRSpecification} lists the operations in the three IRs. 
+In HIR, the model is represented as a collection of binary trees. This abstraction
+allows the implementation of optimizations that require the manipulation of the model
+or its constituent trees. We augment the existing \TreebeardOLD{} infrastructure with loop rewrites on 
 the HIR that are implemented through the scheduling language (Section \ref{sec:schedule}).
@@
 Additionally, the \TreebeardOLD{} infrastructure implements HIR transformations to reorder and pad 
-trees. It also implements tree tiling transformations on the HIR \cite{Treebeard}. We reuse the 
-reordering and padding transformations on the HIR for GPUs. However, we found that 
-tiling trees was not beneficial for GPUs. This is because the tiling transformation
-introduces redundant computation inorder to vectorize computation on CPUs where 
-all lanes need to follow the same control flow. However, on SIMT GPUs, we find that 
-the benefits of tiling (coalescing memory accesses) do not outweigh the cost of
-redundant computation. We leave an investigation of this for future work.
+trees. 
+% Figure \ref{Fig:HIRExample} shows this representation for a model 
+% with four trees and how the model can be transformed. The model contains two trees of 
+% depth 1 and two trees of depth 2 (\circled{A}). The right side of the diagram shows the models after 
+% padding and reordering (\circled{B}). Padding makes all leaves in a tree the same depth (Trees 2 and
+% 4 are padded). Trees are reordered so that trees of the same depth are together (Trees 
+% 2 and 3 are swapped). We use this model as a running example for the rest of the section.
+% After these model-level optimizations are performed on the HIR, the 
+% code is lowered to the mid-level IR (MIR) as dictated by a  
+% \emph{\textbf{schedule}}. The schedule determines how to traverse
+% the iteration space that goes over the trees and input rows 
+% by specifying how loops are to be tiled, 
+% parallelized, mapped to GPU grid and block dimensions etc.
+% (Details in Section \ref{sec:schedule}). While MIR is a loop-based IR that 
+% explicitly encodes details of the iteration space has to be traversed, 
+% it still abstracts details about the in-memory representation of the model.
 
-\subsection{Optimizations on Mid-Level IR}
+In the lowered MIR, the compiler uses the \op{reduce} op from 
+the \op{reduction} dialect we design (details in Section \ref{sec:reduction})
+to represent reduction operations. The lowering of the \op{reduce} operation 
+involves introducing temporary buffers and splitting the reduction 
+to correctly implement reduction in the presence of 
+parallel loops. This process, that we call \textbf{\emph{legalization}}, is 
+described in Section \ref{sec:reduction}.
+
 The original \TreebeardOLD{} infrastructure implements optimizations like 
@@
 unrolling and tree-walk interleaving optimizations on the MIR for GPUs.
-
-While building \Treebeard{}, we found that one of the performance bottlenecks on the 
+For example, while building \Treebeard{}, we found that one of the performance bottlenecks on the 
 GPU was that warps spent significant time being stalled. Since GPUs 
@@
 performance on the GPU.  
-\TODO{Should we talk about how interleaving is implemented as a 
-statemachine and therefore it can be used across representations 
-and tile traversal techniques?}
+% \TODO{Should we talk about how interleaving is implemented as a 
+% statemachine and therefore it can be used across representations 
+% and tile traversal techniques?}
 
-\subsection{A Note on Low-level IR}
+The MIR is then further lowered to a low-level IR (LIR). 
 Significant changes to the original \TreebeardOLD{} design were required to get
@@
 is implemented differently etc.).
+
+This is the
+level at which the compilation pipeline diverges for different targets. In 
+the GPU compilation pipeline, the required memory transfers and kernel
+invocations are inserted into the LIR. Additionally, buffers 
+to hold model values are inserted and abstract tree operations are lowered to
+explicitly refer to these buffers. This lowering is controlled by 
+a plugin mechanism which enables different in-memory representations to 
+be added to the compiler by implementing an interface 
+(Section \ref{sec:representations}). Vectorization of tree traversals
+is also explicitly represented in LIR.
+
+% \subsection{Optimizations on High-Level IR}
+% We augment the existing \TreebeardOLD{} infrastructure with loop rewrites on 
+% the HIR that are implemented through the scheduling language (Section \ref{sec:schedule}).
+% We use these to implement the automatic scheduling described in Section \ref{sec:exploring}. 
+% Additionally, the \TreebeardOLD{} infrastructure implements HIR transformations to reorder and pad 
+% trees. It also implements tree tiling transformations on the HIR \cite{Treebeard}. We reuse the 
+% reordering and padding transformations on the HIR for GPUs. However, we found that 
+% tiling trees was not beneficial for GPUs. This is because the tiling transformation
+% introduces redundant computation inorder to vectorize computation on CPUs where 
+% all lanes need to follow the same control flow. However, on SIMT GPUs, we find that 
+% the benefits of tiling (coalescing memory accesses) do not outweigh the cost of
+% redundant computation. We leave an investigation of this for future work.
+
+% \subsection{Optimizations on Mid-Level IR}
+% The original \TreebeardOLD{} infrastructure implements optimizations like 
+% tree-walk unrolling, tree-walk interleaving, and parallelization
+% on the MIR. These optimizations are beneficial for GPUs as well.
+% and the design of \TreebeardOLD{} allows us to reuse the tree-walk 
+% unrolling and tree-walk interleaving optimizations on the MIR for GPUs.
+
+% While building \Treebeard{}, we found that one of the performance bottlenecks on the 
+% GPU was that warps spent significant time being stalled. Since GPUs 
+% implement scoreboarding~\cite{HennesseyPatterson}, we were able to alleviate this bottleneck by
+% interleaving tree walks. This significantly improved performance of generated 
+% code. We found it surprising that the use of ILP could benefit 
+% performance on the GPU.  
+% \TODO{Should we talk about how interleaving is implemented as a 
+% statemachine and therefore it can be used across representations 
+% and tile traversal techniques?}
+
+% \subsection{A Note on Low-level IR}
+% Significant changes to the original \TreebeardOLD{} design were required to get
+% LIR to correctly lower to GPU code. The most important of these was 
+% the change to how the compiler implements support for in-memory 
+% representations of models (Section \ref{sec:representations}).
+% With these design changes, we were able to reuse much of the CPU 
+% implementation while customizing some parts for GPUs (for example,
+% buffers need to be allocated differently for CPU and GPU, caching 
+% is implemented differently etc.).

### 2024-04-16 docs/SOSP 2024/motivation.tex
Edit schedule and IR description sections. Remove Overview

@@
 \section{Motivation}
-
+\label{sec:motivation}
 In this section, we first motivate the need for a scheduling language 
@@
 performance can vary across these variants for real benchmarks.
+
+\begin{figure}[htb]
+  \centering
+  \includegraphics[width=\linewidth]{figures/HIR.PNG}
+  \caption{The representation of a model in high-level IR and the model 
+  after trees are padded and reordered. }
+  \label{Fig:HIRExample}
+\end{figure}
 

### 2024-04-16 docs/SOSP 2024/schedule.tex
Edit schedule and IR description sections. Remove Overview

@@
 \label{sec:schedule}
-As we articulated in Section \ref{sec:intro}, there are several 
+As established in Section \ref{sec:motivation}, there are several 
 different configurations and optimizations strategies for decision 
-tree inference. The best ones are significantly different across
+tree inference. The best ones vary significantly across
 models, batch sizes and hardware platforms. Therefore, designing
-any one hard-coded lowering strategy is not feasible as this would 
-make portable performance impossible. To address this problem, we 
+any one hard-coded strategy is not feasible as this would 
+make portable performance impossible. To address this, we 
 design a scheduling language for \Treebeard{}. The scheduling 
 language provides an abstract way to specify loop structure and 
-other optimizations as an input to the compiler. Making the 
-scheduling specification external to the compiler allows to build 
-auto-schedulers and auto-tuners (Section \ref{sec:exploring}).
-Using a scheduling language will also significantly simplify 
-adding support for new hardware as this will likely require
-different locality optimizations and loop transformations.
+other optimizations as an input to the compiler. The specified 
+schedule controls the lowering of model inference to a set of 
+loop nests. This configurability provided by the schedule allows us 
+to build auto-schedulers and auto-tuners (Section \ref{sec:exploring}).
+% Using a scheduling language will also significantly simplify 
+% adding support for new hardware as this will likely require
+% different locality optimizations and loop transformations.
 
-The goal of \Treebeard{}'s scheduling language is to declaratively
-express loop structures and the application of other optimizations 
-(tree walk unrolling, tree walk interleaving etc.).
+% The goal of \Treebeard{}'s scheduling language is to declaratively
+% express loop structures and the application of other optimizations 
+% (tree walk unrolling, tree walk interleaving etc.).
 
-\subsection{Language Definition}
+% \subsection{Language Definition}
 The core construct of \Treebeard{}'s scheduling language is an 
@@
 
-\Treebeard{} represents loops (index variables) as nodes in a tree where the children
-of a node represent immediately contained loops. Each schedule 
-primitive modifies this tree in some way. The compiler tracks 
-the lineage of each of the loops. This allows the compiler to 
-automatically infer the ranges for all loops.
-
 \Treebeard{}'s scheduling language has three classes of directives. The first is a set 
@@
 (Table \ref{Tab:ReductionOpts}).
+
+The set of loops (index variables) are internally represented as nodes in a tree where the children
+of a node represent immediately contained loops. Each schedule 
+primitive modifies this tree in some way. The compiler tracks 
+the lineage of each of the loops. This allows the compiler to 
+automatically infer the ranges for all loops.
 
@@
                                                &   &  \\
-   \midrule
-   \multirow{4}{*}{\texttt{specialize}} & \textbf{indexVar}  & \multirow{4}{*}{\parbox{0.55\linewidth}{Generate specialized code for each iteration of the loop. 
-   Useful when different iterations of a loop need to execute different code.}} \\
-                                               &   & \\
-                                               &   &  \\
-                                               &   &  \\
+  %  \midrule
+  %  \multirow{4}{*}{\texttt{specialize}} & \textbf{indexVar}  & \multirow{4}{*}{\parbox{0.55\linewidth}{Generate specialized code for each iteration of the loop. 
+  %  Useful when different iterations of a loop need to execute different code.}} \\
+  %                                              &   & \\
+  %                                              &   &  \\
+  %                                              &   &  \\
    \midrule
@@
 
-We now show how the implementation strategies of several existing systems can 
-be encoded in \Treebeard{}'s scheduling language. First, we note that the 
+\Treebeard{}'s scheduling language is expressive enough to represent a wide range of
+strategies used in existing systems. We show examples of how it can be used to 
+represent XGBoost and Tahoe's strategies. Before presenting the examples, we note that \Treebeard{}'s 
 default loop order is [\op{batch}, \op{tree}], i.e, for each row in the input
@@
 have been walked for this set of rows, the next set of rows is 
-taken up. Also, different sets of rows are processed in parallel.
-This schedule can be implemented in \Treebeard{}'s scheduling language 
-as follows.
+taken up. Different sets of rows are processed in parallel.
+The following schedule expresses XGBoost's strategy.
 \begin{lstlisting}[style=c++]
@@
 We show how two of these strategies can be encoded using \Treebeard{}'s scheduling language.
-The rest can be encoded similarly. 
+%The rest can be encoded similarly. 
 \begin{itemize}
-  \item \textbf{Direct Method}: In this strategy, a single GPU thread walks all trees
+  \item In the \emph{direct method}, a single GPU thread walks all trees
   for a given input row. The schedule for this strategy is as follows.
@@
   Here, \op{ROWS\_PER\_TB} is the number of rows that are processed by a single thread block.
-  \item \textbf{Shared Data}: In this strategy, a thread block walks all the trees 
-  for a given row in parallel. If threads walk multiple trees, each thread accumulates
-  partial results. Finally, a thread block wide reduction is performed to compute 
+  \item In the \emph{shared data} strategy, a thread block walks all the trees 
+  for a given row in parallel. 
+  % If threads walk multiple trees, each thread accumulates partial results. 
+  Then, a thread block wide reduction is performed to compute 
   the prediction. The schedule for this strategy is as follows.
@@
 
-There are some simplifying assumptions and limitations in the current design
-of the scheduling language. Mainly, tree traversals are considered atomic and
-accumulation of tree predictions is done immediately (as opposed to, 
-for example, collecting all predictions and performing a reduction later).
-However, we find that these are not significant limitations in practice as
-the current design is able to express most strategies of interest.
+% There are some simplifying assumptions and limitations in the current design
+% of the scheduling language. Mainly, tree traversals are considered atomic and
+% accumulation of tree predictions is done immediately (as opposed to, 
+% for example, collecting all predictions and performing a reduction later).
+% However, we find that these are not significant limitations in practice as
+% the current design is able to express most strategies of interest.

### 2024-04-16 docs/SOSP 2024/hir-mir-opts.tex
Edit IR description section

@@
 
-\Treebeard{} takes a serialized decision tree ensemble as input (E.g.:
+\Treebeard{} takes a serialized decision tree ensemble as input (
 XGBoost JSON, ONNX etc.) and automatically generates an optimized inference function
@@
 Since the \TreebeardOLD{} infrastructure was originally designed to target CPUs,
-significant extensions  (shown using dark boxes in Figure \ref{Fig:CompilerStructure})
-were required to support schedule guided compilation across CPUs and GPUs. 
+significant extensions  (shown as dark boxes in Figure \ref{Fig:CompilerStructure})
+were required to support schedule guided compilation for CPUs and GPUs. 
 % However, it implements several optimizations on the HIR and MIR that 
@@
 allows the implementation of optimizations that require the manipulation of the model
-or its constituent trees. We augment the existing \TreebeardOLD{} infrastructure with loop rewrites on 
+or its constituent trees. We extend the \TreebeardOLD{} infrastructure with loop rewrites on 
 the HIR that are implemented through the scheduling language (Section \ref{sec:schedule}).
@@
 
+The HIR is lowered to MIR as dictated by the \emph{schedule}.
 In the lowered MIR, the compiler uses the \op{reduce} op from 
@@
 to represent reduction operations. The lowering of the \op{reduce} operation 
-involves introducing temporary buffers and splitting the reduction 
+involves introducing temporary buffers and splitting the operation  
 to correctly implement reduction in the presence of 
@@
 
-The original \TreebeardOLD{} infrastructure implements optimizations like 
+\TreebeardOLD{} implements optimizations like 
 tree-walk unrolling, tree-walk interleaving, and parallelization
@@
 unrolling and tree-walk interleaving optimizations on the MIR for GPUs.
-For example, while building \Treebeard{}, we found that one of the performance bottlenecks on the 
+While building \Treebeard{}, we found that one of the performance bottlenecks on the 
 GPU was that warps spent significant time being stalled. Since GPUs 
@@
 interleaving tree walks. This significantly improved performance of generated 
-code. We found it surprising that the use of ILP could benefit 
-performance on the GPU.  
+code. The fact that exploting ILP improved performance on GPUs was surprising.
+
 % \TODO{Should we talk about how interleaving is implemented as a 
@@
 representations of models (Section \ref{sec:representations}).
-With these design changes, we were able to reuse much of the CPU 
-implementation while customizing some parts for GPUs (for example,
-buffers need to be allocated differently for CPU and GPU, caching 
-is implemented differently etc.).
-
-This is the
+% With these design changes, we were able to reuse much of the CPU 
+% implementation while customizing some parts for GPUs (for example,
+% buffers need to be allocated differently for CPU and GPU, caching 
+% is implemented differently etc.). 
+Also, this is the
 level at which the compilation pipeline diverges for different targets. In 
@@
 to hold model values are inserted and abstract tree operations are lowered to
-explicitly refer to these buffers. This lowering is controlled by 
-a plugin mechanism which enables different in-memory representations to 
-be added to the compiler by implementing an interface 
-(Section \ref{sec:representations}). Vectorization of tree traversals
-is also explicitly represented in LIR.
-
+explicitly refer to these buffers.
+% This lowering is controlled by 
+% a plugin mechanism which enables different in-memory representations to 
+% be added to the compiler by implementing an interface 
+% (Section \ref{sec:representations}). Vectorization of tree traversals
+% is also explicitly represented in LIR.
+Subsequently, the LIR is lowered to LLVM and then JIT'ed to the 
+specified target processor.
 % \subsection{Optimizations on High-Level IR}

### 2024-04-16 results/python/analyze_gpu_speedups_results.py
Ignore batch size 256

@@ -121,6 +121,7 @@ def plot_bar_graph_speedups_for_batch_size(df, batch_size, kwargs=None):
 
 def plot_line_graph_geomean_speedups_over_batch_size(df, tb_time_col, compare_time_cols, line_labels, gpu_name, time_measured):
     batch_sizes = df['Batch size'].unique()
+    batch_sizes = [batch_size for batch_size in batch_sizes if batch_size != 256]
     speedups = []
 
     assert(len(compare_time_cols) == len(line_labels))

### 2024-04-16 docs/SOSP 2024/results.tex
Add abstract and XGB comparison
Add Tahoe vs TB on T400 graph
Edit schedule epxloration section
Add new graphs

@@
 \section{Experimental Evaluation}
+\label{sec:results}
+
+\begin{figure*}[ht]
+  \centering
+  \begin{subfigure}[b]{.45\textwidth}
+    \subcaptionbox*{}{\includegraphics[width=\textwidth]{figures/RandomModels/kernel_speedup_b512_depth8.png}}
+    \caption{Batch size 512, depth 8}
+  \end{subfigure}
+  \begin{subfigure}[b]{.45\textwidth}
+    \subcaptionbox*{}{\includegraphics[width=\textwidth]{figures/RandomModels/kernel_speedup_b4096_depth6.png}}
+    \caption{Batch size 4096, depth 6}
+  \end{subfigure}
+  \hfill
+  \caption{\Treebeard{} vs RAPIDs Kernel Time Speedup on NVIDIA RTX 4060 for several randomly generated models.}
+\end{figure*}
 
@@
   \centering
-  \includegraphics[width=\linewidth]{figures/TBvsRAPIDs_4060_Speedup.png}
-  \caption{\Treebeard{} vs RAPIDs Speedup on NVIDIA RTX 4060.}
-  \label{Fig:TBvsRAPIDs_4060_Speedup}
+  \includegraphics[width=0.75\linewidth]{figures/geomean_speedup_4060_kernel_time.png}
+  \caption{\Treebeard{} vs RAPIDs and Tahoe Kernel Time Speedup on NVIDIA RTX 4060}
+  \label{Fig:TBvsRAPIDsTahoe_4060_Speedup}
 \end{figure}
@@
   \centering
-  \includegraphics[width=\linewidth]{figures/TBvsTahoe_4060_KernelTimeSpeedup.png}
-  \caption{\Treebeard{} vs Tahoe Kernel Time Speedup.}
-  \label{Fig:TBvsTahoe_4060_KernelTimeSpeedup}
+  \includegraphics[width=0.75\linewidth]{figures/geomean_speedup_4060_total_time.png}
+  \caption{\Treebeard{} vs RAPIDs Total Time Speedup on NVIDIA RTX 4060.}
+  \label{Fig:TBvsRAPIDs_4060_TotalTimeSpeedup}
 \end{figure}
+
+\begin{figure*}[ht]
+  \centering
+  \begin{subfigure}[b]{.45\textwidth}
+    \subcaptionbox*{}{\includegraphics[width=\textwidth]{figures/speedup_bar_graph_1024.png}}
+    \caption{Batch size 1024}
+  \end{subfigure}
+  \begin{subfigure}[b]{.45\textwidth}
+    \subcaptionbox*{}{\includegraphics[width=\textwidth]{figures/speedup_bar_graph_8192.png}}
+    \caption{Batch size 8192}
+  \end{subfigure}
+  \hfill
+  \caption{Kernel time speedup of \Treebeard{} vs RAPIDs on NVIDIA RTX 4060. Numbers on the bars are 
+  inference times per sample in $\mu$s for RAPIDs and Tahoe.}
+\end{figure*}
 
@@
   \centering
-  \includegraphics[width=\linewidth]{figures/TBvsRAPIDs_AllBatchSizes_T400.png}
-  \caption{\Treebeard{} vs RAPIDs Speedup on T400.}
-  \label{Fig:TBvsRAPIDs_AllBatchSizes_T400}
+  \includegraphics[width=0.75\linewidth]{figures/geomean_speedup_T400_kernel_time.png}
+  \caption{\Treebeard{} vs RAPIDs and Tahoe Kernel Time Speedup on NVIDIA T400.}
+  \label{Fig:TBvsRAPIDsTahoe_T400_Speedup}
 \end{figure}
+
+\begin{figure*}[ht]
+  \centering
+  \begin{subfigure}[b]{.45\textwidth}
+    \subcaptionbox*{}{\includegraphics[width=\textwidth]{figures/abs_times_bar_graph_1024.png}}
+    \caption{Batch size 1024.}
+  \end{subfigure}
+  \begin{subfigure}[b]{.45\textwidth}
+    \subcaptionbox*{}{\includegraphics[width=\textwidth]{figures/abs_times_bar_graph_8192.png}}
+    \caption{Batch size 8192.}
+  \end{subfigure}
+  \hfill
+  \caption{\Treebeard{} vs RAPIDs total time comparison on NVIDIA RTX 4060. Numbers on the bars are the times 
+  per sample in $\mu$s for \Treebeard{} and RAPIDs. Times for each benchmark are normalized w.r.t the RAPIDs time for that benchmark.}
+\end{figure*}
 
@@
   \centering
-  \includegraphics[width=\linewidth]{figures/AutotuningSpeedupvs4060Sched_T400.png}
+  \includegraphics[width=0.75\linewidth]{figures/AutotuningSpeedupvs4060Sched_T400.png}
   \caption{Autotuning heuristics speedup vs best 4060 schedule on T400.}
@@
   \centering
-  \includegraphics[width=\linewidth]{figures/HeuristicVsFullExplore_Speedup.png}
+  \includegraphics[width=0.75\linewidth]{figures/HeuristicVsFullExplore_Speedup.png}
   \caption{Autotuning heuristic compile time speedup vs full schedule exploration.}
@@
   \centering
-  \includegraphics[width=\linewidth]{figures/AMD_MI210_ATHeuristicVs4060Sched_speedup.png}
+  \includegraphics[width=0.75\linewidth]{figures/TBvsXGB_TotalTime.png}
+  \caption{\Treebeard{} vs XGBoost Speedup on RTX 4060.}
+  \label{Fig:TBvsXGBoost_Speedup}
+\end{figure}
+
+\begin{figure}[htb]
+  \centering
+  \includegraphics[width=0.75\linewidth]{figures/AMD_MI210_ATHeuristicVs4060Sched_speedup.png}
   \caption{Autotuning heuristics speedup vs best 4060 schedule on MI210.}
@@
 \end{figure}
-
-\begin{figure*}[ht]
-  \centering
-  \begin{subfigure}[b]{.3\textwidth}
-    \subcaptionbox*{}{\includegraphics[width=\textwidth]{figures/RandomModels/kernel_speedup_b512_depth6.png}}
-    \caption{Kernel Speedup for batch size 512, depth 6.}
-  \end{subfigure}
-  \hfill  % Add horizontal space between subfigures
-  \begin{subfigure}[b]{.3\textwidth}
-    \subcaptionbox*{}{\includegraphics[width=\textwidth]{figures/RandomModels/kernel_speedup_b512_depth7.png}}
-    \caption{Kernel Speedup for batch size 512, depth 7.}
-  \end{subfigure}
-  \hfill
-  \begin{subfigure}[b]{.3\textwidth}
-    \subcaptionbox*{}{\includegraphics[width=\textwidth]{figures/RandomModels/kernel_speedup_b512_depth8.png}}
-    \caption{Kernel Speedup for batch size 512, depth 8.}
-  \end{subfigure}
-  \\  % Line break for the second row
-  \begin{subfigure}[b]{.3\textwidth}
-    \subcaptionbox*{}{\includegraphics[width=\textwidth]{figures/RandomModels/kernel_speedup_b4096_depth6.png}}
-    \caption{Kernel Speedup for batch size 4096, depth 6.}
-  \end{subfigure}
-  \hfill
-  \begin{subfigure}[b]{.3\textwidth}
-    \subcaptionbox*{}{\includegraphics[width=\textwidth]{figures/RandomModels/kernel_speedup_b4096_depth7.png}}
-    \caption{Kernel Speedup for batch size 4096, depth 7.}
-  \end{subfigure}
-  \hfill
-  \begin{subfigure}[b]{.3\textwidth}
-    \subcaptionbox*{}{\includegraphics[width=\textwidth]{figures/RandomModels/kernel_speedup_b4096_depth8.png}}
-    \caption{Kernel Speedup for batch size 4096, depth 8.}
-  \end{subfigure}
-\end{figure*}

### 2024-04-16 results/python/analyze_random_model_benchmark_results.py
Add new graphs for random model speedups

@@ -1,9 +1,9 @@
 import os
 
-batch_size = 4096
+batch_size = 512
 # file_path = "/home/ashwin/mlir-build/llvm-project/mlir/examples/treebeard/results/thedeep/gpu/random_models/20240320/random_cmp_output_mp.txt"
-# file_path = "/home/ashwin/mlir-build/llvm-project/mlir/examples/treebeard/results/thedeep/gpu/random_models/20240320/random_cmp_output_mp_b512.txt"
-file_path = "/home/ashwin/mlir-build/llvm-project/mlir/examples/treebeard/results/thedeep/gpu/random_models/20240320/random_cmp_output_mp_b4k_ft500.txt"
+file_path = "/home/ashwin/mlir-build/llvm-project/mlir/examples/treebeard/results/thedeep/gpu/random_models/20240320/random_cmp_output_mp_b512.txt"
+# file_path = "/home/ashwin/mlir-build/llvm-project/mlir/examples/treebeard/results/thedeep/gpu/random_models/20240320/random_cmp_output_mp_b4k_ft500.txt"
 
 # read the file line by line and filter out lines that don't start with "/"
 with open(file_path, "r") as file:
@@ -42,20 +42,20 @@ def get_tree_depth(model_path):
 
 df["tree_depth"] = df["model_path"].apply(get_tree_depth)
 
-print(df.shape)
+# print(df.shape)
 # print the min, max and geomean of the tb_kernel_speedup column
-print(df["tb_kernel_speedup"].min())
-print(df["tb_kernel_speedup"].max())
+# print(df["tb_kernel_speedup"].min())
+# print(df["tb_kernel_speedup"].max())
 # print the geometric mean of the tb_kernel_speedup column
 # compute geomean using scipy.stats.gmean
 import scipy.stats as stats
-print(stats.gmean(df["tb_kernel_speedup"]))
+# print(stats.gmean(df["tb_kernel_speedup"]))
 #print(df["tb_kernel_speedup"].prod() ** (1.0 / df.shape[0]))
 
 for depth in range(6, 9):
     # get entries with tree_depth == depth
     df_depth = df[df["tree_depth"] == depth]
-    print(df_depth.shape)
+    # print(df_depth.shape)
 
     # plot a 3-d graph of num_trees, num_features and tb_kernel_speedup
     import matplotlib.pyplot as plt
[…]

### 2024-04-16 docs/SOSP 2024/results.tex
Add new graphs and some points to results section

@@
 \label{sec:results}
+
+We evaluate \Treebeard{} on several machines. The first machine has an AMD Ryzen 9 7950X 16-Core processor,
+128 GB of RAM, an NVIDIA RTX 4060 GPU with 8 GB of RAM and runs Ubuntu 22.04.2 LTS with CUDA 11.5. The 
+second has an Intel Core i9-11900K (Rocket Lake) processor with 8 physical cores, 128 GB of RAM, an 
+NVIDIA T400 with 2GB of RAM and runs Ubuntu 20.04.3 LTS with CUDA 11.8. We also evaluate \Treebeard{} on 
+an AMD MI210 GPU. To establish the efficacy of \Treebeard{}, we compare it with NVIDIA RAPIDs v23.10,
+Tahoe, and XGBoost v1.7.6. We use the same benchmark models as in the \TreebeardOLD{} paper\cite{Treebeard}.
+We also evaluate \Treebeard{} on a set of randomly generated models.
+
+\subsection{Performance on Random Models}
+\begin{itemize}
+  \item Generated a set of over 700 random models with varying depths (6, 7, 8), number of trees (100 - 1000 in steps of 100), 
+  and number of features (powers of 2 between 8 to 1024 inclusive).
+  \item Compared the kernel time speedup of \Treebeard{} vs RAPIDs on the RTX 4060. The schedule used was the one picked by the auto-tuner. 
+  \item Speedups range between 1.5$\times$ and 8$\times$. \Treebeard{} outperforms RAPIDs on all models and batch sizes tested.
+  \item Figure \ref{fig:randomModels4060} shows the speedup for models of depth 8 and batch size 512 and depth 6 and batch size 4096.
+  Trends for other batch sizes and depths are similar.
+\end{itemize}
 
@@
   \hfill
-  \caption{\Treebeard{} vs RAPIDs Kernel Time Speedup on NVIDIA RTX 4060 for several randomly generated models.}
+  \caption{\label{fig:randomModels4060}\Treebeard{} vs RAPIDs Kernel Time Speedup on NVIDIA RTX 4060 for several randomly generated models.}
 \end{figure*}
+
+\subsection{Comparison with RAPIDs, Tahoe and XGBoost}
+\begin{itemize}
+  \item Measured both the kernel time and total time (time including transferring data to the 
+  GPU and results back) for RAPIDs and \Treebeard{}.
+  \item Tahoe only allows us to measure the kernel time since it is written as an executable 
+  that performs inference repeatedly on the same data that is transferred to the GPU once.
+  \item Tahoe does not support multiclass models. We just ran the multiclass models (\op{covtype} and \op{letters})
+  as regression models for the comparison. Tahoe also gives wrong results (as reported by its own tests) for 
+  \op{letters} and \op{year}. In these cases, we pick the time of the fastest variant that gives the correct results.
+  \item Compared the kernel time and total time speedup of \Treebeard{} vs RAPIDs on the RTX 4060 and T400.
+  \item Compared the total time speedup of \Treebeard{} vs XGBoost on the RTX 4060.
+  \item The schedule used was the one picked by the auto-tuner.
+  \item \Treebeard{} outperforms RAPIDs at all batch sizes as shown in Figure \ref{Fig:TBvsRAPIDsTahoe_4060_Speedup}. 
+  \item \Treebeard{} outperforms Tahoe on all models and batch sizes tested. Individual benchmark speedups range between 1.1$\times$ and 16$\times$.
+  \item \Treebeard{} is faster than XGBoost by more than an order of magnitude. Results are not shown because the speedups don't fit on the same graph.
+  \item Figure \ref{Fig:TBvsRAPIDs_4060_TotalTimeSpeedup} shows that \Treebeard{} offers substantial speedup over RAPIDs even when 
+  data needs to be transferred to the GPU and results need to be transferred back.
+  \item Figure \ref{Fig:TBvsRAPIDsTahoe_T400_Speedup} shows that these speedups are also observed on the T400 thus showing 
+  that \Treebeard{} offers portable performance across different GPUs.
+  \item Figure \ref{Fig:KernelTimeIndividualBenchmarks4060} shows that \Treebeard{} outperforms RAPIDs and Tahoe on 
+  individual benchmarks on the RTX 4060 at small (1024) and large (8192) batch sizes. These batch sizes 
+  require different schedules, but the \Treebeard{} auto-tuning heuristic is able to find them.
+  \item Figure \ref{Fig:TotalTimeIndividualBenchmarks4060} compares the total time of \Treebeard{} vs RAPIDs on individual benchmarks. 
+  It shows that even with the overhead of data transfers, \Treebeard{} offers significant speedup. The graph in Figure \ref{Fig:TotalTimeIndividualBenchmarks4060} 
+  shows the breakup of the total time into kernel time and data transfer time.
+\end{itemize}
 
@@
   \hfill
-  \caption{Kernel time speedup of \Treebeard{} vs RAPIDs on NVIDIA RTX 4060. Numbers on the bars are 
+  \caption{\label{Fig:KernelTimeIndividualBenchmarks4060}Kernel time speedup of \Treebeard{} vs RAPIDs on NVIDIA RTX 4060. Numbers on the bars are 
   inference times per sample in $\mu$s for RAPIDs and Tahoe.}
@@
   \hfill
-  \caption{\Treebeard{} vs RAPIDs total time comparison on NVIDIA RTX 4060. Numbers on the bars are the times 
+  \caption{\label{Fig:TotalTimeIndividualBenchmarks4060}\Treebeard{} vs RAPIDs total time comparison on NVIDIA RTX 4060. Numbers on the bars are the times 
   per sample in $\mu$s for \Treebeard{} and RAPIDs. Times for each benchmark are normalized w.r.t the RAPIDs time for that benchmark.}
@@
 
+\begin{figure}[htb]
+  \centering
+  \includegraphics[width=0.75\linewidth]{figures/TBvsXGB_TotalTime.png}
+  \caption{\Treebeard{} vs XGBoost Speedup on RTX 4060.}
+  \label{Fig:TBvsXGBoost_Speedup}
+\end{figure}
+
+\subsection{Autotuning Heuristic}
 \begin{figure}[htb]
@@
 
-\begin{figure}[htb]
-  \centering
-  \includegraphics[width=0.75\linewidth]{figures/TBvsXGB_TotalTime.png}
-  \caption{\Treebeard{} vs XGBoost Speedup on RTX 4060.}
-  \label{Fig:TBvsXGBoost_Speedup}
-\end{figure}
-
+\subsection{AMD GPU}
 \begin{figure}[htb]
@@
 \end{figure}
+
+\subsection{CPU Improvements}

### 2024-04-16 docs/SOSP 2024/exploring.tex
search

@@
 % \end{itemize}
-
 The set of schedules that can be constructed using the scheduling language described in 
-Section \ref{sec:schedule} is unbounded. Searching this schedule space to find a 
-high-performance schedule is a non-trivial task. To simplify this process, we design a
-template schedule for GPUs that encompasses several strategies published in prior work.
-Our template schedule assigns a configurable number of rows to each thread block and to each thread.
-It distributes the trees across a specified number of threads and can cache trees and 
-input rows if required while unrolling and interleaving of tree walks.
-Table \ref{tab:schedparams} lists the parameter values we tried. 
+Section \ref{sec:schedule} is unbounded.
+\TODO{kr: next few lines need more work}
+Even if one were to search for a good schedule offline, before deploying the model on a specific hardware, additional tools are needed help search for a 
+high-performance schedule. We anticipate that even for the most complex models, one would like to spend at most a few minutes doing this search.
+We propose a set of heuristics to guide the search for a good schedule and meet this goal. We do so in two steps, by first defining a bounded search space and then pruning this search space further.
+\subsection{Bounding the search space}
+We bound the space using a few meta parameters that together define the primitives used to construct a schedule.
+We do so while making sure that the space (at-least) covers the known strategies published in prior work.
+Specifically, we use 4 numeric parameters and 4 boolean parameters listed in Table \ref{tab:schedparams}.
+%Note that even this template has a large number ($1728$) of schedules, and requires further pruning as we discuss later.
+The first three numeric parameters assign a configurable number of rows to each thread block, to each thread, and determine how trees are distributed across a specified number of threads.
+These parameters together define the loop schedule primitives to use (including the arguments to pass) from Table~\ref{Tab:LoopModifiers} and the parallelization strategy.
+The next four parameters determine the caching strategy, unroll factor{\footnote{We choose to always unroll trees completely, while its possible to partially unroll loops we did not observe any performance benefits from doing so.}} and how many trees to traverse simultaneously \TODO{Should we say, in the kernel/within a single thread?}.
+(the remaining primitives from Table~\ref{Tab:Optimizations}).
+The final parameter determines what reduction options to use (Table~\ref{Tab:ReductionOpts}).
+%Table \ref{tab:schedparams} lists the parameter values we tried. 
 % \begin{itemize}
@@
 
-While the template schedule simplifies code generation, finding a good set of 
-parameter values is still hard.
+We evaluated all the schedules in this space on several real world models and observed that there is huge variation in performance with different parameter values. 
 Figure \ref{Fig:ExecTimeDistribution} shows the distribution of normalized execution times
-for all benchmark models with different parameter values for the template schedule (inference
+for some real-world models with different parameter values for the template schedule (inference
 times normalized w.r.t fastest time for that model).
-There is a significant amount of variation in performance even within the variants of the
-template schedule,. Very few schedules perform close to the best while a vast majority of
+As can be seen very few schedules perform close to the best while a vast majority of
 schedules perform poorly.
 
-Exploring the schedule space extensively even for a reasonable set of parameter values
-is very expensive. We explored the set of parameter values listed in Table \ref{tab:schedparams}
-for our benchmarks and found that it took anywhere between thirty minutes up to a few hours
-to explore the entire space for each model. 
+Exhaustive exploration over this bounded space ($1728$ schedules) is still very expensive, it took anywhere between thirty minutes to a few hours
+to explore the entire space for a given model. 
+%Exploring the space for every model is infeasible.
+%\TODO{comment on online vs offline exploration? Set a goal for the heuristic?}
+\subsection{Pruning the search space}
+%Exploring the schedule space extensively even for a reasonable set of parameter values is very expensive.
+
 %Performing this extensive search for every model being compiled is infeasible in practise. 
-We therefore
+A careful analysis of the search space reveals that certain schedules are not likely to perform well as they either do not exploit the parallelism available in the model or do not take advantage of the hardware features.
+We use a combination of 3 strategies to further prune the search space.
+Algorithm \ref{alg:heuristic} presents our final heuristic to find a good schedule.
 %need a better mechanism to guide the search for a good schedule.
-design a heuristic to narrow down the set of schedules to explore based on the following 
-observations.
-\begin{itemize}
-  \item For small batch sizes, the best schedules tend to have a small number of rows
-  per thread block and partition the trees across a larger number of threads. This is 
-  intuitive since the amount of data parallelism across the rows is limited for small batch sizes.
-  \item Always cache rows in shared memory and never cache trees. We find that caching rows 
+
+\subsubsection*{Insufficient parallelism}
+ Some configurations do not expose sufficient parallelism, we prune these out.
+  For example when the batch size is small, it does not make sense to have a many rows
+  per thread block or to partition the trees across a few threads. 
+  We therefore limit the combinations used based on batch size (see lines 3---8).
+\subsubsection*{Utilizing shared memory}
+  Shared memory is a critical resource on GPUs and it helps to only bring objects that would be reused into it. 
+  We observe that tree nodes have limited reuse and the one time cost of loading trees is not sufficiently amortized when the whole tree is not accessed during inference. 
+  On the other hand caching rows 
   %when possible (i.e., when the number of features is small enough to fit in shared memory)
-  almost always improves performance. Caching trees on the other hand almost always degrades 
-  performance. This is because the one time cost of loading trees into shared memory is 
-  not sufficiently amortized when the whole of the tree is not accessed during inference. 
-  \item Models with a large number of features tend to benefit from partitioning the 
-  trees across more threads even at larger batch sizes. This is because processing fewer rows at a time 
-  allows us to keep them in shared memory.
+  almost always improves performance. Further when the inputs have many features, it may not be possible to cache all rows.
+  In this scenario it helps to retain a few rows in memory, and pick schedules similar to the small batch size case (lines 15,3---5).
+
+  %trees across more threads even at larger batch sizes. This is because processing fewer rows at a time 
+  %allows us to keep them in shared memory.
   % We empirically find that the threshold for when 
@@
   % is greater than 100. 
-  \item We find that when a model prefers schedules with shared reduction, the same schedules 
-  without shared reduction are among the best performing schedules without shared reduction.
+  \subsubsection*{Orthogonal parameters} We find that some parameters like reduction type are orthogonal to the rest. 
+  We therefore break the exploration into two phases, first we pick the best schedules without reduction and then evaluate reduction options on the best schedules.
+    %We find that when a model prefers schedules with shared reduction, the same schedules 
+  %without shared reduction are among the best performing schedules without shared reduction.
   % We therefore are able to separate the evaluation of shared reduction by collecting the 
   % best schedules without shared reduction and only evaluating shared reduction on them. 
-  Evaluating the top 3 schedules for shared reduction is sufficient in practice.  
-\end{itemize}
+  Evaluating the top 3 schedules for reduction is sufficient in practice.
 
+\begin{comment}
 \Treebeard{} uses these observations to narrow down the set of schedules to explore. 
@@
 \op{TBConfigs}. A set of schedules based on these thread block configurations 
-is then computed (\op{schedules}). The model is compiled with each of these schedules 
-and then the resulting inference code is profiled. The three best performing schedules 
-are collected and shared reduction is enabled on them and the resulting schedules evaluated.
+is then computed (\op{schedules}). 
+\end{comment}
+
+\noindent
+\Treebeard{} performs an exhaustive search over the pruned schedules from Algorithm~\ref{alg:heuristic} to find the best schedule.
+The model is compiled with each of these schedules and evaluated on a few input batches. 
+%The three best performing schedules 
+%are collected and shared reduction is enabled on them and the resulting schedules evaluated.
 The best schedule among all the evaluated schedules is selected as the schedule to use.
-We find that this heuristic is able to find schedules that are close to the best schedules
-but improves the search time by two orders of magnitude as we show in Section \ref{sec:results}.
+We report that this heuristic is able to find schedules that are close to the best schedules
+while improving the search time by two orders of magnitude (see Section \ref{sec:results}).
 

### 2024-04-17 docs/SOSP 2024/schedule.tex
Update scheduling language section

@@
 As established in Section \ref{sec:motivation}, there are several 
-different configurations and optimizations strategies for decision 
-tree inference. The best ones vary significantly across
+configurations and optimizations strategies for decision 
+tree inference. The best one to use varies significantly across
 models, batch sizes and hardware platforms. Therefore, designing
@@
 other optimizations as an input to the compiler. The specified 
-schedule controls the lowering of model inference to a set of 
-loop nests. This configurability provided by the schedule allows us 
+\textbf{\emph{schedule}} controls the lowering of model inference to a set of 
+loop nests. The configurability provided by the schedule allows us 
 to build auto-schedulers and auto-tuners (Section \ref{sec:exploring}).
@@
 \textbf{\emph{index variable}} which abstractly represents a loop. 
-The language then provides directives to manipulate these index 
+Each index variable has a range of values that it can take along 
+with a step. 
+The language provides directives to manipulate these index 
 variables. There are two special index variables -- \op{batch} and
@@
 new index variables from these root index variables by applying
-directives.
+directives. \TODO{Should we somehow explain what tree and batch loop are?}
 
@@
 walk the iteration space (Table \ref{Tab:LoopModifiers}). The second is a set of 
-directives that enable optimizations on a loop (Table \ref{Tab:Optimizations}). 
+directives that enable optimizations (Table \ref{Tab:Optimizations}). 
 Finally, we have a class of attributes that enable reduction specific optimizations
@@
 
-The set of loops (index variables) are internally represented as nodes in a tree where the children
+The compiler internally represents loops (index variables) 
+as nodes in a tree where the children
 of a node represent immediately contained loops. Each schedule 
@@
    \multirow{6}{*}{\texttt{split}} & \textbf{indexVar}  & \multirow{6}{*}{\parbox{0.55\linewidth}{Fiss the loop represented by \textbf{indexVar}
-   at iteration \textbf{splitIter}. Resulting loops will be represented by \textbf{first} and \textbf{second}. Returns a maps from nested 
-   index variables to new ones created by splitting.}} \\
+   at iteration \textbf{splitIter}. Resulting loops will be represented by \textbf{first} and \textbf{second}.}} \\
                                                &  \textbf{first} & \\
@@
    \multirow{4}{*}{\texttt{reorder}} & \textbf{indices[]}  & \multirow{4}{*}{\parbox{0.55\linewidth}{Permute loops corresponding to the specified index variables.
-   The loops must be perfectly nested in the current loop structure.}} \\
+   The loops must be perfectly nested.}} \\
                                                &   & \\
@@
 strategies used in existing systems. We show examples of how it can be used to 
-represent XGBoost and Tahoe's strategies. Before presenting the examples, we note that \Treebeard{}'s 
-default loop order is [\op{batch}, \op{tree}], i.e, for each row in the input
-batch, go over all trees.
+represent XGBoost and Tahoe's strategies. 
+% Before presenting the examples, we note that \Treebeard{}'s 
+% default loop order is [\op{batch}, \op{tree}], i.e, for each row in the input
+% batch, go over all trees.
 
@@
 over a fixed number of rows (64 in the previous version)
-for every tree and then moving to the next tree. When all trees 
-have been walked for this set of rows, the next set of rows is 
-taken up. Different sets of rows are processed in parallel.
+for every tree and then moving to the next tree. It moves to the 
+next set of rows when all trees have been walked for the current 
+set of rows. Different sets of rows are processed in parallel.
 The following schedule expresses XGBoost's strategy.
@@
 We show how two of these strategies can be encoded using \Treebeard{}'s scheduling language.
-%The rest can be encoded similarly. 
-\begin{itemize}
-  \item In the \emph{direct method}, a single GPU thread walks all trees
+The rest can be encoded similarly.
+
+In the \emph{direct method}, a single GPU thread walks all trees
 for a given input row. The schedule for this strategy is as follows.
@@
 Here, \op{ROWS\_PER\_TB} is the number of rows that are processed by a single thread block.
-  \item In the \emph{shared data} strategy, a thread block walks all the trees 
+
+In the \emph{shared data} strategy, a thread block walks all the trees 
 for a given row in parallel. 
@@
 % \end{lstlisting}
-\end{itemize}
 
@@
 % the current design is able to express most strategies of interest.
+
+In summary, \Treebeard{}'s scheduling language provides a convenient way to
+encode a wide range of strategies and to control how the compiler lowers 
+the inference routine to optimized target code. The next section 
+discusses how the rest of the compiler is structured.

### 2024-04-17 docs/SOSP 2024/hir-mir-opts.tex
Edit compiler overview section

@@
 Since the \TreebeardOLD{} infrastructure was originally designed to target CPUs,
-significant extensions  (shown as dark boxes in Figure \ref{Fig:CompilerStructure})
-were required to support schedule guided compilation for CPUs and GPUs. 
+significant extensions were required to support schedule-guided compilation for CPUs and GPUs. 
+These parts of the system are shown as shaded boxes in Figure \ref{Fig:CompilerStructure}.
 % However, it implements several optimizations on the HIR and MIR that 
@@
 the HIR that are implemented through the scheduling language (Section \ref{sec:schedule}).
-We use these to implement the automatic scheduling described in Section \ref{sec:exploring}. 
-Additionally, the \TreebeardOLD{} infrastructure implements HIR transformations to reorder and pad 
-trees. 
+The \emph{schedule} is implemented as an MLIR attribute on the \op{predictEnsemble} operation.
+We use this object to implement the automatic scheduling described in Section \ref{sec:exploring}. 
+We reuse HIR transformations to reorder and pad trees from \TreebeardOLD{}.
+The tree transformations that reorder and pad trees are used in conjunction 
+with loop transformations like splitting to specialize inference code as the example 
+in Section \ref{sec:motivation} shows. Also, we enable tree padding only if the schedule 
+specifies unrolling of tree walks.
 % Figure \ref{Fig:HIRExample} shows this representation for a model 
@@
 The HIR is lowered to MIR as dictated by the \emph{schedule}.
-In the lowered MIR, the compiler uses the \op{reduce} op from 
+Optimizations like tree-walk unrolling and tree-walk interleaving
+are performed on the MIR. 
+% These optimizations are beneficial for GPUs as well.
+% and the design of \TreebeardOLD{} allows us to reuse the tree-walk 
+% unrolling and tree-walk interleaving optimizations on the MIR for GPUs.
+% While building \Treebeard{}, we found that
+One surprising thing we found while developing \Treebeard{} was that 
+we could use ILP to improve performance on GPUs. 
+One of the performance bottlenecks in inference code targeted 
+to GPUs was that warps spent significant time being stalled. Since GPUs 
+implement scoreboarding~\cite{HennesseyPatterson}, we were able to 
+alleviate this bottleneck by interleaving tree walks. 
+% This significantly improved performance of generated 
+% code. The fact that exploiting ILP improved performance on GPUs was surprising.
+
+In the generated MIR, the compiler uses the \op{reduce} op from 
 the \op{reduction} dialect we design (details in Section \ref{sec:reduction})
@@
 
-\TreebeardOLD{} implements optimizations like 
-tree-walk unrolling, tree-walk interleaving, and parallelization
-on the MIR. These optimizations are beneficial for GPUs as well.
-and the design of \TreebeardOLD{} allows us to reuse the tree-walk 
-unrolling and tree-walk interleaving optimizations on the MIR for GPUs.
-While building \Treebeard{}, we found that one of the performance bottlenecks on the 
-GPU was that warps spent significant time being stalled. Since GPUs 
-implement scoreboarding~\cite{HennesseyPatterson}, we were able to alleviate this bottleneck by
-interleaving tree walks. This significantly improved performance of generated 
-code. The fact that exploting ILP improved performance on GPUs was surprising.
-
 % \TODO{Should we talk about how interleaving is implemented as a 
@@
 
-The MIR is then further lowered to a low-level IR (LIR). 
+The MIR is further lowered to a low-level IR (LIR). 
 Significant changes to the original \TreebeardOLD{} design were required to get
 LIR to correctly lower to GPU code. The most important of these was 
-the change to how the compiler implements support for in-memory 
+changing how the compiler implements support for in-memory 
 representations of models (Section \ref{sec:representations}).
@@
 % is implemented differently etc.). 
-Also, this is the
-level at which the compilation pipeline diverges for different targets. In 
-the GPU compilation pipeline, the required memory transfers and kernel
+Also, when the target processor is a GPU, the required memory transfers and kernel
 invocations are inserted into the LIR. Additionally, buffers 

### 2024-04-17 results/python/analyze_gpu_auto_schedule_results.py
Fix normalized time histogram

@@ -39,10 +39,16 @@ def plot_normalized_time_histogram(df_sorted, df_best, col_name):
     # print(df_sorted)
     # plot a histogram of the normalized times and save it to a file
     df_sorted.hist(column='normalized_'+col_name, bins=100)
+    plt.xlabel('Normalized Kernel Time')
+    plt.ylabel('Frequency')
+    plt.title('')
     plt.savefig(f'normalized_{col_name}_histogram.png')
 
     # plot a histogram of values with normalized times < 5
     df_sorted[df_sorted['normalized_'+col_name] < 5].hist(column='normalized_'+col_name, bins=500)
+    plt.xlabel('Normalized Kernel Time')
+    plt.ylabel('Frequency')
+    plt.title('')
     plt.savefig(f'normalized_{col_name}_histogram_lt5.png')
 
 def write_best_config_maps_to_file(df_best_total_time, benchmark_names, batch_sizes, file_name):
@@ -258,11 +264,11 @@ for benchmark in benchmark_names:
 
 # print(df_best_kernel_time)
 
-print('Comparison across batch sizes')
-compare_across_batch_sizes(df_best_kernel_time, df_sorted_by_kernel_time, batch_sizes, benchmark_names)
+# print('Comparison across batch sizes')
+# compare_across_batch_sizes(df_best_kernel_time, df_sorted_by_kernel_time, batch_sizes, benchmark_names)
 
-print()
-print('Comparison across benchmarks')
-compare_across_benchmarks(df_best_kernel_time, df_sorted_by_kernel_time, batch_sizes, benchmark_names)
-# plot_normalized_time_histogram(df_sorted_by_kernel_time, df_best_kernel_time, 'kernel')
+# print()
+# print('Comparison across benchmarks')
+# compare_across_benchmarks(df_best_kernel_time, df_sorted_by_kernel_time, batch_sizes, benchmark_names)
+plot_normalized_time_histogram(df_sorted_by_kernel_time, df_best_kernel_time, 'kernel')
 # write_best_config_maps_to_file(df_best_total_time, benchmark_names, batch_sizes, "best_configs.txt")

### 2024-04-17 results/python/analyze_gpu_speedups_results.py
1. Improve existing graphs - make fonts more visible 2. Add new line graphs for speedups

@@ -7,8 +7,17 @@ from scipy.stats import gmean
 # Read the CSV file into a dataframe
 results_4060_df = pd.read_csv('../4060_thedeep_results.csv')
 results_t400_df = pd.read_csv('../T400_holmes_results.csv')
+fullexp_vs_at_df = pd.read_csv('../fullexp_vs_at.csv')
+
+results_4060_df = results_4060_df[results_4060_df['Batch size'] != 256]
+results_t400_df = results_t400_df[results_t400_df['Batch size'] != 256]
+fullexp_vs_at_df = fullexp_vs_at_df[fullexp_vs_at_df['Batch size'] != 256]
 
 PLOT_FILE_EXTENSION = 'png'
+AXIS_FONT_SIZE = 15
+LABEL_FONT_SIZE = 15
+LEGEND_FONT_SIZE = 12
+BAR_LABEL_FONT_SIZE = 8
 
 def plot_bar_graph_of_abs_times_for_batch_size(df, batch_size):
     filtered_df = df[df['Batch size'] == batch_size]
@@ -88,19 +97,25 @@ def plot_bar_graph_speedups_for_batch_size(df, batch_size, kwargs=None):
     _, ax = plt.subplots()
 
     # Plot the bar graph
-    width = 0.6
-    gap = 0.4
-    pos = np.array([float(i) + i * gap for i in range(len(bar_names))])
+    width = BAR_LABEL_FONT_SIZE
+    gap = BAR_LABEL_FONT_SIZE * 0.3
+    pos = np.array([float(2 * i * width + i * gap) for i in range(len(bar_names))])
 
     ax.bar(pos, rapids_speedup, width, label='RAPIDS')
     ax.bar(pos+width, tahoe_speedup, width, label='Tahoe')
 
+    # enable grid
+    # ax.grid(True, which='both', axis='y')
+
+    ax.set_ylim(0, max(max(rapids_speedup), max(tahoe_speedup)) + 1)
+
     # Add labels and title
-    ax.set_xlabel('Benchmark')
-    ax.set_ylabel('Speedup')
+    ax.set_xlabel('Benchmark', fontsize=LABEL_FONT_SIZE)
+    ax.set_ylabel('Speedup', fontsize=LABEL_FONT_SIZE)
     # ax.set_title(f'Speedups for Batch Size {batch_size} (4060)')
     ax.set_xticks(pos + width / 2)
[…]

### 2024-04-17 results/python/analyze_schedule_sensitivity_data.py

@@ -53,9 +53,14 @@ def plot_batch_sensitivity_for_single_benchmark(df_depth, benchmark_name):
         dims=("Batch Size", "Comparison Batch Size"),
     )
 
-    da.plot(xscale="log", yscale="log")
+    fg = da.plot(xscale="log", yscale="log", aspect=1.3, size=20)
+    fg.colorbar.ax.tick_params(labelsize=40)
+    for axis in [fg.axes.xaxis, fg.axes.yaxis]:
+        axis.label.set_size(60)
+        axis.set_tick_params(labelsize=60)
+
     # plt.show()
-    plt.title(f"{benchmark_name} Batch Sensitivity")
+    # plt.title(f"{benchmark_name} Batch Sensitivity")
     # plt.colorbar(im, label='tb_kernel_speedup')
     plt.savefig(f'batch_sensitivity_{benchmark_name}.png')
     plt.close('all')
@@ -82,19 +87,20 @@ def plot_model_sensitivity_for_single_benchmark(df_depth, model_name):
         coords={"Model": x_positions, "Comparison Model": y_positions},
         dims=("Model", "Comparison Model"),
     )
-    da.plot(aspect=1.5, size=10)
+    fg = da.plot(aspect=1.3, size=20)
+    fg.colorbar.ax.tick_params(labelsize=40)
     plt.axis('off')
 
     for name, x_pos in zip(x, x_positions):
         name = name if name != "year_prediction_msd" else "year" 
-        plt.text(x_pos, -1.5, name, rotation=90, fontsize=10, wrap=True)
+        plt.text(x_pos, 2, name, rotation=90, fontsize=60, va = 'top')
     
     for name, y_pos in zip(y, y_positions):
         name = name if name != "year_prediction_msd" else "year"
-        plt.text(-1, y_pos, name, fontsize=10, wrap=True)
+        plt.text(2, y_pos, name, fontsize=60, ha='right')
     
     # plt.show()
-    plt.title(f"Batch Size {batch_size} Model Sensitivity")
+    # plt.title(f"Batch Size {batch_size} Model Sensitivity")
     plt.tight_layout()
     # plt.colorbar(im, label='tb_kernel_speedup')
     plt.savefig(f'model_sensitivity_{batch_size}.png')

### 2024-04-17 docs/SOSP 2024/results.tex
Add bullets for CPU experiments and update kernel time dist histogram

@@
   require different schedules, but the \Treebeard{} auto-tuning heuristic is able to find them.
-  \item Figure \ref{Fig:TotalTimeIndividualBenchmarks4060} compares the total time of \Treebeard{} vs RAPIDs on individual benchmarks. 
-  It shows that even with the overhead of data transfers, \Treebeard{} offers significant speedup. The graph in Figure \ref{Fig:TotalTimeIndividualBenchmarks4060} 
-  shows the breakup of the total time into kernel time and data transfer time.
+  % \item Figure \ref{Fig:TotalTimeIndividualBenchmarks4060} compares the total time of \Treebeard{} vs RAPIDs on individual benchmarks. 
+  % It shows that even with the overhead of data transfers, \Treebeard{} offers significant speedup. The graph in Figure \ref{Fig:TotalTimeIndividualBenchmarks4060} 
+  % shows the breakup of the total time into kernel time and data transfer time.
 \end{itemize}
@@
 
-\begin{figure*}[ht]
-  \centering
-  \begin{subfigure}[b]{.45\textwidth}
-    \subcaptionbox*{}{\includegraphics[width=\textwidth]{figures/abs_times_bar_graph_1024.png}}
-    \caption{Batch size 1024.}
-  \end{subfigure}
-  \begin{subfigure}[b]{.45\textwidth}
-    \subcaptionbox*{}{\includegraphics[width=\textwidth]{figures/abs_times_bar_graph_8192.png}}
-    \caption{Batch size 8192.}
-  \end{subfigure}
-  \hfill
-  \caption{\label{Fig:TotalTimeIndividualBenchmarks4060}\Treebeard{} vs RAPIDs total time comparison on NVIDIA RTX 4060. Numbers on the bars are the times 
-  per sample in $\mu$s for \Treebeard{} and RAPIDs. Times for each benchmark are normalized w.r.t the RAPIDs time for that benchmark.}
-\end{figure*}
+% \begin{figure*}[ht]
+%   \centering
+%   \begin{subfigure}[b]{.45\textwidth}
+%     \subcaptionbox*{}{\includegraphics[width=\textwidth]{figures/abs_times_bar_graph_1024.png}}
+%     \caption{Batch size 1024.}
+%   \end{subfigure}
+%   \begin{subfigure}[b]{.45\textwidth}
+%     \subcaptionbox*{}{\includegraphics[width=\textwidth]{figures/abs_times_bar_graph_8192.png}}
+%     \caption{Batch size 8192.}
+%   \end{subfigure}
+%   \hfill
+%   \caption{\label{Fig:TotalTimeIndividualBenchmarks4060}\Treebeard{} vs RAPIDs total time comparison on NVIDIA RTX 4060. Numbers on the bars are the times 
+%   per sample in $\mu$s for \Treebeard{} and RAPIDs. Times for each benchmark are normalized w.r.t the RAPIDs time for that benchmark.}
+% \end{figure*}
 
@@
 \subsection{CPU Improvements}
+\begin{itemize}
+  \item The ability of \Treebeard{} to parallelize across both trees and rows has a significant impact on CPU performance.
+  \item We run experiments on the Intel Core i9-11900K processor to evaluate the impact of parallelizing across trees with small batch sizes. 
+  \item At batch size 32, we find that the average speedup over all 8 models is 2.2$\times$ with a max speedup of 5$\times$.
+  \item At batch size 64, we find that the average speedup over all 8 models is 1.1$\times$ with a max speedup of 2$\times$. 
+  However, we find that 2 of the 8 models show slowdowns in this case. Again, this highlights the need for schedule exploration on CPUs.
+  \item For small batch sizes, parallelizing across rows does not work great since there is limited reuse of trees in L1 cache.
+  Also, the amount of work per thread is very small leading to high overheads. 
+  \item Parallelizing across trees is more effective since there is more reuse in L1 cache and the amount of work per thread is higher.  
+\end{itemize}

### 2024-04-17 docs/SOSP 2024/results.tex
Edit results

@@
   \centering
-  \includegraphics[width=0.75\linewidth]{figures/geomean_speedup_4060_kernel_time.png}
-  \caption{\Treebeard{} vs RAPIDs and Tahoe Kernel Time Speedup on NVIDIA RTX 4060}
+  \includegraphics[width=0.75\linewidth]{figures/geomean_speedup_4060_kernel_time_total_time.png}
+  \caption{\Treebeard{} vs RAPIDs and Tahoe kernel time and total time speedup on NVIDIA RTX 4060}
   \label{Fig:TBvsRAPIDsTahoe_4060_Speedup}
@@
 
-\begin{figure}[htb]
-  \centering
-  \includegraphics[width=0.75\linewidth]{figures/geomean_speedup_4060_total_time.png}
-  \caption{\Treebeard{} vs RAPIDs Total Time Speedup on NVIDIA RTX 4060.}
-  \label{Fig:TBvsRAPIDs_4060_TotalTimeSpeedup}
-\end{figure}
+% \begin{figure}[htb]
+%   \centering
+%   \includegraphics[width=0.75\linewidth]{figures/geomean_speedup_4060_total_time.png}
+%   \caption{\Treebeard{} vs RAPIDs Total Time Speedup on NVIDIA RTX 4060.}
+%   \label{Fig:TBvsRAPIDs_4060_TotalTimeSpeedup}
+% \end{figure}
 
@@
   \centering
+  \includegraphics[width=0.75\linewidth]{figures/geomean_speedup_4060_best_vs_at.png}
+  \caption{Autotuning heuristics speedup vs best 4060 schedule on NVIDIA RTX 4060}
+  \label{Fig:AutotuningSpeedupvs4060Sched}
+\end{figure}
+
+\begin{figure}[htb]
+  \centering
   \includegraphics[width=0.75\linewidth]{figures/AutotuningSpeedupvs4060Sched_T400.png}
@@
   \centering
-  \includegraphics[width=0.75\linewidth]{figures/HeuristicVsFullExplore_Speedup.png}
+  \includegraphics[width=0.75\linewidth]{figures/geomean_speedup_4060_full_exp_vs_at.png}
   \caption{Autotuning heuristic compile time speedup vs full schedule exploration.}
@@
 \end{figure}
+
+\begin{itemize}
+  \item We compare the schedule found by the schedule exploration heuristic on the RTX 4060 with the best schedule found by
+  extensive exploration. Figure \ref{Fig:HeuristicVsFullExplore_Speedup} shows that the heuristic is able to find schedules that are
+  very close to the best schedule found by exhaustive exploration.
+  \item The speedup of the heuristic schedule vs the best 4060 schedule is shown in Figure \ref{Fig:AutotuningSpeedupvs4060Sched}.
+  Geomean speedup range from 1.1$\times$ to 1.2$\times$.
+  \item 
+\end{itemize}
 

### 2024-04-17 results/python/analyze_gpu_speedups_results.py
Add amd graphs, update other graphs after discussion

@@ -8,10 +8,12 @@ from scipy.stats import gmean
 results_4060_df = pd.read_csv('../4060_thedeep_results.csv')
 results_t400_df = pd.read_csv('../T400_holmes_results.csv')
 fullexp_vs_at_df = pd.read_csv('../fullexp_vs_at.csv')
+results_amdmi2160_df = pd.read_csv('../AMDMI2160_results.csv')
 
 results_4060_df = results_4060_df[results_4060_df['Batch size'] != 256]
 results_t400_df = results_t400_df[results_t400_df['Batch size'] != 256]
 fullexp_vs_at_df = fullexp_vs_at_df[fullexp_vs_at_df['Batch size'] != 256]
+results_amdmi2160_df = results_amdmi2160_df[results_amdmi2160_df['Batch size'] != 256]
 
 PLOT_FILE_EXTENSION = 'png'
 AXIS_FONT_SIZE = 15
@@ -136,14 +138,16 @@ def plot_bar_graph_speedups_for_batch_size(df, batch_size, kwargs=None):
 
 
 class LineGraphPlotter:
-    def __init__(self, gpu_name, file_name_suffix, lower_lim = 1, upper_lim = 1) -> None:
+    def __init__(self, gpu_name, file_name_suffix, lower_lim = 1, upper_lim = 1, round_up=True) -> None:
         self.gpu_name = gpu_name
         self.file_name_suffix = file_name_suffix
         _, self.ax = plt.subplots()
         self.ylim = (lower_lim, upper_lim)
+        self.round_up = round_up
 
     def plot_geomean_speedups_over_batch_size(self, df, tb_time_col, compare_time_cols, line_labels):
         batch_sizes = df['Batch size'].unique()
+        batch_sizes.sort()
         speedups = []
 
         assert(len(compare_time_cols) == len(line_labels))
@@ -167,7 +171,8 @@ class LineGraphPlotter:
         min_speedup = min([min(speedup) for speedup in speedups])
         diff = min_speedup - np.floor(min_speedup)
         lower_lim = np.floor(min_speedup) if diff <= 0.5 else np.floor(min_speedup) + 0.5
-        upper_lim = round(max([max(speedup) for speedup in speedups]) + 0.5, 1)
[…]

### 2024-04-17 docs/SOSP 2024/motivation.tex
Edit results and motivation

@@
 \label{sec:motivation}
-In this section, we first motivate the need for a scheduling language 
-by showing how a model can be compiled in different ways
-and subsequently, we show how drastically 
-performance can vary across these variants for real benchmarks.
+In this section, we first motivate our work by showing how a model can be compiled in different ways
+and subsequently, show the drastic performance difference across these variants for real benchmarks.
 
+\subsection{Motivating Example}
 \begin{figure}[htb]
@@
 branching by unrolling tree walks. However, it doesn't exploit  
-any parallelism and is therefore ill-suited for parallel processors.
+any parallelism. 
+%% RG
+%%  and is therefore ill-suited for parallel processors.
+Further, the \op{batch} for loop can be tiled to get locality benefits.
 
 One form of parallelism that can be exploited is to process rows in 
-parallel. However, with massively parallel processors like GPUs,
-this may not yield sufficient parallel work. Another option is to also 
-parallelize across trees. A possible strategy to accomplish this is 
-encoded in the following schedule.
+parallel. 
+%% RG:
+While this may work for multi-core CPUs, with massively parallel processors like GPUs,
+this strategy may not yield sufficient parallel work. Another option is to also 
+parallelize across trees. 
+%% RG:
+% With GPUs, the performance of generated code would also vary with the 
+% kernel configuration, which specifies the number of thread blocks\footnote{We 
+% use NVIDIA CUDA terminology in the discussion, but as we demonstrate in 
+% Section~\ref{sec:results}, our framework can generate code for AMD GPUs as well.}, 
+% the size of thread blocks and so on. 
+A possible strategy to accomplish this is encoded in the following schedule.
 \begin{lstlisting}[style=c++]
@@
 In the case of this schedule, the \op{reduce}
-operation needs special consideration. In order to correctly generate 
-code for this schedule, the compiler needs to determine that parallel 
-iterations of the \op{t1} 
+operation needs special consideration.  
+%% RG
+%% to correctly generate code for this schedule, the compiler needs 
+To generate the correct code for this schedule, the compiler needs to determine
+the required synchronization and temporary storage for reductions 
+given the parallelization strategy specified. In the given example, it 
+has to identify that parallel iterations of the \op{t1} 
 loop accumulate into the same element of the \op{result} array.
@@
 for simple schedules. Writing these routines by hand 
-is error-prone and time-consuming.  We believe that designing a 
+is error-prone and time-consuming.  
+%% RG 
+%% We beleive 
+Hence designing a 
 scheduling language to encapsulate these strategies and a principled code 
-generator to automatically generate code based on the schedule is the 
-best approach.
+generator to automatically generate code guided by the schedule is the 
+best approach. We promote such an approach in this paper.
+
+\subsection{Performance of Different Schedules}
 
@@
 
-To further complicate matters, we find that these different strategies can have significantly different
-performance. Figures \ref{fig:sensitivitya} and \ref{fig:sensitivityb} show the
-variation in performance when the same schedules are used across batch sizes 
-and models respectively. These diagrams show that the best schedules to use 
-vary across batch sizes and models. The largest slowdown is 5$\times$ when 
-schedules are used across different batch sizes and 6$\times$ when schedules
-are used across different models. Therefore, no one strategy can be used 
-across all models and batch sizes. 
-A system that is capable of specializing generated code for both batch 
+Do different schedules and the code generated based on them matter? 
+%% RG 
+%% To further complicate matters, we 
+We find that these different strategies can have significantly different
+performance. Figures~\ref{fig:sensitivitya} and~\ref{fig:sensitivityb} show the
+variation in performance when the same schedules are used across different batch sizes 
+and different models, respectively. 
+These diagrams show that the best schedules to use vary across batch sizes and models. 
+%% RG
+% Each chart shows the performance of different schedules~-- slowdown compared to the best
+% performing schedule~-- as a heatmap. 
+The largest slowdown is 5$\times$ 
+%% RG
+when the best-performing schedule for a model is  
+%% schedules are 
+used across different batch sizes.  %% and 
+It increases to 6$\times$ when schedules are used across different models. 
+%% Therefore, no one strategy can be used across all models and batch sizes. 
+Clearly, using a single strategy across models and batch sizes leaves significant performance on the table. Hence, 
+%% 
+a system that is capable of specializing generated code for both batch 
 size and model is required for the best performance.
@@
 cache on the CPU) and finally be able to generate target specific code. 
+%% RG
+Lastly, we propose a simple heuristic to find a high-performance schedule 
+for a given model and batch size to remove this burden from the user. 
+We defer the task of implementing an auto-tuner to future work. 
+%%
 The rest of the paper describes these challenges in detail and how we

### 2024-04-17 docs/SOSP 2024/results.tex
Edit results and motivation

@@
 
-\begin{figure}[htb]
-  \centering
-  \includegraphics[width=0.75\linewidth]{figures/TBvsXGB_TotalTime.png}
-  \caption{\Treebeard{} vs XGBoost Speedup on RTX 4060.}
-  \label{Fig:TBvsXGBoost_Speedup}
-\end{figure}
+% \begin{figure}[htb]
+%   \centering
+%   \includegraphics[width=0.75\linewidth]{figures/TBvsXGB_TotalTime.png}
+%   \caption{\Treebeard{} vs XGBoost Speedup on RTX 4060.}
+%   \label{Fig:TBvsXGBoost_Speedup}
+% \end{figure}
 
@@
   \centering
-  \includegraphics[width=0.75\linewidth]{figures/geomean_speedup_4060_best_vs_at.png}
+  \includegraphics[width=0.75\linewidth]{figures/geomean_speedup_T400_4060_vs_T400.png}
   \caption{Autotuning heuristics speedup vs best 4060 schedule on NVIDIA RTX 4060}
@@
 
-\begin{figure}[htb]
-  \centering
-  \includegraphics[width=0.75\linewidth]{figures/AutotuningSpeedupvs4060Sched_T400.png}
-  \caption{Autotuning heuristics speedup vs best 4060 schedule on T400.}
-  \label{Fig:AutotuningSpeedupvs4060Sched_T400}
-\end{figure}
+% \begin{figure}[htb]
+%   \centering
+%   \includegraphics[width=0.75\linewidth]{figures/AutotuningSpeedupvs4060Sched_T400.png}
+%   \caption{Autotuning heuristics speedup vs best 4060 schedule on T400.}
+%   \label{Fig:AutotuningSpeedupvs4060Sched_T400}
+% \end{figure}
 
@@
   very close to the best schedule found by exhaustive exploration.
-  \item The speedup of the heuristic schedule vs the best 4060 schedule is shown in Figure \ref{Fig:AutotuningSpeedupvs4060Sched}.
-  Geomean speedup range from 1.1$\times$ to 1.2$\times$.
-  \item 
+  \item The speedup of the heuristic schedule vs the best 4060 schedule when run on T400 is shown in Figure \ref{Fig:AutotuningSpeedupvs4060Sched}.
+  Geomean speedup ranges from 1.05$\times$ to 1.2$\times$. 
+  \item Unsurprisingly, there is some variation in the best schedule for a model even across NVIDIA GPUs.
 \end{itemize}
@@
   \centering
-  \includegraphics[width=0.75\linewidth]{figures/AMD_MI210_ATHeuristicVs4060Sched_speedup.png}
+  \includegraphics[width=0.75\linewidth]{figures/geomean_speedup_AMDMI2160_4060_vs_MI2160.png}
   \caption{Autotuning heuristics speedup vs best 4060 schedule on MI210.}
@@
 \end{figure}
+
+\begin{itemize}
+  \item \Treebeard{} is able to compile code to AMD GPUs. We run generated code on an AMD MI210 GPU.
+  \item None of the other systems we compare against support running on AMD GPUs. This shows the portability of \Treebeard{}.
+  \item The speedup of the heuristic schedule vs the best 4060 schedule when run on the MI210 is shown in Figure \ref{Fig:AMD_MI210_ATHeuristicVs4060Sched_speedup}.
+  \item We see that the geomean speedup over all benchmarks is 1.5$\times$ at batch size 16k.
+  \item The maximum speedup is 2$\times$ for the \op{letters} benchmark at batch size 16k.
+\end{itemize}
 

### 2024-04-17 docs/SOSP 2024/hir-mir-opts.tex
Edit overview and schedule sections

@@
-\section{\Treebeard{} Multi-Level Compiler}
+\section{Overview of \Treebeard{} Multi-Level Compilation}
 
@@
 \Treebeard{} is built using the open-source \TreebeardOLD{} infrastructure~\cite{Treebeard}.
-Since the \TreebeardOLD{} infrastructure was originally designed to target CPUs,
-significant extensions were required to support schedule-guided compilation for CPUs and GPUs. 
-These parts of the system are shown as shaded boxes in Figure \ref{Fig:CompilerStructure}.
+% Since the \TreebeardOLD{} infrastructure was originally designed to target CPUs,
+% significant extensions were required to support schedule-guided compilation for CPUs and GPUs. 
+The \TreebeardOLD{} infrastructure was originally designed to target CPUs.
+It lacks a scheduling language and does not support generating code for 
+different implementation strategies. We extend \TreebeardOLD{} significantly 
+to support schedule-guided compilation for CPUs and GPUs.
+The parts \Treebeard{} that are new or significantly different compared to \TreebeardOLD{}
+are shown as shaded boxes in Figure \ref{Fig:CompilerStructure}.
 % However, it implements several optimizations on the HIR and MIR that 
@@
 One of the performance bottlenecks in inference code targeted 
-to GPUs was that warps spent significant time being stalled. Since GPUs 
-implement scoreboarding~\cite{HennesseyPatterson}, we were able to 
+to GPUs was that warps spent significant time being stalled. We were able to 
 alleviate this bottleneck by interleaving tree walks. 

### 2024-04-17 docs/SOSP 2024/schedule.tex
Edit overview and schedule sections

@@
 \label{sec:schedule}
-As established in Section \ref{sec:motivation}, there are several 
-configurations and optimizations strategies for decision 
-tree inference. The best one to use varies significantly across
-models, batch sizes and hardware platforms. Therefore, designing
-any one hard-coded strategy is not feasible as this would 
-make portable performance impossible. To address this, we 
-design a scheduling language for \Treebeard{}. The scheduling 
+% As established in Section \ref{sec:motivation}, there are several 
+% configurations and optimizations strategies for decision 
+% tree inference. The best one to use varies significantly across
+% models, batch sizes and hardware platforms. Therefore, designing
+% any one hard-coded strategy is not feasible as this would 
+% make portable performance impossible. 
+We design a scheduling language for \Treebeard{}
+to address the problems discussed in Section \ref{sec:motivation}. The scheduling 
 language provides an abstract way to specify loop structure and 
@@
 
-% \subsection{Language Definition}
+\subsection{Language Definition}
 The core construct of \Treebeard{}'s scheduling language is an 
@@
 
+\subsection{Expressiveness of the Scheduling Language}
 \Treebeard{}'s scheduling language is expressive enough to represent a wide range of
 strategies used in existing systems. We show examples of how it can be used to 
-represent XGBoost and Tahoe's strategies. 
+represent XGBoost~\cite{XGBoost} and Tahoe's strategies~\cite{Tahoe}. 
 % Before presenting the examples, we note that \Treebeard{}'s 
@@
 
-% \subsection{The XBoost Schedule}
 XGBoost\cite{XGBoost} implements inference on the CPU by going 
@@
 set of rows. Different sets of rows are processed in parallel.
-The following schedule expresses XGBoost's strategy.
+This is expressed in \Treebeard{}'s scheduling language as:
 \begin{lstlisting}[style=c++]
@@
 
-In the \emph{direct method}, a single GPU thread walks all trees
+In the \emph{direct method}~\cite{Tahoe}, a single GPU thread walks all trees
 for a given input row. The schedule for this strategy is as follows.
@@
 
-In the \emph{shared data} strategy, a thread block walks all the trees 
+In the \emph{shared data} strategy~\cite{Tahoe}, a thread block walks all the trees 
 for a given row in parallel. 

### 2024-04-17 docs/SOSP 2024/exploring.tex
Minor edits to exploration section

@@
 \TODO{kr: next few lines need more work}
-Even if one were to search for a good schedule offline, before deploying the model on a specific hardware, additional tools are needed help search for a 
-high-performance schedule. We anticipate that even for the most complex models, one would like to spend at most a few minutes doing this search.
-We propose a set of heuristics to guide the search for a good schedule and meet this goal. We do so in two steps, by first defining a bounded search space and then pruning this search space further.
+Even if one were to search for a good schedule offline, before deploying the model 
+on a specific hardware, additional tools are needed to help search for a 
+high-performance schedule. We anticipate that even for the most complex models, 
+% one would like to spend at most a few minutes doing this search.
+spending more than a few minutes on this search is not acceptable.
+We propose a set of heuristics to guide the search for a good schedule 
+and meet this goal. These heuristics work by first defining a 
+bounded search space and then pruning this search space further
+based on the batch size and model characteristics.
+
 \subsection{Bounding the search space}
-We bound the space using a few meta parameters that together define the primitives used to construct a schedule.
-We do so while making sure that the space (at-least) covers the known strategies published in prior work.
+We bound the space using a few meta-parameters that together define 
+the primitives used to construct a schedule. We do so while making sure 
+that the space (at-least) covers the known strategies published in prior work.
 Specifically, we use 4 numeric parameters and 4 boolean parameters listed in Table \ref{tab:schedparams}. 
 %Note that even this template has a large number ($1728$) of schedules, and requires further pruning as we discuss later.
-The first three numeric parameters assign a configurable number of rows to each thread block, to each thread, and determine how trees are distributed across a specified number of threads.
-These parameters together define the loop schedule primitives to use (including the arguments to pass) from Table~\ref{Tab:LoopModifiers} and the parallelization strategy.
-The next four parameters determine the caching strategy, unroll factor{\footnote{We choose to always unroll trees completely, while its possible to partially unroll loops we did not observe any performance benefits from doing so.}} and how many trees to traverse simultaneously \TODO{Should we say, in the kernel/within a single thread?}.
-(the remaining primitives from Table~\ref{Tab:Optimizations}).
+The first three numeric parameters assign a configurable number of rows to each 
+thread block, to each thread, and determine how trees are distributed across a
+specified number of threads. These parameters together define the loop schedule 
+primitives to use (including the arguments to pass) from Table~\ref{Tab:LoopModifiers} 
+and the parallelization strategy. The next four parameters determine the caching strategy, 
+unroll factor{\footnote{We choose to always unroll trees completely, while its possible 
+to partially unroll loops we did not observe any performance benefits from doing so.}} 
+and how many trees to traverse simultaneously \TODO{Should we say, in the kernel/within
+a single thread?}.  (the remaining primitives from Table~\ref{Tab:Optimizations}).  
 The final parameter determines what reduction options to use (Table~\ref{Tab:ReductionOpts}).
@@
 
-We evaluated all the schedules in this space on several real world models and observed that there is huge variation in performance with different parameter values. 
+We evaluated all the schedules in this space on several real world models and observed that 
+there is huge variation in performance with different parameter values. 
 Figure \ref{Fig:ExecTimeDistribution} shows the distribution of normalized execution times
@@
 
-Exhaustive exploration over this bounded space ($1728$ schedules) is still very expensive, it took anywhere between thirty minutes to a few hours
-to explore the entire space for a given model. 
+Exhaustive exploration over this bounded space ($5184$ schedules) is still very expensive, taking
+anywhere between thirty minutes to a few hours to explore the entire space for a given model. 
 %Exploring the space for every model is infeasible.
 %\TODO{comment on online vs offline exploration? Set a goal for the heuristic?}
+
 \subsection{Pruning the search space}
@@
 %Performing this extensive search for every model being compiled is infeasible in practise. 
-A careful analysis of the search space reveals that certain schedules are not likely to perform well as they either do not exploit the parallelism available in the model or do not take advantage of the hardware features.
+A careful analysis of the search space reveals that certain schedules are not likely to 
+perform well as they either do not exploit the parallelism available in the model or do not take advantage of the hardware features.
 We use a combination of 3 strategies to further prune the search space.
@@
   \subsubsection*{Orthogonal parameters} We find that some parameters like reduction type are orthogonal to the rest. 
-  We therefore break the exploration into two phases, first we pick the best schedules without reduction and then evaluate reduction options on the best schedules.
+  We therefore break the exploration into two phases, first we pick the best schedules without reduction
+  optimizations and then evaluate reduction options on the best schedules.
   %We find that when a model prefers schedules with shared reduction, the same schedules 

### 2024-04-17 docs/SOSP 2024/motivation.tex
Edit motivation

@@
 
-As an example, consider a model with four trees, two complete
+Consider a model with four trees, two 
 trees of depth 1 and two of depth 2 (Figure \ref{Fig:HIRExample}).
 % We use this model as our running example throughout this section.
-We first describe a simple strategy that processes one tree at a time 
-for all input rows and unrolls all tree walks. 
-The loop over the trees is split into two loops -- one that
+We first describe a simple schedule that processes one tree at a time 
+for all input rows. The scheduling constructs are fairly intuitive. We 
+defer a detailed explanation of the scheduling language to Section~\ref{sec:scheduling}. 
+The loop over the trees is split into two -- one that
 iterates over the first two trees (Trees 1 and 2 with depth 1) and 
 the second that iterates over the last two trees (Trees 3 and 4 with
-depth 2). The \Treebeard{} schedule then unrolls the tree walks for each tree.
+depth 2). Straightforward traversal of trees involves branching. 
+One way to avoid this as done in this schedule is to unroll the 
+tree walks for each tree. 
+When the schedule specifies that the tree walks should be unrolled, 
+\Treebeard{} pads the trees so that all leaves are at the same depth.
+
 \begin{lstlisting}[style=c++]
@@
 \end{lstlisting}
-The concrete implementation of this schedule (in one of \Treebeard{}'s IRs) 
+The concrete implementation of this schedule (in \Treebeard{}'s IRs) 
 is as follows.
@@
 This schedule is ideally suited for a single-core CPU. It maximizes 
-the reuse of trees in the L1 cache and also minimizes the amount of
-branching by unrolling tree walks. However, it doesn't exploit  
+the reuse of trees in the L1 cache. However, it doesn't exploit  
 any parallelism. 
@@
 %%  and is therefore ill-suited for parallel processors.
-Further, the \op{batch} for loop can be tiled to get locality benefits.
+% Further, the \op{batch} for loop can be tiled to get locality benefits.
 
@@
 While this may work for multi-core CPUs, with massively parallel processors like GPUs,
-this strategy may not yield sufficient parallel work. Another option is to also 
-parallelize across trees. 
+this strategy may not yield sufficient parallel work. To expose more parallelism, 
+we can additionally parallelize across trees as done in the schedule below. 
 %% RG:
@@
 % the size of thread blocks and so on. 
-A possible strategy to accomplish this is encoded in the following schedule.
+% A possible strategy to accomplish this is encoded in the following schedule.
 \begin{lstlisting}[style=c++]
@@
 
-In the case of this schedule, the \op{reduce}
-operation needs special consideration.  
-%% RG
-%% to correctly generate code for this schedule, the compiler needs 
-To generate the correct code for this schedule, the compiler needs to determine
-the required synchronization and temporary storage for reductions 
-given the parallelization strategy specified. In the given example, it 
-has to identify that parallel iterations of the \op{t1} 
-loop accumulate into the same element of the \op{result} array.
-One possible solution is to rewrite the reduction so that each parallel 
-iteration accumulates into a different array element by introducing 
-a temporary buffer (\op{temp}) as follows.
-\begin{lstlisting}[style=c++]
-  float temp[2][BATCH_SIZE]
-  model = ensemble(...)
-  par.for batch = 0 to BATCH_SIZE step 1 <grid.x> {
-    par.for t1 = 0 to 2 step 1 <block.x> {
-      for t0_depth1 = 0 to 2 step 2 {
-        T = getTree(ensemble, t0_depth1 + t1)
-        treePred = walkDecisionTree(T, 
-                      input[batch]) <unrollDepth = 1>
-        reduce(temp[t1][batch], treePred)
-      }
-      for t0_depth2 = 2 to 4 step 2 {
-        T = getTree(ensemble, t0_depth2 + t1)
-        treePred = walkDecisionTree(T,
-                      input[batch]) <unrollDepth = 2>
-        reduce(temp[t1][batch], treePred) <'+', 0.0>
-      }
-    }
-    result[batch] = reduce_dimension(temp[:][batch], 0)
-  }
-\end{lstlisting}
-Here, partial results are accumulated into \op{temp} and then
-reduced across the \op{t1} dimension to get the final result.
+% In the case of this schedule, the \op{reduce}
+% operation needs special consideration.  
+% %% RG
+% %% to correctly generate code for this schedule, the compiler needs 
+% To generate the correct code for this schedule, the compiler needs to determine
+% the required synchronization and temporary storage for reductions 
+% given the parallelization strategy specified. In the given example, it 
+% has to identify that parallel iterations of the \op{t1} 
+% loop accumulate into the same element of the \op{result} array.
+% One possible solution is to rewrite the reduction so that each parallel 
+% iteration accumulates into a different array element by introducing 
+% a temporary buffer (\op{temp}) as follows.
+% \begin{lstlisting}[style=c++]
+%   float temp[2][BATCH_SIZE]
+%   model = ensemble(...)
+%   par.for batch = 0 to BATCH_SIZE step 1 <grid.x> {
+%     par.for t1 = 0 to 2 step 1 <block.x> {
+%       for t0_depth1 = 0 to 2 step 2 {
+%         T = getTree(ensemble, t0_depth1 + t1)
+%         treePred = walkDecisionTree(T, 
+%                       input[batch]) <unrollDepth = 1>
+%         reduce(temp[t1][batch], treePred)
+%       }
+%       for t0_depth2 = 2 to 4 step 2 {
+%         T = getTree(ensemble, t0_depth2 + t1)
+%         treePred = walkDecisionTree(T,
+%                       input[batch]) <unrollDepth = 2>
+%         reduce(temp[t1][batch], treePred) <'+', 0.0>
+%       }
+%     }
+%     result[batch] = reduce_dimension(temp[:][batch], 0)
+%   }
+% \end{lstlisting}
+% Here, partial results are accumulated into \op{temp} and then
+% reduced across the \op{t1} dimension to get the final result.
+Note that while unrolling helps avoid branching, it likely increases 
+the total amount of computation. Another option is to not unroll 
+and let the GPU manage the branching. The schedules with and 
+without unrolling place different constraints on the target processor,
+and the best choice depends on the characteristics of the model
+and micro-architectural features like register file size, handling 
+of branch divergence etc.~\cite{FungMicro11, SomethingElse}.
 
-As is evident from these examples, it is possible to optimize 
-the inference routine in different ways. Also, the structure of the loop 
-nest in the inference routine can get quite complex even 
-for simple schedules. Writing these routines by hand 
-is error-prone and time-consuming.  
-%% RG 
-%% We beleive 
-Hence designing a 
-scheduling language to encapsulate these strategies and a principled code 
-generator to automatically generate code guided by the schedule is the 
-best approach. We promote such an approach in this paper.
+% As is evident from these examples, it is possible to optimize 
+% the inference routine in different ways. Also, the structure of the loop 
+% nest in the inference routine can get quite complex even 
+% for simple schedules. Writing these routines by hand 
+% is error-prone and time-consuming.  
+% %% RG 
+% %% We beleive 
+% Hence designing a 
+% scheduling language to encapsulate these strategies and a principled code 
+% generator to automatically generate code guided by the schedule is the 
+% best approach. We promote such an approach in this paper.
 
@@
 
-Do different schedules and the code generated based on them matter? 
-%% RG 
-%% To further complicate matters, we 
-We find that these different strategies can have significantly different
-performance. Figures~\ref{fig:sensitivitya} and~\ref{fig:sensitivityb} show the
-variation in performance when the same schedules are used across different batch sizes 
-and different models, respectively. 
-These diagrams show that the best schedules to use vary across batch sizes and models. 
-%% RG
-% Each chart shows the performance of different schedules~-- slowdown compared to the best
-% performing schedule~-- as a heatmap. 
-The largest slowdown is 5$\times$ 
-%% RG
-when the best-performing schedule for a model is  
-%% schedules are 
-used across different batch sizes.  %% and 
-It increases to 6$\times$ when schedules are used across different models. 
-%% Therefore, no one strategy can be used across all models and batch sizes. 
-Clearly, using a single strategy across models and batch sizes leaves significant performance on the table. Hence, 
-%% 
-a system that is capable of specializing generated code for both batch 
-size and model is required for the best performance.
+To establish the importance of choosing the right schedule, we
+compare the performance of the schedules generated by \Treebeard{}
+on several real-world benchmarks. 
+Figure~\ref{fig:sensitivitya} shows the
+variation in performance when the best schedule for a given batch 
+size is used across different batch sizes for one of the models. 
+Figure~\ref{fig:sensitivityb} shows the variation when schedules 
+are used across models at a fixed batch size.
+% These diagrams show that the best schedules to use vary across batch sizes and models. 
+As can be seen, performance degrades by 2$\times$ when the best schedule for a
+smaller batch size is used for a larger batch size and vice-versa.
+Across all our benchmarks, the largest slowdown is 5$\times$.  %% and 
+The degradation is much worse when schedules are used across different models.
+In many instances (~20\%), using the best schedule for one model 
+to compile another results in a 5$\times$ slowdown.
+As we report in Section \ref{sec:results}, reusing schedules across
+different architecture also leads to significant slowdowns.
+Clearly, using a single strategy across models, batch sizes and targets
+leaves significant performance on the table.
 
+These performance considerations, coupled with 
+the importance of running ML applications on a diverse set of hardware targets,
+motivates the need for a retargetable compiler for decision tree inference.
 Building such a configurable compiler and supporting code generation for CPUs and GPUs 
-required us to solve several fundamental problems. We had to enable the
-compiler to represent and optimize reductions, deal uniformly with different
-in-memory representations of the model, design optimizations to effectively 
-use the memory hierarchy of the target processor (shared memory on GPUs and 
-cache on the CPU) and finally be able to generate target specific code. 
-%% RG
-Lastly, we propose a simple heuristic to find a high-performance schedule 
-for a given model and batch size to remove this burden from the user. 
-We defer the task of implementing an auto-tuner to future work. 
-%%
+required us to solve several fundamental problems. 
+% We had to enable the
+% compiler to represent and optimize reductions, deal uniformly with different
+% in-memory representations of the model, design optimizations to effectively 
+% use the memory hierarchy of the target processor (shared memory on GPUs and 
+% cache on the CPU) and finally be able to generate target specific code. 
+% %% RG
+% Lastly, we propose a simple heuristic to find a high-performance schedule 
+% for a given model and batch size to remove this burden from the user. 
+% We defer the task of implementing an auto-tuner to future work. 
+% %%
 The rest of the paper describes these challenges in detail and how we

### 2024-04-17 docs/SOSP 2024/results.tex
partial edit results

@@
 
-We evaluate \Treebeard{} on several machines. The first machine has an AMD Ryzen 9 7950X 16-Core processor,
-128 GB of RAM, an NVIDIA RTX 4060 GPU with 8 GB of RAM and runs Ubuntu 22.04.2 LTS with CUDA 11.5. The 
-second has an Intel Core i9-11900K (Rocket Lake) processor with 8 physical cores, 128 GB of RAM, an 
-NVIDIA T400 with 2GB of RAM and runs Ubuntu 20.04.3 LTS with CUDA 11.8. We also evaluate \Treebeard{} on 
-an AMD MI210 GPU. To establish the efficacy of \Treebeard{}, we compare it with NVIDIA RAPIDs v23.10,
-Tahoe, and XGBoost v1.7.6. We use the same benchmark models as in the \TreebeardOLD{} paper\cite{Treebeard}.
-We also evaluate \Treebeard{} on a set of randomly generated models.
+We evaluate \Treebeard{} on four different machines and compare performance against four other systems. The first machine 
+%has an AMD Ryzen 9 7950X 16-Core processor, 128 GB of RAM, 
+has an NVIDIA RTX 4060 GPU with 8 GB of RAM and runs Ubuntu 22.04.2 LTS with CUDA 11.5. The 
+second has 
+%an Intel Core i9-11900K (Rocket Lake) processor with 8 physical cores, 128 GB of RAM, 
+an NVIDIA T400 GPU with 2GB of RAM and runs Ubuntu 20.04.3 LTS with CUDA 11.8. We also evaluate \Treebeard{} on 
+an AMD MI210 GPU and an Intel Core i9-11900K CPU. 
+We compare \Treebeard{} with NVIDIA RAPIDs v23.10,
+Tahoe, XGBoost v1.7.6 and TreebeardOLD{}. We measure both kernel time and total time (including data transfer to the GPU and results back) for RAPIDs and \Treebeard{}.
+Tahoe only allows us to measure the kernel time since it is written as an executable that performs inference repeatedly on the same data that is transferred to the GPU once.
 
-\subsection{Performance on Random Models}
-\begin{itemize}
-  \item Generated a set of over 700 random models with varying depths (6, 7, 8), number of trees (100 - 1000 in steps of 100), 
-  and number of features (powers of 2 between 8 to 1024 inclusive).
-  \item Compared the kernel time speedup of \Treebeard{} vs RAPIDs on the RTX 4060. The schedule used was the one picked by the auto-tuner. 
-  \item Speedups range between 1.5$\times$ and 8$\times$. \Treebeard{} outperforms RAPIDs on all models and batch sizes tested.
-  \item Figure \ref{fig:randomModels4060} shows the speedup for models of depth 8 and batch size 512 and depth 6 and batch size 4096.
-  Trends for other batch sizes and depths are similar.
-\end{itemize}
+We use two sets of benchmarks to perform the comparison.
+We use 8 real-world models trained on data from the Intel Machine Learning Benchmark suite~\cite{}. These models were also used in the \TreebeardOLD{} paper\cite{Treebeard}.
+To enable more exhaustive evaluation we trained 700 random models with varying number of trees (100---1000) and features (powers of two in the range 8---1024).
+Each tree has leaves at depths 2 to a maximum depth of 6, 7 or 8.
 
+\subsection{Performance comparisons on RTX 4060}
+We evaluated \Treebeard{} and RAPIDS on all 700 random models with varying input batch sizes, on multiple hardware.   
+Figure~\ref{fig:randomModels4060} shows two representative samples of the results, at batch sizes 512 and 4096 on all models. 
+Each plot is a 3D histogram of the speedup of \Treebeard{} over RAPIDS, with the x-axis representing the number of trees, the y-axis the number of features and the z-axis the speedup. 
+The schedule used by \Treebeard{} was automatically chosen using the algorithm described in Section~\ref{sec:exploring}. \TODO{kr: report exploration time?}.
+While the exact trends at different batch sizes vary it can be seen that \Treebeard{} consistently outperforms RAPIDS by $1.5-8\times$. 
+%\Treebeard{} consistently outperforms RAPIDS on all models and batch sizes tested.
+% \begin{itemize}
+%   \item Generated a set of over 700 random models with varying depths (6, 7, 8), number of trees (100 - 1000 in steps of 100), 
+%   and number of features (powers of 2 between 8 to 1024 inclusive).
+%   \item Compared the kernel time speedup of \Treebeard{} vs RAPIDs on the RTX 4060. The schedule used was the one picked by the auto-tuner. 
+%   \item Speedups range between 1.5$\times$ and 8$\times$. \Treebeard{} outperforms RAPIDs on all models and batch sizes tested.
+%   \item Figure \ref{fig:randomModels4060} shows the speedup for models of depth 8 and batch size 512 and depth 6 and batch size 4096.
+%   Trends for other batch sizes and depths are similar.
+% \end{itemize}
+We also did more detailed performance analysis on the real-world benchmarks.
+Figure~\ref{Fig:TBvsRAPIDsTahoe_4060_Speedup} shows the geomean speedup of \Treebeard{} over RAPIDs and Tahoe at different batch sizes.
+We do not report results for XGBoost since the speedups are an order of magnitude higher and too large to fit on the same graph.
+The plot has lines for kernel time and total time speedup over RAPIDS and kernel time speedup over Tahoe.
+As can been seen \Treebeard{} is uniformly faster than Tahoe by $2-3\times$ at all batch sizes.
+The relative performance of RAPIDs improves with batch size, but \Treebeard{} is still faster by $1.5-2\times$ evan at very high batch sizes.
+The plot also shows that the speedups are significant even when data needs to be transferred to the GPU and results need to be transferred back.
 \begin{figure*}[ht]
@@
     \subcaptionbox*{}{\includegraphics[width=\textwidth]{figures/RandomModels/kernel_speedup_b512_depth8.png}}
-    \caption{Batch size 512, depth 8}
+    \caption{Batch size 512}
+
   \end{subfigure}
@@
     \subcaptionbox*{}{\includegraphics[width=\textwidth]{figures/RandomModels/kernel_speedup_b4096_depth6.png}}
-    \caption{Batch size 4096, depth 6}
+    \caption{Batch size 4096}
   \end{subfigure}
   \hfill
-  \caption{\label{fig:randomModels4060}\Treebeard{} vs RAPIDs Kernel Time Speedup on NVIDIA RTX 4060 for several randomly generated models.}
+  \caption{\label{fig:randomModels4060}\Treebeard{} vs RAPIDs Kernel Time Speedup for several randomly generated models.}
 \end{figure*}
-
+\begin{comment}
 \subsection{Comparison with RAPIDs, Tahoe and XGBoost}
@@
 \end{itemize}
-
+\end{comment}
 \begin{figure}[htb]

### 2024-04-18 docs/SOSP 2024/results.tex
results

@@
 
-We evaluate \Treebeard{} on four different machines and compare performance against four other systems. The first machine 
-%has an AMD Ryzen 9 7950X 16-Core processor, 128 GB of RAM, 
-has an NVIDIA RTX 4060 GPU with 8 GB of RAM and runs Ubuntu 22.04.2 LTS with CUDA 11.5. The 
-second has 
-%an Intel Core i9-11900K (Rocket Lake) processor with 8 physical cores, 128 GB of RAM, 
-an NVIDIA T400 GPU with 2GB of RAM and runs Ubuntu 20.04.3 LTS with CUDA 11.8. We also evaluate \Treebeard{} on 
-an AMD MI210 GPU and an Intel Core i9-11900K CPU. 
-We compare \Treebeard{} with NVIDIA RAPIDs v23.10,
-Tahoe, XGBoost v1.7.6 and TreebeardOLD{}. We measure both kernel time and total time (including data transfer to the GPU and results back) for RAPIDs and \Treebeard{}.
+We evaluate \Treebeard{} on four different machines, an NVIDIA RTX 4060 GPU (8GB RAM, CUDA 11.5), an NVIDIA T400 GPU (2GB RAM, CUDA 11.5), an AMD MI210 GPU (16?GB RAM, ROCm 4.5?) and an Intel Core i9-11900K CPU (16?
+virtual cores, 128?
+GB RAM).
+We compare \Treebeard{} with four other systems, NVIDIA RAPIDs v23.10, Tahoe, XGBoost v1.7.6 and \TreebeardOLD{} CPU.
+We measure both kernel time and total time (including data transfer to the GPU and results back) for RAPIDs and \Treebeard{}.
 Tahoe only allows us to measure the kernel time since it is written as an executable that performs inference repeatedly on the same data that is transferred to the GPU once.
@@
 We use two sets of benchmarks to perform the comparison.
-We use 8 real-world models trained on data from the Intel Machine Learning Benchmark suite~\cite{}. These models were also used in the \TreebeardOLD{} paper\cite{Treebeard}.
-To enable more exhaustive evaluation we trained 700 random models with varying number of trees (100---1000) and features (powers of two in the range 8---1024).
+We use 8 real-world models trained on data from the Intel Machine Learning Benchmark suite~\cite{}. These models were also used to \TreebeardOLD{}\cite{Treebeard}.
+To enable more exhaustive evaluation we generated 700 random models with varying number of trees (100---1000) and features (powers of two in the range 8---1024).
 Each tree has leaves at depths 2 to a maximum depth of 6, 7 or 8.
+%The schedule used by \Treebeard{} was automatically chosen by the algorithm described in Section~\ref{sec:exploring}.
+\subsection{Performance comparisons on NVIDIA GPUs}
+\subsubsection*{Real-world models}
+We performed a detailed performance comparison between \Treebeard{}, Tahoe, RAPIDS and XGBoost on the real-world benchmarks.
+Figure~\ref{Fig:TBvsRAPIDsTahoe_4060_Speedup} shows the geomean speedup of \Treebeard{} at different batch sizes on RTX 4060. 
+We do not show results for XGBoost since the speedups are an order of magnitude higher and too large to fit on the same graph.  
+  The plot has lines for kernel time and total time speedup over RAPIDS and kernel time speedup over Tahoe.
+As can been seen \Treebeard{} is uniformly faster than Tahoe{\footnote{Tahoe does not support multiclass models and to enable comparison wet ran the multiclass models (\op{covtype} and \op{letters}) as regression models.
+We also noticed that some variants of Tahoe produce wrong results (as reported by its own tests) for
+  \op{letters} and \op{year}. In these cases, we pick the time of the fastest variant that gives the correct results.
+}} by $2-3\times$ at all batch sizes. 
+Compared to RAPIDS, \Treebeard{} is about $4\times$ faster at batch size 512.
+The relative performance of RAPIDs improves with batch size, but \Treebeard{} is still faster by $1.5-2\times$ evan at very high ($16K$) batch sizes.
+The plot also shows that the speedups are significant even when data needs to be transferred to the GPU and results need to be transferred back.
+They are lower than kernel time speedup as both systems have a constant additional transfer overhead.
 
-\subsection{Performance comparisons on RTX 4060}
-We evaluated \Treebeard{} and RAPIDS on all 700 random models with varying input batch sizes, on multiple hardware.   
-Figure~\ref{fig:randomModels4060} shows two representative samples of the results, at batch sizes 512 and 4096 on all models. 
-Each plot is a 3D histogram of the speedup of \Treebeard{} over RAPIDS, with the x-axis representing the number of trees, the y-axis the number of features and the z-axis the speedup. 
-The schedule used by \Treebeard{} was automatically chosen using the algorithm described in Section~\ref{sec:exploring}. \TODO{kr: report exploration time?}.
-While the exact trends at different batch sizes vary it can be seen that \Treebeard{} consistently outperforms RAPIDS by $1.5-8\times$. 
-%\Treebeard{} consistently outperforms RAPIDS on all models and batch sizes tested.
-% \begin{itemize}
-%   \item Generated a set of over 700 random models with varying depths (6, 7, 8), number of trees (100 - 1000 in steps of 100), 
-%   and number of features (powers of 2 between 8 to 1024 inclusive).
-%   \item Compared the kernel time speedup of \Treebeard{} vs RAPIDs on the RTX 4060. The schedule used was the one picked by the auto-tuner. 
-%   \item Speedups range between 1.5$\times$ and 8$\times$. \Treebeard{} outperforms RAPIDs on all models and batch sizes tested.
-%   \item Figure \ref{fig:randomModels4060} shows the speedup for models of depth 8 and batch size 512 and depth 6 and batch size 4096.
-%   Trends for other batch sizes and depths are similar.
-% \end{itemize}
-We also did more detailed performance analysis on the real-world benchmarks.
-Figure~\ref{Fig:TBvsRAPIDsTahoe_4060_Speedup} shows the geomean speedup of \Treebeard{} over RAPIDs and Tahoe at different batch sizes.
-We do not report results for XGBoost since the speedups are an order of magnitude higher and too large to fit on the same graph.
-The plot has lines for kernel time and total time speedup over RAPIDS and kernel time speedup over Tahoe.
-As can been seen \Treebeard{} is uniformly faster than Tahoe by $2-3\times$ at all batch sizes.
-The relative performance of RAPIDs improves with batch size, but \Treebeard{} is still faster by $1.5-2\times$ evan at very high batch sizes.
-The plot also shows that the speedups are significant even when data needs to be transferred to the GPU and results need to be transferred back.
+Figure~\ref{Fig:KernelTimeIndividualBenchmarks4060} shows per benchmark results at two different batch sizes.
+\Treebeard{} is able to find better schedules than both RAPIDS and Tahoe on each of the benchmarks.
+It consistently outperforms both systems, achieving speedups in the range $1.1-12\times$, with about half the benchmarks achieving a speedup of $2\times$ or more  over both baselines at batch size 8192.
 \begin{figure*}[ht]
@@
 \end{figure*}
-\begin{comment}
-\subsection{Comparison with RAPIDs, Tahoe and XGBoost}
-\begin{itemize}
-  \item Measured both the kernel time and total time (time including transferring data to the 
-  GPU and results back) for RAPIDs and \Treebeard{}.
-  \item Tahoe only allows us to measure the kernel time since it is written as an executable 
-  that performs inference repeatedly on the same data that is transferred to the GPU once.
-  \item Tahoe does not support multiclass models. We just ran the multiclass models (\op{covtype} and \op{letters})
-  as regression models for the comparison. Tahoe also gives wrong results (as reported by its own tests) for 
-  \op{letters} and \op{year}. In these cases, we pick the time of the fastest variant that gives the correct results.
-  \item Compared the kernel time and total time speedup of \Treebeard{} vs RAPIDs on the RTX 4060 and T400.
-  \item Compared the total time speedup of \Treebeard{} vs XGBoost on the RTX 4060.
-  \item The schedule used was the one picked by the auto-tuner.
-  \item \Treebeard{} outperforms RAPIDs at all batch sizes as shown in Figure \ref{Fig:TBvsRAPIDsTahoe_4060_Speedup}. 
-  \item \Treebeard{} outperforms Tahoe on all models and batch sizes tested. Individual benchmark speedups range between 1.1$\times$ and 16$\times$.
-  \item \Treebeard{} is faster than XGBoost by more than an order of magnitude. Results are not shown because the speedups don't fit on the same graph.
-  \item Figure \ref{Fig:TBvsRAPIDs_4060_TotalTimeSpeedup} shows that \Treebeard{} offers substantial speedup over RAPIDs even when 
-  data needs to be transferred to the GPU and results need to be transferred back.
-  \item Figure \ref{Fig:TBvsRAPIDsTahoe_T400_Speedup} shows that these speedups are also observed on the T400 thus showing 
-  that \Treebeard{} offers portable performance across different GPUs.
-  \item Figure \ref{Fig:KernelTimeIndividualBenchmarks4060} shows that \Treebeard{} outperforms RAPIDs and Tahoe on 
-  individual benchmarks on the RTX 4060 at small (1024) and large (8192) batch sizes. These batch sizes 
-  require different schedules, but the \Treebeard{} auto-tuning heuristic is able to find them.
-  % \item Figure \ref{Fig:TotalTimeIndividualBenchmarks4060} compares the total time of \Treebeard{} vs RAPIDs on individual benchmarks. 
-  % It shows that even with the overhead of data transfers, \Treebeard{} offers significant speedup. The graph in Figure \ref{Fig:TotalTimeIndividualBenchmarks4060} 
-  % shows the breakup of the total time into kernel time and data transfer time.
-\end{itemize}
-\end{comment}
+
+\subsubsection*{Synthetic models}
+To establish that \Treebeard{} can consistently find betters schedules, we performed an exhaustive comparison between \Treebeard{} and RAPIDS on all 700 randomly generated models.   
+Figure~\ref{fig:randomModels4060} shows two representative samples of the results, at batch sizes 512 and 4096 on all models, on RTX 4060. 
+Each plot is a 3D histogram of the speedup of \Treebeard{} over RAPIDS, with the x-axis representing the number of trees, the y-axis the number of features and the z-axis the speedup. 
+While the exact trends at different batch sizes vary it can be seen that \Treebeard{} consistently outperforms RAPIDS by $1.5-8\times$. 
+%\Treebeard{} consistently outperforms RAPIDS on all models and batch sizes tested.
+% \begin{itemize}
+%   \item Generated a set of over 700 random models with varying depths (6, 7, 8), number of trees (100 - 1000 in steps of 100), 
+%   and number of features (powers of 2 between 8 to 1024 inclusive).
+%   \item Compared the kernel time speedup of \Treebeard{} vs RAPIDs on the RTX 4060. The schedule used was the one picked by the auto-tuner. 
+%   \item Speedups range between 1.5$\times$ and 8$\times$. \Treebeard{} outperforms RAPIDs on all models and batch sizes tested.
+%   \item Figure \ref{fig:randomModels4060} shows the speedup for models of depth 8 and batch size 512 and depth 6 and batch size 4096.
+%   Trends for other batch sizes and depths are similar.
+% \end{itemize}
+\subsubsection*{Comparison on T400}
+Figure~\ref{Fig:TBvsRAPIDsTahoe_T400_Speedup} shows that Treebeard{} is also able to generate better inferences routines than both Tahoe and RAPIDS on a different class of NVIDIA GPUs.
+The trends are similar to what we saw on RTX 4060.
+% \subsection{Comparison with RAPIDs, Tahoe and XGBoost}
+% \begin{itemize}
+%   \item Measured both the kernel time and total time (time including transferring data to the 
+%   GPU and results back) for RAPIDs and \Treebeard{}.
+%   \item Tahoe only allows us to measure the kernel time since it is written as an executable 
+%   that performs inference repeatedly on the same data that is transferred to the GPU once.
+%   \item Tahoe does not support multiclass models. We just ran the multiclass models (\op{covtype} and \op{letters})
+%   as regression models for the comparison. Tahoe also gives wrong results (as reported by its own tests) for 
+%   \op{letters} and \op{year}. In these cases, we pick the time of the fastest variant that gives the correct results.
+%   \item Compared the kernel time and total time speedup of \Treebeard{} vs RAPIDs on the RTX 4060 and T400.
+%   \item Compared the total time speedup of \Treebeard{} vs XGBoost on the RTX 4060.
+%   \item The schedule used was the one picked by the auto-tuner.
+%   \item \Treebeard{} outperforms RAPIDs at all batch sizes as shown in Figure \ref{Fig:TBvsRAPIDsTahoe_4060_Speedup}. 
+%   \item \Treebeard{} outperforms Tahoe on all models and batch sizes tested. Individual benchmark speedups range between 1.1$\times$ and 16$\times$.
+%   \item \Treebeard{} is faster than XGBoost by more than an order of magnitude. Results are not shown because the speedups don't fit on the same graph.
+%   \item Figure \ref{Fig:TBvsRAPIDs_4060_TotalTimeSpeedup} shows that \Treebeard{} offers substantial speedup over RAPIDs even when 
+%   data needs to be transferred to the GPU and results need to be transferred back.
+%   \item Figure \ref{Fig:TBvsRAPIDsTahoe_T400_Speedup} shows that these speedups are also observed on the T400 thus showing 
+%   that \Treebeard{} offers portable performance across different GPUs.
+%   \item Figure \ref{Fig:KernelTimeIndividualBenchmarks4060} shows that \Treebeard{} outperforms RAPIDs and Tahoe on 
+%   individual benchmarks on the RTX 4060 at small (1024) and large (8192) batch sizes. These batch sizes 
+%   require different schedules, but the \Treebeard{} auto-tuning heuristic is able to find them.
+%   % \item Figure \ref{Fig:TotalTimeIndividualBenchmarks4060} compares the total time of \Treebeard{} vs RAPIDs on individual benchmarks. 
+%   % It shows that even with the overhead of data transfers, \Treebeard{} offers significant speedup. The graph in Figure \ref{Fig:TotalTimeIndividualBenchmarks4060} 
+%   % shows the breakup of the total time into kernel time and data transfer time.
+% \end{itemize}
 \begin{figure}[htb]
@@
 
-\subsection{Autotuning Heuristic}
+\subsection{Schedule sensitivity}
+\subsubsection*{Quality of schedules generated by the heuristic}
+We compare the schedule found by the schedule exploration heuristic on the RTX 4060 with the best schedule found by extensive exploration.
+Figure~\ref{Fig:HeuristicVsFullExplore_Speedup} shows that the heuristic is able to find schedules that are very close to the best schedule found by exhaustive exploration.
+\subsubsection*{Portability of schedules}
+Section~\ref{sec:motivation} already showed that even on the same GPU the best schedule can vary with change in model or batch size.
+We further evaluate the portability of schedules across different NVIDIA GPUs by comparing the speedup of the heuristic schedule vs the best 4060 schedule when run on the T400.
+Geomean speedup ranges from 1.05$\times$ to 1.2$\times$.
+Unsurprisingly, there is some variation in the best schedule for a model even across NVIDIA GPUs.
 \begin{figure}[htb]
@@
 
-\begin{itemize}
-  \item We compare the schedule found by the schedule exploration heuristic on the RTX 4060 with the best schedule found by
-  extensive exploration. Figure \ref{Fig:HeuristicVsFullExplore_Speedup} shows that the heuristic is able to find schedules that are
-  very close to the best schedule found by exhaustive exploration.
-  \item The speedup of the heuristic schedule vs the best 4060 schedule when run on T400 is shown in Figure \ref{Fig:AutotuningSpeedupvs4060Sched}.
-  Geomean speedup ranges from 1.05$\times$ to 1.2$\times$. 
-  \item Unsurprisingly, there is some variation in the best schedule for a model even across NVIDIA GPUs.
-\end{itemize}
+% \begin{itemize}
+%   \item We compare the schedule found by the schedule exploration heuristic on the RTX 4060 with the best schedule found by
+%   extensive exploration. Figure \ref{Fig:HeuristicVsFullExplore_Speedup} shows that the heuristic is able to find schedules that are
+%   very close to the best schedule found by exhaustive exploration.
+%   \item The speedup of the heuristic schedule vs the best 4060 schedule when run on T400 is shown in Figure \ref{Fig:AutotuningSpeedupvs4060Sched}.
+%   Geomean speedup ranges from 1.05$\times$ to 1.2$\times$. 
+%   \item Unsurprisingly, there is some variation in the best schedule for a model even across NVIDIA GPUs.
+% \end{itemize}
 
@@
 \end{figure}
-
-\begin{itemize}
-  \item \Treebeard{} is able to compile code to AMD GPUs. We run generated code on an AMD MI210 GPU.
-  \item None of the other systems we compare against support running on AMD GPUs. This shows the portability of \Treebeard{}.
-  \item The speedup of the heuristic schedule vs the best 4060 schedule when run on the MI210 is shown in Figure \ref{Fig:AMD_MI210_ATHeuristicVs4060Sched_speedup}.
-  \item We see that the geomean speedup over all benchmarks is 1.5$\times$ at batch size 16k.
-  \item The maximum speedup is 2$\times$ for the \op{letters} benchmark at batch size 16k.
-\end{itemize}
+We demonstrate that \Treebeard{} is also able to compile code to AMD GPUs and achieve portable performance.
+We run generated code on an AMD MI210 GPU.
+None of the other systems we compare against support running on AMD GPUs.
+We also report the speedup of the heuristic schedule on AMD GPUs vs the best 4060 schedule when run on the MI210 in Figure \ref{Fig:AMD_MI210_ATHeuristicVs4060Sched_speedup}.
+We see that the geomean speedup over all benchmarks is 1.5$\times$ at batch size 16k.
+The maximum speedup is 2$\times$ for the \op{letters} benchmark at batch size 16k.
+% \begin{itemize}
+%   \item \Treebeard{} is able to compile code to AMD GPUs. We run generated code on an AMD MI210 GPU.
+%   \item None of the other systems we compare against support running on AMD GPUs. This shows the portability of \Treebeard{}.
+%   \item The speedup of the heuristic schedule vs the best 4060 schedule when run on the MI210 is shown in Figure \ref{Fig:AMD_MI210_ATHeuristicVs4060Sched_speedup}.
+%   \item We see that the geomean speedup over all benchmarks is 1.5$\times$ at batch size 16k.
+%   \item The maximum speedup is 2$\times$ for the \op{letters} benchmark at batch size 16k.
+% \end{itemize}
 

### 2024-04-18 docs/SOSP 2024/results.tex
Edit results

@@
 
-We evaluate \Treebeard{} on four different machines, an NVIDIA RTX 4060 GPU (8GB RAM, CUDA 11.5), an NVIDIA T400 GPU (2GB RAM, CUDA 11.5), an AMD MI210 GPU (16?GB RAM, ROCm 4.5?) and an Intel Core i9-11900K CPU (16?
-virtual cores, 128?
-GB RAM).
-We compare \Treebeard{} with four other systems, NVIDIA RAPIDs v23.10, Tahoe, XGBoost v1.7.6 and \TreebeardOLD{} CPU.
-We measure both kernel time and total time (including data transfer to the GPU and results back) for RAPIDs and \Treebeard{}.
-Tahoe only allows us to measure the kernel time since it is written as an executable that performs inference repeatedly on the same data that is transferred to the GPU once.
+We evaluate \Treebeard{} on four different machines and compare performance against four other systems. The first machine 
+%has an AMD Ryzen 9 7950X 16-Core processor, 128 GB of RAM, 
+has an NVIDIA RTX 4060 GPU with 8 GB of RAM and runs Ubuntu 22.04.2 LTS with CUDA 11.5. The 
+second has 
+%an Intel Core i9-11900K (Rocket Lake) processor with 8 physical cores, 128 GB of RAM, 
+an NVIDIA T400 GPU with 2GB of RAM and runs Ubuntu 20.04.3 LTS with CUDA 11.8. We also evaluate \Treebeard{} on 
+an AMD MI210 GPU and an Intel Core i9-11900K CPU. 
+We compare \Treebeard{} with NVIDIA RAPIDs v23.10,
+Tahoe, XGBoost v1.7.6 and \TreebeardOLD{}. We measure both kernel time and total time (including data transfer 
+to the GPU and results back) for RAPIDs and \Treebeard{}.
+Tahoe only allows us to measure the kernel time since it is written as an executable that performs inference
+repeatedly on the same data that is transferred to the GPU once\footnote{Tahoe also does not support 
+multi-class models. We let Tahoe run the multi-class models in our benchmarks (\op{covtype} and \op{letters})
+as regression models. Also, we find that for a few models, some variants of Tahoe produce wrong results. 
+In such cases, we pick the time for the fastest correct variant.}.
 
 We use two sets of benchmarks to perform the comparison.
-We use 8 real-world models trained on data from the Intel Machine Learning Benchmark suite~\cite{}. These models were also used to \TreebeardOLD{}\cite{Treebeard}.
-To enable more exhaustive evaluation we generated 700 random models with varying number of trees (100---1000) and features (powers of two in the range 8---1024).
-Each tree has leaves at depths 2 to a maximum depth of 6, 7 or 8.
-%The schedule used by \Treebeard{} was automatically chosen by the algorithm described in Section~\ref{sec:exploring}.
-\subsection{Performance comparisons on NVIDIA GPUs}
-\subsubsection*{Real-world models}
-We performed a detailed performance comparison between \Treebeard{}, Tahoe, RAPIDS and XGBoost on the real-world benchmarks.
-Figure~\ref{Fig:TBvsRAPIDsTahoe_4060_Speedup} shows the geomean speedup of \Treebeard{} at different batch sizes on RTX 4060. 
-We do not show results for XGBoost since the speedups are an order of magnitude higher and too large to fit on the same graph.  
+We use 8 real-world models trained on data from the 
+Intel Machine Learning Benchmark suite~\cite{MLBenchmarks}. These models were 
+also used in the \TreebeardOLD{} paper\cite{Treebeard}.
+To enable more exhaustive evaluation we trained 700 random models with varying number 
+of trees (100---1000) and features (powers of two in the range 8---1024). Each tree has leaves at depths 2 to a maximum depth of 6, 7 or 8.
+
+\subsection{Performance comparisons on RTX 4060}
+We evaluated \Treebeard{} and RAPIDS on all 700 random models with varying input batch sizes, on multiple hardware.   
+Figure~\ref{fig:randomModels4060} shows two representative samples of the results, at batch sizes 512 and 4096 on all models. 
+Each plot is a 3D histogram of the speedup of \Treebeard{} over RAPIDS, with the x-axis representing the number of trees, the y-axis the number of features and the z-axis the speedup. 
+The schedule used by \Treebeard{} was automatically chosen using the algorithm described in Section~\ref{sec:exploring}. \TODO{kr: report exploration time?}.
+While the exact trends at different batch sizes vary it can be seen that \Treebeard{} consistently outperforms RAPIDS by $1.5-8\times$. 
+%\Treebeard{} consistently outperforms RAPIDS on all models and batch sizes tested.
+% \begin{itemize}
+%   \item Generated a set of over 700 random models with varying depths (6, 7, 8), number of trees (100 - 1000 in steps of 100), 
+%   and number of features (powers of 2 between 8 to 1024 inclusive).
+%   \item Compared the kernel time speedup of \Treebeard{} vs RAPIDs on the RTX 4060. The schedule used was the one picked by the auto-tuner. 
+%   \item Speedups range between 1.5$\times$ and 8$\times$. \Treebeard{} outperforms RAPIDs on all models and batch sizes tested.
+%   \item Figure \ref{fig:randomModels4060} shows the speedup for models of depth 8 and batch size 512 and depth 6 and batch size 4096.
+%   Trends for other batch sizes and depths are similar.
+% \end{itemize}
+We also did more detailed performance analysis on the real-world benchmarks.
+Figure~\ref{Fig:TBvsRAPIDsTahoe_4060_Speedup} shows the geomean speedup of \Treebeard{} over RAPIDs 
+and Tahoe at different batch sizes. We do not report results for XGBoost since the speedups are an 
+order of magnitude higher and too large to fit on the same graph.
 The plot has lines for kernel time and total time speedup over RAPIDS and kernel time speedup over Tahoe. 
-As can been seen \Treebeard{} is uniformly faster than Tahoe{\footnote{Tahoe does not support multiclass models and to enable comparison wet ran the multiclass models (\op{covtype} and \op{letters}) as regression models.
-We also noticed that some variants of Tahoe produce wrong results (as reported by its own tests) for
-  \op{letters} and \op{year}. In these cases, we pick the time of the fastest variant that gives the correct results.
-}} by $2-3\times$ at all batch sizes. 
-Compared to RAPIDS, \Treebeard{} is about $4\times$ faster at batch size 512.
-The relative performance of RAPIDs improves with batch size, but \Treebeard{} is still faster by $1.5-2\times$ evan at very high ($16K$) batch sizes.
-The plot also shows that the speedups are significant even when data needs to be transferred to the GPU and results need to be transferred back.
-They are lower than kernel time speedup as both systems have a constant additional transfer overhead.
-
-Figure~\ref{Fig:KernelTimeIndividualBenchmarks4060} shows per benchmark results at two different batch sizes.
-\Treebeard{} is able to find better schedules than both RAPIDS and Tahoe on each of the benchmarks.
-It consistently outperforms both systems, achieving speedups in the range $1.1-12\times$, with about half the benchmarks achieving a speedup of $2\times$ or more  over both baselines at batch size 8192.
+As can been seen, \Treebeard{} is uniformly faster than Tahoe by $2-3\times$ at all batch sizes. 
+The relative performance of RAPIDs improves with batch size, but \Treebeard{} is still faster 
+by close to $3\times$ even at very high batch sizes. The plot also shows that the speedups are significant 
+even when data needs to be transferred to the GPU and results need to be transferred back.
+Figure \ref{Fig:KernelTimeIndividualBenchmarks4060} shows the kernel time speedup 
+of \Treebeard{} on individual benchmarks at two different batch sizes (1024 and 8192). This plot 
+shows that \Treebeard{} is consistently faster than RAPIDs and Tahoe over a wide range of benchmarks
+and batch sizes. These different batch sizes require different schedules, but the \Treebeard{} 
+schedule exploration heuristic is able to find them.
 \begin{figure*}[ht]
@@
 
-\subsubsection*{Synthetic models}
-To establish that \Treebeard{} can consistently find betters schedules, we performed an exhaustive comparison between \Treebeard{} and RAPIDS on all 700 randomly generated models.   
-Figure~\ref{fig:randomModels4060} shows two representative samples of the results, at batch sizes 512 and 4096 on all models, on RTX 4060. 
-Each plot is a 3D histogram of the speedup of \Treebeard{} over RAPIDS, with the x-axis representing the number of trees, the y-axis the number of features and the z-axis the speedup. 
-While the exact trends at different batch sizes vary it can be seen that \Treebeard{} consistently outperforms RAPIDS by $1.5-8\times$. 
-%\Treebeard{} consistently outperforms RAPIDS on all models and batch sizes tested.
-% \begin{itemize}
-%   \item Generated a set of over 700 random models with varying depths (6, 7, 8), number of trees (100 - 1000 in steps of 100), 
-%   and number of features (powers of 2 between 8 to 1024 inclusive).
-%   \item Compared the kernel time speedup of \Treebeard{} vs RAPIDs on the RTX 4060. The schedule used was the one picked by the auto-tuner. 
-%   \item Speedups range between 1.5$\times$ and 8$\times$. \Treebeard{} outperforms RAPIDs on all models and batch sizes tested.
-%   \item Figure \ref{fig:randomModels4060} shows the speedup for models of depth 8 and batch size 512 and depth 6 and batch size 4096.
-%   Trends for other batch sizes and depths are similar.
-% \end{itemize}
-\subsubsection*{Comparison on T400}
-Figure~\ref{Fig:TBvsRAPIDsTahoe_T400_Speedup} shows that Treebeard{} is also able to generate better inferences routines than both Tahoe and RAPIDS on a different class of NVIDIA GPUs.
-The trends are similar to what we saw on RTX 4060.
+% \begin{comment}
 % \subsection{Comparison with RAPIDs, Tahoe and XGBoost}
@@
 % \end{itemize}
+% \end{comment}
+
 \begin{figure}[htb]
@@
 
-\begin{figure}[htb]
-  \centering
-  \includegraphics[width=0.75\linewidth]{figures/geomean_speedup_T400_kernel_time.png}
-  \caption{\Treebeard{} vs RAPIDs and Tahoe Kernel Time Speedup on NVIDIA T400.}
-  \label{Fig:TBvsRAPIDsTahoe_T400_Speedup}
-\end{figure}
-
 % \begin{figure*}[ht]
@@
 
-\subsection{Schedule sensitivity}
-\subsubsection*{Quality of schedules generated by the heuristic}
-We compare the schedule found by the schedule exploration heuristic on the RTX 4060 with the best schedule found by extensive exploration.
-Figure~\ref{Fig:HeuristicVsFullExplore_Speedup} shows that the heuristic is able to find schedules that are very close to the best schedule found by exhaustive exploration.
-\subsubsection*{Portability of schedules}
-Section~\ref{sec:motivation} already showed that even on the same GPU the best schedule can vary with change in model or batch size.
-We further evaluate the portability of schedules across different NVIDIA GPUs by comparing the speedup of the heuristic schedule vs the best 4060 schedule when run on the T400.
-Geomean speedup ranges from 1.05$\times$ to 1.2$\times$.
-Unsurprisingly, there is some variation in the best schedule for a model even across NVIDIA GPUs.
+\subsection{Schedule Exploration Heuristic}
+We evaluated the schedule exploration heuristic described in Section \ref{sec:exploring} on several fronts.
+First, we determine whether the heuristic is able to find schedules that are close to the best 
+schedule found by exhaustive exploration. We then run experiments to test whether the heuristic works well 
+across different GPUs. Figure \ref{Fig:AutotuningSpeedupvs4060Sched} shows the speedup of the schedule 
+found by the heuristic vs the best schedule found by exhaustive exploration on the RTX 4060. The line 
+for RTX 4060 shows that the heuristic is able to find schedules that are very close to the best schedule.
+For T400, the heuristic is able to find schedules that are slightly better. The geomean speedup ranges 
+from 1.05$\times$ to 1.2$\times$ with the maximum speedup being $2\times$ (\op{epsilon} at batch size 512).
+Finally, we measure the improvement in exploration time that the heuristic provides 
+compared to exhaustive exploration. Figure \ref{Fig:HeuristicVsFullExplore_Speedup} shows that the heuristic
+is consistently close to two orders of magnitude faster than exhaustive exploration. 
+These results show that our heuristic is able to quickly find schedules that are close to the best schedule.
+
 \begin{figure}[htb]
@@
   \includegraphics[width=0.75\linewidth]{figures/geomean_speedup_T400_4060_vs_T400.png}
-  \caption{Autotuning heuristics speedup vs best 4060 schedule on NVIDIA RTX 4060}
+  \caption{Geomean speedup of schedule found by exploration heuristic vs best 4060 schedule on NVIDIA RTX 4060 and NVIDIA T400 across batch sizes}
   \label{Fig:AutotuningSpeedupvs4060Sched}
@@
   \includegraphics[width=0.75\linewidth]{figures/geomean_speedup_4060_full_exp_vs_at.png}
-  \caption{Autotuning heuristic compile time speedup vs full schedule exploration.}
+  \caption{Schedule exploration heuristic exploration time speedup vs full schedule exploration}
   \label{Fig:HeuristicVsFullExplore_Speedup}
@@
 
-\subsection{AMD GPU}
+\subsection{Performance on Other GPUs}
+\begin{figure}[htb]
+  \centering
+  \includegraphics[width=0.75\linewidth]{figures/geomean_speedup_T400_kernel_time.png}
+  \caption{\Treebeard{} vs RAPIDs and Tahoe kernel time and total time speedup on NVIDIA T400.}
+  \label{Fig:TBvsRAPIDsTahoe_T400_Speedup}
+\end{figure}
+
 \begin{figure}[htb]
@@
   \includegraphics[width=0.75\linewidth]{figures/geomean_speedup_AMDMI2160_4060_vs_MI2160.png}
-  \caption{Autotuning heuristics speedup vs best 4060 schedule on MI210.}
+  \caption{Speedup of schedule exploration heuristic vs best 4060 schedule on MI210}
   \label{Fig:AMD_MI210_ATHeuristicVs4060Sched_speedup}
 \end{figure}
-We demonstrate that \Treebeard{} is also able to compile code to AMD GPUs and achieve portable performance.
-We run generated code on an AMD MI210 GPU.
-None of the other systems we compare against support running on AMD GPUs.
-We also report the speedup of the heuristic schedule on AMD GPUs vs the best 4060 schedule when run on the MI210 in Figure \ref{Fig:AMD_MI210_ATHeuristicVs4060Sched_speedup}.
-We see that the geomean speedup over all benchmarks is 1.5$\times$ at batch size 16k.
+
+To test the portability of \Treebeard{}'s techniques, we compare the performance 
+of \Treebeard{} on the T400 GPU with RAPIDs and Tahoe. \Treebeard{} is also 
+able to generate high-performance code for an AMD MI210 GPU. Figure \ref{Fig:TBvsRAPIDsTahoe_T400_Speedup}
+shows the speedup of code generated by \Treebeard{} vs RAPIDs and Tahoe on the T400.
+We observe that the trends are similar to trends on the RTX 4060. Even on the smaller T400 GPU, \Treebeard{} is able to outperform RAPIDs
+and Tahoe consistently across batch sizes with geomean speedups at a batch size ranging from $1.5\times$ to $3.3\times$.
+We also observe that \Treebeard{} significantly outperforms RAPIDs even when transfer times are 
+taken into account.
+
+Unlike other systems we compare to, \Treebeard{} is not limited to NVIDIA GPUs. We 
+run our benchmark suite on an AMD MI210 GPU and find that \Treebeard{} is able to
+generate competitive code for this GPU. Figure \ref{Fig:AMD_MI210_ATHeuristicVs4060Sched_speedup}
+shows the speedup of the schedule found by our exploration heuristic compared to the best 
+4060 schedule when run on the MI210. We use this baseline since none of the other systems
+we compare against support running on AMD GPUs. We find that the geomean speedup over all benchmarks is 1.5$\times$ at batch size 16k.
 The maximum speedup is 2$\times$ for the \op{letters} benchmark at batch size 16k.
+
 % \begin{itemize}
@@
 \subsection{CPU Improvements}
-\begin{itemize}
-  \item The ability of \Treebeard{} to parallelize across both trees and rows has a significant impact on CPU performance.
-  \item We run experiments on the Intel Core i9-11900K processor to evaluate the impact of parallelizing across trees with small batch sizes. 
-  \item At batch size 32, we find that the average speedup over all 8 models is 2.2$\times$ with a max speedup of 5$\times$.
-  \item At batch size 64, we find that the average speedup over all 8 models is 1.1$\times$ with a max speedup of 2$\times$. 
-  However, we find that 2 of the 8 models show slowdowns in this case. Again, this highlights the need for schedule exploration on CPUs.
-  \item For small batch sizes, parallelizing across rows does not work great since there is limited reuse of trees in L1 cache.
+To further establish the cross target nature of the techniques described in this paper,
+we compare the performance of \Treebeard{} with that of \TreebeardOLD{} on machine with an  
+Intel Core i9-11900K processor and 128GB of RAM. Specifically, we evaluate the impact of parallelizing
+across trees. We find that the ability to parallelize across trees improves performance 
+significantly at small batch sizes. At batch size 32, we find that the geomean speedup over 
+all benchmark models is 2.2$\times$ with a max speedup of 5$\times$. At batch size 64, the average speedup
+is 1.1$\times$ with a max speedup of 2$\times$. At batch size 32, parallelizing across trees is faster 
+for all models and at batch size 64 the \TreebeardOLD{} schedule that parallelizes across 
+rows is faster for only 2 of the 8 models. For small batch sizes, parallelizing across rows does 
+not offer the best performance as there is limited reuse of trees in L1 cache.
 Also, the amount of work per thread is very small leading to high overheads. 
-  \item Parallelizing across trees is more effective since there is more reuse in L1 cache and the amount of work per thread is higher.  
-\end{itemize}
+Parallelizing across trees fixes both these problems.
+
+% \begin{itemize}
+%   \item The ability of \Treebeard{} to parallelize across both trees and rows has a significant impact on CPU performance.
+%   \item We run experiments on the Intel Core i9-11900K processor to evaluate the impact of parallelizing across trees with small batch sizes. 
+%   \item At batch size 32, we find that the average speedup over all 8 models is 2.2$\times$ with a max speedup of 5$\times$.
+%   \item At batch size 64, we find that the average speedup over all 8 models is 1.1$\times$ with a max speedup of 2$\times$. 
+%   However, we find that 2 of the 8 models show slowdowns in this case. Again, this highlights the need for schedule exploration on CPUs.
+%   \item For small batch sizes, parallelizing across rows does not work great since there is limited reuse of trees in L1 cache.
+%   Also, the amount of work per thread is very small leading to high overheads. 
+%   \item Parallelizing across trees is more effective since there is more reuse in L1 cache and the amount of work per thread is higher.  
+% \end{itemize}
+
+Overall, our evaluation shows that \Treebeard{} is able to efficiently generate 
+high-performance code for processors ranging from NVIDIA and AMD GPUs to Intel CPUs.
+On all platforms and models that we tested on, \Treebeard{} significantly outperforms
+state-of-the-art systems like RAPIDs, Tahoe and \TreebeardOLD{}.

### 2024-04-18 docs/SOSP 2024/results.tex

@@
 schedule exploration heuristic is able to find them.
+We evaluate \Treebeard{} on four different machines, an NVIDIA RTX 4060 GPU (8GB RAM, CUDA 11.5), an NVIDIA T400 GPU (2GB RAM, CUDA 11.5), an AMD MI210 GPU (16?GB RAM, ROCm 4.5?) and an Intel Core i9-11900K CPU (16?
+virtual cores, 128?
+GB RAM).
+We compare \Treebeard{} with four other systems, NVIDIA RAPIDs v23.10, Tahoe, XGBoost v1.7.6 and \TreebeardOLD{} CPU.
+We measure both kernel time and total time (including data transfer to the GPU and results back) for RAPIDs and \Treebeard{}.
+Tahoe only allows us to measure the kernel time since it is written as an executable that performs inference repeatedly on the same data that is transferred to the GPU once.
+
+We use two sets of benchmarks to perform the comparison.
+We use 8 real-world models trained on data from the Intel Machine Learning Benchmark suite~\cite{}. These models were also used to \TreebeardOLD{}\cite{Treebeard}.
+To enable more exhaustive evaluation we generated 700 random models with varying number of trees (100---1000) and features (powers of two in the range 8---1024).
+Each tree has leaves at depths 2 to a maximum depth of 6, 7 or 8.
+%The schedule used by \Treebeard{} was automatically chosen by the algorithm described in Section~\ref{sec:exploring}.
+\subsection{Performance comparisons on NVIDIA GPUs}
+\subsubsection*{Real-world models}
+We performed a detailed performance comparison between \Treebeard{}, Tahoe, RAPIDS and XGBoost on the real-world benchmarks.
+Figure~\ref{Fig:TBvsRAPIDsTahoe_4060_Speedup} shows the geomean speedup of \Treebeard{} at different batch sizes on RTX 4060. 
+We do not show results for XGBoost since the speedups are an order of magnitude higher and too large to fit on the same graph.  
+  The plot has lines for kernel time and total time speedup over RAPIDS and kernel time speedup over Tahoe.
+As can been seen \Treebeard{} is uniformly faster than Tahoe{\footnote{Tahoe does not support multiclass models and to enable comparison wet ran the multiclass models (\op{covtype} and \op{letters}) as regression models.
+We also noticed that some variants of Tahoe produce wrong results (as reported by its own tests) for
+  \op{letters} and \op{year}. In these cases, we pick the time of the fastest variant that gives the correct results.
+}} by $2-3\times$ at all batch sizes. 
+Compared to RAPIDS, \Treebeard{} is about $4\times$ faster at batch size 512.
+The relative performance of RAPIDs improves with batch size, but \Treebeard{} is still faster by $1.5-2\times$ evan at very high ($16K$) batch sizes.
+The plot also shows that the speedups are significant even when data needs to be transferred to the GPU and results need to be transferred back.
+They are lower than kernel time speedup as both systems have a constant additional transfer overhead.
+
+Figure~\ref{Fig:KernelTimeIndividualBenchmarks4060} shows per benchmark results at two different batch sizes.
+\Treebeard{} is able to find better schedules than both RAPIDS and Tahoe on each of the benchmarks.
+It consistently outperforms both systems, achieving speedups in the range $1.1-12\times$, with about half the benchmarks achieving a speedup of $2\times$ or more  over both baselines at batch size 8192.
 \begin{figure*}[ht]
@@
 % \begin{comment}
+\subsubsection*{Synthetic models}
+To establish that \Treebeard{} can consistently find betters schedules, we performed an exhaustive comparison between \Treebeard{} and RAPIDS on all 700 randomly generated models.   
+Figure~\ref{fig:randomModels4060} shows two representative samples of the results, at batch sizes 512 and 4096 on all models, on RTX 4060. 
+Each plot is a 3D histogram of the speedup of \Treebeard{} over RAPIDS, with the x-axis representing the number of trees, the y-axis the number of features and the z-axis the speedup. 
+While the exact trends at different batch sizes vary it can be seen that \Treebeard{} consistently outperforms RAPIDS by $1.5-8\times$. 
+%\Treebeard{} consistently outperforms RAPIDS on all models and batch sizes tested.
+% \begin{itemize}
+%   \item Generated a set of over 700 random models with varying depths (6, 7, 8), number of trees (100 - 1000 in steps of 100), 
+%   and number of features (powers of 2 between 8 to 1024 inclusive).
+%   \item Compared the kernel time speedup of \Treebeard{} vs RAPIDs on the RTX 4060. The schedule used was the one picked by the auto-tuner. 
+%   \item Speedups range between 1.5$\times$ and 8$\times$. \Treebeard{} outperforms RAPIDs on all models and batch sizes tested.
+%   \item Figure \ref{fig:randomModels4060} shows the speedup for models of depth 8 and batch size 512 and depth 6 and batch size 4096.
+%   Trends for other batch sizes and depths are similar.
+% \end{itemize}
+\subsubsection*{Comparison on T400}
+Figure~\ref{Fig:TBvsRAPIDsTahoe_T400_Speedup} shows that Treebeard{} is also able to generate better inferences routines than both Tahoe and RAPIDS on a different class of NVIDIA GPUs.
+The trends are similar to what we saw on RTX 4060.
 % \subsection{Comparison with RAPIDs, Tahoe and XGBoost}
@@
 
+\subsection{Schedule sensitivity}
+\subsubsection*{Quality of schedules generated by the heuristic}
+We compare the schedule found by the schedule exploration heuristic on the RTX 4060 with the best schedule found by extensive exploration.
+Figure~\ref{Fig:HeuristicVsFullExplore_Speedup} shows that the heuristic is able to find schedules that are very close to the best schedule found by exhaustive exploration.
+\subsubsection*{Portability of schedules}
+Section~\ref{sec:motivation} already showed that even on the same GPU the best schedule can vary with change in model or batch size.
+We further evaluate the portability of schedules across different NVIDIA GPUs by comparing the speedup of the heuristic schedule vs the best 4060 schedule when run on the T400.
+Geomean speedup ranges from 1.05$\times$ to 1.2$\times$.
+Unsurprisingly, there is some variation in the best schedule for a model even across NVIDIA GPUs.
 \begin{figure}[htb]
@@
 
+We demonstrate that \Treebeard{} is also able to compile code to AMD GPUs and achieve portable performance.
+We run generated code on an AMD MI210 GPU.
+None of the other systems we compare against support running on AMD GPUs.
+We also report the speedup of the heuristic schedule on AMD GPUs vs the best 4060 schedule when run on the MI210 in Figure \ref{Fig:AMD_MI210_ATHeuristicVs4060Sched_speedup}.
+We see that the geomean speedup over all benchmarks is 1.5$\times$ at batch size 16k.
+The maximum speedup is 2$\times$ for the \op{letters} benchmark at batch size 16k.
 % \begin{itemize}

### 2024-04-18 docs/SOSP 2024/results.tex
Edit result section

--- 2024-04-18 docs/SOSP 2024/results.tex
@@
 
-We evaluate \Treebeard{} on four different machines, an NVIDIA RTX 4060 GPU (8GB RAM, CUDA 11.5), an NVIDIA T400 GPU (2GB RAM, CUDA 11.5), an AMD MI210 GPU (16?GB RAM, ROCm 4.5?) and an Intel Core i9-11900K CPU (16?
-virtual cores, 128?
-GB RAM).
-We compare \Treebeard{} with four other systems, NVIDIA RAPIDs v23.10, Tahoe, XGBoost v1.7.6 and \TreebeardOLD{} CPU.
-We measure both kernel time and total time (including data transfer to the GPU and results back) for RAPIDs and \Treebeard{}.
-Tahoe only allows us to measure the kernel time since it is written as an executable that performs inference repeatedly on the same data that is transferred to the GPU once.
+We evaluate \Treebeard{} on four different target processors, an NVIDIA RTX 4060 GPU (8GB RAM, CUDA 11.5),
+an NVIDIA T400 GPU (2GB RAM, CUDA 11.5), an AMD MI210 GPU (64GB RAM, ROCm 4.5?) and an 
+Intel Core i9-11900K CPU (16 virtual cores, 128 GB RAM). We compare \Treebeard{} with 
+four other systems, NVIDIA RAPIDs v23.10, Tahoe, XGBoost v1.7.6 and \TreebeardOLD{} CPU. 
+We measure both kernel time and total time (including data transfer to the GPU and results back) 
+for RAPIDs and \Treebeard{}. Tahoe only allows us to measure the kernel time since it is written
+as an executable that performs inference repeatedly on the same data that is transferred to the GPU once.
 
 We use two sets of benchmarks to perform the comparison.
-We use 8 real-world models trained on data from the Intel Machine Learning Benchmark suite~\cite{}. These models were also used to \TreebeardOLD{}\cite{Treebeard}.
-To enable more exhaustive evaluation we generated 700 random models with varying number of trees (100---1000) and features (powers of two in the range 8---1024).
+We use 8 real-world models trained on data from the Intel Machine 
+Learning Benchmark suite~\cite{MLBenchmarks}. These models were also
+used to evaluate \TreebeardOLD{}\cite{Treebeard}.
+To enable more exhaustive evaluation we generated 700 random models with
+varying number of trees (100---1000) and features (powers of two in the range 8---1024). 
 Each tree has leaves at depths 2 to a maximum depth of 6, 7 or 8.
-%The schedule used by \Treebeard{} was automatically chosen by the algorithm described in Section~\ref{sec:exploring}.
+
 \subsection{Performance comparisons on NVIDIA GPUs}
@@
 It consistently outperforms both systems, achieving speedups in the range $1.1-12\times$, with about half the benchmarks achieving a speedup of $2\times$ or more  over both baselines at batch size 8192.
-\begin{figure*}[ht]
-  \centering
-  \begin{subfigure}[b]{.45\textwidth}
-    \subcaptionbox*{}{\includegraphics[width=\textwidth]{figures/RandomModels/kernel_speedup_b512_depth8.png}}
-    \caption{Batch size 512}
+\begin{figure}[htb]
+  \begin{minipage}[t]{.475\linewidth}
+    \includegraphics[width=\linewidth]{figures/RandomModels/kernel_speedup_b512_depth8.png}
+    %\caption{Batch size 512}
+  \end{minipage}
+  \begin{minipage}[t]{.475\linewidth}
+    \includegraphics[width=\linewidth]{figures/RandomModels/kernel_speedup_b4096_depth6.png}
+    %\caption{Batch size 4096}
+  \end{minipage}
+  \caption{\label{fig:randomModels4060}\Treebeard{} vs RAPIDs Kernel Time Speedup for several randomly generated models.}
+\end{figure}
 
-  \end{subfigure}
-  \begin{subfigure}[b]{.45\textwidth}
-    \subcaptionbox*{}{\includegraphics[width=\textwidth]{figures/RandomModels/kernel_speedup_b4096_depth6.png}}
-    \caption{Batch size 4096}
-  \end{subfigure}
-  \hfill
-  \caption{\label{fig:randomModels4060}\Treebeard{} vs RAPIDs Kernel Time Speedup for several randomly generated models.}
-\end{figure*}
-
+% \begin{comment}
 \subsubsection*{Synthetic models}
@@
 \subsubsection*{Comparison on T400}
-Figure~\ref{Fig:TBvsRAPIDsTahoe_T400_Speedup} shows that Treebeard{} is also able to generate better inferences routines than both Tahoe and RAPIDS on a different class of NVIDIA GPUs.
-The trends are similar to what we saw on RTX 4060.
+To test the portability of \Treebeard{}'s techniques, we compare the performance 
+of \Treebeard{} on the T400 GPU with RAPIDs and Tahoe. Figure \ref{Fig:TBvsRAPIDsTahoe_T400_Speedup}
+shows the speedup of code generated by \Treebeard{} vs RAPIDs and Tahoe on the T400.
+Even on the smaller T400 GPU, \Treebeard{} is able to outperform RAPIDs
+and Tahoe consistently. We observe similar trends to those on the RTX 4060.
+
 % \subsection{Comparison with RAPIDs, Tahoe and XGBoost}
@@
 
-% \begin{figure}[htb]
-%   \centering
-%   \includegraphics[width=0.75\linewidth]{figures/geomean_speedup_4060_total_time.png}
-%   \caption{\Treebeard{} vs RAPIDs Total Time Speedup on NVIDIA RTX 4060.}
-%   \label{Fig:TBvsRAPIDs_4060_TotalTimeSpeedup}
-% \end{figure}
+\begin{figure}[htb]
+  \centering
+  \includegraphics[width=0.75\linewidth]{figures/geomean_speedup_T400_kernel_time.png}
+  \caption{\Treebeard{} vs RAPIDs and Tahoe kernel time and total time speedup on NVIDIA T400.}
+  \label{Fig:TBvsRAPIDsTahoe_T400_Speedup}
+\end{figure}
 
@@
 \end{figure*}
-
-\begin{figure}[htb]
-  \centering
-  \includegraphics[width=0.75\linewidth]{figures/geomean_speedup_T400_kernel_time.png}
-  \caption{\Treebeard{} vs RAPIDs and Tahoe Kernel Time Speedup on NVIDIA T400.}
-  \label{Fig:TBvsRAPIDsTahoe_T400_Speedup}
-\end{figure}
 
@@
 
-\subsection{Schedule sensitivity}
-\subsubsection*{Quality of schedules generated by the heuristic}
-We compare the schedule found by the schedule exploration heuristic on the RTX 4060 with the best schedule found by extensive exploration.
-Figure~\ref{Fig:HeuristicVsFullExplore_Speedup} shows that the heuristic is able to find schedules that are very close to the best schedule found by exhaustive exploration.
-\subsubsection*{Portability of schedules}
-Section~\ref{sec:motivation} already showed that even on the same GPU the best schedule can vary with change in model or batch size.
-We further evaluate the portability of schedules across different NVIDIA GPUs by comparing the speedup of the heuristic schedule vs the best 4060 schedule when run on the T400.
-Geomean speedup ranges from 1.05$\times$ to 1.2$\times$.
-Unsurprisingly, there is some variation in the best schedule for a model even across NVIDIA GPUs.
+\subsection{Performance on AMD GPUs}
+
+While existing systems only support NVIDIA GPUs, we demonstrate that \Treebeard{} 
+is able to generate competitive code for AMD GPUs. Our objective here is not to compare directly 
+between AMD and NVIDIA GPUs, but we do find that as the MI210 is a more powerful GPU, 
+it achieves better inference times on most benchmarks at large batch sizes ($\geq8$k).
+
+\subsection{Schedule Exploration Heuristic}
+We evaluated the schedule exploration heuristic described in Section \ref{sec:exploring} on several fronts.
+Figure \ref{Fig:AutotuningSpeedupvs4060Sched} compares the best schedule found by exhaustive exploration on the RTX 4060 
+with the schedule found by the exploration heuristic on RTX 4060, T400 and AMD MI210.
+% First, we determine whether the heuristic is able to find schedules that are close to the best 
+% schedule found by exhaustive exploration. Figure \ref{Fig:AutotuningSpeedupvs4060Sched} shows the speedup of the schedule 
+% found by the heuristic vs the best schedule found by exhaustive exploration on the RTX 4060.
+With this we want to establish two things described below.
+\begin{enumerate}
+\item \textbf{\emph{Quality of Heuristic Schedules:}}
+As can be seen from the plot  
+for RTX 4060 the heuristic is able to find schedules that are very close to the best schedule (well 
+within 5\%).
+\item \textbf{\emph{Schedule Sensitivity Across GPUs:}}
+The remaining two lines show that the best schedule varies across target processors.
+For T400, the heuristic is able to find schedules that are 1.05$\times$ to 1.2$\times$ better 
+with the maximum difference being $2\times$ (\op{epsilon} at batch size 512).
+We find that there is a much larger variation in performance 
+between the schedules \Treebeard{} finds on the MI210 compared to the best 4060 schedules. For example, 
+the geomean speedup over all benchmarks is 1.5$\times$ at batch size 16k and 
+The maximum speedup is 2$\times$ for the \op{letters} benchmark at batch size 16k. 
+\end{enumerate}
+
+\subsubsection*{Exploration Time}
+Finally, we measure the improvement in exploration time that the heuristic provides 
+compared to exhaustive exploration. Figure \ref{Fig:HeuristicVsFullExplore_Speedup} shows that the heuristic
+is consistently close to two orders of magnitude faster than exhaustive exploration. 
+The exploration time ranges between $6$ and $167$ seconds for the heuristic with a mean of $28.7$ seconds.
+These results show that our heuristic is able to quickly find schedules that are close to the best schedule.
+
 \begin{figure}[htb]
@@
 
-% \begin{figure}[htb]
-%   \centering
-%   \includegraphics[width=0.75\linewidth]{figures/AutotuningSpeedupvs4060Sched_T400.png}
-%   \caption{Autotuning heuristics speedup vs best 4060 schedule on T400.}
-%   \label{Fig:AutotuningSpeedupvs4060Sched_T400}
-% \end{figure}
+\begin{figure}[htb]
+  \centering
+  \includegraphics[width=0.75\linewidth]{figures/geomean_speedup_AMDMI2160_4060_vs_MI2160.png}
+  \caption{Speedup of schedule exploration heuristic vs best 4060 schedule on MI210}
+  \label{Fig:AMD_MI210_ATHeuristicVs4060Sched_speedup}
+\end{figure}
 
@@
 
-\subsection{AMD GPU}
-\begin{figure}[htb]
-  \centering
-  \includegraphics[width=0.75\linewidth]{figures/geomean_speedup_AMDMI2160_4060_vs_MI2160.png}
-  \caption{Autotuning heuristics speedup vs best 4060 schedule on MI210.}
-  \label{Fig:AMD_MI210_ATHeuristicVs4060Sched_speedup}
-\end{figure}
-We demonstrate that \Treebeard{} is also able to compile code to AMD GPUs and achieve portable performance.
-We run generated code on an AMD MI210 GPU.
-None of the other systems we compare against support running on AMD GPUs.
-We also report the speedup of the heuristic schedule on AMD GPUs vs the best 4060 schedule when run on the MI210 in Figure \ref{Fig:AMD_MI210_ATHeuristicVs4060Sched_speedup}.
-We see that the geomean speedup over all benchmarks is 1.5$\times$ at batch size 16k.
-The maximum speedup is 2$\times$ for the \op{letters} benchmark at batch size 16k.
 % \begin{itemize}
@@
 \subsection{CPU Improvements}
-\begin{itemize}
-  \item The ability of \Treebeard{} to parallelize across both trees and rows has a significant impact on CPU performance.
-  \item We run experiments on the Intel Core i9-11900K processor to evaluate the impact of parallelizing across trees with small batch sizes. 
-  \item At batch size 32, we find that the average speedup over all 8 models is 2.2$\times$ with a max speedup of 5$\times$.
-  \item At batch size 64, we find that the average speedup over all 8 models is 1.1$\times$ with a max speedup of 2$\times$. 
-  However, we find that 2 of the 8 models show slowdowns in this case. Again, this highlights the need for schedule exploration on CPUs.
-  \item For small batch sizes, parallelizing across rows does not work great since there is limited reuse of trees in L1 cache.
+The enhancements made to the compiler enable \Treebeard{} to explore additional schedules on the CPU
+than \TreebeardOLD{}. 
+In particular, we find that the ability to parallelize across trees improves performance 
+significantly at small batch sizes. At batch size 32, we find that the geomean speedup over 
+all benchmark models is 2.2$\times$ with a max speedup of 5$\times$. At batch size 64, the average speedup
+is 1.1$\times$ with a max speedup of 2$\times$. At batch size 32, parallelizing across trees is faster 
+for all models and at batch size 64 the \TreebeardOLD{} schedule that parallelizes across 
+rows is faster for only 2 of the 8 models. For small batch sizes, parallelizing across rows does 
+not offer the best performance as there is limited reuse of trees in L1 cache.
 Also, the amount of work per thread is very small leading to high overheads. 
-  \item Parallelizing across trees is more effective since there is more reuse in L1 cache and the amount of work per thread is higher.  
-\end{itemize}
+Parallelizing across trees fixes both these problems.
+
+% \begin{itemize}
+%   \item The ability of \Treebeard{} to parallelize across both trees and rows has a significant impact on CPU performance.
+%   \item We run experiments on the Intel Core i9-11900K processor to evaluate the impact of parallelizing across trees with small batch sizes. 
+%   \item At batch size 32, we find that the average speedup over all 8 models is 2.2$\times$ with a max speedup of 5$\times$.
+%   \item At batch size 64, we find that the average speedup over all 8 models is 1.1$\times$ with a max speedup of 2$\times$. 
+%   However, we find that 2 of the 8 models show slowdowns in this case. Again, this highlights the need for schedule exploration on CPUs.
+%   \item For small batch sizes, parallelizing across rows does not work great since there is limited reuse of trees in L1 cache.
+%   Also, the amount of work per thread is very small leading to high overheads. 
+%   \item Parallelizing across trees is more effective since there is more reuse in L1 cache and the amount of work per thread is higher.  
+% \end{itemize}
+
+Overall, our evaluation shows that \Treebeard{} is able to efficiently generate 
+high-performance code for processors ranging from NVIDIA and AMD GPUs to Intel CPUs.
+On all platforms and models that we tested on, \Treebeard{} significantly outperforms
+state-of-the-art systems like RAPIDs, Tahoe and \TreebeardOLD{}.

### 2024-04-18 docs/SOSP 2024/motivation.tex
Minor edit to motivation
Put sensitivity figures next to each other

@@
 the second that iterates over the last two trees (Trees 3 and 4 with
-depth 2). Straightforward traversal of trees involves branching. 
+depth 2). Straightforward traversal of trees requires a \op{while} loop
+and involves branching to check if a leaf has been reached. 
 One way to avoid this as done in this schedule is to unroll the 
@@
 
-\begin{figure}[]
+\begin{figure}[htb]
+  \begin{subfigure}[t]{.45\linewidth}
     \centering
-  \begin{subfigure}[b]{.3\textwidth}
-    \subcaptionbox*{}{\includegraphics[width=\textwidth]{figures/batch_sensitivity_covtype.png}}
-    \caption{\label{fig:sensitivitya} Batch sensitivity for \op{covtype}. Each point shows the slowdown when best schedule for the 
-    x-axis batch size is used for the y-axis batch size.}
+    \includegraphics[width=\linewidth]{figures/batch_sensitivity_covtype.png}
+    \caption{\label{fig:sensitivitya}\op{covtype} batch sensitivity}
   \end{subfigure}
-  % \hfill  % Add horizontal space between subfigures
-  \begin{subfigure}[b]{.3\textwidth}
-    \subcaptionbox*{}{\includegraphics[width=\textwidth]{figures/model_sensitivity_4096.png}}
-    \caption{\label{fig:sensitivityb} Model sensitivity for batch size 4096. Each point shows the slowdown when best schedule for the
-    x-axis model is used for the y-axis model.}
+  \begin{subfigure}[t]{.45\linewidth}
+    \centering
+    \includegraphics[width=\linewidth]{figures/model_sensitivity_4096.png}
+    \caption{\label{fig:sensitivityb}Model sensitivity at batch size 4096}
 \end{subfigure}
+  \caption{Batch and model sensitivity plots. Each point shows the slowdown when best schedule for the
+          x-axis batch size (model) is used for the y-axis batch size (model).}
 \end{figure}
+
+% \begin{figure}[htb]
+%   \begin{minipage}[t]{.475\linewidth}
+%     \subcaptionbox*{}{\includegraphics[width=\linewidth]{figures/batch_sensitivity_covtype.png}}
+%     \caption{\label{fig:sensitivitya}Batch sensitivity for \op{covtype}}
+%   \end{minipage}
+%   \begin{minipage}[t]{.475\linewidth}
+%     \subcaptionbox*{}{\includegraphics[width=\linewidth]{figures/model_sensitivity_4096.png}}
+%     \caption{\label{fig:sensitivityb}Model sensitivity at batch size 4096}
+%   \end{minipage}
+%   \caption{Batch and model sensitivity plots. Each point shows the slowdown when best schedule for the
+%           x-axis batch size (model) is used for the y-axis batch size (model).}
+% \end{figure}
 

### 2024-04-18 docs/SOSP 2024/results.tex
1. Update results captions and images 2. Merge schedule heuristic vs 4060 schedule graph

@@
   \centering
-  \includegraphics[width=0.75\linewidth]{figures/geomean_speedup_T400_4060_vs_T400.png}
-  \caption{Geomean speedup of schedule found by exploration heuristic vs best 4060 schedule on NVIDIA RTX 4060 and NVIDIA T400 across batch sizes}
+  \includegraphics[width=0.75\linewidth]{figures/geomean_speedup_T400_4060_vs_T400_vs_MI2160.png}
+  \caption{Geomean speedup of schedule found by exploration heuristic vs best 4060 schedule on NVIDIA RTX 4060, NVIDIA T400 and AMD MI2160 across batch sizes}
   \label{Fig:AutotuningSpeedupvs4060Sched}
-\end{figure}
-
-\begin{figure}[htb]
-  \centering
-  \includegraphics[width=0.75\linewidth]{figures/geomean_speedup_AMDMI2160_4060_vs_MI2160.png}
-  \caption{Speedup of schedule exploration heuristic vs best 4060 schedule on MI210}
-  \label{Fig:AMD_MI210_ATHeuristicVs4060Sched_speedup}
 \end{figure}
@@
 %   \item None of the other systems we compare against support running on AMD GPUs. This shows the portability of \Treebeard{}.
-%   \item The speedup of the heuristic schedule vs the best 4060 schedule when run on the MI210 is shown in Figure \ref{Fig:AMD_MI210_ATHeuristicVs4060Sched_speedup}.
+%   \item The speedup of the heuristic schedule vs the best 4060 schedule when run on the MI210 is shown in Figure \ref{Fig:AutotuningSpeedupvs4060Sched}.
 %   \item We see that the geomean speedup over all benchmarks is 1.5$\times$ at batch size 16k.

### 2024-04-18 results/python/analyze_gpu_speedups_results.py

@@ -214,15 +214,12 @@ line_graph_plotter.plot_geomean_speedups_over_batch_size(results_t400_df, 'TBKer
 line_graph_plotter.plot_geomean_speedups_over_batch_size(results_t400_df, 'TB(AT)', ['RAPIDS(Total)'], ['RAPIDS(total)'])
 line_graph_plotter.save_plot()
 
-line_graph_plotter = LineGraphPlotter('T400', '4060_vs_T400', lower_lim=1)
+line_graph_plotter = LineGraphPlotter('T400', '4060_vs_T400_vs_MI2160', lower_lim=1)
 line_graph_plotter.plot_geomean_speedups_over_batch_size(results_t400_df, 'TBKernel(AT)', ['TBKernel(4060)'], ['Scheduling heuristic(T400) vs Best 4060 schedule'])
 line_graph_plotter.plot_geomean_speedups_over_batch_size(results_4060_df, 'TBKernel(AT)', ['TBKernel'], ['Scheduling heuristic(4060) vs Best 4060 schedule'])
+line_graph_plotter.plot_geomean_speedups_over_batch_size(results_amdmi2160_df, 'TBKernel(MI2160)', ['TBKernel(4060)'], ['Scheduling heuristic(MI2160) vs Best 4060 schedule'])
 line_graph_plotter.save_plot()
 
 line_graph_plotter = LineGraphPlotter('4060', 'full_exp_vs_at', lower_lim=1, upper_lim=110)
-line_graph_plotter.plot_geomean_speedups_over_batch_size(fullexp_vs_at_df, 'AT', ['FullExplore'], [' Auto-Tuning vs Full Explore'])
-line_graph_plotter.save_plot()
-
-line_graph_plotter = LineGraphPlotter('AMDMI2160', '4060_vs_MI2160', lower_lim=1, upper_lim=1.6, round_up=False)
-line_graph_plotter.plot_geomean_speedups_over_batch_size(results_amdmi2160_df, 'TBKernel(MI2160)', ['TBKernel(4060)'], ['Scheduling heuristic(MI2160) vs Best 4060 schedule'])
+line_graph_plotter.plot_geomean_speedups_over_batch_size(fullexp_vs_at_df, 'AT', ['FullExplore'], ['Scheduling heuristic vs Full Exploration'])
 line_graph_plotter.save_plot()

### 2024-04-18 results/python/analyze_gpu_speedups_results.py
Update graph legeneds and patterns on bar graph

@@ -103,8 +103,8 @@ def plot_bar_graph_speedups_for_batch_size(df, batch_size, kwargs=None):
     gap = BAR_LABEL_FONT_SIZE * 0.3
     pos = np.array([float(2 * i * width + i * gap) for i in range(len(bar_names))])
 
-    ax.bar(pos, rapids_speedup, width, label='RAPIDS')
-    ax.bar(pos+width, tahoe_speedup, width, label='Tahoe')
+    ax.bar(pos, rapids_speedup, width, label='Speedup of SilvanForge vs RAPIDS', hatch='//////')
+    ax.bar(pos+width, tahoe_speedup, width, label='Speedup of SilvanForge vs Tahoe', hatch='......')
 
     # enable grid
     # ax.grid(True, which='both', axis='y')
@@ -112,7 +112,7 @@ def plot_bar_graph_speedups_for_batch_size(df, batch_size, kwargs=None):
     ax.set_ylim(0, max(max(rapids_speedup), max(tahoe_speedup)) + 1)
 
     # Add labels and title
-    ax.set_xlabel('Benchmark', fontsize=LABEL_FONT_SIZE)
+    # ax.set_xlabel('Benchmark', fontsize=LABEL_FONT_SIZE)
     ax.set_ylabel('Speedup', fontsize=LABEL_FONT_SIZE)
     # ax.set_title(f'Speedups for Batch Size {batch_size} (4060)')
     ax.set_xticks(pos + width / 2)
@@ -204,14 +204,14 @@ plot_bar_graph_speedups_for_batch_size(results_4060_df, 1024)
 plot_bar_graph_speedups_for_batch_size(results_4060_df, 8192)
 
 # plot geomean speedup over batch sizes
-line_graph_plotter = LineGraphPlotter('4060', 'kernel_time_total_time')
-line_graph_plotter.plot_geomean_speedups_over_batch_size(results_4060_df, 'TBKernel(AT)', ['RAPIDS(kernel)', 'Tahoe(kernel)'], ['RAPIDS(kernel)', 'Tahoe(kernel)'])
-line_graph_plotter.plot_geomean_speedups_over_batch_size(results_4060_df, 'TB (AT)', ['RAPIDS (Total)'], ['RAPIDS(total)'])
[…]

### 2024-04-18 results/python/analyze_schedule_sensitivity_data.py
Align graphs

@@ -62,6 +62,7 @@ def plot_batch_sensitivity_for_single_benchmark(df_depth, benchmark_name):
     # plt.show()
     # plt.title(f"{benchmark_name} Batch Sensitivity")
     # plt.colorbar(im, label='tb_kernel_speedup')
+    plt.tight_layout()
     plt.savefig(f'batch_sensitivity_{benchmark_name}.png')
     plt.close('all')
 

### 2024-04-18 docs/SOSP 2024/reduction.tex
Significantly shorted reduction section

@@
 multi-dimensional array (represented by an MLIR \op{memref}).
-The following example shows how the \op{reduce} op can be used to 
-sum up the elements of an array in parallel.
+Consider the following \Treebeard{} schedule.
+In our example, \op{N\_t} is the number of trees 
+and \op{batch\_size} is the batch size. The schedule tiles the 
+tree loop and parallelizes the resulting outer loop.
+\begin{lstlisting}[style=c++]
+tile(tree, t0, t1, N_t/2);
+reorder(t0, t1, batch);
+parallel(t0);
+\end{lstlisting}
 
+The MIR generated by \Treebeard{} for the above schedule is as follows. 
 \begin{lstlisting}[style=c++]
-  float arr[10], result[1]
-  par.for i0 = 0  to 10 step 5 {
-    for i1 = 0 to 5 
-      reduce(result[0], arr[i0 + i1]) <"+", 0.0>
+  float result[batch_size]
+  model = ensemble(...) 
+  par.for t0 = 0 to N_t step N_t/2 {
+    for t1 = 0 to N_t/2 {
+      for batch = 0 to batch_size {
+        t = getTree(model, t0 + t1) 
+        p = walkDecisionTree(t, rows[batch])
+        reduce(result[batch], p)
+      }
+    }
   }  
@@
 
+The compiler simply generates a \op{reduce} op to perform
+the required parallel reduction. 
 The semantics of the \op{reduce} op is exactly the semantics of 
@@
 can be reduced at the end of the reduction loop it was inserted for. 
-\TODO{We cannot do better than this in terms of memory usage} \TODO{Need a proof}.
+%\TODO{We cannot do better than this in terms of memory usage} \TODO{Need a proof}.
 
-In our example above, parallel iterations of the \op{i0} loop accumulate 
-into the same element of the \op{result} array.
-We would therefore privatize the \op{result} array for each iteration of 
-the \op{i0} loop as follows. 
+In our example, \Treebeard{} determines that the \op{t0} loop is a reduction 
+loop w.r.t the \op{result} array and therefore legalizes 
+the reduction by inserting a privatized array 
+\op{partResults}. The privatized dimension of this array 
+is reduced at the end of the \op{t0} loop.
+
 \begin{lstlisting}[style=c++]
-  float arr[10], result[1]
-  float resultPriv[2][1]
-  par.for i0 = 0  to 10 step 5 {
-    for i1 = 0 to 5 
-      reduce(resultPriv[i0/5][0], arr[i0 + i1]) <"+", 0.0>
+  float result[batch_size], partResults[2][batch_size]
+  model = ensemble(...) 
+  par.for t0 = 0 to N_t step N_t/2 {
+    for t1 = 0 to N_t/2 {
+      for batch = 0 to batch_size {
+        t = getTree(model, t0 + t1) 
+        p = walkDecisionTree(t, rows[batch])
+        reduce(partResults[t0/(N_t/2)][batch], p)
       }
-  result = reduce_dimension(resultPriv, 0)
+    }
+  }
+  results = reduceDimension(partResults[:, :], 0)
 \end{lstlisting}
 
-The op \op{reduce\_dimension} reduces values across the specified
-dimension of an n-dimensional array. In the above example, 
-the \op{reduce\_dimension} op is reducing across all elements 
-of the first dimension (dimension 0). Therefore, in this case, it 
-produces a result memref with a single element (the first dimension
-with size 2 is collapsed).
-
-% \begin{definition}
-%   A parallel loop surrounding one or more \op{reduce} ops is 
-%   a \textbf{conflicting loop} for a target multi-dimensional array if this 
-%   loop has a non-zero dependence distance for the dependence between
-%   any of the contained \op{reduce} ops.
-% \end{definition}
-
-% \begin{definition}
-%   \op{\textbf{reduce\_dimension(targetMemref, memref, dim, [indices], [rangeStart], [rangeEnd])}}:
-%    Computes the reduction over the dimension specified by \op{dimension} and stores the 
-%    result in \op{targetMemref}. \op{[indices]} must be a vector of \op{dim} elements
-%    (or empty if the dimension being reduced is the first dimension). \op{[rangeStart]} 
-%    and \op{[rangeEnd]} represent the range of indices following the reduction dimension and 
-%    must have the same number of elements. If both are \op{null} (not passed), 
-%    all elements of these dimensions are reduced. The computation performed by the op is as follows.
-
-%   $targetMemref[\vec{\boldsymbol{indices}}, \vec{\boldsymbol{k}}] = \sum_{i=0}^{shape[dim]} memref[\vec{\boldsymbol{indices}}, i, \vec{\boldsymbol{k}}]\quad   \forall \vec{\boldsymbol{k}} \in \left[[rangeStart_0, rangeEnd_0), ...
-, [rangeStart_n, rangeEnd_n)\right]$
-% \end{definition}
-
-% Consider the following code with nested parallel loops. 
-% (A situation where trees are split across both
-% threads and thread blocks could result in such generated 
-% code in \Treebeard{}.)
-% \begin{lstlisting}[style=c++]
-%   builtin.func @ReduceVector(%arr: memref<num_elemsxf64>, %result: memref<1xf64>) -> void {
-%     par.for i0 = range(0 : num_elems/2 : num_elems) {
-%       par.for i1 = range(0 : num_elems/4 : num_elems/2) {
-%         for i2 = range(0 : num_elems/4) 
-%           reduce(%result, 0, arr[i0 + i1 + i2]) <"+", 0.0>
-%       }
-%     }
-%   }
-% \end{lstlisting}
-
-% Here, the \op{i0} and \op{i1} loops are conflicting loops wrt 
-% the \op{result} memref. We \textbf{legalize} the reduction 
-% by privatizing the \op{result} array wrt the \op{i0} and \op{i1} loops.
-% However, there are now two privatized dimensions and therefore, 
-% two dimensions need to be reduced to compute the final result.
-% This multi-stage reduction is what enables us to model 
-% hierarchical reductions.
-
-% The following code shows how the reduction above is legalized. 
-% We introduce a new op, \op{reduce\_dimension\_inplace} which 
-% reduces a dimension of the input memref and stores results 
-% in the same array. This helps saves memory by removing the 
-% need to create multiple intermediate arrays to store results.
-% Only the final dimension reduction uses the \op{reduce\_dimension} op.
-
-% \begin{lstlisting}[style=c++]
-%   builtin.func @ReduceVector(%arr: memref<num_elemsxf64>, %result: memref<1xf64>) -> void {
-%     results_1 = memref<2x2x1xf64>
-%     par.for i0 = range(0 : num_elems/2 : num_elems) {
-%       par.for i1 = range(0 : num_elems/4 : num_elems/2) {
-%         for i2 = range(0 : num_elems/4) 
-%           index0 = i0/(num_elems/2)
-%           index1 = i1/(num_elems/4)
-%           reduce(%result_1[index0, index1, 0], arr[i0 + i1 + i2]) <"+", 0.0>
-%       }
-%       // result_1[i0/(num_elems/2), 0] = sum(result_1[i0/(num_elems/2), :])
-%       reduce_dimension_inplace(%result_1, 1, i0/(num_elems/2)) 
-%     }
-%     // result = sum(result[:, 0])
-%     %result = reduce_dimension(%result_1, 0)
-%   }
-% \end{lstlisting}
+The op \op{reduceDimension} reduces values across the specified
+dimension of an n-dimensional array. Here, 
+it reduces all elements of the first dimension (dimension 0). 
+and produces a result memref with a single dimension of size \op{batch\_size}.
 
 To reduce the amount of memory used by arrays introduced for reduction,
-we introduce the \op{reduce\_dimension\_inplace} operation. It is similar to the 
-\op{reduce\_dimension} op except that it updates the input array inplace
+we introduce the \op{reduceDimInplace} operation. It is similar to the 
+\op{reduceDimension} op except that it updates the input array inplace
 rather than writing results to a target array. It writes results to the 
-zeroth index of the dimension being reduced.
+zeroth index of the dimension being reduced. We use this op to 
+compute intermediate results when several reduction loops are identified.
 
@@
 \subsection{Lowering Reduction Operations}
-We implement lowering of the operations defined above to both the CPU and GPU.
-Since the lowering pipeline from MIR to LIR are different for CPU and GPU 
-compilation, we implement lowering and optimization of our reduction dialect to CPUs and
-GPUs simply using different MLIR rewrite patterns. In this section, we briefly describe 
-how these operations are lowered to the CPU and GPU.
+% We implement lowering of the operations defined above to both the CPU and GPU.
+% Since the lowering pipeline from MIR to LIR are different for CPU and GPU 
+% compilation, we implement lowering and optimization of our reduction dialect to CPUs and
+% GPUs simply using different MLIR rewrite patterns. 
+In this section, we briefly describe 
+how we lower the reduction operations to the CPU and GPU.
 
-\subsubsection{Lowering to CPU}
-The lowering of the reduction operations to CPU is fairly straightforward. We lower 
-\op{reduce\_dimension\_inplace} and 
-\op{reduce\_dimension} to a simple loop nest that goes over the specified
+For both CPU and GPU, \op{reduce} is lowered to a sequence 
+of load, compute and write operations. This is possible because 
+legalization ensures that parallel threads do not write to the same 
+array element. 
+%\subsubsection{Lowering to CPU}
+
+For CPUs, we lower \op{reduceDimInplace} and 
+\op{reduceDimension} to a simple loop nest that goes over the specified
 subset of the input array, performs the reduction and writes 
@@
 If the schedule specifies that the reduction is to be vectorized,
-then as many elements as specified by the vector width are read 
-from the input array as a vector, accumulated as a vector, and 
-finally written back to the target array.
+the lowering passes generate vector (LLVM IR) instructions for 
+the reduction.
+% then as many elements as specified by the vector width are read 
+% from the input array as a vector, accumulated as a vector, and 
+% finally written back to the target array.
 
@@
 
-\subsubsection{Lowering to GPU}
-The same abstractions can be lowered to efficient GPU implementations
-and therefore, simplify higher-level code generation. The lowering for 
-the inplace and non-inplace operations are essentially the same, except 
-for the target array and we do not distinguish between them except 
-for finally storing the result.
-
-The lowering of the \op{reduce\_dimension\_*} ops can either exploit
+% \subsubsection{Lowering to GPU}
+The same reduction abstractions can be lowered to efficient GPU implementations
+and therefore, simplify higher-level code generation. 
+% The lowering for 
+% the inplace and non-inplace operations are essentially the same, except 
+% for the target array and we do not distinguish between them except 
+% for finally storing the result. 
+%
+% The lowering of the \op{reduceDim*} 
+\Treebeard{} can lower these ops to either exploit
 parallelism across the independent reductions or 
 the inherent parallelism in the reduction by performing a divide and conquer 
-reduction. If there are enough independent reductions to keep all threads
-in a thread block busy, then the lowering pass can generate code that performs one (or 
-multiple) reductions in each thread. If, however, there are not 
-enough independent reductions, then the lowering pass generates a tree 
-style reduction where multiple threads cooperate to perform a single reduction
-using inter-thread shuffles.
+reduction depending on whether there are enough independent reductions to keep all threads
+in a thread block busy.
+% then the lowering pass can generate code that performs one (or 
+% multiple) reductions in each thread. If, however, there are not 
+% enough independent reductions, then the lowering pass generates a tree 
+% style reduction where multiple threads cooperate to perform a single reduction
+% using inter-thread shuffles.
 
-Another feature specific to GPU reductions is the use of shared memory. 
-If the schedule specifies that the reduction needs to be performed 
+% Another feature specific to GPU reductions is the use of shared memory. 
+Additionally, if the schedule specifies that the reduction needs to be performed 
 using shared memory, the privatized buffer is allocated in shared memory. 
-The compiler only allocates as much shared memory 
-as needed to hold values processed by a single thread-block.
+% The compiler only allocates as much shared memory 
+% as needed to hold values processed by a single thread-block.
 %  and 
@@
 % the indexing of the target memref and the shared memory array.
-Our abstractions allow our lowering passes to be written completely 
-independent of whether we use shared memory and therefore allow 
+Our abstractions for reductions allow our lowering passes to be 
+agnostic of whether we use shared memory and therefore allow 
 us to enable or disable shared memory use independently from the other 
@@
 
-\subsection{Use in \Treebeard{}}
-We now show how \Treebeard{} uses the reduction dialect to generate 
-code for decision tree inference using an example. 
-In our example, \op{N\_t} is the number of trees 
-and \op{batch\_size} is the batch size. The schedule tiles both the 
-batch and tree loops and parallelizes the outer batch 
-and tree loops. The schedule with which code is generated is as follows.
-% In the context of \Treebeard{} we note that, the set of reduction loops is exactly the set 
-% of surrounding parallel loops that are iterating over trees. The 
-% results can be privatized for each loop iteratively 
-% and reductions along each privatized dimension can be inserted  
-% immediately following the loop the dimension was inserted due to.
-
-\begin{lstlisting}[style=c++]
-tile(batch, i0, i1, batch_size/2);
-tile(tree, t0, t1, N_t/2);
-reorder({i0, t0, t1, i1});
-parallel(t0);
-parallel(i0);
-\end{lstlisting}
-
-The MIR generated by \Treebeard{} for the above schedule is as follows. 
-\begin{lstlisting}[style=c++]
-  float result[batch_size]
-  model = ensemble(...) 
-  par.for i0 = 0 to batch_size step batch_size/2 {
-    par.for t0 = 0 to N_t step N_t/2 {
-      for t1 = 0 to N_t/2 {
-        for i1 = 0 to batch_size/2 {
-          t = getTree(model, t0 + t1) 
-          p = walkDecisionTree(t, rows[i0+i1])
-          reduce(result[i0+i1], p)
-        }
-      }
-    }
-  }
-\end{lstlisting}
-
-\Treebeard{} determines that the \op{t0} loop is a reduction 
-loop w.r.t the \op{result} array and therefore legalizes 
-the reduction by inserting a privatized array 
-\op{partResults}. The privatized dimension of this array 
-is reduced at the end of the \op{t0} loop.
-
-\begin{lstlisting}[style=c++]
-  float result[batch_size], partResults[2][batch_size]
-  model = ensemble(...) 
-  par.for i0 = 0 to batch_size step batch_size/2 {
-    par.for t0 = 0 to N_t step N_t/2 {
-      for t1 = 0 to N_t/2 {
-        for i1 = 0 to batch_size/2 {
-          t = getTree(model, t0 + t1) 
-          p = walkDecisionTree(t, rows[i0+i1])
-          reduce(result[i0+i1], p)
-        }
-      }
-    }
-    results[i0:i0+batch_size/2] = reduce_dimension(partResults[:, i0:i0+batch_size/2], 0)
-  }
-\end{lstlisting}
-
-While legalizing the reduction, the compiler determines that the 
-\op{reduce\_dimension} operation can only compute a subset of the final 
-result (the subset that is computed within the current parallel iteration of the 
-\op{i0} loop). 
-% Once this process is complete, the \op{reduce} ops in 
-% the result IR can be lowered to a simple ``read-accumulate-write''
-% sequence of instructions
-
-Finally, we note that in our experiments, we found that our 
-current implementation of lowering the reduction operations 
-was sufficient and reduction is not the bottleneck in our 
-generated code. However, we believe this approach to enabling 
-higher level code generators to easily generate reductions 
-through simple abstractions and then having the compiler 
-automatically lower them to efficient implementation is an
-important area for future work with applicability in several 
-domains.
+% We find that our current implementation of lowering reduction operations 
+% is sufficient for \Treebeard{} and reduction is not the bottleneck in the
+% generated code. However, we believe this approach to enabling 
+% higher level code generators to easily generate reductions 
+% through simple abstractions and then having the compiler 
+% automatically lower them to efficient implementation is an
+% important area for future work with applicability in several 
+% domains.
 

### 2024-04-18 results/python/analyze_random_model_benchmark_results.py
Add cummulative freq plots for random model speedups

@@ -1,4 +1,5 @@
 import os
+import matplotlib.pyplot as plt
 
 batch_size = 512
 # file_path = "/home/ashwin/mlir-build/llvm-project/mlir/examples/treebeard/results/thedeep/gpu/random_models/20240320/random_cmp_output_mp.txt"
@@ -52,30 +53,47 @@ import scipy.stats as stats
 # print(stats.gmean(df["tb_kernel_speedup"]))
 #print(df["tb_kernel_speedup"].prod() ** (1.0 / df.shape[0]))
 
-for depth in range(6, 9):
-    # get entries with tree_depth == depth
-    df_depth = df[df["tree_depth"] == depth]
-    # print(df_depth.shape)
+# for depth in range(6, 9):
+#     # get entries with tree_depth == depth
+#     df_depth = df[df["tree_depth"] == depth]
+#     # print(df_depth.shape)
 
-    # plot a 3-d graph of num_trees, num_features and tb_kernel_speedup
-    import matplotlib.pyplot as plt
-    from mpl_toolkits.mplot3d import Axes3D
+#     # plot a 3-d graph of num_trees, num_features and tb_kernel_speedup
+#     import matplotlib.pyplot as plt
+#     from mpl_toolkits.mplot3d import Axes3D
 
-    fig = plt.figure()
-    ax = fig.add_subplot(111, projection='3d')
+#     fig = plt.figure()
+#     ax = fig.add_subplot(111, projection='3d')
 
-    # 3D scatter plot
-    # ax.scatter3D(df_depth["num_trees"], df_depth["num_features"], df_depth["tb_kernel_speedup"], c='red')
-    surf = ax.plot_trisurf(df_depth["num_trees"], df_depth["num_features"], df_depth["tb_kernel_speedup"], cmap='viridis')
+#     # 3D scatter plot
+#     # ax.scatter3D(df_depth["num_trees"], df_depth["num_features"], df_depth["tb_kernel_speedup"], c='red')
+#     surf = ax.plot_trisurf(df_depth["num_trees"], df_depth["num_features"], df_depth["tb_kernel_speedup"], cmap='viridis')
 
-    ax.set_xlabel('Number of Trees')
-    ax.set_ylabel('Number of Features')
-    ax.set_zlabel('Speedup')
-    min_speedup = df_depth["tb_kernel_speedup"].min()
[…]

### 2024-04-18 results/python/analyze_gpu_speedups_results.py
Add more graphs for geomean speedup for benchmarks + double yaxis graph

@@ -136,6 +136,103 @@ def plot_bar_graph_speedups_for_batch_size(df, batch_size, kwargs=None):
     # plt.show()
     plt.savefig(f'speedup_bar_graph_{batch_size}.{PLOT_FILE_EXTENSION}')
 
+def plot_bar_graph_geomean_speedups_for_models(df, gpu_name):
+    # Prepare data for plotting
+    bar_names = df['Benchmark'].unique()
+    bar_names.sort()
+    batch_sizes = df['Batch size'].unique()
+    batch_sizes.sort()
+    speedups = [[],[]]
+
+    for benchmark in bar_names:
+        filtered_df = df[df['Benchmark'] == benchmark]
+        rapids_speedups = []
+        tahoe_speedups = []
+        for batch_size in batch_sizes:
+            batch_size_df = filtered_df[filtered_df['Batch size'] == batch_size]
+            assert(batch_size_df.shape[0] == 1)
+
+            tb_time = batch_size_df['TBKernel(AT)'].values[0]
+            rapids_time = batch_size_df['RAPIDS(kernel)'].values[0]
+            tahoe_time = batch_size_df['Tahoe(kernel)'].values[0]
+            rapids_speedups.append(rapids_time / tb_time)
+            tahoe_speedups.append(tahoe_time / tb_time)
+        
+        speedups[0].append(gmean(rapids_speedups))
+        speedups[1].append(gmean(tahoe_speedups))
+
+    # Plot the bar graph
+    width = BAR_LABEL_FONT_SIZE
+    gap = BAR_LABEL_FONT_SIZE * 0.3
+    pos = np.array([float(2 * i * width + i * gap) for i in range(len(bar_names))])
+
+    _, ax = plt.subplots()
+    
+    ax.bar(pos, speedups[0], width, label='Speedup of SilvanForge vs RAPIDS', hatch='//////')
+    ax.bar(pos + width, speedups[1], width, label='Speedup of SilvanForge vs Tahoe', hatch='......')
+
+    # Add labels and title
+    ax.set_xlabel('Benchmark', fontsize=AXIS_FONT_SIZE)
+    ax.set_ylabel('Speedup', fontsize=AXIS_FONT_SIZE)
+    ax.set_xticks(pos + width / 2)
[…]

### 2024-04-18 docs/SOSP 2024/motivation.tex
Align graphs
nit

@@
   \caption{The representation of a model in high-level IR and the model 
-  after trees are padded and reordered. }
+  after trees are padded and reordered.\TODO{drop reorder, make 1,2 depth 1 and 3,4 depth 2.} }
   \label{Fig:HIRExample}
@@
 When the schedule specifies that the tree walks should be unrolled, 
-\Treebeard{} pads the trees so that all leaves are at the same depth.
+\Treebeard{} pads the trees so that all leaves are at the same depth (the padded trees are also shown in Figure~\ref{Fig:HIRExample}).
 
@@
 any parallelism. 
-%% RG
+%% RG, 
 %%  and is therefore ill-suited for parallel processors.
@@
 %% RG:
-While this may work for multi-core CPUs, with massively parallel processors like GPUs,
+While this may work for multi-core CPUs\footnote{As we report in Section~\ref{sec:results}, other ways to parallelize can benefit inference on CPUs as well.}, with massively parallel processors like GPUs,
  this strategy may not yield sufficient parallel work. To expose more parallelism, 
@@
 \subsection{Performance of Different Schedules} 
-
 \begin{figure}[htb]
   \begin{subfigure}[t]{.45\linewidth}
+    \vspace{0pt}
     \centering
     \includegraphics[width=\linewidth]{figures/batch_sensitivity_covtype.png}
+    \vspace{2pt}
     \caption{\label{fig:sensitivitya}\op{covtype} batch sensitivity}
-  \end{subfigure}
-  \begin{subfigure}[t]{.45\linewidth}
+  \end{subfigure}\hfill
+  \begin{subfigure}[t]{.52\linewidth}
+    \vspace{0pt}
     \centering
@@
   \end{subfigure}
-  \caption{Batch and model sensitivity plots. Each point shows the slowdown when best schedule for the
-          x-axis batch size (model) is used for the y-axis batch size (model).}
+  \caption{Batch and model sensitivity plots. Each point shows the slowdown when the best schedule for the x-axis batch size (model) is used for the y-axis batch size (model).}
 \end{figure}
@@
 The degradation is much worse when schedules are used across different models.
-In many instances (~20\%), using the best schedule for one model 
+In many (\~20\%) instances, using the best schedule for one model 
 to compile another results in a 5$\times$ slowdown.

### 2024-04-18 docs/SOSP 2024/results.tex
Add new graphs

@@
 We evaluate \Treebeard{} on four different target processors, an NVIDIA RTX 4060 GPU (8GB RAM, CUDA 11.5),
-an NVIDIA T400 GPU (2GB RAM, CUDA 11.5), an AMD MI210 GPU (64GB RAM, ROCm 4.5?) and an 
+an NVIDIA T400 GPU (2GB RAM, CUDA 11.5), an AMD MI210 GPU (64GB RAM, ROCm 4.5\TODO{Check the version}) and an 
 Intel Core i9-11900K CPU (16 virtual cores, 128 GB RAM). We compare \Treebeard{} with 
@@
 Each tree has leaves at depths 2 to a maximum depth of 6, 7 or 8.
+For \Treebeard{}, we use schedules found using the schedule exploration
+heuristic (Section \ref{sec:exploring}) unless otherwise specified.
 
@@
 We do not show results for XGBoost since the speedups are an order of magnitude higher and too large to fit on the same graph.  
-  The plot has lines for kernel time and total time speedup over RAPIDS and kernel time speedup over Tahoe.
-As can been seen \Treebeard{} is uniformly faster than Tahoe{\footnote{Tahoe does not support multiclass models and to enable comparison wet ran the multiclass models (\op{covtype} and \op{letters}) as regression models.
-We also noticed that some variants of Tahoe produce wrong results (as reported by its own tests) for
-  \op{letters} and \op{year}. In these cases, we pick the time of the fastest variant that gives the correct results.
-}} by $2-3\times$ at all batch sizes. 
-Compared to RAPIDS, \Treebeard{} is about $4\times$ faster at batch size 512.
-The relative performance of RAPIDs improves with batch size, but \Treebeard{} is still faster by $1.5-2\times$ evan at very high ($16K$) batch sizes.
-The plot also shows that the speedups are significant even when data needs to be transferred to the GPU and results need to be transferred back.
-They are lower than kernel time speedup as both systems have a constant additional transfer overhead.
+The plot has lines for kernel time and total time speedup over RAPIDS and kernel time speedup over Tahoe. As can been seen 
+\Treebeard{} is uniformly faster than Tahoe{\footnote{Tahoe does not support multiclass models. To enable comparison, we ran 
+the multiclass models (\op{covtype} and \op{letters}) as regression models only with Tahoe. We also noticed that some variants
+of Tahoe produce wrong results (as reported by its own tests) for \op{letters} and \op{year}. In these cases, we pick the 
+time of the fastest variant that gives the correct results.}} by $2-3\times$ at all batch sizes. 
+Compared to RAPIDS, \Treebeard{} is about $4\times$ faster at batch size 512. The relative performance of RAPIDs improves 
+with batch size, but \Treebeard{} is still faster by $1.5-2\times$ even at very high ($16K$) batch sizes. 
+The plot also shows that the speedups are significant even when data needs to be transferred to the GPU and results need to
+be transferred back. They are lower than kernel time speedup as both systems have a constant additional transfer overhead.
 
@@
   \caption{\Treebeard{} vs RAPIDs and Tahoe kernel time and total time speedup on NVIDIA RTX 4060}
+  \label{Fig:TBvsRAPIDsTahoe_4060_Speedup}
+\end{figure}
+
+\begin{figure}[htb]
+  \centering
+  \includegraphics[width=0.75\linewidth]{figures/geomean_speedup_4060_bar_graph.png}
+  \caption{\Treebeard{} vs RAPIDs and Tahoe kernel time and total time speedup on NVIDIA RTX 4060 for individual benchmarks (geomean over all batch sizes).
+  \TODO{Add bar for total time. Add whiskers to each bar to show range of speedups.}}
   \label{Fig:TBvsRAPIDsTahoe_4060_Speedup}
@@
   \centering
-  \includegraphics[width=0.75\linewidth]{figures/geomean_speedup_T400_4060_vs_T400_vs_MI2160.png}
-  \caption{Geomean speedup of schedule found by exploration heuristic vs best 4060 schedule on NVIDIA RTX 4060, NVIDIA T400 and AMD MI2160 across batch sizes}
-  \label{Fig:AutotuningSpeedupvs4060Sched}
+  \includegraphics[width=0.75\linewidth]{figures/speedup_vs_norm_time_line_graph_4060.png}
+  \caption{Speedup of schedule found by the heuristic vs best schedule found through exhaustive search on NVIDIA RTX 4060 and 
+  normalized schedule exploration heuristic exploration time speedup w.r.t full schedule exploration}
+  \label{Fig:HeuristicVsFullExplore_Speedup}
 \end{figure}
@@
   \centering
-  \includegraphics[width=0.75\linewidth]{figures/geomean_speedup_4060_full_exp_vs_at.png}
-  \caption{Schedule exploration heuristic exploration time speedup vs full schedule exploration}
-  \label{Fig:HeuristicVsFullExplore_Speedup}
+  \includegraphics[width=0.75\linewidth]{figures/geomean_speedup_T400_4060_vs_T400_vs_MI2160.png}
+  \caption{Geomean speedup of schedule found by exploration heuristic on NVIDIA T400 and AMD MI2160 vs 
+  schedule found by the heuristic on NVIDIA RTX 4060 across batch sizes\TODO{Change the plot labels to say vs 4060 heuristic schedule}}
+  \label{Fig:AutotuningSpeedupvs4060Sched}
 \end{figure}

### 2024-04-18 results/python/analyze_random_model_benchmark_results.py
Change 3d plot axis limit

@@ -1,9 +1,9 @@
 import os
 import matplotlib.pyplot as plt
 
-batch_size = 512
-# file_path = "/home/ashwin/mlir-build/llvm-project/mlir/examples/treebeard/results/thedeep/gpu/random_models/20240320/random_cmp_output_mp.txt"
-file_path = "/home/ashwin/mlir-build/llvm-project/mlir/examples/treebeard/results/thedeep/gpu/random_models/20240320/random_cmp_output_mp_b512.txt"
+batch_size = 4096
+file_path = "/home/ashwin/mlir-build/llvm-project/mlir/examples/treebeard/results/thedeep/gpu/random_models/20240320/random_cmp_output_mp.txt"
+# file_path = "/home/ashwin/mlir-build/llvm-project/mlir/examples/treebeard/results/thedeep/gpu/random_models/20240320/random_cmp_output_mp_b512.txt"
 # file_path = "/home/ashwin/mlir-build/llvm-project/mlir/examples/treebeard/results/thedeep/gpu/random_models/20240320/random_cmp_output_mp_b4k_ft500.txt"
 
 # read the file line by line and filter out lines that don't start with "/"
@@ -53,47 +53,46 @@ import scipy.stats as stats
 # print(stats.gmean(df["tb_kernel_speedup"]))
 #print(df["tb_kernel_speedup"].prod() ** (1.0 / df.shape[0]))
 
-# for depth in range(6, 9):
-#     # get entries with tree_depth == depth
-#     df_depth = df[df["tree_depth"] == depth]
-#     # print(df_depth.shape)
+for depth in range(6, 9):
+    # get entries with tree_depth == depth
+    df_depth = df[df["tree_depth"] == depth]
+    # print(df_depth.shape)
 
-#     # plot a 3-d graph of num_trees, num_features and tb_kernel_speedup
-#     import matplotlib.pyplot as plt
-#     from mpl_toolkits.mplot3d import Axes3D
+    # plot a 3-d graph of num_trees, num_features and tb_kernel_speedup
 
-#     fig = plt.figure()
-#     ax = fig.add_subplot(111, projection='3d')
+    fig = plt.figure()
+    ax = fig.add_subplot(111, projection='3d')
 
-#     # 3D scatter plot
[…]

### 2024-04-18 docs/SOSP 2024/results.tex
Edit results

@@
 \subsection{Performance comparisons on NVIDIA GPUs}
+
+\begin{figure}[htb]
+  \centering
+  \includegraphics[width=0.75\linewidth]{figures/geomean_speedup_4060_kernel_time_total_time.png}
+  \caption{\Treebeard{} vs RAPIDs and Tahoe kernel time and total time speedup on NVIDIA RTX 4060}
+  \label{Fig:TBvsRAPIDsTahoe_4060_Speedup}
+\end{figure}
+
+% \begin{figure}[htb]
+%   \centering
+%   \includegraphics[width=0.75\linewidth]{figures/geomean_speedup_4060_bar_graph.png}
+%   \caption{\Treebeard{} vs RAPIDs and Tahoe kernel time and total time speedup on NVIDIA RTX 4060 for individual benchmarks (geomean over all batch sizes).
+%   \TODO{Add bar for total time. Add whiskers to each bar to show range of speedups.}}
+%   \label{Fig:TBvsRAPIDsTahoe_4060_Speedup}
+% \end{figure}
+
+\begin{figure}[htb]
+  \centering
+  \includegraphics[width=0.75\linewidth]{figures/geomean_speedup_T400_kernel_time.png}
+  \caption{\Treebeard{} vs RAPIDs and Tahoe kernel time and total time speedup on NVIDIA T400.}
+  \label{Fig:TBvsRAPIDsTahoe_T400_Speedup}
+\end{figure}
+
+\begin{figure*}[ht]
+  \centering
+  \begin{subfigure}[b]{.45\textwidth}
+    \subcaptionbox*{}{\includegraphics[width=\textwidth]{figures/speedup_bar_graph_1024.png}}
+    \caption{Batch size 1024}
+  \end{subfigure}
+  \begin{subfigure}[b]{.45\textwidth}
+    \subcaptionbox*{}{\includegraphics[width=\textwidth]{figures/speedup_bar_graph_8192.png}}
+    \caption{Batch size 8192}
+  \end{subfigure}
+  \hfill
+  \caption{\label{Fig:KernelTimeIndividualBenchmarks4060}Kernel time speedup of \Treebeard{} vs RAPIDs on NVIDIA RTX 4060. Numbers on the bars are 
+  inference times per sample in $\mu$s for RAPIDs and Tahoe.}
+\end{figure*}
+
 \subsubsection*{Real-world models}
@@
 Figure~\ref{Fig:KernelTimeIndividualBenchmarks4060} shows per benchmark results at two different batch sizes. 
-\Treebeard{} is able to find better schedules than both RAPIDS and Tahoe on each of the benchmarks.
-It consistently outperforms both systems, achieving speedups in the range $1.1-12\times$, with about half the benchmarks achieving a speedup of $2\times$ or more  over both baselines at batch size 8192.
+\Treebeard{} is able to find better schedules than both RAPIDS and Tahoe on each of the benchmarks. It consistently
+outperforms both systems, achieving speedups in the range $1.1-12\times$, with about half the benchmarks achieving a speedup of $2\times$ or more  over both baselines at batch size 8192.
 \begin{figure}[htb]
@@
   \end{minipage}
-  \caption{\label{fig:randomModels4060}\Treebeard{} vs RAPIDs Kernel Time Speedup for several randomly generated models.}
+  \caption{\label{fig:randomModels4060}\Treebeard{} vs RAPIDs Kernel Time Speedup for random models
+  of max depth 8 at batch size 512 and max depth 6 at batch size 4096.}
 \end{figure}
@@
 % \end{itemize}
+\subsubsection*{Utility of Schedule Constructs}
+A detailed analysis of the generated schedules indicate that caching, in-memory representation 
+and reduction optimizations are crucial for performance. We observe that all three representations 
+are necessary, array representation is used 60\% of the time, sparse 30\% of the time and reorg 10\% of the time. 
+Shared reduction significantly improves performance especially for multi-class models. For example, 
+the \op{letters} benchmark achieves a speedup of 1.5$\times$ with shared reduction. 
+Shared reduction increases the memory pressure (on shared memory) and not all schedules use it.
+While GPUs have a hardware managed cache, explicitly caching the rows using the scheduling primitives 
+helps exploit the reuse of rows across trees. We find that caching rows is beneficial for many models
+with speedups as large as $1.5\times$ over the best schedule without caching.
+
 \subsubsection*{Comparison on T400}
@@
 
-\begin{figure}[htb]
-  \centering
-  \includegraphics[width=0.75\linewidth]{figures/geomean_speedup_4060_kernel_time_total_time.png}
-  \caption{\Treebeard{} vs RAPIDs and Tahoe kernel time and total time speedup on NVIDIA RTX 4060}
-  \label{Fig:TBvsRAPIDsTahoe_4060_Speedup}
-\end{figure}
-
-\begin{figure}[htb]
-  \centering
-  \includegraphics[width=0.75\linewidth]{figures/geomean_speedup_4060_bar_graph.png}
-  \caption{\Treebeard{} vs RAPIDs and Tahoe kernel time and total time speedup on NVIDIA RTX 4060 for individual benchmarks (geomean over all batch sizes).
-  \TODO{Add bar for total time. Add whiskers to each bar to show range of speedups.}}
-  \label{Fig:TBvsRAPIDsTahoe_4060_Speedup}
-\end{figure}
-
-\begin{figure}[htb]
-  \centering
-  \includegraphics[width=0.75\linewidth]{figures/geomean_speedup_T400_kernel_time.png}
-  \caption{\Treebeard{} vs RAPIDs and Tahoe kernel time and total time speedup on NVIDIA T400.}
-  \label{Fig:TBvsRAPIDsTahoe_T400_Speedup}
-\end{figure}
-
-\begin{figure*}[ht]
-  \centering
-  \begin{subfigure}[b]{.45\textwidth}
-    \subcaptionbox*{}{\includegraphics[width=\textwidth]{figures/speedup_bar_graph_1024.png}}
-    \caption{Batch size 1024}
-  \end{subfigure}
-  \begin{subfigure}[b]{.45\textwidth}
-    \subcaptionbox*{}{\includegraphics[width=\textwidth]{figures/speedup_bar_graph_8192.png}}
-    \caption{Batch size 8192}
-  \end{subfigure}
-  \hfill
-  \caption{\label{Fig:KernelTimeIndividualBenchmarks4060}Kernel time speedup of \Treebeard{} vs RAPIDs on NVIDIA RTX 4060. Numbers on the bars are 
-  inference times per sample in $\mu$s for RAPIDs and Tahoe.}
-\end{figure*}
-
 % \begin{figure*}[ht]
@@
 \subsection{Performance on AMD GPUs}
-
 While existing systems only support NVIDIA GPUs, we demonstrate that \Treebeard{} 
-is able to generate competitive code for AMD GPUs. Our objective here is not to compare directly 
-between AMD and NVIDIA GPUs, but we do find that as the MI210 is a more powerful GPU, 
-it achieves better inference times on most benchmarks at large batch sizes ($\geq8$k).
+is able to generate competitive code for AMD GPUs. \Treebeard{} can target AMD GPUs 
+because it generates a combination of MLIR's \op{gpu} dialect and LLVM IR 
+and these can be JIT'ed to AMD GPUs. While our objective here is not to compare directly 
+between AMD and NVIDIA GPUs, we do find that the MI210 achieves better inference times 
+on most benchmarks at large batch sizes ($\geq8$k). For example, at a batch size of 16k,
+the MI210 is $2\times$ faster than the RTX 4060 on the \op{letters} benchmark.
 
@@
 We evaluated the schedule exploration heuristic described in Section \ref{sec:exploring} on several fronts.
-Figure \ref{Fig:AutotuningSpeedupvs4060Sched} compares the best schedule found by exhaustive exploration on the RTX 4060 
-with the schedule found by the exploration heuristic on RTX 4060, T400 and AMD MI210.
-% First, we determine whether the heuristic is able to find schedules that are close to the best 
-% schedule found by exhaustive exploration. Figure \ref{Fig:AutotuningSpeedupvs4060Sched} shows the speedup of the schedule 
-% found by the heuristic vs the best schedule found by exhaustive exploration on the RTX 4060.
-With this we want to establish two things described below.
-\begin{enumerate}
-\item \textbf{\emph{Quality of Heuristic Schedules:}}
-As can be seen from the plot  
-for RTX 4060 the heuristic is able to find schedules that are very close to the best schedule (well 
-within 5\%).
-\item \textbf{\emph{Schedule Sensitivity Across GPUs:}}
-The remaining two lines show that the best schedule varies across target processors.
+Below we describe the major conclusions we can draw from the evaluation.
+
+\subsubsection*{Efficacy and Speed of Scheduling Heuristic}
+Figure \ref{Fig:HeuristicVsFullExplore_Speedup} compares the best schedule found by exhaustive exploration on the RTX 4060 
+with the schedule found by the exploration heuristic on RTX 4060.
+The plot establishes that the 
+heuristic is able to find schedules that are very close to the best schedule (well 
+within 5\%) on 4060. The heuristic
+is consistently close to two orders of magnitude faster than exhaustive exploration. 
+The exploration time ranges between $6$ and $167$ seconds for the heuristic with a mean of $28.7$ seconds.
+These results show that our heuristic is able to quickly find schedules that are close to the best schedule.
+
+\subsubsection*{Schedule Sensitivity Across GPUs}
+Figure \ref{Fig:AutotuningSpeedupvs4060Sched}
+compares the schedule found by the heuristic on T400 and AMD MI210 with the schedule found on the 4060.
+It shows that schedule exploration needs to be done on each target to achieve the best performance
+and schedules found on one GPU do not always carry over to other GPUs.
 For T400, the heuristic is able to find schedules that are 1.05$\times$ to 1.2$\times$ better 
@@
 We find that there is a much larger variation in performance 
-between the schedules \Treebeard{} finds on the MI210 compared to the best 4060 schedules. For example, 
+between the schedules \Treebeard{} finds on the MI210 compared to the 4060 schedules. For example, 
 the geomean speedup over all benchmarks is 1.5$\times$ at batch size 16k and 
 The maximum speedup is 2$\times$ for the \op{letters} benchmark at batch size 16k.
-\end{enumerate}
 
-\subsubsection*{Exploration Time}
-Finally, we measure the improvement in exploration time that the heuristic provides 
-compared to exhaustive exploration. Figure \ref{Fig:HeuristicVsFullExplore_Speedup} shows that the heuristic
-is consistently close to two orders of magnitude faster than exhaustive exploration. 
-The exploration time ranges between $6$ and $167$ seconds for the heuristic with a mean of $28.7$ seconds.
-These results show that our heuristic is able to quickly find schedules that are close to the best schedule.
+% \subsubsection*{Exploration Time}
+% Finally, we measure the improvement in exploration time that the heuristic provides 
+% compared to exhaustive exploration. Figure \ref{Fig:HeuristicVsFullExplore_Speedup} shows that the heuristic
+% is consistently close to two orders of magnitude faster than exhaustive exploration. 
+% The exploration time ranges between $6$ and $167$ seconds for the heuristic with a mean of $28.7$ seconds.
+% These results show that our heuristic is able to quickly find schedules that are close to the best schedule.
 
@@
 
+\subsection{Discussion}
+\subsubsection*{Cost Performance Tradeoff Across Hardware Platforms}
+Two of the three GPU targets we use are commodity GPUs that cost less than \$400. They 
+are infact much cheaper than the CPU we evaluate on. Despite this, performance on GPUs
+is significantly better than on CPUs (by an order of magnitude on RTX 4060).
+We believe \Treebeard{} can help enable data scientists to better leverage 
+available processors on their systems.
+
+\subsubsection*{A Note on Batch Size}
+Batch size is an important parameter dictated by application requirements. It 
+exposes a tradeoff between latency and throughput. It is not possible to use 
+large batch sizes in latency sensitive applications. Ideally, one would like 
+to maximize performance while respecting the batch size constraint imposed 
+by the application. \Treebeard{} enables users to explore this tradeoff.
+
 % \begin{itemize}

### 2024-04-18 docs/SOSP 2024/motivation.tex
Edit figures and references

@@
   \includegraphics[width=\linewidth]{figures/HIR.PNG}
-  \caption{The representation of a model in high-level IR and the model 
-  after trees are padded and reordered.\TODO{drop reorder, make 1,2 depth 1 and 3,4 depth 2.} }
+  \caption{Model for motivating example. The model has four trees, two of depth 1 and two of depth 2.}
   \label{Fig:HIRExample}
@@
 Consider a model with four trees, two 
-trees of depth 1 and two of depth 2 (Figure \ref{Fig:HIRExample}).
+trees of depth 1 and two of depth 2 (\circled{A} in Figure \ref{Fig:HIRExample}).
 % We use this model as our running example throughout this section.
@@
 for all input rows. The scheduling constructs are fairly intuitive. We 
-defer a detailed explanation of the scheduling language to Section~\ref{sec:scheduling}. 
+defer a detailed explanation of the scheduling language to Section~\ref{sec:schedule}. 
 The loop over the trees is split into two -- one that
@@
 When the schedule specifies that the tree walks should be unrolled, 
-\Treebeard{} pads the trees so that all leaves are at the same depth (the padded trees are also shown in Figure~\ref{Fig:HIRExample}).
+\Treebeard{} pads the trees so that all leaves are at the same depth
+(\circled{B} in Figure~\ref{Fig:HIRExample}).
 
@@
   for t_depth1 = 0 to 2 step 1 {
-    T = getTree(ensemble, t_depth1)
+    T = getTree(model, t_depth1)
     for batch = 0 to BATCH_SIZE step 1 {
@@
   for t_depth2 = 2 to 4 step 1 {
-    T = getTree(ensemble, t_depth2)
+    T = getTree(model, t_depth2)
     for batch = 0 to BATCH_SIZE step 1 {
@@
       for t0_depth1 = 0 to 2 step 2 {
-        T = getTree(ensemble, t0_depth1 + t1)
+        T = getTree(model, t0_depth1 + t1)
         treePred = walkDecisionTree(T, 
@@
       for t0_depth2 = 2 to 4 step 2 {
-        T = getTree(ensemble, t0_depth2 + t1)
+        T = getTree(model, t0_depth2 + t1)
         treePred = walkDecisionTree(T,
@@
 and micro-architectural features like register file size, handling 
-of branch divergence etc.~\cite{FungMicro11, SomethingElse}.
+of branch divergence etc.~\cite{FungMicro, MilindDivergence}.
 

### 2024-04-18 docs/SOSP 2024/motivation.tex
Add figures for schedules

@@
 % We use this model as our running example throughout this section.
-We first describe a simple schedule that processes one tree at a time 
+We first describe the simple schedule shown in Figure \ref{fig:schedulea} that processes one tree at a time 
 for all input rows. The scheduling constructs are fairly intuitive. We 
@@
 
-\begin{lstlisting}[style=c++]
-  reorder(tree, batch)
-  // Fiss tree loop so trees with equal depth 
-  // are processed together  
-  split(tree, t_depth1, t_depth2, 2)
-  // Unroll the tree walks
-  unrollWalk(t_depth1, 1)
-  unrollWalk(t_depth2, 2)
-\end{lstlisting}
-The concrete implementation of this schedule (in \Treebeard{}'s IRs) 
-is as follows.
-\begin{lstlisting}[style=c++]
-  model = ensemble(...)
-  for t_depth1 = 0 to 2 step 1 {
-    T = getTree(model, t_depth1)
-    for batch = 0 to BATCH_SIZE step 1 {
-      treePred = walkDecisionTree(T, 
-                    input[batch]) <unrollDepth = 1>
-      reduce(result[batch], treePred)
-    }
-  }
-  for t_depth2 = 2 to 4 step 1 {
-    T = getTree(model, t_depth2)
-    for batch = 0 to BATCH_SIZE step 1 {
-      treePred = walkDecisionTree(T,
-                    input[batch]) <unrollDepth = 2>
-      reduce(result[batch], treePred) <'+', 0.0>
-    }
-  }
-\end{lstlisting}
+\begin{figure*}[]
+  \begin{subfigure}[t]{.495\linewidth}
+    \vspace{0pt}
+    \centering
+    \includegraphics[width=\linewidth]{figures/MotivatingExample1.PNG}
+    \caption{\label{fig:schedulea}}
+  \end{subfigure}\hfill
+  \begin{subfigure}[t]{.495\linewidth}
+    \vspace{0pt}
+    \centering
+    \includegraphics[width=\linewidth]{figures/MotivatingExample2.PNG}
+    \caption{\label{fig:scheduleb}}
+  \end{subfigure}
+  \caption{Possible schedules to compile the motivating example and the generated \Treebeard{} IR.}
+  \label{Fig:Schedules}
+\end{figure*}
+
+% \begin{lstlisting}[style=c++]
+%   reorder(tree, batch)
+%   // Fiss tree loop so trees with equal depth 
+%   // are processed together  
+%   split(tree, t_depth1, t_depth2, 2)
+%   // Unroll the tree walks
+%   unrollWalk(t_depth1, 1)
+%   unrollWalk(t_depth2, 2)
+% \end{lstlisting}
+The concrete implementation of this schedule (in \Treebeard{}'s IR) 
+is as shown in Figure \ref{fig:schedulea}.
+% \begin{lstlisting}[style=c++]
+%   model = ensemble(...)
+%   for t_depth1 = 0 to 2 step 1 {
+%     T = getTree(model, t_depth1)
+%     for batch = 0 to BATCH_SIZE step 1 {
+%       treePred = walkDecisionTree(T, 
+%                     input[batch]) <unrollDepth = 1>
+%       reduce(result[batch], treePred)
+%     }
+%   }
+%   for t_depth2 = 2 to 4 step 1 {
+%     T = getTree(model, t_depth2)
+%     for batch = 0 to BATCH_SIZE step 1 {
+%       treePred = walkDecisionTree(T,
+%                     input[batch]) <unrollDepth = 2>
+%       reduce(result[batch], treePred) <'+', 0.0>
+%     }
+%   }
+% \end{lstlisting}
 This schedule is ideally suited for a single-core CPU. It maximizes 
@@
 %% RG:
-While this may work for multi-core CPUs\footnote{As we report in Section~\ref{sec:results}, other ways to parallelize can benefit inference on CPUs as well.}, with massively parallel processors like GPUs,
+While this may work for multi-core CPUs\footnote{As we report in Section~\ref{sec:results}, 
+other ways to parallelize can benefit inference on CPUs as well.}, with massively parallel processors like GPUs,
  this strategy may not yield sufficient parallel work. To expose more parallelism, 
-we can additionally parallelize across trees as done in the schedule below. 
+we can additionally parallelize across trees as done in the schedule in Figure \ref{fig:scheduleb}. 
 %% RG:
@@
 % A possible strategy to accomplish this is encoded in the following schedule.
-\begin{lstlisting}[style=c++]
-  // Split the trees into two sets
-  tile(tree, t0, t1, 2)
-  reorder(batch, t1, t0)
-  // Fiss loop so that trees with equal 
-  // depth are processed together
-  split(t0, t0_depth1, t0_depth2, 2)
-  unrollWalk(t0_depth1, 1)
-  unrollWalk(t0_depth2, 2)
-  // Configure the GPU kernel dimensions
-  gpuDimension(batch, grid.x)
-  gpuDimension(t1, block.x)
-\end{lstlisting}
-This schedule generates an inference function that runs on the GPU. 
+% \begin{lstlisting}[style=c++]
+%   // Split the trees into two sets
+%   tile(tree, t0, t1, 2)
+%   reorder(batch, t1, t0)
+%   // Fiss loop so that trees with equal 
+%   // depth are processed together
+%   split(t0, t0_depth1, t0_depth2, 2)
+%   unrollWalk(t0_depth1, 1)
+%   unrollWalk(t0_depth2, 2)
+%   // Configure the GPU kernel dimensions
+%   gpuDimension(batch, grid.x)
+%   gpuDimension(t1, block.x)
+% \end{lstlisting}
+This schedule generates an inference function that runs on the GPU as shown in 
+the IR in Figure \ref{fig:scheduleb}. 
 The inference routine processes one input row per thread block (since the \op{batch}
@@
 Each of the two sets is processed in parallel. We unroll the tree walks 
-for each tree. The IR generated is as follows. 
-\begin{lstlisting}[style=c++]
-  model = ensemble(...)
-  par.for batch = 0 to BATCH_SIZE step 1 <grid.x> {
-    par.for t1 = 0 to 2 step 1 <block.x> {
-      for t0_depth1 = 0 to 2 step 2 {
-        T = getTree(model, t0_depth1 + t1)
-        treePred = walkDecisionTree(T, 
-                        input[batch]) <unrollDepth = 1>
-        reduce(result[batch], treePred)
-      }
-      for t0_depth2 = 2 to 4 step 2 {
-        T = getTree(model, t0_depth2 + t1)
-        treePred = walkDecisionTree(T,
-                        input[batch]) <unrollDepth = 2>
-        reduce(result[batch], treePred) <'+', 0.0>
-      }
-    }
-  }
-\end{lstlisting}
+for each tree.
+% The IR generated is as follows. 
+% \begin{lstlisting}[style=c++]
+%   model = ensemble(...)
+%   par.for batch = 0 to BATCH_SIZE step 1 <grid.x> {
+%     par.for t1 = 0 to 2 step 1 <block.x> {
+%       for t0_depth1 = 0 to 2 step 2 {
+%         T = getTree(model, t0_depth1 + t1)
+%         treePred = walkDecisionTree(T, 
+%                         input[batch]) <unrollDepth = 1>
+%         reduce(result[batch], treePred)
+%       }
+%       for t0_depth2 = 2 to 4 step 2 {
+%         T = getTree(model, t0_depth2 + t1)
+%         treePred = walkDecisionTree(T,
+%                         input[batch]) <unrollDepth = 2>
+%         reduce(result[batch], treePred) <'+', 0.0>
+%       }
+%     }
+%   }
+% \end{lstlisting}
 

### 2024-04-18 docs/SOSP 2024/related.tex
Shorten related work

@@
 \emph{Decision Tree Inference Systems:} 
-Tahoe\cite{Tahoe} is a system that implements high-performance library routines and a 
-performance model for tree inference on GPUs. Tahoe is a library-based system that picks 
+Tahoe~\cite{Tahoe} is a library-based system that picks 
 between four predefined strategies to implement decision tree inference on GPUs.
-In comparison, \Treebeard{} explores a much larger set of implementation options 
-because it is a compiler. \Treebeard{} can also explore different in-memory 
-representations for models. Also, \Treebeard{} generates code that is specific to 
-a particular model, specializing both the parallelism (by deciding the thread block 
-structure on a per model basis) and the kernel code itself by 
-performing optimizations like tree walk unrolling and interleaving.
+In comparison, \Treebeard{} explores a much larger set of implementation options and 
+in-memory representations for models and picks the best.
+% Also, \Treebeard{} generates code that is specific to 
+% a particular model, specializing both the parallelism (by deciding the thread block 
+% structure on a per model basis) and the kernel code itself by 
+% performing optimizations like tree walk unrolling and interleaving.
 % In contrast, Tahoe 
@@
 % like \Treebeard{} does.
-
-RAPIDS FIL\cite{FIL} is a library that implements decision tree inference on GPUs
-and is the most widely used production system for decision tree inference. While 
-FIL does implement some heuristics to pick a good configuration for every model, 
-these techniques are limited and the library essentially uses a single strategy 
+RAPIDS FIL\cite{FIL}, the most widely used GPU library for decision tree inference
+implements some heuristics to pick a good configuration for every model, 
+but these techniques are limited and the library uses a single strategy 
 and in-memory representation for all models. XGBoost \cite{XGBoost} also implements GPU 
@@
 and improve cache performance of decision tree ensembles on CPUs\cite{CacheConscious1, CacheConscious2}
-have been proposed in prior work. However, these systems are limited to CPUs.
+have been proposed in prior work.
 Some systems have been proposed to parallelize decision tree training 
@@
 % target GPUs is not possible.
-
 \TreebeardOLD{} is the work most closely related to \Treebeard{}. While we 
@@
 % order to generate code for multiple target processors.
-Specifically, we introduce the scheduling language and schedule exploration 
-while also enhancing the IRs and support for parallelizing 
-across trees through the implementation of a novel MLIR reduction dialect.
+% Specifically, we introduce the scheduling language and schedule exploration 
+% while also enhancing the IRs and support for parallelizing 
+% across trees through the implementation of a novel MLIR reduction dialect.
 % It cannot perform any of the optimizations \Treebeard{} is designed to perform. 
-
 Hummingbird\cite{Hummingbird} is a compiler that compiles traditional ML models
 to tensor operations, thereby enabling them to be run on tensor-based frameworks like
-TensorFlow\cite{TensorFlow}. Hummingbird can target both CPUs
-and GPUs, but, as was shown earlier~\cite{Treebeard},
-tensor operations are not the most efficient way to implement decision tree inference
-and the performance of Hummingbird is significantly lower than 
-that of other frameworks.
+TensorFlow\cite{TensorFlow}. While it can target both CPUs
+and GPUs, its performance is lower than that of other frameworks~\cite{Treebeard}.
 
@@
 inference. However, this virtual machine is itself implemented by 
-hand on different target processors. This is clearly 
-more expensive than \Treebeard{}'s approach.
+hand on different target processors.
+%This is clearly more expensive than \Treebeard{}'s approach.
 % Additionally, even though they perform layout optimizations, their system does
@@
 from the schedule was pioneered by Halide and has since been adopted 
-by several other systems~\cite{TVM,Tiramisu,GraphIt}. However, to 
-the best of our knowledge, \Treebeard{} is the first system to design
-a scheduling language for decision tree inference optimization
-and to build a system capable of state-of-the-art performance 
-across different processors.
+by several other systems~\cite{TVM,Tiramisu,GraphIt}. 
+% However, to 
+% the best of our knowledge, \Treebeard{} is the first system to design
+% a scheduling language for decision tree inference optimization
+% and to build a system capable of state-of-the-art performance 
+% across different processors.
 % {\sc {Cortex}} transforms 
@@
 have also been developed.
+However, \Treebeard{} is the first system that provides state-of-the-art performance
+across targets by implementing a scheduling language for decision tree inference.
 % BLIS\cite{BLIS} and ATLAS\cite{atlas_sc98}
@@
 \emph{Reductions:} CUB\cite{CUB} and Thrust\cite{Thrust} are libraries 
-that implement high-performance parallel reductions on GPUs. While they 
-provide highly-tuned implementations to perform large reductions, 
-it is not possible to fuse these functions with other computations  
+that implement high-performance parallel reductions on GPUs. 
+However, it is not possible to fuse these functions with other computations  
 as required in \Treebeard{}. Reddy et. al.~\cite{ChandanReduction}
@@
 to express reductions and to represent and optimize them using the polyhedral 
-framework. It is not clear how these techniques can be fused with 
-other computations in arbitrary loop nests as required in \Treebeard{}.
-Additionally, their system does not express the hierarchical nature of 
+framework. 
+% It is not clear how these techniques can be fused with 
+% other computations in arbitrary loop nests as required in \Treebeard{}.
+%Additionally, 
+Their system does not express the hierarchical nature of 
 reductions and also only targets GPUs. Suriana et. al.~\cite{HalideReductions}

### 2024-04-18 docs/SOSP 2024/schedule.tex
Shrink tables and algorithm listing
pass

@@
 % make portable performance impossible. 
-We design a scheduling language for \Treebeard{}
-to address the problems discussed in Section \ref{sec:motivation}. The scheduling 
-language provides an abstract way to specify loop structure and 
+This section discusses the scheduling language that enables exploration of different strategies for decision tree inference. The language provides an abstract way to specify loop structure and 
 other optimizations as an input to the compiler. The specified 
@@
 \begin{table}[htb]
+  \footnotesize
   \centering

### 2024-04-19 docs/SOSP 2024/motivation.tex
pass

@@
 defer a detailed explanation of the scheduling language to Section~\ref{sec:schedule}. 
-The loop over the trees is split into two -- one that
+In this schedule, the loop over the trees is split into two -- one that
 iterates over the first two trees (Trees 1 and 2 with depth 1) and 
@@
 and involves branching to check if a leaf has been reached. 
-One way to avoid this as done in this schedule is to unroll the 
+One way to avoid this, as done in this schedule, is to unroll the 
 tree walks for each tree. 
@@
   \end{subfigure}
-  \caption{Possible schedules to compile the motivating example and the generated \Treebeard{} IR.}
+  \caption{Some possible schedules to compile the motivating example and the generated \Treebeard{} IR.}
   \label{Fig:Schedules}
@@
 % \end{lstlisting}
-This schedule generates an inference function that runs on the GPU as shown in 
-the IR in Figure \ref{fig:scheduleb}. 
-The inference routine processes one input row per thread block (since the \op{batch}
-loop is mapped directly to \op{grid.x}).
-It also splits the trees into two sets by tiling the \op{tree} loop.
-Each of the two sets is processed in parallel. We unroll the tree walks 
-for each tree.
+The corresponding GPU inference routine is shown in the same figure. 
+This schedule generates a routine that processes one input row per thread block (\op{batch}
+loop is mapped to \op{grid.x}).
+The schedule also tiles the \op{tree} loop into a nest of two loops with indices $t0$ and $t1$. 
+It then additionally parallelizes across the $t1$ loop.
+Finally, it splits $t0$ and unrolls the walks for each tree depth.
+%It also splits the trees into two sets by tiling the \op{tree} loop.
+%Each of the two sets is processed in parallel. We unroll the tree walks 
+%for each tree.
 % The IR generated is as follows. 
@@
 % reduced across the \op{t1} dimension to get the final result.
-Note that while unrolling helps avoid branching, it likely increases 
+Note that while unrolling helps avoid branching, it increases 
 the total amount of computation. Another option is to not unroll 
@@
 and the best choice depends on the characteristics of the model
-and micro-architectural features like register file size, handling 
-of branch divergence etc.~\cite{FungMicro, MilindDivergence}.
-
+and micro-architectural features like register file size and handling 
+of branch divergence~\cite{FungMicro, MilindDivergence}.
+These schedules just use a combination of 4-8 constructs and already generate schedules 
+that are different from what library based systems like XGBoost, Tahoe and RAPIDs FIL use. 
+As one can imagine several other schedules with different trade-offs can be generating using the constructs.
 % As is evident from these examples, it is possible to optimize 

### 2024-04-19 docs/SOSP 2024/schedule.tex
pass

@@
 % make portable performance impossible. 
-This section discusses the scheduling language that enables exploration of different strategies for decision tree inference. The language provides an abstract way to specify loop structure and 
+This section presents all constructs in the scheduling language. 
+The language provides an abstract way to specify loop structure and 
 other optimizations as an input to the compiler. The specified 
@@
   \caption{\label{Tab:Optimizations} List of optimization directives in \Treebeard{}'s scheduling language. 
-  We use \emph{index variable} and \emph{loop} interchangeably in descriptions for clarity of exposition.}
+ % We use \emph{index variable} and \emph{loop} interchangeably in descriptions for clarity of exposition.
+ }
 \end{table}
@@
   \caption{\label{Tab:ReductionOpts} List of reduction optimization directives in \Treebeard{}'s scheduling language. 
-  We use \emph{index variable} and \emph{loop} interchangeably in descriptions for clarity of exposition.}
+  %We use \emph{index variable} and \emph{loop} interchangeably in descriptions for clarity of exposition.
+  }
 \end{table}
@@
 XGBoost\cite{XGBoost} implements inference on the CPU by going 
-over a fixed number of rows (64 in the previous version)
+over a fixed number of rows (64 in the current version)
 for every tree and then moving to the next tree. It moves to the 
@@
 set of rows. Different sets of rows are processed in parallel.
-This is expressed in \Treebeard{}'s scheduling language as:
+This is expressed in \Treebeard{}'s as:
 \begin{lstlisting}[style=c++]
@@
 In the \emph{direct method}~\cite{Tahoe}, a single GPU thread walks all trees
-for a given input row. The schedule for this strategy is as follows.
+for a given input row. The schedule for this strategy is:
 \begin{lstlisting}[style=c++]
@@
 Then, a thread block wide reduction is performed to compute 
-the prediction. The schedule for this strategy is as follows.
+the prediction. The schedule for this strategy is:
 \begin{lstlisting}[style=c++]
@@
 gpuDimension(tree, block.x)
-cache(batch)
+reduce(batch)
 \end{lstlisting}

### 2024-04-19 docs/SOSP 2024/results.tex
pass

@@
 Figure~\ref{fig:randomModels4060} shows two representative samples of the results, at batch sizes 512 and 4096 on all models, on RTX 4060. 
-Each plot is a 3D histogram of the speedup of \Treebeard{} over RAPIDS, with the x-axis representing the number of trees, the y-axis the number of features and the z-axis the speedup. 
+Each plot shows the speedup of \Treebeard{} over RAPIDS along the z-axis, with the x-axis representing the number of trees and the y-axis the number of features. 
 While the exact trends at different batch sizes vary it can be seen that \Treebeard{} consistently outperforms RAPIDS by $1.5-8\times$.
+Along the tree dimension we find that the speedup is very high with fewer trees and stabilizes at around $2\times$ from 300 onwards.
+One of our real-world benchmarks $letters$ has 26K trees and still achieves a speedup of around $2\times$ over RAPIDS.
 %\Treebeard{} consistently outperforms RAPIDS on all models and batch sizes tested.

### 2024-04-19 docs/SOSP 2024/results.tex
Minor edits to results

@@
 
-We evaluate \Treebeard{} on four different target processors, an NVIDIA RTX 4060 GPU (8GB RAM, CUDA 11.5),
-an NVIDIA T400 GPU (2GB RAM, CUDA 11.5), an AMD MI210 GPU (64GB RAM, ROCm 4.5\TODO{Check the version}) and an 
+We evaluate \Treebeard{} on four different target processors, an NVIDIA RTX 4060 GPU (8GB RAM, CUDA 11.8),
+an NVIDIA T400 GPU (2GB RAM, CUDA 11.5), an AMD MI210 GPU (64GB RAM, ROCm 6.0.2) and an 
 Intel Core i9-11900K CPU (16 virtual cores, 128 GB RAM). We compare \Treebeard{} with 
@@
   \includegraphics[width=0.75\linewidth]{figures/geomean_speedup_4060_kernel_time_total_time.png}
-  \caption{\Treebeard{} vs RAPIDs and Tahoe kernel time and total time speedup on NVIDIA RTX 4060}
+  \caption{\Treebeard{} vs RAPIDs and Tahoe kernel time and total time speedup (geomean over real-world 
+  benchmarks) across batch sizes on NVIDIA RTX 4060}
   \label{Fig:TBvsRAPIDsTahoe_4060_Speedup}
@@
   \includegraphics[width=0.75\linewidth]{figures/geomean_speedup_T400_kernel_time.png}
-  \caption{\Treebeard{} vs RAPIDs and Tahoe kernel time and total time speedup on NVIDIA T400.}
+  \caption{\Treebeard{} vs RAPIDs and Tahoe kernel time and total time speedup (geomean over real-world 
+  benchmarks) across batch sizes on NVIDIA T400.}
   \label{Fig:TBvsRAPIDsTahoe_T400_Speedup}
@@
 Compared to RAPIDS, \Treebeard{} is about $4\times$ faster at batch size 512. The relative performance of RAPIDs improves 
-with batch size, but \Treebeard{} is still faster by $1.5-2\times$ even at very high ($16K$) batch sizes. 
+with batch size, but \Treebeard{} is still faster by $1.5-2\times$ even at very high batch sizes ($16k$). 
 The plot also shows that the speedups are significant even when data needs to be transferred to the GPU and results need to
@@
 \subsubsection*{Synthetic models}
-To establish that \Treebeard{} can consistently find betters schedules, we performed an exhaustive comparison between \Treebeard{} and RAPIDS on all 700 randomly generated models.   
+To establish that \Treebeard{} can consistently find betters schedules, we performed an exhaustive comparison between \Treebeard{} and RAPIDS on 700 randomly generated models.   
 Figure~\ref{fig:randomModels4060} shows two representative samples of the results, at batch sizes 512 and 4096 on all models, on RTX 4060. 
 Each plot shows the speedup of \Treebeard{} over RAPIDS along the z-axis, with the x-axis representing the number of trees and the y-axis the number of features. 
-While the exact trends at different batch sizes vary it can be seen that \Treebeard{} consistently outperforms RAPIDS by $1.5-8\times$.
-Along the tree dimension we find that the speedup is very high with fewer trees and stabilizes at around $2\times$ from 300 onwards.
-One of our real-world benchmarks $letters$ has 26K trees and still achieves a speedup of around $2\times$ over RAPIDS.
+While the exact trends at different batch sizes vary it can be seen that \Treebeard{} consistently outperforms RAPIDS by $1.5-8\times$. Along the tree 
+dimension we find that the speedup is very high with fewer trees and stabilizes at around $2\times$ from 300 onwards. We note that \Treebeard{} continues to scale well 
+when the number of trees are increased even further. It achieves a speedup of around $2\times$ compared to RAPIDS on \op{letters}, a real-world benchmark
+with 26K trees.
 %\Treebeard{} consistently outperforms RAPIDS on all models and batch sizes tested.
@@
 and reduction optimizations are crucial for performance. We observe that all three representations 
-are necessary, array representation is used 60\% of the time, sparse 30\% of the time and reorg 10\% of the time. 
+are necessary, array representation is used 60\% of the time, sparse 30\% of the time and reorg 
+10\% of the time for the best schedules on the NVIDIA 4060.  
 Shared reduction significantly improves performance especially for multi-class models. For example, 
-the \op{letters} benchmark achieves a speedup of 1.5$\times$ with shared reduction. 
+the \op{letters} and \op{covtype} benchmarks achieves speedups of 1.5$\times$ with shared reduction. 
 Shared reduction increases the memory pressure (on shared memory) and not all schedules use it.
@@
   \includegraphics[width=0.75\linewidth]{figures/speedup_vs_norm_time_line_graph_4060.png}
-  \caption{Speedup of schedule found by the heuristic vs best schedule found through exhaustive search on NVIDIA RTX 4060 and 
-  normalized schedule exploration heuristic exploration time speedup w.r.t full schedule exploration}
+  \caption{Speedup of schedules found by the heuristic vs best schedule found through exhaustive search on NVIDIA RTX 4060 and 
+  heuristic schedule exploration time normalized w.r.t full schedule exploration time \TODO{Change orange line label 
+  to ``Heuristic's Normalized Exploration Time''} \TODO{Change other label to say ``Heuristic Schedule vs Best 4060 Schedule''}}
   \label{Fig:HeuristicVsFullExplore_Speedup}
@@
   \caption{Geomean speedup of schedule found by exploration heuristic on NVIDIA T400 and AMD MI2160 vs 
-  schedule found by the heuristic on NVIDIA RTX 4060 across batch sizes\TODO{Change the plot labels to say vs 4060 heuristic schedule}}
+  the heuristic schedule on NVIDIA RTX 4060 across batch sizes\TODO{Change the plot labels to say vs 4060 heuristic schedule}}
   \label{Fig:AutotuningSpeedupvs4060Sched}
@@
 The enhancements made to the compiler enable \Treebeard{} to explore additional schedules on the CPU
-than \TreebeardOLD{}. 
+compared to \TreebeardOLD{}. 
 In particular, we find that the ability to parallelize across trees improves performance 
 significantly at small batch sizes. At batch size 32, we find that the geomean speedup over 
-all benchmark models is 2.2$\times$ with a max speedup of 5$\times$. At batch size 64, the average speedup
+all 8 real-world benchmark models is 2.2$\times$ with a max speedup of 5$\times$. At batch size 64, the average speedup
 is 1.1$\times$ with a max speedup of 2$\times$. At batch size 32, parallelizing across trees is faster 

### 2024-04-19 docs/SOSP 2024/intro.tex
Edits to intro and motivation

@@
 % We observe that while scalable performance on neural network models has received significant attention, decision tree inference has not seen the same level of acceleration.
-A recent survey~\cite{KaggleSurvey} found that about three quarters of data scientists use decision tree
-based models. An analysis of ML workloads at a large scale web company found that these models are most widely used~\cite{LookingGlass}. 
+The Kaggle AI report for 2023 \cite{ABCD} highlights the continuing \emph{dominance of 
+gradient boosted trees as the algorithm of choice} for tabular data. It also estimates that \emph{between
+50\% and 90\% of practicing data scientists use tabular data as their primary type of data in their professional setting.}
+% An analysis of ML workloads at a large scale web company found that these models are widely used~\cite{LookingGlass}.
+This has also been observed in other surveys~\cite{KaggleSurvey,LookingGlass}.
 Recent work has noted that the cost of inference is the most critical factor in the overall cost of 
@@
 This paper presents \Treebeard{}, a novel schedule guided compilation infrastructure
-for decision tree inference on multiple target hardware. \Treebeard{} is able to 
+for decision tree inference on multiple hardware targets. \Treebeard{} is able to 
 generate high-performance code for decision tree inference by exploring a large optimization space. 
@@
 the optimization space. We demonstrate that the language is 
-sufficient to express the various optimizations proposed by prior work
-and that our schedule exploration heuristic can quickly find a near 
+is capable of expressing all optimizations proposed by prior work and more.
+Further, our schedule exploration heuristic can quickly find a near 
 optimal schedule for the model being compiled. 
 %% Further we generalizes them and incorporates several new optimizations. 
-\TODO{RG: it would be good to expand on this and also talk about the retargetable component. 
-We could also say that the schedule framework and the retargetable compiler work in an 
-interwined manner, each benefitting from the other.}  
+% \TODO{RG: it would be good to expand on this and also talk about the retargetable component. 
+% We could also say that the schedule framework and the retargetable compiler work in an 
+% interwined manner, each benefitting from the other.}  
 The second component of \Treebeard{} is a \emph{retargetable} multi-level compiler that can generate efficient code 
 for any specified schedule for both CPUs and GPUs. For this purpose, we re-architect \TreebeardOLD{} to
-support schedule-guided code generation, and incorporate several new optimizations. These two components 
-of the proposed \Treebeard{} compiler infrastructure are interwined, each benefitting from the other.
+support schedule-guided code generation, and incorporate several new optimizations that are critical to target GPUs. 
+These two components of the proposed \Treebeard{} compiler infrastructure are intertwined, each benefitting from the other.
 %Last, to address the challenge of selecting a near-optimal schedule from the vast optimization space, 
@@
 %we design and implement a heuristic that is able to quickly find high-performance schedules for the model being compiled.
-\TODO{Performance evaluation summary}
+
+We evaluated \Treebeard{} on a large number of synthetic and real-world 
+benchmark models and compared against state-of-the-art systems RAPIDS~\cite{RAPIDS}, Tahoe~\cite{Tahoe}
+, XGBoost~\cite{XGBoost} and \TreebeardOLD{}~\cite{Treebeard}. While these systems can only run on 
+specific hardware, \Treebeard{} outperforms all these systems while providing portable performance across 
+a range of target hardware. The geomean speedup (across all benchmarks) of \Treebeard{} over
+Tahoe and RAPIDS is in the range $2$--$5\times$ over several batch sizes. The geomean speedup over
+XGBoost is more than $10\times$. \Treebeard{} also enables better scalability on CPUs by parallelizing 
+across multiple dimensions. We also demonstrate that \Treebeard{}'s scheduling heuristic is able to 
+find near-optimal schedules in less than 30 seconds on average.
+
 %% RG: This may not be needed -- can be removed to save space!
 %% \TODO{Oganization}
-\subsection{Contributions}
 \begin{itemize}
-  \item We present the design of a multi-target compiler infrastructure for decision tree 
-   inferencing  and implement several optimizations within this framework.
-   We are also the first to implement an optimizing compiler for decision tree 
-   inference on GPUs. \TODO{AP: Given that there is Hummingbird, can we really say this?}
-  \item We identify that an extensive optimization space exists for the problem of 
-  decision tree inference. We design a scheduling language that 
-  allows us to effectively represent this solution space abstractly. This 
-  scheduling language is expressive enough to represent a wide range of  
-  implementation strategies,  %%.  proposed by prior work.
-%% RG: added CPUs and GPUs. 
-  such as different data layouts, caching strategies, parallel reduction schemes, 
-  that work across CPUs and GPUs. 
-  \item To the best of our knowledge, we perform the first extensive characterization
-  of the optimization space for decision tree inference 
-  on GPUs. Using some of the characteristics we identify, we design and implement 
-  a heuristic that is able to quickly find high-performance 
-  schedules for the model being compiled. 
-  \item We design and implement a general framework for expressing and optimizing
-  reductions within MLIR. To the best of our knowledge, this is the first
-  such framework.
-  \item We evaluate our implementation by comparing it against RAPIDs and Tahoe, 
-  the state-of-the-art decision tree inference frameworks for GPU and 
-  report significant speedups. We also show that our compiler can effectively 
-  target different GPUs, including both NVIDIA and AMD GPUs.
+  \item We identify that there are many different ways to implement  
+  decision tree inference and that the best implementation depends both on the model
+  and the target. We design a scheduling language that 
+  allows us to represent this solution space. 
+%   This 
+%   scheduling language is expressive enough to represent a wide range of  
+%   implementation strategies,  %%.  proposed by prior work.
+% %% RG: added CPUs and GPUs. 
+%   such as different data layouts, caching strategies, parallel reduction schemes, 
+%   that work across CPUs and GPUs. 
+  \item We propose a retargetable compiler that can generate code as specified by 
+  the schedule. The compiler also includes new abstractions for caching,
+  reduction and representing models in memory. These are critical to generate 
+  efficient GPU code.
+  \item The scheduling language can represent an unbounded number of schedules. 
+  We design and implement techniques to search over this space and find high-performance 
+  schedules in tens of seconds. 
+  \item We exhaustively evaluate \Treebeard{} and report that it achieves 
+  portable performance across a range of hardware targets and consistently outperforms
+  state-of-the-art decision tree inference frameworks for both CPU and GPU.
 \end{itemize}

### 2024-04-19 docs/SOSP 2024/schedule.tex
Edits to intro and motivation

@@
 In the \emph{shared data} strategy~\cite{Tahoe}, a thread block walks all the trees 
-for a given row in parallel. 
+for a given row in parallel. Input rows are cached in shared memory.
 % If threads walk multiple trees, each thread accumulates partial results. 
-Then, a thread block wide reduction is performed to compute 
-the prediction. The schedule for this strategy is:
+% Then, a thread block wide reduction is performed to compute 
+% the prediction.
+The schedule for this strategy is:
 \begin{lstlisting}[style=c++]
@@
 gpuDimension(tree, block.x)
-reduce(batch)
+cache(batch)
 \end{lstlisting}

### 2024-04-19 docs/SOSP 2024/results.tex
Update graph legends

@@
   \caption{Speedup of schedules found by the heuristic vs best schedule found through exhaustive search on NVIDIA RTX 4060 and 
-  heuristic schedule exploration time normalized w.r.t full schedule exploration time \TODO{Change orange line label 
-  to ``Heuristic's Normalized Exploration Time''} \TODO{Change other label to say ``Heuristic Schedule vs Best 4060 Schedule''}}
+  heuristic schedule exploration time normalized w.r.t full schedule exploration time}
   \label{Fig:HeuristicVsFullExplore_Speedup}
@@
   \caption{Geomean speedup of schedule found by exploration heuristic on NVIDIA T400 and AMD MI2160 vs 
-  the heuristic schedule on NVIDIA RTX 4060 across batch sizes\TODO{Change the plot labels to say vs 4060 heuristic schedule}}
+  the heuristic schedule on NVIDIA RTX 4060 across batch sizes}
   \label{Fig:AutotuningSpeedupvs4060Sched}

### 2024-04-19 docs/SOSP 2024/intro.tex
Minor edits to intro

@@
 
-Decision forest models are composed of a large collection of decision trees (100-1000), and inference involves 
+Decision forest models are composed of a large collection of decision trees (typically 100--1000), and inference involves 
 traversing down each tree in the forest and aggregating the predictions. Inference is typically done in a batched
@@
 %% \TODO{Oganization}
+In summary, this paper makes the following contributions.
 \begin{itemize}

### 2024-04-19 docs/SOSP 2024/exploring.tex
Shrink tables and algorithm listing
Shrink some figures
Add representation values to table
pass

@@
 that the space (at-least) covers the known strategies published in prior work.
-Specifically, we use 4 numeric parameters and 4 boolean parameters listed in Table \ref{tab:schedparams}. 
+Specifically, we use 4 numeric parameters, 4 boolean parameters and one ternary parameter as listed in Table \ref{tab:schedparams}.
+We picked the specific values listed through experimentation.
 %Note that even this template has a large number ($1728$) of schedules, and requires further pruning as we discuss later.
@@
 a single thread?}.  (the remaining primitives from Table~\ref{Tab:Optimizations}).  
-The final parameter determines what reduction options to use (Table~\ref{Tab:ReductionOpts}).
+The last two parameters determine reduction option (Table~\ref{Tab:ReductionOpts}) and the layout{\footnote{This does not have a schedule primitive.
+We always explore all three layout options.}} to use.
 %Table \ref{tab:schedparams} lists the parameter values we tried. 
@@
 \begin{table}[htb]
+  \footnotesize
   \centering
-  \resizebox{\linewidth}{!}{
+  % \resizebox{\linewidth}{!}{
   \begin{tabularx}{\linewidth}{c | c }
@@
     \textbf{Shared memory reduction} & $\{True, False\}$ \\
+    \textbf{Representation} & $\{array, sparse, reorg\}$ \\
     \bottomrule
   \end{tabularx}
-  }
+  % }
   \vskip 5pt
@@
   \centering
-  \includegraphics[width=\linewidth]{figures/normalized_kernel_histogram_lt5.png}
+  \includegraphics[width=0.7\linewidth]{figures/normalized_kernel_histogram_lt5.png}
   \caption{Distribution of normalized execution times for all benchmark models
@@
 times normalized w.r.t fastest time for that model).
-As can be seen very few schedules perform close to the best while a vast majority of
-schedules perform poorly.
+As can be seen very few schedules perform close to the best while a vast majority perform poorly.
 
 Exhaustive exploration over this bounded space ($5184$ schedules) is still very expensive, taking
-anywhere between thirty minutes to a few hours to explore the entire space for a given model. 
+ between $30-600$ minutes to explore the entire space for a given model. 
 %Exploring the space for every model is infeasible.
@@
  Some configurations do not expose sufficient parallelism, we prune these out.
-  For example when the batch size is small, it does not make sense to have a many rows
+  For example when the batch size is small, it does not make sense to have many rows
   per thread block or to partition the trees across a few threads. 
@@
   \label{alg:heuristic}
+  \small{
   \begin{algorithmic}[1]
@@
   \end{algorithmic} 
+  }
 \end{algorithm}

### 2024-04-19 docs/SOSP 2024/abstract.tex
Update abstract. Add page numbers

@@
 
-  This paper is motivated by the growing demand for the increased performance
-  of machine learning applications on different hardware platforms
-  including CPUs and GPUs. We focus on accelerating the inference of decision tree based models, 
-  which are the most popular models for tabular data. Existing solutions do not achieve 
-  the highest possible performance because they do not explore different optimization
-  configurations. And since these systems are hand-written, they are not portable either.
+  The rapid proliferation of machine learning together with the accelerating evolution of 
+  the hardware ecosystem has led to a surge in the demand for model inference on a variety
+  of hardware. This paper is motivated by the problems encountered when targeting 
+  inference of decision tree based models, the most popular models on tabular data,
+  to run at peak performance on diverse CPU and GPU targets. Existing solutions are neither
+  portable nor achieve the best possible performance for the specific hardware they 
+  target because they do not explore different optimization configurations.
 
-  We address these problems by designing \Treebeard{}, a \emph{schedule-guided},
-  \emph{retargetable} compiler infrastructure for decision tree based models.  
+  We present \Treebeard{}, a \emph{schedule-guided},
+  \emph{retargetable} compiler for decision tree based models
+  that searches over several optimization choices and automatically generates 
+  high-performance inference routines for CPUs and GPUs.  
   \Treebeard{} has two core components. The first is a scheduling language 
-  that encapsulates the large optimization space for decision tree inference, 
-  and techniques to efficiently explore this space. \TODO{Change large optimization space}.
-  The second is an optimizing retargetable multi-level compiler that 
-  can generate code for any specified schedule. \Treebeard{}'s retargetability 
-  is based not only on being able to generate code for different target architectures (CPU vs. GPU),
-  but also on its ability to use different data layouts, caching strategies,
-  parallel reduction schemes etc. To accomplish this level of configurability, we re-architect 
-  and significantly extend the open-source \TreebeardOLD{} CPU compiler
-  to support (i) schedule-guided compilation, (ii) retargetable GPU code
-  generation, and (iii) GPU-specific optimizations. 
+  that encapsulates the optimization space for decision tree inference, 
+  and techniques to efficiently explore this space. 
+  The second is an optimizing retargetable compiler that 
+  can generate code for any specified schedule. \Treebeard{}'s ability to use different data layouts, 
+  loop structures and caching strategies enables it to achieve portable performance.   
+  % To accomplish this level of configurability, we re-architect 
+  % and significantly extend the open-source \TreebeardOLD{} CPU compiler
+  % to support (i) schedule-guided compilation, (ii) retargetable GPU code
+  % generation, and (iii) GPU-specific optimizations. 
   %The scheduler and retargetable code generator components work hand-in-glove, synergizing and mutually benefiting from each other.
 
-  We demonstrate that \Treebeard{} can generate high-performance inference 
-  code, for several hundred decision tree models across different batch sizes
-  and target architectures. Our scheduling heuristic is able to quickly find near-optimal schedules 
-  \TODO{[how do we argue that the schedule is near-optimal? Do we have some 
-  exptl. evidence that we can show in the results section?]}
-  schedule while searching over a small number ($\sim$50) of schedules.
+  We evaluate \Treebeard{} on several hundred decision tree models across different batch sizes
+  and target architectures. We find that our schedule exploration strategy is able to quickly
+  find near-optimal schedules.
   In terms of performance, \Treebeard{} generated code is an order of magnitude faster than XGBoost and
-  about 2-3$\times$ faster on average than RAPIDs FIL and Tahoe. While these systems only 
-  target NVIDIA GPUs, \Treebeard{} achieves competent performance on AMD GPUs as well. 
-  On CPUs, \Treebeard{} achieves better scaling compared to \TreebeardOLD{}.
-  For models where \TreebeardOLD{} was only able to achieve diminishing returns with an 
-  increasing number of threads, \Treebeard{} is able to scale linearly with the number of threads.
-  \TODO{(numbers for CPU performance?)}
+  about 2-5$\times$ faster on average than RAPIDS FIL and Tahoe over several batch sizes.
+  While these systems only target NVIDIA GPUs, \Treebeard{} achieves competent performance on AMD GPUs as well. 
+  On CPUs, \Treebeard{} is able to outperform \TreebeardOLD{} by up to $5\times$ by utilizing 
+  additional sources of parallelism at small batch sizes.
 \end{abstract}

### 2024-04-19 docs/SOSP 2024/results.tex
Edit results. Reduce algorithm size

@@
 Intel Core i9-11900K CPU (16 virtual cores, 128 GB RAM). We compare \Treebeard{} with 
-four other systems, NVIDIA RAPIDs v23.10, Tahoe, XGBoost v1.7.6 and \TreebeardOLD{} CPU. 
+four other systems, NVIDIA RAPIDS\cite{RAPIDS} v23.10, Tahoe\cite{Tahoe}, XGBoost\cite{XGBoost}
+v1.7.6 and \TreebeardOLD{}\cite{Treebeard} CPU. 
 We measure both kernel time and total time (including data transfer to the GPU and results back) 
-for RAPIDs and \Treebeard{}. Tahoe only allows us to measure the kernel time since it is written
+for RAPIDS and \Treebeard{}. Tahoe only allows us to measure the kernel time since it is written
 as an executable that performs inference repeatedly on the same data that is transferred to the GPU once.
+For \Treebeard{}, we use schedules found using the schedule exploration
+heuristic (Section \ref{sec:exploring}) unless otherwise specified.
 
@@
 Each tree has leaves at depths 2 to a maximum depth of 6, 7 or 8.
-For \Treebeard{}, we use schedules found using the schedule exploration
-heuristic (Section \ref{sec:exploring}) unless otherwise specified.
 
 \subsection{Performance comparisons on NVIDIA GPUs}
+
+\subsubsection*{Real-world models}
+Figure~\ref{Fig:TBvsRAPIDSTahoe_4060_Speedup} shows the geomean speedup of \Treebeard{} over RAPIDS and Tahoe at different batch sizes on RTX 4060. 
+We do not show results for XGBoost since the speedups are an order of magnitude higher ($9\times$-$40\times$) and too large to fit on the same graph.  
+The plot has lines for kernel time and total time speedup over RAPIDS and kernel time speedup over Tahoe. As can been seen 
+\Treebeard{} is uniformly faster than Tahoe{\footnote{Tahoe does not support multiclass models. To enable comparison, we ran 
+the multiclass models (\op{covtype} and \op{letters}) as regression models only with Tahoe. We also noticed that some variants
+of Tahoe produce wrong results (as reported by its own tests) for \op{letters} and \op{year}. In these cases, we pick the 
+time of the fastest variant that gives the correct results.}} by $2-2.5\times$ at all batch sizes. 
+Compared to RAPIDS, \Treebeard{} is about $4\times$ faster at batch size 512. The relative performance of RAPIDS improves 
+with batch size as RAPIDS is optimized for larger batches, but \Treebeard{} is still faster by $1.5-2\times$ even at very high batch sizes ($16k$). 
+The plot also shows that the speedups are significant even when the overhead of data transfer to and from the GPU
+is included. They are lower than kernel time speedup as both systems have a constant additional transfer overhead.
 
@@
   \includegraphics[width=0.75\linewidth]{figures/geomean_speedup_4060_kernel_time_total_time.png}
-  \caption{\Treebeard{} vs RAPIDs and Tahoe kernel time and total time speedup (geomean over real-world 
+  \caption{\Treebeard{} vs RAPIDS and Tahoe kernel time and total time speedup (geomean over real-world 
   benchmarks) across batch sizes on NVIDIA RTX 4060}
-  \label{Fig:TBvsRAPIDsTahoe_4060_Speedup}
+  \label{Fig:TBvsRAPIDSTahoe_4060_Speedup}
 \end{figure}
@@
 %   \includegraphics[width=0.75\linewidth]{figures/geomean_speedup_4060_bar_graph.png}
-%   \caption{\Treebeard{} vs RAPIDs and Tahoe kernel time and total time speedup on NVIDIA RTX 4060 for individual benchmarks (geomean over all batch sizes).
+%   \caption{\Treebeard{} vs RAPIDS and Tahoe kernel time and total time speedup on NVIDIA RTX 4060 for individual benchmarks (geomean over all batch sizes).
 %   \TODO{Add bar for total time. Add whiskers to each bar to show range of speedups.}}
-%   \label{Fig:TBvsRAPIDsTahoe_4060_Speedup}
+%   \label{Fig:TBvsRAPIDSTahoe_4060_Speedup}
 % \end{figure}
@@
   \includegraphics[width=0.75\linewidth]{figures/geomean_speedup_T400_kernel_time.png}
-  \caption{\Treebeard{} vs RAPIDs and Tahoe kernel time and total time speedup (geomean over real-world 
+  \caption{\Treebeard{} vs RAPIDS and Tahoe kernel time and total time speedup (geomean over real-world 
   benchmarks) across batch sizes on NVIDIA T400.}
-  \label{Fig:TBvsRAPIDsTahoe_T400_Speedup}
+  \label{Fig:TBvsRAPIDSTahoe_T400_Speedup}
 \end{figure}
@@
   \hfill
-  \caption{\label{Fig:KernelTimeIndividualBenchmarks4060}Kernel time speedup of \Treebeard{} vs RAPIDs on NVIDIA RTX 4060. Numbers on the bars are 
-  inference times per sample in $\mu$s for RAPIDs and Tahoe.}
+  \caption{\label{Fig:KernelTimeIndividualBenchmarks4060}Kernel time speedup of \Treebeard{} vs RAPIDS on NVIDIA RTX 4060. Numbers on the bars are 
+  inference times per sample in $\mu$s for RAPIDS and Tahoe.}
 \end{figure*}
-
-\subsubsection*{Real-world models}
-We performed a detailed performance comparison between \Treebeard{}, Tahoe, RAPIDS and XGBoost on the real-world benchmarks.
-Figure~\ref{Fig:TBvsRAPIDsTahoe_4060_Speedup} shows the geomean speedup of \Treebeard{} at different batch sizes on RTX 4060. 
-We do not show results for XGBoost since the speedups are an order of magnitude higher and too large to fit on the same graph.  
-The plot has lines for kernel time and total time speedup over RAPIDS and kernel time speedup over Tahoe. As can been seen 
-\Treebeard{} is uniformly faster than Tahoe{\footnote{Tahoe does not support multiclass models. To enable comparison, we ran 
-the multiclass models (\op{covtype} and \op{letters}) as regression models only with Tahoe. We also noticed that some variants
-of Tahoe produce wrong results (as reported by its own tests) for \op{letters} and \op{year}. In these cases, we pick the 
-time of the fastest variant that gives the correct results.}} by $2-3\times$ at all batch sizes. 
-Compared to RAPIDS, \Treebeard{} is about $4\times$ faster at batch size 512. The relative performance of RAPIDs improves 
-with batch size, but \Treebeard{} is still faster by $1.5-2\times$ even at very high batch sizes ($16k$). 
-The plot also shows that the speedups are significant even when data needs to be transferred to the GPU and results need to
-be transferred back. They are lower than kernel time speedup as both systems have a constant additional transfer overhead.
 
@@
 \Treebeard{} is able to find better schedules than both RAPIDS and Tahoe on each of the benchmarks. It consistently
-outperforms both systems, achieving speedups in the range $1.1-12\times$, with about half the benchmarks achieving a speedup of $2\times$ or more  over both baselines at batch size 8192.
+outperforms both systems, achieving a speedup of upto $12\times$ an average speedup of over $2\times$.
+For about half the benchmarks, the speedup is $2\times$ or more over both baselines at batch size 8192.
+
+% \begin{comment}
+\subsubsection*{Synthetic models}
+To establish that \Treebeard{} can consistently find betters schedules, we performed an exhaustive comparison between \Treebeard{} and RAPIDS on 700 randomly generated models.   
+Figure~\ref{fig:randomModels4060} the results at batch sizes 512 and 4096 on all models, on RTX 4060. 
+Each plot shows the speedup of \Treebeard{} over RAPIDS along the z-axis, with the x-axis and y-axis representing the number of trees and the number of features. 
+While the exact trends at different batch sizes vary it can be seen that \Treebeard{} consistently outperforms RAPIDS by $1.5-8\times$. Along the tree 
+dimension we find that the speedup is very high with fewer trees and stabilizes at around $2\times$ from 300 onwards. We note that \Treebeard{} continues to scale well 
+when the number of trees are increased even further. It achieves a speedup of around $2\times$ compared to RAPIDS on \op{letters}, a real-world benchmark
+with 26K trees.
+
 \begin{figure}[htb]
@@
   \end{minipage}
-  \caption{\label{fig:randomModels4060}\Treebeard{} vs RAPIDs Kernel Time Speedup for random models
+  \caption{\label{fig:randomModels4060}\Treebeard{} vs RAPIDS Kernel Time Speedup for random models
   of max depth 8 at batch size 512 and max depth 6 at batch size 4096.}
@@
 
-% \begin{comment}
-\subsubsection*{Synthetic models}
-To establish that \Treebeard{} can consistently find betters schedules, we performed an exhaustive comparison between \Treebeard{} and RAPIDS on 700 randomly generated models.   
-Figure~\ref{fig:randomModels4060} shows two representative samples of the results, at batch sizes 512 and 4096 on all models, on RTX 4060. 
-Each plot shows the speedup of \Treebeard{} over RAPIDS along the z-axis, with the x-axis representing the number of trees and the y-axis the number of features. 
-While the exact trends at different batch sizes vary it can be seen that \Treebeard{} consistently outperforms RAPIDS by $1.5-8\times$. Along the tree 
-dimension we find that the speedup is very high with fewer trees and stabilizes at around $2\times$ from 300 onwards. We note that \Treebeard{} continues to scale well 
-when the number of trees are increased even further. It achieves a speedup of around $2\times$ compared to RAPIDS on \op{letters}, a real-world benchmark
-with 26K trees.
-%\Treebeard{} consistently outperforms RAPIDS on all models and batch sizes tested.
-% \begin{itemize}
-%   \item Generated a set of over 700 random models with varying depths (6, 7, 8), number of trees (100 - 1000 in steps of 100), 
-%   and number of features (powers of 2 between 8 to 1024 inclusive).
-%   \item Compared the kernel time speedup of \Treebeard{} vs RAPIDs on the RTX 4060. The schedule used was the one picked by the auto-tuner. 
-%   \item Speedups range between 1.5$\times$ and 8$\times$. \Treebeard{} outperforms RAPIDs on all models and batch sizes tested.
-%   \item Figure \ref{fig:randomModels4060} shows the speedup for models of depth 8 and batch size 512 and depth 6 and batch size 4096.
-%   Trends for other batch sizes and depths are similar.
-% \end{itemize}
 \subsubsection*{Utility of Schedule Constructs}
@@
 10\% of the time for the best schedules on the NVIDIA 4060.  
-Shared reduction significantly improves performance especially for multi-class models. For example, 
+Next, shared reduction significantly improves performance especially for multi-class models. For example, 
 the \op{letters} and \op{covtype} benchmarks achieves speedups of 1.5$\times$ with shared reduction. 
 Shared reduction increases the memory pressure (on shared memory) and not all schedules use it.
-While GPUs have a hardware managed cache, explicitly caching the rows using the scheduling primitives 
+While GPUs have a hardware managed cache, explicitly caching the rows in shared memory using the scheduling primitives 
 helps exploit the reuse of rows across trees. We find that caching rows is beneficial for many models
@@
 To test the portability of \Treebeard{}'s techniques, we compare the performance 
-of \Treebeard{} on the T400 GPU with RAPIDs and Tahoe. Figure \ref{Fig:TBvsRAPIDsTahoe_T400_Speedup}
-shows the speedup of code generated by \Treebeard{} vs RAPIDs and Tahoe on the T400.
-Even on the smaller T400 GPU, \Treebeard{} is able to outperform RAPIDs
-and Tahoe consistently. We observe similar trends to those on the RTX 4060.
+of \Treebeard{} on the T400 GPU with RAPIDS and Tahoe. Figure \ref{Fig:TBvsRAPIDSTahoe_T400_Speedup}
+shows the speedup of code generated by \Treebeard{} vs RAPIDS and Tahoe on the T400.
+Even on the smaller T400 GPU, \Treebeard{} is able to outperform RAPIDS
+and Tahoe consistently (speedup greater than $1.5\times$). We observe similar trends to those on the RTX 4060.
 
-% \subsection{Comparison with RAPIDs, Tahoe and XGBoost}
+% \subsection{Comparison with RAPIDS, Tahoe and XGBoost}
 % \begin{itemize}
 %   \item Measured both the kernel time and total time (time including transferring data to the 
-%   GPU and results back) for RAPIDs and \Treebeard{}.
+%   GPU and results back) for RAPIDS and \Treebeard{}.
 %   \item Tahoe only allows us to measure the kernel time since it is written as an executable 
@@
 %   \op{letters} and \op{year}. In these cases, we pick the time of the fastest variant that gives the correct results.
-%   \item Compared the kernel time and total time speedup of \Treebeard{} vs RAPIDs on the RTX 4060 and T400.
+%   \item Compared the kernel time and total time speedup of \Treebeard{} vs RAPIDS on the RTX 4060 and T400.
 %   \item Compared the total time speedup of \Treebeard{} vs XGBoost on the RTX 4060.
 %   \item The schedule used was the one picked by the auto-tuner.
-%   \item \Treebeard{} outperforms RAPIDs at all batch sizes as shown in Figure \ref{Fig:TBvsRAPIDsTahoe_4060_Speedup}. 
+%   \item \Treebeard{} outperforms RAPIDS at all batch sizes as shown in Figure \ref{Fig:TBvsRAPIDSTahoe_4060_Speedup}. 
 %   \item \Treebeard{} outperforms Tahoe on all models and batch sizes tested. Individual benchmark speedups range between 1.1$\times$ and 16$\times$.
 %   \item \Treebeard{} is faster than XGBoost by more than an order of magnitude. Results are not shown because the speedups don't fit on the same graph.
-%   \item Figure \ref{Fig:TBvsRAPIDs_4060_TotalTimeSpeedup} shows that \Treebeard{} offers substantial speedup over RAPIDs even when 
+%   \item Figure \ref{Fig:TBvsRAPIDS_4060_TotalTimeSpeedup} shows that \Treebeard{} offers substantial speedup over RAPIDS even when 
 %   data needs to be transferred to the GPU and results need to be transferred back.
-%   \item Figure \ref{Fig:TBvsRAPIDsTahoe_T400_Speedup} shows that these speedups are also observed on the T400 thus showing 
+%   \item Figure \ref{Fig:TBvsRAPIDSTahoe_T400_Speedup} shows that these speedups are also observed on the T400 thus showing 
 %   that \Treebeard{} offers portable performance across different GPUs.
-%   \item Figure \ref{Fig:KernelTimeIndividualBenchmarks4060} shows that \Treebeard{} outperforms RAPIDs and Tahoe on 
+%   \item Figure \ref{Fig:KernelTimeIndividualBenchmarks4060} shows that \Treebeard{} outperforms RAPIDS and Tahoe on 
 %   individual benchmarks on the RTX 4060 at small (1024) and large (8192) batch sizes. These batch sizes 
 %   require different schedules, but the \Treebeard{} auto-tuning heuristic is able to find them.
-%   % \item Figure \ref{Fig:TotalTimeIndividualBenchmarks4060} compares the total time of \Treebeard{} vs RAPIDs on individual benchmarks. 
+%   % \item Figure \ref{Fig:TotalTimeIndividualBenchmarks4060} compares the total time of \Treebeard{} vs RAPIDS on individual benchmarks. 
 %   % It shows that even with the overhead of data transfers, \Treebeard{} offers significant speedup. The graph in Figure \ref{Fig:TotalTimeIndividualBenchmarks4060} 
@@
 %   \hfill
-%   \caption{\label{Fig:TotalTimeIndividualBenchmarks4060}\Treebeard{} vs RAPIDs total time comparison on NVIDIA RTX 4060. Numbers on the bars are the times 
-%   per sample in $\mu$s for \Treebeard{} and RAPIDs. Times for each benchmark are normalized w.r.t the RAPIDs time for that benchmark.}
+%   \caption{\label{Fig:TotalTimeIndividualBenchmarks4060}\Treebeard{} vs RAPIDS total time comparison on NVIDIA RTX 4060. Numbers on the bars are the times 
+%   per sample in $\mu$s for \Treebeard{} and RAPIDS. Times for each benchmark are normalized w.r.t the RAPIDS time for that benchmark.}
 % \end{figure*}
@@
 \subsection{Performance on AMD GPUs}
-While existing systems only support NVIDIA GPUs, we demonstrate that \Treebeard{} 
+While Tahoe, RAPIDS and XGBoost only support NVIDIA GPUs, we demonstrate that \Treebeard{} 
 is able to generate competitive code for AMD GPUs. \Treebeard{} can target AMD GPUs 
@@
 
-\subsection{Schedule Exploration Heuristic}
+\subsection{Evaluation of the Schedule Exploration Heuristic}
 We evaluated the schedule exploration heuristic described in Section \ref{sec:exploring} on several fronts.
@@
 heuristic is able to find schedules that are very close to the best schedule (well 
-within 5\%) on 4060. The heuristic
-is consistently close to two orders of magnitude faster than exhaustive exploration. 
+within 5\%) on 4060. 
+% The heuristic is consistently close to two orders of magnitude faster than exhaustive exploration. 
+However, the heuristic method is able to find good schedules in a fraction($1/80$-$1/100$) of the time taken by exhaustive exploration.
 The exploration time ranges between $6$ and $167$ seconds for the heuristic with a mean of $28.7$ seconds.
@@
 compares the schedule found by the heuristic on T400 and AMD MI210 with the schedule found on the 4060.
-It shows that schedule exploration needs to be done on each target to achieve the best performance
-and schedules found on one GPU do not always carry over to other GPUs.
+It shows that evaluation of the small number of schedules explored by the heuristic needs to be 
+done on each target to achieve the best performance.
+Schedules found on one GPU do not always carry over to other GPUs and often result in sub-optimal performance.
 For T400, the heuristic is able to find schedules that are 1.05$\times$ to 1.2$\times$ better 
@@
 significantly at small batch sizes. At batch size 32, we find that the geomean speedup over 
-all 8 real-world benchmark models is 2.2$\times$ with a max speedup of 5$\times$. At batch size 64, the average speedup
+all 8 real-world benchmark models is 2.2$\times$ with a max speedup of 5$\times$ (details omitted 
+due to space constraints). At batch size 64, the average speedup
 is 1.1$\times$ with a max speedup of 2$\times$. At batch size 32, parallelizing across trees is faster 
@@
 Also, the amount of work per thread is very small leading to high overheads. 
-Parallelizing across trees fixes both these problems.
+Parallelizing across trees addresses both these problems.
 
@@
 On all platforms and models that we tested on, \Treebeard{} significantly outperforms
-state-of-the-art systems like RAPIDs, Tahoe and \TreebeardOLD{}.
+state-of-the-art systems like RAPIDS, Tahoe and \TreebeardOLD{}.

### 2024-04-19 results/python/analyze_gpu_speedups_results.py
Revert to previous file names

@@ -230,7 +230,7 @@ def plot_double_y_axis_line_graph(df_4060, full_exp_df, gpu_name):
     ax1.legend(fontsize=LEGEND_FONT_SIZE, loc='upper right')
     ax2.legend(fontsize=LEGEND_FONT_SIZE, loc='lower right')
     plt.tight_layout()
-    plt.savefig(f'speedup_over_norm_time_line_graph_{gpu_name}.png')
+    plt.savefig(f'speedup_vs_norm_time_line_graph_{gpu_name}.png')
 
 
 
@@ -311,12 +311,12 @@ line_graph_plotter.plot_geomean_speedups_over_batch_size(results_t400_df, 'TBKer
 line_graph_plotter.plot_geomean_speedups_over_batch_size(results_t400_df, 'TB(AT)', ['RAPIDS(Total)'], ['Speedup of SilvanForge over RAPIDS(total)'])
 line_graph_plotter.save_plot()
 
-line_graph_plotter = LineGraphPlotter('T400', '4060_over_T400_over_MI2160', lower_lim=1, upper_lim=1.6)
+line_graph_plotter = LineGraphPlotter('T400', '4060_vs_T400_vs_MI2160', lower_lim=1, upper_lim=1.6)
 line_graph_plotter.plot_geomean_speedups_over_batch_size(results_t400_df, 'TBKernel(AT)', ['TBKernel(4060)'], ['Scheduling heuristic(T400) over 4060 heuristic schedule'])
 line_graph_plotter.plot_geomean_speedups_over_batch_size(results_amdmi2160_df, 'TBKernel(MI2160)', ['TBKernel(4060)'], ['Scheduling heuristic(MI2160) over 4060 heuristic schedule'])
 line_graph_plotter.save_plot(leg_font_size=11)
 
-line_graph_plotter = LineGraphPlotter('4060', 'full_exp_over_at', lower_lim=1, upper_lim=110)
+line_graph_plotter = LineGraphPlotter('4060', 'full_exp_vs_at', lower_lim=1, upper_lim=110)
 line_graph_plotter.plot_geomean_speedups_over_batch_size(fullexp_over_at_df, 'AT', ['FullExplore'], ['Scheduling heuristic over Full Exploration'])
 line_graph_plotter.save_plot()
 

### 2024-04-19 results/python/analyze_gpu_speedups_results.py
rotate bar labels 90 deg and make it slightly bigger

@@ -19,7 +19,7 @@ PLOT_FILE_EXTENSION = 'png'
 AXIS_FONT_SIZE = 15
 LABEL_FONT_SIZE = 15
 LEGEND_FONT_SIZE = 11.5
-BAR_LABEL_FONT_SIZE = 8
+BAR_LABEL_FONT_SIZE = 11
 
 def plot_bar_graph_of_abs_times_for_batch_size(df, batch_size):
     filtered_df = df[df['Batch size'] == batch_size]
@@ -109,7 +109,7 @@ def plot_bar_graph_speedups_for_batch_size(df, batch_size, kwargs=None):
     # enable grid
     # ax.grid(True, which='both', axis='y')
 
-    ax.set_ylim(0, max(max(rapids_speedup), max(tahoe_speedup)) + 1)
+    ax.set_ylim(0, max(max(rapids_speedup), max(tahoe_speedup)) + 2)
 
     # Add labels and title
     # ax.set_xlabel('Benchmark', fontsize=LABEL_FONT_SIZE)
@@ -127,11 +127,11 @@ def plot_bar_graph_speedups_for_batch_size(df, batch_size, kwargs=None):
         rt = round(rt, 2)
         tt = round(tt, 2)
 
-        plt.text(benchmark, rs + 0.2 , str(rt), ha='center', fontsize=BAR_LABEL_FONT_SIZE)
-        plt.text(benchmark + width, ts + 0.2, str(tt), ha='center', fontsize=BAR_LABEL_FONT_SIZE)
+        plt.text(benchmark, rs + 0.2 , str(rt), ha='center', fontsize=BAR_LABEL_FONT_SIZE, rotation=90)
+        plt.text(benchmark + width, ts + 0.2, str(tt), ha='center', fontsize=BAR_LABEL_FONT_SIZE, rotation=90)
 
     # Show the legend and plot
-    plt.legend(fontsize=LEGEND_FONT_SIZE)
+    plt.legend(fontsize=LEGEND_FONT_SIZE, loc='upper left')
     plt.tight_layout()
     # plt.show()
     plt.savefig(f'speedup_bar_graph_{batch_size}.{PLOT_FILE_EXTENSION}')

### 2024-04-19 docs/SOSP 2024/results.tex
Shrink figures in results

@@
 
-\begin{figure}[htb]
-  \centering
-  \includegraphics[width=0.75\linewidth]{figures/geomean_speedup_4060_kernel_time_total_time.png}
-  \caption{\Treebeard{} vs RAPIDS and Tahoe kernel time and total time speedup (geomean over real-world 
-  benchmarks) across batch sizes on NVIDIA RTX 4060}
-  \label{Fig:TBvsRAPIDSTahoe_4060_Speedup}
-\end{figure}
+\begin{figure*}[htb]
+    \begin{subfigure}[t]{.32\linewidth}
+      \includegraphics[width=\textwidth]{figures/geomean_speedup_4060_kernel_time_total_time.png}
+      \caption{\label{Fig:TBvsRAPIDSTahoe_4060_Speedup}Speedup vs batch size}      
+    \end{subfigure}
+    \begin{subfigure}[t]{.32\linewidth}
+      \includegraphics[width=\textwidth]{figures/speedup_bar_graph_1024.png}
+      \caption{\label{Fig:KernelTimeIndividualBenchmarks4060a} Batch size 1024}
+    \end{subfigure}
+    \begin{subfigure}[t]{.32\linewidth}
+      \includegraphics[width=\textwidth]{figures/speedup_bar_graph_8192.png}
+      \caption{\label{Fig:KernelTimeIndividualBenchmarks4060b} Batch size 8192}
+    \end{subfigure}
+    % \hfill
+    \caption{The first graph shows kernel and total time speedup of \Treebeard{} over RAPIDS and Tahoe (geomean over real-world 
+    benchmarks) across batch sizes on NVIDIA RTX 4060. The bar graphs show kernel time speedup of \Treebeard{} over RAPIDS and Tahoe 
+    on NVIDIA RTX 4060. Numbers on the bars are inference times per sample in $\mu$s for RAPIDS and Tahoe.}
+\end{figure*}
+
+% \begin{figure}[htb]
+%   \centering
+%   \includegraphics[width=0.75\linewidth]{figures/geomean_speedup_4060_kernel_time_total_time.png}
+%   \caption{\Treebeard{} vs RAPIDS and Tahoe kernel time and total time speedup (geomean over real-world 
+%   benchmarks) across batch sizes on NVIDIA RTX 4060}
+%   \label{Fig:TBvsRAPIDSTahoe_4060_Speedup}
+% \end{figure}
 
@@
 
-\begin{figure}[htb]
-  \centering
-  \includegraphics[width=0.75\linewidth]{figures/geomean_speedup_T400_kernel_time.png}
-  \caption{\Treebeard{} vs RAPIDS and Tahoe kernel time and total time speedup (geomean over real-world 
-  benchmarks) across batch sizes on NVIDIA T400.}
-  \label{Fig:TBvsRAPIDSTahoe_T400_Speedup}
-\end{figure}
+% \begin{figure*}[ht]
+%   \centering
+%   \begin{subfigure}[b]{.45\textwidth}
+%     \subcaptionbox*{}{\includegraphics[width=\textwidth]{figures/speedup_bar_graph_1024.png}}
+%     \caption{Batch size 1024}
+%   \end{subfigure}
+%   \begin{subfigure}[b]{.45\textwidth}
+%     \subcaptionbox*{}{\includegraphics[width=\textwidth]{figures/speedup_bar_graph_8192.png}}
+%     \caption{Batch size 8192}
+%   \end{subfigure}
+%   \hfill
+%   \caption{\label{Fig:KernelTimeIndividualBenchmarks4060}Kernel time speedup of \Treebeard{} vs RAPIDS on NVIDIA RTX 4060. Numbers on the bars are 
+%   inference times per sample in $\mu$s for RAPIDS and Tahoe.}
+% \end{figure*}
 
-\begin{figure*}[ht]
-  \centering
-  \begin{subfigure}[b]{.45\textwidth}
-    \subcaptionbox*{}{\includegraphics[width=\textwidth]{figures/speedup_bar_graph_1024.png}}
-    \caption{Batch size 1024}
-  \end{subfigure}
-  \begin{subfigure}[b]{.45\textwidth}
-    \subcaptionbox*{}{\includegraphics[width=\textwidth]{figures/speedup_bar_graph_8192.png}}
-    \caption{Batch size 8192}
-  \end{subfigure}
-  \hfill
-  \caption{\label{Fig:KernelTimeIndividualBenchmarks4060}Kernel time speedup of \Treebeard{} vs RAPIDS on NVIDIA RTX 4060. Numbers on the bars are 
-  inference times per sample in $\mu$s for RAPIDS and Tahoe.}
-\end{figure*}
-
-Figure~\ref{Fig:KernelTimeIndividualBenchmarks4060} shows per benchmark results at two different batch sizes. 
+Figures~\ref{Fig:KernelTimeIndividualBenchmarks4060a} and ~\ref{Fig:KernelTimeIndividualBenchmarks4060b} 
+show per benchmark results at two different batch sizes. 
 \Treebeard{} is able to find better schedules than both RAPIDS and Tahoe on each of the benchmarks. It consistently
-outperforms both systems, achieving a speedup of upto $12\times$ an average speedup of over $2\times$.
+outperforms both systems, achieving a speedup of upto $12\times$ and average speedups of over $2\times$.
 For about half the benchmarks, the speedup is $2\times$ or more over both baselines at batch size 8192.
@@
 \begin{figure}[htb]
-  \begin{minipage}[t]{.475\linewidth}
+  \begin{subfigure}[t]{.475\linewidth}
     \includegraphics[width=\linewidth]{figures/RandomModels/kernel_speedup_b512_depth8.png}
-    %\caption{Batch size 512}
-  \end{minipage}
-  \begin{minipage}[t]{.475\linewidth}
+    \caption{}
+  \end{subfigure}
+  \begin{subfigure}[t]{.475\linewidth}
     \includegraphics[width=\linewidth]{figures/RandomModels/kernel_speedup_b4096_depth6.png}
-    %\caption{Batch size 4096}
-  \end{minipage}
-  \caption{\label{fig:randomModels4060}\Treebeard{} vs RAPIDS Kernel Time Speedup for random models
+    \caption{}
+  \end{subfigure}
+  \caption{\label{fig:randomModels4060}Kernel time speedup of \Treebeard{} over RAPIDS for random models
   of max depth 8 at batch size 512 and max depth 6 at batch size 4096.}
@@
 
-% \subsection{Comparison with RAPIDS, Tahoe and XGBoost}
-% \begin{itemize}
-%   \item Measured both the kernel time and total time (time including transferring data to the 
-%   GPU and results back) for RAPIDS and \Treebeard{}.
-%   \item Tahoe only allows us to measure the kernel time since it is written as an executable 
-%   that performs inference repeatedly on the same data that is transferred to the GPU once.
-%   \item Tahoe does not support multiclass models. We just ran the multiclass models (\op{covtype} and \op{letters})
-%   as regression models for the comparison. Tahoe also gives wrong results (as reported by its own tests) for 
-%   \op{letters} and \op{year}. In these cases, we pick the time of the fastest variant that gives the correct results.
-%   \item Compared the kernel time and total time speedup of \Treebeard{} vs RAPIDS on the RTX 4060 and T400.
-%   \item Compared the total time speedup of \Treebeard{} vs XGBoost on the RTX 4060.
-%   \item The schedule used was the one picked by the auto-tuner.
-%   \item \Treebeard{} outperforms RAPIDS at all batch sizes as shown in Figure \ref{Fig:TBvsRAPIDSTahoe_4060_Speedup}. 
-%   \item \Treebeard{} outperforms Tahoe on all models and batch sizes tested. Individual benchmark speedups range between 1.1$\times$ and 16$\times$.
-%   \item \Treebeard{} is faster than XGBoost by more than an order of magnitude. Results are not shown because the speedups don't fit on the same graph.
-%   \item Figure \ref{Fig:TBvsRAPIDS_4060_TotalTimeSpeedup} shows that \Treebeard{} offers substantial speedup over RAPIDS even when 
-%   data needs to be transferred to the GPU and results need to be transferred back.
-%   \item Figure \ref{Fig:TBvsRAPIDSTahoe_T400_Speedup} shows that these speedups are also observed on the T400 thus showing 
-%   that \Treebeard{} offers portable performance across different GPUs.
-%   \item Figure \ref{Fig:KernelTimeIndividualBenchmarks4060} shows that \Treebeard{} outperforms RAPIDS and Tahoe on 
-%   individual benchmarks on the RTX 4060 at small (1024) and large (8192) batch sizes. These batch sizes 
-%   require different schedules, but the \Treebeard{} auto-tuning heuristic is able to find them.
-%   % \item Figure \ref{Fig:TotalTimeIndividualBenchmarks4060} compares the total time of \Treebeard{} vs RAPIDS on individual benchmarks. 
-%   % It shows that even with the overhead of data transfers, \Treebeard{} offers significant speedup. The graph in Figure \ref{Fig:TotalTimeIndividualBenchmarks4060} 
-%   % shows the breakup of the total time into kernel time and data transfer time.
-% \end{itemize}
-% \end{comment}
-
-% \begin{figure*}[ht]
-%   \centering
-%   \begin{subfigure}[b]{.45\textwidth}
-%     \subcaptionbox*{}{\includegraphics[width=\textwidth]{figures/abs_times_bar_graph_1024.png}}
-%     \caption{Batch size 1024.}
-%   \end{subfigure}
-%   \begin{subfigure}[b]{.45\textwidth}
-%     \subcaptionbox*{}{\includegraphics[width=\textwidth]{figures/abs_times_bar_graph_8192.png}}
-%     \caption{Batch size 8192.}
-%   \end{subfigure}
-%   \hfill
-%   \caption{\label{Fig:TotalTimeIndividualBenchmarks4060}\Treebeard{} vs RAPIDS total time comparison on NVIDIA RTX 4060. Numbers on the bars are the times 
-%   per sample in $\mu$s for \Treebeard{} and RAPIDS. Times for each benchmark are normalized w.r.t the RAPIDS time for that benchmark.}
-% \end{figure*}
-
-% \begin{figure}[htb]
-%   \centering
-%   \includegraphics[width=0.75\linewidth]{figures/TBvsXGB_TotalTime.png}
-%   \caption{\Treebeard{} vs XGBoost Speedup on RTX 4060.}
-%   \label{Fig:TBvsXGBoost_Speedup}
-% \end{figure}
+\begin{figure}[htb]
+  \centering
+  \includegraphics[width=0.65\linewidth]{figures/geomean_speedup_T400_kernel_time.png}
+  \caption{Speedup of \Treebeard{} over RAPIDS and Tahoe (geomean over real-world 
+  benchmarks) across batch sizes on NVIDIA T400.}
+  \label{Fig:TBvsRAPIDSTahoe_T400_Speedup}
+\end{figure}
 
@@
 
-% \subsubsection*{Exploration Time}
-% Finally, we measure the improvement in exploration time that the heuristic provides 
-% compared to exhaustive exploration. Figure \ref{Fig:HeuristicVsFullExplore_Speedup} shows that the heuristic
-% is consistently close to two orders of magnitude faster than exhaustive exploration. 
-% The exploration time ranges between $6$ and $167$ seconds for the heuristic with a mean of $28.7$ seconds.
-% These results show that our heuristic is able to quickly find schedules that are close to the best schedule.
-
 \begin{figure}[htb]
   \centering
-  \includegraphics[width=0.75\linewidth]{figures/speedup_vs_norm_time_line_graph_4060.png}
-  \caption{Speedup of schedules found by the heuristic vs best schedule found through exhaustive search on NVIDIA RTX 4060 and 
+  \includegraphics[width=0.65\linewidth]{figures/speedup_vs_norm_time_line_graph_4060.png}
+  \caption{Slowdown of heuristic schedule vs best schedule on NVIDIA RTX 4060 and 
   heuristic schedule exploration time normalized w.r.t full schedule exploration time}
@@
   \centering
-  \includegraphics[width=0.75\linewidth]{figures/geomean_speedup_T400_4060_vs_T400_vs_MI2160.png}
-  \caption{Geomean speedup of schedule found by exploration heuristic on NVIDIA T400 and AMD MI2160 vs 
+  \includegraphics[width=0.65\linewidth]{figures/geomean_speedup_T400_4060_vs_T400_vs_MI2160.png}
+  \caption{Geomean speedup of heuristic schedule on NVIDIA T400 and AMD MI2160 vs 
   the heuristic schedule on NVIDIA RTX 4060 across batch sizes}
@@
 \end{figure}
-
-% \begin{itemize}
-%   \item We compare the schedule found by the schedule exploration heuristic on the RTX 4060 with the best schedule found by
-%   extensive exploration. Figure \ref{Fig:HeuristicVsFullExplore_Speedup} shows that the heuristic is able to find schedules that are
-%   very close to the best schedule found by exhaustive exploration.
-%   \item The speedup of the heuristic schedule vs the best 4060 schedule when run on T400 is shown in Figure \ref{Fig:AutotuningSpeedupvs4060Sched}.
-%   Geomean speedup ranges from 1.05$\times$ to 1.2$\times$. 
-%   \item Unsurprisingly, there is some variation in the best schedule for a model even across NVIDIA GPUs.
-% \end{itemize}
-
-% \begin{itemize}
-%   \item \Treebeard{} is able to compile code to AMD GPUs. We run generated code on an AMD MI210 GPU.
-%   \item None of the other systems we compare against support running on AMD GPUs. This shows the portability of \Treebeard{}.
-%   \item The speedup of the heuristic schedule vs the best 4060 schedule when run on the MI210 is shown in Figure \ref{Fig:AutotuningSpeedupvs4060Sched}.
-%   \item We see that the geomean speedup over all benchmarks is 1.5$\times$ at batch size 16k.
-%   \item The maximum speedup is 2$\times$ for the \op{letters} benchmark at batch size 16k.
-% \end{itemize}
 
@@
 
-% \begin{itemize}
-%   \item The ability of \Treebeard{} to parallelize across both trees and rows has a significant impact on CPU performance.
-%   \item We run experiments on the Intel Core i9-11900K processor to evaluate the impact of parallelizing across trees with small batch sizes. 
-%   \item At batch size 32, we find that the average speedup over all 8 models is 2.2$\times$ with a max speedup of 5$\times$.
-%   \item At batch size 64, we find that the average speedup over all 8 models is 1.1$\times$ with a max speedup of 2$\times$. 
-%   However, we find that 2 of the 8 models show slowdowns in this case. Again, this highlights the need for schedule exploration on CPUs.
-%   \item For small batch sizes, parallelizing across rows does not work great since there is limited reuse of trees in L1 cache.
-%   Also, the amount of work per thread is very small leading to high overheads. 
-%   \item Parallelizing across trees is more effective since there is more reuse in L1 cache and the amount of work per thread is higher.  
-% \end{itemize}
-
 Overall, our evaluation shows that \Treebeard{} is able to efficiently generate 

### 2024-04-19 docs/SOSP 2024/reduction.tex
switch to python style loops to save space

@@
   model = ensemble(...) 
-  par.for t0 = 0 to N_t step N_t/2 {
-    for t1 = 0 to N_t/2 {
-      for batch = 0 to batch_size {
+  par.for t0 = 0 to N_t step N_t/2:
+    for t1 = 0 to N_t/2:
+      for batch = 0 to batch_size:
         t = getTree(model, t0 + t1) 
@@
         reduce(result[batch], p)
-      }
-    }
-  }  
 \end{lstlisting}
@@
         reduce(partResults[t0/(N_t/2)][batch], p)
-      }
-    }
-  }
+
   results = reduceDimension(partResults[:, :], 0)

### 2024-04-19 docs/SOSP 2024/results.tex
Some more latex hacking to save space

@@
 
-\begin{figure}[htb]
-  \centering
-  \includegraphics[width=0.65\linewidth]{figures/geomean_speedup_T400_kernel_time.png}
-  \caption{Speedup of \Treebeard{} over RAPIDS and Tahoe (geomean over real-world 
-  benchmarks) across batch sizes on NVIDIA T400.}
-  \label{Fig:TBvsRAPIDSTahoe_T400_Speedup}
-\end{figure}
-
 \subsection{Performance on AMD GPUs}
@@
 
-\begin{figure}[htb]
+\begin{figure*}[htb]
+\begin{minipage}[t]{.3\linewidth}
   \centering
-  \includegraphics[width=0.65\linewidth]{figures/speedup_vs_norm_time_line_graph_4060.png}
+  \includegraphics[width=\linewidth]{figures/geomean_speedup_T400_kernel_time.png}
+  \caption{Speedup of \Treebeard{} over RAPIDS and Tahoe (geomean over real-world 
+  benchmarks) across batch sizes on NVIDIA T400.}
+  \label{Fig:TBvsRAPIDSTahoe_T400_Speedup}
+\end{minipage}
+\hspace{0.5cm}
+\begin{minipage}[t]{.3\linewidth}
+  \centering
+  \includegraphics[width=\linewidth]{figures/speedup_vs_norm_time_line_graph_4060.png}
   \caption{Slowdown of heuristic schedule vs best schedule on NVIDIA RTX 4060 and 
@@
   \label{Fig:HeuristicVsFullExplore_Speedup}
-\end{figure}
-
-\begin{figure}[htb]
+\end{minipage}
+\hspace{0.5cm}
+\begin{minipage}[t]{.3\linewidth}
   \centering
-  \includegraphics[width=0.65\linewidth]{figures/geomean_speedup_T400_4060_vs_T400_vs_MI2160.png}
+  \includegraphics[width=\linewidth]{figures/geomean_speedup_T400_4060_vs_T400_vs_MI2160.png}
   \caption{Geomean speedup of heuristic schedule on NVIDIA T400 and AMD MI2160 vs 
@@
   \label{Fig:AutotuningSpeedupvs4060Sched}
-\end{figure}
+\end{minipage}
+\end{figure*}
 

### 2024-04-19 docs/SOSP 2024/caching.tex
Shrink some figures
Fix some typos

@@
   \centering
-  \includegraphics[width=\linewidth]{figures/representations.png}
+  \includegraphics[width=0.7\linewidth]{figures/representations.png}
   \caption{The three representations supported by \Treebeard{}.}
@@
 this interface thus hiding details of the actual representation from the core 
-compiler. Curcially, the interface abstracts how and what buffers 
+compiler. Crucially, the interface abstracts how and what buffers 
 are allocated, how to move from a node to its child, how trees 
@@
 When the MIR is lowered to LIR, the cache ops are lowered to target-specific code. Each of 
-the two caching operations is lowered differently for the CPU and the GPU. On CPU, the cahce operations are lowered to preteches while
+the two caching operations is lowered differently for the CPU and the GPU. On CPU, the cache operations are lowered to prefetches while
 on the GPU they are lowered to reads into shared memory.

### 2024-04-19 docs/SOSP 2024/abstract.tex
Edit abstract and intro

@@
   portable nor achieve the best possible performance for the specific hardware they 
-  target because they do not explore different optimization configurations.
+  target because they do not explore and customize optimization configurations to the model 
+  being used.
 
@@
   \Treebeard{} has two core components. The first is a scheduling language 
-  that encapsulates the optimization space for decision tree inference, 
-  and techniques to efficiently explore this space. 
-  The second is an optimizing retargetable compiler that 
+  that encapsulates the optimization space, and techniques to efficiently
+  explore this space. The second is an optimizing retargetable compiler that 
   can generate code for any specified schedule. \Treebeard{}'s ability to use different data layouts, 
-  loop structures and caching strategies enables it to achieve portable performance.   
+  loop structures and caching strategies enables it to achieve portable performance
+  across a range of CPUs and GPUs.   
   % To accomplish this level of configurability, we re-architect 

### 2024-04-19 docs/SOSP 2024/intro.tex
Edit abstract and intro

@@
 Despite the simplicity of the model and the availability of multiple sources of coarse-grain parallelism (parallelism 
-across inputs in a batch and parallelism across trees), existing systems do not consistently scale well across different
+across inputs in a batch and parallelism across trees), existing systems do not consistently perform well across different
 models even on the limited set of targets they support. 
@@
 and typically only target a single platform. 
-XGBoost~\cite{XGBoost} uses a sparse representation for the model and a loop structure that processes one tree for a block of rows before moving to 
-the next tree. RAPIDS FIL~\cite{FIL} uses a reorg representation and partitions trees across a fixed number of threads. 
+On the CPU, XGBoost~\cite{XGBoost} uses a sparse representation for the model 
+and a loop structure that processes one tree for a block of rows before moving to 
+the next tree. On the GPU, RAPIDS FIL~\cite{FIL} uses a reorg representation~\cite{FIL}
+and partitions trees across a fixed number of threads. 
 %% RG: 
@@
 %%
-Tahoe~\cite{Tahoe} uses a
-variation of the reorg representation and has four predefined inference strategies from which it picks one based on an analytical model.
-All these systems are CUDA based and only work on NVIDIA based GPUs.
+Tahoe~\cite{Tahoe} uses a variation of the reorg representation and 
+has four predefined inference strategies from which it picks one based on an analytical model.
+The GPU systems are CUDA based and only work on NVIDIA GPUs.
 %These systems are all library based and are hard to extend. 
@@
 built using the MLIR infrastructure~\cite{MLIR}, supports two fixed loop structures 
-and does not scale well with increasing number of threads. Additionally, it lacks 
-GPU specific optimizations that are critical to scale performance to massive number of threads.
-\TODO{Shouldn't we reverse this? First say no GPU support and then the scaling issue.}
+% and does not scale well with increasing number of threads. Additionally, it 
+and lacks GPU specific optimizations that are critical to scale performance to massive number of threads.
+% \TODO{Shouldn't we reverse this? First say no GPU support and then the scaling issue.}
 
-This paper presents \Treebeard{}, a novel schedule guided compilation infrastructure
+This paper presents \Treebeard{}, a novel schedule-guided compiler
 for decision tree inference on multiple hardware targets. \Treebeard{} is able to 
@@
 support schedule-guided code generation, and incorporate several new optimizations that are critical to target GPUs. 
-These two components of the proposed \Treebeard{} compiler infrastructure are intertwined, each benefitting from the other.
+These two components of \Treebeard{} are intertwined, each benefitting from the other.
 %Last, to address the challenge of selecting a near-optimal schedule from the vast optimization space, 
@@
 benchmark models and compared against state-of-the-art systems RAPIDS~\cite{RAPIDS}, Tahoe~\cite{Tahoe}
-, XGBoost~\cite{XGBoost} and \TreebeardOLD{}~\cite{Treebeard}. While these systems can only run on 
-specific hardware, \Treebeard{} outperforms all these systems while providing portable performance across 
-a range of target hardware. The geomean speedup (across all benchmarks) of \Treebeard{} over
+, XGBoost~\cite{XGBoost} and \TreebeardOLD{}~\cite{Treebeard}. 
+% While these systems can only run on 
+% specific hardware, \Treebeard{} outperforms all these systems while providing portable performance across 
+% a range of target hardware.
+The geomean speedup (across all benchmarks) of \Treebeard{} over
 Tahoe and RAPIDS is in the range $2$--$5\times$ over several batch sizes. The geomean speedup over
-XGBoost is more than $10\times$. \Treebeard{} also enables better scalability on CPUs by parallelizing 
+XGBoost is more than $10\times$. While other systems can only run on specific hardware,
+\Treebeard{} can provide portable performance across a range of target hardware
+including AMD GPUs and CPUs. \Treebeard{} also enables better scalability on CPUs by parallelizing 
 across multiple dimensions. We also demonstrate that \Treebeard{}'s scheduling heuristic is able to 
@@
   decision tree inference and that the best implementation depends both on the model
-  and the target. We design a scheduling language that 
+  and the target processor. We design a scheduling language that 
   allows us to represent this solution space. 
-%   This 
-%   scheduling language is expressive enough to represent a wide range of  
-%   implementation strategies,  %%.  proposed by prior work.
-% %% RG: added CPUs and GPUs. 
-%   such as different data layouts, caching strategies, parallel reduction schemes, 
-%   that work across CPUs and GPUs. 
   \item We propose a retargetable compiler that can generate code as specified by 
@@
   efficient GPU code.
-  \item The scheduling language can represent an unbounded number of schedules. 
-  We design and implement techniques to search over this space and find high-performance 
+  \item Since the scheduling language can represent an unbounded number of schedules, 
+  we design and implement techniques to search over this space and find high-performance 
   schedules in tens of seconds. 

### 2024-04-19 docs/SOSP 2024/motivation.tex
Edit motivation

@@
 We first describe the simple schedule shown in Figure \ref{fig:schedulea} that processes one tree at a time 
-for all input rows. The scheduling constructs are fairly intuitive. We 
-defer a detailed explanation of the scheduling language to Section~\ref{sec:schedule}. 
+for all input rows. \footnote{Since the scheduling constructs are fairly intuitive, we 
+defer a detailed explanation of the scheduling language to Section~\ref{sec:schedule}.} 
 In this schedule, the loop over the trees is split into two -- one that
@@
 When the schedule specifies that the tree walks should be unrolled, 
-\Treebeard{} pads the trees so that all leaves are at the same depth
+\Treebeard{} pads the trees so that all leaves of a tree are at the same depth
 (\circled{B} in Figure~\ref{Fig:HIRExample}). 
+The concrete implementation of this schedule (in \Treebeard{}'s IR) 
+is as shown in Figure \ref{fig:schedulea}.
+This schedule is ideally suited for a single-core CPU. It maximizes 
+the reuse of trees in the L1 cache. However, it doesn't exploit  
+any parallelism.
 
@@
 
-% \begin{lstlisting}[style=c++]
-%   reorder(tree, batch)
-%   // Fiss tree loop so trees with equal depth 
-%   // are processed together  
-%   split(tree, t_depth1, t_depth2, 2)
-%   // Unroll the tree walks
-%   unrollWalk(t_depth1, 1)
-%   unrollWalk(t_depth2, 2)
-% \end{lstlisting}
-The concrete implementation of this schedule (in \Treebeard{}'s IR) 
-is as shown in Figure \ref{fig:schedulea}.
-% \begin{lstlisting}[style=c++]
-%   model = ensemble(...)
-%   for t_depth1 = 0 to 2 step 1 {
-%     T = getTree(model, t_depth1)
-%     for batch = 0 to BATCH_SIZE step 1 {
-%       treePred = walkDecisionTree(T, 
-%                     input[batch]) <unrollDepth = 1>
-%       reduce(result[batch], treePred)
-%     }
-%   }
-%   for t_depth2 = 2 to 4 step 1 {
-%     T = getTree(model, t_depth2)
-%     for batch = 0 to BATCH_SIZE step 1 {
-%       treePred = walkDecisionTree(T,
-%                     input[batch]) <unrollDepth = 2>
-%       reduce(result[batch], treePred) <'+', 0.0>
-%     }
-%   }
-% \end{lstlisting}
-This schedule is ideally suited for a single-core CPU. It maximizes 
-the reuse of trees in the L1 cache. However, it doesn't exploit  
-any parallelism. 
-%% RG, 
-%%  and is therefore ill-suited for parallel processors.
-% Further, the \op{batch} for loop can be tiled to get locality benefits.
-
-One form of parallelism that can be exploited is to process rows in 
-parallel. 
-%% RG:
+One form of parallelism that can be exploited is to process rows in parallel. 
 While this may work for multi-core CPUs\footnote{As we report in Section~\ref{sec:results}, 
@@
 we can additionally parallelize across trees as done in the schedule in Figure \ref{fig:scheduleb}. 
-%% RG:
-% With GPUs, the performance of generated code would also vary with the 
-% kernel configuration, which specifies the number of thread blocks\footnote{We 
-% use NVIDIA CUDA terminology in the discussion, but as we demonstrate in 
-% Section~\ref{sec:results}, our framework can generate code for AMD GPUs as well.}, 
-% the size of thread blocks and so on. 
-% A possible strategy to accomplish this is encoded in the following schedule.
-% \begin{lstlisting}[style=c++]
-%   // Split the trees into two sets
-%   tile(tree, t0, t1, 2)
-%   reorder(batch, t1, t0)
-%   // Fiss loop so that trees with equal 
-%   // depth are processed together
-%   split(t0, t0_depth1, t0_depth2, 2)
-%   unrollWalk(t0_depth1, 1)
-%   unrollWalk(t0_depth2, 2)
-%   // Configure the GPU kernel dimensions
-%   gpuDimension(batch, grid.x)
-%   gpuDimension(t1, block.x)
-% \end{lstlisting}
 The corresponding GPU inference routine is shown in the same figure. 
 This schedule generates a routine that processes one input row per thread block (\op{batch}
-loop is mapped to \op{grid.x}).
-The schedule also tiles the \op{tree} loop into a nest of two loops with indices $t0$ and $t1$. 
-It then additionally parallelizes across the $t1$ loop.
+loop is mapped to \op{grid.x}). The schedule also tiles the \op{tree} loop into a
+nest of two loops with indices $t0$ and $t1$. It then additionally parallelizes across the $t1$ loop.
 Finally, it splits $t0$ and unrolls the walks for each tree depth.
-%It also splits the trees into two sets by tiling the \op{tree} loop.
-%Each of the two sets is processed in parallel. We unroll the tree walks 
-%for each tree.
-% The IR generated is as follows. 
-% \begin{lstlisting}[style=c++]
-%   model = ensemble(...)
-%   par.for batch = 0 to BATCH_SIZE step 1 <grid.x> {
-%     par.for t1 = 0 to 2 step 1 <block.x> {
-%       for t0_depth1 = 0 to 2 step 2 {
-%         T = getTree(model, t0_depth1 + t1)
-%         treePred = walkDecisionTree(T, 
-%                         input[batch]) <unrollDepth = 1>
-%         reduce(result[batch], treePred)
-%       }
-%       for t0_depth2 = 2 to 4 step 2 {
-%         T = getTree(model, t0_depth2 + t1)
-%         treePred = walkDecisionTree(T,
-%                         input[batch]) <unrollDepth = 2>
-%         reduce(result[batch], treePred) <'+', 0.0>
-%       }
-%     }
-%   }
-% \end{lstlisting}
 
-% In the case of this schedule, the \op{reduce}
-% operation needs special consideration.  
-% %% RG
-% %% to correctly generate code for this schedule, the compiler needs 
-% To generate the correct code for this schedule, the compiler needs to determine
-% the required synchronization and temporary storage for reductions 
-% given the parallelization strategy specified. In the given example, it 
-% has to identify that parallel iterations of the \op{t1} 
-% loop accumulate into the same element of the \op{result} array.
-% One possible solution is to rewrite the reduction so that each parallel 
-% iteration accumulates into a different array element by introducing 
-% a temporary buffer (\op{temp}) as follows.
-% \begin{lstlisting}[style=c++]
-%   float temp[2][BATCH_SIZE]
-%   model = ensemble(...)
-%   par.for batch = 0 to BATCH_SIZE step 1 <grid.x> {
-%     par.for t1 = 0 to 2 step 1 <block.x> {
-%       for t0_depth1 = 0 to 2 step 2 {
-%         T = getTree(ensemble, t0_depth1 + t1)
-%         treePred = walkDecisionTree(T, 
-%                       input[batch]) <unrollDepth = 1>
-%         reduce(temp[t1][batch], treePred)
-%       }
-%       for t0_depth2 = 2 to 4 step 2 {
-%         T = getTree(ensemble, t0_depth2 + t1)
-%         treePred = walkDecisionTree(T,
-%                       input[batch]) <unrollDepth = 2>
-%         reduce(temp[t1][batch], treePred) <'+', 0.0>
-%       }
-%     }
-%     result[batch] = reduce_dimension(temp[:][batch], 0)
-%   }
-% \end{lstlisting}
-% Here, partial results are accumulated into \op{temp} and then
-% reduced across the \op{t1} dimension to get the final result.
 Note that while unrolling helps avoid branching, it increases 
@@
 of branch divergence~\cite{FungMicro, MilindDivergence}.
-These schedules just use a combination of 4-8 constructs and already generate schedules 
+
+These \Treebeard{} schedules just use a combination of 4-8 constructs and already generate strategies 
 that are different from what library based systems like XGBoost, Tahoe and RAPIDs FIL use. 
-As one can imagine several other schedules with different trade-offs can be generating using the constructs.
-% As is evident from these examples, it is possible to optimize 
-% the inference routine in different ways. Also, the structure of the loop 
-% nest in the inference routine can get quite complex even 
-% for simple schedules. Writing these routines by hand 
-% is error-prone and time-consuming.  
-% %% RG 
-% %% We beleive 
-% Hence designing a 
-% scheduling language to encapsulate these strategies and a principled code 
-% generator to automatically generate code guided by the schedule is the 
-% best approach. We promote such an approach in this paper.
+As one can imagine, several other schedules with different trade-offs can be generating using 
+these constructs.
 
@@
   \end{subfigure}
-  \caption{Batch and model sensitivity plots. Each point shows the slowdown when the best schedule for the x-axis batch size (model) is used for the y-axis batch size (model).}
+  \caption{Batch and model sensitivity plots. Each point shows the slowdown when the best 
+  schedule for the x-axis batch size (model) is used for the y-axis batch size (model).}
 \end{figure}
-
-% \begin{figure}[htb]
-%   \begin{minipage}[t]{.475\linewidth}
-%     \subcaptionbox*{}{\includegraphics[width=\linewidth]{figures/batch_sensitivity_covtype.png}}
-%     \caption{\label{fig:sensitivitya}Batch sensitivity for \op{covtype}}
-%   \end{minipage}
-%   \begin{minipage}[t]{.475\linewidth}
-%     \subcaptionbox*{}{\includegraphics[width=\linewidth]{figures/model_sensitivity_4096.png}}
-%     \caption{\label{fig:sensitivityb}Model sensitivity at batch size 4096}
-%   \end{minipage}
-%   \caption{Batch and model sensitivity plots. Each point shows the slowdown when best schedule for the
-%           x-axis batch size (model) is used for the y-axis batch size (model).}
-% \end{figure}
 
@@
 The degradation is much worse when schedules are used across different models.
-In many (\~20\%) instances, using the best schedule for one model 
+In many ($\sim20$\%) instances, using the best schedule for one model 
 to compile another results in a 5$\times$ slowdown.

### 2024-04-19 docs/SOSP 2024/reduction.tex
Change titles to save space
Use ttfamily for snippets

@@
-\section{Reductions : Representation, Optimization and Lowering}
+\section{Representing and Optimizing Reductions}
 \label{sec:reduction}
@@
 \begin{lstlisting}[style=c++]
-  float result[batch_size], partResults[2][batch_size]
+  float result[batch_size];
+  float partResults[2][batch_size];
   model = ensemble(...) 

### 2024-04-19 docs/SOSP 2024/schedule.tex
switch to python style loops to save space
Minor edits to schedule section

@@
-\section{Scheduling Language}
+\section{\Treebeard{}'s Scheduling Language}
 \label{sec:schedule}
@@
 % make portable performance impossible. 
-This section presents all constructs in the scheduling language. 
-The language provides an abstract way to specify loop structure and 
+\Treebeard{}'s scheduling language provides an abstract way to specify loop structure and 
 other optimizations as an input to the compiler. The specified 
@@
 variables. There are two special index variables -- \op{batch} and
-\op{tree} that are used to represent the batch and tree loops and all 
+\op{tree} that are used to represent the batch and tree loops\footnote{
+The canonical inference routine has a batch loop that goes over 
+all rows and a loop that goes over all trees nested within it.} and all 
 other index variables are derived from these. A schedule derives 
 new index variables from these root index variables by applying
-directives. \TODO{Should we somehow explain what tree and batch loop are?}
+directives.
 
@@
 of a node represent immediately contained loops. Each schedule 
-primitive modifies this tree in some way. The compiler tracks 
-the lineage of each of the loops. This allows the compiler to 
-automatically infer the ranges for all loops.
+primitive modifies this tree in some way. Also, the compiler 
+automatically infers the ranges of all loops by tracking
+its lineage.
 
@@
    \midrule
-   \multirow{6}{*}{\texttt{split}} & \textbf{indexVar}  & \multirow{6}{*}{\parbox{0.55\linewidth}{Fiss the loop represented by \textbf{indexVar}
+   \multirow{4}{*}{\texttt{split}} & \textbf{indexVar}  & \multirow{4}{*}{\parbox{0.55\linewidth}{Fiss the loop represented by \textbf{indexVar}
    at iteration \textbf{splitIter}. Resulting loops will be represented by \textbf{first} and \textbf{second}.}} \\
@@
                                                &  \textbf{splitIter} &  \\
-                                               & & \\
-                                               & & \\
-
    \midrule
-   \multirow{4}{*}{\texttt{reorder}} & \textbf{indices[]}  & \multirow{4}{*}{\parbox{0.55\linewidth}{Permute loops corresponding to the specified index variables.
+   \multirow{3}{*}{\texttt{reorder}} & \textbf{indices[]}  & \multirow{3}{*}{\parbox{0.55\linewidth}{Permute loops corresponding to the specified index variables.
    The loops must be perfectly nested.}} \\
-                                               &   & \\
                                                &   & \\
@@
    \midrule
-   \multirow{3}{*}{\texttt{gpuDimension}} & \textbf{indexVar}  & \multirow{3}{*}{\parbox{0.55\linewidth}{Map the passed index variable to a dimension of either the grid
+   \multirow{2}{*}{\texttt{gpuDimension}} & \textbf{indexVar}  & \multirow{2}{*}{\parbox{0.55\linewidth}{Map the passed index variable to a dimension of either the grid
    or thread block.}} \\
                                                & \textbf{gpuDim}  & \\
-                                               &   &  \\
 
@@
    \midrule
-   \multirow{4}{*}{\texttt{cache}} & \textbf{indexVar}  & \multirow{4}{*}{\parbox{0.55\linewidth}{Cache the working set of one iteration of the specified loop. 
+   \multirow{3}{*}{\texttt{cache}} & \textbf{indexVar}  & \multirow{3}{*}{\parbox{0.55\linewidth}{Cache the working set of one iteration of the specified loop. 
    Cache rows for a batch loop and trees for a tree loop.}} \\
-                                               &  &  \\
                                                &  &  \\
@@
    \midrule                                               
-   \multirow{3}{*}{\texttt{interleave}} & \textbf{indexVar}  & \multirow{3}{*}{\parbox{0.55\linewidth}{Interleave tree walks within the specified 
+   \multirow{2}{*}{\texttt{interleave}} & \textbf{indexVar}  & \multirow{2}{*}{\parbox{0.55\linewidth}{Interleave tree walks within the specified 
    loop (must be innermost loop).}} \\
-                                               &  & \\
                                                &  & \\
@@
    \midrule
-   \multirow{3}{*}{\texttt{sharedReduce}} & \textbf{indexVar}  & \multirow{3}{*}{\parbox{0.55\linewidth}{Specifies that intermediate
+   \multirow{2}{*}{\texttt{sharedReduce}} & \textbf{indexVar}  & \multirow{2}{*}{\parbox{0.55\linewidth}{Specifies that intermediate
    results are to be stored in shared memory (GPU only).}} \\
-                                               &  & \\
                                                &  & \\

### 2024-04-19 docs/SOSP 2024/motivation.tex
Edit motivation figure

@@
 \begin{figure*}[]
-  \begin{subfigure}[t]{.495\linewidth}
+  \begin{subfigure}[t]{.5\linewidth}
     \vspace{0pt}
@@
   \end{subfigure}\hfill
-  \begin{subfigure}[t]{.495\linewidth}
+  \begin{subfigure}[t]{.5\linewidth}
     \vspace{0pt}
@@
   \end{subfigure}
-  \caption{Some possible schedules to compile the motivating example and the generated \Treebeard{} IR.}
+  \caption{Some possible schedules to compile the motivating example and the generated \Treebeard{} IR.
+  The arrows connect entities in the schedule to related entities in the IR. }
   \label{Fig:Schedules}
@@
 that are different from what library based systems like XGBoost, Tahoe and RAPIDs FIL use. 
-As one can imagine, several other schedules with different trade-offs can be generating using 
+As one can imagine, several other schedules with different trade-offs can be generated using 
 these constructs.

### 2024-04-19 docs/SOSP 2024/abstract.tex
pass

@@
 \begin{abstract}
-
-  The rapid proliferation of machine learning together with the accelerating evolution of 
+  The proliferation of machine learning together with the rapid evolution of 
   the hardware ecosystem has led to a surge in the demand for model inference on a variety
-  of hardware. This paper is motivated by the problems encountered when targeting 
-  inference of decision tree based models, the most popular models on tabular data,
-  to run at peak performance on diverse CPU and GPU targets. Existing solutions are neither
+  of hardware. 
+  Decision tree based models are the most popular models on tabular data.
+  This paper is motivated by the problems encountered when targeting 
+  inference of these models to run at peak performance on CPU and GPU targets.
+  Existing solutions are neither
   portable nor achieve the best possible performance for the specific hardware they 
-  target because they do not explore and customize optimization configurations to the model 
-  being used.
+  target. This is because they do not explore and customize optimization 
+  configurations to the model being used.
 
@@
   loop structures and caching strategies enables it to achieve portable performance
-  across a range of CPUs and GPUs.   
+  across a range of targets.   
   % To accomplish this level of configurability, we re-architect 

### 2024-04-19 docs/SOSP 2024/intro.tex
pass

@@
 %Parallelism, loop structure and model representation together determine the accesses patterns and the working set. 
-Existing systems today are mostly library based, and only support a predefined combination of optimizations
+Existing systems today are mostly library based, only support a predefined combination of optimizations
 and typically only target a single platform. 
@@
 %% RG
-built using the MLIR infrastructure~\cite{MLIR}, supports two fixed loop structures 
+, supports two fixed loop structures 
 % and does not scale well with increasing number of threads. Additionally, it 

### 2024-04-19 docs/SOSP 2024/conclusion.tex
Add conclusion

\section{Conclusions}
\label{sec:conclusion}
Two trends motivate the need for systems that provide portable 
performance for ML inference -- machine learning 
is becoming more ubiquitous and hardware is getting more diverse.
This paper discusses the challenges in targeting decision tree models
to run at peak performance on CPUs and GPUs. To address these, we 
design \Treebeard{}, a schedule-guided, retargetable compiler for
decision tree inference. We demonstrate that code generated 
by \Treebeard{} is significantly faster than existing systems like XGBoost, 
RAPIDS FIL and Tahoe. This is because our scheduling language is able 
express more combinations of optimization strategies and our 
schedule exploration technique is able to quickly find high-performance schedules.

### 2024-04-19 docs/SOSP 2024/caching.tex
Improve anonymization

@@
 the three representations shown in Figure \ref{Fig:Representations}. The array and sparse 
-representations are the ones described in the \TreebeardOLD{} paper\cite{Treebeard}. The reorg 
+representations are as proposed in \TreebeardOLD{}~\cite{Treebeard}. The reorg 
 representation is the representation used by the RAPIDs library\cite{FIL}. 
@@
 
-One of the major changes we make to the original design of \TreebeardOLD{} \cite{Treebeard} is to
+We change the design of \TreebeardOLD{}~\cite{Treebeard} to
 separate the implementation of representations from the rest of the compiler. This allows

### 2024-04-19 docs/SOSP 2024/hir-mir-opts.tex
pass
Change titles to save space
Improve anonymization

@@
-\section{Overview of \Treebeard{} Multi-Level Compilation}
+\section{Overview of \Treebeard{}}
 
@@
 % significant extensions were required to support schedule-guided compilation for CPUs and GPUs. 
-The \TreebeardOLD{} infrastructure was originally designed to target CPUs.
+The \TreebeardOLD{} infrastructure targets CPUs.
 It lacks a scheduling language and does not support generating code for 
-different implementation strategies. We extend \TreebeardOLD{} significantly 
+different implementation strategies. We enhance \TreebeardOLD{}  
 to support schedule-guided compilation for CPUs and GPUs.
@@
 The MIR is further lowered to a low-level IR (LIR). 
-Significant changes to the original \TreebeardOLD{} design were required to get
+Significant changes to \TreebeardOLD{} were required to get
 LIR to correctly lower to GPU code. The most important of these was 

### 2024-04-19 docs/SOSP 2024/reduction.tex
Improve anonymization

@@
 \begin{lstlisting}[style=c++]
-  float result[batch_size];
-  float partResults[2][batch_size];
+  float result[batch_size]
+  float partResults[2][batch_size]
   model = ensemble(...) 

### 2024-04-19 docs/SOSP 2024/motivation.tex
Add conclusion
Add IR disclaimer

@@
 We first describe the simple schedule shown in Figure \ref{fig:schedulea} that processes one tree at a time 
-for all input rows. \footnote{Since the scheduling constructs are fairly intuitive, we 
-defer a detailed explanation of the scheduling language to Section~\ref{sec:schedule}.} 
+for all input rows\footnote{Since the scheduling constructs are fairly intuitive, we 
+defer a detailed explanation of the scheduling language to Section~\ref{sec:schedule}.}.
 In this schedule, the loop over the trees is split into two -- one that
@@
 The concrete implementation of this schedule (in \Treebeard{}'s IR) 
-is as shown in Figure \ref{fig:schedulea}.
+is as shown in Figure \ref{fig:schedulea}\footnote{We do not show 
+the full MLIR listing of the IR. Instead, we present inference routines 
+as pseudo-code for the sake of clarity.}.
 This schedule is ideally suited for a single-core CPU. It maximizes 
@@
 In many ($\sim20$\%) instances, using the best schedule for one model 
-to compile another results in a 5$\times$ slowdown.
+on another results in a 5$\times$ slowdown.
 As we report in Section \ref{sec:results}, reusing schedules across

### 2024-04-19 docs/SOSP 2024/related.tex
Edits to related work

@@
 % optimization space and can generate code that is tailored to a specific model.
-
+As shown in Section \ref{sec:results}, \Treebeard{} outperforms
+these systems by a significant margin.
 % Finally, all of Tahoe's optimizations are designed to address GPU specific problems. 
@@
 TensorFlow\cite{TensorFlow}. While it can target both CPUs
-and GPUs, its performance is lower than that of other frameworks~\cite{Treebeard}.
+and GPUs, its performance is shown to be lower than that of other frameworks~\cite{Treebeard}.
 
@@
 image processing applications. The concept of separating the computation 
-from the schedule was pioneered by Halide and has since been adopted 
+from the schedule was effectively utilized by Halide and has since been adopted 
 by several other systems~\cite{TVM,Tiramisu,GraphIt}. 

### 2024-04-19 docs/SOSP 2024/hir-mir-opts.tex
Minor edits

@@
 with loop transformations like splitting to specialize inference code as the example 
-in Section \ref{sec:motivation} shows. Also, we enable tree padding only if the schedule 
-specifies unrolling of tree walks.
+in Section \ref{sec:motivation} shows. 
+% Also, we enable tree padding only if the schedule 
+% specifies unrolling of tree walks.
 % Figure \ref{Fig:HIRExample} shows this representation for a model 
@@
 Also, when the target processor is a GPU, the required memory transfers and kernel
-invocations are inserted into the LIR. Additionally, buffers 
+invocations are inserted into the LIR. Buffers 
 to hold model values are inserted and abstract tree operations are lowered to

### 2024-04-19 docs/SOSP 2024/results.tex
results pass

@@
 
-We evaluate \Treebeard{} on four different target processors, an NVIDIA RTX 4060 GPU (8GB RAM, CUDA 11.8),
-an NVIDIA T400 GPU (2GB RAM, CUDA 11.5), an AMD MI210 GPU (64GB RAM, ROCm 6.0.2) and an 
-Intel Core i9-11900K CPU (16 virtual cores, 128 GB RAM). We compare \Treebeard{} with 
+We evaluate \Treebeard{} on four different target processors, a low resource NVIDIA T400 GPU with 2GB RAM, a medium-tier NVIDIA RTX 4060 GPU with 8GB RAM, a large AMD MI210 GPU with 64GB RAM, and an 
+Intel Core i9 CPU  with 16 virtual cores and 128 GB RAM. We compare \Treebeard{} with 
 four other systems, NVIDIA RAPIDS\cite{RAPIDS} v23.10, Tahoe\cite{Tahoe}, XGBoost\cite{XGBoost}
@@
 We measure both kernel time and total time (including data transfer to the GPU and results back) 
-for RAPIDS and \Treebeard{}. Tahoe only allows us to measure the kernel time since it is written
+for all GPU systems except Tahoe. Tahoe only allows us to measure the kernel time since it is written
 as an executable that performs inference repeatedly on the same data that is transferred to the GPU once.
@@
 Figure~\ref{Fig:TBvsRAPIDSTahoe_4060_Speedup} shows the geomean speedup of \Treebeard{} over RAPIDS and Tahoe at different batch sizes on RTX 4060. 
-We do not show results for XGBoost since the speedups are an order of magnitude higher ($9\times$-$40\times$) and too large to fit on the same graph.  
+We do not show results for XGBoost since the kernel time speedups are an order of magnitude higher ($9\times$---$40\times$) and too large to fit on the same graph.  
 The plot has lines for kernel time and total time speedup over RAPIDS and kernel time speedup over Tahoe. As can been seen 
@@
 Compared to RAPIDS, \Treebeard{} is about $4\times$ faster at batch size 512. The relative performance of RAPIDS improves 
-with batch size as RAPIDS is optimized for larger batches, but \Treebeard{} is still faster by $1.5-2\times$ even at very high batch sizes ($16k$). 
+with batch size from 512---4096 as RAPIDS is optimized for larger batches. \Treebeard{} is still consistently faster by $1.5-2\times$ all the way up to a very large batch size of $16k$. 
 The plot also shows that the speedups are significant even when the overhead of data transfer to and from the GPU
@@
 outperforms both systems, achieving a speedup of upto $12\times$ and average speedups of over $2\times$.
-For about half the benchmarks, the speedup is $2\times$ or more over both baselines at batch size 8192.
+%For about half the benchmarks, the speedup is $2\times$ or more over both baselines at batch size 8192.
 
@@
 Next, shared reduction significantly improves performance especially for multi-class models. For example, 
-the \op{letters} and \op{covtype} benchmarks achieves speedups of 1.5$\times$ with shared reduction. 
+the \op{letters} and \op{covtype} benchmarks achieve speedups of 1.5$\times$ with shared reduction (over the best schedule without shared reduction). 
 Shared reduction increases the memory pressure (on shared memory) and not all schedules use it.
-While GPUs have a hardware managed cache, explicitly caching the rows in shared memory using the scheduling primitives 
+Finally, while GPUs have a hardware managed cache, explicitly caching the rows in shared memory using the scheduling primitives 
 helps exploit the reuse of rows across trees. We find that caching rows is beneficial for many models
@@
 \subsubsection*{Comparison on T400}
-To test the portability of \Treebeard{}'s techniques, we compare the performance 
-of \Treebeard{} on the T400 GPU with RAPIDS and Tahoe. Figure \ref{Fig:TBvsRAPIDSTahoe_T400_Speedup}
-shows the speedup of code generated by \Treebeard{} vs RAPIDS and Tahoe on the T400.
-Even on the smaller T400 GPU, \Treebeard{} is able to outperform RAPIDS
-and Tahoe consistently (speedup greater than $1.5\times$). We observe similar trends to those on the RTX 4060.
+To test the portability of \Treebeard{}'s techniques, we compare \Treebeard{} on the T400 GPU with RAPIDS and Tahoe. 
+Figure \ref{Fig:TBvsRAPIDSTahoe_T400_Speedup} shows 
+that even on the smaller GPU, \Treebeard{} is able to outperform them consistently (speedup greater than $1.5\times$). 
+As we adjust the batch size, we notice trends that are similar to those observed on the RTX 4060.
 
@@
 and these can be JIT'ed to AMD GPUs. While our objective here is not to compare directly 
-between AMD and NVIDIA GPUs, we do find that the MI210 achieves better inference times 
+between AMD and NVIDIA GPUs, we do find that the MI210 (which is more powerful) achieves better inference times 
 on most benchmarks at large batch sizes ($\geq8$k). For example, at a batch size of 16k,
@@
 Figure \ref{Fig:HeuristicVsFullExplore_Speedup} compares the best schedule found by exhaustive exploration on the RTX 4060 
-with the schedule found by the exploration heuristic on RTX 4060.
+with the schedule found by the exploration heuristic.
 The plot establishes that the 
 heuristic is able to find schedules that are very close to the best schedule (well 
-within 5\%) on 4060. 
+within 5\%). 
 % The heuristic is consistently close to two orders of magnitude faster than exhaustive exploration. 
-However, the heuristic method is able to find good schedules in a fraction($1/80$-$1/100$) of the time taken by exhaustive exploration.
+Importantly, the heuristic method is able to find good schedules in a fraction($1/80$-$1/100$) of the time taken by exhaustive exploration.
 The exploration time ranges between $6$ and $167$ seconds for the heuristic with a mean of $28.7$ seconds.
@@
 \subsection{Discussion}
-\subsubsection*{Cost Performance Tradeoff Across Hardware Platforms}
+Beyond portable performance, our evaluation indicates that \Treebeard{} has several additional advantages over existing systems.
+\subsubsection*{Cost Performance trade-off Across Hardware Platforms}
 Two of the three GPU targets we use are commodity GPUs that cost less than \$400. They 
@@
 is significantly better than on CPUs (by an order of magnitude on RTX 4060).
-We believe \Treebeard{} can help enable data scientists to better leverage 
+\Treebeard{} can therefore help data scientists make the best use of 
 available processors on their systems.
 
-\subsubsection*{A Note on Batch Size}
-Batch size is an important parameter dictated by application requirements. It 
-exposes a tradeoff between latency and throughput. It is not possible to use 
-large batch sizes in latency sensitive applications. Ideally, one would like 
+\subsubsection*{Performance considerations for batch size}
+Batch size, exposes a trade-off between latency and throughput. It is not possible to use 
+large batch sizes in latency sensitive applications. 
+Ideally, one would like 
 to maximize performance while respecting the batch size constraint imposed 
-by the application. \Treebeard{} enables users to explore this tradeoff.
+by the application. \Treebeard{} enables users to explore this trade-off.
 
+\noindent{}
 Overall, our evaluation shows that \Treebeard{} is able to efficiently generate 

### 2024-04-19 docs/SOSP 2024/results.tex
pass

@@
     \caption{The first graph shows kernel and total time speedup of \Treebeard{} over RAPIDS and Tahoe (geomean over real-world 
-    benchmarks) across batch sizes on NVIDIA RTX 4060. The bar graphs show kernel time speedup of \Treebeard{} over RAPIDS and Tahoe 
-    on NVIDIA RTX 4060. Numbers on the bars are inference times per sample in $\mu$s for RAPIDS and Tahoe.}
+    benchmarks) across batch sizes on NVIDIA RTX 4060.
+The bar graphs shows the kernel time speedup of \Treebeard{} per benchmark.
+Numbers on the bars are inference times per sample in $\mu$s for RAPIDS and Tahoe.}
 \end{figure*}
@@
     \includegraphics[width=\linewidth]{figures/RandomModels/kernel_speedup_b512_depth8.png}
-    \caption{}
+    \caption{Batch size 512}
   \end{subfigure}
@@
     \includegraphics[width=\linewidth]{figures/RandomModels/kernel_speedup_b4096_depth6.png}
-    \caption{}
+    \caption{Batch size 4096}
   \end{subfigure}
-  \caption{\label{fig:randomModels4060}Kernel time speedup of \Treebeard{} over RAPIDS for random models
-  of max depth 8 at batch size 512 and max depth 6 at batch size 4096.}
+  \caption{\label{fig:randomModels4060}Kernel time speedup of \Treebeard{} over RAPIDS for random models.
+  }
 \end{figure}
@@
 
-\subsection{Discussion}
-Beyond portable performance, our evaluation indicates that \Treebeard{} has several additional advantages over existing systems.
+\subsection{Additional Observations}
+%Beyond portable performance, our evaluation indicates that \Treebeard{} has several additional advantages over existing systems.
 \subsubsection*{Cost Performance trade-off Across Hardware Platforms}
@@
 
-\subsubsection*{Performance considerations for batch size}
+\subsubsection*{A Note on batch size}
 Batch size, exposes a trade-off between latency and throughput. It is not possible to use 
@@
 by the application. \Treebeard{} enables users to explore this trade-off.  
-
-\noindent{}
+%\subsection*{}
+\bigbreak
+\noindent
 Overall, our evaluation shows that \Treebeard{} is able to efficiently generate 
@@
 On all platforms and models that we tested on, \Treebeard{} significantly outperforms
-state-of-the-art systems like RAPIDS, Tahoe and \TreebeardOLD{}.
+state-of-the-art systems like RAPIDS, Tahoe, XGBoost and \TreebeardOLD{}.

### 2024-04-19 docs/SOSP 2024/exploring.tex
Edit results. Reduce algorithm size
Fix some typos
Edit exploration section

@@
 Section \ref{sec:schedule} is unbounded.
-\TODO{kr: next few lines need more work}
-Even if one were to search for a good schedule offline, before deploying the model 
-on a specific hardware, additional tools are needed to help search for a 
-high-performance schedule. We anticipate that even for the most complex models, 
+%\TODO{kr: next few lines need more work}
+% Even if one were to search for a good schedule offline, before deploying the model 
+% on a specific hardware, additional tools are needed to help search for a 
+% high-performance schedule. 
+Additional tools are required to help search for a high-performance schedule.
+We anticipate that even for the most complex models, 
 % one would like to spend at most a few minutes doing this search.
@@
 to partially unroll loops we did not observe any performance benefits from doing so.}} 
-and how many trees to traverse simultaneously \TODO{Should we say, in the kernel/within
-a single thread?}.  (the remaining primitives from Table~\ref{Tab:Optimizations}).  
-The last two parameters determine reduction option (Table~\ref{Tab:ReductionOpts}) and the layout{\footnote{This does not have a schedule primitive.
-We always explore all three layout options.}} to use.
+and how many trees to traverse simultaneously within
+a single thread (the remaining primitives from Table~\ref{Tab:Optimizations}).  
+The last two parameters determine reduction optimizations (Table~\ref{Tab:ReductionOpts}) 
+and the layout{\footnote{This does not have a schedule primitive. We always explore all three layout options.}} to use.
 %Table \ref{tab:schedparams} lists the parameter values we tried. 
@@
   \centering
-  \includegraphics[width=0.7\linewidth]{figures/normalized_kernel_histogram_lt5.png}
-  \caption{Distribution of normalized execution times for all benchmark models
-  with the template schedule using parameter values as shown in Table \ref{tab:schedparams}}. 
+  \includegraphics[width=0.85\linewidth]{figures/normalized_kernel_histogram_lt5.png}
+  \caption{Distribution of normalized execution times for real-world benchmark models
+  with different combinations of parameter values shown in Table \ref{tab:schedparams}}. 
   \label{Fig:ExecTimeDistribution}
@@
 Figure \ref{Fig:ExecTimeDistribution} shows the distribution of normalized execution times
-for some real-world models with different parameter values for the template schedule (inference
+for our real-world benchmark models with different parameter values (inference
 times normalized w.r.t fastest time for that model).
-As can be seen very few schedules perform close to the best while a vast majority perform poorly.
+As can be seen, very few schedules perform close to the best while a vast majority perform poorly.
 
 Exhaustive exploration over this bounded space ($5184$ schedules) is still very expensive, taking
- between $30-600$ minutes to explore the entire space for a given model. 
+between $30-300$ minutes to explore the entire space for a given model. 
 %Exploring the space for every model is infeasible.
@@
 \noindent
-\Treebeard{} performs an exhaustive search over the pruned schedules from Algorithm~\ref{alg:heuristic} to find the best schedule.
+\Treebeard{} performs an exhaustive search over the pruned schedules to find the best schedule.
 The model is compiled with each of these schedules and evaluated on a few input batches. 
@@
   \label{alg:heuristic}
-  \small{
+  \footnotesize{
   \begin{algorithmic}[1]

### 2024-04-19 docs/SOSP 2024/results.tex
pass

@@
 \Treebeard{} is able to find better schedules than both RAPIDS and Tahoe on each of the benchmarks. It consistently
-outperforms both systems, achieving a speedup of upto $12\times$ and average speedups of over $2\times$.
-%For about half the benchmarks, the speedup is $2\times$ or more over both baselines at batch size 8192.
-
+outperforms both systems, achieving a speedup of upto $12\times$.
+For about half the benchmarks, the speedup is $2\times$ or more over both baselines at batch size 8192.
 % \begin{comment}
 \subsubsection*{Synthetic models}
-To establish that \Treebeard{} can consistently find betters schedules, we performed an exhaustive comparison between \Treebeard{} and RAPIDS on 700 randomly generated models.   
+To establish that \Treebeard{} can consistently find betters schedules, we performed an exhaustive comparison between \Treebeard{} and RAPIDS on the 700 randomly generated models.   
 Figure~\ref{fig:randomModels4060} the results at batch sizes 512 and 4096 on all models, on RTX 4060. 
@@
 Figure \ref{Fig:TBvsRAPIDSTahoe_T400_Speedup} shows 
-that even on the smaller GPU, \Treebeard{} is able to outperform them consistently (speedup greater than $1.5\times$). 
+that even on the smaller GPU, \Treebeard{} is able to outperform them consistently (speedup always greater than $1.5\times$). 
 As we adjust the batch size, we notice trends that are similar to those observed on the RTX 4060.
@@
 While Tahoe, RAPIDS and XGBoost only support NVIDIA GPUs, we demonstrate that \Treebeard{} 
-is able to generate competitive code for AMD GPUs. \Treebeard{} can target AMD GPUs 
+is able to generate competitive code for AMD GPUs as well. \Treebeard{} can target AMD GPUs 
 because it generates a combination of MLIR's \op{gpu} dialect and LLVM IR 
@@
 \subsubsection*{Schedule Sensitivity Across GPUs}
-Figure \ref{Fig:AutotuningSpeedupvs4060Sched}
-compares the schedule found by the heuristic on T400 and AMD MI210 with the schedule found on the 4060.
-It shows that evaluation of the small number of schedules explored by the heuristic needs to be 
-done on each target to achieve the best performance.
-Schedules found on one GPU do not always carry over to other GPUs and often result in sub-optimal performance.
-For T400, the heuristic is able to find schedules that are 1.05$\times$ to 1.2$\times$ better 
+We designed additional experiments to evaluate whether the best schedule (as picked by the heuristic) from one GPU can be used on another.
+Figure \ref{Fig:AutotuningSpeedupvs4060Sched} reports the Geomean performance improvements of using the best schedule on T400 and MI210 compared to using the best schedule found on the RTX 4060 (for a given batch size).
+As can be seen the T400 specific schedule is 1.05$\times$ to 1.2$\times$ better 
 with the maximum difference being $2\times$ (\op{epsilon} at batch size 512).
-We find that there is a much larger variation in performance 
-between the schedules \Treebeard{} finds on the MI210 compared to the 4060 schedules. For example, 
-the geomean speedup over all benchmarks is 1.5$\times$ at batch size 16k and 
-The maximum speedup is 2$\times$ for the \op{letters} benchmark at batch size 16k.
+There is a much larger variation in performance on MI120. For example, 
+the geomean speedup over all benchmarks is 1.5$\times$ at batch size 16k (maxiumum 2$\times$ for the \op{letters} benchmark).
+In conclusion, schedules found on one GPU do not carry over to other GPUs and often result in sub-optimal performance.
+It is necessary tune schedules on each target to achieve the best performance.
 
@@
   \caption{Slowdown of heuristic schedule vs best schedule on NVIDIA RTX 4060 and 
-  heuristic schedule exploration time normalized w.r.t full schedule exploration time}
+  heuristic schedule exploration time normalized w.r.t full schedule exploration time.}
   \label{Fig:HeuristicVsFullExplore_Speedup}
@@
   \includegraphics[width=\linewidth]{figures/geomean_speedup_T400_4060_vs_T400_vs_MI2160.png}
-  \caption{Geomean speedup of heuristic schedule on NVIDIA T400 and AMD MI2160 vs 
-  the heuristic schedule on NVIDIA RTX 4060 across batch sizes}
+  \caption{Geomean speedup of using a hardware specific schedule on NVIDIA T400 and AMD MI210 over the best schedule from a different hardware (NVIDIA RTX 4060).}
   \label{Fig:AutotuningSpeedupvs4060Sched}

### 2024-04-20 docs/SOSP 2024/conclusion.tex
Put conclusions in past tense and more refinement

@@
 \label{sec:conclusion}
-Two trends motivate the need for systems that provide portable 
-performance for ML inference -- machine learning 
-is becoming more ubiquitous and hardware is getting more diverse.
-This paper discusses the challenges in targeting decision tree models
-to run at peak performance on CPUs and GPUs. To address these, we 
-design \Treebeard{}, a schedule-guided, retargetable compiler for
-decision tree inference. We demonstrate that code generated 
-by \Treebeard{} is significantly faster than existing systems like XGBoost, 
-RAPIDS FIL and Tahoe. This is because our scheduling language is able 
-express more combinations of optimization strategies and our 
-schedule exploration technique is able to quickly find high-performance schedules.
+Two trends motivate the need for systems that provide portable performance for
+ML inference -- machine learning is becoming more ubiquitous and hardware is
+getting more diverse.  This paper discussed the challenges in targeting decision
+tree models to run at peak performance on CPUs and GPUs. To address these, we
+designed \Treebeard{}, a schedule-guided, retargetable compiler for decision
+tree inference. We demonstrated that code generated by \Treebeard{} is
+significantly faster than existing systems like XGBoost, RAPIDS FIL and Tahoe.
+We obtained such improvements because our scheduling language was able express
+more combinations of optimization strategies, and our schedule exploration
+technique was able to quickly find high-performance schedules.

### 2024-04-20 docs/SOSP 2024/results.tex
Full stop after captions

--- 2024-04-19 docs/SOSP 2024/results.tex
@@
       \includegraphics[width=\textwidth]{figures/geomean_speedup_4060_kernel_time_total_time.png}
-      \caption{\label{Fig:TBvsRAPIDSTahoe_4060_Speedup}Speedup vs batch size}      
+      \caption{\label{Fig:TBvsRAPIDSTahoe_4060_Speedup}Speedup vs batch size.}      
     \end{subfigure}
@@
       \includegraphics[width=\textwidth]{figures/speedup_bar_graph_1024.png}
-      \caption{\label{Fig:KernelTimeIndividualBenchmarks4060a} Batch size 1024}
+      \caption{\label{Fig:KernelTimeIndividualBenchmarks4060a} Batch size 1024.}
     \end{subfigure}
@@
       \includegraphics[width=\textwidth]{figures/speedup_bar_graph_8192.png}
-      \caption{\label{Fig:KernelTimeIndividualBenchmarks4060b} Batch size 8192}
+      \caption{\label{Fig:KernelTimeIndividualBenchmarks4060b} Batch size 8192.}
     \end{subfigure}
@@
     \includegraphics[width=\linewidth]{figures/RandomModels/kernel_speedup_b512_depth8.png}
-    \caption{Batch size 512}
+    \caption{Batch size 512.}
   \end{subfigure}
@@
     \includegraphics[width=\linewidth]{figures/RandomModels/kernel_speedup_b4096_depth6.png}
-    \caption{Batch size 4096}
+    \caption{Batch size 4096.}
   \end{subfigure}
@@
   \caption{Geomean speedup of heuristic schedule on NVIDIA T400 and AMD MI2160 vs 
-  the heuristic schedule on NVIDIA RTX 4060 across batch sizes}
+  the heuristic schedule on NVIDIA RTX 4060 across batch sizes.}
   \label{Fig:AutotuningSpeedupvs4060Sched}

### 2024-04-20 docs/SOSP 2024/motivation.tex
Minor edit

@@
 is as shown in Figure \ref{fig:schedulea}\footnote{We do not show 
-the full MLIR listing of the IR. Instead, we present inference routines 
+the full listing of the IR. Instead, we present inference routines 
 as pseudo-code for the sake of clarity.}.

### 2024-04-20 docs/SOSP 2024/related.tex
Changes to related work

@@
 \label{Sec:Related}
-While several optimization strategies for decision tree based models have been 
-studied in the literature, to the best of our knowledge, no systems that are 
-capable of exploring the full optimization space exist. We describe related work 
-and compare these systems to \Treebeard{} in this section.
+% While several optimization strategies for decision tree based models have been 
+% studied in the literature, to the best of our knowledge, no systems that are 
+% capable of exploring the full optimization space exist.
+This section discusses prior work related to \Treebeard{}.
 
@@
 between four predefined strategies to implement decision tree inference on GPUs.
-In comparison, \Treebeard{} explores a much larger set of implementation options and 
-in-memory representations for models and picks the best.
 % Also, \Treebeard{} generates code that is specific to 
@@
 RAPIDS FIL\cite{FIL}, the most widely used GPU library for decision tree inference
-implements some heuristics to pick a good configuration for every model, 
-but these techniques are limited and the library uses a single strategy 
-and in-memory representation for all models. XGBoost \cite{XGBoost} also implements GPU 
-support\cite{XGBGPU} but uses a single strategy and in-memory representation. 
+% implements some heuristics to pick a good configuration for every model, 
+% but these techniques are limited and the library uses a single strategy 
+% and in-memory representation for all models. 
+and XGBoost's\cite{XGBoost} GPU library\cite{XGBGPU} 
+use a single strategy and in-memory representation.
+In comparison, \Treebeard{} explores a much larger set of implementation options and 
+in-memory representations for models and picks the best.
 % In contrast, \Treebeard{} is a compiler that can explore a much larger
@@
 % This can be the subject of a separate future work.
+
+Some compilers for decision tree ensembles have been proposed in the 
+literature \cite{Treelite, Treebeard, Hummingbird}. \TreebeardOLD{} and Treelite
+exclusively target CPUs and all their optimizations are designed purely for 
+performance on CPUs. 
+Treelite\cite{Treelite} is a model compiler that only  
+generates \op{if-else} code for each tree in the model. 
+\TreebeardOLD{} is the work most closely related to \Treebeard{}. While we 
+build on top of \TreebeardOLD{}, \Treebeard{} is a significant enhancement 
+over \TreebeardOLD{}. 
+Hummingbird\cite{Hummingbird} is a compiler that compiles traditional ML models
+to tensor operations, thereby enabling them to be run using tensor-based frameworks like
+PyTorch\cite{PyTorch} on CPUs and GPUs. As reported in their paper, Hummingbird's performance
+on decision tree models is comparable to that of RAPIDS on GPUs. On CPU, \TreebeardOLD{} 
+is faster~\cite{Treebeard}.
+% While it can target both CPUs
+% and GPUs, its performance has been shown to be lower than that of other frameworks~\cite{Treebeard}.
+
 On CPUs, XGBoost\cite{XGBoost}, LightGBM\cite{LightGBM} and
-scikit-learn\cite{Sklearn} are extremely popular. However, 
-as mentioned in Section \ref{sec:intro}, none of these systems
-provide portable performance across different target machines.
-\TODO{Write about the PACT paper and whether our scheduling language can represent 
-all the schedules they propose.}
+scikit-learn\cite{Sklearn} are extremely popular.
+A recent paper~\cite{PACTVanLunteren} describes an adaptive mechanism 
+to pick one of a few predefined parallelization and vectorization strategies.
 Other systems that hide dependency stalls by interleaving tree walks\cite{VPred},
@@
 on CPUs and GPUs\cite{Jansson2014gpuRFAG, Nasridinov2013DecisionTC}.
-
-\emph{Decision Tree Ensemble Compilers:}
-Several compilers for decision tree ensembles have been proposed in the 
-literature \cite{Treelite, Treebeard, Hummingbird}. \TreebeardOLD{} and Treelite
-exclusively target CPUs and all their optimizations are designed purely for 
-performance on CPUs. 
-Treelite\cite{Treelite} is a model compiler that only  
-generates \op{if-else} code for each tree in the model. 
-% Its code generator is hard-coded 
-% to expand trees in an ensemble into a series of \op{if-else} statements (one for 
-% each node in a tree). Due to this, extending Treelite and reusing it to 
-% target GPUs is not possible.
-\TreebeardOLD{} is the work most closely related to \Treebeard{}. While we 
-build on top of \TreebeardOLD{}, \Treebeard{} is a significant enhancement 
-over \TreebeardOLD{}. 
-% Most importantly, we re-architect \TreebeardOLD{} in 
-% order to generate code for multiple target processors.
-% Specifically, we introduce the scheduling language and schedule exploration 
-% while also enhancing the IRs and support for parallelizing 
-% across trees through the implementation of a novel MLIR reduction dialect.
-% It cannot perform any of the optimizations \Treebeard{} is designed to perform. 
-Hummingbird\cite{Hummingbird} is a compiler that compiles traditional ML models
-to tensor operations, thereby enabling them to be run on tensor-based frameworks like
-TensorFlow\cite{TensorFlow}. While it can target both CPUs
-and GPUs, its performance is shown to be lower than that of other frameworks~\cite{Treebeard}.
-
-% Hummingbird\cite{Hummingbird} compiles traditional ML models to make use of tensor primitives so that
-% they can be integrated into tensor-based frameworks like TensorFlow \cite{TensorFlow}.
-% While Hummingbird does compile decision tree based models to both CPUs and GPUs, the primary aim of 
-% the system is to leverage progress in tensor compilers by representing ML models as tensor operations. 
-% Consequently, Hummingbird does not perform any optimizations tailored to decision trees. It 
-% also does not perform model specific optimizations that are enabled by \Treebeard{}'s abstractions.
-
-% Asadi et. al.\cite{VPred} optimize tree walks by hiding dependency stalls 
-% by interleaving tree walks. In contrast, \Treebeard{} is an extensible 
-% optimizing compiler that carefully implements this and many other
-% optimizations at different levels of abstraction.
-% QuickScorer\cite{QuickScorer, QuickScorer1} is an algorithm that uses 
-% bit manipulation to compute tree predictions. Even though QuickScorer is 
-% extremely fast for smaller models, it does not scale well to larger 
-% models\cite{ProbBasedLayout}. The goal of QuickScorer is orthogonal 
-% to the goals of \Treebeard{} and the QuickScorer algorithm can easily  
-% be integrated into \Treebeard{} as another traversal strategy for the 
-% system to explore. Tang et. al.\cite{CacheConscious1} and Jin et. al.\cite{CacheConscious2}
-% build models to predict cache performance of decision tree ensembles on CPUs.
-% This work is again orthogonal to the work described in this paper. Some systems have been proposed 
-% to parallelize decision tree training on CPUs and GPUs\cite{Jansson2014gpuRFAG, Nasridinov2013DecisionTC}.
-
-%Two recently published systems explore 
-%using GPUs for decision tree inference\cite{Tahoe, Hummingbird}.
+However, none of these systems provide portable performance across different target machines.
+While the specifics may vary, \Treebeard{} supports compiler optimizations that 
+achieve similar results as techniques in these systems.
 
@@
 intermediate language and a virtual machine to enable vector execution of decision tree
-inference. However, this virtual machine is itself implemented by 
+inference. However, this virtual machine is implemented by 
 hand on different target processors.
@@
 Jo et. al.\cite{MilindTreeVectorization} describe code transformations and runtime 
-techniques that help vectorize tree-based applications. However, they do not 
-study optimizations specific to decision trees.
+techniques that vectorize tree-based applications but these 
+optimizations are not specific to decision trees.
 % Both these works vectorize tree walks by performing different tree walks on each
@@
 % therefore need repeated effort to implement on each target. 
-Inspector-executor systems \cite{TaoOfParallelism,HybridCPUGPU} have been 
-developed to parallelize tree walks but are not a good fit 
+Inspector-executor systems \cite{TaoOfParallelism,HybridCPUGPU}  
+parallelize tree walks but are not a good fit 
 for decision tree inference as the individual node predicates are 
@@
 
-\emph{Code Generation Systems from Other Domains:}
-Several optimizing compilers and code generation techniques have been developed 
-for other domains. TVM\cite{TVM}, Tiramisu\cite{Tiramisu}, and Tensor 
+\emph{Code Generation Systems for Other Domains:}
+% Several optimizing compilers and code generation techniques have been developed 
+% for other domains. 
+TVM\cite{TVM}, Tiramisu\cite{Tiramisu}, and Tensor 
 Comprehensions\cite{TensorComprehensions} are optimizing compilers 

### 2024-04-20 docs/SOSP 2024/related.tex
Fix typos

@@
 to tensor operations, thereby enabling them to be run using tensor-based frameworks like
-PyTorch\cite{PyTorch} on CPUs and GPUs. As reported in their paper, Hummingbird's performance
+PyTorch\cite{NEURIPS2019_9015} on CPUs and GPUs. As reported in their paper, Hummingbird's performance
 on decision tree models is comparable to that of RAPIDS on GPUs. On CPU, \TreebeardOLD{} 

### 2024-04-20 docs/SOSP 2024/results.tex
Change section title. Move figures in results
Fix typos

--- 2024-04-19 docs/SOSP 2024/results.tex
@@
 \label{sec:results}
+
+\begin{figure*}[htb]
+  \begin{subfigure}[t]{.32\linewidth}
+    \includegraphics[width=\textwidth]{figures/geomean_speedup_4060_kernel_time_total_time.png}
+    \caption{\label{Fig:TBvsRAPIDSTahoe_4060_Speedup}Speedup vs batch size.}      
+  \end{subfigure}
+  \begin{subfigure}[t]{.32\linewidth}
+    \includegraphics[width=\textwidth]{figures/speedup_bar_graph_1024.png}
+    \caption{\label{Fig:KernelTimeIndividualBenchmarks4060a} Batch size 1024.}
+  \end{subfigure}
+  \begin{subfigure}[t]{.32\linewidth}
+    \includegraphics[width=\textwidth]{figures/speedup_bar_graph_8192.png}
+    \caption{\label{Fig:KernelTimeIndividualBenchmarks4060b} Batch size 8192.}
+  \end{subfigure}
+  % \hfill
+  \caption{The first graph shows kernel and total time speedup of \Treebeard{} over RAPIDS and Tahoe (geomean over real-world 
+  benchmarks) across batch sizes on NVIDIA RTX 4060.
+The bar graphs shows the kernel time speedup of \Treebeard{} per benchmark.
+Numbers on the bars are inference times per sample in $\mu$s for RAPIDS and Tahoe.}
+\end{figure*}
 
@@
 is included. They are lower than kernel time speedup as both systems have a constant additional transfer overhead.
-
-\begin{figure*}[htb]
-    \begin{subfigure}[t]{.32\linewidth}
-      \includegraphics[width=\textwidth]{figures/geomean_speedup_4060_kernel_time_total_time.png}
-      \caption{\label{Fig:TBvsRAPIDSTahoe_4060_Speedup}Speedup vs batch size}      
-    \end{subfigure}
-    \begin{subfigure}[t]{.32\linewidth}
-      \includegraphics[width=\textwidth]{figures/speedup_bar_graph_1024.png}
-      \caption{\label{Fig:KernelTimeIndividualBenchmarks4060a} Batch size 1024}
-    \end{subfigure}
-    \begin{subfigure}[t]{.32\linewidth}
-      \includegraphics[width=\textwidth]{figures/speedup_bar_graph_8192.png}
-      \caption{\label{Fig:KernelTimeIndividualBenchmarks4060b} Batch size 8192}
-    \end{subfigure}
-    % \hfill
-    \caption{The first graph shows kernel and total time speedup of \Treebeard{} over RAPIDS and Tahoe (geomean over real-world 
-    benchmarks) across batch sizes on NVIDIA RTX 4060.
-The bar graphs shows the kernel time speedup of \Treebeard{} per benchmark.
-Numbers on the bars are inference times per sample in $\mu$s for RAPIDS and Tahoe.}
-\end{figure*}
 
@@
 \end{figure}
+
+\begin{figure*}[h]
+  \begin{minipage}[t]{.3\linewidth}
+    \centering
+    \includegraphics[width=\linewidth]{figures/geomean_speedup_T400_kernel_time.png}
+    \caption{Speedup of \Treebeard{} over RAPIDS and Tahoe (geomean over real-world 
+    benchmarks) across batch sizes on NVIDIA T400.}
+    \label{Fig:TBvsRAPIDSTahoe_T400_Speedup}
+  \end{minipage}
+  \hspace{0.5cm}
+  \begin{minipage}[t]{.3\linewidth}
+    \centering
+    \includegraphics[width=\linewidth]{figures/speedup_vs_norm_time_line_graph_4060.png}
+    \caption{Slowdown of heuristic schedule vs best schedule on NVIDIA RTX 4060 and 
+    heuristic schedule exploration time normalized w.r.t full schedule exploration time.}
+    \label{Fig:HeuristicVsFullExplore_Speedup}
+  \end{minipage}
+  \hspace{0.5cm}
+  \begin{minipage}[t]{.3\linewidth}
+    \centering
+    \includegraphics[width=\linewidth]{figures/geomean_speedup_T400_4060_vs_T400_vs_MI2160.png}
+    \caption{Geomean speedup of using a hardware specific schedule on NVIDIA T400 and AMD MI210 over the best schedule from a different hardware (NVIDIA RTX 4060).}
+    \label{Fig:AutotuningSpeedupvs4060Sched}
+  \end{minipage}
+  \end{figure*}
 
@@
 There is a much larger variation in performance on MI120. For example, 
-the geomean speedup over all benchmarks is 1.5$\times$ at batch size 16k (maxiumum 2$\times$ for the \op{letters} benchmark).
+the geomean speedup over all benchmarks is 1.5$\times$ at batch size 16k (maximum 2$\times$ for the \op{letters} benchmark).
 In conclusion, schedules found on one GPU do not carry over to other GPUs and often result in sub-optimal performance.
 It is necessary tune schedules on each target to achieve the best performance.
-
-\begin{figure*}[htb]
-\begin{minipage}[t]{.3\linewidth}
-  \centering
-  \includegraphics[width=\linewidth]{figures/geomean_speedup_T400_kernel_time.png}
-  \caption{Speedup of \Treebeard{} over RAPIDS and Tahoe (geomean over real-world 
-  benchmarks) across batch sizes on NVIDIA T400.}
-  \label{Fig:TBvsRAPIDSTahoe_T400_Speedup}
-\end{minipage}
-\hspace{0.5cm}
-\begin{minipage}[t]{.3\linewidth}
-  \centering
-  \includegraphics[width=\linewidth]{figures/speedup_vs_norm_time_line_graph_4060.png}
-  \caption{Slowdown of heuristic schedule vs best schedule on NVIDIA RTX 4060 and 
-  heuristic schedule exploration time normalized w.r.t full schedule exploration time.}
-  \label{Fig:HeuristicVsFullExplore_Speedup}
-\end{minipage}
-\hspace{0.5cm}
-\begin{minipage}[t]{.3\linewidth}
-  \centering
-  \includegraphics[width=\linewidth]{figures/geomean_speedup_T400_4060_vs_T400_vs_MI2160.png}
-  \caption{Geomean speedup of using a hardware specific schedule on NVIDIA T400 and AMD MI210 over the best schedule from a different hardware (NVIDIA RTX 4060).}
-  \label{Fig:AutotuningSpeedupvs4060Sched}
-\end{minipage}
-\end{figure*}
 
@@
 Two of the three GPU targets we use are commodity GPUs that cost less than \$400. They 
-are infact much cheaper than the CPU we evaluate on. Despite this, performance on GPUs
+are in fact much cheaper than the CPU we evaluate on. Despite this, performance on GPUs
 is significantly better than on CPUs (by an order of magnitude on RTX 4060).

### 2024-04-20 docs/SOSP 2024/hir-mir-opts.tex
Change section title. Move figures in results
Minor edits

@@
-\section{Overview of \Treebeard{}}
+\section{Overview of \Treebeard{} Compiler}
 
@@
 finally lowered to LLVM and then JIT'ed to the specified target processor. 
-\Treebeard{} is built using the open-source \TreebeardOLD{} infrastructure~\cite{Treebeard}.
+\Treebeard{} is built using the open-source \TreebeardOLD{} compiler~\cite{Treebeard}
 % Since the \TreebeardOLD{} infrastructure was originally designed to target CPUs,
 % significant extensions were required to support schedule-guided compilation for CPUs and GPUs. 
-The \TreebeardOLD{} infrastructure targets CPUs.
-It lacks a scheduling language and does not support generating code for 
+that targets CPUs.
+\TreebeardOLD{} lacks a scheduling language and does not support code generation for 
 different implementation strategies. We enhance \TreebeardOLD{}  
 to support schedule-guided compilation for CPUs and GPUs.
-The parts \Treebeard{} that are new or significantly different compared to \TreebeardOLD{}
+The parts of \Treebeard{} that are new or significantly different compared to \TreebeardOLD{}
 are shown as shaded boxes in Figure \ref{Fig:CompilerStructure}.
@@
 allows the implementation of optimizations that require the manipulation of the model
-or its constituent trees. We extend the \TreebeardOLD{} infrastructure with loop rewrites on 
+or its constituent trees. We extend \TreebeardOLD{} with loop rewrites on 
 the HIR that are implemented through the scheduling language (Section \ref{sec:schedule}).

### 2024-04-20 docs/SOSP 2024/results.tex
Fix typos

@@
   benchmarks) across batch sizes on NVIDIA RTX 4060.
-The bar graphs shows the kernel time speedup of \Treebeard{} per benchmark.
+The bar graphs show the kernel time speedup of \Treebeard{} per benchmark.
 Numbers on the bars are inference times per sample in $\mu$s for RAPIDS and Tahoe.}
@@
 
-\subsubsection*{Real-world models}
+\subsubsection*{Real-world Models}
 Figure~\ref{Fig:TBvsRAPIDSTahoe_4060_Speedup} shows the geomean speedup of \Treebeard{} over RAPIDS and Tahoe at different batch sizes on RTX 4060. 
@@
 % \begin{comment}
-\subsubsection*{Synthetic models}
+\subsubsection*{Synthetic Models}
 To establish that \Treebeard{} can consistently find betters schedules, we performed an exhaustive comparison between \Treebeard{} and RAPIDS on the 700 randomly generated models.   
@@
 We designed additional experiments to evaluate whether the best schedule (as picked by the heuristic) from one GPU can be used on another.
-Figure \ref{Fig:AutotuningSpeedupvs4060Sched} reports the Geomean performance improvements of using the best schedule on T400 and MI210 compared to using the best schedule found on the RTX 4060 (for a given batch size).
+Figure \ref{Fig:AutotuningSpeedupvs4060Sched} reports the geomean performance improvements of using the best schedule on T400 and MI210 compared to using the best schedule found on the RTX 4060 (for a given batch size).
 As can be seen the T400 specific schedule is 1.05$\times$ to 1.2$\times$ better 
@@
 In conclusion, schedules found on one GPU do not carry over to other GPUs and often result in sub-optimal performance.
-It is necessary tune schedules on each target to achieve the best performance.
+It is necessary to tune schedules on each target to achieve the best performance.
 
@@
 %Beyond portable performance, our evaluation indicates that \Treebeard{} has several additional advantages over existing systems.
-\subsubsection*{Cost Performance trade-off Across Hardware Platforms}
+\subsubsection*{Cost Performance Trade-off Across Hardware Platforms}
 Two of the three GPU targets we use are commodity GPUs that cost less than \$400. They 
@@
 
-\subsubsection*{A Note on batch size}
+\subsubsection*{A Note on Batch Size}
 Batch size, exposes a trade-off between latency and throughput. It is not possible to use 

### 2024-04-22 docs/SOSP 2024/exploring.tex
Final submission

@@
 Additional tools are required to help search for a high-performance schedule.
-We anticipate that even for the most complex models, 
+Like with other schedule-guided ML compilers~\cite{TVM},
+we expect schedule exploration will be performed offline (before a model deployment),
+% But still, even for the most complex models, 
 % one would like to spend at most a few minutes doing this search.
-spending more than a few minutes on this search is not acceptable.
+but spending more than a few minutes on this search is likely not acceptable.
 We propose a set of heuristics to guide the search for a good schedule 
@@
 that the space (at-least) covers the known strategies published in prior work.
-Specifically, we use 4 numeric parameters, 4 boolean parameters and one ternary parameter as listed in Table \ref{tab:schedparams}.
+Specifically, we use the 9 parameters listed in Table \ref{tab:schedparams}.
 We picked the specific values listed through experimentation.

### 2024-04-22 docs/SOSP 2024/related.tex
Minor edits
Final submission

@@
 % optimization space and can generate code that is tailored to a specific model.
-As shown in Section \ref{sec:results}, \Treebeard{} outperforms
-these systems by a significant margin.
+As shown in Section \ref{sec:results}, \Treebeard{} outperforms these systems
+by a significant margin.
 % Finally, all of Tahoe's optimizations are designed to address GPU specific problems. 
@@
 and improve cache performance of decision tree ensembles on CPUs\cite{CacheConscious1, CacheConscious2}
-have been proposed in prior work.
+have been proposed.
 Some systems have been proposed to parallelize decision tree training 
@@
 However, none of these systems provide portable performance across different target machines.
-While the specifics may vary, \Treebeard{} supports compiler optimizations that 
-achieve similar results as techniques in these systems.
+While the specifics may vary, \Treebeard{} supports retargetable compiler optimizations that 
+achieve similar results. % as techniques in these systems.
 
@@
 \emph{Reductions:} CUB\cite{CUB} and Thrust\cite{Thrust} are libraries 
-that implement high-performance parallel reductions on GPUs. 
+that implement reductions on GPUs. 
 However, it is not possible to fuse these functions with other computations  
@@
 describe language constructs in {\sc {Pencil}}~\cite{Pencil}
-to express reductions and to represent and optimize them using the polyhedral 
+to express reductions and optimize them using the polyhedral 
 framework. 
@@
 reductions and also only targets GPUs. Suriana et. al.~\cite{HalideReductions}
-extend Halide to add support for factoring reductions in the Halide 
+extend Halide to add support for reductions in the Halide 
 scheduling language and to synthesize reduction operators. De Gonzalo 

### 2024-07-23 docs/SOSP 2024/rebuttal.txt
Add rebuttal

Q1: How does SilvanForge’s decision-forest-based ‘schedule space’ and scheduling language differ from prior languages for other domains, e.g., TVM’s? For instance, what operators are new to SilvanForge?

The novelty of SivanForge's scheduling language lies in the non-trivial realization that a suitably enhaced loop centric scheduling language elegantly encapsulates the optimization space for an irregular application like decision tree inference. SilvanForge’s scheduling language naturally adds scheduling primitives that are specific to decision tree inference (to represent walk interleaving, walk unrolling, reductions etc.) similar to how TVM extended Halide's scheduling language with tensor specific primitives, namely tensorization and cooperative data loading. Importantly, we note that none of the several prior systems that address decision tree inference (RAPIDs, Tahoe, Hummingbird, XGBoost etc) have a scheduling language. In contrast, SilvanForge’s scheduling language is general enough to represent the specific optimizations that are used in several previous systems and therefore unifies several existing techniques. We believe that the design of SilvanForge's scheduling language is conceptually satisfying because it shows that extending a core set of loop transformation primitives with a few domain specific primitives can capture the optimization space of a complex application like decision tree inference (similar to how TVM extensions to the scheduling language were sufficient to compile tensor based applications).

TODO: Can we say the core contribution is the realization that a schedule space exists for decision tree inference, showing that there is huge variation in performance and then realizing that a scheduling language is the right way to capture this space?

Q2: Which SilvanForge optimizations are least likely to be supported by PyTorch (in the context of a decision forest compiled by modern Hummingbird), and why?
* HB executes one walk per thread and fully walks all trees for a row in a single thread (as far as I can see from the paper). Therefore, SilvanForge will be able to find better schedules. 
* Expressing tree walks as a set of tensor operations makes it impossible for the compiler to perform optimizations like walk unrolling and walk interleaving. Also, this would necessitate several kernel calls (for each step of the tree walk and then subseqently for reductions). This representation also means that HB cannot perform optimizations like reductions in shared memory which SilvanForge can. 
* HB is hard-coded to use one kind of representation which is not ideal for all models and target machines. 
* TODO should we also say how TB cannot match up on the CPU? And focus on how SF improves even on TB with the schedule?

Q: You claim that SilvanForge decouples schedule from the backend architecture  but it seems both of these would change substantially with CPU? What fraction of your compiler changes are 'portable' across platforms, and how well does the CPU search heuristic work?
A: The scheduling language, HIR and MIR parts of the compiler are fully shared. Most of the backend is shared as well. Maybe a good idea to talk about how we share low level code gen between CPU and GPU? 
We found that a search heuristic is not needed on the GPU. Exploring only a few configurations is sufficient to find the best schedule.

Q: How much do different optimizations contribute to overall performance of code generated by SILVANFORGE?
A: TODO Mention what we've said about perf improvements due to shared reduction and row caching?

Q: There might be other metrics besides kernel time speedup. For instance, the memory footprint of the resulting model.
While we did not evaluate such metrics, it should not be difficult to optimize for any other such metrics. We believe that the scheduling language is expressive enough to optimize for other metrics.
-------------------------------------------------------------------------

Reviewer A

* [Done] There is very little conceptual novelty; similar approaches are widely used for neural network optimization and graphics code (Halide).
* Very limited evaluation (only 8 real-world benchmarks), comparison to important related work appears to be missing.
* Past work has focused on throughput as opposed to latency, and I think this paper seems to conflate them somewhat. I think the evaluation would be stronger if it separately looked at throughput and latency; it may be possible that optimizations in the scheduling space optimize one at the cost of the other.
* [Done] The main issue, in my view, is that the paper is missing a comparison against Hummingbird

Reviewer B
* [Done] Table 5 (scheduler search space) and Algorithm 1 (search heuristic) seem rather GPU specific. You claim that SilvanForge decouples schedule from the backend architecture  but it seems both of these would change substantially with CPU? What fraction of your compiler changes are 'portable' across platforms, and how well does the CPU search heuristic work?

Reviewer C
* [Done] Not clear about novelty over other scheduling languages
* Autotuning is ad hoc

Reviewer D
* Evaluated only on benchmarks. Unclear how well it will work in a practical setup. Specifically, its speedup diminishes on large batch sizes.
* Can you show the absolute time? All the numbers are shown as relative speedups. But if the absolute time is already fast enough, how much do practitioners care about the improved inference performance?

Reviewer E
* [Done] Unclear how the ideas here are different from ideas presented in other programmer-directed scheduling systems like Halide.
* [Done] There might be other metrics besides kernel time speedup. For instance, the memory footprint of the resulting model. In particular, the choice of representation (in Section 6) might affect this footprint. I would have liked to see an evaluation of this.
* [Done] How much do different optimizations contribute to overall performance of code generated by SILVANFORGE?
* [Done] Could you comment on the main decision-tree-specific insights in the scheduling language here in comparison to prior scheduling languages for other domains?

### 2024-07-23 docs/SOSP 2024/rebuttal.txt
Minor edits to rebuttal

@@
 The novelty of SivanForge's scheduling language lies in the non-trivial realization that a suitably enhaced loop centric scheduling language elegantly encapsulates the optimization space for an irregular application like decision tree inference.
-SilvanForge’s scheduling language naturally adds scheduling primitives that are specific to decision tree inference (to represent walk interleaving, walk unrolling, reductions etc.) similar to how TVM extended Halide's scheduling language with tensor specific primitives, namely tensorization and cooperative data loading.
+SilvanForge’s scheduling language naturally adds scheduling primitives that are specific to decision tree inference (to represent walk interleaving, walk unrolling, reductions etc.) which cannot be supported with existing scheduling languages.
+
+This is similar to how TVM provided tensor specific primitives, namely tensorization and cooperative data loading.
 Importantly, we note that none of the several prior systems that address decision tree inference (RAPIDs, Tahoe, Hummingbird, XGBoost etc) have a scheduling language.
@@
 Q2: Which SilvanForge optimizations are least likely to be supported by PyTorch (in the context of a decision forest compiled by modern Hummingbird), and why?
+* Mention how tensor representations are not a suitable abstraction to express all the transformations that SF can do
 * HB executes one walk per thread and fully walks all trees for a row in a single thread (as far as I can see from the paper). Therefore, SilvanForge will be able to find better schedules. 
@@
 While we did not evaluate such metrics, it should not be difficult to optimize for any other such metrics. We believe that the scheduling language is expressive enough to optimize for other metrics.
+
+Q: It's really surprising to me that MLIR doesn't support reductions well -- they seem pretty fundamental to many kernels that other compilers targeting MLIR would see.
+But this section doesn't really explain if there's anything specific to the decision tree world here, or if this is just a general-purpose "add reductions to MLIR" thing.
+Either is probably interesting, but I didn't get enough context from the paper to know the answer.
+
+Q: Specifically, its speedup diminishes on large batch sizes.
+
 -------------------------------------------------------------------------

### 2024-07-24 docs/SOSP 2024/rebuttal.txt
Edit rebuttal

@@
 
-The novelty of SivanForge's scheduling language lies in the non-trivial realization that a suitably enhaced loop centric scheduling language elegantly encapsulates the optimization space for an irregular application like decision tree inference.
+One of our paper's core contributions is identifying that a schedule space exists for decision tree inference, showing that there is a huge variation in performance depending on which schedule is picked and  then designing a scheduling language that cleanly encapsulates this space on both CPUs and GPUs.
+To the best of our knowledge, ours is the first work to study the design of a scheduling language for the domain of decision tree inference.
+While several prior works proposed various techniques to optimize decision tree inference, all of them (RAPIDs, Tahoe, XGBoost etc) use a fixed set of techniques.
+In contrast, SilvanForge’s scheduling language is general enough to represent the specific optimizations that are used in several previous systems and therefore unifies several existing techniques.
+
+SilvanForge's scheduling language is novel because of the non-obvious realization that a suitably enhaced loop centric scheduling language elegantly encapsulates the optimization space for an irregular application like decision tree inference.
 SilvanForge’s scheduling language naturally adds scheduling primitives that are specific to decision tree inference (to represent walk interleaving, walk unrolling, reductions etc.) which cannot be supported with existing scheduling languages.
-
-This is similar to how TVM provided tensor specific primitives, namely tensorization and cooperative data loading.
-Importantly, we note that none of the several prior systems that address decision tree inference (RAPIDs, Tahoe, Hummingbird, XGBoost etc) have a scheduling language.
-In contrast, SilvanForge’s scheduling language is general enough to represent the specific optimizations that are used in several previous systems and therefore unifies several existing techniques.
+This is reminiscent of how TVM provided tensor specific primitives, namely tensorization and cooperative data loading.
 We believe that the design of SilvanForge's scheduling language is conceptually satisfying because it shows that extending a core set of loop transformation primitives with a few domain specific primitives can capture the optimization space of a complex application like decision tree inference (similar to how TVM extensions to the scheduling language were sufficient to compile tensor based applications).
-
-TODO: Can we say the core contribution is the realization that a schedule space exists for decision tree inference, showing that there is huge variation in performance and then realizing that a scheduling language is the right way to capture this space?
 
@@
 
-Q: You claim that SilvanForge decouples schedule from the backend architecture  but it seems both of these would change substantially with CPU?
+RB Q: You claim that SilvanForge decouples schedule from the backend architecture  but it seems both of these would change substantially with CPU?
 What fraction of your compiler changes are 'portable' across platforms, and how well does the CPU search heuristic work?
@@
 
-Q: How much do different optimizations contribute to overall performance of code generated by SILVANFORGE?
-A: TODO Mention what we've said about perf improvements due to shared reduction and row caching?
+RC Q: Although I wonder if at least some of the explanation is that so many people have moved on from decision tree models to the hot new generative thing.
 
-Q: There might be other metrics besides kernel time speedup. For instance, the memory footprint of the resulting model.
-While we did not evaluate such metrics, it should not be difficult to optimize for any other such metrics. We believe that the scheduling language is expressive enough to optimize for other metrics.
-
-Q: It's really surprising to me that MLIR doesn't support reductions well -- they seem pretty fundamental to many kernels that other compilers targeting MLIR would see.
+RC Q: It's really surprising to me that MLIR doesn't support reductions well -- they seem pretty fundamental to many kernels that other compilers targeting MLIR would see.
 But this section doesn't really explain if there's anything specific to the decision tree world here, or if this is just a general-purpose "add reductions to MLIR" thing.
@@
 
-Q: Specifically, its speedup diminishes on large batch sizes.
+RE Q: How much do different optimizations contribute to overall performance of code generated by SILVANFORGE?
+A: TODO Mention what we've said about perf improvements due to shared reduction and row caching?
 
+RE Q: There might be other metrics besides kernel time speedup. For instance, the memory footprint of the resulting model.
+While we did not evaluate such metrics, it should not be difficult to optimize for any other such metrics. We believe that the scheduling language is expressive enough to optimize for other metrics.
+
+RD Q: ...speedup diminishes on large batch sizes.
+
+RC Q:  It would be helpful to see results at batch size 1, for example, even if they're not great.
+TODO: Should we say that we would use CPUs for such small batch sizes?
 -------------------------------------------------------------------------

### 2024-07-24 docs/SOSP 2024/rebuttal.txt
Add HB answer to rebuttal

@@
 Q2: Which SilvanForge optimizations are least likely to be supported by PyTorch (in the context of a decision forest compiled by modern Hummingbird), and why?
-* Mention how tensor representations are not a suitable abstraction to express all the transformations that SF can do
-* HB executes one walk per thread and fully walks all trees for a row in a single thread (as far as I can see from the paper). Therefore, SilvanForge will be able to find better schedules. 
-* Expressing tree walks as a set of tensor operations makes it impossible for the compiler to perform optimizations like walk unrolling and walk interleaving.
-Also, this would necessitate several kernel calls (for each step of the tree walk and then subseqently for reductions).
-This representation also means that HB cannot perform optimizations like reductions in shared memory which SilvanForge can.
-* HB is hard-coded to use one kind of representation which is not ideal for all models and target machines. 
-* TODO should we also say how TB cannot match up on the CPU? And focus on how SF improves even on TB with the schedule?
+Hummingbird represents the decision tree inference computation as a PyTorch graph using tensor operations in order to take advantage of sytems like PyTorch and TVM.
+While it is possible to express decision tree inference as tensor operations, the tensor abstraction is insufficient to express optimizations that SilvanForge performs.
+It is very difficult (or even impossible) to express optimizations like tree walk interleaving, tree walk unrolling (especially when trees have different depths), caching trees and rows and reductions in shared memory using tensor operations.
+It is also not possible to express the various configurations of tree and row parallelism and caching that SilvanForge can express with just tensor abstractions.
+Additionally, expressing tree walks as tensor operations hides the semantics of the computation making it impossible for the tensor compiler to automatically perform these optimizations.
+We also expect that the tensor representation will necessitate multiple kernel calls (at least one for the tree walk and then subseqently for the reduction) which will be inefficient compared to the single kernel call that SilvanForge can generate not only because of the kernel call overhead, but also because intermediate results will need to be written to memory.
+The lowering strategy described in the Hummingbird paper is to use one tensor element to represent the current node of each walk.
+This is the SilvanForge schedule where one thread walks one tree and is only one of the many possible schedules that SilvanForge can generate (TODO check this).
+Another aspect where SilvanForge and Hummingbird differ is support for different representations of the model.
+SilvanForge can generate code for different representations of the model (array, sparse, reorg) while Hummingbird only supports a single representation.
+As noted in the paper, different representations can have different performance characteristics and SilvanForge can generate code for the representation that is best suited for the model being compiled.
+On the CPU, it has already been established by prior work that Hummingbird is not as performant as systems like Treebeard.
 

### 2024-07-24 docs/SOSP 2024/rebuttal.txt
pass on first 2 answers

--- 2024-07-24 docs/SOSP 2024/rebuttal.txt
@@
 
-One of our paper's core contributions is identifying that a schedule space exists for decision tree inference, showing that there is a huge variation in performance depending on which schedule is picked and  then designing a scheduling language that cleanly encapsulates this space on both CPUs and GPUs.
-To the best of our knowledge, ours is the first work to study the design of a scheduling language for the domain of decision tree inference.
-While several prior works proposed various techniques to optimize decision tree inference, all of them (RAPIDs, Tahoe, XGBoost etc) use a fixed set of techniques.
-In contrast, SilvanForge’s scheduling language is general enough to represent the specific optimizations that are used in several previous systems and therefore unifies several existing techniques.
+Unlike other domains where scheduling languages are used, SilvanForge targets an irregular application, one that involves traversing multiple decision trees.
+One of the paper's core contributions is in identifying that a schedule space exists for this domain and showing that there is a huge variation in performance across schedules.
+The proposed scheduling language suitably enhances a loop centric scheduling language to target SilvanForge specific operators and perform tree traversal specific optimizations like walk interleaving, walk unrolling, reductions etc.
+These cannot be supported with existing scheduling languages.
+This is reminiscent of how TVM extends Halide to tensor specific primitives like tensorization and cooperative data loading.
+Despite the fact that TVM and Halide operate on similar domains (deep learning vs.
+image processing), such extensions were necessary for TVM to perform optimizations that were not expressible in Halide.
 
-SilvanForge's scheduling language is novel because of the non-obvious realization that a suitably enhaced loop centric scheduling language elegantly encapsulates the optimization space for an irregular application like decision tree inference.
-SilvanForge’s scheduling language naturally adds scheduling primitives that are specific to decision tree inference (to represent walk interleaving, walk unrolling, reductions etc.) which cannot be supported with existing scheduling languages.
-This is reminiscent of how TVM provided tensor specific primitives, namely tensorization and cooperative data loading.
-We believe that the design of SilvanForge's scheduling language is conceptually satisfying because it shows that extending a core set of loop transformation primitives with a few domain specific primitives can capture the optimization space of a complex application like decision tree inference (similar to how TVM extensions to the scheduling language were sufficient to compile tensor based applications).
+We believe that the design of SilvanForge's scheduling language is conceptually satisfying because it shows that by extending a core set of loop transformation primitives with a few domain specific primitives can capture the optimization space of a complex application like decision tree inference.
+Prior work has proposed a small fixed set of optimization techniques and as discussed in the paper, the language is powerful enough to express all these optimizations and more.
 
 Q2: Which SilvanForge optimizations are least likely to be supported by PyTorch (in the context of a decision forest compiled by modern Hummingbird), and why?
-* Mention how tensor representations are not a suitable abstraction to express all the transformations that SF can do
-* HB executes one walk per thread and fully walks all trees for a row in a single thread (as far as I can see from the paper). Therefore, SilvanForge will be able to find better schedules. 
-* Expressing tree walks as a set of tensor operations makes it impossible for the compiler to perform optimizations like walk unrolling and walk interleaving.
-Also, this would necessitate several kernel calls (for each step of the tree walk and then subseqently for reductions).
-This representation also means that HB cannot perform optimizations like reductions in shared memory which SilvanForge can.
-* HB is hard-coded to use one kind of representation which is not ideal for all models and target machines. 
-* TODO should we also say how TB cannot match up on the CPU? And focus on how SF improves even on TB with the schedule?
+Hummingbird represents the decision tree inference computation as a PyTorch graph using tensor operations in order to take advantage of systems like PyTorch and TVM.
+While it is possible to express decision tree inference as tensor operations, the tensor abstraction is insufficient to express rewrites that SilvanForge performs.
+It is very difficult (or even impossible) to express many of our optimizations including tree walk interleaving , tree walk unrolling (especially when trees have different depths), caching trees/rows and reductions in shared memory, using tensor operations.
+ \TODO{THis sounds too dismissive.
+Can we add a line saying what can be done with tensors.
+What would nail this is to show the tensor computation from HB and argue why its not possible to generate code corresponding to interleaved walk (or another optimization for which we show code) }
+This is partially validated by prior experiments on CPUs where a direct comparison between Treebeard (that performs optimizations like unrolling and interleaving) and Hummingbird reveals a massive performance gap.
+It is also not possible to express the various GPU specific configurations that SilvanForge can express, like parallelism on both tree and row, caching and reduction, with just tensor abstractions.
+
+Additionally, representing tree traversals through tensor operations obscures the underlying computation's semantics, rendering it unfeasible for the tensor compiler to perform these optimizations.
+We expect that the tensor representation will necessitate multiple kernel calls (at least one for the tree walk and then subsequently for the reduction) which will be inefficient compared to the single kernel call that SilvanForge can generate not only because of the kernel call overhead, but also because intermediate results will need to be written to memory.
+
+The translation described in the Hummingbird paper uses one tensor element to represent the current node of each walk.
+This corresponds to the SilvanForge schedule where one thread walks one tree, and is only one of the many possible schedules that SilvanForge can generate (TODO check this).
+Another aspect where SilvanForge and Hummingbird differ is support for different representations of the model.
+SilvanForge can generate code for different representations of the model (array, sparse, reorg) while Hummingbird only supports a single representation.
+As noted in the paper, different representations can have different performance characteristics and SilvanForge can tune the code to use the representation that is best suited for the model being compiled.
 

### 2024-07-25 docs/SOSP 2024/rebuttal.txt
pass on first 2 answers

@@
 Unlike other domains where scheduling languages are used, SilvanForge targets an irregular application, one that involves traversing multiple decision trees.
-One of the paper's core contributions is in identifying that a schedule space exists for this domain and showing that there is a huge variation in performance across schedules.
+One of the paper's key contributions lies in recognizing the existence of a scheduling space for a domain as diverse as ours and demonstrating the significant performance variations across different schedules.
 The proposed scheduling language suitably enhances a loop centric scheduling language to target SilvanForge specific operators and perform tree traversal specific optimizations like walk interleaving, walk unrolling, reductions etc.
 These cannot be supported with existing scheduling languages.
-This is reminiscent of how TVM extends Halide to tensor specific primitives like tensorization and cooperative data loading.
+{Can we show an example.}
+This has parallels to how TVM extends Halide to tensor specific primitives like tensorization and cooperative data loading.
 Despite the fact that TVM and Halide operate on similar domains (deep learning vs.
@@
 Hummingbird represents the decision tree inference computation as a PyTorch graph using tensor operations in order to take advantage of systems like PyTorch and TVM.
-While it is possible to express decision tree inference as tensor operations, the tensor abstraction is insufficient to express rewrites that SilvanForge performs.
+While this enables GPU compilation and re-targetability, the tensor abstraction is not sufficient to express rewrites that SilvanForge performs.
 It is very difficult (or even impossible) to express many of our optimizations including tree walk interleaving , tree walk unrolling (especially when trees have different depths), caching trees/rows and reductions in shared memory, using tensor operations.

### 2024-07-25 docs/SOSP 2024/rebuttal.txt
Edit rebuttal. Add answers to other questions

@@
 Q2: Which SilvanForge optimizations are least likely to be supported by PyTorch (in the context of a decision forest compiled by modern Hummingbird), and why?
+* HB only exposes limited parallelism in all its algorithms currently.
+This is a very small subset of the SilvanForge schedule space and as discussed in Section 2 of the paper, this amount of parallelism is likely insufficient to fully utilize the GPU.
+Further, the algorithms rely on gathering individual features as and when required and it is very unlikely that the underlying tensor compiler will decide to read input rows into shared memory as SilvanForge does.
+The analysis of memory access and reuse is complicated due to the indirect accesses of trees and rows.
+SF gets around this with domain specific abstractions and knowledge.
+* We find that in some cases, padding trees is very detrimental to performance (almost 2x in some cases).
+HB currently always pads trees to the same depth.
+SilvanForge can generate code that does not pad trees.
+Additionally, even when trees are padded, HB forces all trees in the model to have the same depth while this is not the case with SilvanForge.
+* Even if one were to expose all forms of parallelism to the tensor compiler, flattening out this parallelism correctly into high-performance kernels is non-trivial and likely beyond the ability of current tensor compilers.
+As concrete examples, it would be nearly impossible (without domain specific knowledge) to restructure the computation to correctly determine the optimal number of threads to use for tree and row parallelism, or to determine whether or not to interleave walks and to what depth.
+
+* Systems challenges 
+    -- no access to lower level details like shared memory, thread synchronization, etc at the tensor abstraction
+    -- kernels are generated dynamically from the tensor graph and there is no scope for controlling how these kernels are optimized at the level of the tensor graph.
+TVM scheduling language is lower level.
+
 Hummingbird represents the decision tree inference computation as a PyTorch graph using tensor operations in order to take advantage of systems like PyTorch and TVM.
@@
 What fraction of your compiler changes are 'portable' across platforms, and how well does the CPU search heuristic work?
-A: The scheduling language, HIR and MIR parts of the compiler are fully shared.
-Most of the backend is shared as well.
-Maybe a good idea to talk about how we share low level code gen between CPU and GPU?
-We found that a search heuristic is not needed on the GPU. Exploring only a few configurations is sufficient to find the best schedule.
+A: The scheduling language, high-level IR and mid-level IR are fully shared.
+As described in the paper, except for a few target specific primitives, the scheduling language is common for both target processors.
+The same scheduling language is used to specify the structure of inference code on the CPU.
+When a schedule doesn't mark any loops as targetting a GPU dimension, CPU code is generated.
+The high-level and mid-level IR and optimizations on them are fully shared between the CPU and GPU compilation pipelines.
+The inference computation is represented as loops in both these IRs and therefore the, the representations and transformations on them are common.
+The compilation pipelines diverge starting starting at the lowering of MIR to low-level IR (as described in the paper).
+Even so, we are able to share significant amount of code between the CPU and GPU lowering due to how in-memory representations are abstracted in SilvanForge (Section 6).
+We found that a search heuristic is not needed on the CPU.
+Exploring only a few configurations is sufficient to find the best schedule.
+For the CPUs we tested on, we found that applying all optimizations is always beneficial and only a few configurations of the loop-nest needs to be explored.
 
 RC Q: Although I wonder if at least some of the explanation is that so many people have moved on from decision tree models to the hot new generative thing.
+A: As we establish in Section 1, decision trees are still widely used in practice and are likely to remain so for the foreseeable future.
+Also, the fact that their popularity has not declined over the last few years is evident from the last few Kaggle surveys.
+We believe that the retargetable performance that SilvanForge provides will be valuable to the large number of practitioners who use decision trees.
 
@@
 Either is probably interesting, but I didn't get enough context from the paper to know the answer.
+A: As mentioned in the paper, the parallel reduction support in MLIR currently is either for value types (scf.reduce) or for tensors (linalg).
+Both of these do not fit the requirements of generating code for decision tree inference.
+With scf.reduce, every accumulation creates a logical copy and eliminating these copies correctly is a non-trivial problem.
+Additionally, MLIR does not have a lowering path from scf.reduce to a GPU implementation.
+With linalg, the entire tensor of values that are to be reduced need to be written to memory first thus making it impossible to fuse the reduction into the inner-most loops.
+
+The reduction support we implement in SilvanForge is fairly general.
+Only the process of identifying reduction loops (Section 5) is specific to decision tree inference (tree loops are identified as reduction loops).
+All lowering of the ops in the reduction dialect are general and can be used for any other application that requires reductions.
+\TODO{Should we say that we can replace the reduction loop identification with a more general mechanism?}
+
+RE Q: There might be other metrics besides kernel time speedup. For instance, the memory footprint of the resulting model.
+A: While we did not evaluate such metrics, it should not be difficult to optimize for them using SilvanForge.
+We believe that the generality of the scheduling language encapsulates a vast enough space and finding variants that are optimized for other metrics should be possible.
+
+-------------------------------------------------------------------------
 
@@
 A: TODO Mention what we've said about perf improvements due to shared reduction and row caching?
-
-RE Q: There might be other metrics besides kernel time speedup. For instance, the memory footprint of the resulting model.
-While we did not evaluate such metrics, it should not be difficult to optimize for any other such metrics. We believe that the scheduling language is expressive enough to optimize for other metrics.
-
 RD Q: ...speedup diminishes on large batch sizes.
@@
 TODO: Should we say that we would use CPUs for such small batch sizes?
--------------------------------------------------------------------------
 

### 2024-07-25 docs/SOSP 2024/rebuttal.txt
Edits to Q1 rebuttal

@@
 
-We believe that the design of SilvanForge's scheduling language is conceptually satisfying because it shows that by extending a core set of loop transformation primitives with a few domain specific primitives can capture the optimization space of a complex application like decision tree inference.
-Prior work has proposed a small fixed set of optimization techniques and as discussed in the paper, the language is powerful enough to express all these optimizations and more.
+We believe that the design of SilvanForge's scheduling language is conceptually satisfying because it shows that adding the required domain-specific primitives to a core set of loop transformation primitives can capture the optimization space of a complex application like decision tree inference.
+Prior work has proposed a small fixed set of optimization techniques and as discussed in the paper, SilvanForge’s scheduling language is powerful enough to express all these optimizations and more.
 

### 2024-07-25 docs/SOSP 2024/rebuttal.txt
Some edits to HB question

@@
 Q2: Which SilvanForge optimizations are least likely to be supported by PyTorch (in the context of a decision forest compiled by modern Hummingbird), and why?
-* HB only exposes limited parallelism in all its algorithms currently.
-This is a very small subset of the SilvanForge schedule space and as discussed in Section 2 of the paper, this amount of parallelism is likely insufficient to fully utilize the GPU.
-Further, the algorithms rely on gathering individual features as and when required and it is very unlikely that the underlying tensor compiler will decide to read input rows into shared memory as SilvanForge does.
+A: 
+Hummingbird represents the decision tree inference computation as a PyTorch graph using tensor operations in order to take advantage of systems like PyTorch and TVM.
+While this enables GPU compilation and re-targetability, the tensor abstraction is not sufficient to express rewrites that SilvanForge performs.
+Hummingbird exposes all forms of parallelism to the tensor compiler.
+The generated code performs all tree walks in parallel, stores the results of individual tree walks in a tensor and finally performs a reduction.
+Flattening out this parallelism correctly into high-performance kernels is non-trivial and likely beyond the ability of current tensor compilers.
+While Hummingbird does perform a limited form of tree-walk unrolling (it pads all trees to the same depth and unrolls walks), it is very difficult to express other optimizations including tree walk interleaving, caching trees/rows and reductions in shared memory, using tensor operations.
+
+Representing tree traversals through tensor operations obscures the underlying computation's semantics, rendering it unfeasible for the tensor compiler to perform these optimizations.
+We expect that the tensor representation will necessitate multiple kernel calls (at least one for the tree walk and then subsequently for the reduction) which will be inefficient compared to the single kernel call that SilvanForge can generate not only because of the kernel call overhead, but also because intermediate results will need to be written to memory.
+
+Additionally, the generated tensor code relies on gathering individual features when required.
+It is very unlikely that the underlying tensor compiler will decide to read input rows into shared memory as SilvanForge does.
 The analysis of memory access and reuse is complicated due to the indirect accesses of trees and rows.
-SF gets around this with domain specific abstractions and knowledge.
+SilvanForge gets around this with domain specific abstractions and knowledge.
+
+* As concrete examples, it would be nearly impossible (without domain specific knowledge) to restructure the computation to correctly determine the optimal number of threads to use for tree and row parallelism, or to determine whether or not to interleave walks and to what depth.
+
 * We find that in some cases, padding trees is very detrimental to performance (almost 2x in some cases).
@@
 Additionally, even when trees are padded, HB forces all trees in the model to have the same depth while this is not the case with SilvanForge.
-* Even if one were to expose all forms of parallelism to the tensor compiler, flattening out this parallelism correctly into high-performance kernels is non-trivial and likely beyond the ability of current tensor compilers.
-As concrete examples, it would be nearly impossible (without domain specific knowledge) to restructure the computation to correctly determine the optimal number of threads to use for tree and row parallelism, or to determine whether or not to interleave walks and to what depth.
 
@@
 
-Hummingbird represents the decision tree inference computation as a PyTorch graph using tensor operations in order to take advantage of systems like PyTorch and TVM.
-While this enables GPU compilation and re-targetability, the tensor abstraction is not sufficient to express rewrites that SilvanForge performs.
-It is very difficult (or even impossible) to express many of our optimizations including tree walk interleaving , tree walk unrolling (especially when trees have different depths), caching trees/rows and reductions in shared memory, using tensor operations.
+I
  \TODO{THis sounds too dismissive.
@@
 
-Additionally, representing tree traversals through tensor operations obscures the underlying computation's semantics, rendering it unfeasible for the tensor compiler to perform these optimizations.
-We expect that the tensor representation will necessitate multiple kernel calls (at least one for the tree walk and then subsequently for the reduction) which will be inefficient compared to the single kernel call that SilvanForge can generate not only because of the kernel call overhead, but also because intermediate results will need to be written to memory.
+Additionally,
 

### 2024-07-25 docs/SOSP 2024/rebuttal.txt
pass on q2

@@
 A: 
-Hummingbird represents the decision tree inference computation as a PyTorch graph using tensor operations in order to take advantage of systems like PyTorch and TVM.
-While this enables GPU compilation and re-targetability, the tensor abstraction is not sufficient to express rewrites that SilvanForge performs.
-Hummingbird exposes all forms of parallelism to the tensor compiler.
+Hummingbird utilizes PyTorch by translating decision tree inference computation to a PyTorch graph.
+While this enables GPU compilation and re-targetability, as we discuss below the tensor abstraction is not sufficient to express rewrites that SilvanForge performs.
+The Hummingbird translation exposes all forms of parallelism to the tensor compiler.
 The generated code performs all tree walks in parallel, stores the results of individual tree walks in a tensor and finally performs a reduction.
-Flattening out this parallelism correctly into high-performance kernels is non-trivial and likely beyond the ability of current tensor compilers.
-While Hummingbird does perform a limited form of tree-walk unrolling (it pads all trees to the same depth and unrolls walks), it is very difficult to express other optimizations including tree walk interleaving, caching trees/rows and reductions in shared memory, using tensor operations.
+In the process Hummingbird performs a limited form of tree-walk unrolling (it pads all trees to the same depth and unrolls walks).
+This as discussed in our paper is not always useful (say when?).
+It is very difficult to express other optimizations including tree walk interleaving, caching trees/rows and reductions in shared memory, using tensor operations.
 
-Representing tree traversals through tensor operations obscures the underlying computation's semantics, rendering it unfeasible for the tensor compiler to perform these optimizations.
-We expect that the tensor representation will necessitate multiple kernel calls (at least one for the tree walk and then subsequently for the reduction) which will be inefficient compared to the single kernel call that SilvanForge can generate not only because of the kernel call overhead, but also because intermediate results will need to be written to memory.
+Further, representing tree traversals through tensor operations obscures the underlying computation's semantics, rendering it unfeasible for the tensor compiler to perform these optimizations either.
+Flattening out the different sources of parallelism correctly into high-performance kernels is non-trivial and likely beyond the ability of current tensor compilers (this line is not convincing).
+Note that the generated tensor code relies on gathering individual features when required.
+With this representation the analysis of memory access and reuse is complicated due to the indirect accesses of trees and rows.
+This makes it impossible for the tensor compiler to read input rows into shared memory as SilvanForge does.
+FInally, we expect that the tensor representation will necessitate multiple kernel calls (at least one for the tree walk and then subsequently for the reduction) which will be inefficient compared to the single kernel call that SilvanForge can generate not only because of the kernel call overhead, but also because intermediate results will need to be written to memory.
 
-Additionally, the generated tensor code relies on gathering individual features when required.
-It is very unlikely that the underlying tensor compiler will decide to read input rows into shared memory as SilvanForge does.
-The analysis of memory access and reuse is complicated due to the indirect accesses of trees and rows.
-SilvanForge gets around this with domain specific abstractions and knowledge.
+SilvanForge gets around all these with domain specific abstractions and knowledge.
 

### 2024-07-26 docs/SOSP 2024/rebuttal.txt
Edits to Q1 and Q2

@@
 {Can we show an example.}
-This has parallels to how TVM extends Halide to tensor specific primitives like tensorization and cooperative data loading.
+This has parallels to how TVM extends Halide with DNN compilation specific primitives like tensorization and cooperative data loading.
 Despite the fact that TVM and Halide operate on similar domains (deep learning vs.
@@
 
-Further, representing tree traversals through tensor operations obscures the underlying computation's semantics, rendering it unfeasible for the tensor compiler to perform these optimizations either.
-Flattening out the different sources of parallelism correctly into high-performance kernels is non-trivial and likely beyond the ability of current tensor compilers (this line is not convincing).
-Note that the generated tensor code relies on gathering individual features when required.
+Representing tree traversals through tensor operations obscures the underlying computation's semantics, rendering it unfeasible for the tensor compiler to perform these optimizations either.
+More specifically, the generated tensor code relies on gathering individual features when required.
 With this representation the analysis of memory access and reuse is complicated due to the indirect accesses of trees and rows.
 This makes it impossible for the tensor compiler to read input rows into shared memory as SilvanForge does.
-FInally, we expect that the tensor representation will necessitate multiple kernel calls (at least one for the tree walk and then subsequently for the reduction) which will be inefficient compared to the single kernel call that SilvanForge can generate not only because of the kernel call overhead, but also because intermediate results will need to be written to memory.
+Second, to the best of our knowledge, tensor compilers cannot perform interleaving as SilvanForge does (this is not part of their schedule space).
+Third, we expect that the tensor representation will necessitate multiple kernel calls (at least one for the tree walk and then subsequently for the reduction) which will be inefficient compared to the single kernel call that SilvanForge can generate not only because of the kernel call overhead, but also because intermediate results will need to be written to memory (and not to shared memory as in the case of SilvanForge).
+Finally, it would be nearly impossible to restructure the computation to correctly determine the optimal number of threads to use for tree and row parallelism since no template schedules can be specified for the kernels finally generated by the tensor compiler.
+SilvanForge addresses these problems with domain specific abstractions and knowledge.
 
-SilvanForge gets around all these with domain specific abstractions and knowledge.
+The above discussion brings up a more practical system level problem ... 
+* Systems challenges 
+    -- no access to lower level details like shared memory, thread synchronization, etc at the tensor abstraction
+    -- kernels are generated dynamically from the tensor graph and there is no scope for controlling how these kernels are optimized at the level of the tensor graph.
+TVM scheduling language is lower level.
 
-* As concrete examples, it would be nearly impossible (without domain specific knowledge) to restructure the computation to correctly determine the optimal number of threads to use for tree and row parallelism, or to determine whether or not to interleave walks and to what depth.
+Another aspect where SilvanForge and Hummingbird differ is support for different representations of the model.
+SilvanForge can generate code for different representations of the model (array, sparse, reorg) while Hummingbird only supports a single representation.
+As noted in the paper, different representations can have different performance characteristics and SilvanForge can tune the code to use the representation that is best suited for the model being compiled.
+
+These views are partially validated by prior experiments on CPUs where a direct comparison between Treebeard (that performs optimizations like unrolling and interleaving) and Hummingbird reveals a massive performance gap.
+It is also not possible to express the various GPU specific configurations that SilvanForge can express, like parallelism on both tree and row, caching and reduction, with just tensor abstractions.
 
@@
 Additionally, even when trees are padded, HB forces all trees in the model to have the same depth while this is not the case with SilvanForge.
-
-* Systems challenges 
-    -- no access to lower level details like shared memory, thread synchronization, etc at the tensor abstraction
-    -- kernels are generated dynamically from the tensor graph and there is no scope for controlling how these kernels are optimized at the level of the tensor graph.
-TVM scheduling language is lower level.
-
-I
- \TODO{THis sounds too dismissive.
-Can we add a line saying what can be done with tensors.
-What would nail this is to show the tensor computation from HB and argue why its not possible to generate code corresponding to interleaved walk (or another optimization for which we show code) }
-This is partially validated by prior experiments on CPUs where a direct comparison between Treebeard (that performs optimizations like unrolling and interleaving) and Hummingbird reveals a massive performance gap.
-It is also not possible to express the various GPU specific configurations that SilvanForge can express, like parallelism on both tree and row, caching and reduction, with just tensor abstractions.
-
-Additionally,
-
-The translation described in the Hummingbird paper uses one tensor element to represent the current node of each walk.
-This corresponds to the SilvanForge schedule where one thread walks one tree, and is only one of the many possible schedules that SilvanForge can generate (TODO check this).
-Another aspect where SilvanForge and Hummingbird differ is support for different representations of the model.
-SilvanForge can generate code for different representations of the model (array, sparse, reorg) while Hummingbird only supports a single representation.
-As noted in the paper, different representations can have different performance characteristics and SilvanForge can tune the code to use the representation that is best suited for the model being compiled.
 

### 2024-07-26 docs/SOSP 2024/rebuttal.txt
Edit answer to Q2 and the questions.

@@
-Q1: How does SilvanForge’s decision-forest-based ‘schedule space’ and scheduling language differ from prior languages for other domains, e.g., TVM’s?
-For instance, what operators are new to SilvanForge?
+Q1: Novelty of scheduling language.
 
@@
 These cannot be supported with existing scheduling languages.
-{Can we show an example.}
-This has parallels to how TVM extends Halide with DNN compilation specific primitives like tensorization and cooperative data loading.
+{Can we show an example.} This has parallels to how TVM extends Halide with domain-specific primitives for DNN compilation like tensorization and cooperative data loading.
 Despite the fact that TVM and Halide operate on similar domains (deep learning vs.
@@
 
-Q2: Which SilvanForge optimizations are least likely to be supported by PyTorch (in the context of a decision forest compiled by modern Hummingbird), and why?
-A: 
+Q2: Comparison with Hummingbird
+
 Hummingbird utilizes PyTorch by translating decision tree inference computation to a PyTorch graph.
@@
 
-The above discussion brings up a more practical system level problem ... 
-* Systems challenges 
-    -- no access to lower level details like shared memory, thread synchronization, etc at the tensor abstraction
-    -- kernels are generated dynamically from the tensor graph and there is no scope for controlling how these kernels are optimized at the level of the tensor graph.
-TVM scheduling language is lower level.
+The above discussion brings up a more practical system level problem.
+Expressing the tree inference problem as a PyTorch graph means that the decision tree compiler has no access to lower level details like shared memory, thread synchronization, etc.
+Kernels are generated dynamically from the tensor graph and there is no scope for controlling how these kernels are optimized at the level of the tensor graph.
+For example, there is no way to specify a TVM schedule on the tensor graph.
 
@@
 
-* We find that in some cases, padding trees is very detrimental to performance (almost 2x in some cases).
-HB currently always pads trees to the same depth.
-SilvanForge can generate code that does not pad trees.
-Additionally, even when trees are padded, HB forces all trees in the model to have the same depth while this is not the case with SilvanForge.
-
-RB Q: You claim that SilvanForge decouples schedule from the backend architecture  but it seems both of these would change substantially with CPU?
-What fraction of your compiler changes are 'portable' across platforms, and how well does the CPU search heuristic work?
-A: The scheduling language, high-level IR and mid-level IR are fully shared.
+Rev#B: Are your compiler changes 'portable' across platforms?
+The scheduling language, high-level IR and mid-level IR are fully shared.
 As described in the paper, except for a few target specific primitives, the scheduling language is common for both target processors.
@@
 
-RC Q: Although I wonder if at least some of the explanation is that so many people have moved on from decision tree models to the hot new generative thing.
-A: As we establish in Section 1, decision trees are still widely used in practice and are likely to remain so for the foreseeable future.
+Rev#C: Have people moved on from decision tree models?
+As we establish in Section 1, decision trees are still widely used in practice and are likely to remain so for the foreseeable future.
 Also, the fact that their popularity has not declined over the last few years is evident from the last few Kaggle surveys.
@@
 
-RC Q: It's really surprising to me that MLIR doesn't support reductions well -- they seem pretty fundamental to many kernels that other compilers targeting MLIR would see.
-But this section doesn't really explain if there's anything specific to the decision tree world here, or if this is just a general-purpose "add reductions to MLIR" thing.
-Either is probably interesting, but I didn't get enough context from the paper to know the answer.
-A: As mentioned in the paper, the parallel reduction support in MLIR currently is either for value types (scf.reduce) or for tensors (linalg).
+Rev#C: Support for reductions.
+As mentioned in the paper, the parallel reduction support in MLIR currently is either for value types (scf.reduce) or for tensors (linalg).
 Both of these do not fit the requirements of generating code for decision tree inference.
@@
 
-RE Q: There might be other metrics besides kernel time speedup. For instance, the memory footprint of the resulting model.
-A: While we did not evaluate such metrics, it should not be difficult to optimize for them using SilvanForge.
+Rev#E: Other metrics besides kernel time speedup. 
+While we did not evaluate such metrics, it should not be difficult to optimize for them using SilvanForge.
 We believe that the generality of the scheduling language encapsulates a vast enough space and finding variants that are optimized for other metrics should be possible.

### 2024-07-26 docs/SOSP 2024/rebuttal.txt
More edits to Q2 answer

@@
 These cannot be supported with existing scheduling languages.
-{Can we show an example.} This has parallels to how TVM extends Halide with domain-specific primitives for DNN compilation like tensorization and cooperative data loading.
+This has parallels to how TVM extends Halide with domain-specific primitives for DNN compilation like tensorization and cooperative data loading.
 Despite the fact that TVM and Halide operate on similar domains (deep learning vs.
@@
 In the process Hummingbird performs a limited form of tree-walk unrolling (it pads all trees to the same depth and unrolls walks).
-This as discussed in our paper is not always useful (say when?).
+This as discussed in our paper is not always useful (Section 2).
 It is very difficult to express other optimizations including tree walk interleaving, caching trees/rows and reductions in shared memory, using tensor operations.
@@
 Representing tree traversals through tensor operations obscures the underlying computation's semantics, rendering it unfeasible for the tensor compiler to perform these optimizations either.
-More specifically, the generated tensor code relies on gathering individual features when required.
+Consider the example of walk-interleaving.
+SilvanForge’s tree-walk interleaving relies on unrolling only the trees whose walks are being interleaved to the same depth.
+It also reorders trees so that trees with the same depth are processed together.
+While it maybe possible to extend tensor compilers to perform interleaving, it will be very hard to mimic what SilvanForge does.
+Current tensor compilers optimize dense matrix computation and do not have the optimizations necessary to interleave instructions across iterations.
+Even if they did, picking the right tree-walks (same tree different rows, different trees same row etc.) to interleave would be a non-trivial problem.
+
+The tensor compiler cannot perform the required caching and reduction optimizations either.
+The generated tensor code relies on gathering individual features when required.
 With this representation the analysis of memory access and reuse is complicated due to the indirect accesses of trees and rows.
 This makes it impossible for the tensor compiler to read input rows into shared memory as SilvanForge does.
-Second, to the best of our knowledge, tensor compilers cannot perform interleaving as SilvanForge does (this is not part of their schedule space).
-Third, we expect that the tensor representation will necessitate multiple kernel calls (at least one for the tree walk and then subsequently for the reduction) which will be inefficient compared to the single kernel call that SilvanForge can generate not only because of the kernel call overhead, but also because intermediate results will need to be written to memory (and not to shared memory as in the case of SilvanForge).
-Finally, it would be nearly impossible to restructure the computation to correctly determine the optimal number of threads to use for tree and row parallelism since no template schedules can be specified for the kernels finally generated by the tensor compiler.
-SilvanForge addresses these problems with domain specific abstractions and knowledge.
+Additionally, we expect that the tensor representation will necessitate multiple kernel calls (at least one for the tree walk and then subsequently for the reduction) which will be inefficient compared to the single kernel call that SilvanForge can generate.
 
@@
 
-These views are partially validated by prior experiments on CPUs where a direct comparison between Treebeard (that performs optimizations like unrolling and interleaving) and Hummingbird reveals a massive performance gap.
-It is also not possible to express the various GPU specific configurations that SilvanForge can express, like parallelism on both tree and row, caching and reduction, with just tensor abstractions.
+These views are partially validated by experiments in previous literature where, on CPUs, a direct comparison between Treebeard (that performs optimizations like unrolling and interleaving) and Hummingbird reveals a massive performance gap.
 
@@
 -------------------------------------------------------------------------
+
+ Finally, it would be nearly impossible to restructure the computation to correctly determine the optimal number of threads to use for tree and row parallelism since no template schedules can be specified for the kernels finally generated by the tensor compiler.
+SilvanForge addresses these problems with domain specific abstractions and knowledge.
 

### 2024-07-26 docs/SOSP 2024/rebuttal.txt
Minor edits

@@
 One of the paper's key contributions lies in recognizing the existence of a scheduling space for a domain as diverse as ours and demonstrating the significant performance variations across different schedules.
-The proposed scheduling language suitably enhances a loop centric scheduling language to target SilvanForge specific operators and perform tree traversal specific optimizations like walk interleaving, walk unrolling, reductions etc.
+The proposed scheduling language suitably enhances a loop centric scheduling language to target SilvanForge specific operators and perform tree traversal specific optimizations like tree-walk interleaving, tree-walk unrolling, reductions etc.
 These cannot be supported with existing scheduling languages.
@@
 We believe that the design of SilvanForge's scheduling language is conceptually satisfying because it shows that adding the required domain-specific primitives to a core set of loop transformation primitives can capture the optimization space of a complex application like decision tree inference.
-Prior work has proposed a small fixed set of optimization techniques and as discussed in the paper, SilvanForge’s scheduling language is powerful enough to express all these optimizations and more.
+Prior work has proposed a small fixed set of optimization techniques, and as discussed in the paper, SilvanForge’s scheduling language is powerful enough to express all these optimizations and more.
 
@@
 
-Hummingbird utilizes PyTorch by translating decision tree inference computation to a PyTorch graph.
-While this enables GPU compilation and re-targetability, as we discuss below the tensor abstraction is not sufficient to express rewrites that SilvanForge performs.
-The Hummingbird translation exposes all forms of parallelism to the tensor compiler.
-The generated code performs all tree walks in parallel, stores the results of individual tree walks in a tensor and finally performs a reduction.
-In the process Hummingbird performs a limited form of tree-walk unrolling (it pads all trees to the same depth and unrolls walks).
-This as discussed in our paper is not always useful (Section 2).
+Hummingbird translates the decision tree inference into a series of tensor operations to use PyTorch.
+While this enables re-targetability, the tensor abstraction is insufficient to express all SilvanForge optimizations.
+Hummingbird exposes all forms of parallelism to the tensor compiler.
+The generated tensor code performs all tree walks in parallel, stores the individual tree predictions in a tensor and finally performs a reduction.
+In the process, Hummingbird performs a limited form of tree-walk unrolling (it pads all trees to the same depth and unrolls walks).
+This, as discussed in our paper, is not always useful (Section 2).
 It is very difficult to express other optimizations including tree walk interleaving, caching trees/rows and reductions in shared memory, using tensor operations.
 
-Representing tree traversals through tensor operations obscures the underlying computation's semantics, rendering it unfeasible for the tensor compiler to perform these optimizations either.
+Representing tree traversals through tensor operations obscures the underlying computation's semantics, rendering it infeasible for the tensor compiler to perform these optimizations.
 Consider the example of walk-interleaving.
@@
 The tensor compiler cannot perform the required caching and reduction optimizations either.
-The generated tensor code relies on gathering individual features when required.
-With this representation the analysis of memory access and reuse is complicated due to the indirect accesses of trees and rows.
+Since the generated tensor code relies on gathering individual features when required, the indirect accesses of trees and rows complicates the analysis of memory access and reuse.
 This makes it impossible for the tensor compiler to read input rows into shared memory as SilvanForge does.
-Additionally, we expect that the tensor representation will necessitate multiple kernel calls (at least one for the tree walk and then subsequently for the reduction) which will be inefficient compared to the single kernel call that SilvanForge can generate.
+Additionally, we expect that the tensor representation will necessitate multiple kernel calls (at least one for the tree walk and then subsequently for the reduction) which will be inefficient compared to the single kernel call that SilvanForge generates.
 
 The above discussion brings up a more practical system level problem.
-Expressing the tree inference problem as a PyTorch graph means that the decision tree compiler has no access to lower level details like shared memory, thread synchronization, etc.
-Kernels are generated dynamically from the tensor graph and there is no scope for controlling how these kernels are optimized at the level of the tensor graph.
-For example, there is no way to specify a TVM schedule on the tensor graph.
+Expressing the tree inference using PyTorch means that the decision tree compiler has no control over lower level details like shared memory, thread synchronization, etc.
+Kernels are generated dynamically and there is no scope for controlling how these kernels are optimized.
+Systems like TorchDynamo and TorchInductor do not provide this level of control to the PyTorch program.
 
-Another aspect where SilvanForge and Hummingbird differ is support for different representations of the model.
-SilvanForge can generate code for different representations of the model (array, sparse, reorg) while Hummingbird only supports a single representation.
-As noted in the paper, different representations can have different performance characteristics and SilvanForge can tune the code to use the representation that is best suited for the model being compiled.
-
-These views are partially validated by experiments in previous literature where, on CPUs, a direct comparison between Treebeard (that performs optimizations like unrolling and interleaving) and Hummingbird reveals a massive performance gap.
+These views are partially validated by experiments in previous literature where, on CPUs, a direct comparison between Treebeard (that performs optimizations like unrolling and interleaving) and Hummingbird reveals an order-of-magnitude performance gap.
 
@@
 The scheduling language, high-level IR and mid-level IR are fully shared.
-As described in the paper, except for a few target specific primitives, the scheduling language is common for both target processors.
-The same scheduling language is used to specify the structure of inference code on the CPU.
+The same scheduling language is used to specify the structure of inference code on CPUs and GPUs.
 When a schedule doesn't mark any loops as targetting a GPU dimension, CPU code is generated.
@@
 The inference computation is represented as loops in both these IRs and therefore the, the representations and transformations on them are common.
-The compilation pipelines diverge starting starting at the lowering of MIR to low-level IR (as described in the paper).
+The compilation pipelines diverge starting starting at the lowering of MIR to low-level IR.
 Even so, we are able to share significant amount of code between the CPU and GPU lowering due to how in-memory representations are abstracted in SilvanForge (Section 6).
@@
 -------------------------------------------------------------------------
+
+Another aspect where SilvanForge and Hummingbird differ is support for different representations of the model.
+SilvanForge can generate code for different representations of the model (array, sparse, reorg) while Hummingbird only supports a single representation.
+As noted in the paper, different representations can have different performance characteristics and SilvanForge can tune the code to use the representation that is best suited for the model being compiled.
 

### 2024-07-26 docs/SOSP 2024/rebuttal.txt
Edit later answers

@@
 Even so, we are able to share significant amount of code between the CPU and GPU lowering due to how in-memory representations are abstracted in SilvanForge (Section 6).
-We found that a search heuristic is not needed on the CPU.
-Exploring only a few configurations is sufficient to find the best schedule.
-For the CPUs we tested on, we found that applying all optimizations is always beneficial and only a few configurations of the loop-nest needs to be explored.
+For CPUs, exploring only a few configurations is sufficient to find the best schedule as most optimizations are almost always beneficial.
+Therefore, a heuristic is not needed.
 
 Rev#C: Have people moved on from decision tree models?
-As we establish in Section 1, decision trees are still widely used in practice and are likely to remain so for the foreseeable future.
+As we discuss in the paper, decision trees are widely used in practice.
 Also, the fact that their popularity has not declined over the last few years is evident from the last few Kaggle surveys.
@@
 Rev#C: Support for reductions.
-As mentioned in the paper, the parallel reduction support in MLIR currently is either for value types (scf.reduce) or for tensors (linalg).
-Both of these do not fit the requirements of generating code for decision tree inference.
+MLIR currently supports parallel reductions for value types (scf.reduce) and reducing a dimension of a tensor (linalg.reduce).
+Neither of these meets the requirements for generating code for decision tree inference.
 With scf.reduce, every accumulation creates a logical copy and eliminating these copies correctly is a non-trivial problem.
-Additionally, MLIR does not have a lowering path from scf.reduce to a GPU implementation.
-With linalg, the entire tensor of values that are to be reduced need to be written to memory first thus making it impossible to fuse the reduction into the inner-most loops.
+With linalg.reduce, the reduction cannot be fused with another loop as the entire tensor of values that are to be reduced need to be written to memory first.
 
@@
 All lowering of the ops in the reduction dialect are general and can be used for any other application that requires reductions.
-\TODO{Should we say that we can replace the reduction loop identification with a more general mechanism?}
-
-Rev#E: Other metrics besides kernel time speedup. 
-While we did not evaluate such metrics, it should not be difficult to optimize for them using SilvanForge.
-We believe that the generality of the scheduling language encapsulates a vast enough space and finding variants that are optimized for other metrics should be possible.
-
 -------------------------------------------------------------------------
@@
 SilvanForge addresses these problems with domain specific abstractions and knowledge.
+
+Rev#E: Other metrics besides kernel time speedup. 
+While we did not evaluate such metrics, it should not be difficult to optimize for them using SilvanForge.
+We believe that the generality of the scheduling language encapsulates a vast enough space and finding variants that are optimized for other metrics should be possible.
 

### 2024-07-26 docs/SOSP 2024/rebuttal.txt
Minor edits

@@
 Unlike other domains where scheduling languages are used, SilvanForge targets an irregular application, one that involves traversing multiple decision trees.
-One of the paper's key contributions lies in recognizing the existence of a scheduling space for a domain as diverse as ours and demonstrating the significant performance variations across different schedules.
-The proposed scheduling language suitably enhances a loop centric scheduling language to target SilvanForge specific operators and perform tree traversal specific optimizations like tree-walk interleaving, tree-walk unrolling, reductions etc.
+One of the paper's key contributions lies in recognizing the existence of a scheduling space for this domain and demonstrating the significant performance variations across different schedules.
+The proposed scheduling language suitably enhances a loop centric scheduling language to target SilvanForge specific operators and perform tree traversal specific optimizations like tree-walk interleaving, tree-walk unrolling, reduction optimizations etc.
 These cannot be supported with existing scheduling languages.
@@
 
-Hummingbird translates the decision tree inference into a series of tensor operations to use PyTorch.
+Hummingbird translates decision tree inference into a series of tensor operations to use PyTorch.
 While this enables re-targetability, the tensor abstraction is insufficient to express all SilvanForge optimizations.
-Hummingbird exposes all forms of parallelism to the tensor compiler.
 The generated tensor code performs all tree walks in parallel, stores the individual tree predictions in a tensor and finally performs a reduction.
-In the process, Hummingbird performs a limited form of tree-walk unrolling (it pads all trees to the same depth and unrolls walks).
+Hummingbird performs a limited form of tree-walk unrolling (it pads all trees to the same depth and unrolls walks).
 This, as discussed in our paper, is not always useful (Section 2).
-It is very difficult to express other optimizations including tree walk interleaving, caching trees/rows and reductions in shared memory, using tensor operations.
+However.
+it is very difficult to express other optimizations including tree walk interleaving, caching trees/rows and reductions in shared memory, using tensor operations.
 
-Representing tree traversals through tensor operations obscures the underlying computation's semantics, rendering it infeasible for the tensor compiler to perform these optimizations.
+Using tensor operations to represent tree traversals obscures the computation's semantics, making it difficult for the tensor compiler to perform these optimizations.
 Consider the example of walk-interleaving.
@@
 The tensor compiler cannot perform the required caching and reduction optimizations either.
-Since the generated tensor code relies on gathering individual features when required, the indirect accesses of trees and rows complicates the analysis of memory access and reuse.
+Since the generated tensor code relies on gathering individual features when required, the indirect accesses to trees and rows complicates the analysis of memory access and reuse.
 This makes it impossible for the tensor compiler to read input rows into shared memory as SilvanForge does.
-Additionally, we expect that the tensor representation will necessitate multiple kernel calls (at least one for the tree walk and then subsequently for the reduction) which will be inefficient compared to the single kernel call that SilvanForge generates.
+Additionally, we expect that the tensor representation will necessitate multiple kernel calls (at least one for the tree walk and one for the reduction) which will be inefficient compared to the single kernel call that SilvanForge generates.
 

### 2024-07-26 docs/SOSP 2024/rebuttal.txt
Minor edits and shortening

@@
 
-Using tensor operations to represent tree traversals obscures the computation's semantics, making it difficult for the tensor compiler to perform these optimizations.
+Further, using tensor operations to represent tree traversals obscures the computation's semantics, making it difficult for the tensor compiler to perform these optimizations.
 Consider the example of walk-interleaving.
@@
 Current tensor compilers optimize dense matrix computation and do not have the optimizations necessary to interleave instructions across iterations.
-Even if they did, picking the right tree-walks (same tree different rows, different trees same row etc.) to interleave would be a non-trivial problem.
+Even if this support is added, picking the right tree-walks (same tree different rows, different trees same row etc.) to interleave would be a non-trivial problem.
 
@@
 Rev#B: Are your compiler changes 'portable' across platforms?
-The scheduling language, high-level IR and mid-level IR are fully shared.
+The scheduling language, high-level IR and mid-level IR and optimizations on them are fully shared.
 The same scheduling language is used to specify the structure of inference code on CPUs and GPUs.
 When a schedule doesn't mark any loops as targetting a GPU dimension, CPU code is generated.
-The high-level and mid-level IR and optimizations on them are fully shared between the CPU and GPU compilation pipelines.
-The inference computation is represented as loops in both these IRs and therefore the, the representations and transformations on them are common.
 The compilation pipelines diverge starting starting at the lowering of MIR to low-level IR.
-Even so, we are able to share significant amount of code between the CPU and GPU lowering due to how in-memory representations are abstracted in SilvanForge (Section 6).
-For CPUs, exploring only a few configurations is sufficient to find the best schedule as most optimizations are almost always beneficial.
-Therefore, a heuristic is not needed.
-
-Rev#C: Have people moved on from decision tree models?
-As we discuss in the paper, decision trees are widely used in practice.
-Also, the fact that their popularity has not declined over the last few years is evident from the last few Kaggle surveys.
-We believe that the retargetable performance that SilvanForge provides will be valuable to the large number of practitioners who use decision trees.
+Even so, we are able to share significant amount of code between the CPU and GPU lowering because of how in-memory representations are abstracted (Section 6).
 
@@
 With linalg.reduce, the reduction cannot be fused with another loop as the entire tensor of values that are to be reduced need to be written to memory first.
+The reduction support we implement in SilvanForge is fairly general.
+Only the process of identifying reduction loops is currently specific to decision tree inference.
+All lowering of the ops in the reduction dialect are general.
 
-The reduction support we implement in SilvanForge is fairly general.
-Only the process of identifying reduction loops (Section 5) is specific to decision tree inference (tree loops are identified as reduction loops).
-All lowering of the ops in the reduction dialect are general and can be used for any other application that requires reductions.
 -------------------------------------------------------------------------
@@
 SilvanForge addresses these problems with domain specific abstractions and knowledge.
+
+Rev#C: Have people moved on from decision tree models?
+As we discuss in the paper, decision trees are widely used in practice.
+Also, the fact that their popularity has not declined over the last few years is evident from the last few Kaggle surveys.
+We believe that the retargetable performance that SilvanForge provides will be valuable to the large number of practitioners who use decision trees.
 

### 2024-07-26 docs/SOSP 2024/rebuttal.txt
Minor fixes

@@
 
-We believe that the design of SilvanForge's scheduling language is conceptually satisfying because it shows that adding the required domain-specific primitives to a core set of loop transformation primitives can capture the optimization space of a complex application like decision tree inference.
-Prior work has proposed a small fixed set of optimization techniques, and as discussed in the paper, SilvanForge’s scheduling language is powerful enough to express all these optimizations and more.
+We believe that the design of SilvanForge's scheduling language is conceptually satisfying because it shows that adding a few required domain-specific primitives to a core set of loop transformation primitives can capture the optimization space of a complex application like decision tree inference.
+Prior work (e.g.
+Tahoe, RAPIDs) has proposed a small fixed set of optimization techniques, and as discussed in the paper, SilvanForge’s scheduling language is powerful enough to express all these optimizations and more.
 
@@
 However.
-it is very difficult to express other optimizations including tree walk interleaving, caching trees/rows and reductions in shared memory, using tensor operations.
+it is very difficult to express other SilvanForge optimizations including tree walk interleaving, caching trees/rows and reductions in shared memory, using tensor operations.
 
@@
 When a schedule doesn't mark any loops as targetting a GPU dimension, CPU code is generated.
-The compilation pipelines diverge starting starting at the lowering of MIR to low-level IR.
+The compilation pipelines diverge starting at the lowering of MIR to low-level IR.
 Even so, we are able to share significant amount of code between the CPU and GPU lowering because of how in-memory representations are abstracted (Section 6).

### 2024-07-26 docs/SOSP 2024/rebuttal.txt
Add line on HB vs SF
Add question on practical applicability

@@
 These views are partially validated by experiments in previous literature where, on CPUs, a direct comparison between Treebeard (that performs optimizations like unrolling and interleaving) and Hummingbird reveals an order-of-magnitude performance gap.
+We expect that a future comparison of Hummingbird and SilvanForge on GPUs will show similar results.
+
+Rev#D Practical applicability.
+There is interest from industry practitioners to adopt SilvanForge based on the results we demonstrate.
+We gather that there are a wide variety of application specific use-cases where maximizing performance at both low and high batch sizes is important.
 

### 2024-07-26 docs/SOSP 2024/rebuttal.txt
Grammarly suggestions

--- 2024-07-26 docs/SOSP 2024/rebuttal.txt
@@
+We thank the reviewers for their insightful comments.
+
 Q1: Novelty of scheduling language.
@@
 
-Q2: Comparison with Hummingbird
+Q2: Comparison with Hummingbird, PyTorch etc.
 
@@
 
-The above discussion brings up a more practical system level problem.
-Expressing the tree inference using PyTorch means that the decision tree compiler has no control over lower level details like shared memory, thread synchronization, etc.
+The above discussion brings up a more practical system-level problem.
+Expressing the tree inference using PyTorch means that the decision tree compiler has no control over lower-level details like shared memory, thread synchronization, etc.
 Kernels are generated dynamically and there is no scope for controlling how these kernels are optimized.
@@
 These views are partially validated by experiments in previous literature where, on CPUs, a direct comparison between Treebeard (that performs optimizations like unrolling and interleaving) and Hummingbird reveals an order-of-magnitude performance gap.
+We expect that a future comparison of Hummingbird and SilvanForge on GPUs will show similar results.
+
+Rev#D Practical applicability.
+
+Industry practitioners are interested in adopting SilvanForge based on the results we demonstrate.
+We gather that there are a wide variety of application-specific use-cases where maximizing performance at both low and high batch sizes is important.
 
 Rev#B: Are your compiler changes 'portable' across platforms?
+
 The scheduling language, high-level IR and mid-level IR and optimizations on them are fully shared.
@@
 The compilation pipelines diverge starting at the lowering of MIR to low-level IR.
-Even so, we are able to share significant amount of code between the CPU and GPU lowering because of how in-memory representations are abstracted (Section 6).
+Even so, we are able to share a significant amount of code between the CPU and GPU lowering because of how in-memory representations are abstracted (Section 6).
 
 Rev#C: Support for reductions.
+
 MLIR currently supports parallel reductions for value types (scf.reduce) and reducing a dimension of a tensor (linalg.reduce).

### 2024-07-27 docs/SOSP 2024/rebuttal.txt
minor edits

@@
 This, as discussed in our paper, is not always useful (Section 2).
-However.
-it is very difficult to express other SilvanForge optimizations including tree walk interleaving, caching trees/rows and reductions in shared memory, using tensor operations.
+However, it is very difficult to express other SilvanForge optimizations including tree walk interleaving, caching trees/rows and reductions in shared memory, using tensor operations.
 
@@
 Kernels are generated dynamically and there is no scope for controlling how these kernels are optimized.
-Systems like TorchDynamo and TorchInductor do not provide this level of control to the PyTorch program.
+Systems like TorchDynamo and the recently introduced TorchInductor (as per the ASPLOS 2024 paper and other docs) do not provide this level of control to the PyTorch program.
 
@@
 Industry practitioners are interested in adopting SilvanForge based on the results we demonstrate.
+We have an ongoing discussion with relevant stakeholders.
 We gather that there are a wide variety of application-specific use-cases where maximizing performance at both low and high batch sizes is important.

### 2024-07-27 docs/SOSP 2024/rebuttal.txt
Add answer to question on benchmarks

@@
-We thank the reviewers for their insightful comments.
+We thank the reviewers for their detailed and insightful reviews.
 
-Q1: Novelty of scheduling language.
+_Q1: Novelty of scheduling language._
 
@@
 
-Q2: Comparison with Hummingbird, PyTorch etc.
+_Q2: Comparison with Hummingbird, PyTorch etc._
 
@@
 
-Rev#D Practical applicability.
+_Rev#D: Practical applicability._
 
@@
 
-Rev#B: Are your compiler changes 'portable' across platforms?
+_Rev#A: Benchmarks, absolute times._
+
+We used the benchmark set from previous work and included a large number of synthetic models with variations in depth, number of trees, and number of features to provide additional coverage.
+Our results are consistent across all these models, indicating robustness.
+Regarding the execution times, the reported numbers represent the amortized execution time per row averaged over the batch.
+On a GPU, multiple operations are performed in parallel, making it challenging to directly correlate the reported times with single operations.
+
+_Rev#B: Portability._
 
@@
 
-Rev#C: Support for reductions.
+_Rev#C: Support for reductions._
 

### 2024-07-27 docs/SOSP 2024/rebuttal.txt
Shorten, split question on benchmarks and time

@@
-We thank the reviewers for their detailed and insightful reviews.
+We thank the reviewers for their insightful reviews.
 
@@
 This has parallels to how TVM extends Halide with domain-specific primitives for DNN compilation like tensorization and cooperative data loading.
-Despite the fact that TVM and Halide operate on similar domains (deep learning vs.
-image processing), such extensions were necessary for TVM to perform optimizations that were not expressible in Halide.
+Despite the fact that TVM and Halide operate on similar domains, such extensions were necessary for TVM to perform optimizations that were not expressible in Halide.
 
 We believe that the design of SilvanForge's scheduling language is conceptually satisfying because it shows that adding a few required domain-specific primitives to a core set of loop transformation primitives can capture the optimization space of a complex application like decision tree inference.
-Prior work (e.g.
-Tahoe, RAPIDs) has proposed a small fixed set of optimization techniques, and as discussed in the paper, SilvanForge’s scheduling language is powerful enough to express all these optimizations and more.
+Prior work has proposed a small fixed set of optimization techniques, and as discussed in the paper, SilvanForge’s scheduling language is powerful enough to express all these optimizations and more.
 
@@
 The generated tensor code performs all tree walks in parallel, stores the individual tree predictions in a tensor and finally performs a reduction.
-Hummingbird performs a limited form of tree-walk unrolling (it pads all trees to the same depth and unrolls walks).
-This, as discussed in our paper, is not always useful (Section 2).
-However, it is very difficult to express other SilvanForge optimizations including tree walk interleaving, caching trees/rows and reductions in shared memory, using tensor operations.
+Hummingbird performs a limited form of tree-walk unrolling (it pads all trees to the same depth and unrolls walks), which is not always useful (Section 2).
+However, it is very difficult to express other SilvanForge optimizations including tree walk interleaving, caching and reduction optimizations, using tensor operations.
 
@@
 This makes it impossible for the tensor compiler to read input rows into shared memory as SilvanForge does.
-Additionally, we expect that the tensor representation will necessitate multiple kernel calls (at least one for the tree walk and one for the reduction) which will be inefficient compared to the single kernel call that SilvanForge generates.
+Additionally, we expect that the tensor representation will necessitate multiple kernel calls (at least one for the tree walks and one for the reduction) which will be inefficient compared to the single kernel call that SilvanForge generates.
 
 The above discussion brings up a more practical system-level problem.
-Expressing the tree inference using PyTorch means that the decision tree compiler has no control over lower-level details like shared memory, thread synchronization, etc.
+Expressing inference using PyTorch means that the decision tree compiler has no control over lower-level details like shared memory, thread synchronization, etc.
 Kernels are generated dynamically and there is no scope for controlling how these kernels are optimized.
@@
 These views are partially validated by experiments in previous literature where, on CPUs, a direct comparison between Treebeard (that performs optimizations like unrolling and interleaving) and Hummingbird reveals an order-of-magnitude performance gap.
-We expect that a future comparison of Hummingbird and SilvanForge on GPUs will show similar results.
+We expect a future comparison of Hummingbird and SilvanForge on GPUs to show similar results.
 
@@
 
-_Rev#A: Benchmarks, absolute times._
+_Rev#A: Benchmarks._
 
@@
 Our results are consistent across all these models, indicating robustness.
-Regarding the execution times, the reported numbers represent the amortized execution time per row averaged over the batch.
-On a GPU, multiple operations are performed in parallel, making it challenging to directly correlate the reported times with single operations.
+
+_Rev#A: Absolute times._
+
+The reported numbers are the mean execution time per row averaged over the batch.
+Since several rows are processed in parallel on the GPU, it is challenging to directly correlate the reported times with single operations.
 
@@
 
-The scheduling language, high-level IR and mid-level IR and optimizations on them are fully shared.
+The scheduling language, HIR and MIR and optimizations on them are fully shared.
 The same scheduling language is used to specify the structure of inference code on CPUs and GPUs.
 When a schedule doesn't mark any loops as targetting a GPU dimension, CPU code is generated.
-The compilation pipelines diverge starting at the lowering of MIR to low-level IR.
+The compilation pipelines diverge starting at the lowering of MIR to LIR.
 Even so, we are able to share a significant amount of code between the CPU and GPU lowering because of how in-memory representations are abstracted (Section 6).
@@
 
-MLIR currently supports parallel reductions for value types (scf.reduce) and reducing a dimension of a tensor (linalg.reduce).
+MLIR currently supports parallel reductions for value types (scf.reduce) and reducing dimensions of tensors (linalg.reduce).
 Neither of these meets the requirements for generating code for decision tree inference.

### 2024-07-27 docs/SOSP 2024/rebuttal.txt
Final minor edits

@@
 While this enables re-targetability, the tensor abstraction is insufficient to express all SilvanForge optimizations.
-The generated tensor code performs all tree walks in parallel, stores the individual tree predictions in a tensor and finally performs a reduction.
+In the latest Hummingbird version, the generated tensor code performs all tree walks in parallel, stores the individual tree predictions in a tensor and finally performs a reduction.
 Hummingbird performs a limited form of tree-walk unrolling (it pads all trees to the same depth and unrolls walks), which is not always useful (Section 2).
@@
 
-Industry practitioners are interested in adopting SilvanForge based on the results we demonstrate.
-We have an ongoing discussion with relevant stakeholders.
-We gather that there are a wide variety of application-specific use-cases where maximizing performance at both low and high batch sizes is important.
+We are currently engaging with industry practitioners and library code owners who are keen to adopt SilvanForge, inspired by the promising results we’ve demonstrated.
+Their requirements include maximizing performance across a range of batch sizes, as their application use-cases vary.
 

### 2024-08-12 docs/SOSP 2024/caching.tex

@@
 are allocated, how to move from a node to its child, how trees 
-are cached, reading the value of leaves and now threshold and 
+are cached, reading the value of leaves and how threshold and 
 feature indices are read from the allocated buffers.
@@
 the user can specify that the working set of an iteration of a loop
-needs to be cached using the \op{cache} directive. This 
-provides a unified way to specify caching of both trees and input rows.
+needs to be cached using the \op{cache} directive. 
+% This provides a unified way to specify caching of both trees and input rows.
 \Treebeard{} implements caching at the granularity of a tree or a row.
@@
 Caching is encoded in the mid-level IR using the \op{cacheTrees} and \op{cacheRows} operations (Table \ref{Tab:IRSpecification}). 
-These operations are generated when the HIR lowered to MIR and \op{cache} is specified on an 
-index variable in the schedule. While the HIR is being lowered and a cached index variable is 
+% These operations are generated when the HIR lowered to MIR and \op{cache} is specified on an 
+% index variable in the schedule. 
+When the HIR is being lowered and a cached index variable is 
 encountered, the compiler generates a \op{cacheTrees} or \op{cacheRows} operation depending on 
 whether the index variable is a tree or a batch index variable.
-\Treebeard{} also determines the working set of the loop  
+\Treebeard{} also determines the working set of one iteration of the loop  
 and generates a caching operation with the appropriate limits.
@@
 because the input is currently assumed to be a dense
-array format. The lowering for the \op{cacheRows} operation is implemented directly in the \Treebeard{} compiler.
+array format. The lowering for the \op{cacheRows} operation is 
+implemented directly in the \Treebeard{} compiler as a series of 
+coalesced loads into shared memory.
 
@@
 provides a lowering to the target-specific code 
-generator to lower the \op{cacheTrees} op when
-that representation is used.
+generator to lower the \op{cacheTrees} op.
 %  However, the \Treebeard{} infrastructure does provide helpers 

### 2024-08-12 docs/SOSP 2024/intro.tex
Fix typos

@@
 the optimization space. We demonstrate that the language is 
-is capable of expressing all optimizations proposed by prior work and more.
+capable of expressing all optimizations proposed by prior work and more.
 Further, our schedule exploration heuristic can quickly find a near 
@@
 %% \TODO{Oganization}
-In summary, this paper makes the following contributions.
+In summary, this paper makes the following important contributions.
 \begin{itemize}

### 2024-08-12 docs/SOSP 2024/overview.tex

@@
 \label{Sec:Overview}
-\Treebeard{} takes a serialized decision tree ensemble as input (For example
-XGBoost JSON, ONNX etc.) and generates an optimized inference function. 
+\Treebeard{} takes a serialized decision tree ensemble as input (E.g.:
+XGBoost JSON, ONNX etc.) and automatically generates an optimized inference function. 
 \Treebeard{} automatically generates an optimized inference function from 
@@
 or its constituent trees. Figure \ref{Fig:HIRExample} shows this representation for a model 
-with four trees and how the model can be transformed. We use this model as a running example 
-for the rest of the section.
+with four trees and how the model can be transformed. The model contains two trees of 
+depth 1 and two trees of depth 2 (\circled{A}). The right side of the diagram shows the models after 
+padding and reordering (\circled{B}). Padding makes all leaves in a tree the same depth (Trees 2 and
+4 are padded). Trees are reordered so that trees of the same depth are together (Trees 
+2 and 3 are swapped). We use this model as a running example for the rest of the section.
 After these model-level optimizations are performed on the HIR, the 
-code is lowered to the mid-level IR (MIR) as dictated by a user 
-specified \emph{\textbf{schedule}}. The schedule determines how
-the iteration space that goes over the trees and input rows is to be traversed
-by specifying how loops are to be tiled, which loops are to be
-parallelized, which loops are to be mapped to GPU grid and block dimensions etc.
+code is lowered to the mid-level IR (MIR) as dictated by a  
+\emph{\textbf{schedule}}. The schedule determines how to traverse
+the iteration space that goes over the trees and input rows 
+by specifying how loops are to be tiled, 
+parallelized, mapped to GPU grid and block dimensions etc.
 (Details in Section \ref{sec:schedule}). While MIR is a loop-based IR that 
@@
 to correctly implement reduction in the presence of 
-parallel loops. This process, that we call \emph{legalization}, is 
+parallel loops. This process, that we call \textbf{\emph{legalization}}, is 
 described in Section \ref{sec:reduction}.
@@
 invocations are inserted into the LIR. Additionally, buffers 
-to hold model values are inserted and tree operations are lowered to
+to hold model values are inserted and abstract tree operations are lowered to
 explicitly refer to these buffers. This lowering is controlled by 
-a plugin mechanism where different in-memory representations can 
+a plugin mechanism which enables different in-memory representations to 
 be added to the compiler by implementing an interface 
@@
 from Figure \ref{Fig:HIRExample} how different schedules can be used to 
-lower the model to MIR and LIR. We also show how the \op{reduce} operation
+lower the model to MIR. We also show how the \op{reduce} operation
 is lowered and legalized.
 
-First, we consider the the schedule that processes one tree at a time 
-for all input rows and unrolls the tree walks for all trees in the model. 
+First, we consider the schedule that processes one tree at a time 
+for all input rows and unrolls all tree walks. 
 We assume that the trees have been reordered so that all trees with 
@@
 The schedule splits the loop over the trees into two loops -- one that
-iterates over the first two trees and the second that iterates over the
-last two trees. The schedule the unrolls the tree walks for each tree.
+iterates over the first two trees (Trees 1 and 3 with depth 1) and 
+the second that iterates over the last two trees (Trees 2 and 4 with
+depth 2). The schedule then unrolls the tree walks for each tree.
 \begin{lstlisting}[style=c++]
@@
 
-\Treebeard{} represents the set of loops as a tree of loops, where the children 
-of a loop node are the nodes for the immediately contained loops. Each schedule 
-primitive modifies this tree in some way. Additionally,
-the lineage of each of the loops is maintained. This allows 
+\Treebeard{} represents loops as nodes in a tree where the children
+of a node represent immediately contained loops. Each schedule 
+primitive modifies this tree in some way. For example, the \op{split}
+primitive above duplicates the subtree rooted at the node that is 
+being split. The compiler tracks the lineage of each of the loops. This allows 
 the compiler to automatically infer the ranges for all loops. To lower the
-HIR to MIR, the compiler traverses the schedule tree and generates the
-corresponding loops. The MIR for the example model, generated using the above   
+HIR to MIR, the compiler traverses the tree of loops and generates the
+appropriate loops. The MIR for the example model, generated using the above   
 schedule is as follows. 
@@
 Subsequent lowering rewrites the \op{walkDecisionTree} operations to a 
-series of \op{traverseTile} operations, in this case one for the 
+series of \op{traverseTile} operations -- one for the 
 first instance and two for the second (as specified by the unroll depth
@@
 
-We now show how we would compile our example model to a GPU. 
-Additionally, we divide the trees into two sets and process each set
-in parallel. We also unroll the tree walks for each tree.
-The schedule to accomplish this is as follows.
+We now show how easily the same model can be compiled to target a GPU 
+using \Treebeard{}'s scheduling language. 
+A possible schedule to accomplish this is as follows.
 \begin{lstlisting}[style=c++]
@@
 \end{lstlisting}
-This schedule processes on input row per thread block 
-and splits the trees into two sets, each of which contains 
-two trees. The MIR generated is as follows. 
+This schedule processes one input row per thread block (since the \op{batch}
+loop is mapped directly to \op{grid.x}).
+It also splits the trees into two sets by tiling the \op{tree} loop.
+Each of the two sets is processed in parallel. We unroll the tree walks 
+for each tree. The MIR generated is as follows. 
 \begin{lstlisting}[style=c++]
@@
 
-In the case of this example, the lowering passes for the \op{reduce}
-operation determines that several paralel iterations of the \op{t1} 
+In the case of this schedule, the lowering passes for the \op{reduce}
+operation determines that parallel iterations of the \op{t1} 
 loop accumulate into the same element of the \op{result} array.
@@
 iteration accumulates into a different array element by introducing 
-a temporary buffer as follows.
+a temporary buffer (\op{tempResults}) as follows.
 \begin{lstlisting}[style=c++]
@@
     }
-    result[batch] = reduce_dimension(result[0], 1)
+    result[batch] = reduce_dimension(tempResults[:][batch], 0)
   }
@@
 reduced across the \op{t1} dimension (represented by the \op{reduce\_dimension}
-where the second argument represents the dimension number to reduce across)
-to get the final result.
-
-Finally, since this computation is being targeted to the GPU, the 
-parallel loops are rewritten into kernel calls and the  required 
-GPU memory allocations and memory transfers are introduced.
+operation) to get the final result. Finally, since this computation is 
+being targeted to the GPU, the parallel loops are rewritten into kernel calls 
+and the required GPU memory allocations and memory transfers are introduced.
 
@@
 nest in the inference routine can get quite complex even 
-for simple optimizations. Writing these routines by hand 
-is error prone and time-consuming. Building a principled code 
-generator for these routines is the best way to explore the
-vast design space. 
-% Show simple one tree lowering for CPU with walk unrolling
-%    Schedule, MIR, 
-% Show parallelization across trees and GPU target
+for simple schedules. Writing these routines by hand 
+is error-prone and time-consuming. Building a principled code 
+generator to automatically generate these routines is the 
+best way to explore the vast design space. The key features in 
+the design of \Treebeard{} that enable us to build such a code
+generator are as follows.
 
-To reiterate, the following are the salient points of \Treebeard{}'s design.
 \begin{enumerate}
@@
   is specified as an input to the compiler using a scheduling language that is 
-  specialized for decision tree inference computations (Section \ref{Sec:SchedulingLang}).
+  specialized for decision tree inference computations (Section \ref{sec:schedule}).
   This separation allows us to build optimizations and schedule exploration 
-  mechanisms independent of the core compiler (Section \ref{AutoTuneHeuristic}). 
-  The scheduling language also exposes details like parallelization, optimization
-  of reductions (vectorization, use atomics, shared memory on GPUs etc.).
+  mechanisms independent of the core compiler (Section \ref{sec:exploring}).
 
@@
   capabilities to the code generator as a plugin. This interface abstracts several details
-  on how model values are stored. In particular, it abstracts the actual layout of 
-  the memory buffers, how to load model values like thresholds and feature indices, how 
-  to move from a node to its children and determining whether a node is a leaf.
-  This design allows us to write each memory representation as a standalone plugin 
-  and reuse the rest of the compiler infrastructure.
+  on how model values are stored (Details in Section \ref{sec:representations}). This 
+  design allows us to write each in-memory representation as a standalone plugin 
+  and reuse the rest of the compiler infrastructure. \TODO{Should we also say it lets us
+  easily search over multiple representations?}
 \end{enumerate}

### 2024-08-12 docs/SOSP 2024/reduction.tex

@@
 To address this gap, we design an MLIR dialect that allows us to
-specify accumulating values into an element of a multi-dimensional array
+specify accumulating values into an element of a multidimensional array
 and can be lowered to CPU or GPU.
@@
 atomically accumulating values into an element of a 
-multi-dimensional array (represented by an MLIR \op{memref}).
+multidimensional array (represented by an MLIR \op{memref}).
 Consider the following \Treebeard{} schedule.
@@
 reduction operations with a well-defined initial value. The 
-reduction operator and the initial value are attributes applied
+reduction operator and the initial value are attributes
 on the \op{reduce} op.
@@
 % GPUs simply using different MLIR rewrite patterns. 
-In this section, we briefly describe 
-how we lower the reduction operations to the CPU and GPU.
+% In this section, we briefly describe 
+% how we lower the reduction operations to the CPU and GPU.
 
@@
 % using inter-thread shuffles.
-
 % Another feature specific to GPU reductions is the use of shared memory. 

### 2024-08-13 results/python/analyze_gpu_auto_schedule_results.py
Delete raw results

@@ -1,274 +1,0 @@
-from matplotlib import pyplot as plt
-import pandas as pd
-import numpy as np
-import os
-
-def time_diff_computation(x):
-    five_diff = (x.iloc[9] - x.iloc[0])/x.iloc[0]
-    ten_diff = (x.iloc[9] - x.iloc[0])/x.iloc[0]
-    twentyfive_diff = (x.iloc[25] - x.iloc[0])/x.iloc[0]
-    return pd.Series({ 'five_diff': five_diff, 'ten_diff': ten_diff, 'twentyfive_diff': twentyfive_diff})
-
-def compute_rank_in_other_table(df_sorted, df_best, benchmark_names, batch_sizes):
-    for benchmark in benchmark_names:
-        for batch_size in batch_sizes:
-            results_df = df_sorted[(df_sorted['model'] == benchmark) & (df_sorted['batch_size'] == batch_size)]
-            best_schedule = df_best[(df_best['model'] == benchmark) & (df_best['batch_size'] == batch_size)]
-            assert (best_schedule.shape[0] == 1)
-            best_schedule_name = best_schedule.iloc[0]['schedule']
-            best_schedule_rep = best_schedule.iloc[0]['representation']
-            # reset the indices for results_df
-            results_df = results_df.reset_index(drop=True)
-            # find the index of the entry with schedule=best_schedule_name and representation=best_schedule_rep in results_df
-            best_schedule_index = results_df.index[(results_df['schedule'] == best_schedule_name) & (results_df['representation'] == best_schedule_rep)].tolist()[0]
-            # print(best_schedule_index)
-            # add best_schedule_index to df_best_total_time as a new column "kernelRank"
-            df_best.loc[(df_best['model'] == benchmark) & (df_best['batch_size'] == batch_size), 'kernelRank'] = best_schedule_index
-
-def plot_normalized_time_histogram(df_sorted, df_best, col_name):
-    # for every row in df_sorted, find the row in df_best with the same benchmark and batch_size and 
-    # divide the total time of the row in df_sorted by the total time of the row in df_best
-    # plot a histogram of the normalized times
-    for index, row in df_sorted.iterrows():
-        benchmark = row['model']
-        batch_size = row['batch_size']
-        best_schedule = df_best[(df_best['model'] == benchmark) & (df_best['batch_size'] == batch_size)]
-        assert (best_schedule.shape[0] == 1)
-        best_schedule_total_time = best_schedule.iloc[0][col_name]
[…]

### 2024-08-13 results/python/analyze_gpu_schedule_exploration_results.py

@@ -1,92 +1,0 @@
-import pandas as pd
-import numpy as np
-
-def time_diff_computation(x):
-    five_diff = (x.iloc[9] - x.iloc[0])/x.iloc[0]
-    ten_diff = (x.iloc[9] - x.iloc[0])/x.iloc[0]
-    twentyfive_diff = (x.iloc[25] - x.iloc[0])/x.iloc[0]
-    return pd.Series({ 'five_diff': five_diff, 'ten_diff': ten_diff, 'twentyfive_diff': twentyfive_diff})
-
-def compute_rank_in_other_table(df_sorted, df_best, benchmark_names, batch_sizes):
-    for benchmark in benchmark_names:
-        for batch_size in batch_sizes:
-            results_df = df_sorted[(df_sorted['benchmark'] == benchmark) & (df_sorted['batchSize'] == batch_size)]
-            best_schedule = df_best[(df_best['benchmark'] == benchmark) & (df_best['batchSize'] == batch_size)]
-            assert (best_schedule.shape[0] == 1)
-            best_schedule_name = best_schedule.iloc[0]['schedule']
-            best_schedule_rep = best_schedule.iloc[0]['representation']
-            # reset the indices for results_df
-            results_df = results_df.reset_index(drop=True)
-            # find the index of the entry with schedule=best_schedule_name and representation=best_schedule_rep in results_df
-            best_schedule_index = results_df.index[(results_df['schedule'] == best_schedule_name) & (results_df['representation'] == best_schedule_rep)].tolist()[0]
-            # print(best_schedule_index)
-            # add best_schedule_index to df_best_total_time as a new column "kernelTimeRank"
-            df_best.loc[(df_best['benchmark'] == benchmark) & (df_best['batchSize'] == batch_size), 'kernelTimeRank'] = best_schedule_index
-
-results_file = "/home/ashwin/mlir-build/llvm-project/mlir/examples/treebeard/results/thedeep/gpu/exploration/gpu_results_thedeep_nopad.txt"
-# results_file = "/home/ashwin/mlir-build/llvm-project/mlir/examples/treebeard/results/thedeep/gpu/exploration/gpu_results_thedeep_try2.txt"
-# results_file = "/home/ashwin/mlir-build/llvm-project/mlir/examples/treebeard/results/holmes/gpu/exploration/gpu_results_holmes.txt"
-# read results file into pandas dataframe
[…]

### 2024-08-13 results/python/analyze_gpu_speedups_results.py

@@ -1,326 +1,0 @@
-import matplotlib.pyplot as plt
-import numpy as np
-import pandas as pd
-
-from scipy.stats import gmean
-
-# Read the CSV file into a dataframe
-results_4060_df = pd.read_csv('../4060_thedeep_results.csv')
-results_t400_df = pd.read_csv('../T400_holmes_results.csv')
-fullexp_over_at_df = pd.read_csv('../fullexp_vs_at.csv')
-results_amdmi2160_df = pd.read_csv('../AMDMI2160_results.csv')
-
-results_4060_df = results_4060_df[results_4060_df['Batch size'] != 256]
-results_t400_df = results_t400_df[results_t400_df['Batch size'] != 256]
-fullexp_over_at_df = fullexp_over_at_df[fullexp_over_at_df['Batch size'] != 256]
-results_amdmi2160_df = results_amdmi2160_df[results_amdmi2160_df['Batch size'] != 256]
-
-PLOT_FILE_EXTENSION = 'png'
-AXIS_FONT_SIZE = 15
-LABEL_FONT_SIZE = 15
-LEGEND_FONT_SIZE = 11.5
-BAR_LABEL_FONT_SIZE = 11
-
-def plot_bar_graph_of_abs_times_for_batch_size(df, batch_size):
-    filtered_df = df[df['Batch size'] == batch_size]
-    bar_names = filtered_df['Benchmark'].to_list()
-    tb_kernel_time = filtered_df['TBKernel(AT)'].to_list()
-    rapids_kernel_time = filtered_df['RAPIDS(kernel)'].to_list()
-    tb_total_time = filtered_df['TB (AT)'].to_list()
-    rapids_total_time = filtered_df['RAPIDS (Total)'].to_list()
-
-    # Calculate the difference between kernel time and total time
-    tb_diff = [tb_total - tb_kernel for tb_total, tb_kernel in zip(tb_total_time, tb_kernel_time)]
-    rapids_diff = [rapids_total - rapids_kernel for rapids_total, rapids_kernel in zip(rapids_total_time, rapids_kernel_time)]
-
-    # Normalize w.r.t rapids_total_time
-    norm_tb_kernel_time = [time/ rapids_total for time, rapids_total in zip(tb_kernel_time, rapids_total_time)]
-    norm_rapids_kernel_time = [time/ rapids_total for time, rapids_total in zip(rapids_kernel_time, rapids_total_time)]
[…]

### 2024-08-13 results/python/analyze_random_model_benchmark_results.py

@@ -1,98 +1,0 @@
-import os
-import matplotlib.pyplot as plt
-
-batch_size = 4096
-file_path = "/home/ashwin/mlir-build/llvm-project/mlir/examples/treebeard/results/thedeep/gpu/random_models/20240320/random_cmp_output_mp.txt"
-# file_path = "/home/ashwin/mlir-build/llvm-project/mlir/examples/treebeard/results/thedeep/gpu/random_models/20240320/random_cmp_output_mp_b512.txt"
-# file_path = "/home/ashwin/mlir-build/llvm-project/mlir/examples/treebeard/results/thedeep/gpu/random_models/20240320/random_cmp_output_mp_b4k_ft500.txt"
-
-# read the file line by line and filter out lines that don't start with "/"
-with open(file_path, "r") as file:
-    lines = file.readlines()
-    lines = [line for line in lines if line.startswith("/")]
-
-# split each line by the space char
-lines = [line.split(" ") for line in lines]
-
-# construct a pandas dataframe from the list of lists
-import pandas as pd
-cols = ["model_path", "batch_size", "fil_time", "tb_time", "total_speedup", "fil_kernel_time", "tb_kernel_time", "tb_kernel_speedup"]
-df = pd.DataFrame(lines, columns=cols)
-
-# cast the tb_kernel_speedup and total_speedup column to float
-df["tb_kernel_speedup"] = df["tb_kernel_speedup"].astype(float)
-df["total_speedup"] = df["total_speedup"].astype(float)
-
-# compute it from the model_path as follows
-# - split the model_path by the "/" char and get the file name
-# - split the file name by the "_" char and get the third element
-def get_num_trees(model_path):
-    return int(model_path.split("/")[-1].split("_")[2])
-
-# add a column to the dataframe called "num_trees". 
-# This column contains the number of trees in the model
-df["num_trees"] = df["model_path"].apply(get_num_trees)
-
-def get_num_features(model_path):
-    return int(model_path.split("/")[-1].split("_")[3])
-
-df["num_features"] = df["model_path"].apply(get_num_features)
-
-def get_tree_depth(model_path):
-    return int(model_path.split("/")[-1].split("_")[4])
-
-df["tree_depth"] = df["model_path"].apply(get_tree_depth)
-
[…]

### 2024-08-13 results/python/analyze_schedule_sensitivity_data.py

@@ -1,156 +1,0 @@
-# plot a 3-d graph of num_trees, num_features and tb_kernel_speedup
-import pandas as pd
-import numpy as np
-import xarray as xr
-import matplotlib.pyplot as plt
-from mpl_toolkits.mplot3d import Axes3D
-
-# get the path of the current script
-import os
-script_path = os.path.dirname(os.path.realpath(__file__))
-# append ../thedeep/gpu/sensitivity to the path
-sensitivity_path = os.path.join(script_path, "..", "thedeep", "gpu", "sensitivity")
-
-def get_degradation_vals(df_benchmark, X, Y, x_col="cur_batch_size", y_col="cmp_batch_size"):
-    # get the degradation values for the given benchmark
-    degradation_vals = np.zeros((len(X), len(Y)))
-    for i in range(len(X)):
-        for j in range(len(Y)):
-            x = X[i]
-            y = Y[j]
-            df_depth = df_benchmark[(df_benchmark[x_col] == x) & (df_benchmark[y_col] == y)]
-            # assert df_depth.shape[0] == 1, f"Expected 1 row for {x}, {y}, got {df_depth.shape[0]}"
-            if df_depth.shape[0] == 0:
-                degradation_vals[i, j] = -1
-                continue
-            degradation_vals[i, j] = df_depth["degradation"].values[0]
-            # if x < y:
-            #     degradation_vals[i, j] = 10 * degradation_vals[i, j]
-
-    max_val = np.max(degradation_vals)
-    for deg_row in degradation_vals:
-        for i in range(len(deg_row)):
-            if deg_row[i] == -1:
-                deg_row[i] = max_val
-    return degradation_vals
-
-def plot_batch_sensitivity_for_single_benchmark(df_depth, benchmark_name):
-    # draw an intensity plot of the degradation values
-    # x axis is cur_batch_size, y axis is cmp_batch_size, intensity is degradation
-    # x and y axis in log scale
-
-    # get a numpy array for the x, y and z axis
-    x = df_depth["cur_batch_size"].unique()
-    y = df_depth["cmp_batch_size"].unique()
-    x.sort()
-    y.sort()
-    Z = get_degradation_vals(df_depth, x, y) # df_depth['degradation'].values.reshape(len(y), len(x))
-    # print(Z)
-
-    da = xr.DataArray(
[…]

### 2024-08-13 results/python/filter_exploration_results.py

@@ -1,16 +1,0 @@
-import sys
-
-# Get the input file name as the first argument and the output file name as the second argument
-input_file = "/home/ashwin/mlir-build/llvm-project/mlir/examples/treebeard/results/thedeep/gpu/exploration/20240229/exploration_b512-2k_stderr_letters.txt"
-output_file = input_file.replace(".txt", "_filtered.txt")
-
-# Open the input file, read it line by line and split the string by ","
-with open(input_file, "r") as file:
-    lines = file.readlines()
-    # Filter only the lines that contain 13 parts when split by ","
-    lines = [line for line in lines if len(line.split(",")) == 13]
-
-# Write the resulting lines to the output file
-with open(output_file, "w") as file:
-    file.writelines(lines)
-    
